---
title: "Match PAC Classifications"
output: html_notebook
---

# Summary

This notebook computes an industry distribution for each PAC by:
  - assigning under $5k contributors the sector from `industry_override_under_5k` from the manual PAC review
  - assigning `industry_override_under_5k` to entire amount contributed by a federal PAC
    - we don't have access to all of their received contributions, so just classify the entire PAC when possible
  - looking up individually classified over $5k contributors from final_classifications
  - leaving PACs flagged `additional_review_req` or `PENDING` as uncategorized

# Load data and packages

```{r}
library(dplyr)
library(readr)
library(stringr)
library(digest)
library(googlesheets4)
```

```{r}
hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(
    digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16
  ))
}
```

```{r}
over_5k_tagged <- read_csv("../01_contributor_pipeline/01_outputs/over_5k_tagged_09-08-26.csv",
                                 col_types = cols(.default = "c", contributor_amount = "d"))

pac_classifiability <- read_csv("../01_contributor_pipeline/01_outputs/pac_classifiability_09-08-26.csv",
                                 col_types = cols(.default = "c", total_received = "d"))

pac_class_input <- read_csv("../01_contributor_pipeline/01_outputs/pac_classification_input_09-08-26.csv",
                                 col_types = cols(.default = "c"))

ordered_pacs <- read_csv("../01_contributor_pipeline/01_outputs/pac_traversal_order_09-08-26.csv") %>%
  arrange(order) %>% pull(pac_id) %>% as.character()

stuck_pacs <- read_csv("../01_contributor_pipeline/01_outputs/pac_stuck_pacs_09-08-26.csv") %>%
  pull(pac_id) %>% as.character()

pac_final <- read_csv("04_outputs/final_classifications_full_08-23-25.csv",
                                 col_types = cols(.default = "c")) # final classifications after manual review (including pac and direct contributors)

pac_review <- read_csv("04_inputs/pac_manual_review.csv",
                                 col_types = cols(.default = "c", under_5k_amount = "d")) # manual review feedback 
```


```{r}
# add paths to any additional PAC review files here (oldest -> newest)
# each will be merged into pac_review in order, with conflicts flagged.
supplemental_review_paths <- c(
  "04_inputs/pac_manual_review_09-08-26.csv"
)

supplemental_reviews <- lapply(supplemental_review_paths, function(p) {
  read_csv(p, col_types = cols(.default = "c", under_5k_amount = "d")) %>%
    mutate(review_source = basename(p), pac_id = as.character(pac_id))
}) %>% bind_rows()
```


### combine pac review into one dataframe

```{r}
# union subcategory labels that map to the broader "Unions" override in newer reviews.
union_subcategory_overrides <- c(
  "Misc Unions", "Unions", "Public Safety Unions", "Public Sector Unions",
  "Social Services Unions", "Trade Unions"
)

pac_review <- pac_review %>% mutate(pac_id = as.character(pac_id))

all_major_donors     <- list()
all_new_pending      <- list()
all_id_conflicts     <- list()
all_name_conflicts   <- list()
all_pending_resolved <- list()

for (src in unique(supplemental_reviews$review_source)) {
  supp <- supplemental_reviews %>% filter(review_source == src)
  cat(sprintf("\n=== Merging: %s ===\n", src))

  # 1. flag major donors
  major_donor_rows <- supp %>%
    filter(grepl("major donor", notes, ignore.case = TRUE))
  all_major_donors[[src]] <- major_donor_rows %>%
    mutate(review_source = src) %>%
    select(review_source, pac_id, Contributor.Name, notes)
  supp_clean <- supp %>%
    filter(!grepl("major donor", notes, ignore.case = TRUE))

  # 2. pendings pacs
  new_pending <- supp_clean %>%
    filter(industry_override_under_5k == "PENDING" |
           unresolvable_reason == "additional_review_req")
  all_new_pending[[src]] <- new_pending %>%
    mutate(review_source = src) %>%
    select(review_source, pac_id, Contributor.Name,
           industry_override_under_5k, unresolvable_reason)

  # 3. fill pac_id for no-id rows where reviewer coded one (notes = "coded id" or "coded id possible")
  coded_id_rows <- supp_clean %>%
    filter(grepl("coded id", notes, ignore.case = TRUE),
           !is.na(pac_id), pac_id != "") %>%
    distinct(Contributor.Name, .keep_all = TRUE) %>%
    select(Contributor.Name, new_pac_id = pac_id)

  if (nrow(coded_id_rows) > 0) {
    pac_review <- pac_review %>%
      left_join(coded_id_rows, by = "Contributor.Name") %>%
      mutate(pac_id = if_else(
        (is.na(pac_id) | pac_id == "") & !is.na(new_pac_id),
        new_pac_id, pac_id
      )) %>%
      select(-new_pac_id)
    cat(sprintf("  Filled pac_id for %d previously no-id row(s).\n", nrow(coded_id_rows)))
  }

  # 4. detect conflicts (flagged for awareness; new review wins on all overlaps)
  # deduplicate to one override per pac_id/name before joining to avoid many-to-many warnings
  old_id_overrides <- pac_review %>%
    filter(!is.na(pac_id), pac_id != "",
           !is.na(industry_override_under_5k), industry_override_under_5k != "") %>%
    distinct(pac_id, override_old = industry_override_under_5k)

  new_id_overrides <- supp_clean %>%
    filter(!is.na(pac_id), pac_id != "",
           !is.na(industry_override_under_5k), industry_override_under_5k != "") %>%
    distinct(pac_id, override_new = industry_override_under_5k)

  id_conflicts <- old_id_overrides %>%
    inner_join(new_id_overrides, by = "pac_id", relationship = "many-to-many") %>%
    filter(
      override_old != override_new,
      override_old != "PENDING",
      !(override_old %in% union_subcategory_overrides & override_new == "Unions")
    ) %>%
    distinct() %>%
    mutate(review_source = src)
  all_id_conflicts[[src]] <- id_conflicts

  old_name_overrides <- pac_review %>%
    filter(is.na(pac_id) | pac_id == "",
           !is.na(industry_override_under_5k), industry_override_under_5k != "") %>%
    distinct(Contributor.Name, override_old = industry_override_under_5k)

  new_name_overrides <- supp_clean %>%
    filter(is.na(pac_id) | pac_id == "",
           !is.na(industry_override_under_5k), industry_override_under_5k != "") %>%
    distinct(Contributor.Name, override_new = industry_override_under_5k)

  name_conflicts <- old_name_overrides %>%
    inner_join(new_name_overrides, by = "Contributor.Name", relationship = "many-to-many") %>%
    filter(
      override_old != override_new,
      override_old != "PENDING",
      !(override_old %in% union_subcategory_overrides & override_new == "Unions")
    ) %>%
    distinct() %>%
    mutate(review_source = src)
  all_name_conflicts[[src]] <- name_conflicts

  cat(sprintf("  Major donors: %d | PENDING/add'l review: %d | pac_id conflicts: %d | name conflicts: %d\n",
              nrow(major_donor_rows), nrow(new_pending),
              nrow(id_conflicts), nrow(name_conflicts)))

  # 5. Apply new review's overrides (new review wins for all overlapping rows)
  # - industry_override_under_5k: updated where new review has a valid (non-empty, non-PENDING) value
  # - fec_id / is_federal_pac: updated for ALL overlapping rows (new review is more reliable)
  # - unresolvable_reason: cleared for rows that now have a valid override, so they are
  #   not incorrectly skipped by skip_pac_ids downstream

  pending_pac_ids_before <- pac_review %>%
    filter(!is.na(pac_id), pac_id != "", industry_override_under_5k == "PENDING") %>%
    pull(pac_id) %>% unique()

  # id-based: industry override — only where new review has a valid value
  new_id_override <- supp_clean %>%
    filter(!is.na(pac_id), pac_id != "",
           !is.na(industry_override_under_5k),
           industry_override_under_5k != "",
           industry_override_under_5k != "PENDING") %>%
    group_by(pac_id) %>% slice(1) %>% ungroup() %>%
    select(pac_id, new_override = industry_override_under_5k)

  # id-based: federal fields — all overlapping rows regardless of override status
  new_id_federal_base <- supp_clean %>%
    filter(!is.na(pac_id), pac_id != "") %>%
    group_by(pac_id) %>% slice(1) %>% ungroup()
  fed_cols <- intersect(c("fec_id", "is_federal_pac"), names(new_id_federal_base))
  new_id_federal <- new_id_federal_base %>%
    select(pac_id, all_of(fed_cols)) %>%
    rename_with(~ paste0("new_", .), all_of(fed_cols))

  pac_review <- pac_review %>%
    left_join(new_id_override, by = "pac_id") %>%
    mutate(industry_override_under_5k = if_else(
      !is.na(pac_id) & pac_id != "" & !is.na(new_override),
      new_override, industry_override_under_5k
    )) %>%
    select(-new_override) %>%
    left_join(new_id_federal, by = "pac_id")

  if ("new_fec_id" %in% names(pac_review))
    pac_review <- pac_review %>%
      mutate(fec_id = if_else(!is.na(pac_id) & pac_id != "" & !is.na(new_fec_id),
                              new_fec_id, fec_id)) %>%
      select(-new_fec_id)
  if ("new_is_federal_pac" %in% names(pac_review))
    pac_review <- pac_review %>%
      mutate(is_federal_pac = if_else(!is.na(pac_id) & pac_id != "" & !is.na(new_is_federal_pac),
                                      new_is_federal_pac, is_federal_pac)) %>%
      select(-new_is_federal_pac)

  # name-based: industry override only (federal fields not reliable for name-matched rows)
  new_name_override <- supp_clean %>%
    filter(is.na(pac_id) | pac_id == "",
           !is.na(industry_override_under_5k),
           industry_override_under_5k != "",
           industry_override_under_5k != "PENDING") %>%
    group_by(Contributor.Name) %>% slice(1) %>% ungroup() %>%
    select(Contributor.Name, new_override = industry_override_under_5k)

  pac_review <- pac_review %>%
    left_join(new_name_override, by = "Contributor.Name") %>%
    mutate(industry_override_under_5k = if_else(
      (is.na(pac_id) | pac_id == "") & !is.na(new_override),
      new_override, industry_override_under_5k
    )) %>%
    select(-new_override)

  # clear unresolvable_reason = "additional_review_req" for rows that now have a valid override
  pac_review <- pac_review %>%
    mutate(unresolvable_reason = if_else(
      !is.na(industry_override_under_5k) &
      industry_override_under_5k != "" &
      industry_override_under_5k != "PENDING" &
      coalesce(unresolvable_reason, "") == "additional_review_req",
      NA_character_,
      unresolvable_reason
    ))

  # track resolved pending pacs
  all_pending_resolved[[src]] <- pac_review %>%
    filter(pac_id %in% pending_pac_ids_before,
           !is.na(industry_override_under_5k),
           industry_override_under_5k != "PENDING",
           industry_override_under_5k != "") %>%
    distinct(pac_id, override_new = industry_override_under_5k) %>%
    mutate(review_source = src)

  cat(sprintf("  Applied new overrides to overlapping rows (%d PENDING resolved).\n",
              nrow(all_pending_resolved[[src]])))

  # 6. append new rows
  old_pac_ids_now <- pac_review %>% filter(!is.na(pac_id), pac_id != "") %>% pull(pac_id) %>% unique()
  old_names_noid  <- pac_review %>% filter(is.na(pac_id) | pac_id == "") %>% pull(Contributor.Name) %>% unique()

  # for no-id rows, also skip if any name_variant already appears in old review
  new_rows_noid <- supp_clean %>% filter(is.na(pac_id) | pac_id == "")
  if ("name_variants" %in% names(new_rows_noid) && nrow(new_rows_noid) > 0) {
    new_rows_noid <- new_rows_noid %>%
      rowwise() %>%
      mutate(covered_by_variant = !is.na(name_variants) && name_variants != "" &&
               any(trimws(strsplit(name_variants, "\\|")[[1]]) %in% old_names_noid)) %>%
      ungroup() %>%
      filter(!Contributor.Name %in% old_names_noid, !covered_by_variant) %>%
      select(-covered_by_variant)
  } else {
    new_rows_noid <- new_rows_noid %>% filter(!Contributor.Name %in% old_names_noid)
  }

  new_rows <- bind_rows(
    supp_clean %>% filter(!is.na(pac_id), pac_id != "", !pac_id %in% old_pac_ids_now),
    new_rows_noid
  )

  for (col in intersect(names(new_rows), names(pac_review))) {
    if (is.numeric(pac_review[[col]]) && !is.numeric(new_rows[[col]]))
      new_rows[[col]] <- as.numeric(new_rows[[col]])
    else if (is.character(pac_review[[col]]) && !is.character(new_rows[[col]]))
      new_rows[[col]] <- as.character(new_rows[[col]])
  }

  pac_review <- bind_rows(pac_review, new_rows)
  cat(sprintf("  Added %d new row(s). pac_review now has %d rows total.\n",
              nrow(new_rows), nrow(pac_review)))
}

cat(sprintf("\nFinal pac_review: %d rows, %d with no pac_id (name-matched)\n",
            nrow(pac_review),
            sum(is.na(pac_review$pac_id) | pac_review$pac_id == "")))
```

### post-merge: fix federal flags and propagate overrides by fec_id

```{r}
# 1. propagate fec_id within pac_id groups
#    same pac_id can span multiple rows where only one row got a coded fec_id from review
fec_by_pac_id <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "", !is.na(fec_id), fec_id != "") %>%
  group_by(pac_id) %>%
  slice(1) %>%
  ungroup() %>%
  select(pac_id, fec_id_fill = fec_id)

pac_review <- pac_review %>%
  left_join(fec_by_pac_id, by = "pac_id") %>%
  mutate(fec_id = coalesce(fec_id, fec_id_fill)) %>%
  select(-fec_id_fill)

# 2. any row with a non-NA fec_id is a federal PAC
pac_review <- pac_review %>%
  mutate(is_federal_pac = if_else(
    !is.na(fec_id) & fec_id != "",
    "TRUE", is_federal_pac
  ))

# 2b. propagate is_federal_pac = "TRUE" across all rows in the same pac_id group
fed_pac_ids <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "", coalesce(is_federal_pac, "") == "TRUE") %>%
  distinct(pac_id)

pac_review <- pac_review %>%
  mutate(is_federal_pac = if_else(
    !is.na(pac_id) & pac_id != "" & pac_id %in% fed_pac_ids$pac_id,
    "TRUE", is_federal_pac
  ))

# 3. propagate industry override within fec_id groups
#    if multiple pac_ids share a fec_id, they should have consistent overrides.
#    where there is exactly one distinct non-PENDING override -> fill into rows that are missing/PENDING.
#    where overrides genuinely conflict -> warn and print for manual review.
fec_canonical <- pac_review %>%
  filter(!is.na(fec_id), fec_id != "",
         !is.na(industry_override_under_5k),
         industry_override_under_5k != "",
         industry_override_under_5k != "PENDING") %>%
  group_by(fec_id) %>%
  summarize(
    n_distinct = n_distinct(industry_override_under_5k),
    canonical  = dplyr::first(industry_override_under_5k),
    .groups = "drop"
  )

fec_id_conflicts <- fec_canonical %>% filter(n_distinct > 1)
if (nrow(fec_id_conflicts) > 0) {
  warning(sprintf(
    "%d fec_id group(s) have conflicting industry overrides across pac_ids — manual review needed.",
    nrow(fec_id_conflicts)
  ))
  print(pac_review %>%
    filter(fec_id %in% fec_id_conflicts$fec_id) %>%
    select(fec_id, pac_id, Contributor.Name, industry_override_under_5k) %>%
    arrange(fec_id))
}

fec_propagate <- fec_canonical %>%
  filter(n_distinct == 1) %>%
  select(fec_id, fec_canonical_override = canonical)

pac_review <- pac_review %>%
  left_join(fec_propagate, by = "fec_id") %>%
  mutate(industry_override_under_5k = if_else(
    !is.na(fec_id) & fec_id != "" & !is.na(fec_canonical_override) &
    (is.na(industry_override_under_5k) | industry_override_under_5k == "" |
     industry_override_under_5k == "PENDING"),
    fec_canonical_override,
    industry_override_under_5k
  )) %>%
  select(-fec_canonical_override)

cat(sprintf(
  "Rows with fec_id but is_federal_pac != 'TRUE' remaining: %d\n",
  sum(!is.na(pac_review$fec_id) & pac_review$fec_id != "" &
      coalesce(pac_review$is_federal_pac, "") != "TRUE", na.rm = TRUE)
))
cat(sprintf(
  "fec_id override propagation: %d unambiguous fec_id group(s) propagated, %d conflict(s) flagged\n",
  nrow(fec_propagate), nrow(fec_id_conflicts)
))
```

### deduplicate pac_review to one row per pac_id

```{r}
# check for conflicting industry overrides across duplicate rows before collapsing
pac_id_override_conflicts <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  group_by(pac_id) %>%
  summarise(
    n_valid_overrides = n_distinct(industry_override_under_5k[
      !is.na(industry_override_under_5k) &
      industry_override_under_5k != "" &
      industry_override_under_5k != "PENDING"
    ]),
    conflicting_values = paste(sort(unique(industry_override_under_5k[
      !is.na(industry_override_under_5k) &
      industry_override_under_5k != "" &
      industry_override_under_5k != "PENDING"
    ])), collapse = " | "),
    .groups = "drop"
  ) %>%
  filter(n_valid_overrides > 1)

if (nrow(pac_id_override_conflicts) > 0) {
  warning(sprintf(
    "%d pac_id(s) have conflicting industry overrides across duplicate rows — manual resolution needed.",
    nrow(pac_id_override_conflicts)
  ))
  print(pac_review %>%
    filter(pac_id %in% pac_id_override_conflicts$pac_id) %>%
    select(pac_id, Contributor.Name, industry_override_under_5k, is_federal_pac, fec_id) %>%
    arrange(pac_id))
} else {
  cat("No industry override conflicts across duplicate pac_id rows.\n")
}

# collapse to one row per pac_id, combining best values across duplicates
pac_review_noid <- pac_review %>% filter(is.na(pac_id) | pac_id == "")

pac_review <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  group_by(pac_id) %>%
  summarise(
    across(everything(), first),
    # "TRUE" wins if any row in the group has it
    is_federal_pac = if_else(
      any(coalesce(is_federal_pac, "") == "TRUE"), "TRUE", first(is_federal_pac)
    ),
    # take first non-NA, non-empty, non-PENDING override; fall back to first value
    industry_override_under_5k = {
      valid <- industry_override_under_5k[
        !is.na(industry_override_under_5k) &
        industry_override_under_5k != "" &
        industry_override_under_5k != "PENDING"
      ]
      if (length(valid) > 0) valid[1] else first(industry_override_under_5k)
    },
    .groups = "drop"
  ) %>%
  bind_rows(pac_review_noid)

# merge no-id rows into id-matched rows where name matches (case-insensitive)
id_names_upper <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  mutate(name_key = toupper(trimws(Contributor.Name))) %>%
  select(matched_pac_id = pac_id, name_key)

noid_candidates <- pac_review %>%
  filter(is.na(pac_id) | pac_id == "") %>%
  mutate(name_key = toupper(trimws(Contributor.Name))) %>%
  inner_join(id_names_upper, by = "name_key") %>%
  select(-pac_id) %>%
  rename(pac_id = matched_pac_id)

if (nrow(noid_candidates) > 0) {
  cat(sprintf("%d no-id row(s) matched to an id-matched row by name — merging and removing duplicates:\n",
              nrow(noid_candidates)))
  print(noid_candidates %>%
    select(name_key, pac_id, industry_override_under_5k, is_federal_pac, fec_id))

  # fill missing override on id row from no-id row where id row has none
  noid_fill <- noid_candidates %>%
    filter(!is.na(industry_override_under_5k),
           industry_override_under_5k != "",
           industry_override_under_5k != "PENDING") %>%
    distinct(pac_id, .keep_all = TRUE) %>%
    select(pac_id, noid_override = industry_override_under_5k)

  pac_review <- pac_review %>%
    left_join(noid_fill, by = "pac_id") %>%
    mutate(industry_override_under_5k = if_else(
      !is.na(pac_id) & pac_id != "" & !is.na(noid_override) &
      (is.na(industry_override_under_5k) | industry_override_under_5k == "" |
       industry_override_under_5k == "PENDING"),
      noid_override, industry_override_under_5k
    )) %>%
    select(-noid_override)

  # drop the no-id rows that were merged into id rows
  merged_name_keys <- noid_candidates %>% pull(name_key) %>% unique()
  pac_review <- pac_review %>%
    filter(
      !is.na(pac_id) & pac_id != "" |
      !toupper(trimws(Contributor.Name)) %in% merged_name_keys
    )
} else {
  cat("No no-id rows matched to id-matched rows by name.\n")
}

cat(sprintf(
  "pac_review after dedup: %d rows (%d with pac_id, %d without)\n",
  nrow(pac_review),
  sum(!is.na(pac_review$pac_id) & pac_review$pac_id != ""),
  sum(is.na(pac_review$pac_id) | pac_review$pac_id == "")
))

# warn if any pac_no_id / pac_unmatched_id contributor in over_5k_tagged was assigned a pac_id
# during review but that id is not in pac_classifiability — meaning 0803 never traversed it
# and the distribution loop has no contributor data for it.
untraversed_review_ids <- over_5k_tagged %>%
  filter(contributor_type %in% c("pac_no_id", "pac_unmatched_id")) %>%
  mutate(name_key = toupper(trimws(contributor_name))) %>%
  distinct(contributor_name, contributor_type, name_key) %>%
  inner_join(
    pac_review %>%
      filter(!is.na(pac_id), pac_id != "",
             !grepl("major donor", coalesce(notes, ""), ignore.case = TRUE)) %>%
      mutate(name_key = toupper(trimws(Contributor.Name)),
             pac_id   = as.character(pac_id)) %>%
      select(name_key, pac_id, industry_override_under_5k, is_federal_pac),
    by = "name_key"
  ) %>%
  filter(
    !pac_id %in% as.character(pac_classifiability$pac_id),
    coalesce(is_federal_pac, "") != "TRUE",
    trimws(coalesce(industry_override_under_5k, "")) != "Candidates/Political Parties"
  )

if (nrow(untraversed_review_ids) > 0) {
  cat(sprintf(
    "WARNING: %d pac_no_id/pac_unmatched_id contributor(s) were assigned a pac_id in review but that ID is not in pac_classifiability — 0803 needs re-run to traverse them:\n",
    nrow(untraversed_review_ids)
  ))
  print(untraversed_review_ids %>%
    select(contributor_name, contributor_type, pac_id, industry_override_under_5k))
} else {
  cat("All reviewer-assigned pac_ids for pac_no_id/pac_unmatched_id contributors are present in pac_classifiability.\n")
}
```


### audit: PACs without a review entry

```{r}
# PACs in pac_classifiability that have no matching row in pac_review
classifiability_pac_ids <- pac_classifiability %>%
  mutate(pac_id = as.character(pac_id)) %>%
  pull(pac_id) %>%
  unique()

reviewed_pac_ids <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  pull(pac_id) %>%
  unique()

unreviewed_pac_ids <- setdiff(classifiability_pac_ids, reviewed_pac_ids)

unreviewed_detail <- pac_classifiability %>%
  mutate(pac_id = as.character(pac_id), pct_over_5k = as.numeric(pct_over_5k)) %>%
  filter(pac_id %in% unreviewed_pac_ids)

# major donor/IE committees are classified via individual classification path, not pac review
major_donor_unreviewed <- unreviewed_detail %>%
  filter(coalesce(is_major_donor_ie, "FALSE") == "TRUE")

if (nrow(major_donor_unreviewed) > 0) {
  cat(sprintf("INFO: %d major donor/IE committee(s) in pac_classifiability have no review entry — expected, classified via individual path.\n",
              nrow(major_donor_unreviewed)))
}

unreviewed_non_md <- unreviewed_detail %>%
  filter(coalesce(is_major_donor_ie, "FALSE") != "TRUE")

# PACs fully covered by $5k+ classified contributors need no review entry
fully_covered <- unreviewed_non_md %>%
  filter(!is.na(pct_over_5k) & pct_over_5k >= 1 & status != "unresolved")

needs_review <- unreviewed_non_md %>%
  filter(is.na(pct_over_5k) | pct_over_5k < 1 | status == "unresolved")

if (nrow(needs_review) > 0) {
  cat(sprintf("WARNING: %d PAC(s) have no review entry and cannot be fully classified by $5k+ contributors:\n",
              nrow(needs_review)))
  print(needs_review %>%
    select(pac_id, total_received, status, pct_over_5k) %>%
    arrange(desc(as.numeric(total_received))))
} else {
  cat("All PACs without a review entry are fully classifiable by their $5k+ contributors.\n")
}

if (nrow(fully_covered) > 0) {
  cat(sprintf("INFO: %d unreviewed PAC(s) are fully covered by $5k+ contributors (no review needed).\n",
              nrow(fully_covered)))
}
```


### data checks (can skip to next section)

```{r}
# major donors
bind_rows(all_major_donors)

# inspect conflicts
bind_rows(all_id_conflicts)

# inspect name conflicts
bind_rows(all_name_conflicts)

# inspect newly resolved pacs
bind_rows(all_pending_resolved)
```

```{r}
# make sure fed flag persisted
fed_ids <- supplemental_reviews %>% filter(is_federal_pac == "TRUE", !is.na(pac_id)) %>% pull(pac_id)
pac_review %>% filter(pac_id %in% fed_ids) %>% distinct(is_federal_pac)

#pac_review %>% filter(!is.na(fec_id) & coalesce(is_federal_pac, "") != "TRUE") #%>% distinct(is_federal_pac)
#pac_review %>% filter(!is.na(pac_id)) %>% group_by(pac_id) %>% filter(n_distinct(industry_override_under_5k) > 1)
pac_review %>% filter(!is.na(fec_id)) %>% group_by(fec_id) %>% filter(n_distinct(industry_override_under_5k) > 1)
#pac_review %>% filter(!is.na(pac_id)) %>% group_by(pac_id) %>% filter(n_distinct(fec_id) > 1 | n_distinct(is_federal_pac) > 1) %>% arrange(pac_id)
```



# Sector mapping

`industry_override_under_5k` is a manually assigned sector description.
The description needs to be matched to a code. Match against
the granular `subsector_description` first then to the broader
`sector_description`

Read in google sheet that tracks sector descriptions and codes. 

```{r}
gs4_deauth()
sector_map_full <- read_sheet(
  "https://docs.google.com/spreadsheets/d/11QHvNJsdtMlc1YKo_iNvMB_Jfn5Ui-iYdlWhFYjSm9g/edit?gid=605391596#gid=605391596"
) %>%
  mutate(across(everything(), as.character))

subsector_lookup <- sector_map_full %>%
  filter(!is.na(subsector_description)) %>%
  distinct(subsector_description, .keep_all = TRUE) %>%
  select(subsector_description, sector)

sector_lookup <- sector_map_full %>%
  filter(!is.na(sector_description)) %>%
  distinct(sector_description, .keep_all = TRUE) %>%
  select(sector_description, sector)
```

# pac_review lookup tables

```{r}
# PACs that cannot be classified yet — leave their full amount as uncategorized
# rows with no pac_id excluded here
skip_pac_ids <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  filter(
    unresolvable_reason == "additional_review_req" |
    is.na(industry_override_under_5k)              |
    industry_override_under_5k == ""               |
    industry_override_under_5k == "PENDING"
  ) %>%
  pull(pac_id) %>%
  as.character()
length(skip_pac_ids)
```


```{r}
pac_override <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "", !pac_id %in% skip_pac_ids) %>%
  left_join(subsector_lookup, by = c("industry_override_under_5k" = "subsector_description")) %>%
  left_join(sector_lookup, by = c("industry_override_under_5k" = "sector_description"),
            suffix = c("", "_broad")) %>%
  mutate(pac_id = as.character(pac_id)) %>%
  
  # use pac_classifiability for amounts — should reflect the most recent data pull
  left_join(
    pac_classifiability %>% transmute(
      pac_id   = as.character(pac_id),
      # unitemized + entities that contributed <$5k in aggregate counted as under_5k
      under_5k = coalesce(as.numeric(under_5k_amount), 0) +
                 coalesce(as.numeric(unitemized_amount), 0)
    ),
    by = "pac_id"
  ) %>%
  mutate(
    override_sector = coalesce(sector, sector_broad),
    is_federal      = (!is.na(fec_id) & trimws(coalesce(fec_id, "")) != "") |
                      coalesce(is_federal_pac, "") == "TRUE"
  ) %>%
  select(pac_id, override_sector, industry_override_under_5k, under_5k, is_federal)

# warning for descriptions that didn't match a sector code
unmapped <- pac_override %>% filter(is.na(override_sector))
if (nrow(unmapped) > 0) {
  cat("Warning: no sector code found for these industry_override_under_5k values:\n")
  print(unmapped %>% select(pac_id, industry_override_under_5k))
}

pac_override_map <- setNames(seq_len(nrow(pac_override)), pac_override$pac_id)
```


```{r}
# name-based override for pac_review rows with no pac_id 
# normalize to uppercase for case-insensitive matching against contributor_name in over_5k_tagged
skip_pac_names <- pac_review %>%
  filter(is.na(pac_id) | pac_id == "") %>%
  filter(
    unresolvable_reason == "additional_review_req" |
    is.na(industry_override_under_5k)              |
    industry_override_under_5k == ""               |
    industry_override_under_5k == "PENDING"
  ) %>%
  mutate(name_key = toupper(trimws(Contributor.Name))) %>%
  pull(name_key)

pac_name_override <- pac_review %>%
  filter(is.na(pac_id) | pac_id == "") %>%
  mutate(name_key = toupper(trimws(Contributor.Name))) %>%
  filter(!name_key %in% skip_pac_names) %>%
  left_join(subsector_lookup, by = c("industry_override_under_5k" = "subsector_description")) %>%
  left_join(sector_lookup, by = c("industry_override_under_5k" = "sector_description"),
            suffix = c("", "_broad")) %>%
  mutate(
    override_sector = coalesce(sector, sector_broad),
    under_5k        = coalesce(under_5k_amount, 0) + coalesce(as.numeric(unitemized_amount), 0),
    is_federal      = (!is.na(fec_id) & trimws(coalesce(fec_id, "")) != "") |
                      coalesce(is_federal_pac, "") == "TRUE",
    is_candidate    = trimws(coalesce(industry_override_under_5k, "")) == "Candidates/Political Parties"
  ) %>%
  filter(!is.na(override_sector)) %>%
  distinct(name_key, .keep_all = TRUE) %>%
  select(name = Contributor.Name, name_key, override_sector, industry_override_under_5k, under_5k, is_federal, is_candidate)

# keyed on uppercase name for case-insensitive lookup in the distribution loop
pac_name_override_map <- setNames(seq_len(nrow(pac_name_override)), pac_name_override$name_key)

unmapped_names <- pac_review %>%
  filter(is.na(pac_id) | pac_id == "") %>%
  mutate(name_key = toupper(trimws(Contributor.Name))) %>%
  filter(!name_key %in% skip_pac_names,
         !name_key %in% pac_name_override$name_key)
if (nrow(unmapped_names) > 0) {
  cat("Warning: no sector code for these name-only pac_review entries:\n")
  print(unmapped_names %>% select(Contributor.Name, industry_override_under_5k))
}
```

```{r}
# broader name-to-override covering ALL resolved pac_review entries (both id-based and no-id).
# used in the distribution loop for pac_no_id / pac_unmatched_id contributors whose
# pac_review entry may have gained a pac_id since over_5k_tagged was generated —
# those entries are now in pac_override (keyed by id), not pac_name_override (no-id only),
# so they would otherwise fall through to "99" on a name lookup.
all_pac_review_by_name <- pac_review %>%
  filter(
    !(coalesce(unresolvable_reason, "") == "additional_review_req" |
      is.na(industry_override_under_5k)                            |
      trimws(coalesce(industry_override_under_5k, "")) %in% c("", "PENDING"))
  ) %>%
  mutate(name_key = toupper(trimws(coalesce(Contributor.Name, "")))) %>%
  filter(name_key != "") %>%
  left_join(subsector_lookup, by = c("industry_override_under_5k" = "subsector_description")) %>%
  left_join(sector_lookup, by = c("industry_override_under_5k" = "sector_description"),
            suffix = c("", "_broad")) %>%
  mutate(override_sector = coalesce(sector, sector_broad)) %>%
  filter(!is.na(override_sector)) %>%
  distinct(name_key, .keep_all = TRUE) %>%
  select(name_key, override_sector)

all_pac_name_override_map <- setNames(all_pac_review_by_name$override_sector,
                                      all_pac_review_by_name$name_key)
cat(sprintf("Broader name override: %d resolved pac_review entries keyed by name\n",
            nrow(all_pac_review_by_name)))
```


# Over-$5k individual/org classification lookup

`pac_class_input` is data file produced by 0803 that includes the $5k+
contributors to each PAC in the PAC graph that need to be classified. 

Use over_5k_tagged for amounts; build a per-entity
code lookup from pac_class_input + pac_final (best non-"99" code wins when
the same entity_group maps to multiple uuids across different classification runs).


**Note:** There should only be one code per `classification_unit_id`, but there
may be rare exceptions. E.g. if there were two John Smiths who are retired 
they'd have the same `classification_unit_it`, but its determined that these
are two different people and one John Smith is assigned a different code during the manual
review because the reviewer determined he's associated with a different industry (according to 
other contributions he may have made). This is very unlikely, but if there are legitimate conflicts within
`classiciation_unit_id` after a manual review, may need to adapt this part to expand to contribution-level 
records. 


```{r}
# join classifications to pac_class_input at classification_unit_id level
pac_class_coded <- pac_class_input %>%
  filter(contributor_type %in% c("individual_or_org", "major_donor_ie_committee")) %>%
  mutate(
    parent_pac_id = as.character(parent_pac_id),
    Amount        = as.numeric(Amount) # transaction-level amount
  ) %>%
  left_join(pac_final %>% select(unit_id, code_final), by = c("uuid" = "unit_id")) %>%
  mutate(industry = coalesce(code_final, "99"))

# diagnostic: flag classification_unit_ids with >1 distinct non-"99" code — should never happen
unit_code_conflicts <- pac_class_coded %>%
  group_by(parent_pac_id, classification_unit_id) %>%
  filter(n_distinct(industry[industry != "99"]) > 1) %>%
  summarise(codes = paste(sort(unique(industry)), collapse = " | "), .groups = "drop")

if (nrow(unit_code_conflicts) > 0) {
  cat(sprintf("WARNING: %d classification_unit_id(s) with conflicting codes:\n", nrow(unit_code_conflicts)))
  print(unit_code_conflicts)
} else {
  cat("No classification_unit_id code conflicts.\n")
}

# diagnostic: flag entity_groups with >1 distinct non-"99" code across their classification units
# just to see - shouldn't be an issue if this is true in some cases
pac_class_coded %>%
  group_by(parent_pac_id, entity_group) %>%
  filter(n_distinct(industry[industry != "99"]) > 1) %>%
  summarise(
    n_units = n_distinct(classification_unit_id),
    codes   = paste(sort(unique(industry)), collapse = " | "),
    .groups = "drop"
  )
```


```{r}
# one code per (parent_pac_id, classification_unit_id) — most common non-"99" code wins
# but shouldn't be any conflicts
unit_code_lookup <- pac_class_coded %>%
  group_by(parent_pac_id, classification_unit_id) %>%
  summarise(
    entity_group = entity_group[1],
    industry = {
      non_99 <- industry[industry != "99"]
      if (length(non_99) > 0) names(sort(table(non_99), decreasing = TRUE))[1] else "99"
    },
    .groups = "drop"
  )

# amounts per (parent_pac_id, classification_unit_id) — summed from raw transactions
unit_amounts <- pac_class_coded %>%
  group_by(parent_pac_id, classification_unit_id) %>%
  summarise(unit_amount = sum(Amount, na.rm = TRUE), .groups = "drop")

# entity_code_lookup kept for reference (collapses to one code per entity_group)
entity_code_lookup <- unit_code_lookup %>%
  group_by(parent_pac_id, entity_group) %>%
  summarise(
    industry = {
      non_99 <- industry[industry != "99"]
      if (length(non_99) > 0) names(sort(table(non_99), decreasing = TRUE))[1] else "99"
    },
    .groups = "drop"
  )

# over_5k_direct: only individual/org and major donor contributors — PAC-type contributors
# (federal_pac, candidate_pac, pac_no_id, pac_unmatched_id) are handled in the distribution loop
over_5k_direct <- over_5k_tagged %>%
  filter(contributor_type %in% c("individual_or_org", "major_donor_ie_committee")) %>%
  mutate(
    contributor_amount = as.numeric(contributor_amount),
    parent_pac_id      = as.character(parent_pac_id)
  ) %>%
  left_join(
    unit_code_lookup %>% select(parent_pac_id, entity_group, classification_unit_id, industry),
    by = c("parent_pac_id", "entity_group")
  ) %>%
  left_join(unit_amounts, by = c("parent_pac_id", "classification_unit_id")) %>%
  mutate(
    contributor_amount = coalesce(unit_amount, contributor_amount),
    industry           = coalesce(industry, "99")
  ) %>%
  select(-unit_amount)

# diagnostic: individual/org entity_groups with no classification_unit_id match in pac_class_input
missing_classifications <- over_5k_direct %>% filter(is.na(classification_unit_id))
if (nrow(missing_classifications) > 0) {
  cat(sprintf("WARNING: %d individual/org contributor(s) have no classification_unit_id — will be coded '99':\n",
              nrow(missing_classifications)))
  print(missing_classifications %>%
    select(parent_pac_id, entity_group, contributor_name, contributor_amount, contributor_type))
} else {
  cat("All individual/org contributors matched to a classification unit.\n")
}

# amount per pac/industry for direct contributors
direct_by_pac_list <- over_5k_direct %>%
  group_by(parent_pac_id, industry) %>%
  summarise(amount = sum(contributor_amount, na.rm = TRUE), .groups = "drop") %>%
  split(~parent_pac_id)
```

```{r}
# location variants and contribution counts per parent_pac_id x entity_group
# for the PAC contributor audit

egd_input <- pac_class_input %>%
  mutate(parent_pac_id = as.character(parent_pac_id))
for (col in c("Contributor.City", "Contributor.State", "zip_code_processed",
              "processed_employer_name")) {
  if (!col %in% names(egd_input)) egd_input[[col]] <- NA_character_
}

entity_group_details <- egd_input %>%
  group_by(parent_pac_id, entity_group) %>%
  summarise(
    n_contributions      = n(),
    contributor_city     = paste(sort(unique(na.omit(Contributor.City))),        collapse = " | "),
    contributor_state    = paste(sort(unique(na.omit(Contributor.State))),       collapse = " | "),
    contributor_zip      = paste(sort(unique(na.omit(zip_code_processed))),      collapse = " | "),
    contributor_employer = paste(sort(unique(na.omit(processed_employer_name))), collapse = " | "),
    .groups = "drop"
  )
entity_group_details_by_pac <- split(entity_group_details, entity_group_details$parent_pac_id)
```


# Compute PAC industry distributions

```{r}
over_5k_by_pac <- split(
  over_5k_tagged %>% mutate(
    parent_pac_id  = as.character(parent_pac_id),
    contributor_id = as.character(contributor_id)
  ),
  ~ as.character(parent_pac_id)
)

pac_deps <- over_5k_tagged %>%
  filter(contributor_type %in% c("pac"), !is.na(contributor_id)) %>%
  mutate(parent_pac_id  = as.character(parent_pac_id),
         contributor_id = as.character(contributor_id)) %>%
  distinct(parent_pac_id, contributor_id) %>%
  group_by(parent_pac_id) %>%
  summarise(deps = list(contributor_id), .groups = "drop") %>%
  { setNames(.$deps, .$parent_pac_id) }
```

```{r}
industry_distrib <- list()
pac_attribution  <- list()

compute_industry_distrib <- function(pid) {
  pid       <- as.character(pid)
  attr_rows <- list()

  # helper: append one attribution row to attr_rows
  add_attr <- function(contributor_name, contributor_id, contributor_type,
                       contributor_amount, industry, allocated_amount,
                       eg_detail = NULL) {
    attr_rows[[length(attr_rows) + 1]] <<- tibble(
      pac_id               = pid,
      contributor_name     = contributor_name,
      contributor_id       = contributor_id,
      contributor_type     = contributor_type,
      n_contributions      = if (!is.null(eg_detail)) eg_detail$n_contributions[1]  else NA_integer_,
      contributor_city     = if (!is.null(eg_detail)) eg_detail$contributor_city[1] else NA_character_,
      contributor_state    = if (!is.null(eg_detail)) eg_detail$contributor_state[1] else NA_character_,
      contributor_zip      = if (!is.null(eg_detail)) eg_detail$contributor_zip[1]  else NA_character_,
      contributor_employer = if (!is.null(eg_detail)) eg_detail$contributor_employer[1] else NA_character_,
      contributor_amount   = contributor_amount,
      industry             = industry,
      allocated_amount     = allocated_amount
    )
  }

  # (1) pending / needs review — entire amount stays uncategorized
  if (pid %in% skip_pac_ids) {
    pac_row <- pac_classifiability[pac_classifiability$pac_id == pid, ]
    total   <- if (nrow(pac_row) > 0) coalesce(pac_row$total_received[1], 0) else 0
    industry_distrib[[pid]] <<- if (total > 0) c("99" = total) else numeric(0)
    if (total > 0)
      add_attr("(pending/unreviewed)", NA_character_, "unreviewed", total, "99", total)
    pac_attribution[[pid]] <<- bind_rows(attr_rows)
    return(invisible(NULL))
  }

  ovr <- if (pid %in% names(pac_override_map)) pac_override[pac_override_map[[pid]], ] else NULL

  # (2) federal PAC — entire amount classified by override sector
  if (!is.null(ovr) && isTRUE(ovr$is_federal)) {
    pac_row <- pac_classifiability[pac_classifiability$pac_id == pid, ]
    total   <- if (nrow(pac_row) > 0) coalesce(pac_row$total_received[1], 0) else 0
    code    <- coalesce(ovr$override_sector, "99")
    industry_distrib[[pid]] <<- if (total > 0) setNames(total, code) else numeric(0)
    if (total > 0)
      add_attr(NA_character_, NA_character_, "federal_pac_override", total, code, total)
    pac_attribution[[pid]] <<- bind_rows(attr_rows)
    return(invisible(NULL))
  }

  # (3) Candidates/Political Parties: entire amount categorized as candidates/political parties
  if (!is.null(ovr) && !is.na(ovr$industry_override_under_5k) &&
      trimws(ovr$industry_override_under_5k) == "Candidates/Political Parties") {
    pac_row <- pac_classifiability[pac_classifiability$pac_id == pid, ]
    total   <- if (nrow(pac_row) > 0) coalesce(pac_row$total_received[1], 0) else 0
    code    <- coalesce(ovr$override_sector, "99")
    industry_distrib[[pid]] <<- if (total > 0) setNames(total, code) else numeric(0)
    if (total > 0)
      add_attr(NA_character_, NA_character_, "candidate_pac_override", total, code, total)
    pac_attribution[[pid]] <<- bind_rows(attr_rows)
    return(invisible(NULL))
  }

  # (4) non-federal PAC: over-$5k direct + recursive PAC-type + under-$5k override

  direct    <- direct_by_pac_list[[pid]] # get direct contributors
  ind_amounts <- if (!is.null(direct) && nrow(direct) > 0)
    setNames(direct$amount, direct$industry) else numeric(0)

  # direct individual/org contributors: one attribution row per contributor
  eg_lookup <- entity_group_details_by_pac[[pid]]
  direct_contribs <- over_5k_direct %>% filter(parent_pac_id == pid)
  for (i in seq_len(nrow(direct_contribs))) {
    r  <- direct_contribs[i, ]
    eg <- if (!is.null(eg_lookup))
      eg_lookup[eg_lookup$entity_group == r$entity_group, ] else NULL
    add_attr(r$contributor_name, r$contributor_id, r$contributor_type,
             r$contributor_amount, coalesce(r$industry, "99"), r$contributor_amount,
             eg_detail = if (!is.null(eg) && nrow(eg) > 0) eg else NULL)
  }

  pac_rows <- over_5k_by_pac[[pid]]
  if (!is.null(pac_rows)) {
    all_pac_rows <- pac_rows %>% filter(contributor_type == "pac")

    # PAC contributors with a committee id: distribute recursively
    id_rows <- all_pac_rows %>% filter(!is.na(contributor_id) & contributor_id != "")
    for (i in seq_len(nrow(id_rows))) {
      child_id     <- id_rows$contributor_id[i]
      child_amount <- id_rows$contributor_amount[i]
      child_dist   <- industry_distrib[[child_id]]
      if (is.null(child_dist) || length(child_dist) == 0) {
        ind_amounts["99"] <- coalesce(ind_amounts["99"], 0) + child_amount
        add_attr(id_rows$contributor_name[i], child_id, "child_pac",
                 child_amount, "99", child_amount)
        next
      }
      child_total <- sum(child_dist)
      if (child_total <= 0) next
      scaled <- child_dist / child_total * child_amount
      for (ind in names(scaled)) {
        ind_amounts[ind] <- coalesce(ind_amounts[ind], 0) + scaled[ind]
        add_attr(id_rows$contributor_name[i], child_id, "child_pac",
                 child_amount, ind, scaled[ind])
      }
    }

    # PAC contributors with no id: look up override sector by name
    no_id_rows <- all_pac_rows %>% filter(is.na(contributor_id) | contributor_id == "")
    for (i in seq_len(nrow(no_id_rows))) {
      child_name     <- no_id_rows$contributor_name[i]
      child_name_key <- toupper(trimws(child_name))
      child_amount   <- no_id_rows$contributor_amount[i]
      if (child_name_key %in% skip_pac_names || !child_name_key %in% names(pac_name_override_map)) {
        ind_amounts["99"] <- coalesce(ind_amounts["99"], 0) + child_amount
        add_attr(child_name, NA_character_, "child_pac_no_id",
                 child_amount, "99", child_amount)
        next
      }
      code <- pac_name_override$override_sector[pac_name_override_map[[child_name_key]]]
      code <- coalesce(code, "99")
      ind_amounts[code] <- coalesce(ind_amounts[code], 0) + child_amount
      add_attr(child_name, NA_character_, "child_pac_no_id",
               child_amount, code, child_amount)
    }

    # federal_pac, candidate_pac, pac_no_id, pac_unmatched_id contributors: classified via pac_review override
    # lookup order: 
    # (1) contributor_id match in pac_override_map,
    # (2) name match in all_pac_name_override_map — covers formerly-no-id entries that gained a pac_id in review 
    other_pac_rows <- pac_rows %>%
      filter(contributor_type %in% c("federal_pac", "candidate_pac", "pac_no_id", "pac_unmatched_id"))

    for (i in seq_len(nrow(other_pac_rows))) {
      r              <- other_pac_rows[i, ]
      child_id       <- as.character(coalesce(r$contributor_id, ""))
      child_name_key <- toupper(trimws(coalesce(r$contributor_name, "")))
      child_amount   <- r$contributor_amount

      code <- if (child_id != "" && child_id %in% names(pac_override_map)) {
        coalesce(pac_override$override_sector[pac_override_map[[child_id]]], "99")
      } else if (child_name_key %in% names(all_pac_name_override_map)) {
        all_pac_name_override_map[[child_name_key]]
      } else {
        "99"
      }

      ind_amounts[code] <- coalesce(ind_amounts[code], 0) + child_amount
      add_attr(r$contributor_name, r$contributor_id, r$contributor_type,
               child_amount, code, child_amount)
    }
  }

  # under-$5k override
  if (!is.null(ovr) && !is.na(ovr$under_5k) && ovr$under_5k > 0) {
    code <- coalesce(ovr$override_sector, "99")
    ind_amounts[code] <- coalesce(ind_amounts[code], 0) + ovr$under_5k
    add_attr("(under $5k contributors)", NA_character_, "under_5k",
             ovr$under_5k, code, ovr$under_5k)
  } else {
    # PAC not yet in pac_review (no manual override): classify all under-$5k as 99/uncategorized
    pac_row <- pac_classifiability[pac_classifiability$pac_id == pid, ]
    if (nrow(pac_row) > 0) {
      under_5k_amt <- coalesce(as.numeric(pac_row$under_5k_amount[1]), 0) +
                      coalesce(as.numeric(pac_row$unitemized_amount[1]), 0) #+
                      #coalesce(as.numeric(pac_row$small_dollar_amount[1]), 0)
      if (under_5k_amt > 0) {
        ind_amounts["99"] <- coalesce(ind_amounts["99"], 0) + under_5k_amt
        add_attr("(under $5k contributors)", NA_character_, "under_5k",
                 under_5k_amt, "99", under_5k_amt)
      }
    }
  }

  industry_distrib[[pid]] <<- ind_amounts
  pac_attribution[[pid]]  <<- bind_rows(attr_rows)
}
```


```{r}
# federal and candidate PACs run first — status may have been updated during
# manual review after ordered_pacs was written in 0803
federal_candidate_ids <- unique(c(
  pac_classifiability %>%
    filter(
      coalesce(is_federal_pac, "FALSE") == "TRUE" |
      coalesce(is_candidate_pac, "FALSE") == "TRUE"
    ) %>%
    pull(pac_id) %>% as.character(),
  pac_override %>%
    filter(
      coalesce(is_federal, FALSE) |
      trimws(coalesce(industry_override_under_5k, "")) == "Candidates/Political Parties"
    ) %>%
    pull(pac_id) %>% as.character()
))
cat(sprintf("Processing %d federal/candidate PAC(s) first\n", length(federal_candidate_ids)))
for (pid in federal_candidate_ids) compute_industry_distrib(pid)

# major donor/IE committees are classified via individual path (over_5k_direct),
# not the PAC distribution loop — skip them entirely
major_donor_ie_ids <- pac_classifiability %>%
  mutate(pac_id = as.character(pac_id)) %>%
  filter(coalesce(is_major_donor_ie, "FALSE") == "TRUE") %>%
  pull(pac_id)
cat(sprintf("Skipping %d major donor/IE committee(s) — classified via individual path\n",
            length(major_donor_ie_ids)))

# main pass: skip federal/candidate and major donor/IE PACs
for (pid in ordered_pacs) {
  if (pid %in% federal_candidate_ids) next
  if (pid %in% major_donor_ie_ids) next
  compute_industry_distrib(pid)
}

remaining <- setdiff(stuck_pacs, c(federal_candidate_ids, major_donor_ie_ids))
while (length(remaining) > 0) {
  ready <- Filter(function(pid) {
    deps <- pac_deps[[pid]]
    is.null(deps) || !any(deps %in% remaining)
  }, remaining)
  if (length(ready) == 0) ready <- remaining
  for (pid in ready) compute_industry_distrib(pid)
  remaining <- setdiff(remaining, ready)
}
```



# Output

```{r}
pac_name_lookup <- pac_review %>%
  filter(!is.na(pac_id) & pac_id != "") %>%
  mutate(pac_id = as.character(pac_id)) %>%
  group_by(pac_id) %>%
  slice(1) %>%
  ungroup() %>%
  select(pac_id, pac_name = Contributor.Name)

# distributions for pacs w/ ids
id_breakdown <- lapply(names(industry_distrib), function(pid) {
  ind <- industry_distrib[[pid]]
  if (length(ind) == 0) return(NULL)
  tibble(pac_id = pid, industry = names(ind), amount = unname(ind))
}) %>%
  bind_rows() %>%
  left_join(pac_name_lookup, by = "pac_id") %>%
  left_join(
    pac_classifiability %>%
      transmute(pac_id = as.character(pac_id), total_received,
                entity_id_normalized = as.character(entity_id_normalized)),
    by = "pac_id"
  ) %>%
  mutate(pct_of_total = amount / total_received)

# no-id PAC distributions.
# reuses pac_name_override (sector lookups already done, is_federal/is_candidate flags set).
# federal/candidate: pct_of_total = 1 to their override industry (100% classifiable by override).
# all others: industry = "99", pct_of_total = 1 — explicitly uncategorized; contributions to
# candidates cannot be split without traversal data and will remain "99" until the PAC
# receives an ID and 0803 traverses it
# unmapped_names adds entries whose override description didn't map to a sector code —
# they are also explicitly "99" and appear in pac_industry_breakdown for the audit
name_breakdown <- bind_rows(
  pac_name_override %>%
    transmute(
      pac_id               = NA_character_,
      pac_name             = name,
      entity_id_normalized = NA_character_,
      industry             = case_when(
        is_federal | is_candidate ~ coalesce(override_sector, "99"),
        TRUE                      ~ "99"
      ),
      amount         = NA_real_,
      total_received = NA_real_,
      pct_of_total   = 1
    ),
  unmapped_names %>%
    transmute(
      pac_id               = NA_character_,
      pac_name             = Contributor.Name,
      entity_id_normalized = NA_character_,
      industry             = "99",
      amount               = NA_real_,
      total_received       = NA_real_,
      pct_of_total         = 1
    )
)

pac_industry_breakdown <- bind_rows(id_breakdown, name_breakdown) %>%
  select(pac_id, pac_name, entity_id_normalized, industry, amount, total_received, pct_of_total) %>%
  arrange(pac_id, pac_name, desc(amount))

# seed PACs that have a manually
# assigned industry_override_under_5k in pac_review but produced no
# industry_distrib rows. The label is mapped to a sector code via the same
# subsector_lookup / sector_lookup built above, then a 100% row is synthesized
# using total_received from pac_classifiability.
depth0_fallback <- pac_classifiability %>%
  mutate(pac_id = as.character(pac_id)) %>%
  filter(!pac_id %in% pac_industry_breakdown$pac_id) %>%
  inner_join(
    pac_override %>%
      filter(!is.na(industry_override_under_5k),
             industry_override_under_5k != "",
             industry_override_under_5k != "PENDING") %>%
      select(pac_id, industry_override_under_5k),
    by = "pac_id"
  ) %>%
  left_join(subsector_lookup, by = c("industry_override_under_5k" = "subsector_description")) %>%
  left_join(sector_lookup,    by = c("industry_override_under_5k" = "sector_description"),
            suffix = c("", "_broad")) %>%
  mutate(mapped_code = coalesce(sector, sector_broad, "99"),
         entity_id_normalized = as.character(entity_id_normalized)) %>%
  left_join(pac_name_lookup, by = "pac_id") %>%
  transmute(
    pac_id,
    pac_name,
    entity_id_normalized,
    industry       = mapped_code,
    amount         = NA_real_,
    total_received = NA_real_,
    pct_of_total   = 1
  )

pac_industry_breakdown <- bind_rows(pac_industry_breakdown, depth0_fallback)
cat(sprintf("Depth-0 fallback: added %d PACs (%s)\n",
            nrow(depth0_fallback),
            paste(depth0_fallback$pac_id, collapse = ", ")))

# sanity check: every fallback PAC should be either a federal PAC or one with
# no contribution data (total_received is NA or 0). Anything else is unexpected.
fallback_check <- depth0_fallback %>%
  select(pac_id) %>%
  left_join(
    pac_classifiability %>% transmute(pac_id = as.character(pac_id),
                                      total_received = as.numeric(total_received)),
    by = "pac_id"
  ) %>%
  left_join(
    pac_override %>% select(pac_id, is_federal),
    by = "pac_id"
  ) %>%
  filter(!isTRUE(is_federal) & !is.na(total_received) & total_received > 0)

if (nrow(fallback_check) > 0) {
  warning(sprintf(
    "depth0_fallback contains %d unexpected PAC(s) that are non-federal and have contribution data — investigate why they were not processed:\n%s",
    nrow(fallback_check),
    paste(fallback_check$pac_id, collapse = ", ")
  ))
} else {
  cat("Fallback check passed: all fallback PACs are federal or have no contribution data.\n")
}

# warn if federal/candidate PAC contributor IDs from over_5k_tagged are missing from
# pac_classifiability — these will not be processed by compute_industry_distrib and
# will be absent from pac_industry_breakdown. Should resolve once 0803 traverses them.
fed_cand_missing_from_clfy <- over_5k_tagged %>%
  filter(contributor_type %in% c("federal_pac", "candidate_pac"),
         !is.na(contributor_id), contributor_id != "") %>%
  distinct(contributor_id, contributor_type) %>%
  filter(!contributor_id %in% pac_classifiability$pac_id)

if (nrow(fed_cand_missing_from_clfy) > 0) {
  cat(sprintf(
    "WARNING: %d federal/candidate PAC contributor ID(s) in over_5k_tagged are not in pac_classifiability and will be absent from pac_industry_breakdown:\n",
    nrow(fed_cand_missing_from_clfy)
  ))
  print(fed_cand_missing_from_clfy %>% arrange(contributor_type, contributor_id))
} else {
  cat("All federal/candidate PAC contributor IDs in over_5k_tagged are present in pac_classifiability.\n")
}
```


```{r, eval = FALSE}
# data check
depth0_diagnostic <- depth0_fallback %>%
  mutate(
    in_ordered       = pac_id %in% ordered_pacs,
    in_stuck         = pac_id %in% stuck_pacs,
    distrib_len      = sapply(pac_id, function(pid) length(industry_distrib[[pid]])),
    in_pac_override  = pac_id %in% pac_override$pac_id
  ) %>%
  left_join(
    pac_classifiability %>%
      transmute(pac_id = as.character(pac_id),
                under_5k_amount = as.numeric(under_5k_amount),
                unitemized_amount = as.numeric(unitemized_amount),
                total_received),
    by = "pac_id"
  )

depth0_diagnostic %>%
  count(in_ordered, in_stuck, distrib_len > 0, in_pac_override)

```


# WRITE OUTPUT

```{r}
write_csv(pac_industry_breakdown, "04_outputs/pac_industry_breakdown.csv")
```

```{r}
# PAC contributor audit: one row per PAC × contributor × industry
# for child PAC contributors, rows are exploded across industry distribution
# contributor_type values:
#   individual_or_org / major_donor_ie_committee — directly classified contributors
#   child_pac / child_pac_no_id — PAC contributors (with / without a committee id)
#   under_5k — bucket for under-$5k, unitemized, and small-dollar amounts
#   federal_pac_override / candidate_pac_override / unreviewed — whole-PAC override cases

today_str <- format(Sys.Date(), "%m-%d-%y")

pac_contributor_audit <- bind_rows(pac_attribution) %>%
  left_join(
    pac_classifiability %>%
      transmute(pac_id = as.character(pac_id), total_pac_received = total_received),
    by = "pac_id"
  ) %>%
  left_join(pac_name_lookup, by = "pac_id") %>%
  select(pac_id, pac_name, contributor_name, contributor_id, contributor_type,
         n_contributions, contributor_city, contributor_state, contributor_zip,
         contributor_employer, contributor_amount, industry, allocated_amount,
         total_pac_received) %>%
  arrange(pac_id, contributor_type, desc(contributor_amount))

audit_path <- paste0("04_outputs/pac_contributor_audit_", today_str, ".csv")
write_csv(pac_contributor_audit, audit_path)
cat(sprintf("PAC contributor audit: %d rows across %d PACs → %s\n",
            nrow(pac_contributor_audit),
            n_distinct(pac_contributor_audit$pac_id, na.rm = TRUE),
            audit_path))
```
