---
title: "Build Classification Input"
output: html_notebook
---

# Summary
      
This notebook is run after entity resultion and name mapping has been completed. 
It aggregates contributors by their `Normalized` name to determine which
entities meet the $5k threshold for classification. 

Unique ids from previously classified data are matched back, then the notebook
separates PAC contributors (to feed to the pac graph in 0803) from direct
individual/organization contributors that will be classified directly. 

This only looks at DIRECT contributors to the candidates and props. Contributors
to props are handled in 0803. 

# Load packages

```{r}
here::i_am("0102c_build_classification_input.Rmd")
library(here)
library(dplyr)
library(stringr)
library(stringdist)
library(fuzzyjoin)
library(reclin2)
library(phonics)
library(igraph)
library(readr)
library(digest)
library(jsonlite)
library(tidyr)
library(uuid)

source("standardization_helpers.R")
```

## Define helper functions

```{r}
# makes contributor_id
make_row_hash <- function(...) {
  cols <- list(...)
  n <- length(cols[[1]])
  vapply(seq_len(n), function(i) {
    parts <- vapply(cols, function(col) as.character(col[i]), character(1))
    digest::digest(paste(parts, collapse = "|"), algo = "xxhash32")
  }, character(1))
}

# recipient_key for applying threshold checks:
#   PAC contributions        -> Recipient.Committee.ID (each PAC evaluated separately)
#   ballot measure           -> name of Ballot.Measure (groups all committees for same measure/side)
#   candidate                -> "Recipient.Name | Office" key for candidate race (Ben Allen has more than one committee)
add_recipient_key <- function(df) {
  df %>% mutate(
    recipient_key = case_when(
      !is.na(Ballot.Measure) & Ballot.Measure != "" ~ Ballot.Measure,
      !is.na(Office) & Office != ""                 ~ paste(Recipient.Name, Office, sep = " | "),
      TRUE                                          ~ Recipient.Name
    )
  )
}

# look up contributor name in entity registry
# returns entity_id, is_conflict and is_new
lookup_cluster_uuid <- function(raw_names, mapping_df) {
  clean <- unique(toupper(raw_names))
  matched_uuids <- mapping_df %>%
    filter(raw_name %in% clean) %>%
    pull(entity_uuid) %>%
    unique()

  if (length(matched_uuids) == 0) {
    list(entity_id = UUIDgenerate(), is_conflict = FALSE, is_new = TRUE)
  } else if (length(matched_uuids) == 1) {
    list(entity_id = matched_uuids[1], is_conflict = FALSE, is_new = FALSE)
  } else {
    list(entity_id = matched_uuids[1], is_conflict = TRUE, is_new = FALSE)
  }
}

# assigns subgroup-level entity UUIDs to distinct individuals w/ the same name
assign_subgroup_entity_id <- function(df, default_col = "entity_uuid") {
  if (nrow(subgroup_profiles) == 0) return(df)
  df <- df %>% mutate(.match_row_id = row_number())
  candidates <- df %>%
    filter(Normalized %in% subgroup_profiles$canonical_name) %>%
    transmute(
      .match_row_id,
      canonical_name  = Normalized,
      cand_employer   = toupper(str_squish(coalesce(Contributor.Employer, ""))),
      cand_occupation = toupper(str_squish(coalesce(Contributor.Occupation, ""))),
      cand_city       = toupper(str_squish(coalesce(Contributor.City, ""))),
      cand_state      = toupper(str_squish(coalesce(Contributor.State, ""))),
      cand_zip        = coalesce(as.character(Contributor.Zip.Code), ""),
      cand_committee  = coalesce(as.character(Recipient.Committee.ID), "")
    ) %>%
    inner_join(subgroup_profiles, by = "canonical_name", relationship = "many-to-many") %>%
    mutate(
      # look for matching employer, occupation, etc. -> used to distinguish individuals
      is_match =
        (is.na(profile_employer)   | profile_employer   == cand_employer)   &
        (is.na(profile_occupation) | profile_occupation == cand_occupation) &
        (is.na(profile_city)       | profile_city       == cand_city)       &
        (is.na(profile_state)      | profile_state      == cand_state)      &
        (is.na(profile_zip)        | profile_zip        == cand_zip)        &
        (is.na(profile_committee)  | profile_committee  == cand_committee)
    ) %>%
    filter(is_match) %>%
    distinct(.match_row_id, canonical_name, subgroup_id)
  ambiguous <- candidates %>%
    group_by(.match_row_id) %>%
    filter(n_distinct(subgroup_id) > 1) %>%
    ungroup()
  if (nrow(ambiguous) > 0) {
    warning(sprintf(
      "%d contribution record(s) matched more than one disambiguation subgroup - left unresolved (default uuid used)",
      n_distinct(ambiguous$.match_row_id)
    ))
    print(
      df %>% inner_join(ambiguous %>% distinct(.match_row_id), by = ".match_row_id") %>%
        select(Contributor.Name, Contributor.Employer, Contributor.Occupation,
               Contributor.City, Contributor.State, Contributor.Zip.Code, Recipient.Committee.ID)
    )
  }
  resolved <- candidates %>%
    anti_join(ambiguous, by = ".match_row_id") %>%
    distinct(.match_row_id, canonical_name, subgroup_id) %>%
    left_join(subgroup_uuid_lookup, by = c("canonical_name", "subgroup_id")) %>%
    select(.match_row_id, subgroup_entity_uuid = entity_uuid)
  df %>%
    left_join(resolved, by = ".match_row_id") %>%
    mutate(!!default_col := coalesce(subgroup_entity_uuid, .data[[default_col]])) %>%
    select(-.match_row_id, -subgroup_entity_uuid)
}
```

# Load and standardize data

Name of already-classified data file needs to be updated manually
with the most recent classifications. 

```{r}
# full contributions dataset
contributions_full <- read.csv("01_inputs/power_search_contributions_normalized.csv")
nrow(contributions_full)

# already classified contributors
categorized <- read.csv("01_inputs/already_classified_contributions_09-06-26.csv")
nrow(categorized)
```

```{r}
# load ID corrections data
library(googlesheets4)
gs4_deauth()

id_corrections <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1UnO3wldvEo84Md0ac5jbpg4SqWzvPle3yh15e_DGAVk",
  sheet = "corrected_pac_ids",
  col_types = "c"
) %>%
  mutate(
    Contributor.ID = coalesce(Contributor.ID, ""),
    resolved       = !is.na(resolved) & tolower(str_trim(resolved)) != "false",
    is_fed_pac     = tolower(str_trim(coalesce(is_fed_pac, ""))) %in% c("true", "1", "yes")
  )

contributions_full <- apply_contributor_corrections(contributions_full, id_corrections)
```



## Load name entity mapping and build subgroup structures

```{r}
# load name variants map
names_2026 <- read.csv(
  here("01_inputs/2026 Contributor name mapping and descriptions - Names.csv"),
  stringsAsFactors = FALSE
) %>%
  rename(raw_name = Original, canonical_name = Normalized) %>% # rename columns so all dataframes share col names
  filter(!is.na(raw_name), raw_name != "", !is.na(canonical_name), canonical_name != "") %>%
  mutate(
    raw_name       = (toupper(raw_name)), 
    canonical_name = str_squish(canonical_name) # canonical name shouldn't have any double spaces
  )
```


```{r}
# this file contains the uuids associated with each normalized name that has been mapped

NAME_ENTITY_MAPPING_PATH <- here("01_inputs/name_entity_mapping_rebuilt.csv")

# Fixed namespace for deterministic UUID generation (UUID v5 — same Normalized name always
# produces the same UUID across runs, for entities not explicitly tracked in the name map)
NORM_UUID_NAMESPACE <- "887a6583-910c-47b0-b373-d5b2863e8c8a"

name_entity_mapping <- read.csv(NAME_ENTITY_MAPPING_PATH, stringsAsFactors = FALSE)

# canonical name to UUID map (written in 0102)
canonical_uuid_map <- name_entity_mapping %>%
  filter(!is.na(canonical_name), confirmed == TRUE) %>%
  mutate(canonical_name = str_squish(canonical_name)) %>%
  distinct(canonical_name, entity_uuid)

if (!"subgroup_id" %in% names(names_2026)) names_2026$subgroup_id <- NA_character_

# get profile for each invdividual that was added from the "disambiguation" queue
# maps distinct individuals who share the same name
subgroup_profiles <- names_2026 %>%
  filter(!is.na(subgroup_id), subgroup_id != "") %>%
  transmute(
    canonical_name     = canonical_name,
    subgroup_id        = as.character(subgroup_id),
    profile_employer   = na_if(toupper(str_squish(coalesce(Contributor.Employer, ""))), ""),
    profile_occupation = na_if(toupper(str_squish(coalesce(Contributor.Occupation, ""))), ""),
    profile_city       = na_if(toupper(str_squish(coalesce(Contributor.City, ""))), ""),
    profile_state      = na_if(toupper(str_squish(coalesce(Contributor.State, ""))), ""),
    profile_zip        = na_if(coalesce(as.character(Contributor.Zip.Code), ""), ""),
    profile_committee  = na_if(coalesce(as.character(Recipient.Committee.ID), ""), "")
  ) %>%
  distinct()

# subgroup UUIDs — read from registry
subgroup_uuid_lookup <- name_entity_mapping %>%
  filter(source == "subgroup_disambig") %>%
  distinct(raw_name, entity_uuid) %>%
  separate(raw_name, into = c("canonical_name", "subgroup_id"), sep = "::", extra = "merge")
```

## Filter for races and props we are classifying

```{r}
# filter for gov and IC race + ALL props
contributions <- contributions_full %>% 
  rename(
    Ballot.Measure = `Ballot.Measure.s.`
  ) %>% 
  filter(
    Recipient.Name == "BECERRA, XAVIER" | 
    Recipient.Name == "HILTON, STEVE" | 
    (Recipient.Name == "ALLEN, BEN" & Office == "Insurance Commissioner") | 
    Recipient.Name == "KIM, JANE" | 
    (Ballot.Measure != "" & Ballot.Measure != "SUPPORTED: PROPOSITION 050 - ACA 8 (RIVAS) CONGRESSIONAL REDISTRICTING. (RES. CH. 156, 2025)" & Ballot.Measure != "OPPOSED: PROPOSITION 050 - ACA 8 (RIVAS) CONGRESSIONAL REDISTRICTING. (RES. CH. 156, 2025)") # include all ballot measures apart from prop 50 
) 
```


# Data processing

### apply id maps to standardize contributor id

```{r}
# id_map maps XREF_FILER_ID (Contributor.ID) to FILER_ID (Recipinet.Committee.ID)
id_map_early <- read_tsv("01_inputs/FILERNAME_CD.tsv", col_types = cols(.default = "c")) %>%
  mutate(FILER_ID = as.character(FILER_ID), XREF_FILER_ID = as.character(XREF_FILER_ID)) %>%
  filter(!is.na(FILER_ID), !is.na(XREF_FILER_ID)) %>%
  distinct(XREF_FILER_ID, FILER_ID,FILER_TYPE)


contributions <- contributions %>% 
  left_join(id_map_early, by = c("Contributor.ID" = "XREF_FILER_ID")) %>%
  mutate(FILER_ID = as.character(FILER_ID))
```

### manually process edge cases to correctly flag orgs vs individuals

Correct cases that the is_individual() function got wrong. 
Edge cases are identified by looking at which contributors flagged as individuals were missing employer info
and which contributors flagged as organizations had employer info. 

```{r}
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$standardized_name <- "LOVECCHIO, PETER"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$processed_name <- "LOVECCHIO PETER"

contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$standardized_name <- "ZUNILDA, PEREZ"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$processed_name <- "ZUNILDA, PEREZ"

contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$standardized_name <- "LIDELL, JOHN MARK" # PA-C = physician assistant (don't want to add PAC to the suffix list)
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$processed_name <- "LIDELL JOHN MARK"

contributions[which(contributions$Contributor.Name == "Last0b6df6bd, First7606aee3"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lastcbea8f37, Firstd599d312"),]$entity_type <- "individual"
```

### standardize employer + occupation for individuals

For individuals that are missing employer/occupation info, standardize employer/occupation to be `NONE` 
Note: there are edge cases where employer is N/A, but occupation indicates there are employed or self-employed. 
These individuals are classified as major donors. 

```{r}
contributions <- contributions %>%
  mutate(
    standardized_employer_name   = if_else(entity_type == "individual" & standardized_employer_name == "", "NONE", standardized_employer_name),
    processed_employer_name  = if_else(entity_type == "individual" & processed_employer_name == "", "NONE", processed_employer_name),

    standardized_occupation  = if_else(entity_type == "individual" & standardized_occupation == "", "NONE", standardized_occupation),
    processed_occupation = if_else(entity_type == "individual" & processed_occupation == "", "NONE", processed_occupation),    
  )

```




# Aggregate entities by Normalized name

Group direct contributor rows on the `Normalized`
column, with subgroups considered for individuals with the same name
who were determined to be different people through manual review. 

The $5k threshold is applied within each `recipient_key`:
- `Ballot.Measure` for prop contributions (there may be multiple recipient committees per side)
- `Recipient.Name` | `Office` for candidate races (one name spans all of a candidate's
  committees across the dataset - Ben Allen has more than one)

```{r}
# Orgs

orgs <- contributions %>%
  filter(entity_type == "organization") %>%
  mutate(Normalized = str_squish(coalesce(Normalized, ""))) %>%
  #make_name_variants("Contributor.Name") %>%
  add_recipient_key()

# orgs can be grouped directly by Normalized name
org_norm_totals <- orgs %>%
  filter(Normalized != "") %>%
  group_by(Normalized, recipient_key) %>%
  summarise(norm_total = sum(Amount, na.rm = TRUE), .groups = "drop") %>%
  filter(norm_total >= 5000)

# contributor rows for qualifying (Normalized × recipient) entities,
# with entity UUID from canonical_uuid_map
org_classification_input_normalized <- orgs %>%
  filter(Normalized != "") %>%
  inner_join(
    org_norm_totals %>% select(Normalized, recipient_key),
    by = c("Normalized", "recipient_key")
  ) %>%
  left_join(
    canonical_uuid_map %>% select(canonical_name, entity_uuid),
    by = c("Normalized" = "canonical_name")
  ) %>%
  mutate(entity_uuid = coalesce(entity_uuid, uuid::UUIDfromName(NORM_UUID_NAMESPACE, Normalized))) %>%
  rename(entity_id_normalized = entity_uuid)

cat(sprintf(
  "Normalized aggregation (orgs): %d entity x recipient pairs above $5k | %d distinct entities\n",
  nrow(org_norm_totals),
  n_distinct(org_norm_totals$Normalized)
))

# Individuals

individuals <- contributions %>%
  filter(entity_type == "individual") %>%
  mutate(Normalized = str_squish(coalesce(Normalized, ""))) %>%
  add_recipient_key()

# assign entity UUIDs before threshold so subgroup splitting can be applied first

indiv_with_entity_id <- individuals %>%
  filter(Normalized != "") %>%
  left_join(
    canonical_uuid_map %>% select(canonical_name, entity_uuid),
    by = c("Normalized" = "canonical_name")
  ) %>%
  mutate(entity_uuid = coalesce(entity_uuid, uuid::UUIDfromName(NORM_UUID_NAMESPACE, Normalized))) %>%
  assign_subgroup_entity_id(default_col = "entity_uuid") # call function to distinguish tracked individual w/ same name

# threshold is applied per entity_uuid x recipient (not per Normalized x recipient)
indiv_entity_totals <- indiv_with_entity_id %>%
  group_by(entity_uuid, recipient_key) %>%
  summarise(entity_total = sum(Amount, na.rm = TRUE), .groups = "drop") %>%
  filter(entity_total >= 5000)

indiv_classification_input_normalized <- indiv_with_entity_id %>%
  inner_join(
    indiv_entity_totals %>% select(entity_uuid, recipient_key),
    by = c("entity_uuid", "recipient_key")
  ) %>%
  rename(entity_id_normalized = entity_uuid)

cat(sprintf(
  "Normalized aggregation (individuals): %d entity x recipient pairs above $5k | %d distinct entities\n",
  nrow(indiv_entity_totals),
  n_distinct(indiv_entity_totals$entity_uuid)
))

```


### data check for subgroup - can skip to "Combine individuals and orgs" section

```{r}
# data checks
individuals_norm_only <- individuals %>%
  filter(Normalized != "") %>% 
  group_by(Normalized, recipient_key) %>%
  summarise(entity_total = sum(Amount, na.rm = TRUE), .groups = "drop") %>%
  filter(entity_total >= 5000)

setdiff(indiv_entity_totals %>% select(-entity_uuid),individuals_norm_only %>% select(-Normalized))

disambig_names <- subgroup_profiles %>% distinct(canonical_name) %>% pull(canonical_name)

indiv_with_entity_id %>%
  filter(Normalized %in% disambig_names) %>%
  group_by(Normalized, entity_uuid, recipient_key) %>%
  summarise(total = sum(Amount, na.rm = TRUE), n_rows = n(), .groups = "drop") %>%
  group_by(Normalized, recipient_key) %>%
  filter(n_distinct(entity_uuid) > 1) %>%
  ungroup() %>%
  arrange(Normalized, recipient_key, entity_uuid)

```


### Combine individuals and orgs

```{r}
classification_input <- bind_rows(
  org_classification_input_normalized,
  indiv_classification_input_normalized
) %>%
  rename(entity_id = entity_id_normalized)
cat(sprintf("Classification input (normalized method): %d rows\n", nrow(classification_input)))
```


### match back uuid from previous classifications

```{r}
# define normalization functions needed to map 
pad_zip <- function(z) {
  z <- trimws(as.character(z))
  needs_pad <- !is.na(z) & z != "" & str_detect(z, "^[0-9]+$") & nchar(z) < 5
  z[needs_pad] <- paste0(strrep("0", 5 - nchar(z[needs_pad])), z[needs_pad])
  substr(z, 1, 5)
}

normalize_missing <- function(x) {
  x <- trimws(as.character(x))
  if_else(x %in% c("", "N/A", "NA", "None","n/a") | is.na(x), "N/A", x)
}
```

**Note:** the `legacy_match_id` is NOT distinct per contribution record - some
records contain the exact same info across name, id, amount, date, etc. 
`uuid` IS distinct per row, but is assigned back based on `legacy_match_id`
(since dup rows contain the exact same info it shouldn't cause an issue - these
rows are classified as one unit and won't have received different classifications).

```{r}
categorized <- categorized %>%
  mutate(
    legacy_match_id = make_row_hash(
      Contributor.Name, Contributor.ID, as.numeric(Amount),
      as.character(Recipient.Committee.ID),
      Contributor.City, normalize_missing(Contributor.Employer),
      normalize_missing(Contributor.Occupation), pad_zip(Contributor.Zip.Code),
      Start.Date
    )
  ) %>%
  group_by(legacy_match_id) %>%
  mutate(dup_rank = row_number()) %>%
  ungroup()

classification_input <- classification_input %>%
  mutate(
    legacy_match_id = make_row_hash(
      Contributor.Name, Contributor.ID, as.numeric(Amount),
      as.character(Recipient.Committee.ID),
      Contributor.City, normalize_missing(Contributor.Employer),
      normalize_missing(Contributor.Occupation), pad_zip(Contributor.Zip.Code),
      Start.Date
    )
  ) %>%
  group_by(legacy_match_id) %>%
  mutate(dup_rank = row_number()) %>%
  ungroup() %>%
  left_join(
    categorized %>% select(legacy_match_id, dup_rank, uuid),
    by = c("legacy_match_id", "dup_rank")
  ) %>%
  select(-dup_rank)

```


### data checks (can skip to next section)

Make sure everything matched back that should have.
Note that `categorized` includes MOSTLY PAC contributors, so we 
expect most of the uuids to not have matched back to the direct contributors
data. 

The thresholds were applied differently in the past, so this also cause some 
contributors to be classified in earlier iterations that now no longer meet 
the threshold for singular recipient. 

```{r}
# data checks
classification_input %>% count(uuid) %>% filter(n > 1) # should only be NAs (uuids haven't been assigned to new contributors yet)

# make sure match id is unique - should be 0 rows
categorized %>% count(legacy_match_id, dup_rank) %>% filter(n > 1)

length(unique(categorized$uuid))
length(which(!is.na(classification_input$uuid)))
```

```{r}
#direct_recipient_ids <- unique(as.character(contributions_full$Recipient.Committee.ID))
direct_recipient_ids <- contributions %>% select(Recipient.Committee.ID) %>% distinct() %>% pull(Recipient.Committee.ID)
direct_recipient_ids <- as.character(direct_recipient_ids)

categorized_direct <- categorized %>%
  filter(coalesce(as.character(Recipient.Committee.ID), "") %in% direct_recipient_ids)
categorized_pac <- categorized %>%
  filter(!coalesce(as.character(Recipient.Committee.ID), "") %in% direct_recipient_ids)

cat(sprintf(
  "categorized: %d total rows; %d direct contributions (to tracked races/props); %d PAC contributions\n",
  nrow(categorized), nrow(categorized_direct), nrow(categorized_pac)
))

# Direct-contributor mismatches only
categorized_direct %>%
  filter(!(uuid %in% classification_input$uuid) & Contributor.Name != "Unitemized Contributions") %>%
  select(uuid, Contributor.Name, Contributor.Employer, Contributor.Occupation, Contributor.City, Contributor.State, Contributor.Zip.Code,
         Recipient.Committee, Amount, Start.Date)
```

```{r}
unmatched_direct <- categorized_direct %>%
  filter(!(uuid %in% classification_input$uuid))

# current per-contributor/per-committee totals in the new data
current_totals <- contributions_full %>%
  mutate(Recipient.Committee.ID = as.character(Recipient.Committee.ID)) %>%
  group_by(Contributor.Name, Recipient.Committee.ID) %>%
  summarise(current_total = sum(Amount, na.rm = TRUE), .groups = "drop")

threshold_check <- unmatched_direct %>%
  mutate(Recipient.Committee.ID = as.character(Recipient.Committee.ID)) %>%
  left_join(current_totals, by = c("Contributor.Name", "Recipient.Committee.ID")) %>%
  mutate(
    meets_threshold  = coalesce(current_total, 0) >= 5000,
    in_current_data  = !is.na(current_total)
  )

# summary
threshold_check %>%
  count(in_current_data, meets_threshold) %>%
  arrange(in_current_data, meets_threshold)

# rows that don't meet threshold — 
threshold_check %>%
  filter(!meets_threshold) %>%
  select(Contributor.Name, Recipient.Committee.ID, current_total, Amount, Start.Date) %>%
  arrange(desc(coalesce(current_total, 0))) # aggregated total 


# rows that DO meet threshold -> should all be unitemized
threshold_check %>%
  filter(meets_threshold) %>%
  select(Contributor.Name, Recipient.Committee.ID, current_total, Amount, Start.Date) %>%
  arrange(desc(coalesce(current_total, 0)))
```

```{r}
unmatched_direct <- categorized_direct %>%
  filter(!(uuid %in% classification_input$uuid))

# Case 1: legacy_match_id exists in classification_input — dup_rank collision
in_ci_by_hash <- unmatched_direct %>%
  filter(legacy_match_id %in% classification_input$legacy_match_id)

# Case 2: name is in classification_input but hash doesn't match — a field changed
in_ci_by_name <- unmatched_direct %>%
  semi_join(classification_input, by = "Contributor.Name") %>%
  anti_join(classification_input, by = "legacy_match_id")

# Case 3: name not in classification_input at all — row was dropped from the data
absent_from_data <- unmatched_direct %>%
  anti_join(classification_input, by = "Contributor.Name")

cat(sprintf("Case 1 (hash match, dup_rank mismatch): %d\n", nrow(in_ci_by_hash)))
cat(sprintf("Case 2 (name in data, field changed): %d\n", nrow(in_ci_by_name)))
cat(sprintf("Case 3 (completely absent from new data): %d\n", nrow(absent_from_data)))

```

```{r}
categorized_direct %>% filter(Contributor.Name == "Google, LLC and Affiliated Entities" & !(uuid %in% classification_input$uuid)) %>% select(Contributor.Name, Amount, Start.Date, Contributor.City,
         Contributor.Employer, Contributor.Occupation, Contributor.Zip.Code, uuid, Recipient.Committee, Office)
```


```{r}
dropped_name <- "Google, LLC and Affiliated Entities"

contributions_full %>%
  filter(Contributor.Name == dropped_name,
         Recipient.Committee.ID %in% direct_recipient_ids) %>%
  group_by(Recipient.Committee.ID, Recipient.Committee) %>%
  summarise(total = sum(Amount), n = n(), .groups = "drop") %>%
  arrange(desc(total))

```

```{r}
classification_input %>% filter(Contributor.Name == "Blue Diamond Growers") %>% select(Contributor.Name, Amount, Start.Date, Contributor.City,
         Contributor.Employer, Contributor.Occupation, Contributor.Zip.Code, uuid, Recipient.Committee)
contributions_full %>% filter(Contributor.Name == "Blue Diamond Growers", Recipient.Committee == "Ben Allen Senate 2022 Officeholder Account") %>% select(Contributor.Name, Amount, Start.Date, Contributor.City,
         Contributor.Employer, Contributor.Occupation, Contributor.Zip.Code)
contributions_full %>% filter(Contributor.Name == "Blue Diamond Growers")
```

### Add uuids to new contribution records

```{r}
# add remaining uuids
classification_input[which(is.na(classification_input$uuid)),]$uuid <-
  uuid::UUIDgenerate(n = nrow(classification_input[which(is.na(classification_input$uuid)),]))
```


# Look for contributors missing a Contributor.ID 

Not all contributors have a Contributor.ID when they should. 
Look at reference data to flag contributor names without an ID
that match a known name of a recipient committee. 

  
### look for any names that match a recipient committee

```{r}
# data to map contributor ids
id_map_full <- read_tsv("01_inputs/FILERNAME_CD.tsv")
id_map <- unique(id_map_full[,c("XREF_FILER_ID","FILER_ID","FILER_TYPE")]) # get unique xref id / filer ids combos

# convert all ids to character strings
id_map$FILER_ID  <- as.character(id_map$FILER_ID)
id_map$XREF_FILER_ID <- as.character(id_map$XREF_FILER_ID)
classification_input$Contributor.ID  <- as.character(classification_input$Contributor.ID)
classification_input$effective_Contributor.ID <- as.character(classification_input$effective_Contributor.ID)

# map filer id to contributor id using effective_Contributor.ID (includes sheet corrections)
classification_input <- left_join(classification_input, id_map, by = c("effective_Contributor.ID" = "XREF_FILER_ID"))

# check that each has a corresponding filer id
classification_input %>% filter(is.na(FILER_ID) & !is.na(effective_Contributor.ID)) %>% select(Contributor.Name,Contributor.ID,FILER_ID)
```


```{r}
# build name/ID lookup from full dataset
full_id_lookup <- id_map_full %>%
  filter(!is.na(FILER_ID), FILER_ID != "",FILER_TYPE == "RECIPIENT COMMITTEE") %>% # , (FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" | FILER_TYPE == "RECIPIENT COMMITTEE")
  mutate(name_v5 = standardize_names(NAML) %>% 
           str_remove_all("\\s*\\(.*?\\)") %>%
          str_squish()) %>%
  distinct(name_v5, FILER_ID) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(FILER_ID)) %>%
  ungroup()

# propagate IDs to orgs still missing an effective ID after sheet corrections
full_id_propagation <- classification_input %>%
  filter(is.na(effective_Contributor.ID) | effective_Contributor.ID == "") %>%
  mutate(name_v5 = standardize_names(Contributor.Name) %>% 
           str_remove_all("\\s*\\(.*?\\)") %>%
          str_squish()) %>%
  select(row_id, Contributor.Name, name_v5) %>%
  inner_join(full_id_lookup, by = "name_v5") %>%
  mutate(needs_review = n_ids > 1,
         FILER_ID = as.character(FILER_ID)) %>%
  select(row_id, Contributor.Name, propagated_id = FILER_ID, n_ids, needs_review)
full_id_propagation
```

If any contributor doesn't have a Contributor.ID that should, add them to the ID corrections Google Sheet.
Propagated IDs used to be applied automatically (with review) - but now require review of the specific
transaction to confirm that's the correct id. 

```{r}
classification_input$Contributor.ID <- as.character(classification_input$Contributor.ID)
full_id_propagation$propagated_id <- as.character(full_id_propagation$propagated_id)
```


# WRITE OUTPUTS

How entities are classified depends on whether they are a PAC or an org/individual. 
Orgs/individuals are classified directly, while PACs are classified according to their 
contributors. Separate PACs from orgs/individuals in the output. 

PAC input includes:
  - all recipient PACs
  - anything with PAC or federal PAC language in the name (even in cases where there is no contributor ID)
  - major donors only if they have PAC/fed PAC language (very rare); 
    major donors without PAC language go to direct classification even if they have an ID.
    note that FILER_TYPE = `MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE` is misleading. All IE committees
    that receive contributions have FILER_TYPE = `RECIPIENT COMMITTEE`

```{r}
pacs <- classification_input %>%
  filter(
    coalesce(has_pac_language, FALSE) | # anything with pac language processed as a pac
    coalesce(is_fed_pac, FALSE) |
    #  rows that didn't match a FILER_Id won't have a FILER_TYPE -> default to treating as a PAC
    (!is.na(effective_Contributor.ID) & (is.na(FILER_TYPE) | FILER_TYPE != "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE")) 
  ) %>%
  mutate(is_candidate_pac = is_candidate_pac(Contributor.Name)) # flag for candidate committees

direct <- classification_input %>%
  filter(
    !coalesce(has_pac_language, FALSE) &
    !coalesce(is_fed_pac, FALSE) &
    (is.na(effective_Contributor.ID) | FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE")
  )

nrow(classification_input) == sum(nrow(pacs) + nrow(direct))
```


```{r}
today <- format(Sys.Date(), "%m-%d-%y")
write.csv(direct, paste0("01_outputs/classification_input_", today, ".csv"), row.names = FALSE)
write.csv(pacs,   paste0("01_outputs/pac_input_", today, ".csv"),            row.names = FALSE)
```


### data check: entity groups with mixed federal PAC status

Entity groups where at least one name variant is a confirmed federal PAC and at least one is not.
Review whether the non-federal PAC variants should also be flagged in the Google Sheet.

```{r}
name_variant_fed_pac_review <- classification_input %>%
  group_by(entity_id) %>%
  filter(any(coalesce(is_fed_pac, FALSE)) & !all(coalesce(is_fed_pac, FALSE))) %>%
  ungroup() %>%
  select(entity_id, Normalized, Contributor.Name, is_fed_pac, has_pac_language, Amount, recipient_key) %>%
  arrange(entity_id, desc(coalesce(is_fed_pac, FALSE)), Contributor.Name)

name_variant_fed_pac_review
```



