---
title: "0102b - PAC Graph Walk"
output: html_notebook
---

Produces `pac_universe.csv`: the set of PAC FILER_IDs reachable from the seed
PACs in `pac_input.csv` via recursive Contributor.ID -> Recipient.Committee.ID
traversal. No $5k threshold is applied — the universe is used by 0802 to scope
PAC contributor entity resolution before 0803 runs.

PACs classified as Candidates/Political Parties or federal in `pac_manual_review`
are included in the universe but not traversed (their contributors don't require
individual classification).

Run order: after 0802 produces `pac_input.csv`, before 0803.

# Setup

```{r}
library(dplyr)
library(readr)
library(stringr)
library(googlesheets4)
gs4_deauth()

source("standardization_helpers.R")
```

# Load data

```{r}
contributions_full <- read.csv("01_inputs/power_search_contributions_normalized.csv") %>%
  rename(Ballot.Measure = Ballot.Measure.s.) %>%
  mutate(
    Contributor.ID         = as.character(Contributor.ID),
    Recipient.Committee.ID = as.character(Recipient.Committee.ID)
  )
```

```{r}
# FILERNAME_CD maps XREF_FILER_ID → FILER_ID; used to standardize contributor IDs.
id_map <- read_tsv("01_inputs/FILERNAME_CD.tsv", col_types = cols(.default = "c")) %>%
  mutate(FILER_ID = as.character(FILER_ID), XREF_FILER_ID = as.character(XREF_FILER_ID)) %>%
  filter(!is.na(FILER_ID), !is.na(XREF_FILER_ID)) %>%
  distinct(XREF_FILER_ID, FILER_ID, FILER_TYPE)
```


```{r}
id_corrections <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1UnO3wldvEo84Md0ac5jbpg4SqWzvPle3yh15e_DGAVk",
  sheet = "corrected_pac_ids",
  col_types = "c"
) %>%
  mutate(
    Contributor.ID = coalesce(Contributor.ID, ""),
    resolved       = !is.na(resolved) & tolower(str_trim(resolved)) != "false",
    is_fed_pac     = tolower(str_trim(coalesce(is_fed_pac, ""))) %in% c("true", "1", "yes")
  )
```


**update pac_input file**

```{r}
#pac_input <- read.csv("01_outputs/pac_input_09-07-26.csv", colClasses = "character")
pac_input <- read.csv("01_outputs/new_prop_pacs.csv", colClasses = "character")
```

```{r}
# re-apply corrections in case the sheet was updated after pac_input was written
pac_input <- pac_input %>%
  mutate(is_fed_pac = as.logical(coalesce(is_fed_pac, "FALSE")))

pac_input <- apply_contributor_corrections(pac_input, id_corrections)

# re-apply flag
pac_input <- pac_input %>%
  mutate(is_fed_pac = is_fed_pac | has_fed_pac_language(Contributor.Name))

# re-derive FILER_ID and FILER_TYPE for any seeds still missing them after corrections
pac_input <- pac_input %>%
  left_join(
    id_map %>%
      distinct(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>%
      group_by(XREF_FILER_ID) %>% slice(1) %>% ungroup() %>%
      rename(derived_filer_id = FILER_ID, derived_filer_type = FILER_TYPE),
    by = c("effective_Contributor.ID" = "XREF_FILER_ID")
  ) %>%
  mutate(
    FILER_ID   = coalesce(na_if(FILER_ID, ""), derived_filer_id),
    FILER_TYPE = coalesce(na_if(FILER_TYPE, ""), derived_filer_type, "RECIPIENT COMMITTEE")
  ) %>%
  select(-derived_filer_id, -derived_filer_type)

cat(sprintf(
  "pac_input after corrections: %d rows; %d with FILER_ID; %d flagged is_fed_pac\n",
  nrow(pac_input),
  sum(!is.na(pac_input$FILER_ID) & pac_input$FILER_ID != ""),
  sum(coalesce(pac_input$is_fed_pac, FALSE))
))
```


```{r, eval = FALSE}
# data checks
pac_input %>% filter(is.na(FILER_ID) | is.na(FILER_TYPE))
pac_input %>% filter(Contributor.Name == "DRIVE Committee" | Contributor.Name == "Elect Dems Now PAC" |Contributor.Name == "International Association of Sheet Metal, Air, Rail and Transportation Workers Political Action League" ) %>% select(is_fed_pac)
```


```{r}
# real in feedback from manual review of pacs
pac_manual_review <- read_csv(
  "../04_assign_classifications/04_inputs/pac_manual_review.csv",
  col_types = cols(.default = "c")
) %>%
  mutate(
    pac_id           = na_if(trimws(coalesce(pac_id, "")), ""),
    fec_id           = na_if(trimws(coalesce(fec_id, "")), ""),
    is_resolved      = !(
      coalesce(unresolvable_reason, "") == "additional_review_req" |
      is.na(industry_override_under_5k) |
      trimws(coalesce(industry_override_under_5k, "")) %in% c("", "PENDING")
    ),
    is_candidates_pp = trimws(coalesce(industry_override_under_5k, "")) == "Candidates/Political Parties",
    is_fec_federal   = !is.na(fec_id)
  )
```


# Prepare contributions

Apply contributor ID corrections and add PAC flags.

```{r}
contributions <- apply_contributor_corrections(contributions_full, id_corrections)

# Contributor.ID (XREF_FILER_ID) → std_contributor_id (FILER_ID)
contributions <- contributions %>%
  left_join(
    id_map %>% select(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>%
      rename(std_contributor_id = FILER_ID),
    by = c("effective_Contributor.ID" = "XREF_FILER_ID")
  ) %>%
  mutate(std_contributor_id = as.character(std_contributor_id))
```

```{r}
contributions <- contributions %>%
  mutate(
    has_id             = !is.na(effective_Contributor.ID) & effective_Contributor.ID != "",
    is_fed_pac         = has_fed_pac_language(Contributor.Name) | coalesce(is_fed_pac, FALSE),
    is_self            = !is.na(std_contributor_id) & std_contributor_id == Recipient.Committee.ID,
    std_contributor_id = if_else(is_self, NA_character_, std_contributor_id),
    is_pac             = !is_self & has_id & !is_fed_pac,
    is_pac_no_id       = !has_id & !is_fed_pac & (
                           coalesce(has_pac_language, FALSE) |
                           grepl("\\bPAC\\b|COMMITTEE|POLITICAL ACTION|POLITICAL FUND|ACTION FUND",
                                 toupper(Contributor.Name))
                         )
  )
```


# Define no-traverse PAC sets

PACs classified as Candidates/Political Parties or federal are recorded in the
universe but not traversed — we do not classify these PACs by their $5k contributors.

```{r}
no_traverse_from_review <- pac_manual_review %>%
  filter(!is.na(pac_id), is_resolved, is_candidates_pp | is_fec_federal) %>%
  pull(pac_id) %>%
  as.character() %>%
  unique()

# exclude PACs flagged is_fed_pac in pac_input (via corrections or name
# pattern) that may not have been reviewed in pac_manual_review yet.
fed_pac_from_input <- pac_input %>%
  filter(is_fed_pac == TRUE, !is.na(FILER_ID), FILER_ID != "") %>%
  pull(FILER_ID) %>%
  unique()

# exclude candidate PACs identified by recipient name pattern that haven't
# been reviewed in pac_manual_review yet.
candidate_pac_ids_by_name <- contributions %>%
  distinct(Recipient.Committee.ID, Recipient.Committee) %>%
  filter(is_candidate_pac(Recipient.Committee)) %>%
  pull(Recipient.Committee.ID) %>%
  unique()

no_traverse_pac_ids <- unique(c(no_traverse_from_review, fed_pac_from_input, candidate_pac_ids_by_name))

cat(sprintf(
  "No-traverse PACs: %d total (%d from pac_manual_review; %d is_fed_pac; %d candidate by name)\n",
  length(no_traverse_pac_ids),
  length(no_traverse_from_review),
  length(setdiff(fed_pac_from_input, no_traverse_from_review)),
  length(setdiff(candidate_pac_ids_by_name, c(no_traverse_from_review, fed_pac_from_input)))
))
```



# Define PAC graph

Starting from the seed PAC FILER_IDs in `pac_input.csv`, follow non-federal
PAC contributor edges recursively. No threshold applied - need all contributors
to do entity resolution. 

```{r}
walk_pac_graph <- function(seed_filer_ids, prepped, no_traverse_ids,
                           max_depth = 5) {
  seeds   <- seed_filer_ids[!is.na(seed_filer_ids) & seed_filer_ids != ""]
  queue   <- lapply(seeds, function(id) list(id = id, depth = 0L, is_seed = TRUE))
  visited <- tibble(pac_id = character(), depth = integer(), is_seed = logical(),
                    traversed = logical(), has_pac_no_id_contributors = logical())
  visited_ids <- character()

  while (length(queue) > 0) {
    node   <- queue[[1]]; queue <- queue[-1]
    pac_id <- node$id;    depth <- as.integer(node$depth)

    if (pac_id %in% visited_ids || depth > max_depth) next
    visited_ids <- c(visited_ids, pac_id)

    contribs <- prepped %>%
      filter(Recipient.Committee.ID == pac_id)

    # PACs classified as candidates/PP or federal: record but don't traverse
    pac_name <- contribs$Recipient.Committee[1]
    is_federal_by_name   <- !is.na(pac_name) && has_fed_pac_language(pac_name)
    is_candidate_by_name <- !is.na(pac_name) && is_candidate_pac(pac_name)

    if (pac_id %in% no_traverse_ids || is_federal_by_name || is_candidate_by_name) {
      visited <- bind_rows(visited,
                           tibble(pac_id = pac_id, depth = depth,
                                  is_seed = isTRUE(node$is_seed),
                                  traversed = FALSE,
                                  has_pac_no_id_contributors = FALSE))
      next
    }

    # non-federal, non-candidate PAC contributors with a resolved ID are queued
    to_traverse <- contribs %>%
      filter(!is_fed_pac, !is_candidate_pac(Contributor.Name), !is.na(std_contributor_id)) %>%
      pull(std_contributor_id) %>%
      unique()

    for (cid in to_traverse) {
      if (!cid %in% visited_ids)
        queue <- append(queue, list(list(id = cid, depth = depth + 1L, is_seed = FALSE)))
    }

    has_no_id <- any(contribs$is_pac_no_id, na.rm = TRUE)

    visited <- bind_rows(visited,
                         tibble(pac_id = pac_id, depth = depth,
                                is_seed = isTRUE(node$is_seed),
                                traversed = TRUE,
                                has_pac_no_id_contributors = has_no_id))
  }

  visited
}
```

```{r}
# Patch known FILER_ID gaps
pac_input <- pac_input %>%
  mutate(FILER_ID = case_when(
    effective_Contributor.ID == "1479934" ~ "1479934", 
    effective_Contributor.ID == "1049814" ~ "1049814",
    TRUE ~ FILER_ID
  ))

seed_ids <- pac_input %>%
  filter(!is.na(FILER_ID), FILER_ID != "") %>%
  pull(FILER_ID) %>%
  unique()

cat(sprintf("Seeds: %d PAC FILER_IDs from pac_input.csv\n", length(seed_ids)))
```


### Walk PAC graph

```{r}
pac_universe <- walk_pac_graph(seed_ids, contributions, no_traverse_pac_ids,
                               max_depth = 5)

cat(sprintf(
  "PAC universe: %d PACs (%d seeds, %d discovered; %d traversed, %d not traversed)\n",
  nrow(pac_universe),
  sum(pac_universe$is_seed),
  sum(!pac_universe$is_seed),
  sum(pac_universe$traversed),
  sum(!pac_universe$traversed)
))

pac_universe %>%
  count(depth) %>%
  rename(n_pacs = n) %>%
  print()

no_id_pacs <- pac_universe %>% filter(traversed, has_pac_no_id_contributors)
if (nrow(no_id_pacs) > 0) {
  cat(sprintf(
    "Note: %d traversed PAC(s) have contributors with PAC language but no Contributor.ID —",
    nrow(no_id_pacs)
  ))
  cat(" indirect contributors may be missed.\n")
}
```

```{r}
# data check: review candidate PACs
candidate_pac_in_universe <- pac_universe %>%
  filter(!traversed) %>%
  left_join(
    contributions %>% distinct(Recipient.Committee.ID, Recipient.Committee),
    by = c("pac_id" = "Recipient.Committee.ID")
  ) %>%
  filter(is_candidate_pac(coalesce(Recipient.Committee, pac_id))) %>%
  select(pac_id, Recipient.Committee, depth, is_seed)
cat(sprintf("%d PAC(s) in universe flagged as candidate committees\n",
            nrow(candidate_pac_in_universe)))
print(candidate_pac_in_universe)
```


# Look at missing ids

```{r}
no_id_pacs <- pac_universe %>% filter(traversed, has_pac_no_id_contributors)

contributions %>%
  filter(is_pac_no_id, Recipient.Committee.ID %in% no_id_pacs$pac_id) %>%
  distinct(Contributor.Name, Recipient.Committee.ID, Recipient.Committee,Amount) %>%
  arrange(Contributor.Name) %>% distinct(Contributor.Name)

```

```{r}
traversed_ids <- pac_universe %>% filter(traversed) %>% pull(pac_id)

# Candidate/ballot measure committee IDs : what seed PACs contributed to
tracked_committee_ids <- contributions %>%
  filter(std_contributor_id %in% seed_ids) %>%
  pull(Recipient.Committee.ID) %>%
  unique()

# Traversed PACs that gave $5k+ to a relevant downstream recipient
qualifying_recipient_pacs <- contributions %>%
  filter(
    !is.na(std_contributor_id),
    std_contributor_id %in% traversed_ids,
    Recipient.Committee.ID %in% c(traversed_ids, tracked_committee_ids)
  ) %>%
  group_by(std_contributor_id, Recipient.Committee.ID) %>%
  summarise(total_given = sum(as.numeric(Amount), na.rm = TRUE), .groups = "drop") %>%
  filter(total_given >= 5000) %>%
  pull(std_contributor_id) %>%
  unique()

no_id_for_review <- contributions %>%
  filter(
    is_pac_no_id,
    Recipient.Committee.ID %in% qualifying_recipient_pacs
  ) %>%
  group_by(Contributor.Name, Recipient.Committee.ID, Recipient.Committee) %>%
  summarise(total_amount = sum(as.numeric(Amount), na.rm = TRUE), .groups = "drop") %>%
  filter(total_amount >= 5000) %>%
  arrange(desc(total_amount))

```


# Write output

```{r}
# flag candidate and federal PACs so 0802 can exclude them from
# relevant_recipient_ids 
all_candidate_pac_ids <- unique(c(
  pac_manual_review %>% filter(!is.na(pac_id), is_candidates_pp) %>% pull(pac_id),
  candidate_pac_ids_by_name
))

fed_pac_ids_by_name <- contributions %>%
  distinct(Recipient.Committee.ID, Recipient.Committee) %>%
  filter(has_fed_pac_language(Recipient.Committee)) %>%
  pull(Recipient.Committee.ID) %>%
  unique()

all_federal_pac_ids <- unique(c(
  fed_pac_from_input,
  pac_manual_review %>% filter(!is.na(pac_id), is_fec_federal) %>% pull(pac_id),
  fed_pac_ids_by_name
))

pac_universe <- pac_universe %>%
  mutate(
    is_candidate_pac = pac_id %in% all_candidate_pac_ids,
    is_federal_pac   = pac_id %in% all_federal_pac_ids
  )

cat(sprintf(
  "%d PAC(s) flagged is_candidate_pac; %d flagged is_federal_pac\n",
  sum(pac_universe$is_candidate_pac),
  sum(pac_universe$is_federal_pac)
))
```

```{r}
#write.csv(pac_universe, "01_outputs/pac_universe.csv", row.names = FALSE)
write.csv(pac_universe, "01_outputs/pac_universe_new_props.csv", row.names = FALSE)
cat(sprintf("pac_universe.csv written: %d rows\n", nrow(pac_universe)))
```
