---
title: "0802b - PAC Graph Walk"
output: html_notebook
---

Produces `pac_universe.csv`: the set of PAC FILER_IDs reachable from the seed
PACs in `pac_input.csv` via recursive Contributor.ID → Recipient.Committee.ID
traversal. No $5k threshold is applied — the universe is used by 0802 to scope
PAC contributor entity resolution before 0803 runs.

PACs classified as Candidates/Political Parties or federal in `pac_manual_review`
are included in the universe but not traversed (their contributors don't require
individual classification).

Run order: after 0802 produces `pac_input.csv`, before 0803.

# Setup

```{r}
library(dplyr)
library(readr)
library(stringr)
library(googlesheets4)
gs4_deauth()

source("standardization_helpers.R")
```

# Load data

```{r}
contributions_full <- read.csv("08_inputs/power_search_contributions_normalized.csv") %>%
  rename(Ballot.Measure = Ballot.Measure.s.) %>%
  mutate(
    Contributor.ID         = as.character(Contributor.ID),
    Recipient.Committee.ID = as.character(Recipient.Committee.ID)
  )
```

```{r}
# Keep full table for name-based ID propagation (NAML = committee/org name).
# id_map is derived from this so we only read the file once.
id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv", col_types = cols(.default = "c")) %>%
  mutate(FILER_ID = as.character(FILER_ID), XREF_FILER_ID = as.character(XREF_FILER_ID))

id_map <- id_map_full %>%
  filter(!is.na(FILER_ID), !is.na(XREF_FILER_ID)) %>%
  distinct(XREF_FILER_ID, FILER_ID, FILER_TYPE)
```


```{r}
id_corrections <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1UnO3wldvEo84Md0ac5jbpg4SqWzvPle3yh15e_DGAVk",
  sheet = "corrected_pac_ids",
  col_types = "c"
) %>%
  mutate(
    Contributor.ID = coalesce(Contributor.ID, ""),
    resolved       = !is.na(resolved) & tolower(str_trim(resolved)) != "false",
    is_fed_pac     = tolower(str_trim(coalesce(is_fed_pac, ""))) %in% c("true", "1", "yes")
  )
```


```{r}
#pac_input <- read.csv("08_outputs/pac_input_09-07-26.csv", colClasses = "character")
pac_input <- read.csv("08_outputs/new_prop_pacs.csv", colClasses = "character")
```

```{r}
pac_input
```



```{r}
pac_input %>%
  mutate(is_fed_pac_alt = has_fed_pac_language(Contributor.Name)) %>% filter(is_fed_pac_alt != is_fed_pac) %>% select(Contributor.Name,is_fed_pac,is_fed_pac_alt)
```

```{r}
# Re-apply corrections in case the sheet was updated after pac_input was written.
# Coerce is_fed_pac to logical first since it was read as character from CSV.
pac_input <- pac_input %>%
  mutate(is_fed_pac = as.logical(coalesce(is_fed_pac, "FALSE")))

pac_input <- apply_contributor_corrections(pac_input, id_corrections)

# Apply has_fed_pac_language so federal PACs are caught even if 0802 wasn't re-run
pac_input <- pac_input %>%
  mutate(is_fed_pac = is_fed_pac | has_fed_pac_language(Contributor.Name))

# Re-derive FILER_ID and FILER_TYPE for any seeds still missing them after corrections
pac_input <- pac_input %>%
  left_join(
    id_map %>%
      distinct(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>%
      group_by(XREF_FILER_ID) %>% slice(1) %>% ungroup() %>%
      rename(derived_filer_id = FILER_ID, derived_filer_type = FILER_TYPE),
    by = c("effective_Contributor.ID" = "XREF_FILER_ID")
  ) %>%
  mutate(
    FILER_ID   = coalesce(na_if(FILER_ID, ""), derived_filer_id),
    FILER_TYPE = coalesce(na_if(FILER_TYPE, ""), derived_filer_type, "RECIPIENT COMMITTEE")
  ) %>%
  select(-derived_filer_id, -derived_filer_type)

cat(sprintf(
  "pac_input after corrections: %d rows; %d with FILER_ID; %d flagged is_fed_pac\n",
  nrow(pac_input),
  sum(!is.na(pac_input$FILER_ID) & pac_input$FILER_ID != ""),
  sum(coalesce(pac_input$is_fed_pac, FALSE))
))
```

```{r}
pac_input %>% filter(is_fed_pac == TRUE)
```


```{r}
# data checks
pac_input %>% filter(is.na(FILER_ID) | is.na(FILER_TYPE))
pac_input %>% filter(Contributor.Name == "DRIVE Committee" | Contributor.Name == "Elect Dems Now PAC" |Contributor.Name == "International Association of Sheet Metal, Air, Rail and Transportation Workers Political Action League" ) %>% select(is_fed_pac)
```


```{r}
pac_manual_review <- read_csv(
  "../10_assign_classifications/10_inputs/pac_manual_review.csv",
  col_types = cols(.default = "c")
) %>%
  mutate(
    pac_id           = na_if(trimws(coalesce(pac_id, "")), ""),
    fec_id           = na_if(trimws(coalesce(fec_id, "")), ""),
    is_resolved      = !(
      coalesce(unresolvable_reason, "") == "additional_review_req" |
      is.na(industry_override_under_5k) |
      trimws(coalesce(industry_override_under_5k, "")) %in% c("", "PENDING")
    ),
    is_candidates_pp = trimws(coalesce(industry_override_under_5k, "")) == "Candidates/Political Parties",
    is_fec_federal   = !is.na(fec_id)
  )
```


# Prepare contributions

Apply contributor ID corrections and compute PAC flags — same logic as 0803.

```{r}
contributions <- apply_contributor_corrections(contributions_full, id_corrections)

# Contributor.ID (XREF_FILER_ID) → std_contributor_id (FILER_ID)
contributions <- contributions %>%
  left_join(
    id_map %>% select(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>%
      rename(std_contributor_id = FILER_ID),
    by = c("effective_Contributor.ID" = "XREF_FILER_ID")
  ) %>%
  mutate(std_contributor_id = as.character(std_contributor_id))
```

```{r}
contributions <- contributions %>%
  mutate(
    has_id             = !is.na(effective_Contributor.ID) & effective_Contributor.ID != "",
    is_fed_pac         = has_fed_pac_language(Contributor.Name) | coalesce(is_fed_pac, FALSE),
    is_self            = !is.na(std_contributor_id) & std_contributor_id == Recipient.Committee.ID,
    std_contributor_id = if_else(is_self, NA_character_, std_contributor_id),
    is_pac             = !is_self & has_id & !is_fed_pac,
    is_pac_no_id       = !has_id & !is_fed_pac & (
                           coalesce(has_pac_language, FALSE) |
                           grepl("\\bPAC\\b|COMMITTEE|POLITICAL ACTION|POLITICAL FUND|ACTION FUND",
                                 toupper(Contributor.Name))
                         )
  )
```


# Define no-traverse PAC sets

PACs classified as Candidates/Political Parties or federal are recorded in the
universe but not traversed — we do not classify these PACs by their $5k contributors.

```{r}
no_traverse_from_review <- pac_manual_review %>%
  filter(!is.na(pac_id), is_resolved, is_candidates_pp | is_fec_federal) %>%
  pull(pac_id) %>%
  as.character() %>%
  unique()

# Also exclude PACs flagged is_fed_pac in pac_input (via corrections or name
# pattern) that may not have been reviewed in pac_manual_review yet.
fed_pac_from_input <- pac_input %>%
  filter(is_fed_pac == TRUE, !is.na(FILER_ID), FILER_ID != "") %>%
  pull(FILER_ID) %>%
  unique()

# Also exclude candidate PACs identified by recipient name pattern that haven't
# been reviewed in pac_manual_review yet.
candidate_pac_ids_by_name <- contributions %>%
  distinct(Recipient.Committee.ID, Recipient.Committee) %>%
  filter(is_candidate_pac(Recipient.Committee)) %>%
  pull(Recipient.Committee.ID) %>%
  unique()

no_traverse_pac_ids <- unique(c(no_traverse_from_review, fed_pac_from_input, candidate_pac_ids_by_name))

cat(sprintf(
  "No-traverse PACs: %d total (%d from pac_manual_review; %d is_fed_pac; %d candidate by name)\n",
  length(no_traverse_pac_ids),
  length(no_traverse_from_review),
  length(setdiff(fed_pac_from_input, no_traverse_from_review)),
  length(setdiff(candidate_pac_ids_by_name, c(no_traverse_from_review, fed_pac_from_input)))
))
```

- some of the additional include cases that weren't labeled as a federal pac in the manual review data, but had been reviewed.

```{r}
contributions %>% filter(grepl("WIRELESS INFRA", Contributor.Name)) %>% select(is_fed_pac, std_contributor_id,Contributor.Name)
```


```{r}
# data check
setdiff(fed_pac_from_input, no_traverse_from_review)
pac_input %>% filter(FILER_ID == "1048353")
pac_input %>%
  filter(is_fed_pac == TRUE & !(FILER_ID %in% pac_manual_review$pac_id))
```


Duane Morris LLP Government Committee PA ID#7900289 -> not a federal pac, but should treat as such

```{r}
pac_manual_review %>% filter(pac_id %in% no_traverse_pac_ids)
pac_input %>% filter(Contributor.ID %in% no_traverse_pac_ids)
```

- confirmed these are all federal pacs or candidate pacs (at least they were labeled that way in the review)

# Raw graph walk

Starting from the seed PAC FILER_IDs in `pac_input.csv`, follow non-federal
PAC contributor edges recursively. No $5k threshold — all reachable PACs are
included regardless of contribution amount.

```{r}
walk_pac_graph <- function(seed_filer_ids, prepped, no_traverse_ids,
                           max_depth = 5) {
  seeds   <- seed_filer_ids[!is.na(seed_filer_ids) & seed_filer_ids != ""]
  queue   <- lapply(seeds, function(id) list(id = id, depth = 0L, is_seed = TRUE))
  visited <- tibble(pac_id = character(), depth = integer(), is_seed = logical(),
                    traversed = logical(), has_pac_no_id_contributors = logical())
  visited_ids <- character()

  while (length(queue) > 0) {
    node   <- queue[[1]]; queue <- queue[-1]
    pac_id <- node$id;    depth <- as.integer(node$depth)

    if (pac_id %in% visited_ids || depth > max_depth) next
    visited_ids <- c(visited_ids, pac_id)

    contribs <- prepped %>%
      filter(Recipient.Committee.ID == pac_id)

    # PACs classified as candidates/PP or federal: record but don't traverse.
    # Check committee name at visit time to catch PACs not pre-listed in
    # no_traverse_ids (e.g. discovered mid-walk without a pre-identified FILER_ID).
    pac_name <- contribs$Recipient.Committee[1]
    is_federal_by_name   <- !is.na(pac_name) && has_fed_pac_language(pac_name)
    is_candidate_by_name <- !is.na(pac_name) && is_candidate_pac(pac_name)

    if (pac_id %in% no_traverse_ids || is_federal_by_name || is_candidate_by_name) {
      visited <- bind_rows(visited,
                           tibble(pac_id = pac_id, depth = depth,
                                  is_seed = isTRUE(node$is_seed),
                                  traversed = FALSE,
                                  has_pac_no_id_contributors = FALSE))
      next
    }

    # non-federal, non-candidate PAC contributors with a resolved ID are queued
    to_traverse <- contribs %>%
      filter(!is_fed_pac, !is_candidate_pac(Contributor.Name), !is.na(std_contributor_id)) %>%
      pull(std_contributor_id) %>%
      unique()

    for (cid in to_traverse) {
      if (!cid %in% visited_ids)
        queue <- append(queue, list(list(id = cid, depth = depth + 1L, is_seed = FALSE)))
    }

    has_no_id <- any(contribs$is_pac_no_id, na.rm = TRUE)

    visited <- bind_rows(visited,
                         tibble(pac_id = pac_id, depth = depth,
                                is_seed = isTRUE(node$is_seed),
                                traversed = TRUE,
                                has_pac_no_id_contributors = has_no_id))
  }

  visited
}
```


```{r}
# Patch known FILER_ID gaps (mirroring 0803 manual corrections)
pac_input <- pac_input %>%
  mutate(FILER_ID = case_when(
    # don't think we need these anymore
    effective_Contributor.ID == "1479934" ~ "1479934", 
    effective_Contributor.ID == "1049814" ~ "1049814",
    TRUE ~ FILER_ID
  ))

seed_ids <- pac_input %>%
  filter(!is.na(FILER_ID), FILER_ID != "") %>%
  pull(FILER_ID) %>%
  unique()

cat(sprintf("Seeds: %d PAC FILER_IDs from pac_input.csv\n", length(seed_ids)))
```

```{r}
length(unique(pac_input$Contributor.ID))
```


```{r}
nrow(pac_input)
length(unique(pac_input$effective_Contributor.ID))

pac_input %>% filter(is.na(effective_Contributor.ID) | effective_Contributor.ID == "") %>% select(Contributor.Name, Contributor.ID, effective_Contributor.ID, FILER_ID, FILER_TYPE, is_fed_pac, has_pac_language,Contributor.City, Recipient.Committee)
```

### Walk PAC graph

```{r}
pac_universe <- walk_pac_graph(seed_ids, contributions, no_traverse_pac_ids,
                               max_depth = 5)

cat(sprintf(
  "PAC universe: %d PACs (%d seeds, %d discovered; %d traversed, %d not traversed)\n",
  nrow(pac_universe),
  sum(pac_universe$is_seed),
  sum(!pac_universe$is_seed),
  sum(pac_universe$traversed),
  sum(!pac_universe$traversed)
))

pac_universe %>%
  count(depth) %>%
  rename(n_pacs = n) %>%
  print()

no_id_pacs <- pac_universe %>% filter(traversed, has_pac_no_id_contributors)
if (nrow(no_id_pacs) > 0) {
  cat(sprintf(
    "Note: %d traversed PAC(s) have contributors with PAC language but no Contributor.ID —",
    nrow(no_id_pacs)
  ))
  cat(" indirect contributors may be missed. Address via ID propagation in 0803.\n")
}
```
56 seeds -> 1001 total traversal PACs... crazy difference.

```{r}
# PACs in the universe that were not traversed because they matched the candidate
# PAC name pattern. Review these — if any should be traversed, remove them from
# pac_manual_review or update the pattern.
candidate_pac_in_universe <- pac_universe %>%
  filter(!traversed) %>%
  left_join(
    contributions %>% distinct(Recipient.Committee.ID, Recipient.Committee),
    by = c("pac_id" = "Recipient.Committee.ID")
  ) %>%
  filter(is_candidate_pac(coalesce(Recipient.Committee, pac_id))) %>%
  select(pac_id, Recipient.Committee, depth, is_seed)
cat(sprintf("%d PAC(s) in universe flagged as candidate committees\n",
            nrow(candidate_pac_in_universe)))
print(candidate_pac_in_universe)
```
Current ones from new props look good


Ask about these cases: 
- Fresno Families for Affordability Supporting Sandra Celedon for Assembly 2026, sponsored by environmental and labor organizations
- Valley Families for Melissa Hurtado for Senate 2026   

Okay to not include for now.

```{r}
is_candidate_pac("ELECT DEMS NOW PAC FED ID# C00880062")
```


# Automated ID propagation

Propagate IDs for contributors to PACs in the universe that are missing
`effective_Contributor.ID`. Mirrors 0802's `full_id_propagation` approach:
normalize `NAML → name_v5`, join on standardized name, apply unique matches.
Ambiguous and unmatched cases are printed for manual addition to
`corrected_pac_ids`.

```{r}
full_id_lookup <- id_map_full %>%
  filter(!is.na(FILER_ID), FILER_ID != "", FILER_TYPE == "RECIPIENT COMMITTEE") %>%
  mutate(name_v5 = toupper(NAML) %>% # standardized_names(NAML) #str_remove_all("\\s*\\(.*?\\)") %>%
           str_squish()) %>%
  distinct(name_v5, FILER_ID) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(FILER_ID)) %>%
  ungroup()
```

```{r}
universe_pac_ids <- pac_universe %>% pull(pac_id) %>% unique()

no_id_in_universe <- contributions %>%
  filter(
    Recipient.Committee.ID %in% universe_pac_ids,
    is.na(effective_Contributor.ID) | trimws(effective_Contributor.ID) == ""
  ) %>%
  mutate(
    #name_v5 = standardize_names(Contributor.Name) %>%
    #  str_remove_all("\\s*\\(.*?\\)") %>%
    #  str_squish()
    name_v5 = str_squish(toupper(Contributor.Name))
    
  ) %>%
  distinct(Contributor.Name, name_v5)

full_id_propagation <- no_id_in_universe %>%
  inner_join(full_id_lookup, by = "name_v5") %>%
  mutate(needs_review = n_ids > 1, FILER_ID = as.character(FILER_ID)) %>%
  select(Contributor.Name, name_v5, propagated_id = FILER_ID, n_ids, needs_review) # name_v5

cat(sprintf(
  "Auto-propagated: %d name(s) (%d unique, %d ambiguous/needs review)\n",
  nrow(full_id_propagation),
  sum(!full_id_propagation$needs_review),
  sum(full_id_propagation$needs_review)
))
```
- not sure I want to do this - also need to update so it only looks at PACs active during 2025-2026
```{r}
full_id_propagation %>% distinct()
```

- lots of these look like major donors, not sure why they'd get an id b/c it should only be for recipeint committees
- issue would be if these PACs led to new contributors need to be resolved once ids are applied... 
- don't want to propagate without the parentheses - confuses unitemized contributions.

```{r}
id_map_full %>% filter(toupper(NAML) == "ADVOCACY ACTION FUND")
```

```{r}
contributions_full %>% filter(grepl("AMGEN INC",toupper(Contributor.Name))) %>% select(Contributor.Name,Contributor.ID,Recipient.Committee,Recipient.Committee.ID,Start.Date,Amount,Contributor.City)
```

Planned Parenthood Action Fund of the Pacific Southwest PAC	1280724
doesn't have the 496.57 contribution on 09/03 (but does have the 496 contribution) but also don't see it on FEC website

California New Car Dealers Association Issues PAC: only one missing to yes on 50 - amount matches, but the date doesn't...its 6/12 in cal access and 6/16 here
- lots of issues w/ yes on 50
- can we count as a candidate pac? 

California Chiropractic Association PAC
- also missing a contributor.id when donating to yes on 50; date is also not an exact match to ca acces

Amgen Inc State PAC -> not sure what's happening...

```{r}
99466.43 + 533.57
```

progresive era pac: 
almost always has contributor.id = 1449477
in the cases it DOESN'T -> 460s show that it was a federal pac contribution...
https://cal-access.sos.ca.gov/PDFGen/pdfgen.prg?filingid=3156695&amendid=1 
https://cal-access.sos.ca.gov/PDFGen/pdfgen.prg?filingid=3138211&amendid=0

But OTHER filings w/ this name show it w/ the pac id seen in cal access



Yes on 50, The Election Rigging Response Act, Governor Newsom?s Ballot Measure Committee -> why is this in the traversal? can we categorize it directly? 
	SUPPORTED: PROPOSITION 050 - ACA 8 (RIVAS) CONGRESSIONAL REDISTRICTING. (RES. CH. 156, 2025) : how does this come up in our data? \
	
	
### apply id propagations - CAN SKIP! (maybe wait til we deal w/ id propagation in a different way)

```{r}
# Apply unique matches; flag ambiguous for manual review via corrected_pac_ids.
auto_propagated <- full_id_propagation %>%
  filter(!needs_review) %>%
  distinct(Contributor.Name, propagated_id)

contributions <- contributions %>%
  left_join(auto_propagated, by = "Contributor.Name") %>%
  mutate(
    effective_Contributor.ID = coalesce(
      na_if(trimws(effective_Contributor.ID), ""), propagated_id
    ),
    std_contributor_id = coalesce(std_contributor_id, propagated_id)
  ) %>%
  select(-propagated_id)

# Print anything not matched or ambiguous — add manually to corrected_pac_ids
unmatched <- no_id_in_universe %>%
  anti_join(full_id_propagation, by = "Contributor.Name")
if (nrow(unmatched) > 0 || sum(full_id_propagation$needs_review) > 0) {
  cat("Needs manual ID (add to corrected_pac_ids sheet):\n")
  bind_rows(
    unmatched %>% mutate(reason = "no_match"),
    full_id_propagation %>% filter(needs_review) %>%
      select(Contributor.Name, name_v5) %>% mutate(reason = "ambiguous")
  ) %>% print()
}
```

```{r}
# Re-run walk with propagated IDs so newly-identified PACs are traversed.
pac_universe <- walk_pac_graph(seed_ids, contributions, no_traverse_pac_ids,
                               max_depth = 5)
cat(sprintf(
  "PAC universe after propagation: %d PACs (%d seeds, %d discovered; %d traversed)\n",
  nrow(pac_universe), sum(pac_universe$is_seed), sum(!pac_universe$is_seed),
  sum(pac_universe$traversed)
))
```

# Look at missing ids

Don't propagate ids when they were missing b/c of a self-contribution.

```{r}
no_id_pacs <- pac_universe %>% filter(traversed, has_pac_no_id_contributors)

contributions %>%
  filter(is_pac_no_id, Recipient.Committee.ID %in% no_id_pacs$pac_id) %>%
  distinct(Contributor.Name, Recipient.Committee.ID, Recipient.Committee,Amount) %>%
  arrange(Contributor.Name) %>% distinct(Contributor.Name)

```

- does introduce a lot of PACs w/out an id we would need to deal with 

```{r}
contributions_full %>% filter(Contributor.Name == "Teamsters Local Union 350 PAC")
```


```{r}
traversed_ids <- pac_universe %>% filter(traversed) %>% pull(pac_id)

# Candidate/measure committee IDs = what seed PACs contributed to
tracked_committee_ids <- contributions %>%
  filter(std_contributor_id %in% seed_ids) %>%
  pull(Recipient.Committee.ID) %>%
  unique()

# Traversed PACs that gave $5k+ to a relevant downstream recipient
qualifying_recipient_pacs <- contributions %>%
  filter(
    !is.na(std_contributor_id),
    std_contributor_id %in% traversed_ids,
    Recipient.Committee.ID %in% c(traversed_ids, tracked_committee_ids)
  ) %>%
  group_by(std_contributor_id, Recipient.Committee.ID) %>%
  summarise(total_given = sum(as.numeric(Amount), na.rm = TRUE), .groups = "drop") %>%
  filter(total_given >= 5000) %>%
  pull(std_contributor_id) %>%
  unique()

no_id_for_review <- contributions %>%
  filter(
    is_pac_no_id,
    Recipient.Committee.ID %in% qualifying_recipient_pacs
  ) %>%
  group_by(Contributor.Name, Recipient.Committee.ID, Recipient.Committee) %>%
  summarise(total_amount = sum(as.numeric(Amount), na.rm = TRUE), .groups = "drop") %>%
  filter(total_amount >= 5000) %>%
  arrange(desc(total_amount))

```

```{r}
no_id_for_review %>% select(Contributor.Name, Recipient.Committee,Recipient.Committee.ID,total_amount)
```


- don't want to worry about candidate pacs - adds a lot of work -> could we flag candidate pacs now and send to review? can update processing if changed

# Write output

```{r}
# Flag candidate and federal PACs in the universe so 0802 can exclude them from
# relevant_recipient_ids — contributions to these don't scope entity resolution.
all_candidate_pac_ids <- unique(c(
  pac_manual_review %>% filter(!is.na(pac_id), is_candidates_pp) %>% pull(pac_id),
  candidate_pac_ids_by_name
))

fed_pac_ids_by_name <- contributions %>%
  distinct(Recipient.Committee.ID, Recipient.Committee) %>%
  filter(has_fed_pac_language(Recipient.Committee)) %>%
  pull(Recipient.Committee.ID) %>%
  unique()

all_federal_pac_ids <- unique(c(
  fed_pac_from_input,
  pac_manual_review %>% filter(!is.na(pac_id), is_fec_federal) %>% pull(pac_id),
  fed_pac_ids_by_name
))

pac_universe <- pac_universe %>%
  mutate(
    is_candidate_pac = pac_id %in% all_candidate_pac_ids,
    is_federal_pac   = pac_id %in% all_federal_pac_ids
  )

cat(sprintf(
  "%d PAC(s) flagged is_candidate_pac; %d flagged is_federal_pac\n",
  sum(pac_universe$is_candidate_pac),
  sum(pac_universe$is_federal_pac)
))
```

```{r}
nrow(pac_universe)
```

```{r}
cand_pacs <- pac_universe %>% filter(is_candidate_pac == TRUE) %>% pull(pac_id)
contributions_full %>% filter(Recipient.Committee.ID %in% cand_pacs) %>% select(Recipient.Committee) %>% distinct()

fed_pacs <- pac_universe %>% filter(is_federal_pac == TRUE) %>% pull(pac_id)
contributions_full %>% filter(Recipient.Committee.ID %in% fed_pacs) %>% select(Recipient.Committee) %>% distinct()
```
```{r}
length(unique(pac_universe$pac_id))
```

```{r}
head(pac_universe)
```


```{r}
#write.csv(pac_universe, "08_outputs/pac_universe.csv", row.names = FALSE)
write.csv(pac_universe, "08_outputs/pac_universe_new_props.csv", row.names = FALSE)
cat(sprintf("pac_universe.csv written: %d rows\n", nrow(pac_universe)))
```
