---
title: "R Notebook"
output: html_notebook
---


# Load packages

```{r}
here::i_am("0801_entity_resolution.Rmd")
library(here)
library(dplyr)
library(stringr)
library(stringdist)
library(reclin2)
library(phonics)
library(igraph)
library(readr)
library(digest)
library(jsonlite)
library(tidyr)

source("standardization_helpers.R")
```

```{r}
hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16))
}
```


# Load and standardize all contributors data

```{r}
#contributions_full <- read.csv("08_inputs/contributors_data_full.csv")
contributions_full <- read.csv("08_inputs/all_contributors_processed.csv")
nrow(contributions_full)
```



### load already categorized data

```{r}
categorized <- read.csv("08_inputs/already_classified_contributions.csv")
nrow(categorized)
```


### look at aliases by contributor id

```{r, eval = FALSE}
contributions_full %>% group_by(Contributor.ID) %>% mutate(count = n_distinct(Contributor.Name)) %>% arrange(Contributor.ID) %>% filter(count > 1) %>% select(Contributor.Name, Contributor.ID) %>% unique()
```


### filter for contributions data we're interested in

```{r}
contributions <- contributions_full %>% filter(
    Ballot.Measure == "OPPOSED: PROPOSITION 040 - IMPOSES ONE-TIME TAX ON CERTAIN INDIVIDUALS AND TRUSTS. INITIATIVE CONSTITUTIONAL AMENDMENT AND STATUTE." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 040 - IMPOSES ONE-TIME TAX ON CERTAIN INDIVIDUALS AND TRUSTS. INITIATIVE CONSTITUTIONAL AMENDMENT AND STATUTE." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 041 - REQUIRES AUDITS OF PROGRAMS FUNDED BY NEW STATE SPECIAL TAXES. PROHIBITS NEW STATE TAXES THAT ARE EXCLUDED FROM EXISTING VOTER-APPROVED STATE SPENDING LIMIT..." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 042 - PROHIBITS NEW STATE PERSONAL PROPERTY TAXES AND CERTAIN RETROACTIVE STATE TAXES. INITIATIVE CONSTITUTIONAL AMENDMENT." | 
    Recipient.Name == "BECERRA, XAVIER" | 
    Recipient.Name == "HILTON, STEVE" | 
    Office == "Insurance Commissioner"
) %>%
  mutate(
    Amount = as.numeric(Amount)
  ) #%>% 
  #rename(
  #  Ballot.Measure = 'Ballot.Measure(s)'
  #)

nrow(contributions)
```



```{r}
# make sure we' pulled everything (compare to power search)
contributions %>% filter(Recipient.Name == "BECERRA, XAVIER") %>% summarise(total = sum(Amount))
contributions %>% filter(Recipient.Name == "HILTON, STEVE") %>% summarise(total = sum(Amount))
```


```{r}
# flag contributions by race/prop
contributions <- contributions %>%
  mutate(
    race_prop = case_when(
      Ballot.Measure.Contribution == "Y" ~ Ballot.Measure,
      !is.na(Office) & Office != "" ~ Office,
      TRUE ~ Recipient.Name
    )
  )
unique(contributions$race_prop)
```




### standardize contributions data


```{r}
make_row_hash <- function(...) {
  cols <- list(...)
  n <- length(cols[[1]])
  vapply(seq_len(n), function(i) {
    parts <- vapply(cols, function(col) as.character(col[i]), character(1))
    digest::digest(paste(parts, collapse = "|"), algo = "xxhash32")
  }, character(1))
}

```


```{r}
contributions <- contributions %>%
  filter(Contributor.Name != "Unitemized Contributions") %>%
  mutate(
    standardized_name = standardize_names(Contributor.Name) %>% fix_typos(),
    processed_name = standardized_name %>%
      remove_pac_info() %>% remove_unit_info() %>% 
      str_remove_all("\\s*\\(.*?\\)") %>% replace_business_text() %>% str_squish(),

    standardized_employer_name = standardize_names(Contributor.Employer) %>% fix_typos(),
    processed_employer_name = standardized_employer_name %>% standardize_occupation_employer() %>%
      remove_pac_info() %>% remove_unit_info() %>% str_remove_all("\\s*\\(.*?\\)") %>%
      replace_business_text() %>%
      str_squish(),

    standardized_occupation = standardize_names(Contributor.Occupation),
    processed_occupation    = standardize_occupation_employer(standardized_occupation),
    standardized_city       = standardize_names(Contributor.City),
    zip_code_processed      = substr(Contributor.Zip.Code, 1, 5),

    has_pac_language = has_pac_language(Contributor.Name),
    entity_type      = if_else(sapply(standardized_name, is_individual), "individual", "organization"),
    row_id           = row_number(),
    contribution_id  = make_row_hash(Contributor.Name, Contributor.ID, Amount, race_prop,
                                      Contributor.City, Contributor.Employer,
                                      Contributor.Occupation, Contributor.Zip.Code, Start.Date)
  )

```

```{r}
contributions <- read.csv("08_inputs/race_contributions_processed.csv")
```


```{r, eval = FALSE}
contributions %>% group_by(contribution_id) %>% summarise(count = n()) %>% filter(count > 1)
```


## manually process edge cases

Manually verified each of these. 

```{r}
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$standardized_name <- "LOVECCHIO, PETER"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$processed_name <- "LOVECCHIO PETER"

contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$standardized_name <- "ZUNILDA, PEREZ"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$processed_name <- "ZUNILDA, PEREZ"

contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$standardized_name <- "LIDELL, JOHN MARK" # PA-C = physician assistant (don't want to add PAC to suffix list though)
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$processed_name <- "LIDELL JOHN MARK"

contributions[which(contributions$Contributor.Name == "Last0b6df6bd, First7606aee3"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lastcbea8f37, Firstd599d312"),]$entity_type <- "individual"
```



## individuals

- Make sure employer and occupation are NONE for individuals where these columns are blank 
- Could also update employer to UNKNOWN for individuals where employer is N/A, but occupation specifies they're self-employed
  - Are cases where employer is N/A, but occupation indicates they have a job.

```{r}
contributions <- contributions %>%
  mutate(
    standardized_employer_name   = if_else(entity_type == "individual" & standardized_employer_name == "", "NONE", standardized_employer_name),
    processed_employer_name  = if_else(entity_type == "individual" & processed_employer_name == "", "NONE", processed_employer_name),

    standardized_occupation  = if_else(entity_type == "individual" & standardized_occupation == "", "NONE", standardized_occupation),
    processed_occupation = if_else(entity_type == "individual" & processed_occupation == "", "NONE", processed_occupation),    
  )

```


# Load and standardize name maps

```{r}
map1 <- read.csv("08_inputs/dd_names_mapped_to_org_concepts_3_18_2026.csv") %>%
  rename(canonical_name = org_concept_name, alias_group_id = oid) %>%
  mutate(
    record_name    = str_squish(record_name),
    canonical_name = str_squish(canonical_name),
    mapping_source = "dd_name_mapping",
    confirmed      = TRUE
  )

map2 <- read.csv("08_inputs/2024 Contributor name mapping and descriptions - Combined.csv") %>%
  select(record_name = Original, canonical_name = Normalized) %>%
  filter(!is.na(record_name) & record_name != "") %>%
  mutate(
    record_name    = str_squish(record_name),
    canonical_name = str_squish(canonical_name),
    alias_group_id = NA_integer_,
    mapping_source = "2024_combined",
    confirmed      = TRUE
  )

# combine — prefer map1 when both have same record_name
reference_mapping <- bind_rows(map1, map2) %>%
  group_by(record_name) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(record_name_std = standardize_names(record_name) %>% fix_typos())

```


# Organizations

```{r}
make_name_variants <- function(df, col) {
  df %>% mutate(
    name_v1 = standardize_names(.data[[col]]) %>% fix_typos(),
    name_v2 = remove_pac_info(name_v1)        %>% str_squish(), # removing FEC ID, FED PAC ID, MAJOR DONOR ID
    #name_v3 = remove_unit_info(name_v2)        %>% str_squish(),
    name_v4 = str_remove_all(name_v1, "\\s*\\(.*?\\)") %>% str_squish(),
    name_v5 = replace_business_text(name_v4)   %>% str_squish()
    #name_v6 = str_remove_all(name_v5, "\\s*\\(.*?\\)") %>% str_squish()
  )
}

find_exact_variant_pairs <- function(pool, variant_col) {
  candidates <- pool %>%
    filter(!is.na(.data[[variant_col]]), .data[[variant_col]] != "") %>%
    group_by(across(all_of(c(variant_col, "has_pac_language")))) %>%
    filter(n() > 1) %>%
    ungroup()

  if (nrow(candidates) == 0)
    return(tibble(.x = integer(), .y = integer(), match_source = character()))

  candidates %>%
    group_by(across(all_of(c(variant_col, "has_pac_language")))) %>%
    group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>%
    ungroup() %>%
    select(.x, .y) %>%
    mutate(match_source = paste0("exact_", variant_col))
}
```


```{r}
orgs <- contributions %>%
  filter(entity_type == "organization") %>%
  mutate(cont_row_id = row_id) %>%          # preserve row_id before overwriting
  mutate(has_contributor_id = !is.na(Contributor.ID) & str_trim(as.character(Contributor.ID)) != "") %>%
  make_name_variants("Contributor.Name") %>%
  filter(!is.na(name_v1), name_v1 != "") %>%
  mutate(row_id = row_number())
```

## Step 1: ID-based aliases (same Contributor.ID, different names)


```{r}
id_based_aliases <- orgs %>%
  filter(has_contributor_id) %>%
  group_by(Contributor.ID) %>%
  filter(n_distinct(Contributor.Name) > 1) %>%
  ungroup()

id_auto_pairs <- id_based_aliases %>%
  group_by(Contributor.ID) %>%
  group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>%
  ungroup() %>%
  mutate(link_tier = "auto_link", match_source = "same_contributor_id")

cat("ID-based alias pairs:", nrow(id_auto_pairs), "\n")

id_based_aliases %>%
  group_by(Contributor.ID) %>%
  summarise(names = paste(unique(Contributor.Name), collapse = " | "), n = n(), .groups = "drop") %>%
  arrange(desc(n))
```


## Step 2: Exact variant cascade

```{r}
id_resolved_rows <- unique(c(id_auto_pairs$.x, id_auto_pairs$.y))
remaining_orgs   <- orgs %>% filter(!row_id %in% id_resolved_rows)

#variant_cols     <- c("name_v1", "name_v2", "name_v3", "name_v4", "name_v5", "name_v6")
variant_cols     <- c("name_v1", "name_v2", "name_v4", "name_v5") # don't do variants where we remove unit info or words attributed with PACs
exact_pairs_list <- list()

for (v in variant_cols) {
  new_pairs <- find_exact_variant_pairs(remaining_orgs, v)
  if (nrow(new_pairs) > 0) {
    exact_pairs_list[[v]] <- new_pairs
    remaining_orgs <- remaining_orgs %>%
      filter(!row_id %in% unique(c(new_pairs$.x, new_pairs$.y)))
  }
  cat("variant", v, "- new pairs:", nrow(new_pairs), "| remaining:", nrow(remaining_orgs), "\n")
}

exact_pairs <- bind_rows(exact_pairs_list) %>%
  left_join(orgs %>% select(row_id, id_x = Contributor.ID, has_id_x = has_contributor_id), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, id_y = Contributor.ID, has_id_y = has_contributor_id), by = c(".y" = "row_id")) %>%
  filter(!(coalesce(has_id_x, FALSE) & coalesce(has_id_y, FALSE) & coalesce(id_x != id_y, FALSE))) %>%
  mutate(link_tier = "auto_link") %>%
  select(.x, .y, link_tier, match_source)

cat("Exact cascade auto-links:", nrow(exact_pairs), "| remaining for fuzzy:", nrow(remaining_orgs), "\n")
```


### Step 3: internal fuzzy match - Skip in simplified version


## Combine + cluster

Combine all links, including those that need to be reviewed into a single entity
- not including fuzzy pairs and bridge pairs for now

```{r}
# pairs computed through different linkage methods
all_org_pairs <- bind_rows(
  id_auto_pairs %>% select(.x, .y, link_tier, match_source),
  exact_pairs,
  #org_pairs    %>% select(.x, .y, link_tier, match_source, name_jw, name_cosine, name_sim),
  #bridge_only  %>% select(.x, .y, link_tier, match_source, name_jw, name_cosine, name_sim, canonical_name)
)

# build one graph from all non-rejected pairs, treating review links as true
linkable_pairs <- all_org_pairs %>% filter(link_tier %in% c("auto_link", "review"))

if (nrow(linkable_pairs) > 0) {
  g_orgs       <- graph_from_data_frame(linkable_pairs %>% select(.x, .y), directed = FALSE)
  org_clusters <- components(g_orgs)$membership
  org_cluster_map <- tibble(row_id = as.integer(names(org_clusters)), cluster_num = org_clusters) %>%
    left_join(orgs %>% select(row_id, Contributor.ID, name_v5), by = "row_id") %>%
    group_by(cluster_num) %>%
    mutate(member_key = paste0(coalesce(as.character(Contributor.ID), ""), "::", name_v5),
           entity_id  = hash_id(member_key, "org")) %>%
    ungroup() %>%
    select(row_id, entity_id)
} else {
  org_cluster_map <- tibble(row_id = integer(), entity_id = character())
}

orgs_singletons <- orgs %>%
  filter(!row_id %in% org_cluster_map$row_id) %>%
  mutate(member_key = paste0(coalesce(as.character(Contributor.ID), ""), "::", name_v5),
         entity_id  = vapply(member_key, hash_id, character(1), prefix = "org")) %>%
  select(row_id, entity_id)

review_org_entity_ids <- all_org_pairs %>%
  filter(link_tier == "review") %>%
  left_join(bind_rows(org_cluster_map, orgs_singletons), by = c(".x" = "row_id"),
            relationship = "many-to-one") %>%
  filter(!is.na(entity_id)) %>%
  pull(entity_id) %>%
  unique()

orgs_resolved <- orgs %>%
  left_join(bind_rows(org_cluster_map, orgs_singletons), by = "row_id") %>%
  mutate(needs_review = entity_id %in% review_org_entity_ids)

```


# INDIVIDUALS:

```{r, eval = FALSE}
# parse names 
parsed_cols <- parse_names_df(
  contributions %>% filter(entity_type == "individual") %>% pull(standardized_name) # use standardized name here because it has the commas still in it
)
```


```{r}
individuals <- contributions %>%
  filter(entity_type == "individual") %>%
  bind_cols(parsed_cols %>% rename(last_name = last, first_name = first)) %>%
  mutate(
    original_row_id = row_id,       # preserve for final join back to x
    row_id          = row_number(), # sequential for pair_blocking compatibility
    first_initial   = substr(first_name, 1, 1),
    last_meta       = safe_metaphone(last_name)
  ) %>%
  filter(!is.na(last_meta), last_meta != "")

```

### group by name

- Group individuals if they have the same name

```{r}
# assign entity id if they share the same name
individuals <- individuals %>%
  mutate(
    member_key = paste0(standardized_name, "::"), # processed_employer_name
    entity_id  = vapply(member_key, hash_id, character(1), prefix = "indiv")
  ) %>%
  select(-member_key)

individuals <- individuals %>%
  group_by(entity_id, race_prop) %>%
  mutate(
    race_total_amount = sum(Amount, na.rm = TRUE),
    crosses_threshold = race_total_amount >= 5000
  ) %>%
  ungroup()


```


# Classification input

```{r}
orgs_resolved <- orgs_resolved %>%
  group_by(entity_id, race_prop) %>%
  mutate(
    race_total_amount = sum(Amount, na.rm = TRUE),
    crosses_threshold = race_total_amount >= 5000
  ) %>%
  ungroup()

```

```{r}
individuals_classification_input <- individuals %>%
  filter(crosses_threshold)

nrow(individuals_classification_input)
n_distinct(individuals_classification_input$entity_id)

```


```{r}
classification_input <- bind_rows(
  individuals    %>% filter(crosses_threshold),
  orgs_resolved  %>% filter(crosses_threshold)
)

nrow(classification_input)
n_distinct(classification_input$entity_id)

```


### match back uuid from previous classifications

```{r}
pad_zip <- function(z) {
  z <- trimws(as.character(z))
  needs_pad <- !is.na(z) & z != "" & str_detect(z, "^[0-9]+$") & nchar(z) < 5
  z[needs_pad] <- paste0(strrep("0", 5 - nchar(z[needs_pad])), z[needs_pad])
  substr(z, 1, 5)
}
```

```{r}
normalize_missing <- function(x) {
  x <- trimws(as.character(x))
  if_else(x %in% c("", "N/A", "NA", "None","n/a") | is.na(x), "N/A", x)
}
```

```{r}
categorized <- categorized %>%
  mutate(
    race_prop = "Governor",
    legacy_match_id = make_row_hash(
      Contributor.Name, Contributor.ID, as.numeric(Amount), race_prop,
      Contributor.City, normalize_missing(Contributor.Employer),
      normalize_missing(Contributor.Occupation), pad_zip(Contributor.Zip.Code),
      Start.Date
    )
  ) %>%
  group_by(legacy_match_id) %>%
  mutate(dup_rank = row_number()) %>%
  ungroup()

classification_input <- classification_input %>%
  mutate(
    legacy_match_id = make_row_hash(
      Contributor.Name, Contributor.ID, as.numeric(Amount), race_prop,
      Contributor.City, normalize_missing(Contributor.Employer),
      normalize_missing(Contributor.Occupation), pad_zip(Contributor.Zip.Code),
      Start.Date
    )
  ) %>%
  group_by(legacy_match_id) %>%
  mutate(dup_rank = row_number()) %>%
  ungroup() %>%
  left_join(
    categorized %>% select(legacy_match_id, dup_rank, uuid),
    by = c("legacy_match_id", "dup_rank")
  ) %>%
  select(-dup_rank)

```



```{r}
# data checks
nrow(classification_input)
nrow(classification_input_test)
n_distinct(classification_input_test$uuid) == nrow(classification_input_test)

classification_input_test %>% count(uuid) %>% filter(n > 1) # should only be NAs

# make sure match id is unique
categorized %>% count(legacy_match_id, dup_rank) %>% filter(n > 1)

nrow(categorized)
length(which(!is.na(classification_input_test$uuid)))
```

- Middleteon, Geraldine: ended up not aggregating to $5k because of negative contributions that lowered her aggregated spending


```{r}
ambiguous <- categorized %>%
  group_by(legacy_match_id) %>%
  filter(n() > 1) %>%
  ungroup()

nrow(ambiguous)                     
n_distinct(ambiguous$legacy_match_id)   

```


```{r}
# add remaining uuids
namespace <- "887a6583-910c-47b0-b373-d5b2863e8c8a"
classification_input[which(is.na(classification_input$uuid)),]$uuid <-
  uuid::UUIDgenerate(n = nrow(classification_input[which(is.na(classification_input$uuid)),]))
```


# Output PAC list

(1) Look for anything with a missing contributor.id
(2) output list of all rows with a contributor.id
  
### look for any names that match a recipient committee - deal with major donors/IE later on


```{r}
# data to map contributor ids
id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv")
id_map <- unique(id_map_full[,c("XREF_FILER_ID","FILER_ID","FILER_TYPE")]) # get unique xref id / filer ids combos
```

```{r}
# convert all ids to character strings
id_map$FILER_ID         <- as.character(id_map$FILER_ID)   
id_map$XREF_FILER_ID <- as.character(id_map$XREF_FILER_ID)
classification_input$Contributor.ID <- as.character(classification_input$Contributor.ID)
```

```{r}
# map filer id to contributor id (different ids may be used internally vs externally)
classification_input_test <- left_join(classification_input, id_map, by = c("Contributor.ID" = "XREF_FILER_ID"))

# check that each has a corresponding filer id
nrow(classification_input_test[which(is.na(classification_input_test$FILER_ID) & !is.na(classification_input_test$Contributor.ID)),])
```


```{r}
# Build name/ID lookup from full dataset
full_id_lookup <- id_map_full %>%
  filter(!is.na(FILER_ID), FILER_ID != "",FILER_TYPE == "RECIPIENT COMMITTEE") %>% # , (FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" | FILER_TYPE == "RECIPIENT COMMITTEE")
  mutate(name_v5 = standardize_names(NAML) %>% 
           str_remove_all("\\s*\\(.*?\\)") %>%
          str_squish()) %>%
  distinct(name_v5, FILER_ID, FILER_TYPE) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(FILER_ID)) %>%
  ungroup()


```

Different filer types should have different ids

```{r}
# Propagate IDs to orgs that are missing them
full_id_propagation <- classification_input %>%
  filter(is.na(Contributor.ID)) %>%
  select(row_id, Contributor.Name, name_v5) %>%
  inner_join(full_id_lookup, by = "name_v5") %>%
  mutate(needs_review = n_ids > 1) %>%
  select(row_id, Contributor.Name, propagated_id = FILER_ID, n_ids, needs_review, FILER_TYPE)
full_id_propagation
```


```{r}
# apply ids back; Contributor.ID stays raw — effective_Contributor.ID includes propagated ids
classification_input <- classification_input %>%
  left_join(
    full_id_propagation %>% distinct(row_id, propagated_id),
    by = "row_id"
  ) %>%
  mutate(
    effective_Contributor.ID = coalesce(Contributor.ID, propagated_id),
    has_contributor_id       = !is.na(effective_Contributor.ID) & str_trim(as.character(effective_Contributor.ID)) != ""
  ) %>%
  select(-propagated_id)
```





# WRITE OUTPUTS

- Update this so major donors don't go to pac processing unless they have pac language

```{r}
# anything with a contributor id goes to pacs
write.csv(classification_input[which(!is.na(classification_input$effective_Contributor.ID)),],"classification_input.csv",row.names = FALSE)
write.csv(classification_input[which(!is.na(classification_input$effective_Contributor.ID)),],"pac_input.csv",row.names = FALSE)
```







