---
title: "R Notebook"
output: html_notebook
---

# Summary: 
This notebook reads in the data pulled in `0801_data_pull.R` and outputs a dataset of contributors that need to be classified.
- Notebook aggregated contributors to determine which ones meet the $5k threshold.
- Matches back uuids from previously classified contribution records (records that already went through the pipeline)
- Propagates Contributor.ID to contributions with a missing contributor.id
- Outputs a file of org/individual contributors and a file of PAC contributors that needs to go through PAC processing
- This notebook also currently does records linkage with fuzzy matching, name variant matching and a custom record linkage 
  algorithm for individuals to surface name variants to review (this is going to be moved into a separate notebook)

# Load packages

```{r}
here::i_am("0802_entity_resolution_simplified.Rmd")
library(here)
library(dplyr)
library(stringr)
library(stringdist)
library(reclin2)
library(phonics)
library(igraph)
library(readr)
library(digest)
library(jsonlite)
library(tidyr)
library(uuid)

source("standardization_helpers.R")
```

```{r}
safe_metaphone <- function(x) {
  clean <- str_replace_all(coalesce(x, ""), "[^A-Za-z]", "")
  ifelse(clean != "", metaphone(clean), NA_character_)
}

bigram_cosine <- function(a, b) stringsim(a, b, method = "cosine", q = 2)

hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16))
}

normalize_str <- function(x) {
  coalesce(x, "") %>% toupper() %>%
    stringr::str_replace_all("[^A-Z0-9 ]+", " ") %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_trim()
}

# look up contributor name in entity registry
# returns entity_id, is_conflict and is_new

lookup_cluster_uuid <- function(raw_names, mapping_df) {
  clean <- unique(str_squish(toupper(raw_names)))
  matched_uuids <- mapping_df %>%
    filter(raw_name %in% clean) %>%
    pull(entity_uuid) %>%
    unique()

  if (length(matched_uuids) == 0) {
    list(entity_id = UUIDgenerate(), is_conflict = FALSE, is_new = TRUE)
  } else if (length(matched_uuids) == 1) {
    list(entity_id = matched_uuids[1], is_conflict = FALSE, is_new = FALSE)
  } else {
    list(entity_id = matched_uuids[1], is_conflict = TRUE, is_new = FALSE)
  }
}
```

# Load and standardize all contributors data

```{r}
#contributions_full <- read.csv("08_inputs/contributors_data_full.csv")
#contributions_full <- read.csv("08_inputs/all_contributors_processed.csv")
contributions_full <- read.csv("08_inputs/power_search_contributions_normalized.csv")
nrow(contributions_full)
```

- need to deal with unitemized contributions


### load already categorized data

```{r}
categorized <- read.csv("08_inputs/already_classified_contributions.csv")
nrow(categorized)
```

### filter for contributions data we're interested in

```{r}
contributions <- contributions_full %>% 
  rename(
    Ballot.Measure = `Ballot.Measure.s.`
  ) %>% 
  filter(
    Ballot.Measure == "OPPOSED: PROPOSITION 040 - IMPOSES ONE-TIME TAX ON CERTAIN INDIVIDUALS AND TRUSTS. INITIATIVE CONSTITUTIONAL AMENDMENT AND STATUTE." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 040 - IMPOSES ONE-TIME TAX ON CERTAIN INDIVIDUALS AND TRUSTS. INITIATIVE CONSTITUTIONAL AMENDMENT AND STATUTE." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 041 - REQUIRES AUDITS OF PROGRAMS FUNDED BY NEW STATE SPECIAL TAXES. PROHIBITS NEW STATE TAXES THAT ARE EXCLUDED FROM EXISTING VOTER-APPROVED STATE SPENDING LIMIT..." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 042 - PROHIBITS NEW STATE PERSONAL PROPERTY TAXES AND CERTAIN RETROACTIVE STATE TAXES. INITIATIVE CONSTITUTIONAL AMENDMENT." | 
    Recipient.Name == "BECERRA, XAVIER" | 
    Recipient.Name == "HILTON, STEVE" | 
    Office == "Insurance Commissioner"
) #%>%
  #mutate(
  #  Amount = as.numeric(Amount)
  #) #%>% 

nrow(contributions)

# flag contributions by race/prop
#contributions <- contributions %>%
#  mutate(
#    race_prop = case_when(
#      Ballot.Measure.Contribution == "Y" ~ Ballot.Measure,
#      !is.na(Office) & Office != "" ~ Office,
#      TRUE ~ Recipient.Name
#    )
#  )
unique(contributions$race_prop)
```

```{r}
# make sure we' pulled everything (compare to power search)
contributions %>% filter(Recipient.Name == "BECERRA, XAVIER") %>% summarise(total = sum(Amount))
contributions %>% filter(Recipient.Name == "HILTON, STEVE") %>% summarise(total = sum(Amount))
```
- not fully updated maybe? 

# Load PACs from graph traversal

PAC graph traversal is run in 0803 and will need to be completed before loading this.

```{r}
pacs <- read.csv("08_outputs/pac_classifiability.csv")
nrow(pacs)
```

This includes both seed PACs and PACs that contributed to other PACs (surface in graph traversal)

```{r}
#pacs %>% filter(pac_id %in% contributions$Contributor.ID)

# pull all contributors to these PACs
pac_contributors <- contributions_full %>% filter(Recipient.Committee.ID %in% pacs$pac_id)
```


# Combine direct and pac contributor data

```{r}
contributions <- bind_rows(contributions,pac_contributors)
nrow(contributions)
```


### standardize contributions data


```{r}
make_row_hash <- function(...) {
  cols <- list(...)
  n <- length(cols[[1]])
  vapply(seq_len(n), function(i) {
    parts <- vapply(cols, function(col) as.character(col[i]), character(1))
    digest::digest(paste(parts, collapse = "|"), algo = "xxhash32")
  }, character(1))
}

```


## manually process edge cases

Manually verified each of these. 

```{r}
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$standardized_name <- "LOVECCHIO, PETER"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$processed_name <- "LOVECCHIO PETER"

contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$standardized_name <- "ZUNILDA, PEREZ"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$processed_name <- "ZUNILDA, PEREZ"

contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$standardized_name <- "LIDELL, JOHN MARK" # PA-C = physician assistant (don't want to add PAC to suffix list though)
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$processed_name <- "LIDELL JOHN MARK"

contributions[which(contributions$Contributor.Name == "Last0b6df6bd, First7606aee3"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lastcbea8f37, Firstd599d312"),]$entity_type <- "individual"
```

## individuals

- Make sure employer and occupation are NONE for individuals where these columns are blank 
- Could also update employer to UNKNOWN for individuals where employer is N/A, but occupation specifies they're self-employed
  - Are cases where employer is N/A, but occupation indicates they have a job.

```{r}
contributions <- contributions %>%
  mutate(
    standardized_employer_name   = if_else(entity_type == "individual" & standardized_employer_name == "", "NONE", standardized_employer_name),
    processed_employer_name  = if_else(entity_type == "individual" & processed_employer_name == "", "NONE", processed_employer_name),

    standardized_occupation  = if_else(entity_type == "individual" & standardized_occupation == "", "NONE", standardized_occupation),
    processed_occupation = if_else(entity_type == "individual" & processed_occupation == "", "NONE", processed_occupation),    
  )

```

# Load and standardize name maps

```{r}
map1 <- read.csv("08_inputs/dd_names_mapped_to_org_concepts_3_18_2026.csv") %>%
  rename(canonical_name = org_concept_name, alias_group_id = oid) %>%
  mutate(
    record_name    = str_squish(record_name),
    canonical_name = str_squish(canonical_name),
    mapping_source = "dd_name_mapping",
    confirmed      = TRUE
  )

map2 <- read.csv("08_inputs/2026 Contributor name mapping and descriptions - Names.csv") %>%
  select(record_name = Original, canonical_name = Normalized) %>%
  filter(!is.na(record_name) & record_name != "") %>%
  mutate(
    record_name    = str_squish(record_name),
    canonical_name = str_squish(canonical_name),
    alias_group_id = NA_integer_,
    mapping_source = "2024_combined",
    confirmed      = TRUE
  )

# combine — prefer map1 when both have same record_name
reference_mapping <- bind_rows(map1, map2) %>%
  group_by(record_name) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(record_name_std = standardize_names(record_name) %>% fix_typos())

```

# Entity Registry

Syncs the 2026 Names CSV (used for normalization) and the name_entity_mapping.csv (tracks uuids per entity).

```{r}
ENTITY_REGISTRY_PATH     <- here("08_inputs/entity_registry.csv")
NAME_ENTITY_MAPPING_PATH <- here("08_inputs/name_entity_mapping.csv")
FEEDBACK_PATH            <- here("08_inputs/org_name_variant_review_queue_w_review.csv")

# official source of truth
names_2026 <- read.csv(
  here("08_inputs/2026 Contributor name mapping and descriptions - Names.csv"),
  stringsAsFactors = FALSE
) %>%
  rename(raw_name = Original, canonical_name = Normalized) %>%
  filter(!is.na(raw_name), raw_name != "", !is.na(canonical_name), canonical_name != "") %>%
  mutate(
    raw_name       = str_squish(toupper(raw_name)),
    canonical_name = str_squish(canonical_name)
  )

# persistent UUID store — updated by manual review
if (file.exists(NAME_ENTITY_MAPPING_PATH)) {
  name_entity_mapping_existing <- read.csv(NAME_ENTITY_MAPPING_PATH, stringsAsFactors = FALSE) %>%
    mutate(confirmed = as.logical(confirmed))
} else {
  name_entity_mapping_existing <- tibble(raw_name = character(), entity_uuid = character(),
                                          canonical_name = character(), confirmed = logical(),
                                          added_date = character(), source = character())
}
```

```{r}
canonical_uuid_lookup <- name_entity_mapping_existing %>%
  filter(!is.na(canonical_name), confirmed == TRUE) %>%
  mutate(canonical_name = str_squish(canonical_name)) %>%
  group_by(canonical_name) %>% # make sure multiple uuids are never used per canonical name - this needs updates
  slice(1) %>%
  ungroup() %>%
  select(canonical_name, entity_uuid)

canonical_uuid_map <- names_2026 %>%
  mutate(canonical_name = str_squish(canonical_name)) %>%
  distinct(canonical_name) %>%
  left_join(canonical_uuid_lookup, by = "canonical_name") %>%
  mutate(entity_uuid = coalesce(entity_uuid, UUIDgenerate(n = n())))

cat(sum(is.na(left_join(names_2026 %>% distinct(canonical_name),
                         canonical_uuid_lookup, by="canonical_name")$entity_uuid)),
    "new canonical groups in 2026 Names CSV assigned fresh UUIDs\n")
```

New entities added since last time (not by me)


### data checks

```{r}
temp <- left_join(names_2026 %>% distinct(canonical_name),
          canonical_uuid_lookup, by = "canonical_name")
temp %>% filter(is.na(entity_uuid))
```
- others are new..

```{r}
name_entity_mapping_existing %>%
  filter(!is.na(canonical_name), confirmed == TRUE) %>%
  group_by(canonical_name) %>% filter(n_distinct(entity_uuid) > 1) %>% arrange(canonical_name)
```

- Need to address these cases. 
  - think ones i had to address came from mistakes of not updating uuid when i found is actually matched to an existing entity that wasn't originally flagged

```{r}
# Reviewed names registry: tracks ALL names where a review decision was made.
# Covers canonical names and singletons that don't need an entry in name_entity_mapping.
REVIEWED_NAMES_PATH <- here("08_inputs/reviewed_names.csv")

if (file.exists(FEEDBACK_PATH)) {
  feedback_raw <- read.csv(FEEDBACK_PATH, stringsAsFactors = FALSE)
  req_cols     <- c("to_add_status", "entity_name", "canonical_name", "review_group_id", "confirmed_link")
  if (nrow(feedback_raw) > 0 && all(req_cols %in% names(feedback_raw))) {

    # rrefer canonical_uuid_map (canonical_name → stable UUID).
    # for rows not matched to a canonical (rejected links, singletons), assign a UUID per
    # review_group_id — NOT entity_id, which is shared across split sub-groups in the spreadsheet.
    feedback_settled <- feedback_raw %>%
      filter(
        to_add_status %in% c("in existing map", "name is already canonical", "add to map", "don't add"),
        coalesce(confirmed_link, "") %in% c("yes", "no")
      ) %>%
      mutate(raw_name = str_squish(toupper(entity_name))) %>%
      left_join(canonical_uuid_map %>% select(canonical_name, canonical_uuid = entity_uuid),
                by = "canonical_name")

    group_uuid_map <- feedback_settled %>%
      filter(is.na(canonical_uuid)) %>% # no canonical uuid matched from the official registry
      distinct(review_group_id) %>% # new uuids assigned per group id
      mutate(group_uuid = UUIDgenerate(n = n()))

    reviewed_from_feedback <- feedback_settled %>%
      left_join(group_uuid_map, by = "review_group_id") %>%
      transmute(
        raw_name      = raw_name,
        entity_uuid   = coalesce(canonical_uuid, group_uuid),
        reviewed_date = as.character(Sys.Date()),
        source        = "manual_review"
      ) %>%
      filter(!is.na(raw_name), raw_name != "", !is.na(entity_uuid), entity_uuid != "")

  } else {
    reviewed_from_feedback <- tibble(raw_name=character(), entity_uuid=character(),
                                      reviewed_date=character(), source=character())
  }
} else {
  reviewed_from_feedback <- tibble(raw_name=character(), entity_uuid=character(),
                                    reviewed_date=character(), source=character())
}

if (file.exists(REVIEWED_NAMES_PATH)) {
  reviewed_names_existing <- read.csv(REVIEWED_NAMES_PATH, stringsAsFactors = FALSE)
} else {
  reviewed_names_existing <- tibble(raw_name=character(), entity_uuid=character(),
                                     reviewed_date=character(), source=character())
}
reviewed_names <- bind_rows(
  reviewed_from_feedback,
  reviewed_names_existing %>% filter(!raw_name %in% reviewed_from_feedback$raw_name)
)

# Build confirmed mapping rows from 
names_from_b <- names_2026 %>%
  distinct(raw_name, canonical_name) %>%
  left_join(canonical_uuid_map %>% select(canonical_name, entity_uuid), by = "canonical_name") %>%
  mutate(confirmed = TRUE, added_date = as.character(Sys.Date()), source = "2026_names") %>%
  select(raw_name, entity_uuid, canonical_name, confirmed, added_date, source)

extra_from_c <- name_entity_mapping_existing %>%
  filter(!raw_name %in% names_from_b$raw_name)

name_entity_mapping <- bind_rows(names_from_b, extra_from_c)

entity_registry <- name_entity_mapping %>%
  filter(!is.na(canonical_name)) %>%
  distinct(canonical_name, entity_uuid) %>%
  arrange(canonical_name)

# fast lookup
nem_lookup <- name_entity_mapping %>%
  arrange(desc(confirmed)) %>%
  group_by(raw_name) %>%
  slice(1) %>%
  ungroup()

# Also match contributor names against canonical names in the map.
canonical_as_raw <- name_entity_mapping %>%
  filter(!is.na(canonical_name), confirmed == TRUE) %>%
  group_by(canonical_name) %>%
  slice(1) %>%
  ungroup() %>%
  transmute(
    raw_name       = canonical_name,
    entity_uuid    = entity_uuid,
    canonical_name = canonical_name,
    confirmed      = TRUE,
    added_date     = added_date,
    source         = source
  ) %>%
  filter(!raw_name %in% nem_lookup$raw_name)

nem_lookup <- bind_rows(nem_lookup, canonical_as_raw)
```


entity_registry: build from name_entity_mapping (combined existing + names 2026 input)
nem_lookup: raw_name, entity_uuid and canonical_name
name_entity_mapping: combined existing name entity map and any other entries in name_entity_mapping_existing (all of these should be in 2026)
reviewed_from_feedback: source for all reviewed names / decisions

```{r}
# data checks
reviewed_names %>% filter(grepl("UAW",raw_name)) # these now have different uuids
reviewed_names %>% filter(grepl("IBEW LOCAL",raw_name)) 
reviewed_names %>% filter(grepl("LONGSHORE",raw_name)) 
```

```{r}
# shouldnt be any after i transfered mappings over to main tracker
extra_from_c #%>% filter(raw_name == canonical_name)
```

The extra from c are all ones that didn't end up getting added because name = canonical name -> don't need to add

```{r}
# these need to be addressed in names 2026
name_entity_mapping %>% group_by(raw_name) %>% filter(n() > 1) 
```


```{r}
# Write updated files
#write.csv(entity_registry,  ENTITY_REGISTRY_PATH,     row.names = FALSE)
#write.csv(name_entity_mapping, NAME_ENTITY_MAPPING_PATH, row.names = FALSE)
#write.csv(reviewed_names,   REVIEWED_NAMES_PATH,      row.names = FALSE)

cat("Entity registry:", nrow(entity_registry), "canonical entities\n")
cat("Name mapping:", nrow(name_entity_mapping), "raw names",
    "(", nrow(names_from_b), "from 2026 Names CSV +", nrow(extra_from_c), "other)\n")
cat("Reviewed names registry:", nrow(reviewed_names), "settled names\n")
```



```{r}
nrow(entity_registry)
length(unique(map2$canonical_name))
nrow(map2)
```


```{r}
map2 %>% filter(!(canonical_name %in% entity_registry$canonical_name)) %>% pull(canonical_name)
names_2026 %>% filter(!(canonical_name %in% entity_registry$canonical_name)) %>% pull(canonical_name)
```


```{r}
head(entity_registry)
head(name_entity_mapping)
```


# Organizations

```{r}
make_name_variants <- function(df, col) {
  df %>% mutate(
    name_v1 = standardize_names(.data[[col]]) %>% fix_typos(),
    name_v2 = remove_pac_info(name_v1)        %>% str_squish(), # removing FEC ID, FED PAC ID, MAJOR DONOR ID
    #name_v3 = remove_unit_info(name_v2)        %>% str_squish(),
    name_v4 = str_remove_all(name_v1, "\\s*\\(.*?\\)") %>% str_squish(),
    name_v5 = replace_business_text(name_v4)   %>% str_squish()
    #name_v6 = str_remove_all(name_v5, "\\s*\\(.*?\\)") %>% str_squish()
  )
}

find_exact_variant_pairs <- function(pool, variant_col) {
  candidates <- pool %>%
    filter(!is.na(.data[[variant_col]]), .data[[variant_col]] != "") %>%
    group_by(across(all_of(c(variant_col, "has_pac_language")))) %>%
    filter(n() > 1) %>%
    ungroup()

  if (nrow(candidates) == 0)
    return(tibble(.x = integer(), .y = integer(), match_source = character()))

  candidates %>%
    group_by(across(all_of(c(variant_col, "has_pac_language")))) %>%
    group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>%
    ungroup() %>%
    select(.x, .y) %>%
    mutate(match_source = paste0("exact_", variant_col))
}
```


```{r}
orgs <- contributions %>%
  filter(entity_type == "organization") %>%
  mutate(cont_row_id = row_id) %>%          # preserve row_id before overwriting
  mutate(has_contributor_id = !is.na(Contributor.ID) & str_trim(as.character(Contributor.ID)) != "") %>%
  make_name_variants("Contributor.Name") %>%
  filter(!is.na(name_v1), name_v1 != "") %>%
  mutate(row_id = row_number())
```

## Step 1: ID-based aliases (same Contributor.ID, different names)


```{r}
id_based_aliases <- orgs %>%
  filter(has_contributor_id) %>%
  group_by(Contributor.ID) %>%
  filter(n_distinct(Contributor.Name) > 1) %>%
  ungroup()

id_auto_pairs <- id_based_aliases %>%
  group_by(Contributor.ID) %>%
  group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>%
  ungroup() %>%
  mutate(link_tier = "auto_link", match_source = "same_contributor_id")

cat("ID-based alias pairs:", nrow(id_auto_pairs), "\n")

id_based_aliases %>%
  group_by(Contributor.ID) %>%
  summarise(names = paste(unique(Contributor.Name), collapse = " | "), n = n(), .groups = "drop") %>%
  arrange(desc(n))
```


## Step 2: Exact variant cascade

```{r}
id_resolved_rows <- unique(c(id_auto_pairs$.x, id_auto_pairs$.y))
remaining_orgs   <- orgs %>% filter(!row_id %in% id_resolved_rows)

#variant_cols     <- c("name_v1", "name_v2", "name_v3", "name_v4", "name_v5", "name_v6")
variant_cols     <- c("name_v1", "name_v2", "name_v4", "name_v5") # don't do variants where we remove unit info or words attributed with PACs
exact_pairs_list <- list()

for (v in variant_cols) {
  new_pairs <- find_exact_variant_pairs(remaining_orgs, v)
  if (nrow(new_pairs) > 0) {
    exact_pairs_list[[v]] <- new_pairs
    remaining_orgs <- remaining_orgs %>%
      filter(!row_id %in% unique(c(new_pairs$.x, new_pairs$.y)))
  }
  cat("variant", v, "- new pairs:", nrow(new_pairs), "| remaining:", nrow(remaining_orgs), "\n")
}

exact_pairs <- bind_rows(exact_pairs_list) %>%
  left_join(orgs %>% select(row_id, id_x = Contributor.ID, has_id_x = has_contributor_id), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, id_y = Contributor.ID, has_id_y = has_contributor_id), by = c(".y" = "row_id")) %>%
  filter(!(coalesce(has_id_x, FALSE) & coalesce(has_id_y, FALSE) & coalesce(id_x != id_y, FALSE))) %>%
  mutate(link_tier = "auto_link") %>%
  select(.x, .y, link_tier, match_source)

cat("Exact cascade auto-links:", nrow(exact_pairs), "| remaining for fuzzy:", nrow(remaining_orgs), "\n")
```


### Step 3: internal fuzzy match


```{r}
remaining <- orgs %>%
  filter(
    !row_id %in% id_resolved_rows,
    !row_id %in% unique(c(exact_pairs$.x, exact_pairs$.y))
  )

ids <- remaining$row_id
pairs_idx <- combn(length(ids), 2)

org_pairs <- tibble(
  .x = ids[pairs_idx[1, ]],
  .y = ids[pairs_idx[2, ]]
) %>%
  left_join(orgs %>% select(row_id, name_x = name_v5, id_x = Contributor.ID,
                             has_id_x = has_contributor_id, pac_x = has_pac_language),
            by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = name_v5, id_y = Contributor.ID,
                             has_id_y = has_contributor_id, pac_y = has_pac_language),
            by = c(".y" = "row_id")) %>%
  filter(
    !(coalesce(has_id_x, FALSE) & coalesce(has_id_y, FALSE) & coalesce(id_x != id_y, FALSE)),
    pac_x == pac_y
  ) %>%
  mutate(
    name_jw     = stringsim(name_x, name_y, method = "jw", p = 0.1),
    name_cosine = bigram_cosine(name_x, name_y)
  ) %>%
  filter(name_jw >= 0.70, name_cosine >= 0.50) %>%
  select(-name_x, -name_y) %>%
  mutate(
    name_sim     = (name_jw + name_cosine) / 2,
    link_tier    = case_when(
      name_jw >= 0.999 ~ "auto_link",
      name_sim >= 0.90 ~ "auto_link", # reviewd this manually
      name_sim >= 0.85 ~ "review", # was 0.75, but no of those were really close..
      TRUE             ~ "no_link"
    ),
    match_source = "internal_fuzzy"
  )


```

```{r}
nrow(org_pairs)
```


## Step 5: External bridge (map1 / map2)

- update to only use the one map
- but check the other map at the end to make sure the name variant doesn't already exist

```{r}
# standardize external names to name_v5 format for matching
build_bridge_lookup <- function(mapping) {
  mapping %>%
    mutate(name_std = standardize_names(record_name) %>%
             fix_typos() %>%
             #remove_pac_info() %>%
             #remove_unit_info() %>%
             replace_business_text() %>%
             str_remove_all("\\s*\\(.*?\\)") %>%
             str_squish()) %>%
    select(name_std, canonical_name) %>%
    filter(!is.na(name_std), name_std != "", !is.na(canonical_name), canonical_name != "")
}

map1_lookup <- build_bridge_lookup(map1)
```


```{r}
find_bridge_pairs <- function(orgs_df, lookup, source_label) {
  matched <- orgs_df %>%
    inner_join(lookup, by = c("name_v5" = "name_std")) %>%
    select(row_id, name_v5, canonical_name) %>%
    distinct(name_v5, canonical_name, .keep_all = TRUE)  # one row per unique name × canonical (used to be row_id instead of name_v5 here)

  if (nrow(matched) < 2) return(tibble(.x = integer(), .y = integer(),
                                        match_source = character(), canonical_name = character()))

  matched %>%
    rename(.x = row_id, name_v5_x = name_v5) %>%
    inner_join(matched %>% rename(.y = row_id, name_v5_y = name_v5), by = "canonical_name") %>%
    filter(.x < .y, name_v5_x != name_v5_y) %>%
    group_by(.x, .y) %>%
    mutate(
      n_canonicals = n_distinct(canonical_name),
      needs_review = n_canonicals > 1
    ) %>%
    ungroup() %>%
    mutate(match_source = source_label) %>%
    select(.x, .y, match_source, canonical_name, n_canonicals, needs_review)
}

```

```{r}
bridge_map1 <- find_bridge_pairs(orgs, map1_lookup, "map1_bridge")
cat("Bridge pairs — map1:", nrow(bridge_map1), "\n")

bridge_map1 %>%
  mutate(.x = pmin(.x, .y), .y = pmax(.x, .y)) %>%
  distinct(.x, .y, canonical_name, .keep_all = TRUE) %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(name_x, name_y, canonical_name, n_canonicals) %>%
  arrange(name_x, name_y)

```
- get many-to-many warnings here

### combine bridge pairs

- think i can remove part where bridge pairs upgrade fuzzy match

```{r}
# bridge pairs from map1 only
bridge_pairs <- bridge_map1 %>%
  group_by(.x, .y) %>%
  summarise(
    match_source   = paste(sort(unique(match_source)), collapse = "+"),
    canonical_name = paste(sort(unique(canonical_name)), collapse = " / "),
    .groups = "drop"
  )

# canonical ID per name_v5: if ANY contribution row for that name has an ID, record it
# used detect conflicting-ID pairs regardless of which row_id find_bridge_pairs picked.
org_name_canonical_id <- orgs %>%
  filter(has_contributor_id) %>%
  group_by(name_v5) %>%
  summarise(canonical_id = names(sort(table(Contributor.ID), decreasing = TRUE))[1],
            .groups = "drop")

# score bridge pairs on internal similarity 
bridge_scored <- bridge_pairs %>%
  left_join(orgs %>% select(row_id, name_x = name_v5, pac_x = has_pac_language),
            by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = name_v5, pac_y = has_pac_language),
            by = c(".y" = "row_id")) %>%
  # join on name, not row_id in case a row was picked that didn't have a contributor.id
  left_join(org_name_canonical_id %>% rename(name_x = name_v5, id_x = canonical_id), by = "name_x") %>%
  left_join(org_name_canonical_id %>% rename(name_y = name_v5, id_y = canonical_id), by = "name_y") %>% # drop any pair where both orgs have IDs and those IDs differ — they are distinct entities
  filter(!(coalesce(!is.na(id_x), FALSE) & coalesce(!is.na(id_y), FALSE) & coalesce(id_x != id_y, FALSE))) %>%

  mutate(
    name_jw     = stringsim(name_x, name_y, method = "jw", p = 0.1),
    name_cosine = bigram_cosine(name_x, name_y),
    name_sim    = (name_jw + name_cosine) / 2
  ) %>%
  select(-name_x, -name_y, -id_x, -id_y)

# tag internal fuzzy paris thr bridge found— for match_source logging
internal_pair_keys <- org_pairs %>%
  filter(link_tier != "no_link") %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y)))

bridge_pair_keys <- bridge_scored %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  select(pair_key, bridge_source = match_source)

# upgrade internal pairs corroborated by a bridge
BRIDGE_UPGRADE_THRESHOLD <- 0.65

org_pairs <- org_pairs %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  left_join(bridge_pair_keys, by = "pair_key") %>%
  mutate(
    match_source = if_else(!is.na(bridge_source),
                           paste(match_source, bridge_source, sep = "+"),
                           match_source),
    link_tier = case_when(
      link_tier == "review" & !is.na(bridge_source) & name_sim >= BRIDGE_UPGRADE_THRESHOLD ~ "auto_link",
      TRUE ~ link_tier
    )
  ) %>%
  select(-pair_key, -bridge_source)

# bridge-only pairs
bridge_only <- bridge_scored %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  filter(!pair_key %in% internal_pair_keys$pair_key) %>%
  select(-pair_key) %>%
  mutate(
    link_tier = "review"  # bridge alone never auto-links
  )

cat("Bridge-corroborated internal pairs:", sum(!is.na(bridge_pair_keys$bridge_source[match(internal_pair_keys$pair_key, bridge_pair_keys$pair_key)])), "\n")
cat("Bridge-only new pairs for review:", nrow(bridge_only), "\n")
```


## Combine + cluster - definitive links for classification input

Combine all links, including those that need to be reviewed into a single entity
- not including fuzzy pairs and bridge pairs for now

```{r}
# pairs computed through different linkage methods
all_org_pairs <- bind_rows(
  id_auto_pairs %>% select(.x, .y, link_tier, match_source),
  exact_pairs,
  #org_pairs    %>% select(.x, .y, link_tier, match_source, name_jw, name_cosine, name_sim),
  #bridge_only  %>% select(.x, .y, link_tier, match_source, name_jw, name_cosine, name_sim, canonical_name)
)

# build one graph from all non-rejected pairs, treating review links as true
linkable_pairs <- all_org_pairs %>% filter(link_tier %in% c("auto_link", "review"))

if (nrow(linkable_pairs) > 0) {
  g_orgs       <- graph_from_data_frame(linkable_pairs %>% select(.x, .y), directed = FALSE)
  org_clusters <- components(g_orgs)$membership

  org_cluster_raw <- tibble(row_id = as.integer(names(org_clusters)), cluster_num = org_clusters) %>%
    left_join(orgs %>% select(row_id, Contributor.Name, Contributor.ID, name_v5), by = "row_id")

  cluster_uuid_map <- org_cluster_raw %>%
    group_by(cluster_num) %>%
    summarise(raw_names = list(unique(Contributor.Name)), .groups = "drop") %>%
    rowwise() %>%
    mutate(
      result      = list(lookup_cluster_uuid(raw_names, nem_lookup)),
      entity_id   = result$entity_id,
      is_conflict = result$is_conflict,
      is_new      = result$is_new
    ) %>%
    ungroup() %>%
    select(cluster_num, entity_id, is_conflict, is_new)

  org_cluster_map <- org_cluster_raw %>%
    left_join(cluster_uuid_map, by = "cluster_num") %>%
    select(row_id, entity_id, is_conflict, is_new)
} else {
  org_cluster_map <- tibble(row_id = integer(), entity_id = character(),
                             is_conflict = logical(), is_new = logical())
}

orgs_singleton_lookup <- orgs %>%
  filter(!row_id %in% org_cluster_map$row_id) %>%
  distinct(Contributor.Name) %>%
  mutate(raw_name_key = str_squish(toupper(Contributor.Name))) %>%
  left_join(nem_lookup %>% select(raw_name, entity_uuid), by = c("raw_name_key" = "raw_name")) %>%
  mutate(
    is_new    = is.na(entity_uuid),
    entity_id = coalesce(entity_uuid, UUIDgenerate(n = n()))
  ) %>%
  select(Contributor.Name, entity_id, is_new)

orgs_singletons <- orgs %>%
  filter(!row_id %in% org_cluster_map$row_id) %>%
  left_join(orgs_singleton_lookup, by = "Contributor.Name") %>%
  mutate(is_conflict = FALSE) %>%
  select(row_id, entity_id, is_new, is_conflict)

review_org_entity_ids <- all_org_pairs %>%
  filter(link_tier == "review") %>%
  left_join(bind_rows(org_cluster_map, orgs_singletons), by = c(".x" = "row_id"),
            relationship = "many-to-one") %>%
  filter(!is.na(entity_id)) %>%
  pull(entity_id) %>%
  unique()

orgs_resolved <- orgs %>%
  left_join(bind_rows(org_cluster_map, orgs_singletons), by = "row_id") %>%
  mutate(needs_review = entity_id %in% review_org_entity_ids)

```

## Org name variant review queue

For each entity with more than one distinct name, output the names that are not already
in map2 so they can be added to the name variant list. `entity_status` indicates whether
the entity already has some presence in map2 ("adding_to_existing") or is brand new ("new_entity").

```{r}
# Review-only grouping: includes fuzzy + bridge in addition to id/exact.
# Separate from all_org_pairs so classification aggregation is unaffected.
all_org_pairs_review <- bind_rows(
  id_auto_pairs %>% select(.x, .y, link_tier, match_source),
  exact_pairs   %>% select(.x, .y, link_tier, match_source),
  org_pairs     %>% filter(link_tier != "no_link") %>% select(.x, .y, link_tier, match_source),
  bridge_only   %>% select(.x, .y, link_tier, match_source)
)

if (nrow(all_org_pairs_review) > 0) {
  g_org_review     <- graph_from_data_frame(all_org_pairs_review %>% select(.x, .y), directed = FALSE)
  org_review_comps <- components(g_org_review)$membership
  org_review_group_map <- tibble(
    row_id          = as.integer(names(org_review_comps)),
    review_group_id = paste0("org_rg_", org_review_comps)
  )
} else {
  org_review_group_map <- tibble(row_id = integer(), review_group_id = character())
}

# Flagreview groups that contain names with conflicting non-NA Contributor.IDs.
org_cid <- orgs %>%
  select(row_id, cid = Contributor.ID) %>%
  mutate(cid = na_if(str_trim(as.character(cid)), ""))

conflicting_rg <- org_review_group_map %>%
  left_join(org_cid, by = "row_id") %>%
  filter(!is.na(cid)) %>%
  group_by(review_group_id) %>%
  filter(n_distinct(cid) > 1) %>%
  pull(review_group_id) %>%
  unique()

org_review_singletons <- orgs %>%
  filter(!row_id %in% org_review_group_map$row_id) %>%
  transmute(row_id, review_group_id = paste0("org_rg_solo_", row_id))

org_review_groups <- bind_rows(org_review_group_map, org_review_singletons)

# per-row_id lookup of which linkage methods connect it to others in its group
org_row_sources <- bind_rows(
  all_org_pairs_review %>% select(row_id = .x, match_source),
  all_org_pairs_review %>% select(row_id = .y, match_source)
) %>%
  group_by(row_id) %>%
  summarise(link_methods = paste(sort(unique(match_source)), collapse = " | "), .groups = "drop")

org_review_names <- orgs %>%
  left_join(org_review_groups, by = "row_id") %>%
  left_join(orgs_resolved %>% select(row_id, entity_id, is_conflict), by = "row_id") %>%
  left_join(org_row_sources, by = "row_id") %>%
  mutate(
    exported_name  = toupper(Contributor.Name),
    raw_name_key   = str_squish(exported_name),
    contributor_id = na_if(str_trim(as.character(Contributor.ID)), "")
  ) %>%
  left_join(nem_lookup %>% select(raw_name, confirmed, canonical_name),
            by = c("raw_name_key" = "raw_name")) %>%
  # a name is confirmed if it's in name_entity_mapping (variant confirmed) OR reviewed_names
  # (any review decision made — covers canonical names and singletons that don't need normalization)
  mutate(confirmed = coalesce(confirmed, FALSE) | raw_name_key %in% reviewed_names$raw_name) %>%
  group_by(review_group_id) %>%
  filter(n_distinct(exported_name) > 1) %>%
  ungroup() %>%
  group_by(review_group_id, exported_name, contributor_id) %>%
  summarise(
    entity_id       = first(entity_id),
    canonical_name  = first(canonical_name),
    confirmed       = first(confirmed),
    is_conflict     = coalesce(first(is_conflict), FALSE),
    link_methods    = paste(sort(unique(link_methods[!is.na(link_methods)])), collapse = " | "),
    n_contributions = n(),
    total_amount    = sum(Amount, na.rm = TRUE),
    city  = paste(sort(unique(Contributor.City[ !is.na(Contributor.City)  & Contributor.City  != ""])), collapse = " | "),
    state = paste(sort(unique(Contributor.State[!is.na(Contributor.State) & Contributor.State != ""])), collapse = " | "),
    zip   = paste(sort(unique(Contributor.Zip.Code[!is.na(Contributor.Zip.Code) & Contributor.Zip.Code != ""])), collapse = " | "),
    .groups = "drop"
  )

org_review_group_status <- org_review_names %>%
  group_by(review_group_id) %>%
  summarise(
    # all_confirmed takes priority: both same-entity confirmations (same UUID) and
    # not-same-entity confirmations (different UUIDs) suppress a group from review.
    entity_status = case_when(
      all(confirmed)   ~ "all_confirmed",
      any(is_conflict) ~ "merge_conflict",
      any(confirmed)   ~ "adding_to_existing",
      TRUE             ~ "new_entity"
    ),
    conflicting_contributor_ids = any(review_group_id %in% conflicting_rg),
    .groups = "drop"
  )


# Recipients: Recipient.Committee.ID (PACs), Ballot.Measure (props), Recipient.Name (candidates).
# A group qualifies for review only if it crossed $5k to at least one singular recipient.
org_review_group_threshold <- orgs %>%
  left_join(org_review_groups %>% select(row_id, review_group_id), by = "row_id") %>%
  mutate(
    recipient_key = case_when(
      !is.na(Recipient.Committee.ID) & as.character(Recipient.Committee.ID) != "" ~ as.character(Recipient.Committee.ID),
      !is.na(Ballot.Measure) & Ballot.Measure != ""                               ~ Ballot.Measure,
      TRUE                                                                         ~ Recipient.Name
    )
  ) %>%
  group_by(review_group_id, recipient_key) %>%
  summarise(recipient_total = sum(Amount, na.rm = TRUE), .groups = "drop") %>%
  group_by(review_group_id) %>%
  summarise(max_single_recipient = max(recipient_total, na.rm = TRUE), .groups = "drop")

org_name_variant_queue <- org_review_names %>%
  left_join(org_review_group_status, by = "review_group_id") %>%
  left_join(org_review_group_threshold, by = "review_group_id") %>%
  filter(
    entity_status != "all_confirmed",
    max_single_recipient >= 5000
  ) %>%
  select(entity_name = exported_name, review_group_id, entity_id, canonical_name, entity_status,
         conflicting_contributor_ids, confirmed, is_conflict, link_methods, Contributor.ID = contributor_id,
         n_contributions, total_amount, max_single_recipient, city, state, zip) %>%
  arrange(desc(conflicting_contributor_ids), review_group_id)

```

```{r}
nrow(org_name_variant_queue)
length(unique(org_name_variant_queue$entity_name))
length(unique(org_name_variant_queue$entity_id))
```


```{r}
write.csv(org_name_variant_queue, "08_outputs/org_name_variant_review_queue_08-19-26.csv", row.names = FALSE)
cat("Org name variant review queue:", nrow(org_name_variant_queue), "names across",
    n_distinct(org_name_variant_queue$review_group_id), "groups\n")
```



# INDIVIDUALS:

```{r}
# ── Constants ─────────────────────────────────────────────────────────────────
UNINFORMATIVE_VALS        <- c("UNKNOWN", "RETIRED", "NONE", "SELF EMPLOYED") 
FIRST_NAME_MIN            <- 0.85
MATCH_THRESHOLD           <- 0.90
STRICT_THRESHOLD          <- 0.94  # both employer & occupation uninformative
COMMON_NAME_THRESHOLD     <- 0.93  # common name, employer informative
COMMON_NAME_STRICT        <- 0.97  # common name + both uninformative
UNCERTAIN_LOW             <- 0.75
EMPLOYER_CONFLICT_SIM     <- 0.70  # employer sim below this = conflict
#COMMON_NAME_MIN_COUNT     <- 3     # names appearing >= N times treated as common
RARE_NAME_UNCERTAIN_LOW <- 0.65  # tune based on what you see in the review queue
COMMON_LAST_MIN           <- 3
COMMON_FIRST_MIN          <- 3
RARE_NAME_UNCERTAIN_LOW   <- 0.65


```

```{r}
# compute name frequency

# parse distinct name strings from full dataset
full_names_parsed <- contributions_full %>%
  mutate(
    standardized_name = standardize_names(Contributor.Name),
    entity_type = if_else(sapply(standardized_name, is_individual), "individual", "organization")
  ) %>%
  filter(entity_type == "individual") %>%
  distinct(standardized_name) %>%
  bind_cols(parse_names_df(.$standardized_name)) %>%
  filter(!is.na(last), !is.na(first))

last_name_freq  <- full_names_parsed %>% count(last,  name = "last_name_freq")
first_name_freq <- full_names_parsed %>% count(first, name = "first_name_freq")


# parse names 
parsed_cols <- parse_names_df(
  contributions %>% filter(entity_type == "individual") %>% pull(standardized_name) # use standardized name here because it has the commas still in it
)
```

```{r}
#write_csv(full_names_parsed, "08_outputs/individuals_name_parsed_all_contributors.csv")
```



```{r}
individuals <- contributions %>%
  filter(entity_type == "individual") %>%
  bind_cols(parsed_cols %>% rename(last_name = last, first_name = first)) %>%
  mutate(
    original_row_id = row_id,       # preserve for final join back to x
    row_id          = row_number(), # sequential for pair_blocking compatibility
    first_initial   = substr(first_name, 1, 1),
    last_meta       = safe_metaphone(last_name)
  ) #%>%
  #filter(!is.na(last_meta), last_meta != "")

which(is.na(individuals$last_meta))

```

### group by name

- Group individuals if they have the same name

```{r}
# assign entity id — one UUID per standardized_name, looked up from registry if known
# adds entity_id to existing individuals (which already has last_meta, original_row_id, etc.)
indiv_name_lookup <- individuals %>%
  distinct(Contributor.Name, standardized_name) %>%
  mutate(raw_name_key = str_squish(toupper(Contributor.Name))) %>%
  left_join(nem_lookup %>% select(raw_name, entity_uuid), by = c("raw_name_key" = "raw_name"))

indiv_entity_key <- indiv_name_lookup %>%
  group_by(standardized_name) %>%
  summarise(entity_id = coalesce(first(na.omit(entity_uuid)), UUIDgenerate()), .groups = "drop")

individuals <- individuals %>%
  left_join(indiv_entity_key, by = "standardized_name")

individuals <- individuals %>%
  group_by(entity_id, race_prop) %>%
  mutate(
    race_total_amount = sum(Amount, na.rm = TRUE),
    crosses_threshold = race_total_amount >= 5000
  ) %>%
  ungroup()


```

```{r}
contributions %>% filter(toupper(Contributor.Name) != standardized_name & entity_type == "individual") %>% select(Contributor.Name,standardized_name)
```

- periods do make a difference in a lot of the names
Orgs that are individuals: 
- Rolls, Anderson & Rolls: labeled as an individual not an org
- DRS HOLDEN, KNAPP AND LANGSFORD AKA FAMILY OPTOMETRY

### record linkage algorithm

```{r}
# scoring function
compare_and_score_indiv <- function(pairs_obj, df) {
  pairs_obj %>%
    compare_pairs(
      on = c("last_name", "first_name", "zip_code_processed",
             "processed_employer_name", "processed_occupation", "standardized_city"),
      default_comparator = cmp_jarowinkler(),
      comparators = list(
        processed_employer_name = function(a, b) stringsim(a, b, method = "cosine", q = 2),
        zip_code_processed      = cmp_identical(),
        processed_occupation    = cmp_identical()
      )
    ) %>%
    as_tibble() %>%
    # join raw employer/occupation values
    left_join(df %>% select(row_id, emp_a = processed_employer_name, occ_a = processed_occupation,
                             city_a = standardized_city),
              by = c(".x" = "row_id")) %>%
    left_join(df %>% select(row_id, emp_b = processed_employer_name, occ_b = processed_occupation,
                             city_b = standardized_city),
              by = c(".y" = "row_id")) %>%
    mutate(
      emp_a_uninf = emp_a %in% UNINFORMATIVE_VALS | is.na(emp_a) | emp_a == "",
      emp_b_uninf = emp_b %in% UNINFORMATIVE_VALS | is.na(emp_b) | emp_b == "",
      occ_a_uninf = occ_a %in% UNINFORMATIVE_VALS | is.na(occ_a) | occ_a == "",
      occ_b_uninf = occ_b %in% UNINFORMATIVE_VALS | is.na(occ_b) | occ_b == "",
      w_emp  = if_else(emp_a_uninf & emp_b_uninf, 0,    0.25),
      w_occ  = if_else(occ_a_uninf & occ_b_uninf, 0,    0.10),
      w_tot  = 0.30 + 0.25 + w_emp + w_occ + 0.07 + 0.03,
      emp_sim = if_else(emp_a_uninf | emp_b_uninf, 0, coalesce(processed_employer_name, 0)),
      occ_sim = if_else(occ_a_uninf | occ_b_uninf, 0, coalesce(processed_occupation, 0)),
      score  = (0.30 * coalesce(last_name, 0) +
                0.25 * coalesce(first_name, 0) +
                w_emp * emp_sim +
                w_occ * occ_sim +
                0.07 * coalesce(zip_code_processed, 0) +
                0.03 * coalesce(standardized_city, 0)) / w_tot,
      both_uninformative = (emp_a_uninf & emp_b_uninf) & (occ_a_uninf & occ_b_uninf),
    
      employer_substring = !emp_a_uninf & !emp_b_uninf & mapply(
        function(a, b) {
          if (is.na(a) || is.na(b) || a == "" || b == "") return(FALSE)
          a <- toupper(a); b <- toupper(b)
          grepl(b, a, fixed = TRUE) || grepl(a, b, fixed = TRUE)
        },
        emp_a, emp_b,
        USE.NAMES = FALSE
      ),
    
      employer_conflict  = !emp_a_uninf & !emp_b_uninf &
                           !employer_substring &
                           coalesce(processed_employer_name, 0) < EMPLOYER_CONFLICT_SIM,
    
      # city conflict: both cities are known and different (JW < 0.85)
      city_conflict = (!is.na(city_a) & city_a != "") &
                      (!is.na(city_b) & city_b != "") &
                      coalesce(standardized_city, 1) < 0.85
    ) %>%
    select(-occ_a, -occ_b,
           -emp_a_uninf, -emp_b_uninf, -occ_a_uninf, -occ_b_uninf,
           -w_emp, -w_occ, -w_tot, -emp_sim, -occ_sim,
           -city_a, -city_b)
}
```

```{r}
# blocking + scoring
pairs_zip  <- pair_blocking(individuals, on = c("last_meta", "zip_code_processed"), deduplication = TRUE)
pairs_init <- pair_blocking(individuals, on = c("last_meta", "first_initial"),       deduplication = TRUE)

indiv_pairs <- bind_rows(
  compare_and_score_indiv(pairs_zip,  individuals),
  compare_and_score_indiv(pairs_init, individuals)
) %>%
  group_by(.x, .y) %>%
  slice_max(score, n = 1, with_ties = FALSE) %>%
  ungroup()


# name frequency adjustment
indiv_pairs <- indiv_pairs %>%
  left_join(individuals %>% select(row_id, last_name_str = last_name, first_name_str = first_name),
            by = c(".x" = "row_id")) %>%
  left_join(last_name_freq,  by = c("last_name_str"  = "last")) %>%
  left_join(first_name_freq, by = c("first_name_str" = "first")) %>%
  mutate(
    common_last  = coalesce(last_name_freq,  1L) >= COMMON_LAST_MIN,
    common_first = coalesce(first_name_freq, 1L) >= COMMON_FIRST_MIN,
    common_name  = common_last & common_first,
    rare_name    = !common_last & !common_first,
    match_threshold = case_when(
      common_name & both_uninformative ~ COMMON_NAME_STRICT,
      common_name                      ~ COMMON_NAME_THRESHOLD,
      both_uninformative               ~ STRICT_THRESHOLD,
      TRUE                             ~ MATCH_THRESHOLD
    ),
    # unusual names flagged for review at a lower score bar
    review_threshold = if_else(rare_name, RARE_NAME_UNCERTAIN_LOW, UNCERTAIN_LOW)
  ) %>%
  select(-last_name_str, -first_name_str, -last_name_freq, -first_name_freq)

# link tiers
indiv_pairs <- indiv_pairs %>%
  mutate(
    link_tier = case_when(
      coalesce(first_name, 0) < FIRST_NAME_MIN ~ "no_link",
      common_name & both_uninformative ~ "no_link",
      city_conflict & both_uninformative ~ "no_link",
      common_name & city_conflict ~ "no_link",       # city disagreement is enough to reject common names
      common_name & employer_conflict ~ "no_link",   # clearly different employers also enough for common names
      score < review_threshold ~ "no_link",
      score >= match_threshold ~ "auto_link",
      TRUE                                           ~ "review"
    ),
    match_source = "internal_fuzzy"
  )

cat("Individual pairs — auto_link:", sum(indiv_pairs$link_tier == "auto_link"),
    "| review:", sum(indiv_pairs$link_tier == "review"),
    "| employer conflicts in review:", sum(indiv_pairs$employer_conflict & indiv_pairs$link_tier == "review"), "\n")

```


### Cluster individuals

```{r}
# treat all links as true, review after classifications
linkable_indiv_pairs <- indiv_pairs %>% filter(link_tier %in% c("auto_link", "review"))

if (nrow(linkable_indiv_pairs) > 0) {
  g_indiv        <- graph_from_data_frame(linkable_indiv_pairs %>% select(.x, .y), directed = FALSE)
  indiv_clusters <- components(g_indiv)$membership

  indiv_cluster_raw <- tibble(row_id = as.integer(names(indiv_clusters)), cluster_num = indiv_clusters) %>%
    left_join(individuals %>% select(row_id, Contributor.Name, Contributor.ID, standardized_name,
                                     original_row_id), by = "row_id")

  indiv_cluster_uuid_map <- indiv_cluster_raw %>%
    group_by(cluster_num) %>%
    summarise(raw_names = list(unique(Contributor.Name)), .groups = "drop") %>%
    rowwise() %>%
    mutate(
      result      = list(lookup_cluster_uuid(raw_names, nem_lookup)),
      entity_id   = result$entity_id,
      is_conflict = result$is_conflict,
      is_new      = result$is_new
    ) %>%
    ungroup() %>%
    select(cluster_num, entity_id, is_conflict, is_new)

  indiv_cluster_map <- indiv_cluster_raw %>%
    left_join(indiv_cluster_uuid_map, by = "cluster_num") %>%
    select(row_id = original_row_id, entity_id, is_conflict, is_new)
} else {
  indiv_cluster_map <- tibble(row_id = integer(), entity_id = character(),
                               is_conflict = logical(), is_new = logical())
}

indiv_singleton_lookup <- individuals %>%
  filter(!original_row_id %in% indiv_cluster_map$row_id) %>%
  distinct(Contributor.Name) %>%
  mutate(raw_name_key = str_squish(toupper(Contributor.Name))) %>%
  left_join(nem_lookup %>% select(raw_name, entity_uuid), by = c("raw_name_key" = "raw_name")) %>%
  mutate(
    is_new    = is.na(entity_uuid),
    entity_id = coalesce(entity_uuid, UUIDgenerate(n = n()))
  ) %>%
  select(Contributor.Name, entity_id, is_new)

indiv_singletons <- individuals %>%
  filter(!original_row_id %in% indiv_cluster_map$row_id) %>%
  select(-entity_id) %>%
  left_join(indiv_singleton_lookup, by = "Contributor.Name") %>%
  mutate(is_conflict = FALSE) %>%
  select(row_id = original_row_id, entity_id, is_new, is_conflict)

# within-individuals row_id to entity_id map (needed for needs_review flag)
# indiv_pairs uses within-individuals row_id; cluster_map was renamed to original_row_id
indiv_internal_to_entity <- individuals %>%
  select(row_id, original_row_id) %>%
  left_join(bind_rows(indiv_cluster_map, indiv_singletons), by = c("original_row_id" = "row_id"))

review_indiv_entity_ids <- indiv_pairs %>%
  filter(link_tier == "review") %>%
  left_join(indiv_internal_to_entity %>% select(row_id, entity_id), by = c(".x" = "row_id"),
            relationship = "many-to-one") %>%
  filter(!is.na(entity_id)) %>%
  pull(entity_id) %>%
  unique()

# use longer employer name for pairs when one employer is the exact substring of another
employer_normalization <- indiv_pairs %>%
  filter(employer_substring) %>%
  mutate(longer = if_else(nchar(coalesce(emp_a, "")) >= nchar(coalesce(emp_b, "")), emp_a, emp_b)) %>%
  select(.x, .y, longer) %>%
  pivot_longer(c(.x, .y), names_to = "side", values_to = "row_id") %>%
  group_by(row_id) %>%
  slice_max(nchar(coalesce(longer, "")), n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(row_id, normalized_employer = longer)


individuals_resolved <- individuals %>%
  select(-entity_id) %>%   # drop initial name-based entity_id; cluster-based wins
  left_join(bind_rows(indiv_cluster_map, indiv_singletons), by = c("original_row_id" = "row_id")) %>%
  mutate(needs_review = entity_id %in% review_indiv_entity_ids)

individuals_resolved <- individuals_resolved %>%
  left_join(employer_normalization, by = "row_id") %>%
  mutate(classification_employer = coalesce(normalized_employer, processed_employer_name))


```

## Individual name variant review queue

For each pair surfaced by the record linkage algorithm where the two names differ, output
one row per name.

```{r}
# Build connected components from all suspected links (A↔B, B↔C → all three in one group)
linkable_indiv_review <- indiv_pairs %>% filter(link_tier %in% c("auto_link", "review"))

if (nrow(linkable_indiv_review) > 0) {
  g_review     <- graph_from_data_frame(linkable_indiv_review %>% select(.x, .y), directed = FALSE)
  review_comps <- components(g_review)$membership
  review_group_map <- tibble(
    row_id          = as.integer(names(review_comps)),
    entity_group_id = paste0("igroup_", review_comps)
  )
} else {
  review_group_map <- tibble(row_id = integer(), entity_group_id = character())
}

# Attach fields including recipient/amount for threshold filtering
indiv_review_data <- review_group_map %>%
  left_join(
    individuals %>% select(row_id, Contributor.Name, Contributor.Employer,
                           Contributor.Occupation, Contributor.City,
                           Contributor.State, Contributor.Zip.Code,
                           Amount, Recipient.Name, Recipient.Committee.ID, Ballot.Measure),
    by = "row_id"
  ) %>%
  left_join(indiv_internal_to_entity %>% select(row_id, entity_id), by = "row_id") %>%
  mutate(
    exported_name = toupper(Contributor.Name),
    raw_name_key  = str_squish(exported_name),
    recipient_key = case_when(
      !is.na(Recipient.Committee.ID) & as.character(Recipient.Committee.ID) != "" ~ as.character(Recipient.Committee.ID),
      !is.na(Ballot.Measure) & Ballot.Measure != ""                               ~ Ballot.Measure,
      TRUE                                                                         ~ Recipient.Name
    )
  ) %>%
  left_join(nem_lookup %>% select(raw_name, confirmed, canonical_name),
            by = c("raw_name_key" = "raw_name")) %>%
  mutate(confirmed = coalesce(confirmed, FALSE)) %>%
  group_by(entity_group_id) %>%
  filter(n_distinct(exported_name) > 1) %>%
  ungroup()

# $5k threshold: keep groups where any single recipient received >= $5k
indiv_review_group_threshold <- indiv_review_data %>%
  group_by(entity_group_id, recipient_key) %>%
  summarise(recipient_total = sum(Amount, na.rm = TRUE), .groups = "drop") %>%
  group_by(entity_group_id) %>%
  summarise(max_single_recipient = max(recipient_total, na.rm = TRUE), .groups = "drop")

indiv_group_status <- indiv_review_data %>%
  group_by(entity_group_id) %>%
  summarise(
    entity_status = case_when(
      all(confirmed)  ~ "all_confirmed",
      any(confirmed)  ~ "adding_to_existing",
      TRUE            ~ "new_entity"
    ),
    .groups = "drop"
  )

indiv_name_variant_queue <- indiv_review_data %>%
  group_by(entity_group_id, exported_name, Contributor.Employer, Contributor.Occupation) %>%
  summarise(
    entity_id       = first(entity_id),
    canonical_name  = first(canonical_name),
    confirmed       = first(confirmed),
    n_contributions = n(),
    total_amount    = sum(Amount, na.rm = TRUE),
    city      = paste(sort(unique(Contributor.City[!is.na(Contributor.City)   & Contributor.City   != ""])), collapse = " | "),
    state     = paste(sort(unique(Contributor.State[!is.na(Contributor.State)  & Contributor.State  != ""])), collapse = " | "),
    zip       = paste(sort(unique(Contributor.Zip.Code[!is.na(Contributor.Zip.Code) & Contributor.Zip.Code != ""])), collapse = " | "),
    recipient = paste(sort(unique(recipient_key[!is.na(recipient_key) & recipient_key != ""])), collapse = " | "),
    .groups = "drop"
  ) %>%
  left_join(indiv_group_status, by = "entity_group_id") %>%
  left_join(indiv_review_group_threshold, by = "entity_group_id") %>%
  filter(
    entity_status != "all_confirmed",
    max_single_recipient >= 5000
  ) %>%
  select(entity_group_id, entity_name = exported_name, entity_id, canonical_name,
         entity_status, confirmed, Contributor.Employer, Contributor.Occupation,
         n_contributions, total_amount, max_single_recipient, city, state, zip, recipient) %>%
  arrange(entity_group_id)
```

```{r}
indiv_name_variant_queue
```


```{r}
write.csv(indiv_name_variant_queue, "08_outputs/indiv_name_variant_review_queue.csv", row.names = FALSE)
cat("Individual name variant review queue:", n_distinct(indiv_name_variant_queue$entity_group_id),
    "groups across", n_distinct(indiv_name_variant_queue$entity_id), "entities\n")
```



# Provisional registry write-back

Append new names from this run to `name_entity_mapping.csv` with `confirmed = FALSE`.
On the next run they will be loaded and their UUIDs reused, keeping entity_id stable
Confirmed entries are never overwritten.

```{r}
all_resolved <- bind_rows(
  orgs_resolved %>%
    select(Contributor.Name, entity_id, is_new) %>%
    mutate(is_new = coalesce(is_new, FALSE)),
  individuals_resolved %>%
    select(Contributor.Name, entity_id, is_new) %>%
    mutate(is_new = coalesce(is_new, FALSE))
)

new_provisional <- all_resolved %>%
  filter(is_new) %>%
  distinct(Contributor.Name, entity_id) %>%
  mutate(
    raw_name   = str_squish(toupper(Contributor.Name)),
    entity_uuid = entity_id,
    canonical_name = NA_character_,
    confirmed  = FALSE,
    added_date = as.character(Sys.Date()),
    source     = "run_provisional"
  ) %>%
  filter(!raw_name %in% name_entity_mapping$raw_name) %>%   # don't overwrite existing entries
  select(raw_name, entity_uuid, canonical_name, confirmed, added_date, source)

if (nrow(new_provisional) > 0) {
  updated_mapping <- bind_rows(name_entity_mapping, new_provisional)
  write.csv(updated_mapping, NAME_ENTITY_MAPPING_PATH, row.names = FALSE)
  cat("Appended", nrow(new_provisional), "provisional name mappings to registry\n")
} else {
  cat("No new names to append to registry\n")
}
```

# Classification input

```{r}
orgs_resolved <- orgs_resolved %>%
  group_by(entity_id, race_prop) %>%
  mutate(
    race_total_amount = sum(Amount, na.rm = TRUE),
    crosses_threshold = race_total_amount >= 5000
  ) %>%
  ungroup()

individuals_classification_input <- individuals %>%
  filter(crosses_threshold)

nrow(individuals_classification_input)
n_distinct(individuals_classification_input$entity_id)

```


```{r}
classification_input <- bind_rows(
  individuals    %>% filter(crosses_threshold),
  orgs_resolved  %>% filter(crosses_threshold)
)

nrow(classification_input)
n_distinct(classification_input$entity_id)

```



### match back uuid from previous classifications

```{r}
pad_zip <- function(z) {
  z <- trimws(as.character(z))
  needs_pad <- !is.na(z) & z != "" & str_detect(z, "^[0-9]+$") & nchar(z) < 5
  z[needs_pad] <- paste0(strrep("0", 5 - nchar(z[needs_pad])), z[needs_pad])
  substr(z, 1, 5)
}

normalize_missing <- function(x) {
  x <- trimws(as.character(x))
  if_else(x %in% c("", "N/A", "NA", "None","n/a") | is.na(x), "N/A", x)
}
```

```{r}
categorized <- categorized %>%
  mutate(
    race_prop = "Governor",
    legacy_match_id = make_row_hash(
      Contributor.Name, Contributor.ID, as.numeric(Amount), race_prop,
      Contributor.City, normalize_missing(Contributor.Employer),
      normalize_missing(Contributor.Occupation), pad_zip(Contributor.Zip.Code),
      Start.Date
    )
  ) %>%
  group_by(legacy_match_id) %>%
  mutate(dup_rank = row_number()) %>%
  ungroup()

classification_input <- classification_input %>%
  mutate(
    legacy_match_id = make_row_hash(
      Contributor.Name, Contributor.ID, as.numeric(Amount), race_prop,
      Contributor.City, normalize_missing(Contributor.Employer),
      normalize_missing(Contributor.Occupation), pad_zip(Contributor.Zip.Code),
      Start.Date
    )
  ) %>%
  group_by(legacy_match_id) %>%
  mutate(dup_rank = row_number()) %>%
  ungroup() %>%
  left_join(
    categorized %>% select(legacy_match_id, dup_rank, uuid),
    by = c("legacy_match_id", "dup_rank")
  ) %>%
  select(-dup_rank)

```



```{r}
# data checks
nrow(classification_input)
n_distinct(classification_input$uuid) == nrow(classification_input)

classification_input %>% count(uuid) %>% filter(n > 1) # should only be NAs

# make sure match id is unique
categorized %>% count(legacy_match_id, dup_rank) %>% filter(n > 1)

nrow(categorized)
length(which(!is.na(classification_input$uuid)))
```

- Middleteon, Geraldine: ended up not aggregating to $5k because of negative contributions that lowered her aggregated spending

Seems like some don't match - likely because amendments were filed or because 460 trumps 497. 
Re-classify those that didn't match back.

```{r}
categorized %>% filter(!(uuid %in% classification_input$uuid))
categorized %>% filter(Contributor.Name == "Thomas, Drake")
classification_input %>% filter(Contributor.Name == "Thomas, Drake")
contributions_full %>% filter(Contributor.Name == "Thomas, Drake")
```

```{r}
ambiguous <- categorized %>%
  group_by(legacy_match_id) %>%
  filter(n() > 1) %>%
  ungroup()

nrow(ambiguous)                     
n_distinct(ambiguous$legacy_match_id)   

```


```{r}
# add remaining uuids
namespace <- "887a6583-910c-47b0-b373-d5b2863e8c8a"
classification_input[which(is.na(classification_input$uuid)),]$uuid <-
  uuid::UUIDgenerate(n = nrow(classification_input[which(is.na(classification_input$uuid)),]))
```

Okay to reassign these uuids? stuff downstream impacted by reassigned uuids? esp matching contribs to LLM search results

# Output PAC list

(1) Look for anything with a missing contributor.id
(2) output list of all rows with a contributor.id
  
### look for any names that match a recipient committee - deal with major donors/IE later on


```{r}
# data to map contributor ids
id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv")
id_map <- unique(id_map_full[,c("XREF_FILER_ID","FILER_ID","FILER_TYPE")]) # get unique xref id / filer ids combos

# convert all ids to character strings
id_map$FILER_ID         <- as.character(id_map$FILER_ID)   
id_map$XREF_FILER_ID <- as.character(id_map$XREF_FILER_ID)
classification_input$Contributor.ID <- as.character(classification_input$Contributor.ID)

# map filer id to contributor id (different ids may be used internally vs externally)
classification_input<- left_join(classification_input, id_map, by = c("Contributor.ID" = "XREF_FILER_ID"))

# check that each has a corresponding filer id
nrow(classification_input[which(is.na(classification_input$FILER_ID) & !is.na(classification_input$Contributor.ID)),])
```





```{r}
# Build name/ID lookup from full dataset
full_id_lookup <- id_map_full %>%
  filter(!is.na(FILER_ID), FILER_ID != "",FILER_TYPE == "RECIPIENT COMMITTEE") %>% # , (FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" | FILER_TYPE == "RECIPIENT COMMITTEE")
  mutate(name_v5 = standardize_names(NAML) %>% 
           str_remove_all("\\s*\\(.*?\\)") %>%
          str_squish()) %>%
  distinct(name_v5, FILER_ID) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(FILER_ID)) %>%
  ungroup()


```

Different filer types should have different ids

```{r}
temp <- classification_input
```


```{r}
# Propagate IDs to orgs that are missing them
full_id_propagation <- classification_input %>%
  filter(is.na(Contributor.ID)) %>%
  select(row_id, Contributor.Name, name_v5) %>%
  inner_join(full_id_lookup, by = "name_v5") %>%
  mutate(needs_review = n_ids > 1,
         FILER_ID = as.character(FILER_ID)) %>%
  select(row_id, Contributor.Name, propagated_id = FILER_ID, n_ids, needs_review)
full_id_propagation
```

```{r}
classification_input$FILER_TYPE
```

```{r}
classification_input$Contributor.ID <- as.character(classification_input$Contributor.ID)
full_id_propagation$propagated_id <- as.character(full_id_propagation$propagated_id)
```



```{r}
# apply ids back; Contributor.ID stays raw — effective_Contributor.ID includes propagated ids
classification_input <- classification_input %>%
  left_join(
    full_id_propagation %>% distinct(row_id, propagated_id),
    by = "row_id"
  ) %>%
  mutate(
    effective_Contributor.ID = coalesce(Contributor.ID, propagated_id),
    has_contributor_id       = !is.na(effective_Contributor.ID) & str_trim(as.character(effective_Contributor.ID)) != ""
  ) %>%
  select(-propagated_id)
```


```{r}
classification_input <- classification_input %>% 
  mutate(FILER_TYPE = ifelse(is.na(Contributor.ID) & !is.na(effective_Contributor.ID), "RECIPIENT COMMITTEE", FILER_TYPE))
```


```{r}
nrow(classification_input[which(!is.na(classification_input$effective_Contributor.ID)),])
nrow(classification_input[which(is.na(classification_input$effective_Contributor.ID)),])
```


# WRITE OUTPUTS

- Update this so major donors don't go to pac processing unless they have pac language

Make sure to send major donors to be classified directly.

```{r}
classification_input %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(Contributor.Name, has_pac_language,effective_Contributor.ID)
```

- PAC input: 
  - all recipient pacs 
  - major donors with PAC language -> will go to review queue
  
```{r}
pacs <- classification_input %>% filter((!is.na(effective_Contributor.ID) & FILER_TYPE != "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" ) | (FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" & has_pac_language == TRUE))

direct <- classification_input %>% filter(is.na(effective_Contributor.ID) | (FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" & has_pac_language == FALSE))

nrow(classification_input[which(!is.na(classification_input$effective_Contributor.ID)),])
nrow(pacs)
nrow(direct)
nrow(classification_input) == sum(nrow(pacs) + nrow(direct))
```
```{r}
nrow(pacs)
nrow(direct)
```


```{r}
write.csv(direct,"08_outputs/classification_input.csv",row.names = FALSE)
write.csv(pacs,"08_outputs/pac_input.csv",row.names = FALSE)
```

```{r}
direct <- read.csv("08_outputs/classification_input.csv")
pacs <- read.csv("08_outputs/pac_input.csv")
```






