---
title: "R Notebook"
output: html_notebook
---

# Summary: 

This notebook performs entity resolution over contributors and attempts to identify 
name variants of orgs and individuals as well as identify cases where different individuals
share the same name. 

The entity resolution process is designed to be compatible with what is already being done 
for the campaign finance data on the Voter Guide. The approach for the voter guide 
groups contributors by name to determine how much that contributor gave in aggregate. Known name
variants of individuals and orgs are mapped to a `Normalized` name in a Google Sheet.

The notebook uses name variant matching, fuzzy matching and a custom record linkage algorithm
to try to identify name variants. Name variants are manually reviewed and added to the map that
is also used for the Voter Guide. 

For this project, we are not classifying entities that gave less than $5k to a race/prop/pac,
so entity resolution / manual name variant review are only performed for entities that have 
the potential to cross this threshold. Entity resolution also focuses on within-race/pac 
contributions; not all name variants across all recipient committees are tracked since we 
aggregate within each prop/candidate race. 

Processing steps: 

(1) Download the latest 2026 name entity map (https://docs.google.com/spreadsheets/d/1THwQtvw5s9ZO9n7QHncekl4kRMc1Nf3MJHViSYALDps/edit?gid=897118445#gid=897118445) 
    and feedback from manual reviews of name variants (should already be in the directory).
(2) Run name variant linkage and fuzzy matching on all organizations. Surface potential name variants of organizations for manual review.
    After manual review is finished, new name variants are added to the 2026 name map. 
(3) Run record linkage algorithm on individuals to surface name variants for manual review AND surface potential cases
    of diferent individuals who share the same name. After manual review, add name variants and disambiguation cases to the
    2026 name map.
 

# Load packages

```{r}
here::i_am("0802_entity_resolution_simplified.Rmd")
library(here)
library(dplyr)
library(stringr)
library(stringdist)
library(fuzzyjoin)
library(reclin2)
library(phonics)
library(igraph)
library(readr)
library(digest)
library(jsonlite)
library(tidyr)
library(uuid)

source("standardization_helpers.R")
```

## Define helper functions

```{r}
# functions used in record linkage
safe_metaphone <- function(x) {
  clean <- str_replace_all(coalesce(x, ""), "[^A-Za-z]", "")
  ifelse(clean != "", metaphone(clean), NA_character_)
}

bigram_cosine <- function(a, b) stringsim(a, b, method = "cosine", q = 2)

# creates unique hash-based key for contributors
hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16))
}

# makes contributor_id
make_row_hash <- function(...) {
  cols <- list(...)
  n <- length(cols[[1]])
  vapply(seq_len(n), function(i) {
    parts <- vapply(cols, function(col) as.character(col[i]), character(1))
    digest::digest(paste(parts, collapse = "|"), algo = "xxhash32")
  }, character(1))
}

# recipient_key for applying threshold checks:
#   PAC contributions        -> Recipient.Committee.ID (each PAC evaluated separately)
#   ballot measure           -> name of Ballot.Measure (groups all committees for same measure/side)
#   candidate with Office    -> "Recipient.Name | Office" key for candidate race (Ben Allen has more than one committee)
add_recipient_key <- function(df) {
  df %>% mutate(
    recipient_key = case_when(
      coalesce(is_pac_contribution, FALSE)          ~ as.character(Recipient.Committee.ID),
      !is.na(Ballot.Measure) & Ballot.Measure != "" ~ Ballot.Measure,
      !is.na(Office) & Office != ""                 ~ paste(Recipient.Name, Office, sep = " | "),
      TRUE                                          ~ Recipient.Name
    )
  )
}

# look up contributor name in entity registry
# returns entity_id, is_conflict and is_new
lookup_cluster_uuid <- function(raw_names, mapping_df) {
  clean <- unique(toupper(raw_names))
  matched_uuids <- mapping_df %>%
    filter(raw_name %in% clean) %>%
    pull(entity_uuid) %>%
    unique()

  if (length(matched_uuids) == 0) {
    list(entity_id = UUIDgenerate(), is_conflict = FALSE, is_new = TRUE)
  } else if (length(matched_uuids) == 1) {
    list(entity_id = matched_uuids[1], is_conflict = FALSE, is_new = FALSE)
  } else {
    list(entity_id = matched_uuids[1], is_conflict = TRUE, is_new = FALSE)
  }
}
```

# Load and standardize data

**Name of already-classified data file needs to be updated manually with the most recent classifications.**

```{r}
# full contributions dataset generated in 0801
contributions_full <- read.csv("08_inputs/power_search_contributions_normalized.csv")
nrow(contributions_full)

# already classified contributors
categorized <- read.csv("08_inputs/already_classified_contributions_09-06-26.csv")
nrow(categorized)
```

```{r}
# load ID corrections data
library(googlesheets4)
gs4_deauth()

id_corrections <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1UnO3wldvEo84Md0ac5jbpg4SqWzvPle3yh15e_DGAVk",
  sheet = "corrected_pac_ids",
  col_types = "c"
) %>%
  mutate(
    Contributor.ID = coalesce(Contributor.ID, ""),
    resolved       = !is.na(resolved) & tolower(str_trim(resolved)) != "false",
    is_fed_pac     = tolower(str_trim(coalesce(is_fed_pac, ""))) %in% c("true", "1", "yes")
  )

contributions_full <- apply_contributor_corrections(contributions_full, id_corrections)
```


### filter for races and props we are classifying

```{r}
# filter for gov and IC race + ALL props (not including prop 50)
contributions <- contributions_full %>% 
  rename(
    Ballot.Measure = `Ballot.Measure.s.`
  ) %>% 
  filter(
    Recipient.Name == "BECERRA, XAVIER" | 
    Recipient.Name == "HILTON, STEVE" | 
    (Recipient.Name == "ALLEN, BEN" & Office == "Insurance Commissioner") | 
    Recipient.Name == "KIM, JANE" | 
    (Ballot.Measure != "" & Ballot.Measure != "SUPPORTED: PROPOSITION 050 - ACA 8 (RIVAS) CONGRESSIONAL REDISTRICTING. (RES. CH. 156, 2025)" & Ballot.Measure != "OPPOSED: PROPOSITION 050 - ACA 8 (RIVAS) CONGRESSIONAL REDISTRICTING. (RES. CH. 156, 2025)")
) 
```

```{r}
# data check : compare to power search to make sure the amounts match
contributions %>% filter(Recipient.Name == "BECERRA, XAVIER") %>% summarise(total = sum(Amount))
contributions %>% filter(Recipient.Name == "HILTON, STEVE") %>% summarise(total = sum(Amount))
```

```{r}
# id_map maps XREF_FILER_ID (Contributor.ID) -> FILER_ID (Recipinet.Committee.ID)
id_map_early <- read_tsv("08_inputs/FILERNAME_CD.tsv", col_types = cols(.default = "c")) %>%
  mutate(FILER_ID = as.character(FILER_ID), XREF_FILER_ID = as.character(XREF_FILER_ID)) %>%
  filter(!is.na(FILER_ID), !is.na(XREF_FILER_ID)) %>%
  distinct(XREF_FILER_ID, FILER_ID,FILER_TYPE)


contributions <- contributions %>% 
  left_join(id_map_early, by = c("Contributor.ID" = "XREF_FILER_ID")) %>%
  mutate(FILER_ID = as.character(FILER_ID))
```



# Load PACs from previous run 

On the first run `pac_universe.csv` will be stale — it is generated in 0802b.
0802b runs the PAC graph over all PACs that contributed to the gov/IC race or one of the props.
`pac_universe.csv` include ALL PACs, regardless of contribution amount that are 
surfaced in the PAC graph in 0802b (all PACs indirectly connected to one of our 
prop/candidate committees). Matching algorithms are applied across all contributors to these PACs
to make sure we're properly aggregating all name variants of the same individual/organization.

```{r}
if (file.exists("08_outputs/pac_universe.csv")) {
  pac_universe     <- read.csv("08_outputs/pac_universe.csv", colClasses = "character")
  known_pacs       <- pac_universe %>% filter(as.logical(traversed))
  pac_contributors <- contributions_full %>%
    filter(Recipient.Committee.ID %in% known_pacs$pac_id)
  cat(sprintf("Loaded %d traversed PACs from pac_universe.csv (%d contributor rows)\n",
              nrow(known_pacs), nrow(pac_contributors)))
} else {
  known_pacs       <- data.frame(pac_id = character())
  pac_contributors <- contributions_full[0, ]
  cat("pac_universe.csv not found — run 0802b first, then re-run this notebook.\n")
}

# temporary fix to add pacs for props that were not previously being classified
if (file.exists("08_outputs/pac_universe_new_props.csv")) {
  pac_universe_new_props_raw <- read.csv("08_outputs/pac_universe_new_props.csv",
                                          colClasses = "character")
  pac_universe <- bind_rows(pac_universe, pac_universe_new_props_raw) %>%
    group_by(pac_id) %>%
    summarise(
      depth                      = min(as.integer(depth), na.rm = TRUE),
      is_seed                    = any(as.logical(is_seed),                    na.rm = TRUE),
      traversed                  = any(as.logical(traversed),                  na.rm = TRUE),
      has_pac_no_id_contributors = any(as.logical(has_pac_no_id_contributors), na.rm = TRUE),
      is_candidate_pac           = any(coalesce(as.logical(is_candidate_pac), FALSE)),
      is_federal_pac             = any(coalesce(as.logical(is_federal_pac),   FALSE)),
      .groups = "drop"
    )
}

```

```{r}
# derive known_pacs and pac_contributors from merged pac lists
known_pacs <- pac_universe %>%
  filter(
    coalesce(as.logical(traversed),        FALSE),
    !coalesce(as.logical(is_candidate_pac), FALSE),
    !coalesce(as.logical(is_federal_pac),   FALSE)
  )

pac_contributors <- contributions_full %>%
  rename(Ballot.Measure = Ballot.Measure.s.) %>%
  filter(Recipient.Committee.ID %in% known_pacs$pac_id)

cat(sprintf("Merged universe: %d traversed non-candidate/non-federal PACs (%d contributor rows)\n",
            nrow(known_pacs), nrow(pac_contributors)))
```


# Combine direct and pac contributor data

Add PAC contributors to the `contributions` data to do entity resolution. 
It is possible a direct contributor can also be counted under PAC contributions
this happens when a ballot measure committee (e.g. YES ON 41 - CALIFORNIANS FOR TRANSPARENCY AND ACCOUNTABILITY, A COALITION OF SMALL BUSINESSES, TAXPAYERS, GOOD GOVERNMENT ADVOCATES, ACCOUNTANTS AND AUDITING EXPERTS)
contributes to a PAC. 

```{r}
n_direct <- nrow(contributions)
n_pac    <- nrow(pac_contributors)
contributions <- bind_rows(
  contributions    %>% mutate(is_pac_contribution = FALSE), # flag for distinguishing direct contributors from PAC contributors. direct wins
  pac_contributors %>% mutate(is_pac_contribution = TRUE)
) %>%
  distinct(rowid, .keep_all = TRUE)  # rowid is the source row number; both datasets share it b/c they are derived from contributions_full, so this drops PAC rows that duplicate a direct-contributor row
cat(sprintf("%d direct + %d pac → %d combined (%d duplicate rows dropped)\n",
            n_direct, n_pac, nrow(contributions),
            n_direct + n_pac - nrow(contributions)))
nrow(contributions)
```


# Data processing

## manually process edge cases to correctly flag orgs vs individuals

Correct cases that the is_individual() function got wrong. 

```{r}
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$standardized_name <- "LOVECCHIO, PETER"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$processed_name <- "LOVECCHIO PETER"

contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$standardized_name <- "ZUNILDA, PEREZ"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$processed_name <- "ZUNILDA, PEREZ"

contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$standardized_name <- "LIDELL, JOHN MARK" # PA-C = physician assistant (don't want to add PAC to suffix list though)
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$processed_name <- "LIDELL JOHN MARK"

contributions[which(contributions$Contributor.Name == "Last0b6df6bd, First7606aee3"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lastcbea8f37, Firstd599d312"),]$entity_type <- "individual"
```

## standardize employer + occupation for individuals

For individuals that are missing employer/occupation info, standardize employer/occupation to be `NONE` 
Note: there are edge cases where employer is N/A, but occupation indicates there are employed or self-employed. 
These individuals are tagged as major donors. 

```{r}
contributions <- contributions %>%
  mutate(
    standardized_employer_name   = if_else(entity_type == "individual" & standardized_employer_name == "", "NONE", standardized_employer_name),
    processed_employer_name  = if_else(entity_type == "individual" & processed_employer_name == "", "NONE", processed_employer_name),

    standardized_occupation  = if_else(entity_type == "individual" & standardized_occupation == "", "NONE", standardized_occupation),
    processed_occupation = if_else(entity_type == "individual" & processed_occupation == "", "NONE", processed_occupation),    
  )

```

# Load and standardize name maps

Load Digital Democracy's org concept map. 
This is used in entity resolution to tentatively link organizations belonging to the same org concept
that would not otherwise be linked through fuzzy matching or name variant matching. To be consistent
with the Voter Guide, we don't use all of the same groupings as the org concept map (e.g. chapters
of unions are kept separate). 

```{r}
map1 <- read.csv("08_inputs/dd_names_mapped_to_org_concepts_3_18_2026.csv") %>%
  rename(canonical_name = org_concept_name, alias_group_id = oid) %>%
  mutate(
    record_name    = str_squish(record_name),
    canonical_name = str_squish(canonical_name),
    mapping_source = "dd_name_mapping",
    confirmed      = TRUE
  )
```


# Update entity registry and no-link pairs

Syncs the 2026 Names CSV (used for normalization) and the name_entity_mapping.csv (tracks uuids per entity).
Dataset descriptions: 
`2026 Contributor name mapping and descriptions - Names.csv` : Official source of truth - maps record name to normalized name. Shared by the team. 
`entity_registry.csv` : keeps track of each canonical name that links name variants within an entity and the corresponding uuid.
`name_entity_mapping_rebuilt.csv` : essentially a copy of the Name 2026 data, but has a uuid for each row to keep track of unique entities (one uuid per `Normalized` name).
                                    updated manually with each review queue. 
`no_link_registry.csv` : feedback from each manual review. org names that were compared and determined not to be a link added to the registry.
                         note that only names w/in a review group are compared, may miss some name variants that weren't grouped in the same review group 
                         (which is why only names within a review group are added to the no-link registry).
                         
Download latest version of 2026 names map at: https://docs.google.com/spreadsheets/d/1THwQtvw5s9ZO9n7QHncekl4kRMc1Nf3MJHViSYALDps/edit?gid=897118445#gid=897118445 


```{r}
ENTITY_REGISTRY_PATH     <- here("08_inputs/entity_registry.csv") 
NAME_ENTITY_MAPPING_PATH <- here("08_inputs/name_entity_mapping_rebuilt.csv") # updated with each entity resolution review queue. 
NO_LINK_REGISTRY_PATH <- here("08_inputs/no_link_registry.csv")
registry_mtime <- if (file.exists(NO_LINK_REGISTRY_PATH)) file.mtime(NO_LINK_REGISTRY_PATH) else as.POSIXct(0, origin = "1970-01-01")

# only load feedback files newer than the registry's last write
# all historical files are loaded as a one-time backfill
NEW_ORG_FEEDBACK_FILES <- {
  f <- list.files(here("08_inputs"), pattern = "^org_name_variant_review_queue_w_review.*\\.csv$", full.names = TRUE)
  f[file.mtime(f) > registry_mtime]
}
NEW_INDIV_FEEDBACK_FILES <- {
  f <- list.files(here("08_inputs"), pattern = "^indiv_name_variant_review_queue_w_review.*\\.csv$", full.names = TRUE)
  f[file.mtime(f) > registry_mtime]
}

print(NEW_ORG_FEEDBACK_FILES)
print(NEW_INDIV_FEEDBACK_FILES)
```

```{r}
# official source of truth - shared by voter guide team 
names_2026 <- read.csv(
  here("08_inputs/2026 Contributor name mapping and descriptions - Names.csv"),
  stringsAsFactors = FALSE
) %>%
  rename(raw_name = Original, canonical_name = Normalized) %>% # rename columns so all dataframes share col names
  filter(!is.na(raw_name), raw_name != "", !is.na(canonical_name), canonical_name != "") %>%
  mutate(
    raw_name       = (toupper(raw_name)), 
    canonical_name = str_squish(canonical_name) # canonical name shouldn't have any double spaces
  )

# persistent UUID registry — updated by manual review (should be a copy of Names 2026, but tracks uuid)
if (file.exists(NAME_ENTITY_MAPPING_PATH)) {
  name_entity_mapping_existing <- read.csv(NAME_ENTITY_MAPPING_PATH, stringsAsFactors = FALSE) %>%
    mutate(confirmed = as.logical(confirmed))
} else {
  name_entity_mapping_existing <- tibble(raw_name = character(), entity_uuid = character(),
                                          canonical_name = character(), confirmed = logical(),
                                          added_date = character(), source = character())
}
```


```{r}
# create uuid lookup - matches canonical name to its uuid
canonical_uuid_conflicts <- name_entity_mapping_existing %>%
  filter(!is.na(canonical_name), confirmed == TRUE) %>%
  mutate(canonical_name = str_squish(canonical_name)) %>%
  group_by(canonical_name) %>%
  summarise(n_uuids = n_distinct(entity_uuid), uuids = paste(unique(entity_uuid), collapse = " | "), .groups = "drop") %>%
  filter(n_uuids > 1)

if (nrow(canonical_uuid_conflicts) > 0) {
  print(canonical_uuid_conflicts)
  stop(sprintf("%d canonical name(s) are mapped to more than one entity_uuid — resolve conflicts before proceeding.", nrow(canonical_uuid_conflicts)))
}

canonical_uuid_lookup <- name_entity_mapping_existing %>%
  filter(!is.na(canonical_name), confirmed == TRUE) %>%
  mutate(canonical_name = str_squish(canonical_name)) %>%
  distinct(canonical_name, entity_uuid) %>%
  select(canonical_name, entity_uuid)

# add uuids to any new entities added to the 2026 Names map (added by someone else on the team.)
canonical_uuid_map <- names_2026 %>%
  mutate(canonical_name = str_squish(canonical_name)) %>%
  distinct(canonical_name) %>%
  left_join(canonical_uuid_lookup, by = "canonical_name") %>% # join uuids to already known canonical names 
  mutate(entity_uuid = coalesce(entity_uuid, UUIDgenerate(n = n()))) # create new uuids for new canonical names

cat(sum(is.na(left_join(names_2026 %>% distinct(canonical_name),
                         canonical_uuid_lookup, by="canonical_name")$entity_uuid)),
    "new canonical groups in 2026 Names CSV assigned fresh UUIDs\n")

```


```{r}
#write.csv(canonical_uuid_map, "08_outputs/canonical_uuid_map.csv", row.names = FALSE)
```


## Individual disambiguation subgroups

Contributors are aggregated based on their names; individual disambiguation separates individuals 
with the same name who have been manually confirmed to be different people by adding other variables
(employer, occupation, etc.) to the map and assigning a subgroup to these individuals. 

`subgroup_id` (A, B, C, etc.) distinguishes rows *within* one `canonical_name` — the same letter reused
under a different `canonical_name` is unrelated. Each confirmed `(canonical_name, subgroup_id)`
pair is a distinct person; give contributor their own persistent uuid, tracked in `name_entity_mapping.csv`
under a synthetic `canonical_name::subgroup_id` key. 

Records that don't match any subgroup's profile keep the plain canonical name uui.

```{r}
if (!"subgroup_id" %in% names(names_2026)) names_2026$subgroup_id <- NA_character_

subgroup_profiles <- names_2026 %>%
  filter(!is.na(subgroup_id), subgroup_id != "") %>%
  transmute(
    canonical_name     = canonical_name,
    subgroup_id        = as.character(subgroup_id),
    profile_employer   = na_if(toupper(str_squish(coalesce(Contributor.Employer, ""))), ""),
    profile_occupation = na_if(toupper(str_squish(coalesce(Contributor.Occupation, ""))), ""),
    profile_city       = na_if(toupper(str_squish(coalesce(Contributor.City, ""))), ""),
    profile_state      = na_if(toupper(str_squish(coalesce(Contributor.State, ""))), ""),
    profile_zip        = na_if(coalesce(as.character(Contributor.Zip.Code), ""), ""),
    profile_committee  = na_if(coalesce(as.character(Recipient.Committee.ID), ""), "")
  ) %>%
  distinct()

# assign a uuid per (canonical_name, subgroup_id)
subgroup_keys <- subgroup_profiles %>%
  distinct(canonical_name, subgroup_id) %>%
  mutate(raw_name = paste(canonical_name, subgroup_id, sep = "::"))

known_subgroup_uuids <- name_entity_mapping_existing %>%
  filter(source == "subgroup_disambig") %>%
  distinct(raw_name, entity_uuid)

new_subgroup_keys <- subgroup_keys %>%
  filter(!raw_name %in% known_subgroup_uuids$raw_name) %>%
  mutate(
    entity_uuid    = UUIDgenerate(n = n()),
    canonical_name = NA_character_, # composite key, not a real canonical name - excluded from the canonical/uuid checks above
    confirmed      = TRUE,
    added_date     = as.character(Sys.Date()),
    source         = "subgroup_disambig"
  ) %>%
  select(raw_name, entity_uuid, canonical_name, confirmed, added_date, source)

if (nrow(new_subgroup_keys) > 0) {
  name_entity_mapping_existing <- bind_rows(name_entity_mapping_existing, new_subgroup_keys)
  write.csv(name_entity_mapping_existing, NAME_ENTITY_MAPPING_PATH, row.names = FALSE)
  cat("Assigned", nrow(new_subgroup_keys), "new subgroup uuid(s) for disambiguated individuals\n")
}

subgroup_uuid_lookup <- bind_rows(
    known_subgroup_uuids,
    new_subgroup_keys %>% select(raw_name, entity_uuid)
  ) %>%
  separate(raw_name, into = c("canonical_name", "subgroup_id"), sep = "::", extra = "merge")
```

```{r}
# match a contribution record to a subgroup: it matches subgroup S if it agrees with AT LEAST
# ONE of S's profile rows on every field that row has filled in
# records matching more than one subgroup are left unresolved 
# records matching none keep the default canonical-name-level uuid
assign_subgroup_entity_id <- function(df, default_col = "entity_uuid") {
  if (nrow(subgroup_profiles) == 0) return(df)

  df <- df %>% mutate(.match_row_id = row_number())

  candidates <- df %>%
    filter(Normalized %in% subgroup_profiles$canonical_name) %>%
    transmute(
      .match_row_id,
      canonical_name  = Normalized,
      cand_employer   = toupper(str_squish(coalesce(Contributor.Employer, ""))),
      cand_occupation = toupper(str_squish(coalesce(Contributor.Occupation, ""))),
      cand_city       = toupper(str_squish(coalesce(Contributor.City, ""))),
      cand_state      = toupper(str_squish(coalesce(Contributor.State, ""))),
      cand_zip        = coalesce(as.character(Contributor.Zip.Code), ""),
      cand_committee  = coalesce(as.character(Recipient.Committee.ID), "")
    ) %>%
    inner_join(subgroup_profiles, by = "canonical_name", relationship = "many-to-many") %>%
    mutate(
      is_match =
        (is.na(profile_employer)   | profile_employer   == cand_employer)   &
        (is.na(profile_occupation) | profile_occupation == cand_occupation) &
        (is.na(profile_city)       | profile_city       == cand_city)       &
        (is.na(profile_state)      | profile_state       == cand_state)      &
        (is.na(profile_zip)        | profile_zip        == cand_zip)        &
        (is.na(profile_committee)  | profile_committee  == cand_committee)
    ) %>%
    filter(is_match) %>%
    distinct(.match_row_id, canonical_name, subgroup_id)

  ambiguous <- candidates %>%
    group_by(.match_row_id) %>%
    filter(n_distinct(subgroup_id) > 1) %>%
    ungroup()

  if (nrow(ambiguous) > 0) {
    warning(sprintf(
      "%d contribution record(s) matched more than one disambiguation subgroup -- left unresolved (default uuid used). Review:",
      n_distinct(ambiguous$.match_row_id)
    ))
    print(
      df %>% inner_join(ambiguous %>% distinct(.match_row_id), by = ".match_row_id") %>%
        select(Contributor.Name, Contributor.Employer, Contributor.Occupation,
               Contributor.City, Contributor.State, Contributor.Zip.Code, Recipient.Committee.ID)
    )
  }

  resolved <- candidates %>%
    anti_join(ambiguous, by = ".match_row_id") %>%
    distinct(.match_row_id, canonical_name, subgroup_id) %>%
    left_join(subgroup_uuid_lookup, by = c("canonical_name", "subgroup_id")) %>%
    select(.match_row_id, subgroup_entity_uuid = entity_uuid)

  df %>%
    left_join(resolved, by = ".match_row_id") %>%
    mutate(!!default_col := coalesce(subgroup_entity_uuid, .data[[default_col]])) %>%
    select(-.match_row_id, -subgroup_entity_uuid)
}
```

```{r}
# anyhing in names 2026 that's not in entity mapping?
names_2026 %>% filter(!(str_squish(toupper(raw_name)) %in% str_squish(toupper(name_entity_mapping_existing$raw_name))))
```


```{r}
org_req_cols   <- c("to_add_status", "entity_name", "canonical_name", "review_group_id", "confirmed_link")
indiv_req_cols <- c("entity_name", "canonical_name", "entity_group_id", "confirmed_link")

# load feedback from review queues
load_feedback <- function(path, req_cols) {
  if (is.null(path) || !file.exists(path)) return(NULL)
  df <- read.csv(path, stringsAsFactors = FALSE)
  if (nrow(df) == 0 || !all(req_cols %in% names(df))) return(NULL)
  out      <- df[, req_cols]
  out$flag              <- if ("flag"              %in% names(df)) df$flag              else NA_character_
  out$original_group_id <- if ("original_group_id" %in% names(df)) df$original_group_id else NA_character_ # original review groups as an indicator for what was compared
  out
}

# no-link registry: all confirmed pairs from reviews that reviewer determined should
# not be liked under the same registry
# for each reviewed group, any two names in DIFFERENT subgroups are a confirmed not-linked
# pair — this includes singletons vs. sub-entities AND pairs across different sub-entities
# Pairs are stored and used to suppress already-reviewed pairs from future review queues
if (file.exists(NO_LINK_REGISTRY_PATH)) {
  no_link_registry <- read.csv(NO_LINK_REGISTRY_PATH, stringsAsFactors = FALSE)
} else {
  no_link_registry <- tibble(name_a = character(), name_b = character(),
                              entity_type = character(), reviewed_date = character())
}

# each file is processed independently — group IDs are only compared within the same review queue file
# original_group_id column (added to queue outputs going forward) identifies the pre-split
# group; falls back to stripping the reviewer-added suffix letter(s) for older files.
extract_no_link_pairs <- function(fb, group_id_col, entity_type_label) {
  empty <- tibble(name_a = character(), name_b = character(),
                  entity_type = character(), reviewed_date = character())
  if (is.null(fb) || nrow(fb) == 0) return(empty)
  grp <- fb %>%
    filter(tolower(trimws(coalesce(confirmed_link, ""))) != "uncertain") %>%
    mutate(
      raw_name    = toupper(entity_name),
      subgroup_id = .data[[group_id_col]], # subgroup ids reflect which name variants were determined to be linked
      orig_grp    = coalesce(original_group_id, sub("_[a-z]+$", "", .data[[group_id_col]])) # replace added suffixes to reconstruct original review group ids
    ) %>%
    distinct(raw_name, subgroup_id, orig_grp)
  left  <- grp %>% rename(name_a = raw_name, sg_a = subgroup_id)
  right <- grp %>% rename(name_b = raw_name, sg_b = subgroup_id)
  left %>%
    inner_join(right, by = "orig_grp", relationship = "many-to-many") %>%
    filter(sg_a != sg_b, name_a < name_b) %>% 
    transmute(name_a = name_a, name_b = name_b,
              entity_type = entity_type_label, reviewed_date = as.character(Sys.Date())) %>%
    distinct()
}

new_no_link_org_pairs <- bind_rows(lapply(NEW_ORG_FEEDBACK_FILES, function(f) { # apply to each feedback file individually. 
  extract_no_link_pairs(load_feedback(f, org_req_cols), "review_group_id", "org")
}))
new_no_link_indiv_pairs <- bind_rows(lapply(NEW_INDIV_FEEDBACK_FILES, function(f) {
  extract_no_link_pairs(load_feedback(f, indiv_req_cols), "entity_group_id", "indiv")
}))
new_no_link_pairs <- bind_rows(new_no_link_org_pairs, new_no_link_indiv_pairs)

pair_key <- function(df) paste(df$name_a, df$name_b, df$entity_type, sep = "|||") # key to distinguish pairs
no_link_registry <- bind_rows(
  new_no_link_pairs,
  no_link_registry %>% filter(!pair_key(.) %in% pair_key(new_no_link_pairs))
)

# build confirmed mapping rows from 2026 Names + entity registry
names_from_b <- names_2026 %>%
  distinct(raw_name, canonical_name) %>%
  left_join(canonical_uuid_map %>% select(canonical_name, entity_uuid), by = "canonical_name") %>%
  mutate(confirmed = TRUE, added_date = as.character(Sys.Date()), source = "2026_names") %>% 
  select(raw_name, entity_uuid, canonical_name, confirmed, added_date, source)

extra_from_c <- name_entity_mapping_existing %>% # anything not yet added to Names 2026 (typically, it should have been added already)
  filter(!((raw_name) %in% (names_from_b$raw_name)))

name_entity_mapping <- bind_rows(names_from_b, extra_from_c)

entity_registry <- name_entity_mapping %>%
  filter(!is.na(canonical_name)) %>%
  distinct(canonical_name, entity_uuid) %>%
  arrange(canonical_name)

# check for raw_name associated w/ multiple canonical names. 
# duplicate rows with the SAME canonical are harmless - will be collapsed below
raw_name_conflicts <- name_entity_mapping %>%
  filter(!is.na(canonical_name)) %>%
  group_by(raw_name) %>%
  summarise(n_canonical    = n_distinct(canonical_name),
            canonical_names = paste(sort(unique(canonical_name)), collapse = " | "),
            .groups = "drop") %>%
  filter(n_canonical > 1)

if (nrow(raw_name_conflicts) > 0) {
  warning(sprintf("%d raw name(s) map to more than one canonical name — first occurrence used. Review:",
                  nrow(raw_name_conflicts)))
  print(raw_name_conflicts)
}

# check for one uuid per canonical name
uuid_canonical_conflicts <- name_entity_mapping %>%
  filter(!is.na(canonical_name), !is.na(entity_uuid)) %>%
  group_by(entity_uuid) %>%
  summarise(n_canonical    = n_distinct(canonical_name),
            canonical_names = paste(sort(unique(canonical_name)), collapse = " | "),
            .groups = "drop") %>%
  filter(n_canonical > 1)

if (nrow(uuid_canonical_conflicts) > 0) {
  warning(sprintf("%d entity_uuid(s) map to more than one canonical name — first occurrence used. Review:",
                  nrow(uuid_canonical_conflicts)))
  print(uuid_canonical_conflicts)
}

# de-duplicate
nem_lookup <- name_entity_mapping %>%
  distinct(raw_name, .keep_all = TRUE)

# in map, add rows where the raw_name = canonical_name so name variants can be compared against canonical names as well. 
canonical_as_raw <- name_entity_mapping %>%
  filter(!is.na(canonical_name), confirmed == TRUE) %>%
  distinct(canonical_name, .keep_all = TRUE) %>%
  transmute(
    raw_name       = canonical_name,
    entity_uuid    = entity_uuid,
    canonical_name = canonical_name,
    confirmed      = TRUE,
    added_date     = added_date,
    source         = source
  ) %>%
  filter(!raw_name %in% nem_lookup$raw_name)

nem_lookup <- bind_rows(nem_lookup, canonical_as_raw)
```


Manually flag instances of one name being associated with multiple canonical names in the `2026 Names` 
Google Sheet: https://docs.google.com/spreadsheets/d/1THwQtvw5s9ZO9n7QHncekl4kRMc1Nf3MJHViSYALDps/edit?gid=897118445#gid=897118445
(or manually resolve if error came from name variant review queue)



# ENTITY RESOLUTION / NAME VARIANT MATCHING

Apply name variant matching, fuzzy matching and custom record linkage algorithm to identify contributors that are likely the same, but associated with different name variants. 
Name variants will be manually reviewed and if confirmed, they will be added to the Names 2026 mapping. 
After all name variants are added to the map, contributors will then be aggregated by their `Normalized` name. 

## Organizations

1. Name variant linkage: name variants of an org name are compared for exact matches. Name variants include removing business ending, removing parentheses, etc.
2. Fuzzy matching performed on org name to id additional potential pairs.
3. DD org concept maps used to surface potential links between org names. 

```{r}
# make variants of each contributor name
make_name_variants <- function(df, col) {
  df %>% mutate(
    name_v1 = standardize_names(.data[[col]]) %>% fix_typos(), # simple standardization and known typos corrected
    name_v2 = remove_pac_info(name_v1)        %>% str_squish(), # removing FEC ID, FED PAC ID, MAJOR DONOR ID
    # remove name_v3 -> need to rename
    name_v4 = str_remove_all(name_v1, "\\s*\\(.*?\\)") %>% str_squish(), # remove parentheses
    name_v5 = replace_business_text(name_v4)   %>% str_squish() # remove business ending e.g. inc, corporation, etc. 
  )
}

# look for exact matches between name variants. 
find_exact_variant_pairs <- function(pool, variant_col) {
  candidates <- pool %>%
    filter(!is.na(.data[[variant_col]]), .data[[variant_col]] != "") %>%
    group_by(across(all_of(variant_col))) %>%
    filter(n() > 1) %>%
    ungroup()

  if (nrow(candidates) == 0)
    return(tibble(.x = integer(), .y = integer(), match_source = character()))

  candidates %>%
    group_by(across(all_of(variant_col))) %>%
    group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>%
    ungroup() %>%
    select(.x, .y) %>%
    mutate(match_source = paste0("exact_", variant_col))
}
```

```{r}
orgs <- contributions %>%
  filter(entity_type == "organization") %>%
  mutate(cont_row_id = row_id) %>%          # preserve row_id before overwriting
  mutate(has_contributor_id = !is.na(effective_Contributor.ID) & str_trim(as.character(effective_Contributor.ID)) != "") %>%
  make_name_variants("Contributor.Name") %>%
  filter(!is.na(name_v1), name_v1 != "") %>%
  mutate(row_id = row_number())
```


## No-link registry data checks (can skip)

Surface any entities added to the no-link registry that SHOULDN'T be
there. 

```{r}
if (nrow(no_link_registry) > 0) {

  # (1) No exact pairs (name_a == name_b)
  self_pairs <- no_link_registry %>% filter(name_a == name_b)
  if (nrow(self_pairs) > 0) {
    cat(sprintf("WARNING (1): %d self-pair(s) where name_a == name_b:\n", nrow(self_pairs)))
    print(self_pairs)
  } else cat("Check 1 passed: no self-pairs.\n")

  # (2) No shared name variants across a no-link pair
  registry_name_variants <- tibble(name = union(no_link_registry$name_a, no_link_registry$name_b)) %>%
    make_name_variants("name")

  get_variants <- function(nm) {
    registry_name_variants %>%
      filter(name == nm) %>%
      select(name_v1, name_v2, name_v4, name_v5) %>%
      unlist() %>% na.omit() %>% unique() %>% as.character()
  }

  variant_conflicts <- no_link_registry %>%
    rowwise() %>%
    mutate(
      .vars_a     = list(get_variants(name_a)),
      .vars_b     = list(get_variants(name_b)),
      .shared     = list(intersect(.vars_a, .vars_b)),
      .has_shared = length(.shared) > 0
    ) %>%
    filter(.has_shared) %>%
    mutate(shared_variants = paste(.shared, collapse = " | ")) %>%
    select(name_a, name_b, entity_type, shared_variants) %>%
    ungroup()

  if (nrow(variant_conflicts) > 0) {
    cat(sprintf("WARNING (2): %d no-link pair(s) share a name variant:\n", nrow(variant_conflicts)))
    print(variant_conflicts)
  } else cat("Check 2 passed: no shared name variants.\n")

  # (3) No shared Contributor.ID across a no-link pair (org pairs only)
  org_id_lookup <- orgs %>%
    filter(!is.na(effective_Contributor.ID),
           str_trim(as.character(effective_Contributor.ID)) != "") %>%
    mutate(.name_key = toupper(name_v1)) %>%
    select(.name_key, effective_Contributor.ID) %>%
    distinct()

  get_ids <- function(nm) {
    org_id_lookup %>%
      filter(.name_key == nm) %>%
      pull(effective_Contributor.ID) %>%
      as.character() %>% unique()
  }

  id_conflicts <- no_link_registry %>%
    filter(entity_type == "org") %>%
    rowwise() %>%
    mutate(
      .ids_a      = list(get_ids(name_a)),
      .ids_b      = list(get_ids(name_b)),
      .shared     = list(intersect(.ids_a, .ids_b)),
      .has_shared = length(.shared) > 0
    ) %>%
    filter(.has_shared) %>%
    mutate(shared_ids = paste(.shared, collapse = " | ")) %>%
    select(name_a, name_b, entity_type, shared_ids) %>%
    ungroup()

  if (nrow(id_conflicts) > 0) {
    cat(sprintf("WARNING (3): %d no-link pair(s) share a Contributor.ID:\n", nrow(id_conflicts)))
    print(id_conflicts)
  } else cat("Check 3 passed: no shared Contributor.IDs.\n")

  # (4) No shared chapter/unit numbers (digit sequences) across a no-link pair
  chapter_conflicts <- no_link_registry %>%
    rowwise() %>%
    mutate(
      .digits_a   = list(str_extract_all(name_a, "\\d+")[[1]]),
      .digits_b   = list(str_extract_all(name_b, "\\d+")[[1]]),
      .shared     = list(intersect(.digits_a, .digits_b)),
      .has_shared = length(.digits_a) > 0 & length(.digits_b) > 0 & length(.shared) > 0
    ) %>%
    filter(.has_shared) %>%
    mutate(shared_numbers = paste(.shared, collapse = " | ")) %>%
    select(name_a, name_b, entity_type, shared_numbers) %>%
    ungroup()

  if (nrow(chapter_conflicts) > 0) {
    cat(sprintf("NOTE (4): %d no-link pair(s) share digit sequences (check chapter numbers):\n",
                nrow(chapter_conflicts)))
    print(chapter_conflicts)
  } else cat("Check 4 passed: no shared chapter/unit numbers.\n")

} else {
  cat("No-link registry is empty — skipping checks.\n")
}
```



## Step 1: ID-based aliases (same Contributor.ID, different names)

```{r}
# all orgs with more than one name variant per contributor.id
id_based_aliases <- orgs %>%
  filter(has_contributor_id) %>%
  group_by(effective_Contributor.ID) %>%
  filter(n_distinct(Contributor.Name) > 1) %>%
  ungroup()

# list of all name pairs that share the same id
id_auto_pairs <- id_based_aliases %>%
  group_by(effective_Contributor.ID) %>%
  group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>% # combn() generates all pairs of the ids
  ungroup() %>%
  mutate(link_tier = "auto_link", match_source = "same_contributor_id")

cat("ID-based alias pairs:", nrow(id_auto_pairs), "\n")

id_based_aliases %>%
  group_by(effective_Contributor.ID) %>%
  summarise(names = paste(unique(Contributor.Name), collapse = " | "), n = n(), .groups = "drop") %>%
  arrange(desc(n))
```


## Step 2: Exact variant cascade

```{r}
# look for exact matches between name variants. 
find_exact_variant_pairs <- function(pool, variant_col) {
  candidates <- pool %>%
    filter(!is.na(.data[[variant_col]]), .data[[variant_col]] != "") %>%
    group_by(across(all_of(variant_col))) %>%
    filter(n() > 1) %>% # more than one record shares an exact match of the name variant
    ungroup()

  if (nrow(candidates) == 0)
    return(tibble(.x = integer(), .y = integer(), match_source = character()))

  candidates %>%
    group_by(across(all_of(variant_col))) %>%
    group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>% # create pairs
    ungroup() %>%
    select(.x, .y) %>%
    mutate(match_source = paste0("exact_", variant_col))
}
```


```{r}
id_resolved_rows <- unique(c(id_auto_pairs$.x, id_auto_pairs$.y)) # contribution records already linked through contributor.id
#remaining_orgs   <- orgs %>% filter(!row_id %in% id_resolved_rows) # filter out records that were already grouped through a contributor.id match
remaining_orgs <- orgs # include ALL orgs so we can link PACs and non-PACs

variant_cols     <- c("name_v1", "name_v2", "name_v4", "name_v5") # skips name_v3 which removes unit info
seen_pairs       <- tibble(.x = integer(), .y = integer())
exact_pairs_list <- list()

for (v in variant_cols) {
  new_pairs <- find_exact_variant_pairs(orgs, v) %>%
    anti_join(seen_pairs, by = c(".x", ".y"))
  if (nrow(new_pairs) > 0) {
    exact_pairs_list[[v]] <- new_pairs
    seen_pairs <- bind_rows(seen_pairs, new_pairs %>% select(.x, .y))
  }
  cat("variant", v, "- new pairs:", nrow(new_pairs), "\n")
}

exact_pairs <- bind_rows(exact_pairs_list) %>%  # no dedup needed; each pair has its most-specific source
  left_join(orgs %>% select(row_id, id_x = effective_Contributor.ID, has_id_x = has_contributor_id), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, id_y = effective_Contributor.ID, has_id_y = has_contributor_id), by = c(".y" = "row_id")) %>%
  mutate(link_tier = if_else(
    coalesce(has_id_x, FALSE) & coalesce(has_id_y, FALSE) & coalesce(id_x != id_y, FALSE),
    "review", "auto_link"
  )) %>%
  select(.x, .y, link_tier, match_source)

cat(sprintf("Exact cascade: %d auto-links, %d conflicting-ID review pairs\n",
            sum(exact_pairs$link_tier == "auto_link"),
            sum(exact_pairs$link_tier == "review")))

exact_pairs %>%
  filter(link_tier == "review") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name, id_x = effective_Contributor.ID), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name, id_y = effective_Contributor.ID), by = c(".y" = "row_id")) %>%
  select(name_x, id_x, name_y, id_y, match_source) %>%
  distinct() %>% print(n = 50)
```


```{r}
exact_pairs %>%
  filter(link_tier == "review") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name, id_x = effective_Contributor.ID), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name, id_y = effective_Contributor.ID), by = c(".y" = "row_id")) %>%
  select(name_x, name_y,id_x,  id_y, match_source) %>%
  distinct() 
```


### Step 3: internal fuzzy match

The fuzzy match is the most expensive, so it doesn't run on all orgs like the other matching processes do.

1. Group orgs by links that were already estblished through contributor.id matches and name variant matches.
   The most common name per group is chosen as the `name_rep` for the group and used to fuzzy match against.
   The goal is to identify other possible name variants belonging to already-established groups that weren't
   already found during the id alias and name variant step.
2. Perform fuzzy matching on all unresolved org names + the name reps from each org group.
   This step keeps constraint that conflicting-id matches are paired since fuzzy matching is much looser, 
   but could consider removing this constraint. 

```{r}
# id and exact pairs to get one rep per already-established cluster
all_early_pairs <- bind_rows(
  id_auto_pairs %>% select(.x, .y),
  exact_pairs   %>% filter(link_tier == "auto_link") %>% select(.x, .y) # doesn't include contributor.id conflicts
)

if (nrow(all_early_pairs) > 0) {
  g_early       <- graph_from_data_frame(all_early_pairs, directed = FALSE,
                                          vertices = data.frame(name = orgs$row_id))
  early_members <- components(g_early)$membership
  cluster_reps  <- tibble(row_id = as.integer(names(early_members)), cluster = early_members) %>%
    left_join(orgs %>% select(row_id, name_v5), by = "row_id") %>%
    add_count(cluster, name_v5, name = "n_v5") %>%
    group_by(cluster) %>%
    slice_max(n_v5, n = 1, with_ties = TRUE) %>%   # most common name_v5 in cluster
    slice_min(row_id, n = 1, with_ties = FALSE) %>% # tiebreak: lowest row_id
    pull(row_id)
} else {
  cluster_reps <- integer(0)
}

linked_row_ids <- unique(c(all_early_pairs$.x, all_early_pairs$.y))

# name_reps: one rep per name_v5, covering:
#   - cluster reps (most-common name_v5 per linked group — represents whole cluster for fuzzy)
#   - unlinked singletons (not in any early pair)
name_reps <- orgs %>%
  filter(row_id %in% cluster_reps | !row_id %in% linked_row_ids) %>%
  group_by(name_v5) %>% 
  slice_min(row_id, n = 1, with_ties = FALSE) %>% # those with more than one name_v5, not already-linked are those with diff ids i think 
  ungroup()

cat(sprintf("Fuzzy matching input: %d name reps\n", nrow(name_reps)))
```


```{r}
# token Jaccard: exact token-overlap ratio between two name strings
token_jaccard <- function(a, b) {
  tok_a <- str_split(str_squish(a), "\\s+")[[1]]
  tok_b <- str_split(str_squish(b), "\\s+")[[1]]
  n_int <- length(intersect(tok_a, tok_b))
  n_uni <- length(union(tok_a, tok_b))
  if (n_uni == 0) return(0)
  n_int / n_uni
}

cat(sprintf("name_reps: %d rows\n", nrow(name_reps)))

# add chunks to help with memory issues
# candidate pairs via stringdist_inner_join 
# chunks splits the left side into batches of CHUNK_SIZE
# each batch joins against the full right side, so every
# pair is evaluated exactly once (when the lower row_id is on the left).
CHUNK_SIZE <- 500L
name_reps_right <- name_reps %>%
  rename(row_id_y = row_id, name_v5_y = name_v5) %>%
  select(row_id_y, name_v5_y)

chunks <- split(
  name_reps %>% select(row_id, name_v5),
  ceiling(seq_len(nrow(name_reps)) / CHUNK_SIZE)
)

fuzzy_candidates <- bind_rows(lapply(chunks, function(chunk) {
  stringdist_inner_join(
    chunk,
    name_reps_right,
    by           = c("name_v5" = "name_v5_y"),
    method       = "jw", p = 0.1, max_dist = 0.12,
    distance_col = "jw_dist"
  ) %>%
    filter(row_id < row_id_y)
})) %>%
  rename(.x = row_id, .y = row_id_y, name_x = name_v5, name_y = name_v5_y)

rm(chunks, name_reps_right)
gc(verbose = FALSE)

cat(sprintf("Fuzzy candidates (JW dist <= 0.12): %d pairs (from %d reps)\n",
            nrow(fuzzy_candidates), nrow(name_reps)))

# score all candidate pairs and flag conflicting-ID pairs before splitting
org_pairs_scored <- fuzzy_candidates %>%
  left_join(orgs %>% select(row_id, id_x = effective_Contributor.ID, has_id_x = has_contributor_id),
            by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, id_y = effective_Contributor.ID, has_id_y = has_contributor_id),
            by = c(".y" = "row_id")) %>%
  mutate(
    has_id_conflict = coalesce(has_id_x, FALSE) & coalesce(has_id_y, FALSE) & coalesce(id_x != id_y, FALSE),
    jaccard     = mapply(token_jaccard, name_x, name_y),
    name_jw     = 1 - jw_dist,
    name_cosine = bigram_cosine(name_x, name_y)
  ) %>%
  filter(jaccard >= 0.40) %>%   # more lenient threshold 
  mutate(
    name_sim  = (name_jw + name_cosine) / 2,
    link_tier = case_when(
      name_jw  >= 0.999 ~ "auto_link",
      name_sim >= 0.90  ~ "auto_link",
      name_sim >= 0.85  ~ "review",
      TRUE              ~ "no_link"
    ),
    match_source = "internal_fuzzy"
  )

# conflicting-ID pairs that pass scoring thresholds (currently excluded from org_pairs)
#cat("\nFuzzy pairs with conflicting contributor IDs that would be linked:\n")
conflicting_id_links <- org_pairs_scored %>%
  filter(has_id_conflict, link_tier != "no_link") %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".y" = "row_id"), suffix = c("_x", "_y")) %>%
  select(Contributor.Name_x, id_x, Contributor.Name_y, id_y, name_sim, link_tier) 

# org_pairs: conflicting IDs still excluded
org_pairs <- org_pairs_scored %>%
  filter(!has_id_conflict) %>%
  select(.x, .y, id_x, id_y, has_id_x, has_id_y, name_x, name_y,
         jaccard, name_sim, name_jw, name_cosine, link_tier, match_source)

# flag different-chapter pairs: names identical after stripping digits AND the name
# contains a chapter indicator 
CHAPTER_RE <- regex("\\b(LOCAL|LODGE|CHAPTER|COUNCIL|DISTRICT|UNIT|UNION|MEMBER DUES)\\b|\\b[A-Za-z]{2,4}\\s+\\d{3,4}\\b", ignore_case = TRUE)

org_pairs <- org_pairs %>%
  mutate(
    nx_no_digits = str_remove_all(name_x, "\\d+") %>% str_squish(),
    ny_no_digits = str_remove_all(name_y, "\\d+") %>% str_squish(),
    diff_chapter = str_detect(name_x, "\\d") & str_detect(name_y, "\\d") &
                   nx_no_digits == ny_no_digits &
                   (str_detect(name_x, CHAPTER_RE) | str_detect(name_y, CHAPTER_RE))
  ) %>%
  select(-nx_no_digits, -ny_no_digits)

# review before removing — check that nothing legitimate is being caught
diff_chapter_pairs <- org_pairs %>%
  filter(diff_chapter, link_tier != "no_link") %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".y" = "row_id"), suffix = c("_x", "_y")) %>%
  select(Contributor.Name_x, Contributor.Name_y, name_sim, link_tier) 

org_pairs <- org_pairs %>%
  filter(!diff_chapter) %>%
  select(-diff_chapter)

cat(sprintf("\nFuzzy: %d auto-links, %d review pairs, %d no-link (Jaccard >= 0.40)\n",
            sum(org_pairs$link_tier == "auto_link"),
            sum(org_pairs$link_tier == "review"),
            sum(org_pairs$link_tier == "no_link")))
```


#### data checks (can skip)

```{r, eval=FALSE}
# review different chapter pairs that were removed
diff_chapters
```

```{r,eval=FALSE}
org_pairs_scored %>%
  filter(has_id_conflict, link_tier != "no_link") %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".y" = "row_id"), suffix = c("_x", "_y")) %>%
  select(Contributor.Name_x, Contributor.Name_y, id_x,  id_y, name_sim, link_tier)
```

```{r, eval=FALSE}
# review all linkable fuzzy pairs - see if changes need to be made to matching algorithm 
org_pairs %>%
  filter(link_tier != "no_link") %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".y" = "row_id"), suffix = c("_x", "_y")) %>%
  select(Contributor.Name_x, Contributor.Name_y, jaccard, name_sim, name_jw, link_tier) %>%
  arrange(desc(name_sim))
``` 


## Step 5: External bridge map

Use Digital Democracy's org concept mapping to surface additional links. 
This helps identify organizations with very different names that may not 
be caught by fuzzy matching. 

DD org concepts are looser than our links (e.g. all chapters of a union are linked under one 
org concept), so we don't want to automatically treat these as links.

This originally also used the equivalent of Names 2026 as a bridge; but that's no longer needed
since we are now using that directly in the entity resolution process. 

```{r}
# standardize external names to name_v5 format for matching
build_bridge_lookup <- function(mapping) {
  mapping %>%
    mutate(name_std = standardize_names(record_name) %>%
             fix_typos() %>%
             #remove_pac_info() %>%
             #remove_unit_info() %>%
             replace_business_text() %>%
             str_remove_all("\\s*\\(.*?\\)") %>%
             str_squish()) %>%
    select(name_std, canonical_name) %>%
    filter(!is.na(name_std), name_std != "", !is.na(canonical_name), canonical_name != "")
}

map1_lookup <- build_bridge_lookup(map1)

find_bridge_pairs <- function(orgs_df, lookup, source_label) {
  matched <- orgs_df %>%
    inner_join(lookup, by = c("name_v5" = "name_std")) %>%
    select(row_id, name_v5, canonical_name) %>%
    distinct(name_v5, canonical_name, .keep_all = TRUE) 

  if (nrow(matched) < 2) return(tibble(.x = integer(), .y = integer(),
                                        match_source = character(), canonical_name = character()))

  matched %>%
    rename(.x = row_id, name_v5_x = name_v5) %>%
    inner_join(matched %>% rename(.y = row_id, name_v5_y = name_v5), by = "canonical_name") %>%
    filter(.x < .y, name_v5_x != name_v5_y) %>%
    group_by(.x, .y) %>%
    mutate(
      n_canonicals = n_distinct(canonical_name),
      needs_review = n_canonicals > 1
    ) %>%
    ungroup() %>%
    mutate(match_source = source_label) %>%
    select(.x, .y, match_source, canonical_name, n_canonicals, needs_review)
}

```

```{r}
# appply bridge map to ALL orgs
bridge_map1 <- find_bridge_pairs(orgs, map1_lookup, "map1_bridge")
cat("Bridge pairs — map1:", nrow(bridge_map1), "\n")

bridge_map1 %>%
  mutate(.x = pmin(.x, .y), .y = pmax(.x, .y)) %>%
  distinct(.x, .y, canonical_name, .keep_all = TRUE) %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(name_x, name_y, canonical_name, n_canonicals) %>%
  arrange(name_x, name_y)
```


### combine bridge pairs

Note: upgrading fuzzy matches can be removed; this is mostly an artifact of 
the old entity-resolution process.

```{r}
# bridge pairs from map1 only
bridge_pairs <- bridge_map1 %>%
  group_by(.x, .y) %>%
  summarise(
    match_source   = paste(sort(unique(match_source)), collapse = "+"),
    canonical_name = paste(sort(unique(canonical_name)), collapse = " / "),
    .groups = "drop"
  )

# canonical ID per name_v5: if ANY contribution row for that name has an ID, record it
# used detect conflicting-ID pairs regardless of which row_id find_bridge_pairs picked.
# there are cases where name_v5 could have multiple contributor.ids
org_name_canonical_id <- orgs %>%
  filter(has_contributor_id) %>%
  group_by(name_v5) %>%
  summarise(canonical_id = names(sort(table(effective_Contributor.ID), decreasing = TRUE))[1],
            .groups = "drop")

# score bridge pairs on internal similarity
bridge_scored <- bridge_pairs %>%
  left_join(orgs %>% select(row_id, name_x = name_v5),
            by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = name_v5),
            by = c(".y" = "row_id")) %>%
  # join on name, not row_id in case a row was picked that didn't have a contributor.id
  left_join(org_name_canonical_id %>% rename(name_x = name_v5, id_x = canonical_id), by = "name_x") %>%
  left_join(org_name_canonical_id %>% rename(name_y = name_v5, id_y = canonical_id), by = "name_y") %>%
  mutate(
    has_id_conflict = coalesce(!is.na(id_x), FALSE) & coalesce(!is.na(id_y), FALSE) & coalesce(id_x != id_y, FALSE)
  )

# show conflicting-ID bridge pairs before filtering them out
#cat("\nBridge pairs with conflicting contributor IDs (currently excluded):\n")
# maybe change this so i can look at which are filtered out AFTER local units conflicts removed
bridge_id_conflicts <- bridge_scored %>%
  filter(has_id_conflict) %>%
  left_join(orgs %>% select(row_id, raw_name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, raw_name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(raw_name_x, id_x, raw_name_y, id_y, canonical_name) %>%
  arrange(canonical_name) 

bridge_scored <- bridge_scored %>%
  filter(!has_id_conflict) %>% # filters out id-conflicts
  mutate(
    name_jw     = stringsim(name_x, name_y, method = "jw", p = 0.1),
    name_cosine = bigram_cosine(name_x, name_y),
    name_sim    = (name_jw + name_cosine) / 2
  ) %>%
  select(-name_x, -name_y, -id_x, -id_y, -has_id_conflict)

# tag internal fuzzy paris thr bridge found— for match_source logging
internal_pair_keys <- org_pairs %>%
  filter(link_tier != "no_link") %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y)))

bridge_pair_keys <- bridge_scored %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  select(pair_key, bridge_source = match_source)

# upgrade internal pairs corroborated by a bridge
BRIDGE_UPGRADE_THRESHOLD <- 0.65

org_pairs <- org_pairs %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  left_join(bridge_pair_keys, by = "pair_key") %>%
  mutate(
    match_source = if_else(!is.na(bridge_source),
                           paste(match_source, bridge_source, sep = "+"),
                           match_source),
    link_tier = case_when(
      link_tier == "review" & !is.na(bridge_source) & name_sim >= BRIDGE_UPGRADE_THRESHOLD ~ "auto_link",
      TRUE ~ link_tier
    )
  ) %>%
  select(-pair_key, -bridge_source)

# bridge-only pairs
bridge_only <- bridge_scored %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  filter(!pair_key %in% internal_pair_keys$pair_key) %>%
  select(-pair_key) %>%
  mutate(link_tier = "review")  # bridge alone never auto-links

# flag different-chapter pairs in bridge (same logic as fuzzy, with CHAPTER_RE guard)
#CHAPTER_RE <- "\\b(LOCAL|LODGE|CHAPTER|COUNCIL|DISTRICT|UNIT)\\b|^[A-Z]{2,5}\\s+\\d"
CHAPTER_RE <- regex("\\b(LOCAL|LODGE|CHAPTER|COUNCIL|DISTRICT|UNIT|UNION|MEMBER DUES)\\b|\\b[A-Za-z]{2,4}\\s+\\d{3,4}\\b", ignore_case = TRUE)

bridge_only <- bridge_only %>%
  left_join(orgs %>% select(row_id, name_x = name_v5), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = name_v5), by = c(".y" = "row_id")) %>%
  mutate(
    nx_no_digits = str_remove_all(name_x, "\\d+") %>% str_squish(),
    ny_no_digits = str_remove_all(name_y, "\\d+") %>% str_squish(),
    diff_chapter = str_detect(name_x, "\\d") & str_detect(name_y, "\\d") &
                   nx_no_digits == ny_no_digits &
                   (str_detect(name_x, CHAPTER_RE) | str_detect(name_y, CHAPTER_RE))
  )

#cat("\nDifferent-chapter pairs flagged for removal (bridge):\n")
diff_chapter_bridge <- bridge_only %>%
  filter(diff_chapter) %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, Contributor.Name), by = c(".y" = "row_id"), suffix = c("_x", "_y")) %>%
  select(Contributor.Name_x, Contributor.Name_y, canonical_name) 

bridge_only <- bridge_only %>%
  filter(!diff_chapter) %>%
  select(-name_x, -name_y, -nx_no_digits, -ny_no_digits, -diff_chapter)

cat("Bridge-corroborated internal pairs:", sum(!is.na(bridge_pair_keys$bridge_source[match(internal_pair_keys$pair_key, bridge_pair_keys$pair_key)])), "\n")
cat("Bridge-only new pairs for review:", nrow(bridge_only), "\n")
```


#### data check (can skip)


```{r}
# review different chapter conflicts that were removed
diff_chapter_bridge
```



## Combine + cluster - definitive links

```{r}
# pairs joined through definitive links (not fuzzy matchin)
all_org_pairs <- bind_rows(
  id_auto_pairs %>% select(.x, .y, link_tier, match_source),
  exact_pairs,
)

# build one graph from all non-rejected pairs, treating review links as true
linkable_pairs <- all_org_pairs %>% filter(link_tier %in% c("auto_link", "review")) 

if (nrow(linkable_pairs) > 0) {
  g_orgs       <- graph_from_data_frame(linkable_pairs %>% select(.x, .y), directed = FALSE)
  org_clusters <- components(g_orgs)$membership

  org_cluster_raw <- tibble(row_id = as.integer(names(org_clusters)), cluster_num = org_clusters) %>%
    left_join(orgs %>% select(row_id, Contributor.Name, Contributor.ID, name_v5), by = "row_id")

  cluster_uuid_map <- org_cluster_raw %>%
    group_by(cluster_num) %>%
    summarise(raw_names = list(unique(Contributor.Name)), .groups = "drop") %>%
    rowwise() %>%
    mutate(
      result      = list(lookup_cluster_uuid(raw_names, nem_lookup)),
      entity_id   = result$entity_id,
      is_conflict = result$is_conflict,
      is_new      = result$is_new
    ) %>%
    ungroup() %>%
    select(cluster_num, entity_id, is_conflict, is_new)

  org_cluster_map <- org_cluster_raw %>%
    left_join(cluster_uuid_map, by = "cluster_num") %>%
    select(row_id, entity_id, is_conflict, is_new)
} else {
  org_cluster_map <- tibble(row_id = integer(), entity_id = character(),
                             is_conflict = logical(), is_new = logical())
}

orgs_singleton_lookup <- orgs %>%
  filter(!row_id %in% org_cluster_map$row_id) %>%
  distinct(Contributor.Name) %>%
  mutate(raw_name_key = toupper(Contributor.Name)) %>%
  left_join(nem_lookup %>% select(raw_name, entity_uuid), by = c("raw_name_key" = "raw_name")) %>%
  mutate(
    is_new    = is.na(entity_uuid),
    entity_id = coalesce(entity_uuid, UUIDgenerate(n = n()))
  ) %>%
  select(Contributor.Name, entity_id, is_new)

orgs_singletons <- orgs %>%
  filter(!row_id %in% org_cluster_map$row_id) %>%
  left_join(orgs_singleton_lookup, by = "Contributor.Name") %>%
  mutate(is_conflict = FALSE) %>%
  select(row_id, entity_id, is_new, is_conflict)

orgs_resolved <- orgs %>%
  left_join(bind_rows(org_cluster_map, orgs_singletons), by = "row_id")

```


## Org name variant review queue

For each entity with more than one distinct name, output the names that are not already
tracked so they can be added to the name variant list. `entity_status` indicates whether
this is a new name variant of a known entity or a new entity group entirely. 

```{r}
# grouping for name variant review queue: includes fuzzy + bridge in addition to id and exact
all_org_pairs_review <- bind_rows(
  id_auto_pairs %>% select(.x, .y, link_tier, match_source),
  exact_pairs   %>% select(.x, .y, link_tier, match_source),
  org_pairs     %>% filter(link_tier != "no_link") %>% select(.x, .y, link_tier, match_source),
  bridge_only   %>% select(.x, .y, link_tier, match_source)
)

# drop pairs where both names are confirmed as DIFFERENT entities in name_entity_mapping
# uses nem_lookup so UUIDs are stable; names w/out a canonical name stay NA 
# the all_confirmed mechanism handles those groups separately
confirmed_nem <- nem_lookup %>% filter(confirmed == TRUE) %>% select(raw_name, entity_uuid) # all of these should have confirmed = TRUE, but just in case
orgs_known_uuid <- orgs %>%
  transmute(row_id,
            known_uuid = confirmed_nem$entity_uuid[match(toupper(Contributor.Name), confirmed_nem$raw_name)])

n_before_split_filter <- nrow(all_org_pairs_review)
# remove pairs that matched to different uuids (if name is already in map, it means it was verified manually)
all_org_pairs_review <- all_org_pairs_review %>%
  left_join(orgs_known_uuid %>% rename(uuid_x = known_uuid), by = c(".x" = "row_id")) %>%
  left_join(orgs_known_uuid %>% rename(uuid_y = known_uuid), by = c(".y" = "row_id")) %>%
  filter(is.na(uuid_x) | is.na(uuid_y) | uuid_x == uuid_y) %>% # only include pairs where at least one uuid is known OR they have the same uuid 
  select(.x, .y, link_tier, match_source)
cat(sprintf("Split-entity filter: removed %d pairs where both names had confirmed different UUIDs\n",
            n_before_split_filter - nrow(all_org_pairs_review)))

# create graph now that false matches are removed
if (nrow(all_org_pairs_review) > 0) {
  g_org_review     <- graph_from_data_frame(all_org_pairs_review %>% select(.x, .y), directed = FALSE)
  org_review_comps <- components(g_org_review)$membership
  org_review_group_map <- tibble(
    row_id          = as.integer(names(org_review_comps)),
    review_group_id = paste0("org_rg_", org_review_comps)
  )
} else {
  org_review_group_map <- tibble(row_id = integer(), review_group_id = character())
}

# pair filter above removes direct edges between confirmed-different-entity
# names, but transitive paths through unconfirmed names can still group different entities together
# for any component that contains multiple distinct confirmed UUIDs, re-assign confirmed names to
# UUID-based subgroups - unconfirmed names keep the original review_group_id and remain in the queue
orgs_confirmed_uuid <- orgs %>%
  transmute(row_id, raw_name = toupper(Contributor.Name)) %>%
  left_join(confirmed_nem, by = "raw_name") %>%
  select(row_id, confirmed_uuid = entity_uuid)

org_review_group_map <- org_review_group_map %>%
  left_join(orgs_confirmed_uuid, by = "row_id") %>%
  group_by(review_group_id) %>%
  mutate(n_distinct_uuid = n_distinct(confirmed_uuid[!is.na(confirmed_uuid)])) %>%
  ungroup() %>%
  mutate(
    review_group_id = case_when(
      n_distinct_uuid > 1 & !is.na(confirmed_uuid) ~ paste0("org_rg_confirmed_", confirmed_uuid),
      TRUE                                           ~ review_group_id 
    )
  ) %>%
  select(row_id, review_group_id)

cat(sprintf("Post-hoc UUID split: %d groups contain confirmed names re-assigned to UUID-based subgroups\n",
            sum(grepl("^org_rg_confirmed_", org_review_group_map$review_group_id))))

# flag review groups that contain names with conflicting non-NA Contributor.IDs (just a flag for reference in review queue)
org_cid <- orgs %>%
  select(row_id, cid = effective_Contributor.ID) %>%
  mutate(cid = na_if(str_trim(as.character(cid)), ""))

conflicting_rg <- org_review_group_map %>%
  left_join(org_cid, by = "row_id") %>%
  filter(!is.na(cid)) %>%
  group_by(review_group_id) %>%
  filter(n_distinct(cid) > 1) %>%
  pull(review_group_id) %>%
  unique()

org_review_singletons <- orgs %>%
  filter(!row_id %in% org_review_group_map$row_id) %>%
  transmute(row_id, review_group_id = paste0("org_rg_solo_", row_id))

org_review_groups <- bind_rows(org_review_group_map, org_review_singletons)

```

### Apply qualifying PAC fitler

For each review group, sum contributions to relevant recipients treating all links as true.
Only keep orgs that contributed to a qualifying PAC (a PAC whose tentative group gave $5k+
to a tracked race/prop or a traversed PAC)

```{r}
# candidate committees and ballot measure committees
tracked_race_committee_ids <- contributions %>%
  filter(!coalesce(is_pac_contribution, FALSE)) %>%
  pull(Recipient.Committee.ID) %>%
  as.character() %>%
  unique()

# committees that came up in the 0802b PAC graph (no threshold considered);
# exclude candidate and federal PACs — their contributors don't require individual classification
all_universe_pac_ids <- pac_universe %>%
  mutate(
    is_candidate_pac = if ("is_candidate_pac" %in% names(pac_universe)) as.logical(is_candidate_pac) else FALSE,
    is_federal_pac   = if ("is_federal_pac"   %in% names(pac_universe)) as.logical(is_federal_pac)   else FALSE
  ) %>%
  filter(!coalesce(is_candidate_pac, FALSE), !coalesce(is_federal_pac, FALSE)) %>%
  pull(pac_id) %>%
  as.character()

# combine 'direct' committees and PAC graph committees
relevant_recipient_ids <- unique(c(tracked_race_committee_ids, all_universe_pac_ids))
length(relevant_recipient_ids)

# id_map maps XREF_FILER_ID (Contributor.ID) -> FILER_ID (Recipinet.Committee.ID)
id_map_early <- read_tsv("08_inputs/FILERNAME_CD.tsv", col_types = cols(.default = "c")) %>%
  mutate(FILER_ID = as.character(FILER_ID), XREF_FILER_ID = as.character(XREF_FILER_ID)) %>%
  filter(!is.na(FILER_ID), !is.na(XREF_FILER_ID)) %>%
  distinct(XREF_FILER_ID, FILER_ID)

```

```{r}
# data check for id map
unique_rcids <- contributions %>%
  pull(Recipient.Committee.ID) %>%
  as.character() %>%
  unique()

all_filer_ids <- unique(id_map_early$FILER_ID)

# (a) Recipient.Committee.IDs that don't appear as any FILER_ID in FILERNAME_CD
not_a_filer_id <- unique_rcids[!unique_rcids %in% all_filer_ids]
cat(sprintf("(a) %d Recipient.Committee.IDs not found as any FILER_ID in FILERNAME_CD\n",
            length(not_a_filer_id)))
if (length(not_a_filer_id) > 0) print(head(not_a_filer_id, 20))

# (b) Recipient.Committee.IDs that appear as an XREF_FILER_ID where XREF != FILER
non_canonical_rcids <- id_map_early %>%
  filter(XREF_FILER_ID %in% unique_rcids, XREF_FILER_ID != FILER_ID) %>%
  distinct(XREF_FILER_ID, FILER_ID)
cat(sprintf("(b) %d Recipient.Committee.IDs are XREF_FILER_IDs with a different canonical FILER_ID\n",
            nrow(non_canonical_rcids)))
if (nrow(non_canonical_rcids) > 0) print(non_canonical_rcids)
```


```{r}
orgs_grouped <- orgs %>%
  left_join(org_review_groups %>% select(row_id, review_group_id), by = "row_id") %>%
  select(-any_of(c("FILER_ID", "FILER_TYPE"))) %>%
  left_join(id_map_early, by = c("effective_Contributor.ID" = "XREF_FILER_ID")) %>%
  mutate(FILER_ID = as.character(FILER_ID))

not_qualifying_review_groups <- orgs_grouped %>%
  mutate(rc_id = as.character(Recipient.Committee.ID)) %>%
  filter(rc_id %in% relevant_recipient_ids) %>% 
  group_by(review_group_id, rc_id) %>% # is this grouping right??? b/c what about entities w/ MISSING id -> that's okay, recipient should never be missing
  summarise(total = sum(Amount, na.rm = TRUE), .groups = "drop") %>%
  filter(total < 5000) %>%
  pull(review_group_id) %>% # pull all orgs that are part of a tentatively-linked entity that gave $5k+ to a single recipient. 
  unique()
```

```{r}
length(not_qualifying_review_groups)
below <- orgs_grouped %>% # orgs w/ review_group_ids 
  filter(review_group_id %in% not_qualifying_review_groups)
```

```{r}
below %>% group_by(review_group_id,Recipient.Committee.ID) %>% summarise(total = sum(Amount)) %>% arrange(desc(total))
```
reverse doesn't work -> expected? 

```{r}
orgs_grouped %>% filter(review_group_id == "org_rg_383") %>% group_by(Recipient.Committee.ID) %>% summarise(total = sum(Amount) )
```

why many-to-many for join in paired_qualifying? 

```{r}
# map review_group_id -> all raw contributor names in that group; expands beyond the
# single representative row_id chosen during fuzzy/bridge matching so that all
# contribution rows for each name are included in the $5k threshold check.
org_group_name_index <- org_review_group_map %>%
  left_join(
    orgs %>%
      mutate(raw_name_key = str_squish(toupper(Contributor.Name))) %>%
      select(row_id, raw_name_key),
    by = "row_id"
  ) %>%
  distinct(review_group_id, raw_name_key) # all names associated w/ each review group 

# paired groups: expand ALL contribution rows for names in the group via raw name
paired_qualifying <- orgs %>%
  mutate(
    raw_name_key = str_squish(toupper(Contributor.Name)),
    rc_id        = as.character(Recipient.Committee.ID)
  ) %>%
  inner_join(org_group_name_index, by = "raw_name_key", relationship = "many-to-one") %>%
  filter(rc_id %in% relevant_recipient_ids)

# singletons: orgs not in any paired group, kept under their own solo review_group_id
solo_qualifying <- orgs %>%
  mutate(
    raw_name_key = str_squish(toupper(Contributor.Name)),
    rc_id        = as.character(Recipient.Committee.ID)
  ) %>%
  filter(!raw_name_key %in% org_group_name_index$raw_name_key) %>%
  left_join(org_review_singletons %>% select(row_id, review_group_id), by = "row_id") %>%
  filter(rc_id %in% relevant_recipient_ids)

qualifying_expanded <- bind_rows(paired_qualifying, solo_qualifying)

qualifying_review_groups <- qualifying_expanded %>%
  mutate(
    race_recipient_key = case_when(
      coalesce(is_pac_contribution, FALSE)          ~ rc_id,
      !is.na(Ballot.Measure) & Ballot.Measure != "" ~ Ballot.Measure,
      !is.na(Office) & Office != ""                 ~ paste(Recipient.Name, Office, sep = " | "),
      TRUE                                          ~ Recipient.Name
    )
  ) %>%
  group_by(review_group_id, race_recipient_key) %>%
  summarise(total = sum(Amount, na.rm = TRUE), .groups = "drop") %>%
  filter(total >= 5000) %>%
  pull(review_group_id) %>%
  unique()

qualifying_pac_filer_ids <- qualifying_expanded %>%
  select(-any_of(c("FILER_ID", "FILER_TYPE"))) %>% # in case it was added earlier
  filter(review_group_id %in% qualifying_review_groups) %>%
  left_join(id_map_early, by = c("effective_Contributor.ID" = "XREF_FILER_ID")) %>%
  mutate(FILER_ID = as.character(FILER_ID)) %>%
  filter(!is.na(FILER_ID), FILER_ID != "", FILER_ID %in% all_universe_pac_ids) %>% # pac traversal ids
  pull(FILER_ID) %>%
  unique()

cat(sprintf(
  "%d of %d traversed PAC(s) qualify (tentatively gave $5k+ to a relevant recipient)\n",
  length(qualifying_pac_filer_ids), length(unique(known_pacs$pac_id))
))
```

- 777 / 1068 traversed PACs marked as qualifying; didn't change after i moved the expansion up here I don't think 
DOES NOT INCLUDE the ballot measure and candidate PACS. 

```{r}
# data checks
qualifying_expanded %>% group_by(row_id) %>% filter(n() > 1)
orgs %>% filter(!(row_id %in% qualifying_expanded$row_id))
```


NOTE: the entire review group will qualify even if only some of its name variants were the ones that contributed $5k+ to a race.
In other words, an org may be part of a review group even if that specific name variant didn't contribute $5k+ to a race, but other name variants did. 

WAIT! are tracked_race_committee_ids counted??? 

where does orgs_resolved come from? why is this included? only definitive links, but not relevant here
do entity_id, canonical_name etc ever vary in the summarise step? 



```{r}
# flag which linkage methods connect it to others in its group
org_row_sources <- bind_rows(
  all_org_pairs_review %>% select(row_id = .x, match_source),
  all_org_pairs_review %>% select(row_id = .y, match_source)
) %>%
  group_by(row_id) %>%
  summarise(link_methods = paste(sort(unique(match_source)), collapse = " | "), .groups = "drop")

org_review_names_raw <- orgs %>%
  left_join(org_review_groups, by = "row_id") %>%
  left_join(orgs_resolved %>% select(row_id, entity_id), by = "row_id") %>%
  left_join(org_row_sources, by = "row_id") %>%
  mutate(
    exported_name  = toupper(Contributor.Name),
    raw_name_key   = exported_name,
    contributor_id = na_if(str_trim(as.character(effective_Contributor.ID)), "")
  ) %>%
  left_join(nem_lookup %>% select(raw_name, confirmed, canonical_name),
            by = c("raw_name_key" = "raw_name")) %>%
  mutate(confirmed = coalesce(confirmed, FALSE)) %>%
  group_by(review_group_id) %>%
  filter(n_distinct(exported_name) > 1) %>%
  ungroup()

# entity_id should not vary within (review_group_id, exported_name, contributor_id)
entity_id_check <- org_review_names_raw %>%
  group_by(review_group_id, exported_name, contributor_id) %>%
  summarise(n_entity_ids = n_distinct(entity_id[!is.na(entity_id)]), .groups = "drop") %>%
  filter(n_entity_ids > 1)
if (nrow(entity_id_check) > 0) {
  stop(sprintf(
    "entity_id varies within %d (review_group_id, exported_name, contributor_id) group(s) — check orgs_resolved",
    nrow(entity_id_check)
  ))
}

org_review_names <- org_review_names_raw %>%
  group_by(review_group_id, exported_name, contributor_id) %>%
  summarise(
    entity_id       = first(na.omit(entity_id)),
    canonical_name  = first(na.omit(canonical_name)),
    confirmed       = all(confirmed),
    link_methods    = paste(sort(unique(link_methods[!is.na(link_methods)])), collapse = " | "),
    n_contributions = n(),
    total_amount    = sum(Amount, na.rm = TRUE),
    city  = paste(sort(unique(Contributor.City[ !is.na(Contributor.City)  & Contributor.City  != ""])), collapse = " | "),
    state = paste(sort(unique(Contributor.State[!is.na(Contributor.State) & Contributor.State != ""])), collapse = " | "),
    zip   = paste(sort(unique(Contributor.Zip.Code[!is.na(Contributor.Zip.Code) & Contributor.Zip.Code != ""])), collapse = " | "),
    .groups = "drop"
  )

rm(org_review_names_raw)

# look at review group as a whole
# merge_conflict: group contains confirmed names mapping to different canonical entities
org_review_group_status <- org_review_names %>%
  group_by(review_group_id) %>%
  summarise(
    entity_status = case_when(
      all(confirmed)                                                        ~ "all_confirmed",
      n_distinct(canonical_name[confirmed & !is.na(canonical_name)]) > 1   ~ "merge_conflict",
      any(confirmed)                                                        ~ "adding_to_existing",
      TRUE                                                                  ~ "new_entity"
    ),
    conflicting_contributor_ids = any(review_group_id %in% conflicting_rg),
    .groups = "drop"
  )

# A group qualifies for review only if more than one distinct name variant donated to
# the same single recipient AND their combined total is >= $5k.
org_threshold_expanded <- orgs %>%
  filter(
    !coalesce(is_pac_contribution, FALSE) | # direct contributor
    as.character(Recipient.Committee.ID) %in% qualifying_pac_filer_ids # OR a contributor to a $5k+ PAC
  ) %>%
  mutate(
    raw_name_key  = str_squish(toupper(Contributor.Name)),
    exported_name = toupper(Contributor.Name)
  ) %>%
  add_recipient_key() %>%
  inner_join(org_group_name_index, by = "raw_name_key", relationship = "many-to-one") # apply review_group_ids to ALL orgs with name in the review group


# does this work as intended? 
org_review_group_threshold <- org_threshold_expanded %>%
  group_by(review_group_id, recipient_key) %>% # recipient.committee.id
  summarise(
    recipient_total = sum(Amount, na.rm = TRUE),
    n_name_variants = n_distinct(exported_name),
    .groups = "drop"
  ) %>%
  filter(n_name_variants > 1) %>% # more than one name variant contributed to single committee
  group_by(review_group_id) %>%
  summarise(max_single_recipient = max(recipient_total, na.rm = TRUE), .groups = "drop")

# Per-name flag: TRUE if this exported_name donated to any recipient that ALSO
# received from another name variant in the same group. FALSE means the name is
# in the group but its contributions went to different recipients than the others.
same_recipient_name_flag <- org_threshold_expanded %>%
  group_by(review_group_id, recipient_key) %>%
  mutate(n_names_to_recipient = n_distinct(exported_name)) %>%
  filter(n_names_to_recipient > 1) %>%
  ungroup() %>%
  distinct(review_group_id, exported_name) %>%
  mutate(same_recipient_variant = TRUE)

# suppress groups where every within-group name pair already exists in the no-link registry.
# if theres one new name variant in the group, names will be paired again which is annoying but want to compare new names against all possible variants
org_registry_pair_keys <- no_link_registry %>%
  filter(entity_type == "org") %>%
  mutate(pair_key = paste(name_a, name_b, sep = "|||")) %>%
  pull(pair_key)

org_group_name_pairs <- org_review_names %>%
  distinct(review_group_id, name = toupper(exported_name))

org_fully_reviewed_groups <- org_group_name_pairs %>%
  inner_join(org_group_name_pairs, by = "review_group_id", relationship = "many-to-many") %>%
  filter(name.x < name.y) %>%
  mutate(pair_key = paste(name.x, name.y, sep = "|||"),
         reviewed = pair_key %in% org_registry_pair_keys) %>%
  group_by(review_group_id) %>%
  summarise(all_reviewed = all(reviewed), .groups = "drop") %>%
  filter(all_reviewed) %>%
  pull(review_group_id)

org_name_variant_queue <- org_review_names %>%
  left_join(org_review_group_status, by = "review_group_id") %>%
  left_join(org_review_group_threshold, by = "review_group_id") %>%
  left_join(same_recipient_name_flag, by = c("review_group_id", "exported_name")) %>%
  mutate(same_recipient_variant = coalesce(same_recipient_variant, FALSE)) %>%
  filter(
    entity_status != "all_confirmed",
    !review_group_id %in% org_fully_reviewed_groups,
    max_single_recipient >= 5000
  ) %>%
  mutate(original_group_id = review_group_id) %>% # create a copy of review_group_id that will be preserved
  select(entity_name = exported_name, review_group_id, original_group_id,
         entity_id, canonical_name, entity_status,
         conflicting_contributor_ids, confirmed, link_methods, Contributor.ID = contributor_id,
         n_contributions, total_amount, max_single_recipient, same_recipient_variant, city, state, zip) %>%
  arrange(desc(conflicting_contributor_ids), review_group_id)

```

```{r}
today <- format(Sys.Date(), "%m-%d-%y")
ORG_REVIEW_OUTPUT_PATH <- paste0("08_outputs/org_name_variant_review_queue_", today, ".csv")
write_csv(org_name_variant_queue, ORG_REVIEW_OUTPUT_PATH)
cat("Org name variant review queue:", nrow(org_name_variant_queue), "names across",
    n_distinct(org_name_variant_queue$review_group_id), "groups\n")
```


# INDIVIDUALS:

Entity resolution for individuals.Entity resolution (in the context of categorizations) is used to 
determine who meets the $5k threshold to be categorized. For mapping name variants, we want to surface name variants WITHIN
recipient committees, since aggregation won't happen across recipient committees.

Process: 
1. Filter the individuals data to only include contributors to PACs that meet the $5k threshold. 
2. Parse names into first, last and if applicable, middle, prefix and suffix
3. Build record linkage algorithm that scores how similar individuals are based on their name,
   employer, occupation, city, etc. 
4. Create name variant review queue for individuals - contributors that are likely the same person,
   but contributed under slightly different names.
5. Create a "disambiguation" review queue that flags contributors with the exact same name (meaning
   they would normally be aggregated) who were flagged as different people by the record linkage
   algorithm.


**update name of file that's read in here**

```{r}
#  read in latest version of the data w/ individuals names parsed
full_names_parsed <- read.csv("08_outputs/individuals_name_parsed_all_contributors_09-08-26.csv") 
```

```{r}
# filter so we're only doing entity resolution for individuals who contributed directly to one 
# of our candidate committees / props OR contributed to a PAC/entity that meets the $5k threshold
individuals <- contributions %>%
  filter(entity_type == "individual",
         !coalesce(is_pac_contribution, FALSE) |
         as.character(Recipient.Committee.ID) %in% qualifying_pac_filer_ids) %>%
  left_join(full_names_parsed %>% select(standardized_name, last_name = last, first_name = first),
            by = "standardized_name") %>%
  mutate(
    original_row_id = row_id,
    row_id          = row_number(),
    first_initial   = substr(first_name, 1, 1),
    last_meta       = safe_metaphone(last_name)
  )
```

```{r}
# parse new names in the data 
missing_names <- individuals %>%
  filter(is.na(last_name)) %>%
  distinct(standardized_name)
if (nrow(missing_names) > 0) {
  cat(sprintf("Parsing %d name(s) not found in pre-built CSV\n", nrow(missing_names)))
  patched <- parse_names_df(missing_names$standardized_name) %>%
    mutate(standardized_name = missing_names$standardized_name,
           last_name = last, first_name = first) %>%
    select(standardized_name, last_name, first_name)
  individuals <- individuals %>%
    left_join(patched, by = "standardized_name", suffix = c("", "_patch")) %>%
    mutate(
      last_name     = coalesce(last_name,  last_name_patch),
      first_name    = coalesce(first_name, first_name_patch),
      first_initial = substr(coalesce(first_name, ""), 1, 1),
      last_meta     = safe_metaphone(last_name)
    ) %>%
    select(-last_name_patch, -first_name_patch)

  new_rows <- patched %>% rename(last = last_name, first = first_name)
  full_names_parsed <- bind_rows(full_names_parsed, new_rows)
  write.csv(full_names_parsed,
            "08_outputs/individuals_name_parsed_all_contributors.csv",
            row.names = FALSE)
  cat(sprintf("Updated full_names_parsed with %d new name(s)\n", nrow(missing_names)))
}

last_name_freq  <- full_names_parsed %>% count(last,  name = "last_name_freq")
first_name_freq <- full_names_parsed %>% count(first, name = "first_name_freq")
```


```{r}
# assign entity id — one UUID per Normalized name, looked up from registry if known
# adds entity_id to existing individuals (which already has last_meta, original_row_id, etc.)
# NOTE: this early entity_id is used only for the disambiguation queue;
# it is dropped at individuals_resolved where the cluster-based id wins.
indiv_name_lookup <- individuals %>%
  distinct(Contributor.Name, Normalized) %>%
  mutate(raw_name_key = toupper(Contributor.Name)) %>%
  left_join(nem_lookup %>% select(raw_name, entity_uuid), by = c("raw_name_key" = "raw_name"))

indiv_entity_key <- indiv_name_lookup %>%
  group_by(Normalized) %>%
  summarise(entity_id = coalesce(first(na.omit(entity_uuid)), UUIDgenerate()), .groups = "drop")

individuals <- individuals %>%
  left_join(indiv_entity_key, by = "Normalized")
```


### record linkage algorithm

Block on Recipient.Committee.ID so we only compare pairs within recipient committees.
Pairs are compared if:
- they have the same last_meta and zip code (within the same recipient committee)
- they have the same last_meta and first initial (within the same recipient committee)

Note: 
- Doing this record linkage algorithm across the entire dataset was too expensive 
- Limitation is that this will limit our ability to link name variants of individuals across committees
  to make sure they are being categorized consistently. But won't be an issue for contributions that used
  the exact same name across contributions to different committees. 

Threshold logic: 
- Pairs are not linked if the first names are too different: this was added because many linked pairs were husband/wife
or family members where the two people had the same last name, same city/zip, and first name that started with
the same letter (but were clearly different people).
- A stricter threshold is used for people with names that are more common in the data. e.g. Michael Brown should
only be linked to another Michael Brown if info is very similar. Lower threshold used for people with less common names. 
- A stricter threshold is also used for pairs where the employer/occupation are uninformative. E.g. so many people have
'retired' or 'not employed' as their employer, which is not very strong evidence to link people. 

Thresholds were adjusted and additional requirements were added after doing several rounds of review of linked pairs.
The original goal was to have stricter matching requirements in order to err on the side of not linking contributors
if there was uncertainty, but this means it may not link people like Bill Bloomfield who has very little info that
ties his records together apart from his name (uninformative employer/occupation, different zip codes across contributions).

### Scoring function summary

Each candidate pair is scored using a weighted combination of similarities between their fields, 
normalized by the sum of active weights so the scale is preserved when fields drop out:

| Field | Comparator | Base weight |
|---|---|---|
| Last name | Jaro-Winkler | 0.30 |
| First name | Jaro-Winkler | 0.25 |
| Employer | Cosine bigram similarity | 0.25 (drops to 0 if both sides uninformative) |
| Occupation | Exact match | 0.10 (drops to 0 if both sides uninformative) |
| Zip code | Exact match | 0.07 |
| City | Jaro-Winkler | 0.03 |

**Uninformative values**: blank, NA, UNKNOWN, RETIRED, NONE, SELF EMPLOYED. 
When one side of a pair has an uninformative employer/occupation and the other does not, the similarity for that field is set to 0 (penalizing the pair for asymmetry). 
When both sides are uninformative, the field weight drops to 0 and the field is excluded from scoring entirely. 
`any_context_dropped` is TRUE when either field (employer or occupation) is excluded — this triggers a stricter match threshold.

**Threshold tiers** (applied after scoring):

| Condition | Threshold |
|---|---|
| Common name + any context dropped | `COMMON_NAME_STRICT` (0.97) |
| Common name | `COMMON_NAME_THRESHOLD` (0.93) |
| Any context dropped | `STRICT_THRESHOLD` (0.94) |
| Default | `MATCH_THRESHOLD` (0.90) |

Common name = last name appears ≥ 3 times AND first name appears ≥ 3 times in the data. 
Pairs are always `no_link` if first name similarity < 0.85, or if there is a city conflict combined with any dropped context field.

```{r}
# Constants
UNINFORMATIVE_VALS        <- c("UNKNOWN", "RETIRED", "NONE", "SELF EMPLOYED") # uninformative values of employer/occupation (after these variables are processed to standardize N/A types of values)
FIRST_NAME_MIN            <- 0.85 
MATCH_THRESHOLD           <- 0.90
STRICT_THRESHOLD          <- 0.94  # both employer & occupation uninformative
COMMON_NAME_THRESHOLD     <- 0.93  # common name, employer informative
COMMON_NAME_STRICT        <- 0.97  # common name + both uninformative
UNCERTAIN_LOW             <- 0.75
EMPLOYER_CONFLICT_SIM     <- 0.70  # employer similarity below this = conflict
RARE_NAME_UNCERTAIN_LOW <- 0.65 
COMMON_LAST_MIN           <- 3
COMMON_FIRST_MIN          <- 3
RARE_NAME_UNCERTAIN_LOW   <- 0.65
```

```{r}
# scoring function
compare_and_score_indiv <- function(pairs_obj, df) {
  if (nrow(pairs_obj) == 0) return(tibble())
  pairs_obj %>%
    compare_pairs(
      on = c("last_name", "first_name", "zip_code_processed",
             "processed_employer_name", "processed_occupation", "standardized_city"),
      default_comparator = cmp_jarowinkler(),
      comparators = list(
        processed_employer_name = function(a, b) stringsim(a, b, method = "cosine", q = 2),
        zip_code_processed      = cmp_identical(),
        processed_occupation    = cmp_identical()
      )
    ) %>%
    as_tibble() %>%
    
    # join raw employer/occupation values
    left_join(df %>% select(row_id, emp_a = processed_employer_name, occ_a = processed_occupation,
                             city_a = standardized_city),
              by = c(".x" = "row_id")) %>%
    left_join(df %>% select(row_id, emp_b = processed_employer_name, occ_b = processed_occupation,
                             city_b = standardized_city),
              by = c(".y" = "row_id")) %>%
    mutate(
      emp_a_uninf = emp_a %in% UNINFORMATIVE_VALS | is.na(emp_a) | emp_a == "",
      emp_b_uninf = emp_b %in% UNINFORMATIVE_VALS | is.na(emp_b) | emp_b == "",
      occ_a_uninf = occ_a %in% UNINFORMATIVE_VALS | is.na(occ_a) | occ_a == "",
      occ_b_uninf = occ_b %in% UNINFORMATIVE_VALS | is.na(occ_b) | occ_b == "",
      
      w_emp  = if_else(emp_a_uninf & emp_b_uninf, 0,    0.25),
      w_occ  = if_else(occ_a_uninf & occ_b_uninf, 0,    0.10),
      w_tot  = 0.30 + 0.25 + w_emp + w_occ + 0.07 + 0.03,
      emp_sim = if_else(emp_a_uninf | emp_b_uninf, 0, coalesce(processed_employer_name, 0)),
      occ_sim = if_else(occ_a_uninf | occ_b_uninf, 0, coalesce(processed_occupation, 0)),
      score  = (0.30 * coalesce(last_name, 0) +
                0.25 * coalesce(first_name, 0) +
                w_emp * emp_sim +
                w_occ * occ_sim +
                0.07 * coalesce(zip_code_processed, 0) +
                0.03 * coalesce(standardized_city, 0)) / w_tot,
      
      # flag cases where employer and occupation are uninformative (e.g. not employed)
      # don't want uninformative values to increase similarity between pairs
      both_uninformative  = (emp_a_uninf & emp_b_uninf) & (occ_a_uninf & occ_b_uninf),
      # TRUE when either contextual field drops from scoring (both emp uninf OR both occ uninf)
      any_context_dropped = (emp_a_uninf & emp_b_uninf) | (occ_a_uninf & occ_b_uninf),
    
      
      # flag cases where one employer is the substring of the other employer in the pair 
      employer_substring = !emp_a_uninf & !emp_b_uninf & mapply(
        function(a, b) {
          if (is.na(a) || is.na(b) || a == "" || b == "") return(FALSE)
          a <- toupper(a); b <- toupper(b)
          grepl(b, a, fixed = TRUE) || grepl(a, b, fixed = TRUE)
        },
        emp_a, emp_b,
        USE.NAMES = FALSE
      ),
    
      employer_conflict  = !emp_a_uninf & !emp_b_uninf &
                           !employer_substring &
                           coalesce(processed_employer_name, 0) < EMPLOYER_CONFLICT_SIM,
    
      # city conflict: both cities are known and different (JW < 0.85)
      city_conflict = (!is.na(city_a) & city_a != "") &
                      (!is.na(city_b) & city_b != "") &
                      coalesce(standardized_city, 1) < 0.85
    ) %>%
    select(-occ_a, -occ_b,
           -emp_a_uninf, -emp_b_uninf, -occ_a_uninf, -occ_b_uninf,
           -w_emp, -w_occ, -w_tot, -emp_sim, -occ_sim,
           -city_a, -city_b)
}
```

### blocking and scoring

```{r}
# pair_blocking by Recipient.Committee.ID × last_meta

INDIV_PAIRS_TMP <- "08_outputs/indiv_pairs_tmp.csv"
if (file.exists(INDIV_PAIRS_TMP)) file.remove(INDIV_PAIRS_TMP)
first_written <- FALSE

# split into a named list of dataframes — one per (committee × last_meta) group.
indiv_filtered <- individuals %>%
  filter(!is.na(last_meta), last_meta != "", !is.na(Recipient.Committee.ID))

# for hyphenated last names (e.g. "SMITH-JONES"), add extra rows with each component's
# metaphone
hyphen_rows <- indiv_filtered %>%
  filter(str_detect(coalesce(last_name, ""), "-")) %>%
  rowwise() %>%
  mutate(component_metas = list({
    parts <- str_split(last_name, "-")[[1]]
    metas <- unique(sapply(parts, safe_metaphone))
    setdiff(metas[metas != "" & !is.na(metas)], last_meta)
  })) %>%
  ungroup() %>%
  filter(map_lgl(component_metas, ~ length(.x) > 0)) %>%
  unnest(component_metas) %>%
  mutate(last_meta = component_metas) %>%
  select(-component_metas)

indiv_filtered <- bind_rows(indiv_filtered, hyphen_rows) %>%
  mutate(.group_key = paste0(Recipient.Committee.ID, "||", last_meta))

individual_chunks <- split(indiv_filtered, indiv_filtered$.group_key)
rm(indiv_filtered, hyphen_rows)

n_chunks <- length(individual_chunks)
cat(sprintf("Processing %d (committee × last_meta) chunks...\n", n_chunks))

# flush to disk every BATCH_FLUSH chunks with results
# instead of opening the file on every iteration.
BATCH_FLUSH  <- 500L
batch_buf    <- vector("list", BATCH_FLUSH)
batch_i      <- 0L

for (i in seq_along(individual_chunks)) {
  chunk <- individual_chunks[[i]]
  if (nrow(chunk) < 2) next

  global_ids <- chunk$row_id
  chunk <- chunk %>% mutate(row_id = row_number())

  # pairs based on zip code and last name
  pairs_zip <- tryCatch(
    pair_blocking(chunk, on = c("last_meta", "zip_code_processed"), deduplication = TRUE),
    error = function(e) NULL
  )
  
  # pairs based on first initial and last name
  pairs_init <- tryCatch(
    pair_blocking(chunk, on = c("last_meta", "first_initial"), deduplication = TRUE),
    error = function(e) NULL
  )

  # score the pairs that were identified
  chunk_pairs <- bind_rows(
    if (!is.null(pairs_zip))  compare_and_score_indiv(pairs_zip,  chunk) else tibble(),
    if (!is.null(pairs_init)) compare_and_score_indiv(pairs_init, chunk) else tibble()
  )

  if (nrow(chunk_pairs) > 0) {
    chunk_pairs <- chunk_pairs %>% mutate(.x = global_ids[.x], .y = global_ids[.y])
    batch_i <- batch_i + 1L
    batch_buf[[batch_i]] <- chunk_pairs

    if (batch_i >= BATCH_FLUSH) {
      write_csv(bind_rows(batch_buf[seq_len(batch_i)]), INDIV_PAIRS_TMP, append = first_written)
      first_written <- TRUE
      batch_buf <- vector("list", BATCH_FLUSH)
      batch_i   <- 0L
    }
  }

  if (i %% 5000L == 0L) {
    cat(sprintf("  %d / %d chunks\n", i, n_chunks))
    gc(verbose = FALSE)
  }
}

# flush remaining buffer
if (batch_i > 0L) {
  write_csv(bind_rows(batch_buf[seq_len(batch_i)]), INDIV_PAIRS_TMP, append = first_written)
}
rm(individual_chunks, batch_buf)
gc(verbose = FALSE)
cat("Done. Reading pairs from disk...\n")

indiv_pairs <- if (file.exists(INDIV_PAIRS_TMP) && file.size(INDIV_PAIRS_TMP) > 0) {
  read_csv(INDIV_PAIRS_TMP, show_col_types = FALSE)
} else {
  cat("No within-committee matches found.\n")
  tibble(.x = integer(), .y = integer())
}

indiv_pairs <- indiv_pairs %>%
  group_by(.x, .y) %>%
  slice_max(score, n = 1, with_ties = FALSE) %>%
  ungroup()


# name frequency adjustment
# use stricter matching thresholds when names are common
# or when employers/orgs are uninformative
indiv_pairs <- indiv_pairs %>%
  left_join(individuals %>% select(row_id, last_name_str = last_name, first_name_str = first_name),
            by = c(".x" = "row_id")) %>%
  left_join(last_name_freq,  by = c("last_name_str"  = "last")) %>%
  left_join(first_name_freq, by = c("first_name_str" = "first")) %>%
  mutate(
    # flag whether first and last names are relatively common in the data
    common_last  = coalesce(last_name_freq,  1L) >= COMMON_LAST_MIN,
    common_first = coalesce(first_name_freq, 1L) >= COMMON_FIRST_MIN,
    common_name  = common_last & common_first,
    rare_name    = !common_last & !common_first,
    match_threshold = case_when(
      common_name & any_context_dropped ~ COMMON_NAME_STRICT,
      common_name                       ~ COMMON_NAME_THRESHOLD,
      any_context_dropped               ~ STRICT_THRESHOLD,
      TRUE                              ~ MATCH_THRESHOLD
    ),
    # unusual names flagged for review at a lower threshold 
    review_threshold = if_else(rare_name, RARE_NAME_UNCERTAIN_LOW, UNCERTAIN_LOW)
  ) %>%
  select(-last_name_str, -first_name_str, -last_name_freq, -first_name_freq)

# link tiers: auto-link vs no-link vs review
indiv_pairs <- indiv_pairs %>%
  mutate(
    link_tier = case_when(
      coalesce(first_name, 0) < FIRST_NAME_MIN ~ "no_link", # first names have to be relatively similar
      city_conflict & any_context_dropped ~ "no_link",
      common_name & city_conflict ~ "no_link",       # city disagreement is enough to reject common names
      common_name & employer_conflict ~ "no_link",   # clearly different employers also enough for common names
      score < review_threshold ~ "no_link",
      score >= match_threshold ~ "auto_link",
      TRUE                                           ~ "review"
    ),
    match_source = "internal_fuzzy"
  )

cat("Individual pairs — auto_link:", sum(indiv_pairs$link_tier == "auto_link"),
    "| review:", sum(indiv_pairs$link_tier == "review"),
    "| employer conflicts in review:", sum(indiv_pairs$employer_conflict & indiv_pairs$link_tier == "review"), "\n")

```


### Cluster individuals

Form entity clusters based on suspected matches.
Auto-links and review links are treated as true when forming clusters. 
The clusters are then used to flag which individual contributors are likely 
the same person, but used a different `Contributor.Name`.

```{r}
# treat all links as true -> name variants we be reviewed to see if they are real matches
linkable_indiv_pairs <- indiv_pairs %>% filter(link_tier %in% c("auto_link", "review"))

if (nrow(linkable_indiv_pairs) > 0) {
  g_indiv        <- graph_from_data_frame(linkable_indiv_pairs %>% select(.x, .y), directed = FALSE)
  indiv_clusters <- components(g_indiv)$membership

  indiv_cluster_raw <- tibble(row_id = as.integer(names(indiv_clusters)), cluster_num = indiv_clusters) %>%
    left_join(individuals %>% select(row_id, Contributor.Name, Contributor.ID, standardized_name,
                                     original_row_id), by = "row_id")

  indiv_cluster_uuid_map <- indiv_cluster_raw %>%
    group_by(cluster_num) %>%
    summarise(raw_names = list(unique(Contributor.Name)), .groups = "drop") %>%
    rowwise() %>%
    mutate(
      result      = list(lookup_cluster_uuid(raw_names, nem_lookup)),
      entity_id   = result$entity_id,
      is_conflict = result$is_conflict,
      is_new      = result$is_new
    ) %>%
    ungroup() %>%
    select(cluster_num, entity_id, is_conflict, is_new)

  if (any(indiv_cluster_uuid_map$is_conflict, na.rm = TRUE)) {
    cat(sprintf("WARNING: %d cluster(s) have conflicting entity registry UUIDs — check indiv_conflict_review\n",
                sum(indiv_cluster_uuid_map$is_conflict, na.rm = TRUE)))
    indiv_conflict_review <- indiv_cluster_raw %>%
      filter(cluster_num %in% indiv_cluster_uuid_map$cluster_num[indiv_cluster_uuid_map$is_conflict]) %>%
      left_join(indiv_cluster_uuid_map %>% select(cluster_num, entity_id), by = "cluster_num") %>%
      select(cluster_num, entity_id, Contributor.Name, Contributor.ID)
  }

  indiv_cluster_map <- indiv_cluster_raw %>%
    left_join(indiv_cluster_uuid_map, by = "cluster_num") %>%
    select(row_id = original_row_id, entity_id, is_conflict, is_new)
} else {
  indiv_cluster_map <- tibble(row_id = integer(), entity_id = character(),
                               is_conflict = logical(), is_new = logical())
}

# look up entity_id for unique Contributor.Name
indiv_singleton_lookup <- individuals %>%
  filter(!original_row_id %in% indiv_cluster_map$row_id) %>%
  distinct(Contributor.Name) %>%
  mutate(raw_name_key = toupper(Contributor.Name)) %>%
  left_join(nem_lookup %>% select(raw_name, entity_uuid), by = c("raw_name_key" = "raw_name")) %>%
  mutate(
    is_new    = is.na(entity_uuid),
    entity_id = coalesce(entity_uuid, UUIDgenerate(n = n()))
  ) %>%
  select(Contributor.Name, entity_id, is_new)

indiv_singletons <- individuals %>%
  filter(!original_row_id %in% indiv_cluster_map$row_id) %>% # not a part of an entity group
  select(-entity_id) %>%
  left_join(indiv_singleton_lookup, by = "Contributor.Name") %>% # join entity_id that corresponds to this name
  mutate(is_conflict = FALSE) %>%
  select(row_id = original_row_id, entity_id, is_new, is_conflict)

# auto-link only clustering: used for comparison with the method of aggregating by `Normalized` name
# NA auto_cluster means singleton or review link
auto_only_pairs <- indiv_pairs %>% filter(link_tier == "auto_link")

if (nrow(auto_only_pairs) > 0) {
  g_auto        <- graph_from_data_frame(auto_only_pairs %>% select(.x, .y), directed = FALSE)
  auto_clusters <- components(g_auto)$membership

  auto_cluster_map <- tibble(
    row_id       = as.integer(names(auto_clusters)),
    auto_cluster = auto_clusters
  ) %>%
    left_join(individuals %>% select(row_id, original_row_id), by = "row_id") %>%
    select(original_row_id, auto_cluster)
} else {
  auto_cluster_map <- tibble(original_row_id = integer(), auto_cluster = integer())
}

# within-individuals row_id to entity_id map (used by name variant queue to attach entity_id)
# indiv_pairs use within-individuals row_id; cluster_map was renamed to original_row_id
indiv_internal_to_entity <- individuals %>%
  select(row_id, original_row_id) %>%
  left_join(bind_rows(indiv_cluster_map, indiv_singletons), by = c("original_row_id" = "row_id"))

individuals_resolved <- individuals %>%
  select(-entity_id) %>%   # drop initial Normalized name-based entity_id; cluster-based wins
  left_join(bind_rows(indiv_cluster_map, indiv_singletons), by = c("original_row_id" = "row_id")) %>%
  left_join(auto_cluster_map, by = "original_row_id")

``` 

## Individual name variant review queue

For each pair surfaced by the record linkage algorithm where the two names differ, output
one row per name.

To narrow down the scope, we're only reviewing name variants of tentatively-linked entity groups
that meet the $5k threshold. Since we are not classifying entities that contributed less than 
$5k to a race, prop or PAC, tracking the name variants of these entities doesn't matter.

```{r}
# use indiv_clusters from the clustering step - includes auto-links and review links
if (exists("indiv_clusters") && length(indiv_clusters) > 0) {
  review_group_map <- tibble(
    row_id          = as.integer(names(indiv_clusters)),
    entity_group_id = paste0("igroup_", indiv_clusters)
  )
} else {
  review_group_map <- tibble(row_id = integer(), entity_group_id = character())
}

# join directly by row_id — only rows that were actually in a linked pair are included.
indiv_review_data <- review_group_map %>%
  left_join(
    individuals %>%
      select(row_id, standardized_name, Normalized, Contributor.Name, Contributor.Employer,
             Contributor.Occupation, Contributor.City,
             Contributor.State, Contributor.Zip.Code,
             Amount, Recipient.Name, Recipient.Committee.ID, Ballot.Measure,
             Office, is_pac_contribution, race_prop),
    by = "row_id"
  ) %>%
  left_join(indiv_internal_to_entity %>% select(row_id, entity_id), by = "row_id") %>%
  mutate(
    exported_name = toupper(Contributor.Name), # match the format of the names in the 2026 name map
    raw_name_key  = exported_name
  ) %>%
  add_recipient_key() %>%
  left_join(nem_lookup %>% select(raw_name, confirmed, canonical_name),
            by = c("raw_name_key" = "raw_name")) %>%
  mutate(confirmed = coalesce(confirmed, FALSE)) %>%
  group_by(entity_group_id) %>%
  filter(n_distinct(exported_name) > 1) %>% # groups with distinct Contributor.Names (multiple name variants)
  ungroup()

# all individuals contributions by Normalized × recipient — used for both queue thresholds
all_norm_recip_totals <- individuals %>%
  mutate(norm_key = coalesce(Normalized, toupper(Contributor.Name))) %>%
  add_recipient_key() %>%
  group_by(norm_key, recipient_key) %>%
  summarise(norm_total = sum(Amount, na.rm = TRUE), .groups = "drop")

# for each entity_group, identify the (`Normalized` × recipient) combos of its members,
# then pull ALL contributions sharing those `Normalized` values for those same recipients 
# (just in case one wasn't included in one of the matching pairs)

entity_group_norms_recips <- indiv_review_data %>%
  mutate(norm_key = coalesce(Normalized, toupper(Contributor.Name))) %>%
  distinct(entity_group_id, norm_key, recipient_key) # may be more than one normalized name per entity group


indiv_review_group_threshold <- entity_group_norms_recips %>%
  left_join(all_norm_recip_totals, by = c("norm_key", "recipient_key")) %>%
  group_by(entity_group_id, recipient_key) %>%
  summarise(recipient_total = sum(coalesce(norm_total, 0), na.rm = TRUE), .groups = "drop") %>%
  group_by(entity_group_id) %>%
  summarise(max_single_recipient = max(recipient_total, na.rm = TRUE), .groups = "drop")

# flagging whether names are already in the name map or not
indiv_group_status <- indiv_review_data %>%
  group_by(entity_group_id) %>%
  summarise(
    entity_status = case_when(
      all(confirmed)  ~ "all_confirmed",
      any(confirmed)  ~ "adding_to_existing",
      TRUE            ~ "new_entity"
    ),
    .groups = "drop"
  )

# take out groups where every name pair already exists in the no-link registry (they were all reviewed already)
indiv_registry_pair_keys <- no_link_registry %>%
  filter(entity_type == "indiv") %>%
  mutate(pair_key = paste(name_a, name_b, sep = "|||")) %>%
  pull(pair_key)

indiv_group_name_pairs <- indiv_review_data %>%
  distinct(entity_group_id, name = toupper(exported_name))

indiv_fully_reviewed_groups <- indiv_group_name_pairs %>%
  inner_join(indiv_group_name_pairs, by = "entity_group_id", relationship = "many-to-many") %>%
  filter(name.x < name.y) %>%
  mutate(pair_key = paste(name.x, name.y, sep = "|||"),
         reviewed = pair_key %in% indiv_registry_pair_keys) %>%
  group_by(entity_group_id) %>%
  summarise(all_reviewed = all(reviewed), .groups = "drop") %>%
  filter(all_reviewed) %>%
  pull(entity_group_id)

# First determine which entity_groups qualify based on cluster-member analysis
qualifying_groups <- indiv_review_data %>%
  distinct(entity_group_id) %>%
  left_join(indiv_group_status, by = "entity_group_id") %>%
  left_join(indiv_review_group_threshold, by = "entity_group_id") %>%
  filter(
    entity_status != "all_confirmed",
    !entity_group_id %in% indiv_fully_reviewed_groups,
    max_single_recipient >= 5000
  )

# expand to ALL individuals sharing the same (Normalized × recipient) as any group member —
# not just cluster members - the reviewer needs every row associated with the flagged Normalized
# value in that committee, including rows that never directly matched a cluster member
indiv_queue_expanded <- entity_group_norms_recips %>%
  filter(entity_group_id %in% qualifying_groups$entity_group_id) %>%
  left_join(
    individuals %>%
      mutate(
        norm_key      = coalesce(Normalized, toupper(Contributor.Name)),
        exported_name = toupper(Contributor.Name),
        raw_name_key  = exported_name
      ) %>%
      add_recipient_key(),
    by = c("norm_key", "recipient_key")
  ) %>%
  left_join(nem_lookup %>% select(raw_name, confirmed, canonical_name),
            by = c("raw_name_key" = "raw_name")) %>%
  left_join(indiv_internal_to_entity %>% select(row_id, cluster_uuid = entity_id), by = "row_id") %>%
  mutate(confirmed = coalesce(confirmed, FALSE))

indiv_name_variant_queue <- indiv_queue_expanded %>%
  group_by(entity_group_id, exported_name, Contributor.Employer, Contributor.Occupation) %>%
  summarise(
    cluster_uuid    = first(cluster_uuid),
    canonical_name  = first(canonical_name),
    confirmed       = first(confirmed),
    n_contributions = n(),
    total_amount    = sum(Amount, na.rm = TRUE),
    city      = paste(sort(unique(Contributor.City[!is.na(Contributor.City)   & Contributor.City   != ""])), collapse = " | "),
    state     = paste(sort(unique(Contributor.State[!is.na(Contributor.State)  & Contributor.State  != ""])), collapse = " | "),
    zip       = paste(sort(unique(Contributor.Zip.Code[!is.na(Contributor.Zip.Code) & Contributor.Zip.Code != ""])), collapse = " | "),
    recipient = paste(sort(unique(recipient_key[!is.na(recipient_key) & recipient_key != ""])), collapse = " | "),
    .groups = "drop"
  ) %>%
  left_join(qualifying_groups %>% select(entity_group_id, entity_status, max_single_recipient),
            by = "entity_group_id") %>%
  select(entity_group_id,
         entity_name = exported_name, cluster_uuid, canonical_name,
         entity_status, confirmed, Contributor.Employer, Contributor.Occupation,
         n_contributions, total_amount, city, state, zip, recipient, max_single_recipient)
```


```{r}
today <- format(Sys.Date(), "%m-%d-%y")
INDIV_REVIEW_OUTPUT_PATH <- paste0("08_outputs/indiv_name_variant_review_queue", today, ".csv")
write_csv(indiv_name_variant_queue, INDIV_REVIEW_OUTPUT_PATH)
cat("Individual name variant review queue:", n_distinct(indiv_name_variant_queue$entity_group_id),
    "groups,", nrow(indiv_name_variant_queue), "distinct records\n")
```



## Record linkage disambiguation queue

Surfaces cases where two records share the same `Normalized` value but are scored as
`no_link` (confirmed different people) or `review` (uncertain link). Both cases require
disambiguation: `no_link` pairs identify that records grouped together under `Normalized`
may actually be different people; `review` pairs with the same `Normalized` can't be
surfaced by the name variant queue (which requires distinct names) and need a human
decision. Applies the same $5k per-recipient threshold as the individual name variant queue.

```{r}
# no_link and review pairs where both sides share the same Normalized value
same_name_no_link <- indiv_pairs %>%
  filter(link_tier %in% c("no_link", "review")) %>%
  left_join(individuals %>% select(row_id, nm_x = Normalized), by = c(".x" = "row_id")) %>%
  left_join(individuals %>% select(row_id, nm_y = Normalized), by = c(".y" = "row_id")) %>%
  filter(!is.na(nm_x), nm_x == nm_y)

# restrict to the specific (Normalized × committee) combinations from the flagged pairs;
# pulling all committees for a flagged Normalized value incorrectly includes records that
# contributed to different recipients and were never part of any no_link pair
flagged_combos <- individuals %>%
  filter(row_id %in% unique(c(same_name_no_link$.x, same_name_no_link$.y))) %>%
  add_recipient_key() %>%
  distinct(Normalized, recipient_key)

rla_disambig_data <- individuals %>%
  add_recipient_key() %>%
  inner_join(flagged_combos, by = c("Normalized", "recipient_key")) %>%
  mutate(entity_name = toupper(Contributor.Name))

# threshold uses Normalized-level totals from all individuals (reuses all_norm_recip_totals)
rla_disambig_threshold <- rla_disambig_data %>%
  mutate(norm_key = coalesce(Normalized, toupper(Contributor.Name))) %>%
  distinct(Normalized, norm_key) %>%
  left_join(
    all_norm_recip_totals %>%
      group_by(norm_key) %>%
      summarise(max_single_recipient = max(norm_total, na.rm = TRUE), .groups = "drop"),
    by = "norm_key"
  ) %>%
  group_by(Normalized) %>%
  summarise(max_single_recipient = max(coalesce(max_single_recipient, 0), na.rm = TRUE), .groups = "drop") %>%
  filter(max_single_recipient >= 5000)

rla_disambig_queue <- rla_disambig_data %>%
  filter(Normalized %in% rla_disambig_threshold$Normalized) %>%
  select(
    Normalized,
    entity_name,
    entity_id,
    Contributor.Employer,
    Contributor.Occupation,
    Contributor.City,
    Contributor.State,
    Contributor.Zip.Code,
    Recipient.Committee.ID
  ) %>%
  distinct() %>%
  group_by(Normalized) %>%
  filter(n() >= 2) %>%          # remove single-profile groups — same person flagged no_link due to
  ungroup() %>%                 # common name / uninformative fields; already grouped by singleton path
  left_join(rla_disambig_threshold %>% select(Normalized, max_single_recipient),
            by = "Normalized") %>%
  mutate(in_names_map = entity_name %in% names_2026$raw_name)

# flag Normalized × committee combinations where multiple raw names are present —
# indicates name variants contributing under the same Normalized entity
committee_name_variant_flags <- rla_disambig_queue %>%
  group_by(Normalized, Recipient.Committee.ID) %>%
  summarise(has_name_variants = n_distinct(entity_name) > 1, .groups = "drop")

name_committee_totals <- rla_disambig_data %>%
  mutate(entity_name = toupper(Contributor.Name)) %>%
  group_by(entity_name, Contributor.Employer, Contributor.Occupation,
           Contributor.City, Contributor.State, Contributor.Zip.Code,
           Recipient.Committee.ID) %>%
  summarise(
    amount_to_committee = sum(Amount, na.rm = TRUE),
    n_contributions     = n(),
    .groups = "drop"
  )

rla_disambig_queue <- rla_disambig_queue %>%
  left_join(committee_name_variant_flags, by = c("Normalized", "Recipient.Committee.ID")) %>%
  left_join(name_committee_totals,
            by = c("entity_name", "Contributor.Employer", "Contributor.Occupation",
                   "Contributor.City", "Contributor.State", "Contributor.Zip.Code",
                   "Recipient.Committee.ID")) %>%
  arrange(Normalized, entity_name, Contributor.Employer, Contributor.Occupation, Contributor.City)

cat("Record linkage disambiguation queue:", n_distinct(rla_disambig_queue$Normalized),
    "Normalized groups,", nrow(rla_disambig_queue), "distinct records\n")
``` 


```{r}
today <- format(Sys.Date(), "%m-%d-%y")
DISAMBIG_REVIEW_OUTPUT_PATH <- paste0("08_outputs/individual_disambiguation_queue", today, ".csv")
write_csv(rla_disambig_queue, DISAMBIG_REVIEW_OUTPUT_PATH)
```


## Manual disambiguation review queue

Used for names I noticed while looking at the data. 

Each name in `name_list` is matched as a substring of the uppercased contributor name,
so "SMITH, STEVE" captures "SMITH, STEVE R." and "SMITH, STEVE ROBERT".

```{r}
name_list <- c(
  "SMITH, STEVE", "DAVIS, JOHN", "KING, JUSTIN", "SPENCER, RICHARD",
  "FISHER, ROBERT", "ROBERTSTON, CARL", "ROBERTSON, CHIP", "W ROBERTSON, CARL"
)

combined_pattern <- paste(toupper(name_list), collapse = "|")

manual_disambig_queue <- individuals %>%
  mutate(entity_name = toupper(Contributor.Name)) %>%
  filter(grepl(combined_pattern, entity_name)) %>%
  select(
    entity_name,
    Contributor.Employer,
    Contributor.Occupation,
    Contributor.City,
    Contributor.State,
    Contributor.Zip.Code,
    Recipient.Committee.ID,
    Recipient.Committee,
    Amount
  ) %>%
  distinct() %>%
  arrange(entity_name, Contributor.Employer, Contributor.Occupation, Contributor.City)

cat("Manual disambiguation queue:", nrow(manual_disambig_queue), "distinct records across",
    n_distinct(manual_disambig_queue$entity_name), "name variants\n")
```

```{r}
write_csv(bind_rows(manual_disambig_queue,rla_disambig_queue),"08_outputs/individual_disambiguation.csv")
```
 