---
title: "Entity Resolution"
output: html_notebook
---

# Purpose

This notebook aims to group contributors that belong to the same entity for the purpose of: 
1. Aggregating contributions by entity to see which cross the $5k threshold. 
2. Computing one classification per entity for consistency and to minimize the number of classifications needed.

Process: 

Group organizations and committees
1. Look for any contributors that are missing a Contributor.ID by name matching to the full contributions data (across races, props and recipient committees). 
2. Group contributors that share the same Contributor.ID under one entity. 
3. Look for exact name matches between contributors based on different processing of the name (removing parentheses, removing business endings, etc.)
4. Fuzzy match contributors to see if there are any other links
5. Use external references from Digital Democracy and previous campaign finance work to group orgs that were determined to be the same entity.

Individuals: 
1. Group individuals based on record linkage algorithm that assigns different weights according to the importance of each column in distinguishing individuals.
Looks at name, employer, occupation, city, state and zip.

- Contributors are assigned an entity_id and canonical name based on suspected links. Some entities are flagged for review. 
- Data is filtered to only include entities that contributed $5k or more in the classification process. 


# Questions
- are there duplicates in the data or are people really making two of the exact same contributions on the same day?
- why does same contributor name match to several contributor ids (for cases in the full contributions data)
- ask for whatever mapping is used for camp fin this year to group contributors..
- how to treat different unit numbers -> same entity or separate? also depends on what ends up happening with the PACs. 
- ask about whether or not to propagate the contributor.id automatically. 
  - mostly weird in major donor cases where most of the time the contributor id is NA, but some of them do have a contributor id
  - also don't want to get major donors confused with PACs / committees

# To Do 
- add uuid
- output contribution-level data... 
- add process for only processing new contributors
- add process for flagging which entities have already been classified
- add separate output for committees to send to PAC processing
- simplify where possible
- incorporate id_map to see which are major donors if possible (don't want to send these to PAC review)
- add PAC output
- use id map instead of contributions_full for propagating ids (should have more name variants)
- could add a check for anything w/ PAC/fund/etc that doesn't have a contributor id
- could try using fastLink or using splink to see if it works better, esp for individuals
- look at full contib data for cases w/ diff ids per name to see whats happening
- check no negatives after aggregation
- make sure hashing not based on any standardization that could change



# Load packages

```{r}
here::i_am("0802_entity_resolution.Rmd")
library(here)
library(dplyr)
library(stringr)
library(stringdist)
library(reclin2)
library(phonics)
library(igraph)
library(readr)
library(digest)
library(jsonlite)
library(tidyr)

source("standardization_helpers.R")
```

```{r}
safe_metaphone <- function(x) {
  clean <- str_replace_all(coalesce(x, ""), "[^A-Za-z]", "")
  ifelse(clean != "", metaphone(clean), NA_character_)
}

bigram_cosine <- function(a, b) stringsim(a, b, method = "cosine", q = 2)

hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16))
}

normalize_str <- function(x) {
  coalesce(x, "") %>% toupper() %>%
    stringr::str_replace_all("[^A-Z0-9 ]+", " ") %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_trim()
}
```


# Load and standardize all contributors data

```{r}
#contributions_full <- read.csv("08_inputs/contributors_data_full.csv")
contributions_full <- read.csv("08_inputs/all_contributors_processed.csv")
nrow(contributions_full)
```


### load already categorized data

```{r}
categorized <- read.csv("08_inputs/already_classified_contributions.csv")
nrow(categorized)
```

### look at aliases by contributor id

```{r, eval = FALSE}
contributions_full %>% group_by(Contributor.ID) %>% mutate(count = n_distinct(Contributor.Name)) %>% arrange(Contributor.ID) %>% filter(count > 1) %>% select(Contributor.Name, Contributor.ID) %>% unique()
```


### filter for contributions data we're interested in

```{r}
contributions <- contributions_full %>% filter(
    Ballot.Measure == "OPPOSED: PROPOSITION 040 - IMPOSES ONE-TIME TAX ON CERTAIN INDIVIDUALS AND TRUSTS. INITIATIVE CONSTITUTIONAL AMENDMENT AND STATUTE." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 040 - IMPOSES ONE-TIME TAX ON CERTAIN INDIVIDUALS AND TRUSTS. INITIATIVE CONSTITUTIONAL AMENDMENT AND STATUTE." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 041 - REQUIRES AUDITS OF PROGRAMS FUNDED BY NEW STATE SPECIAL TAXES. PROHIBITS NEW STATE TAXES THAT ARE EXCLUDED FROM EXISTING VOTER-APPROVED STATE SPENDING LIMIT..." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 042 - PROHIBITS NEW STATE PERSONAL PROPERTY TAXES AND CERTAIN RETROACTIVE STATE TAXES. INITIATIVE CONSTITUTIONAL AMENDMENT." | 
    Recipient.Name == "BECERRA, XAVIER" | 
    Recipient.Name == "HILTON, STEVE" | 
    Office == "Insurance Commissioner"
) %>%
  mutate(
    Amount = as.numeric(Amount)
  ) #%>% 
  #rename(
  #  Ballot.Measure = 'Ballot.Measure(s)'
  #)

nrow(contributions)
```



```{r}
# make sure we' pulled everything (compare to power search)
contributions %>% filter(Recipient.Name == "BECERRA, XAVIER") %>% summarise(total = sum(Amount))
contributions %>% filter(Recipient.Name == "HILTON, STEVE") %>% summarise(total = sum(Amount))
```


```{r}
# flag contributions by race/prop
contributions <- contributions %>%
  mutate(
    race_prop = case_when(
      Ballot.Measure.Contribution == "Y" ~ Ballot.Measure,
      !is.na(Office) & Office != "" ~ Office,
      TRUE ~ Recipient.Name
    )
  )
unique(contributions$race_prop)
```


### standardize contributions data


```{r}
make_row_hash <- function(...) {
  cols <- list(...)
  n <- length(cols[[1]])
  vapply(seq_len(n), function(i) {
    parts <- vapply(cols, function(col) as.character(col[i]), character(1))
    digest::digest(paste(parts, collapse = "|"), algo = "xxhash32")
  }, character(1))
}

```


```{r}
contributions <- contributions %>%
  filter(Contributor.Name != "Unitemized Contributions") %>%
  mutate(
    standardized_name = standardize_names(Contributor.Name) %>% fix_typos(),
    processed_name = standardized_name %>%
      remove_pac_info() %>% remove_unit_info() %>% 
      str_remove_all("\\s*\\(.*?\\)") %>% replace_business_text() %>% str_squish(),

    standardized_employer_name = standardize_names(Contributor.Employer) %>% fix_typos(),
    processed_employer_name = standardized_employer_name %>% standardize_occupation_employer() %>%
      remove_pac_info() %>% remove_unit_info() %>% str_remove_all("\\s*\\(.*?\\)") %>%
      replace_business_text() %>%
      str_squish(),

    standardized_occupation = standardize_names(Contributor.Occupation),
    processed_occupation    = standardize_occupation_employer(standardized_occupation),
    standardized_city       = standardize_names(Contributor.City),
    zip_code_processed      = substr(Contributor.Zip.Code, 1, 5),

    has_pac_language = has_pac_language(Contributor.Name),
    entity_type      = if_else(sapply(standardized_name, is_individual), "individual", "organization"),
    row_id           = row_number(),
    contribution_id  = make_row_hash(Contributor.Name, Contributor.ID, Amount, race_prop,
                                      Contributor.City, Contributor.Employer,
                                      Contributor.Occupation, Contributor.Zip.Code, Start.Date)
  )

```

```{r}
contributions <- read.csv("08_inputs/race_contributions_processed.csv")
```


```{r, eval = FALSE}
contributions %>% group_by(contribution_id) %>% summarise(count = n()) %>% filter(count > 1)
```


## manually process edge cases

Manually verified each of these. 

```{r}
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$standardized_name <- "LOVECCHIO, PETER"
contributions[which(contributions$Contributor.Name == "Lovecchio, US Navy (R, CDR Peter"),]$processed_name <- "LOVECCHIO PETER"

contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$standardized_name <- "ZUNILDA, PEREZ"
contributions[which(contributions$Contributor.Name == "Title Group Llc Zunilda Perez, Zmar"),]$processed_name <- "ZUNILDA, PEREZ"

contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$standardized_name <- "LIDELL, JOHN MARK" # PA-C = physician assistant (don't want to add PAC to suffix list though)
contributions[which(contributions$Contributor.Name == "Lidell   Pac, John Mark"),]$processed_name <- "LIDELL JOHN MARK"

contributions[which(contributions$Contributor.Name == "Last0b6df6bd, First7606aee3"),]$entity_type <- "individual"
contributions[which(contributions$Contributor.Name == "Lastcbea8f37, Firstd599d312"),]$entity_type <- "individual"
```

### standardize occupation and employer for organizations and individuals


## orgs 

- check employer of non-individuals
- for organizations, could change standardized occupation and employers to be blank for consistency -> but don't know if this is a factor later? 
- comes into play when looking at unique name, employer combos? 
- There are also cases where employer = N/A or none, but occupation = self-employed, earlier, i changed employer to also be self employer, but not sure this matters. maybe for matching, but not much info anyways.
- Not having an occupation would mean these people are send to the no-employer queue

```{r,eval = FALSE}
contributions %>% filter(entity_type == "individual",Contributor.Employer == "N/A",Contributor.Occupation != "N/A") %>% select(Contributor.Name,Contributor.Employer,Contributor.Occupation)
```

```{r}
#contributions %>% filter(entity_type == "organization") %>% select(Contributor.Employer) %>% distinct()
#contributions %>% filter(entity_type == "organization") %>% select(Contributor.Occupation) %>% table()
```


## individuals

- Make sure employer and occupation are NONE for individuals where these columns are blank 
- Could also update employer to UNKNOWN for individuals where employer is N/A, but occupation specifies they're self-employed
  - Are cases where employer is N/A, but occupation indicates they have a job.

```{r}
contributions <- contributions %>%
  mutate(
    standardized_employer_name   = if_else(entity_type == "individual" & standardized_employer_name == "", "NONE", standardized_employer_name),
    processed_employer_name  = if_else(entity_type == "individual" & processed_employer_name == "", "NONE", processed_employer_name),

    standardized_occupation  = if_else(entity_type == "individual" & standardized_occupation == "", "NONE", standardized_occupation),
    processed_occupation = if_else(entity_type == "individual" & processed_occupation == "", "NONE", processed_occupation),    
  )

```



# Load and standardize name maps

```{r}
map1 <- read.csv("08_inputs/dd_names_mapped_to_org_concepts_3_18_2026.csv") %>%
  rename(canonical_name = org_concept_name, alias_group_id = oid) %>%
  mutate(
    record_name    = str_squish(record_name),
    canonical_name = str_squish(canonical_name),
    mapping_source = "dd_name_mapping",
    confirmed      = TRUE
  )

map2 <- read.csv("08_inputs/2024 Contributor name mapping and descriptions - Combined.csv") %>%
  select(record_name = Original, canonical_name = Normalized) %>%
  filter(!is.na(record_name) & record_name != "") %>%
  mutate(
    record_name    = str_squish(record_name),
    canonical_name = str_squish(canonical_name),
    alias_group_id = NA_integer_,
    mapping_source = "2024_combined",
    confirmed      = TRUE
  )

# combine — prefer map1 when both have same record_name
reference_mapping <- bind_rows(map1, map2) %>%
  group_by(record_name) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(record_name_std = standardize_names(record_name) %>% fix_typos())

```

# Organizations

```{r}
make_name_variants <- function(df, col) {
  df %>% mutate(
    name_v1 = standardize_names(.data[[col]]) %>% fix_typos(),
    name_v2 = remove_pac_info(name_v1)        %>% str_squish(), # removing FEC ID, FED PAC ID, MAJOR DONOR ID
    #name_v3 = remove_unit_info(name_v2)        %>% str_squish(),
    name_v4 = str_remove_all(name_v1, "\\s*\\(.*?\\)") %>% str_squish(),
    name_v5 = replace_business_text(name_v4)   %>% str_squish()
    #name_v6 = str_remove_all(name_v5, "\\s*\\(.*?\\)") %>% str_squish()
  )
}

find_exact_variant_pairs <- function(pool, variant_col) {
  candidates <- pool %>%
    filter(!is.na(.data[[variant_col]]), .data[[variant_col]] != "") %>%
    group_by(across(all_of(c(variant_col, "has_pac_language")))) %>%
    filter(n() > 1) %>%
    ungroup()

  if (nrow(candidates) == 0)
    return(tibble(.x = integer(), .y = integer(), match_source = character()))

  candidates %>%
    group_by(across(all_of(c(variant_col, "has_pac_language")))) %>%
    group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>%
    ungroup() %>%
    select(.x, .y) %>%
    mutate(match_source = paste0("exact_", variant_col))
}
```


```{r}
orgs <- contributions %>%
  filter(entity_type == "organization") %>%
  mutate(cont_row_id = row_id) %>%          # preserve row_id before overwriting
  mutate(has_contributor_id = !is.na(Contributor.ID) & str_trim(as.character(Contributor.ID)) != "") %>%
  make_name_variants("Contributor.Name") %>%
  filter(!is.na(name_v1), name_v1 != "") %>%
  mutate(row_id = row_number())


# boilerplate removal for fuzzy blocking key
#all_tokens    <- orgs$name_v5 %>% str_split("\\s+") %>% unlist()
#token_counts  <- table(all_tokens)

#BOILERPLATE_MIN_SHARE <- 0.05  # was 0.01
#boilerplate_words <- names(token_counts)[token_counts > nrow(orgs) * BOILERPLATE_MIN_SHARE]
#boilerplate_words <- names(token_counts)[token_counts > nrow(orgs) * 0.01]
#cat("boilerplate words dropped:", paste(boilerplate_words, collapse = ", "), "\n")

#strip_boilerplate <- function(name) {
#  vapply(str_split(name, "\\s+"), function(toks) {
#    kept <- toks[!toks %in% boilerplate_words]
#    paste(if (length(kept) == 0) toks else kept, collapse = " ")
#  }, character(1))
#}

#orgs <- orgs %>%
#  mutate(
#    core_name   = strip_boilerplate(name_v5),
#    first_token = str_extract(core_name, "^\\S+"),
#    block_key   = safe_metaphone(first_token)
#  )
```



## Look at names in contributor data that match to multiple contributor ids

```{r}
orgs %>% group_by(name_v5) %>%
  filter(n_distinct(Contributor.ID) > 1) %>% arrange(name_v5) %>% select(Contributor.Name,name_v5,Contributor.ID)

orgs %>% group_by(Contributor.Name) %>%
  filter(n_distinct(Contributor.ID) > 1) %>% arrange(Contributor.Name) %>% select(Contributor.Name,name_v5,Contributor.ID)
```

- flag this as an error if it comes up
- pacaso inc. is just a major donor, but ramsey robinson for gov 2025 should defs have a contributor id


## ID Propagation - skip for now

- consider whether to do a name variant like name_v5 or a stricter name match

```{r}
# Build name/ID lookup from full dataset
full_id_lookup <- contributions_full %>%
  filter(!is.na(Contributor.ID), Contributor.ID != "") %>%
  mutate(name_v5 = standardize_names(Contributor.Name) %>% fix_typos() %>%
           replace_business_text() %>% # remove_pac_info() %>% remove_unit_info() %>% 
           str_remove_all("\\s*\\(.*?\\)") %>% str_squish()) %>%
  distinct(name_v5, Contributor.ID) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(Contributor.ID)) %>%
  ungroup()
```


```{r}
full_id_lookup %>% 
  group_by(name_v5) %>%
  filter(n_distinct(Contributor.ID) > 1) %>% arrange(name_v5)
```

- There are many instances where the same name matches to multiple contributor ids.
  - in some cases I'm wondering if its the difference between xrefid and the filer id
  - other times it just looks like a typo in the id which is unfortuante - will be hard to grab these

```{r}
# Propagate IDs to orgs that are missing them
full_id_propagation <- orgs %>%
  filter(!has_contributor_id) %>%
  select(row_id, Contributor.Name, name_v5) %>%
  inner_join(full_id_lookup, by = "name_v5") %>%
  mutate(needs_review = n_ids > 1) %>%
  select(row_id, Contributor.Name, propagated_id = Contributor.ID, n_ids, needs_review)

cat("IDs propagated (unambiguous):", n_distinct(full_id_propagation$row_id[!full_id_propagation$needs_review]), "\n")
cat("IDs to review (multiple matches):", n_distinct(full_id_propagation$row_id[full_id_propagation$needs_review]), "\n")

# apply unambiguous ones back to orgs
orgs <- orgs %>%
  left_join(
    full_id_propagation %>% filter(!needs_review) %>% distinct(row_id, propagated_id),
    by = "row_id"
  ) %>%
  mutate(
    Contributor.ID     = coalesce(Contributor.ID, propagated_id),
    has_contributor_id = !is.na(Contributor.ID) & str_trim(as.character(Contributor.ID)) != ""
  ) %>%
  select(-propagated_id)

# inspect review cases
full_id_propagation %>%
  filter(needs_review) %>%
  arrange(Contributor.Name, propagated_id)

```

Look at which rows ids were added to: 

```{r}
full_id_propagation %>%
  filter(!needs_review) %>%
  arrange(Contributor.Name, propagated_id)
```

- most of these look like major donors as opposed to PACs/ committees. 
  - maybe use id map to flag which are major donors? 

```{r}
contributions_full %>% filter(Contributor.Name == "Twenty-Nine Palms Band of Mission Indians") %>% select(Contributor.Name,Contributor.ID) #%>% unique()
```

- some of these ids may correspond to a major donor id instead of a committee id? e.g. blue shielf of california
- ones that match to diff ids (one being NA are not flagged for review)


## Step 1: ID-based aliases (same Contributor.ID, different names)

- should update this to just group by contributor.id, regardless of whether names are different or not

```{r}
id_based_aliases <- orgs %>%
  filter(has_contributor_id) %>%
  group_by(Contributor.ID) %>%
  filter(n_distinct(Contributor.Name) > 1) %>%
  ungroup()

id_auto_pairs <- id_based_aliases %>%
  group_by(Contributor.ID) %>%
  group_modify(~ as_tibble(t(combn(.x$row_id, 2)), .name_repair = ~ c(".x", ".y"))) %>%
  ungroup() %>%
  mutate(link_tier = "auto_link", match_source = "same_contributor_id")

cat("ID-based alias pairs:", nrow(id_auto_pairs), "\n")

id_based_aliases %>%
  group_by(Contributor.ID) %>%
  summarise(names = paste(unique(Contributor.Name), collapse = " | "), n = n(), .groups = "drop") %>%
  arrange(desc(n))
```

## Step 2: Exact variant cascade

```{r}
id_resolved_rows <- unique(c(id_auto_pairs$.x, id_auto_pairs$.y))
remaining_orgs   <- orgs %>% filter(!row_id %in% id_resolved_rows)

#variant_cols     <- c("name_v1", "name_v2", "name_v3", "name_v4", "name_v5", "name_v6")
variant_cols     <- c("name_v1", "name_v2", "name_v4", "name_v5") # don't do variants where we remove unit info or words attributed w/ PAC for now
exact_pairs_list <- list()

for (v in variant_cols) {
  new_pairs <- find_exact_variant_pairs(remaining_orgs, v)
  if (nrow(new_pairs) > 0) {
    exact_pairs_list[[v]] <- new_pairs
    remaining_orgs <- remaining_orgs %>%
      filter(!row_id %in% unique(c(new_pairs$.x, new_pairs$.y)))
  }
  cat("variant", v, "- new pairs:", nrow(new_pairs), "| remaining:", nrow(remaining_orgs), "\n")
}

exact_pairs <- bind_rows(exact_pairs_list) %>%
  left_join(orgs %>% select(row_id, id_x = Contributor.ID, has_id_x = has_contributor_id), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, id_y = Contributor.ID, has_id_y = has_contributor_id), by = c(".y" = "row_id")) %>%
  filter(!(coalesce(has_id_x, FALSE) & coalesce(has_id_y, FALSE) & coalesce(id_x != id_y, FALSE))) %>%
  mutate(link_tier = "auto_link") %>%
  select(.x, .y, link_tier, match_source)

cat("Exact cascade auto-links:", nrow(exact_pairs), "| remaining for fuzzy:", nrow(remaining_orgs), "\n")
```


### Step 3: internal fuzzy match


```{r}
remaining <- orgs %>%
  filter(
    !row_id %in% id_resolved_rows,
    !row_id %in% unique(c(exact_pairs$.x, exact_pairs$.y))
  )

ids <- remaining$row_id
pairs_idx <- combn(length(ids), 2)

org_pairs <- tibble(
  .x = ids[pairs_idx[1, ]],
  .y = ids[pairs_idx[2, ]]
) %>%
  left_join(orgs %>% select(row_id, name_x = name_v5, id_x = Contributor.ID,
                             has_id_x = has_contributor_id, pac_x = has_pac_language),
            by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = name_v5, id_y = Contributor.ID,
                             has_id_y = has_contributor_id, pac_y = has_pac_language),
            by = c(".y" = "row_id")) %>%
  filter(
    !(coalesce(has_id_x, FALSE) & coalesce(has_id_y, FALSE) & coalesce(id_x != id_y, FALSE)),
    pac_x == pac_y
  ) %>%
  mutate(
    name_jw     = stringsim(name_x, name_y, method = "jw", p = 0.1),
    name_cosine = bigram_cosine(name_x, name_y)
  ) %>%
  filter(name_jw >= 0.70, name_cosine >= 0.50) %>%
  select(-name_x, -name_y) %>%
  mutate(
    name_sim     = (name_jw + name_cosine) / 2,
    link_tier    = case_when(
      name_jw >= 0.999 ~ "auto_link",
      name_sim >= 0.90 ~ "auto_link", # reviewd this manually
      name_sim >= 0.85 ~ "review", # was 0.75, but no of those were really close..
      TRUE             ~ "no_link"
    ),
    match_source = "internal_fuzzy"
  )


```

### Inspect internal fuzzy matches

```{r}
org_fuzzy_inspect <- org_pairs %>%
  filter(link_tier != "no_link") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(link_tier, name_x, name_y, name_jw, name_cosine, name_sim,
         has_id_x, has_id_y, id_x, id_y) %>%
  arrange(link_tier, desc(name_sim))

org_fuzzy_inspect
```


```{r}
# just below review threshold
org_pairs %>%
  filter(name_jw >= 0.70, name_cosine >= 0.50, link_tier == "no_link") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(name_x, name_y, name_jw, name_cosine, name_sim) %>%
  arrange(desc(name_sim))

```

- don't think i want to adjust threshold b/c none of these look good

```{r}
# review
org_pairs %>%
  filter(link_tier == "auto_link") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(name_x, name_y, name_jw, name_cosine, name_sim) %>%
  arrange(desc(name_sim))

```

- one of the auto-links we defs don't want. only happening because one org is missing a contributor.id
- only two of these are worth reviewing and one will get flagged by another one of the matching methods anyways

```{r}
org_fuzzy_inspect %>%
  filter(link_tier == "review") %>%
  arrange(name_sim)

```


## Step 5: External bridge (map1 / map2)

- do i want to use name_v5 here or just use standardized name? 
  - not sure about removing business text, etc. b/c name variants will account for different endings
  - want to be really sure about matching to name variants.

```{r}
# standardize external names to name_v5 format for matching
build_bridge_lookup <- function(mapping) {
  mapping %>%
    mutate(name_std = standardize_names(record_name) %>%
             fix_typos() %>%
             #remove_pac_info() %>%
             #remove_unit_info() %>%
             replace_business_text() %>%
             str_remove_all("\\s*\\(.*?\\)") %>%
             str_squish()) %>%
    select(name_std, canonical_name) %>%
    filter(!is.na(name_std), name_std != "", !is.na(canonical_name), canonical_name != "")
}

map1_lookup <- build_bridge_lookup(map1)
map2_lookup <- build_bridge_lookup(map2)
```

```{r}
find_bridge_pairs <- function(orgs_df, lookup, source_label) {
  matched <- orgs_df %>%
    inner_join(lookup, by = c("name_v5" = "name_std")) %>%
    select(row_id, name_v5, canonical_name) %>%
    distinct(name_v5, canonical_name, .keep_all = TRUE)  # one row per unique name × canonical (used to be row_id instead of name_v5 here)

  if (nrow(matched) < 2) return(tibble(.x = integer(), .y = integer(),
                                        match_source = character(), canonical_name = character()))

  matched %>%
    rename(.x = row_id, name_v5_x = name_v5) %>%
    inner_join(matched %>% rename(.y = row_id, name_v5_y = name_v5), by = "canonical_name") %>%
    filter(.x < .y, name_v5_x != name_v5_y) %>%
    group_by(.x, .y) %>%
    mutate(
      n_canonicals = n_distinct(canonical_name),
      needs_review = n_canonicals > 1
    ) %>%
    ungroup() %>%
    mutate(match_source = source_label) %>%
    select(.x, .y, match_source, canonical_name, n_canonicals, needs_review)
}

```

```{r}
bridge_map1 <- find_bridge_pairs(orgs, map1_lookup, "map1_bridge")
bridge_map2 <- find_bridge_pairs(orgs, map2_lookup, "map2_bridge")
cat("Bridge pairs — map1:", nrow(bridge_map1), "| map2:", nrow(bridge_map2), "\n")
```

```{r, eval=FALSE}
bridge_pairs <- bind_rows(bridge_map1, bridge_map2) %>%
  mutate(
    .x = pmin(.x, .y),
    .y = pmax(.x, .y)
  ) %>%
  distinct(.x, .y, canonical_name, .keep_all = TRUE)

bridge_pairs %>%
  # filter(needs_review) %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(name_x, name_y, canonical_name, n_canonicals) %>%
  arrange(name_x, name_y)

```


```{r}
#x[which(x$Contributor.Name == "Local 770 United Food and Commercial Workers PAC"),]$Contributor.ID
#x[which(x$Contributor.Name == "United Food and Commercial Workers Active Ballot Club (California) (Fed PAC ID# C00002766)"),]$Contributor.ID

#x[grep("association of health",tolower(x$Contributor.Name)),]
```


### combine bridge pairs

```{r}
# combine bridge pairs
# log when both sources agree
bridge_pairs <- bind_rows(bridge_map1, bridge_map2) %>%
  group_by(.x, .y) %>%
  summarise(
    match_source   = paste(sort(unique(match_source)), collapse = "+"),
    canonical_name = paste(sort(unique(canonical_name)), collapse = " / "),
    .groups = "drop"
  )

# canonical ID per name_v5: if ANY contribution row for that name has an ID, record it
# used detect conflicting-ID pairs regardless of which row_id find_bridge_pairs picked.
org_name_canonical_id <- orgs %>%
  filter(has_contributor_id) %>%
  group_by(name_v5) %>%
  summarise(canonical_id = names(sort(table(Contributor.ID), decreasing = TRUE))[1],
            .groups = "drop")

# score bridge pairs on internal similarity - keep?
bridge_scored <- bridge_pairs %>%
  left_join(orgs %>% select(row_id, name_x = name_v5, pac_x = has_pac_language),
            by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = name_v5, pac_y = has_pac_language),
            by = c(".y" = "row_id")) %>%
  # join on name, not row_id in case a row was picked that didn't have a contributor.id
  left_join(org_name_canonical_id %>% rename(name_x = name_v5, id_x = canonical_id), by = "name_x") %>%
  left_join(org_name_canonical_id %>% rename(name_y = name_v5, id_y = canonical_id), by = "name_y") %>% # drop any pair where both orgs have IDs and those IDs differ — they are distinct entities
  filter(!(coalesce(!is.na(id_x), FALSE) & coalesce(!is.na(id_y), FALSE) & coalesce(id_x != id_y, FALSE))) %>%

  mutate(
    name_jw     = stringsim(name_x, name_y, method = "jw", p = 0.1),
    name_cosine = bigram_cosine(name_x, name_y),
    name_sim    = (name_jw + name_cosine) / 2
  ) %>%
  select(-name_x, -name_y, -id_x, -id_y)

# tag internal fuzzy paris thr bridge found— for match_source logging
internal_pair_keys <- org_pairs %>%
  filter(link_tier != "no_link") %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y)))

bridge_pair_keys <- bridge_scored %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  select(pair_key, bridge_source = match_source)

# upgrade internal pairs corroborated by a bridge
BRIDGE_UPGRADE_THRESHOLD <- 0.65

org_pairs <- org_pairs %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  left_join(bridge_pair_keys, by = "pair_key") %>%
  mutate(
    match_source = if_else(!is.na(bridge_source),
                           paste(match_source, bridge_source, sep = "+"),
                           match_source),
    link_tier = case_when(
      link_tier == "review" & !is.na(bridge_source) & name_sim >= BRIDGE_UPGRADE_THRESHOLD ~ "auto_link",
      TRUE ~ link_tier
    )
  ) %>%
  select(-pair_key, -bridge_source)

# bridge-only pairs
bridge_only <- bridge_scored %>%
  mutate(pair_key = paste(pmin(.x, .y), pmax(.x, .y))) %>%
  filter(!pair_key %in% internal_pair_keys$pair_key) %>%
  select(-pair_key) %>%
  mutate(
    link_tier = "review"  # bridge alone never auto-links
  )

cat("Bridge-corroborated internal pairs:", sum(!is.na(bridge_pair_keys$bridge_source[match(internal_pair_keys$pair_key, bridge_pair_keys$pair_key)])), "\n")
cat("Bridge-only new pairs for review:", nrow(bridge_only), "\n")
```

```{r}
bridge_only
```



## Step 4: ID propagation — where one side is missing an ID - this should happen earlier in the process 

**not sure i should do this based on fuzzy matches, but will do for now just to look at review process**

```{r}
# exact cascade — needs all columns joined
exact_id_propagation <- bind_rows(exact_pairs_list) %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name, std_name_x = name_v5,
                             id_x = Contributor.ID, has_id_x = has_contributor_id), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name, std_name_y = name_v5,
                             id_y = Contributor.ID, has_id_y = has_contributor_id), by = c(".y" = "row_id")) %>%
  filter(coalesce(has_id_x, FALSE) != coalesce(has_id_y, FALSE))

# fuzzy — id_x, id_y, has_id_x, has_id_y already in org_pairs; only join names
fuzzy_id_propagation <- org_pairs %>%
  filter(link_tier %in% c("auto_link", "review")) %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name, std_name_x = name_v5), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name, std_name_y = name_v5), by = c(".y" = "row_id")) %>%
  filter(coalesce(has_id_x, FALSE) != coalesce(has_id_y, FALSE))


id_propagation <- bind_rows(
  exact_id_propagation %>% mutate(confidence = "high"),
  fuzzy_id_propagation %>% mutate(confidence = "review")
) %>%
  mutate(
    source_id       = coalesce(id_x, id_y),
    name_with_id    = if_else(coalesce(has_id_x, FALSE), name_x, name_y),
    name_missing_id = if_else(coalesce(has_id_x, FALSE), name_y, name_x),
    std_with_id     = if_else(coalesce(has_id_x, FALSE), std_name_x, std_name_y),
    std_missing_id  = if_else(coalesce(has_id_x, FALSE), std_name_y, std_name_x),
    row_missing_id  = if_else(coalesce(has_id_x, FALSE), .y, .x)
  ) %>%
  select(row_missing_id, name_missing_id, std_missing_id, name_with_id, std_with_id,
         source_id, match_source, confidence)

cat("IDs to propagate:", nrow(id_propagation), "\n")
id_propagation

```


## Combine + cluster

Combine all links, including those that need to be reviewed into a single entity
- not including fuzzy pairs and bridge pairs for now

```{r}
# pairs computed through different linkage methods
all_org_pairs <- bind_rows(
  id_auto_pairs %>% select(.x, .y, link_tier, match_source),
  exact_pairs,
  #org_pairs    %>% select(.x, .y, link_tier, match_source, name_jw, name_cosine, name_sim),
  #bridge_only  %>% select(.x, .y, link_tier, match_source, name_jw, name_cosine, name_sim, canonical_name)
)

# build one graph from all non-rejected pairs, treating review links as true
linkable_pairs <- all_org_pairs %>% filter(link_tier %in% c("auto_link", "review"))

if (nrow(linkable_pairs) > 0) {
  g_orgs       <- graph_from_data_frame(linkable_pairs %>% select(.x, .y), directed = FALSE)
  org_clusters <- components(g_orgs)$membership
  org_cluster_map <- tibble(row_id = as.integer(names(org_clusters)), cluster_num = org_clusters) %>%
    left_join(orgs %>% select(row_id, Contributor.ID, name_v5), by = "row_id") %>%
    group_by(cluster_num) %>%
    mutate(member_key = paste0(coalesce(as.character(Contributor.ID), ""), "::", name_v5),
           entity_id  = hash_id(member_key, "org")) %>%
    ungroup() %>%
    select(row_id, entity_id)
} else {
  org_cluster_map <- tibble(row_id = integer(), entity_id = character())
}

orgs_singletons <- orgs %>%
  filter(!row_id %in% org_cluster_map$row_id) %>%
  mutate(member_key = paste0(coalesce(as.character(Contributor.ID), ""), "::", name_v5),
         entity_id  = vapply(member_key, hash_id, character(1), prefix = "org")) %>%
  select(row_id, entity_id)

review_org_entity_ids <- all_org_pairs %>%
  filter(link_tier == "review") %>%
  left_join(bind_rows(org_cluster_map, orgs_singletons), by = c(".x" = "row_id"),
            relationship = "many-to-one") %>%
  filter(!is.na(entity_id)) %>%
  pull(entity_id) %>%
  unique()

orgs_resolved <- orgs %>%
  left_join(bind_rows(org_cluster_map, orgs_singletons), by = "row_id") %>%
  mutate(needs_review = entity_id %in% review_org_entity_ids)

```

- update to also include match_method for those that need a review and reason for review - could be id propagation or bridge....

```{r}
orgs_resolved %>% filter(needs_review == TRUE) %>% arrange(entity_id) %>% select(entity_id,Contributor.Name)
```


```{r}
# inspect resolved clusters (multi-member only)
orgs_resolved %>%
  group_by(entity_id) %>%
  filter(n() > 1) %>%
  summarise(members = paste(Contributor.Name, collapse = " | "), n = n(), .groups = "drop") %>%
  arrange(desc(n))
```


```{r}
all_org_pairs %>%
  filter(link_tier == "auto_link") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(name_x, name_y, match_source)

```


```{r}
all_org_pairs %>%
  filter(link_tier == "review") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name), by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name), by = c(".y" = "row_id")) %>%
  select(name_x, name_y, match_source)

```


- some that should not match due to local unit nums: e.g. Northern California District Council ILWU (no id) vs Marine Clerks ILWU Local 63- PAC Small Contributor Committee (these were bridged)
- Laborer's Union Local No. 777 PAC: doesn't have an id and it should, so its matching to other PACs w/ diff local unit numbers
- look for names in id_map instead of contributions_full


# Individuals

- Uninformative vals
  - look at whether blank is ever in there for individuals (employer or occupation)
- need to tune what counts as a common name
- take self employed out of uninfomative_vals? its not capturing all insteances of self employed individuals
  
```{r}
# ── Constants ─────────────────────────────────────────────────────────────────
UNINFORMATIVE_VALS        <- c("UNKNOWN", "RETIRED", "NONE", "SELF EMPLOYED") 
FIRST_NAME_MIN            <- 0.85
MATCH_THRESHOLD           <- 0.90
STRICT_THRESHOLD          <- 0.94  # both employer & occupation uninformative
COMMON_NAME_THRESHOLD     <- 0.93  # common name, employer informative
COMMON_NAME_STRICT        <- 0.97  # common name + both uninformative
UNCERTAIN_LOW             <- 0.75
EMPLOYER_CONFLICT_SIM     <- 0.70  # employer sim below this = conflict
#COMMON_NAME_MIN_COUNT     <- 3     # names appearing >= N times treated as common
RARE_NAME_UNCERTAIN_LOW <- 0.65  # tune based on what you see in the review queue
COMMON_LAST_MIN           <- 3
COMMON_FIRST_MIN          <- 3
RARE_NAME_UNCERTAIN_LOW   <- 0.65


```


```{r,eval = FALSE}
# compute name frequency

# parse distinct name strings from full dataset
full_names_parsed <- contributions_full %>%
  mutate(
    standardized_name = standardize_names(Contributor.Name),
    entity_type = if_else(sapply(standardized_name, is_individual), "individual", "organization")
  ) %>%
  filter(entity_type == "individual") %>%
  distinct(standardized_name) %>%
  bind_cols(parse_names_df(.$standardized_name)) %>%
  filter(!is.na(last), !is.na(first))

last_name_freq  <- full_names_parsed %>% count(last,  name = "last_name_freq")
first_name_freq <- full_names_parsed %>% count(first, name = "first_name_freq")

```

```{r}
full_names_parsed <- read.csv("08_inputs/full_names_parsed.csv")
last_name_freq  <- full_names_parsed %>% count(last,  name = "last_name_freq")
first_name_freq <- full_names_parsed %>% count(first, name = "first_name_freq")
```


```{r}
#last_name_freq %>% arrange(desc(last_name_freq))
#first_name_freq %>% arrange(desc(first_name_freq))
```


```{r}
# parse names 
parsed_cols <- parse_names_df(
  contributions %>% filter(entity_type == "individual") %>% pull(standardized_name) # use standardized name here because it has the commas still in it
)

individuals <- contributions %>%
  filter(entity_type == "individual") %>%
  bind_cols(parsed_cols %>% rename(last_name = last, first_name = first)) %>%
  mutate(
    original_row_id = row_id,       # preserve for final join back to x
    row_id          = row_number(), # sequential for pair_blocking compatibility
    first_initial   = substr(first_name, 1, 1),
    last_meta       = safe_metaphone(last_name)
  ) %>%
  filter(!is.na(last_meta), last_meta != "")

```


### temporary 

- group by individual and employer OR group by invidividual, employer and city (and occupation?)
- tricky to group people who are not employed...individuals are hard to determine


```{r}
individuals %>% group_by(standardized_name,processed_employer_name) %>% mutate(count = n(), n_city = n_distinct(standardized_city), total = sum(Amount), n_occ = n_distinct(processed_occupation)) %>% filter(n_occ > 1, total >= 5000) %>% select(Contributor.Name,Contributor.Employer,Contributor.City,Contributor.Occupation,Amount) %>%
  arrange(standardized_name)
```
- St. -> Saint
- Oaklans -> Oakland
- make sure city is removing zip code
- feel fine with all the people with diff cities, but same name and employer -> not many
- retired vs not employed -> don't want to capture that difference i don't think
  - this seems to account for the vast majority of cases
- don't want to aggregate based on occupation -> that could miss people who are the same
- but make sure we classify people with different occupations differently when they are self-employed
- need to double check how this is being handled.... 
- think we just group by employer and occupation for now.


Robert, Anderson Santa Rosa vs SF, writer vs Physician seems like a different person... 

### record linkage algorithm

```{r}
# scoring function
compare_and_score_indiv <- function(pairs_obj, df) {
  pairs_obj %>%
    compare_pairs(
      on = c("last_name", "first_name", "zip_code_processed",
             "processed_employer_name", "processed_occupation", "standardized_city"),
      default_comparator = cmp_jarowinkler(),
      comparators = list(
        processed_employer_name = function(a, b) stringsim(a, b, method = "cosine", q = 2),
        zip_code_processed      = cmp_identical(),
        processed_occupation    = cmp_identical()
      )
    ) %>%
    as_tibble() %>%
    # join raw employer/occupation values
    left_join(df %>% select(row_id, emp_a = processed_employer_name, occ_a = processed_occupation,
                             city_a = standardized_city),
              by = c(".x" = "row_id")) %>%
    left_join(df %>% select(row_id, emp_b = processed_employer_name, occ_b = processed_occupation,
                             city_b = standardized_city),
              by = c(".y" = "row_id")) %>%
    mutate(
      emp_a_uninf = emp_a %in% UNINFORMATIVE_VALS | is.na(emp_a) | emp_a == "",
      emp_b_uninf = emp_b %in% UNINFORMATIVE_VALS | is.na(emp_b) | emp_b == "",
      occ_a_uninf = occ_a %in% UNINFORMATIVE_VALS | is.na(occ_a) | occ_a == "",
      occ_b_uninf = occ_b %in% UNINFORMATIVE_VALS | is.na(occ_b) | occ_b == "",
      w_emp  = if_else(emp_a_uninf & emp_b_uninf, 0,    0.25),
      w_occ  = if_else(occ_a_uninf & occ_b_uninf, 0,    0.10),
      w_tot  = 0.30 + 0.25 + w_emp + w_occ + 0.07 + 0.03,
      emp_sim = if_else(emp_a_uninf | emp_b_uninf, 0, coalesce(processed_employer_name, 0)),
      occ_sim = if_else(occ_a_uninf | occ_b_uninf, 0, coalesce(processed_occupation, 0)),
      score  = (0.30 * coalesce(last_name, 0) +
                0.25 * coalesce(first_name, 0) +
                w_emp * emp_sim +
                w_occ * occ_sim +
                0.07 * coalesce(zip_code_processed, 0) +
                0.03 * coalesce(standardized_city, 0)) / w_tot,
      both_uninformative = (emp_a_uninf & emp_b_uninf) & (occ_a_uninf & occ_b_uninf),
    
      employer_substring = !emp_a_uninf & !emp_b_uninf & mapply(
        function(a, b) {
          if (is.na(a) || is.na(b) || a == "" || b == "") return(FALSE)
          a <- toupper(a); b <- toupper(b)
          grepl(b, a, fixed = TRUE) || grepl(a, b, fixed = TRUE)
        },
        emp_a, emp_b,
        USE.NAMES = FALSE
      ),
    
      employer_conflict  = !emp_a_uninf & !emp_b_uninf &
                           !employer_substring &
                           coalesce(processed_employer_name, 0) < EMPLOYER_CONFLICT_SIM,
    
      # city conflict: both cities are known and different (JW < 0.85)
      city_conflict = (!is.na(city_a) & city_a != "") &
                      (!is.na(city_b) & city_b != "") &
                      coalesce(standardized_city, 1) < 0.85
    ) %>%
    select(-occ_a, -occ_b,
           -emp_a_uninf, -emp_b_uninf, -occ_a_uninf, -occ_b_uninf,
           -w_emp, -w_occ, -w_tot, -emp_sim, -occ_sim,
           -city_a, -city_b)
}
```

```{r}
# blocking + scoring
pairs_zip  <- pair_blocking(individuals, on = c("last_meta", "zip_code_processed"), deduplication = TRUE)
pairs_init <- pair_blocking(individuals, on = c("last_meta", "first_initial"),       deduplication = TRUE)

indiv_pairs <- bind_rows(
  compare_and_score_indiv(pairs_zip,  individuals),
  compare_and_score_indiv(pairs_init, individuals)
) %>%
  group_by(.x, .y) %>%
  slice_max(score, n = 1, with_ties = FALSE) %>%
  ungroup()


# name frequency adjustment
indiv_pairs <- indiv_pairs %>%
  left_join(individuals %>% select(row_id, last_name_str = last_name, first_name_str = first_name),
            by = c(".x" = "row_id")) %>%
  left_join(last_name_freq,  by = c("last_name_str"  = "last")) %>%
  left_join(first_name_freq, by = c("first_name_str" = "first")) %>%
  mutate(
    common_last  = coalesce(last_name_freq,  1L) >= COMMON_LAST_MIN,
    common_first = coalesce(first_name_freq, 1L) >= COMMON_FIRST_MIN,
    common_name  = common_last & common_first,
    rare_name    = !common_last & !common_first,
    match_threshold = case_when(
      common_name & both_uninformative ~ COMMON_NAME_STRICT,
      common_name                      ~ COMMON_NAME_THRESHOLD,
      both_uninformative               ~ STRICT_THRESHOLD,
      TRUE                             ~ MATCH_THRESHOLD
    ),
    # unusual names flagged for review at a lower score bar
    review_threshold = if_else(rare_name, RARE_NAME_UNCERTAIN_LOW, UNCERTAIN_LOW)
  ) %>%
  select(-last_name_str, -first_name_str, -last_name_freq, -first_name_freq)

# link tiers
indiv_pairs <- indiv_pairs %>%
  mutate(
    link_tier = case_when(
      coalesce(first_name, 0) < FIRST_NAME_MIN ~ "no_link",
      common_name & both_uninformative ~ "no_link",
      city_conflict & both_uninformative ~ "no_link",
      common_name & city_conflict ~ "no_link",       # city disagreement is enough to reject common names
      common_name & employer_conflict ~ "no_link",   # clearly different employers also enough for common names
      score < review_threshold ~ "no_link",
      score >= match_threshold ~ "auto_link",
      TRUE                                           ~ "review"
    ),
    match_source = "internal_fuzzy"
  )

cat("Individual pairs — auto_link:", sum(indiv_pairs$link_tier == "auto_link"),
    "| review:", sum(indiv_pairs$link_tier == "review"),
    "| employer conflicts in review:", sum(indiv_pairs$employer_conflict & indiv_pairs$link_tier == "review"), "\n")

```


## Cluster individuals

```{r}
# treat all links as true, review after classifications
linkable_indiv_pairs <- indiv_pairs %>% filter(link_tier %in% c("auto_link", "review"))

if (nrow(linkable_indiv_pairs) > 0) {
  g_indiv        <- graph_from_data_frame(linkable_indiv_pairs %>% select(.x, .y), directed = FALSE)
  indiv_clusters <- components(g_indiv)$membership
  indiv_cluster_map <- tibble(row_id = as.integer(names(indiv_clusters)), cluster_num = indiv_clusters) %>%
    left_join(individuals %>% select(row_id, Contributor.ID, standardized_name), by = "row_id") %>%
    mutate(member_key = paste0(coalesce(as.character(Contributor.ID), ""), "::", standardized_name)) %>%
    group_by(cluster_num) %>%
    mutate(entity_id = hash_id(member_key, "indiv")) %>%
    ungroup() %>%
    left_join(individuals %>% select(row_id, original_row_id), by = "row_id") %>%
    select(row_id = original_row_id, entity_id)
} else {
  indiv_cluster_map <- tibble(row_id = integer(), entity_id = character())
}

indiv_singletons <- individuals %>%
  filter(!original_row_id %in% indiv_cluster_map$row_id) %>%
  mutate(member_key = paste0(coalesce(as.character(Contributor.ID), ""), "::", standardized_name),
         entity_id  = vapply(member_key, hash_id, character(1), prefix = "indiv")) %>%
  select(row_id = original_row_id, entity_id)

# within-individuals row_id to entity_id map (needed for needs_review flag)
# indiv_pairs uses within-individuals row_id; cluster_map was renamed to original_row_id
indiv_internal_to_entity <- individuals %>%
  select(row_id, original_row_id) %>%
  left_join(bind_rows(indiv_cluster_map, indiv_singletons), by = c("original_row_id" = "row_id"))

review_indiv_entity_ids <- indiv_pairs %>%
  filter(link_tier == "review") %>%
  left_join(indiv_internal_to_entity %>% select(row_id, entity_id), by = c(".x" = "row_id"),
            relationship = "many-to-one") %>%
  filter(!is.na(entity_id)) %>%
  pull(entity_id) %>%
  unique()

# use longer employer name for pairs when one employer is the exact substring of another
employer_normalization <- indiv_pairs %>%
  filter(employer_substring) %>%
  mutate(longer = if_else(nchar(coalesce(emp_a, "")) >= nchar(coalesce(emp_b, "")), emp_a, emp_b)) %>%
  select(.x, .y, longer) %>%
  pivot_longer(c(.x, .y), names_to = "side", values_to = "row_id") %>%
  group_by(row_id) %>%
  slice_max(nchar(coalesce(longer, "")), n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(row_id, normalized_employer = longer)


individuals_resolved <- individuals %>%
  left_join(bind_rows(indiv_cluster_map, indiv_singletons), by = c("original_row_id" = "row_id")) %>%
  mutate(needs_review = entity_id %in% review_indiv_entity_ids)

individuals_resolved <- individuals_resolved %>%
  left_join(employer_normalization, by = "row_id") %>%
  mutate(classification_employer = coalesce(normalized_employer, processed_employer_name))


```


# CHOOSE CANONICAL NAMES

```{r}
entity_canonical <- bind_rows(

  # orgs: use longest with contributor id; else longest overall
  orgs_resolved %>%
    group_by(entity_id) %>%
    summarise(
      canonical_name = {
        with_id <- Contributor.Name[has_contributor_id]
        if (length(with_id) > 0)
          with_id[which.max(nchar(with_id))]
        else
          Contributor.Name[which.max(nchar(Contributor.Name))]
      },
      canonical_id = if (any(has_contributor_id))
        names(sort(table(Contributor.ID[has_contributor_id]), decreasing = TRUE))[1]
      else NA_character_,
      entity_type = "organization",
      .groups = "drop"
    ),

  # individuals: use longest name with most complete info; else longest overall
  individuals_resolved %>%
    group_by(entity_id) %>%
    summarise(
      canonical_name = {
        has_info   <- !processed_employer_name %in% c("UNKNOWN", "NONE", "RETIRED", "", NA) &
                       !is.na(standardized_city) & standardized_city != ""
        candidates <- Contributor.Name[has_info]
        if (length(candidates) > 0)
          candidates[which.max(nchar(candidates))]
        else
          Contributor.Name[which.max(nchar(Contributor.Name))]
      },
      canonical_id = NA_character_,
      entity_type  = "individual",
      .groups = "drop"
    )
)

```



# PRODUCE FINAL OUTPUT

uninformative employer vals treated as not counting toward multiple employers
- could drop uninformative employer vals filter in employer_units

```{r}

#UNINFORMATIVE_EMPLOYER_VALS <- c("UNKNOWN", "RETIRED", "NONE", "SELF EMPLOYED", "", NA)

parse_date_safe <- function(x) suppressWarnings(as.Date(x, format = "%Y-%m-%d"))

all_members <- bind_rows(
  orgs_resolved %>%
    transmute(
      entity_id, contribution_id, needs_review,
      entity_type_raw = "organization",
      has_pac_language,
      raw_name = Contributor.Name,
      normalized_name = name_v5, # Do i want to do standardized name instead? 
      #blocking_key = block_key,
      raw_employer = NA_character_, raw_occupation = NA_character_,
      processed_employer_name = NA_character_, processed_occupation = NA_character_,
      raw_city = Contributor.City, raw_state = Contributor.State, raw_zip = Contributor.Zip.Code,
      contribution_amount = Amount,
      race_or_proposition_id = race_prop,
      event_date = parse_date_safe(Start.Date)
    ),
  individuals_resolved %>%
    transmute(
      entity_id, contribution_id, needs_review,
      entity_type_raw = "individual",
      has_pac_language = FALSE,
      raw_name = Contributor.Name,
      normalized_name = standardized_name,
      #blocking_key = paste(last_meta, zip_code_processed, first_initial, sep = "|"),
      raw_employer = Contributor.Employer, raw_occupation = Contributor.Occupation,
      processed_employer_name, processed_occupation,
      raw_city = Contributor.City, raw_state = Contributor.State, raw_zip = Contributor.Zip.Code,
      contribution_amount = Amount,
      race_or_proposition_id = race_prop,
      event_date = parse_date_safe(Start.Date)
    )
)

```



## entity contribution totals by race

```{r}
entity_race_totals <- bind_rows(
  orgs_resolved        %>% select(entity_id, cont_row_id),
  individuals_resolved %>% select(entity_id, cont_row_id = original_row_id)
) %>%
  left_join(contributions %>% select(row_id, race_prop, Amount), by = c("cont_row_id" = "row_id")) %>%
  group_by(entity_id, race_prop) %>%
  summarise(entity_total = sum(Amount, na.rm = TRUE), .groups = "drop")

entity_max_amounts <- entity_race_totals %>%
  group_by(entity_id) %>%
  summarise(max_race_total = max(entity_total, na.rm = TRUE), .groups = "drop")

```

### review matching - individuals

```{r}
indiv_5k_row_ids <- individuals_resolved %>%
  filter(entity_id %in% (entity_max_amounts %>%
           filter(max_race_total >= 5000) %>% pull(entity_id))) %>%
  pull(row_id)

pair_detail_cols <- c("name_x", "name_y", "employer_x", "employer_y",
                      "occ_x", "occ_y", "city_x", "city_y",
                      "score", "match_threshold", "employer_conflict",
                      "common_name", "both_uninformative")

add_names <- function(pairs) {
  pairs %>%
    left_join(individuals %>% select(row_id, name_x = processed_name,
                                      employer_x = processed_employer_name,
                                      occ_x = processed_occupation,
                                      city_x = standardized_city,
                                     zip_x = zip_code_processed),
              by = c(".x" = "row_id")) %>%
    left_join(individuals %>% select(row_id, name_y = processed_name,
                                      employer_y = processed_employer_name,
                                      occ_y = processed_occupation,
                                      city_y = standardized_city,
                                     zip_y = zip_code_processed),
              by = c(".y" = "row_id"))
}

# review individuals that meet the $5k threshold
indiv_review_5k <- indiv_pairs %>%
  filter(link_tier == "review",
         .x %in% indiv_5k_row_ids | .y %in% indiv_5k_row_ids) %>%
  add_names() %>%
  select(all_of(pair_detail_cols)) %>%
  arrange(desc(score))

# auto-links that meet $5k threshold
indiv_auto_5k <- indiv_pairs %>%
  filter(link_tier == "auto_link",
         .x %in% indiv_5k_row_ids | .y %in% indiv_5k_row_ids) %>%
  add_names() %>%
  select(all_of(pair_detail_cols)) %>%
  arrange(score)

```

```{r,eval=FALSE}
indiv_review_5k %>% arrange(score)
#indiv_auto_5k %>% arrange(score) %>% distinct()
```

- commas in names that shouldn't be there
- are still seeing some with scores of 0.75 that are clearly the same person
- employer and occupation both informative, should never tolerate different cities, esp when either name is common
  - Cynthia Anderson is a case where they probs are the same person, but don't think we can make that assumption.. 
  - bill bloomfield is another one with a city conlflict, commong name, no employer or occupation info, but probs the same person
- UC DAVIS vs University of California Davis
- UCLA vs University of California

```{r}
#contributions %>% filter(grepl("Van Nest, Robert",Contributor.Name))
contributions[which(contributions$processed_employer_name == "N/A"),]
```
- this should not be the case...

### review matching -- orgs


```{r}
# org review queue
org_review_queue <- all_org_pairs %>%
  filter(link_tier == "review") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name, city_x = Contributor.City,
                             state_x = Contributor.State),
            by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name, city_y = Contributor.City,
                             state_y = Contributor.State),
            by = c(".y" = "row_id")) %>%
  left_join(orgs_resolved %>% distinct(row_id, entity_id), by = c(".x" = "row_id"),
            relationship = "many-to-one") %>%
  left_join(entity_max_amounts, by = "entity_id") %>%
  select(.x, .y, name_x, name_y, city_x, city_y, state_x, state_y,
         match_source, name_jw, name_cosine, name_sim, canonical_name,
         entity_id, max_race_total) %>%
  arrange(desc(coalesce(name_sim, 0)))

# filter to $5k entities
org_review_queue %>%
  filter(coalesce(max_race_total, 0) >= 5000) %>%
  select(name_x, name_y, city_x, city_y, match_source,
         name_jw, name_cosine, name_sim, canonical_name, max_race_total)

# auto-links
all_org_pairs %>%
  filter(link_tier == "auto_link") %>%
  left_join(orgs %>% select(row_id, name_x = Contributor.Name, city_x = Contributor.City),
            by = c(".x" = "row_id")) %>%
  left_join(orgs %>% select(row_id, name_y = Contributor.Name, city_y = Contributor.City),
            by = c(".y" = "row_id")) %>%
  select(name_x, name_y, city_x, city_y, match_source, name_jw, name_cosine, name_sim, canonical_name) %>%
  arrange(coalesce(name_sim, 1))  # lowest confidence first

```

```{r,eval=FALSE}
contributions[which(contributions$Contributor.Name == "IBEW Local 684 PAC"),]
#contributions[which(contributions$Contributor.Name == "IBEW Local 6 Political Account"),]
contributions[which(contributions$Contributor.Name == "IBEW PAC Educational Fund (FEC ID #C00027342)"),]
```



## entity 

Note that contribution_id is not necessarily unique to contributions. 
- can be more than one contribution with the exact same info.

```{r}
entity_aliases_out <- all_members %>%
  #filter(!entity_id %in% flagged_entity_ids) %>%
  group_by(entity_id, raw_name, normalized_name) %>% # , blocking_key
  summarise(
    occurrence_count = n(),
    source_contribution_ids = paste(contribution_id, collapse = ";"),
    .groups = "drop"
  ) %>%
  mutate(alias_id = hash_id(paste(entity_id, raw_name), "alias")) %>%
  select(alias_id, entity_id, raw_name, normalized_name, 
         occurrence_count, source_contribution_ids) # blocking_key

#write_csv(entity_aliases_out, "entity_aliases.csv")

```


## entity attributes

- not sure about using contribution_id here

```{r}
build_attribute_rows <- function(df, attr_type, raw_col, norm_col) {
  df %>%
    #filter(!entity_id %in% flagged_entity_ids, !is.na(.data[[raw_col]]), .data[[raw_col]] != "") %>%
    filter(!is.na(.data[[raw_col]]), .data[[raw_col]] != "") %>%
    group_by(entity_id, value = .data[[raw_col]], normalized_value = .data[[norm_col]]) %>%
    summarise(
      occurrence_count = n(),
      total_amount_for_this_value = sum(contribution_amount, na.rm = TRUE),
      first_seen_contribution_id = contribution_id[which.min(event_date)],
      last_seen_contribution_id  = contribution_id[which.max(event_date)],
      .groups = "drop"
    ) %>%
    mutate(attribute_type = attr_type, .before = value)
}

entity_attributes <- bind_rows(
  build_attribute_rows(all_members, "city",  "raw_city",  "raw_city"),
  build_attribute_rows(all_members, "state", "raw_state", "raw_state"),
  build_attribute_rows(all_members, "zip", "raw_zip", "raw_zip"),
  build_attribute_rows(all_members %>% filter(entity_type_raw == "individual"),
                        "employer",   "raw_employer",   "processed_employer_name"),
  build_attribute_rows(all_members %>% filter(entity_type_raw == "individual"),
                        "occupation", "raw_occupation", "processed_occupation")
)

#write_csv(entity_attributes, "entity_attributes.csv")

```


# ENTITIES OUTPUT


**change has pac language to has contributor id**


```{r}
entity_type_lookup <- all_members %>%
  distinct(entity_id, entity_type_raw, has_pac_language) %>%
  group_by(entity_id) %>%
  summarise(
    entity_type = case_when(
      first(entity_type_raw) == "individual" ~ "individual",
      any(has_pac_language)                   ~ "committee", # want to do has contributor id here
      TRUE                                     ~ "org"
    ),
    .groups = "drop"
  )

entity_totals <- all_members %>%
  group_by(entity_id) %>%
  summarise(total_contribution_amount = sum(contribution_amount, na.rm = TRUE),
            cluster_size = n(), .groups = "drop")

entity_needs_review <- all_members %>%
  group_by(entity_id) %>%
  summarise(needs_review = any(needs_review), .groups = "drop")

entity_crosses_threshold <- entity_max_amounts %>%
  transmute(entity_id, crosses_threshold = max_race_total >= 5000)


entity_race_totals_nested <- entity_race_totals %>%
  rename(race_or_proposition_id = race_prop, total_amount = entity_total) %>%
  mutate(crosses_threshold = total_amount >= 5000) %>%
  tidyr::nest(race_totals = c(race_or_proposition_id, total_amount, crosses_threshold))

entity_aliases_nested <- all_members %>%
  distinct(entity_id, raw_name, normalized_name, contribution_id) %>% # blocking_key
  group_by(entity_id, raw_name, normalized_name) %>% # blocking_key
  summarise(occurrence_count = n(),
            source_contribution_ids = paste(contribution_id, collapse = ";"),
            .groups = "drop") %>%
  tidyr::nest(aliases = c(raw_name, normalized_name,
                          occurrence_count, source_contribution_ids)) #blocking_key

# change these to be normalize version of city and state
entity_attributes_nested <- bind_rows(
  build_attribute_rows(all_members, "city",  "raw_city",  "raw_city"),
  build_attribute_rows(all_members, "state", "raw_state", "raw_state"),
  build_attribute_rows(all_members, "zip", "raw_zip", "raw_zip"),
  build_attribute_rows(all_members %>% filter(entity_type_raw == "individual"),
                        "occupation", "raw_occupation", "processed_occupation")
) %>%
  tidyr::nest(attributes = c(attribute_type, value, normalized_value, occurrence_count,
                             total_amount_for_this_value,
                             first_seen_contribution_id, last_seen_contribution_id))


# incorporate when available
#prior_entity_classifications <- tryCatch(
#  read_csv("entity_classifications.csv", show_col_types = FALSE),
#  error = function(e) tibble(entity_id = character(), industry_code = character(),
#                              confidence = double(), method = character())
#)


entities <- entity_canonical %>%
  select(entity_id, canonical_name) %>%
  left_join(entity_type_lookup,        by = "entity_id") %>%
  left_join(entity_totals,             by = "entity_id") %>%
  left_join(entity_crosses_threshold,  by = "entity_id") %>%
  left_join(entity_needs_review,       by = "entity_id") %>%
  left_join(entity_aliases_nested,     by = "entity_id") %>%
  left_join(entity_attributes_nested,  by = "entity_id") %>%
  left_join(entity_race_totals_nested, by = "entity_id") %>%
  #left_join(prior_entity_classifications, by = "entity_id") %>%
  mutate(
    crosses_threshold  = coalesce(crosses_threshold, FALSE),
    needs_review       = coalesce(needs_review, FALSE),
    status             = "active",
    name_for_matching  = if_else(entity_type == "individual", NA_character_, canonical_name),
    #industry_code      = if_else(entity_type == "individual", NA_character_, industry_code),
    #already_classified = if_else(entity_type == "individual", NA,           !is.na(industry_code))
  ) %>%
  select(entity_id, entity_type, canonical_name, status, cluster_size,
         total_contribution_amount, crosses_threshold, needs_review,
         name_for_matching, 
         aliases, attributes, race_totals) # already_classified, industry_code


#saveRDS(entities, "entities.rds")                          
#jsonlite::write_json(entities, "entities.json", auto_unbox = TRUE, na = "null", pretty = TRUE)

```


# TAGGED CONTRIBUTORS FILE

```{r}
row_to_entity_all <- bind_rows(
  orgs_resolved        %>% select(row_id = cont_row_id, entity_id),
  individuals_resolved %>% select(row_id = original_row_id, entity_id)
) #%>% distinct()

contributions_tagged <- contributions %>%
  left_join(row_to_entity_all, by = "row_id") %>% # , relationship = "many-to-one"
  left_join(entity_type_lookup, by = "entity_id")

#write_csv(contributions_tagged, "contributions_tagged.csv")

```


# CLASSIFICATION INPUT

```{r}

auto_org_links <- all_org_pairs %>% filter(link_tier == "auto_link") %>% select(.x, .y)
if (nrow(auto_org_links) > 0) {
  auto_org_comp <- components(graph_from_data_frame(auto_org_links, directed = FALSE))$membership
  org_auto_unit_map <- tibble(row_id = as.integer(names(auto_org_comp)), auto_unit = auto_org_comp)
} else {
  org_auto_unit_map <- tibble(row_id = integer(), auto_unit = integer())
}

auto_indiv_links <- indiv_pairs %>% filter(link_tier == "auto_link") %>% select(.x, .y)
if (nrow(auto_indiv_links) > 0) {
  auto_indiv_comp <- components(graph_from_data_frame(auto_indiv_links, directed = FALSE))$membership
  indiv_auto_unit_map <- tibble(row_id = as.integer(names(auto_indiv_comp)), auto_unit = auto_indiv_comp)
} else {
  indiv_auto_unit_map <- tibble(row_id = integer(), auto_unit = integer())
}

# orgs
org_units <- orgs_resolved %>%
  left_join(org_auto_unit_map, by = "row_id") %>%
  mutate(unit_key = if_else(!is.na(auto_unit),
                             paste0(entity_id, "::", auto_unit),
                             paste0(entity_id, "::solo::", row_id))) %>%
  group_by(entity_id, unit_key) %>%
  summarise(
    entity_type    = "organization",
    needs_review   = any(needs_review),
    canonical_name = { n <- Contributor.Name; w <- n[has_contributor_id]
                       if (length(w) > 0) w[which.max(nchar(w))] else n[which.max(nchar(n))] },
    canonical_id   = { ids <- Contributor.ID[has_contributor_id & !is.na(Contributor.ID)]
                       if (length(ids) > 0) names(sort(table(ids), decreasing = TRUE))[1] else NA_character_ },
    employer       = NA_character_,
    occupation     = NA_character_,
    city           = names(sort(table(Contributor.City[!is.na(Contributor.City)]),  decreasing = TRUE))[1],
    state          = names(sort(table(Contributor.State[!is.na(Contributor.State)]), decreasing = TRUE))[1],
    zip          = names(sort(table(Contributor.Zip.Code[!is.na(Contributor.Zip.Code)]), decreasing = TRUE))[1],
    total_amount   = sum(Amount, na.rm = TRUE),
    .groups = "drop"
  )

# individual units: one row per {sub-cluster × employer × occupation}
# conflicting employer or occupation within a cluster leads to separate classification rows
indiv_units <- individuals_resolved %>%
  left_join(indiv_auto_unit_map, by = "row_id") %>%
  mutate(
    auto_cluster = if_else(!is.na(auto_unit),
                           paste0(entity_id, "::", auto_unit),
                           paste0(entity_id, "::solo::", original_row_id)),
    employer_key = coalesce(processed_employer_name, ""),
    occ_key      = coalesce(processed_occupation, ""),
    unit_key     = paste(auto_cluster, employer_key, occ_key, sep = "||")
  ) %>%
  group_by(entity_id, unit_key) %>%
  summarise(
    entity_type    = "individual",
     needs_review   = any(needs_review),
    canonical_name = { has_info <- !is.na(processed_employer_name) & processed_employer_name != "" &
                                   !is.na(standardized_city) & standardized_city != ""
                       cands <- standardized_name[has_info]
                       if (length(cands) > 0) cands[which.max(nchar(cands))]
                       else standardized_name[which.max(nchar(standardized_name))] },
    canonical_id   = { ids <- Contributor.ID[!is.na(Contributor.ID) & Contributor.ID != ""]
                       if (length(ids) > 0) names(sort(table(ids), decreasing = TRUE))[1] else NA_character_ },
    employer       = first(na.omit(processed_employer_name[processed_employer_name != ""])),
    occupation     = first(na.omit(processed_occupation[processed_occupation != ""])),
    city           = names(sort(table(standardized_city[!is.na(standardized_city)]),  decreasing = TRUE))[1],
    state          = names(sort(table(Contributor.State[!is.na(Contributor.State)]), decreasing = TRUE))[1],
    zip          = names(sort(table(Contributor.Zip.Code[!is.na(Contributor.Zip.Code)]), decreasing = TRUE))[1],
    total_amount   = sum(Amount, na.rm = TRUE),
    .groups = "drop"
  )

over_5k_entity_ids <- entity_max_amounts %>% filter(max_race_total >= 5000) %>% pull(entity_id)

classification_input <- bind_rows(org_units, indiv_units) %>%
  filter(entity_id %in% over_5k_entity_ids) %>%
  mutate(classification_unit_id = map_chr(unit_key, ~ hash_id(.x, "unit"))) %>%
  left_join(entity_max_amounts, by = "entity_id") %>%
  mutate(crosses_threshold = max_race_total >= 5000) %>%
  select(classification_unit_id, entity_id, entity_type, needs_review,
         canonical_id, canonical_name, employer, occupation, city, state, zip,
         total_amount, crosses_threshold)
```

```{r}
# One-time seed: build entity_classification_index.csv from already_classified_contributions.csv.
# After the index exists this chunk is a no-op.
if (!file.exists("08_inputs/entity_classification_index.csv")) {
  already_classified <- read_csv("08_inputs/already_classified_contributions.csv",
                                  col_types = cols(.default = "c")) %>%
    mutate(name_norm     = normalize_str(Contributor.Name),
           employer_norm = normalize_str(Contributor.Employer))

  employer_lookup <- bind_rows(
    already_classified %>%
      transmute(employer_norm = name_norm, code_final, code_final_description),
    already_classified %>%
      filter(employer_norm != "") %>%
      transmute(employer_norm, code_final, code_final_description)
  ) %>%
    filter(employer_norm != "") %>%
    distinct(employer_norm, .keep_all = TRUE)

  ci <- classification_input %>%
    mutate(name_norm     = normalize_str(canonical_name),
           employer_norm = normalize_str(coalesce(employer, "")))

  org_prior <- ci %>%
    filter(entity_type == "organization") %>%
    left_join(
      already_classified %>% distinct(name_norm, .keep_all = TRUE) %>%
        select(name_norm, code_final, code_final_description),
      by = "name_norm"
    ) %>%
    filter(!is.na(code_final))

  indiv <- ci %>% filter(entity_type == "individual")

  indiv_direct <- indiv %>%
    left_join(
      already_classified %>%
        distinct(name_norm, employer_norm, .keep_all = TRUE) %>%
        select(name_norm, employer_norm, code_final, code_final_description),
      by = c("name_norm", "employer_norm")
    ) %>%
    filter(!is.na(code_final))

  indiv_via_employer <- indiv %>%
    filter(!entity_id %in% indiv_direct$entity_id, employer_norm != "") %>%
    left_join(employer_lookup, by = "employer_norm") %>%
    filter(!is.na(code_final))

  prior_seed <- bind_rows(org_prior, indiv_direct, indiv_via_employer) %>%
    distinct(classification_unit_id, .keep_all = TRUE) %>%
    transmute(
      classification_unit_id, entity_id, entity_type, canonical_name,
      employer, occupation,
      naics_code             = code_final,
      naics_label            = code_final_description,
      level1_category        = NA_character_,
      level2_category        = NA_character_,
      classification_source  = "prior",
      classification_status  = "classified"
    )

  write_csv(prior_seed, "08_outputs/entity_classification_index.csv")
  message("Seeded entity_classification_index.csv with ", nrow(prior_seed), " prior classifications.")
}
```


```{r}
# load index and write only new entities to classification_input.csv
entity_index <- if (file.exists("08_outputs/entity_classification_index.csv")) {
  read_csv("08_outputs/entity_classification_index.csv", col_types = cols(.default = "c"))
} else {
  tibble(classification_unit_id = character())
}

new_for_classification <- classification_input %>%
  filter(!classification_unit_id %in% entity_index$classification_unit_id)

write_csv(new_for_classification, "08_outputs/classification_input.csv")
message(nrow(new_for_classification), " new entities written to classification_input.csv (",
        nrow(entity_index), " already in index).")
```

^ would want to arrange this to be at the begging so we're no re-computing entity groups each time. Just assign new contribs the existing entity id. 







  
  
  