---
title: "R Notebook"
output: html_notebook
---

### to dos
- fix contradictions between is_individual and the label in the info data
  - manually deal with edge cases
- should we include an indicator for a federal committee? 
- think about whether to do additional check with fuzzy matching / jaccard similarity for dups
- look for SEUI spelled out
- should local unit impact categorization?
- ask thomas about the open_secrets_contributor_id and whether that definitively links people
- add zip code processed to info_non_individuals

### load packages

```{r}
library(tidyr)
```


### load helper functions 

```{r}
source("standardization_helpers.R")
```


### open secrets data

```{r}
# open secrets data
categories <- read.csv("08_inputs/ContributionOpenSecretsCategories.csv") # doesn't have employer, occupation, etc.
info <- read.csv("08_inputs/categorization_additional_info_for_individuals.csv") # data with additional info on individuals including employer, occupation, etc.
```

### clean up data

```{r}
# rename columns to be consistens with 'categories' dataset
info <- info %>% rename(
  level1_category = level_1_catecory,
  level2_category = level_2_category,
  level3_category = level_3_category
)
```


```{r}
# take out unitemized contributions
info <- info[-which(info$open_secrets_contributor_name == "UNITEMIZED DONATIONS"),]
categories <- categories[-which(categories$name == "UNITEMIZED DONATIONS"),]
categories <- categories[-which(categories$name == "UNITEMIZED, UNITEMIZED"),]

# take out uncoded contributions
info <- info[-which(tolower(info$level1_category) == "uncoded"),]
categories <- categories[-which(tolower(categories$level1_category) == "uncoded"),]

# take out 'candidate contributions' and 'non-contributions'
categories <- categories[-which(categories$level1_category == "Candidate Contributions"),]
categories <- categories[-which(categories$level1_category == "Non-Contributions"),]

info <- info[-which(info$level1_category == "Candidate Contributions"),]

# filter so the additional info category only includes individuals
#info <- info[which(info$contributor_type == "Individual"),]
```


### standardize names

```{r}
# process open secrets data
categories <- categories %>%
  mutate(
    standardized_name = standardize_names(name),
      standardized_name = fix_typos(standardized_name),
      processed_name = remove_pac_info(standardized_name),
      processed_name = remove_unit_info(processed_name),
      #processed_name = replace_committee_text(processed_name),
      processed_name = replace_business_text(processed_name),
      processed_name = str_replace_all(processed_name, " CA ", " CALIFORNIA "),
      processed_name = str_remove_all(processed_name, "\\s*\\(.*?\\)"),
      processed_name = str_squish(processed_name)
  ) 

# data with additional info
info <- info %>%
  mutate(
    # contributor name
    standardized_name = standardize_names(open_secrets_contributor_name),
    standardized_name = fix_typos(standardized_name),
    processed_name = remove_pac_info(standardized_name),
    processed_name = remove_unit_info(processed_name),
    #processed_name = replace_committee_text(processed_name),
    processed_name = replace_business_text(processed_name),
    processed_name = str_replace_all(processed_name, " CA ", " CALIFORNIA "),
    processed_name = str_remove_all(processed_name, "\\s*\\(.*?\\)"), # removes everything in parentheses
    processed_name = str_squish(processed_name),
    
    # contributor employer
    standardized_employer_name = standardize_names(employer),
    standardized_employer_name = fix_typos(standardized_employer_name),
    processed_employer_name = remove_pac_info(standardized_employer_name),
    processed_employer_name = remove_unit_info(processed_employer_name),
    #processed_employer_name = replace_committee_text(processed_employer_name),
    processed_employer_name = replace_business_text(processed_employer_name),
    processed_employer_name = standardize_occupation_employer(processed_employer_name), # standardize how 'none', 'not employed', etc are marked
    processed_employer_name = str_replace_all(processed_employer_name, " CA ", " CALIFORNIA "),
    processed_employer_name = str_remove_all(processed_employer_name, "\\s*\\(.*?\\)"), # removes everything in parentheses
    processed_employer_name = str_squish(processed_employer_name),
    
    # contributor occupation
    standardized_occupation = standardize_names(occupation),
    processed_occupation = standardize_occupation_employer(standardized_occupation),
    
    # contributor city
    standardized_city = standardize_names(city),
    
  ) 
```

### standardize zip code

```{r}
info$zip_code_processed <- info$zip
info[which(!is.na(info$zip)),]$zip_code_processed <- sprintf("%05d", info[which(!is.na(info$zip)),]$zip)
```


### label individuals vs orgs

```{r}
categories$is_individual <- sapply(categories$standardized_name, is_individual)
info$is_individual <- sapply(info$standardized_name, is_individual)
```


```{r}
info[which(info$is_individual == FALSE & info$contributor_type == "Individual"),]
info[which(info$is_individual == TRUE & info$contributor_type != "Individual"),]
```
some individuals tagged as non-individuals, but just go with designation for now.

```{r}
info_individuals    <- info %>% filter(contributor_type != "Non-Individual")
info_non_individuals <- info %>% filter(contributor_type == "Non-Individual")
categories_orgs <- categories %>%
  filter(!sapply(standardized_name, is_individual))
```


### format employer

```{r}
# employer and occupation should be blank for non individuals
info_non_individuals$processed_employer <- ""
info_non_individuals$processed_occupation <- ""
```

- Some are not, can look at using this info in same way
  - PAC IDS
  - alternate names
  - can take some out manually


# Check for contradictions between the two datasets


```{r}
# generate name variants with different processing
make_name_variants <- function(df, col) {
  df %>%
    mutate(
      name_v1 = standardize_names(.data[[col]]) %>% fix_typos(),
      name_v2 = remove_pac_info(name_v1)          %>% str_squish(),
      name_v3 = remove_unit_info(name_v2)          %>% str_squish(),
      #name_v4 = replace_committee_text(name_v3)    %>% str_squish(),
      name_v5 = replace_business_text(name_v3)     %>% str_squish(),  # = processed_name
      name_v6 = str_remove_all(name_v5, "\\s*\\(.*?\\)") %>% str_squish()  # + remove parens
    )
}

# add name variants
categories_orgs  <- make_name_variants(categories_orgs,   "name") #%>% mutate(row_id = row_number())
info_non_individuals <- make_name_variants(info_non_individuals, "open_secrets_contributor_name")      #%>% mutate(ref_id = row_number())
```

```{r}
# set to "level1", "level2", or "level3" depending on the most granular level we are going to use
CONTRADICTION_LEVEL <- "level2"

find_and_remove_contradictions <- function(cats, info_nonind, level = "level2") {
  class_col   <- paste0(level, "_category")
  variant_cols <- c("standardized_name", "processed_name","name_v1","name_v2","name_v3","name_v5","name_v6")

  # one row per (entity, variant) for categories
  cats_long <- cats %>%
    select(orig_name = name, all_of(variant_cols), cat_class = !!sym(class_col)) %>%
    pivot_longer(cols = all_of(variant_cols), names_to = "variant_type", values_to = "variant") %>%
    filter(!is.na(variant) & variant != "") %>%
    distinct(orig_name, variant, cat_class)

  # one row per (entity, variant) for info non-individuals
  info_long <- info_nonind %>%
    select(orig_name = open_secrets_contributor_name, all_of(variant_cols), info_class = !!sym(class_col)) %>%
    pivot_longer(cols = all_of(variant_cols), names_to = "variant_type", values_to = "variant") %>%
    filter(!is.na(variant) & variant != "") %>%
    distinct(orig_name, variant, info_class)

  # flag a contradiction for any variant with different classification
  contradictions <- inner_join(cats_long, info_long, by = "variant", relationship = "many-to-many") %>%
    filter(!is.na(cat_class) & !is.na(info_class) & cat_class != info_class) %>%
    rename(cats_name = orig_name.x, info_name = orig_name.y, matched_variant = variant) %>%
    distinct(cats_name, info_name, cat_class, info_class, matched_variant)

  n_pairs <- nrow(contradictions %>% distinct(cats_name, info_name))
  cat(sprintf("Contradictions at %s: %d entity pairs removed from both sources\n", level, n_pairs))
  if (n_pairs > 0) print(contradictions %>% select(cats_name, info_name, cat_class, info_class, matched_variant))

  list(
    cats_clean     = cats       %>% filter(!name                        %in% unique(contradictions$cats_name)),
    info_clean     = info_nonind %>% filter(!open_secrets_contributor_name %in% unique(contradictions$info_name)),
    contradictions = contradictions
  )
}

result <- find_and_remove_contradictions(categories_orgs, info_non_individuals, level = CONTRADICTION_LEVEL)

categories_clean           <- result$cats_clean
info_non_individuals_clean <- result$info_clean
contradictions             <- result$contradictions

```


```{r, eval = FALSE}
# OLD VERSION

# non-individuals from info that might overlap with categories
contradictions <- info_non_individuals %>%
  select(processed_name, info_level1 = level1_category, info_level2 = level2_category, info_level3 = level3_category) %>%
  inner_join(
    categories_orgs %>% select(processed_name, cat_level1 = level1_category, cat_level2 = level2_category, cat_level3 = level3_category),
    by = "processed_name"
  ) %>%
  filter(info_level1 != cat_level1 | info_level2 != cat_level2 | info_level3 != cat_level3)

nrow(contradictions)
# inspect before combining
contradictions %>% select(processed_name, info_level1, cat_level1, info_level2, cat_level2, info_level3, cat_level3) %>% unique()

```


### combine datasets

```{r}
categories %>% filter(!(processed_name %in% info_non_individuals$processed_name))
```
```{r}
info_non_individuals[which(!(info_non_individuals$processed_name %in% categories_orgs$processed_name)),]
```

- these should be changed to individual
- could think about including city and state for employers just b/c we have it
- open_secrets_contributor_id

```{r}
info_non_individuals %>% group_by(open_secrets_contributor_id) %>% mutate(count = n_distinct(level3_category)) %>% filter(count > 1) %>% arrange(open_secrets_contributor_id)
info_individuals %>% group_by(open_secrets_contributor_id) %>% mutate(count = n_distinct(level3_category)) %>% filter(count > 1) %>% arrange(open_secrets_contributor_id)
```

Don't want to keep anything that has conflicting categories within a single contributor id -> don't know what to assign to it
- these don't look wrong though, may just need a new id... only a few cases, may as well get rid of them.. 
- don't want to totally rely on this, but not sure when it will come into play.... 


```{r}
info_individuals %>% group_by(open_secrets_contributor_id) %>% mutate(count = n()) %>% filter(count > 1) %>% arrange(open_secrets_contributor_id) %>% select(open_secrets_contributor_id,open_secrets_contributor_name,level3_category,employer,occupation,city)
```
- same individual id seems to have same name, but can have different employers
- which to include in masterfile?
- it seems to have already linked people that are the same but list different jobs and such
- could use as alias lookup -> would match various employers... 



```{r}
# categories not already covered by info 
categories_unique <- categories_clean %>%
  filter(!processed_name %in% info_non_individuals_clean$processed_name)

running_list <- bind_rows(

  # orgs/committees from info — highest priority
  info_non_individuals_clean %>%
    select(
      name = open_secrets_contributor_name,
      open_secrets_contributor_id,
      processed_name,
      city, state, zip = zip_code_processed,
      level1_category, level2_category, level3_category
    ) %>%
    mutate(source = "opensecrets_additional_info", entity_type = "organization"),

  # orgs from categories — only where info has no entry
  categories_unique %>%
    select(
      name,
      processed_name,
      level1_category, level2_category, level3_category
    ) %>%
    mutate(source = "opensecrets_categories", entity_type = "organization"),

  # individuals — only from info, with disambiguation columns
  
  info_individuals %>%
  select(
    name = open_secrets_contributor_name,
    open_secrets_contributor_id,  
    processed_name,
    processed_employer_name, processed_occupation,
    employer, occupation,
    city = standardized_city, zip = zip_code_processed, state,
    level1_category, level2_category, level3_category
  ) %>%
  mutate(source = "opensecrets_additional_info", entity_type = "individual")


) %>%
  distinct() %>%
  mutate(
    # universal join key 
    name_norm = normalize_name_simple(name),
    # lets a later classification step fall back to matching this person's employer against the organization rows 
    employer_norm = normalize_name_simple(employer)
  )


```


```{r}
# write to data file
write.csv(
  running_list,
  "../../data/03_input/masterfile/running_list_opensecrets_alt.csv",
  row.names = FALSE
)

```


  
  
  
  


