---
title: "R Notebook"
output: html_notebook
---

# TO DOS
- need to make sure we grab $5k donors for an pacs where a pac_id was added in the manual review
  - right now those aren't being fed back here
- update how we're dealing with major donors 
  - resolvable unless there pac language or IE language
  - not sure what to do when it matches to both / one is a recipient committee
  - don't remove parentheses from name when matching
- make sure fed pacs being labeled correctly
- deal with propagated ids
- update pac name pull
- add flag for same name, diff contributor ids in our contributors... 
- make sure temporary patches are resolved
- update contributor_type lookup
- add process to propagate contributor.ids based on an exact name match
  - see how this works with major donors that are individuals
- some PACs in the power search data may be missing a committee id; should fill in where we can
  - not for all PACs, but for the ones that show up in our data


# Read in packages and data

```{r}
library(dplyr)
library(readr)
library(stringr)
library(httr)
library(jsonlite)

source("standardization_helpers.R")
```

```{r}
contributions_full <- read.csv("08_inputs/power_search_contributions_normalized.csv")
nrow(contributions_full)
colnames(contributions_full)

contributions_full <- contributions_full %>% rename(
  Ballot.Measure = Ballot.Measure.s.
)

#contributions_full <- read.csv("08_inputs/contributors_data_full.csv")
#contributions_full$Amount <- as.numeric(contributions_full$Amount)
#colnames(contributions_full)[13] <- "Ballot.Measure"
```

```{r}
# data to map contributor ids
id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv", col_types = cols(.default = "c"))
id_map_full <- id_map_full %>% filter(!is.na(FILER_ID) & !is.na(XREF_FILER_ID))
#id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv")

# get unique xref id / filer ids combos
id_map <- unique(id_map_full[,c("XREF_FILER_ID","FILER_ID","FILER_TYPE")])

# direct pac contributors
x <- read.csv("08_outputs/pac_input.csv")
```

```{r}
x[which(x$effective_Contributor.ID == "1479934"),]$FILER_ID <- "1479934"
x[which(x$effective_Contributor.ID == "1049814"),]$FILER_ID <- "1049814"
```


```{r}
id_map[which(is.na(id_map$XREF_FILER_ID)),]
id_map[which(is.na(id_map$FILER_ID)),]
```
```{r}
id_map_full[which(is.na(id_map_full$FILER_ID)),]
```


filername_cd.tsv info: https://calaccess.californiacivicdata.org/documentation/raw-files/filername-cd/ 
- city, state, zip, phone, fax, email can differ across same filer id
- recipient committee id corresponds to FILER_ID, contributor.id corresponds to XREF_ID
- Map FILER_ID to x based on Contributor.ID matching xref_filer_id

```{r,eval = FALSE}
id_map_full %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(NAML, FILER_ID, XREF_FILER_ID) %>% filter(grepl("POLITICAL", NAML)) %>% distinct()
```

Those i manually searched come up as major donor under entity type in cal-access
- USAA GROUP POLITICAL ACTION COMMITTEE comes up as major donor in cal-accesss
- COALITION FOR NATURAL HEALTH POLITICAL ACTION COMMITTEE, INC., THE: also major donor in cal access, says terminated on FEC website
- COUNTRYWIDE HOME LOANS, INC.; COUNTRYWIDE POLITICAL ACTION COMMITTEE (FEDERAL PAC) -> can't find on FEC even though name says federal PAC
  

# aggregate contributors data to only include $5k+ contributions
- this is imperfect for the first pass (will want to pull from entity-resulution data)
- not doing name match here

```{r}
x[which(x$Contributor.ID == "" | is.na(x$Contributor.ID)),] 
x[which(x$effective_Contributor.ID == "" | is.na(x$effective_Contributor.ID)),] 
```

  
# Clean up data

```{r}
# convert all ids to character strings
id_map$FILER_ID         <- as.character(id_map$FILER_ID)   
id_map$XREF_FILER_ID <- as.character(id_map$XREF_FILER_ID)
x$Contributor.ID <- as.character(x$Contributor.ID)
x$effective_Contributor.ID <- as.character(x$effective_Contributor.ID)
contributions_full$Recipient.Committee.ID <- as.character(contributions_full$Recipient.Committee.ID)
x$FILER_ID <- as.character(x$FILER_ID)
contributions_full$Contributor.ID <- as.character(contributions_full$Contributor.ID)

x$Recipient.Committee.ID <- as.character(x$Recipient.Committee.ID)
```



```{r}
# map filer id to contributor id (different ids may be used internally vs externally)
#x <- left_join(x, id_map, by = c("effective_Contributor.ID" = "XREF_FILER_ID"))

# check that each has a corresponding filer id
nrow(x[which(is.na(x$FILER_ID)),])
nrow(x[which(x$FILER_ID == ""),])
```



```{r}
# manually patch, but should be fixed upstream now
#x[which(x$effective_Contributor.ID == "1049814"),]$FILER_ID <- "1049814"
#nrow(x[which(is.na(x$FILER_ID)),])
```



### contributors missing ids

Manually fix known cases of missing or incorrect ids. 
  
```{r}
contributions <- contributions_full

# manually change contributor ids when id has error or is missing
contributions[which(contributions$Contributor.Name == "California Independent Petroleum Association PAC" & contributions$Contributor.ID == "8222237"),]$Contributor.ID <- "822237"
contributions[which(contributions$Contributor.Name == "Drive Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "880969"
contributions[which(contributions$Contributor.Name == "The Hartford Insurance Group, Inc. Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "930174"
contributions[which(contributions$Contributor.Name == "The Hartford Insurance Group, Inc. PAC (AKA THE HARTFORD ADVOCATES FUND)" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "930174"
contributions[which(contributions$Contributor.Name == "Arrowhead Credit Union (Indep Expenditure Committee)" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1240437"
contributions[which(contributions$Contributor.Name == "Sammons Enterprises, Inc. Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1488229"
contributions[which(contributions$Contributor.Name == "Lockheed Martin Employees' Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1431455"
contributions[which(contributions$Contributor.Name == "Sentry Insurance Company Federal Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1384599"
contributions[which(contributions$Contributor.Name == "Sentry Insurance Company Federal PAC" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1384599"
contributions[which(contributions$Contributor.Name == "DentaQuest PAC TN-C" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1408990"
contributions[which(contributions$Contributor.Name == "DentaQuest PAC-TN-C" & contributions$Contributor.ID == "1408900"),]$Contributor.ID <- "1408990"
contributions[which(contributions$Contributor.Name == "Abbott Laboratories Employee Political Action Committee - C00040279" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1332968"
contributions[which(contributions$Contributor.Name == "Abbott Laboratories Employee Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1332968"
contributions[which(contributions$Contributor.Name == "General Motors Company Political Action Committee (GM PAC)" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "790461"

```

- was not able to find contributor ids for all the PACs missing ids -> flag those for review


Don't think we need this (not currently taking city/state into account during entity resolution)

```{r,eval=FALSE}
# manually fix weird states
contributions <- contributions %>% mutate(
  standardized_state = if_else(standardized_state == "93" & standardized_city == "BAKERSFIELD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "C" & standardized_city == "LOS ANGELES", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "MJ" & standardized_city == "PARSIPPANY", "NJ",standardized_state),
  standardized_state = if_else(standardized_state == "A" & standardized_city == "FOLSOM", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "AC" & standardized_city == "GLENDALE", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "US" & standardized_city == "HERMITAGE", "TN",standardized_state),
  
  standardized_state = if_else(standardized_state == "" & standardized_city == "LOS ANGELES", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "SACRAMENTO", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "INGLEWOOD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "HAYWARD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "PASADENA", "CA",standardized_state),
)
```

- many contributions missing a city and/or state
  - could fill in some cases using zip code, but not sure that's needed

  

# Aggregate contributions by contributor

CALIFORNIA REAL ESTATE POLITICAL ACTION COMMITTEE/FEDERAL - CALIFORNIA ASSOCIATION OF REALTORS (FEDERAL PAC ID #C00083279)
-> this should have been flagged as a federal pac, not sure why we grabbed its contributors...

```{r}
x$Contributor.ID <- as.character(x$Contributor.ID)
contributions$Contributor.ID <- as.character(contributions$Contributor.ID)

# join filer id to this data 
contributions <- contributions %>% 
  left_join(
    id_map %>% select(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>% 
      rename(std_contributor_id = FILER_ID), 
    by = c("Contributor.ID" = "XREF_FILER_ID")
  )
```



```{r}
library(igraph)

# 1. flag and map contributor ids
prepped <- contributions %>%
  mutate(
    has_id       = !is.na(Contributor.ID) & Contributor.ID != "",
    is_fed_pac   = grepl("fed pac|federal pac|#c|id#|fec id|fec report|fed form|committee/federal", tolower(Contributor.Name)),
    is_pac_no_id = !has_id & !is_fed_pac &
                   grepl("\\bPAC\\b|COMMITTEE|POLITICAL ACTION|POLITICAL FUND|ACTION FUND", toupper(Contributor.Name))
  ) %>%
  mutate(
    is_self_contribution = !is.na(std_contributor_id) & std_contributor_id == Recipient.Committee.ID,
    unmatched_id        = has_id & is.na(std_contributor_id),
    std_contributor_id   = if_else(is_self_contribution, NA_character_, std_contributor_id),
    is_pac               = !is_self_contribution & has_id & !is_fed_pac
  )

# build per-PAC entity graph
name_nodes <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "") %>%
  distinct(Recipient.Committee.ID, standardized_name) %>% # distinct standardized_names within a recipient committee
  mutate(node = paste0("n_", row_number()))

id_nodes <- prepped %>%
  filter(!is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, std_contributor_id) %>%
  mutate(node = paste0("i_", row_number()))

# names that appear with >1 distinct (non-null) id within the same PAC are ambiguous
ambiguous_within_pac <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "",
         !is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, standardized_name, std_contributor_id) %>%
  group_by(Recipient.Committee.ID, standardized_name) %>% 
  filter(n() > 1) %>% # more than one contributor with this name, could have diff contributor ids
  ungroup() %>%
  distinct(Recipient.Committee.ID, standardized_name)

# edges only for unambiguous 
edges <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "",
         !is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, standardized_name, std_contributor_id) %>%
  anti_join(ambiguous_within_pac, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(name_nodes, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(id_nodes,   by = c("Recipient.Committee.ID", "std_contributor_id")) %>%
  select(from = node.x, to = node.y)

all_nodes <- c(name_nodes$node, id_nodes$node)
g         <- graph_from_data_frame(edges, directed = FALSE, vertices = all_nodes)
comps     <- components(g)

component_df <- tibble(node = names(comps$membership), entity_group = comps$membership)

name_to_group <- name_nodes %>%
  left_join(component_df, by = "node") %>%
  select(Recipient.Committee.ID, standardized_name, name_entity_group = entity_group)

id_to_group <- id_nodes %>%
  left_join(component_df, by = "node") %>%
  select(Recipient.Committee.ID, std_contributor_id, id_entity_group = entity_group)

# assign entity group 
prepped <- prepped %>%
  left_join(name_to_group, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(id_to_group,   by = c("Recipient.Committee.ID", "std_contributor_id")) %>%
  mutate(
    entity_group = case_when(
      !is.na(std_contributor_id) & std_contributor_id != "" ~ id_entity_group,
      !is.na(name_entity_group)                             ~ name_entity_group,
      TRUE ~ as.numeric(max(comps$membership) + row_number()),
    ),
    name_upper     = toupper(standardized_name),
    employer_upper = toupper(coalesce(standardized_employer_name, "")),
    is_dues        = grepl("UNION MEMBER|MEMBER CONTRIBUTIONS|MEMBER DUES|LESS THAN \\$?100|UNDER \\$?100|CONTRIBUTIONS UNDER|UNITEMIZED DUES|LABORERS DUES|LABORER DUES|MEMBERS DUES|INTERMEDIARY FOR MEMBERS|LABORERS TRUST|UNION NO 78\\s+\\d+", name_upper) |
                     grepl("UNION MEMBER|LABORERS|DISTRICT COUNCIL|EMPLOYEES|LOCAL\\s+\\d+|UNION\\s+\\d+|SMW\\s+\\d+|IBEW\\s+\\d+|SEIU\\s+\\d+|UFCW\\s+\\d+|OLCU\\s+\\d+|UNITE HERE\\s+\\d+|AFSCME\\s+\\d+|IRONWORKER\\s+\\d+|UPTE-CWA\\s+\\d+|UNION NO 78\\s+\\d+", employer_upper)
    
    
  ) %>%
  select(-name_entity_group, -id_entity_group, -name_upper, -employer_upper)

#  aggregate 
contributions_agg <- prepped %>%
  group_by(Recipient.Committee.ID, entity_group) %>%
  summarise(
    Amount             = sum(Amount, na.rm = TRUE),
    standardized_name  = {
      ns <- standardized_name[!is.na(standardized_name) & standardized_name != ""]
      if (length(ns) > 0) names(sort(table(ns), decreasing = TRUE))[1] else NA_character_
    },
    std_contributor_id = {
      ids <- std_contributor_id[!is.na(std_contributor_id) & std_contributor_id != ""]
      if (length(ids) > 0) ids[1] else NA_character_
    },
    FILER_TYPE            = {
      types <- FILER_TYPE[!is.na(FILER_TYPE) & FILER_TYPE != ""]
      if (length(types) > 0) names(sort(table(types), decreasing = TRUE))[1] else NA_character_
    },
    is_fed_pac            = any(is_fed_pac,            na.rm = TRUE),
    is_pac_no_id          = any(is_pac_no_id,          na.rm = TRUE),
    is_pac                = any(is_pac,                na.rm = TRUE),
    unmatched_contributor_id         = any(unmatched_id,         na.rm = TRUE), # should change to them all missing this id? could propagate in some cases..
    has_self_contribution = any(is_self_contribution,  na.rm = TRUE),
    is_dues  = any(is_dues, na.rm = TRUE),
    .groups               = "drop"
  ) %>%
  filter(Amount >= 0) %>%
  mutate(is_pac_no_id = is_pac_no_id & is.na(std_contributor_id)) # address cases where some contribs in entity group do have id, even if other don't



```

```{r}
prepped %>% filter(is_self_contribution == TRUE) %>% select(Recipient.Committee,Contributor.Name,  Recipient.Committee.ID,  Contributor.ID, Contributor.Employer)
```

1037597: typos are causing names not to aggregate: FOOTHILL DE ANZA COMMINITY COLLEGE FACULTY ASSOCIATION; FOOTHILL DE ANZA COMMUNITTY COLLEGE FACULTY ASSOCIATION
1486858: same name (missing, didn't aggregate)


# CREATE PAC GRAPH

### helper functions

```{r}
get_pac_contributors <- function(pac_id, contributions_agg) {
  contributions_agg %>%
    filter(Recipient.Committee.ID == pac_id)
}

categorize_contributor <- function(name,amount) {
  name_upper <- toupper(name)
  case_when(
    name_upper == "UNITEMIZED CONTRIBUTIONS" ~ "unitemized",
    amount <= 100 ~ "small_dollar",
    amount < 5000   ~ "under_5k",
    TRUE  ~ "over_5k" 
  )
}
```

```{r}
# find how much each PAC donated (to compare to how much it received)
pac_total_donated <- contributions_agg %>%
  group_by(std_contributor_id,unmatched_contributor_id) %>% #Contributor.ID, unmatched_contributor_id
  summarise(total_donated_all = sum(Amount, na.rm = TRUE), .groups = "drop")

```


### make graph

```{r}
DATA_SUFFICIENCY_THRESHOLD <- 0.75

traverse_pac_graph <- function(seed_ids, ref, id_map, max_depth = 5) {
  queue         <- lapply(seed_ids, function(id) list(id = id, depth = 0, path = id))
  visited       <- character(0)
  pac_summaries <- list()
  over_5k_contribs <- list()

  while (length(queue) > 0) {
    node   <- queue[[1]]; queue <- queue[-1]
    pac_id <- node$id; depth <- node$depth; path <- node$path

    if (pac_id %in% visited || depth >= max_depth) {
      if (depth >= max_depth && !(pac_id %in% visited)) {
        pac_summaries <- append(pac_summaries, list(tibble(
          pac_id = pac_id, depth = depth, path = path, status = "max_depth_reached",
          total_received = NA_real_, unitemized_amount = NA_real_,
        )))
      }
      next
    }
    visited <- c(visited, pac_id)
    
    contribs <- get_pac_contributors(pac_id, ref)

    total_donated_all <- pac_total_donated %>%
      filter(!unmatched_contributor_id, std_contributor_id == pac_id) %>%
      pull(total_donated_all)
    total_donated_all <- if (length(total_donated_all) == 0) NA_real_ else total_donated_all
    
    if (is.null(contribs) || nrow(contribs) == 0) {
      pac_summaries <- append(pac_summaries, list(tibble(
        pac_id = pac_id, depth = depth, path = path, status = "unresolved",
        total_received = NA_real_, total_donated_all = total_donated_all,
        data_sufficiency = NA_real_,
        unitemized_amount = NA_real_, #member_dues_amount = NA_real_,
        under_5k_amount = NA_real_, over_5k_amount = NA_real_
      )))
      next
    }
    
    # check data sufficiency
    total_received    <- sum(contribs$Amount, na.rm = TRUE)
    data_sufficiency  <- if (!is.na(total_donated_all) && total_donated_all > 0)
                           total_received / total_donated_all
                         else NA_real_

    pac_status <- case_when(
      !is.na(data_sufficiency) & data_sufficiency < DATA_SUFFICIENCY_THRESHOLD ~ "low_data_sufficiency",
      is.na(data_sufficiency)                                                   ~ "unknown_data_sufficiency",
      TRUE                                                                      ~ "resolved" # change this from being resolved...
    )
    
    contribs <- contribs %>%
      mutate(contributor_category = categorize_contributor(standardized_name, Amount))
    
    # summarize PAC based on what kind of donations it received
    pac_summaries <- append(pac_summaries, list(tibble(
      pac_id             = pac_id,
      depth              = depth,
      path               = path,
      status             = "resolved",
      total_received     = sum(contribs$Amount, na.rm = TRUE), 
      total_donated_all = total_donated_all,
      data_sufficiency  = total_received / total_donated_all,   # NA if total_donated_all == 0
      unitemized_amount  = sum(contribs$Amount[contribs$contributor_category == "unitemized"],  na.rm = TRUE),
      small_dollar_amount = sum(contribs$Amount[contribs$contributor_category == "small_dollar"],  na.rm = TRUE),
      under_5k_amount    = sum(contribs$Amount[contribs$contributor_category == "under_5k"],    na.rm = TRUE),
      over_5k_amount     = sum(contribs$Amount[contribs$contributor_category == "over_5k"],     na.rm = TRUE)
    )))

    # only process $5k+ contributors further - still want to add member dues even if its over $5k...
    #for(i in seq_len(nrow(contribs %>% filter(contribs$Amount > 5000)))){ # all contributors over $5k, even if union dues...
    for (i in seq_len(nrow(contribs %>% filter(contributor_category == "over_5k")))) {
      row <- (contribs %>% filter(contributor_category == "over_5k"))[i, ]

      if (isTRUE(row$is_fed_pac)) {
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id      = pac_id,
          path               = path,
          depth              = depth + 1,
          entity_group       = row$entity_group,
          contributor_name   = row$standardized_name, # do i want original name here for matching purposes? Contributor.Name
          contributor_id     = row$std_contributor_id, #row$Contributor.ID,
          contributor_type   = "federal_pac",
          contributor_amount = row$Amount
        )))

      #} else if (isTRUE(row$FILER_TYPE == "RECIPIENT COMMITTEE")) { # do want to include independent expenditure committees
       } else if(isTRUE(!is.na(row$std_contributor_id))|| isTRUE(row$is_pac) || isTRUE(row$is_pac_no_id)) { # look at anything with a contributor id
        #new_id   <- if (!is.na(row$std_contributor_id)) row$std_contributor_id else row$Contributor.ID
        new_id <- row$std_contributor_id
        matched  <- !is.na(row$std_contributor_id)
        new_path <- paste0(path, " -> ", new_id)
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id      = pac_id,
          path               = new_path,
          depth              = depth + 1,
          entity_group       = row$entity_group,
          contributor_name   = row$standardized_name, # Contributor.Name?
          contributor_id     = new_id,
          contributor_type   = if (isTRUE(row$is_pac_no_id)) "pac_no_id"
                               else if (!matched) "pac_unmatched_id"
                               else if (isTRUE(row$FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE")) "major_donor_ie_committee"
                               else "pac",
          #filer_type = if(!is.na(row$FILER_TYPE)) row$FILER_TYPE else NA_character_, # see which may be major donors / IE
          contributor_amount = row$Amount
        )))
        if (!is.na(new_id)) {
          queue <- append(queue, list(list(id = new_id, depth = depth + 1, path = new_path)))
        }

      } else {
        # individual or org — record for industry classification
        # could also miss some major donors - need to propagate ids
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id          = pac_id,
          path                   = path,
          depth                  = depth + 1,
          entity_group           = row$entity_group,
          contributor_name       = row$standardized_name, #row$Contributor.Name,
          contributor_id         = NA_character_, # update this, they may have a major donor id
          contributor_type       = "individual_or_org",
          contributor_amount     = row$Amount,
          #contributor_employer   = row$processed_employer_name,
          #contributor_occupation = row$processed_occupation,
          #contributor_city       = row$standardized_city,
          #contributor_state      = row$standardized_state
        )))
      }
    }
  }

  list(
    pac_summaries    = bind_rows(pac_summaries),
    over_5k_contributors = bind_rows(over_5k_contribs)
  )
}

```

- What do do w/ client, slate mailer orgs, payment to influnce and NA?
- only need recipient committee? 


```{r}
# run PAC graph
seed_ids <- unique(x$FILER_ID)
seed_ids <- seed_ids[!is.na(seed_ids) & seed_ids != ""]

result           <- traverse_pac_graph(seed_ids, contributions_agg, id_map, max_depth = 5)
pac_summaries    <- result$pac_summaries
over_5k_contributors <- result$over_5k_contributors

fed_seed_ids <- x %>%
  filter(grepl("fed pac|federal pac|#c|id#|fec id|fec report|fed form", tolower(Contributor.Name))) %>%
  pull(FILER_ID) %>%
  unique()

# include any PAC encountered during traversal that was labeled federal_pac as a contributor
federal_pac_ids <- union(
  fed_seed_ids,
  over_5k_contributors %>%
    filter(contributor_type == "federal_pac") %>%
    pull(contributor_id) %>%
    na.omit() %>%
    unique()
)

major_donor_ie_ids <- union(
  x %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% pull(FILER_ID) %>% unique(),
  over_5k_contributors %>% filter(contributor_type == "major_donor_ie_committee") %>% pull(contributor_id) %>% na.omit() %>% unique()
)

pac_summaries <- pac_summaries %>%
  mutate(
    is_federal_pac    = pac_id %in% federal_pac_ids,
    is_major_donor_ie = pac_id %in% major_donor_ie_ids
  )

```

```{r}
pac_summaries %>% group_by(pac_id) %>% filter(n() > 1) %>% arrange(pac_id)
```


```{r}
#write.csv(pac_summaries, "08_outputs/pac_summaries.csv",row.names = FALSE)
#write.csv(over_5k_contributors, "08_outputs/over_5k_contribs.csv",row.names = FALSE)
```


```{r}
type_priority <- c(federal_pac = 1, major_donor_ie_committee = 2, pac_unmatched_id = 3, pac_no_id = 4)

contributor_type_lookup <- over_5k_contributors %>%
  filter(!is.na(contributor_id), contributor_type %in% names(type_priority)) %>%
  distinct(contributor_id, contributor_type) %>%
  mutate(priority = type_priority[contributor_type]) %>%
  group_by(contributor_id) %>%
  slice_min(priority, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(contributor_id, contributor_type)

```

- Resolved just means the PAC was found (should change that language)


### check major donors/independent expenditure committees

```{r}
over_5k_contributors %>% filter(contributor_type == "major_donor_ie_committee")
```

- none of the major donors we're seeing have an contributors..

  
### check unmatched ids

```{r}
over_5k_contributors %>% filter(contributor_type == "major_donor_ie_committee")
```

```{r}
id_map_full %>% filter(grepl("arrowhead credit union",tolower(NAML)))
```


### look for pacs that are missing contributor.ids

```{r}
# Build name/ID lookup from full dataset
full_id_lookup <- id_map_full %>%
  filter(!is.na(FILER_ID), FILER_ID != "", (FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" | FILER_TYPE == "RECIPIENT COMMITTEE"), STATUS == "ACTIVE") %>%
  mutate(name_v5 = standardize_names(NAML) %>%
           str_remove_all("[[:punct:]]") %>%
          str_squish()) %>%
  distinct(name_v5, FILER_ID,FILER_TYPE, STATUS) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(FILER_ID)) %>%
  ungroup()


```

- check for more than one filer type as well (although that should correspond to more than one id)

```{r}
over_5k_contributors <- over_5k_contributors %>% mutate(
  name_v5 = standardize_names(contributor_name) %>%
           str_remove_all("[[:punct:]]") %>%
          str_squish()
  )
```


```{r}
# Propagate IDs to orgs that are missing them
full_id_propagation <- over_5k_contributors %>%
  filter(is.na(contributor_id)) %>%
  select(contributor_name, name_v5) %>%
  inner_join(full_id_lookup, by = "name_v5", relationship = "many-to-many") %>%
  mutate(needs_review = n_ids > 1) %>%
  select(contributor_name, name_v5, propagated_id = FILER_ID, n_ids, FILER_TYPE, STATUS, needs_review)
```

```{r}
full_id_propagation %>% filter(FILER_TYPE == "RECIPIENT COMMITTEE")
```


```{r}
full_id_propagation %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE")
```
- best way to incorporate these pac into the PAC graph? or just send them to manual review? 
- Wine Institute -> not a strong indicator its a pac -> is terminated... 

```{r}
full_id_propagation %>% filter(n_ids > 1)
```

- not sure what to do with the INDEP EXPENDITURE ACCOUNTS... 
- some match to more than one major donor id - think that's okay because we won't be using it
- do we need to flag cases where major donor has pac / IE language? 
- 

```{r}
id_map_full %>% filter(FILER_ID == "1018798")
```


```{r}
full_id_propagation %>% filter(n_ids > 1) #filter(FILER_TYPE == "RECIPIENT COMMITTEE")
```

Matched to more than one potential id: 
- CALIFORNIA-NEVADA CONFERENCE OF OPERATING ENGINEERS
- CALIFORNIA CREDIT UNION LEAGUE (CCUL ADMIN)
- CALIFORNIA CREDIT UNION LEAGUE (CCUL INDEPENDENT EXPENDITURE ACCOUNT)
  - can assume this one IE is correct..
  
```{r}
#ie_ids <- full_id_propagation %>% filter(FILER_TYPE == "RECIPIENT COMMITTEE")
#ie_ids <- full_id_propagation %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(propagated_id)
#contributions %>% filter(Recipient.Committee.ID %in% ie_ids$propagated_id)
```


- some of these recipient committees don't show any recieved contributions.
- most are major donor/indepedent expenditure committee as expected
  - look to see if any of these ids have any received contributions
- 8 Recipient committees that didn't have an id
- WINE INSTITUTE, THE (major donor), WINE INSTITUTE (recipient committee) -> can't tell the different in name matching, it matches to recipient committee, but hard to say which is which..

Not sure these all make sense: 
- HOD CARRIERS AND LABORERS LOCAL 783 (none have ids) contributed to 	LABORERS LOCAL UNION NO. 783-PAC and these are that PACs only contributors
  - but recipient committee id is different from id associated with HOD CARRIERS AND LABORERS LOCAL 783 so maybe they're not the same? 
- CALIFORNIA CREDIT UNION LEAGUE (CCUL INDEPENDENT EXPENDITURE ACCOUNT)
  - treat as an indepedent expenditure? 


### Apply propagated IDs

```{r}
# RECIPIENT COMMITTEE takes priority over MAJOR DONOR/IE when a name matches both
propagated_effective <- full_id_propagation %>%
  group_by(contributor_name) %>%
  mutate(has_rc = any(FILER_TYPE == "RECIPIENT COMMITTEE")) %>%
  ungroup() %>%
  filter(
    ( has_rc & FILER_TYPE == "RECIPIENT COMMITTEE") |
    (!has_rc & FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE")
  ) %>%
  group_by(contributor_name) %>%
  mutate(n_same_type = n_distinct(propagated_id)) %>%
  ungroup() %>%
  mutate(
    propagation_type = case_when(
      FILER_TYPE == "RECIPIENT COMMITTEE"                           & n_same_type == 1 ~ "recipient_committee",
      FILER_TYPE == "RECIPIENT COMMITTEE"                           & n_same_type > 1  ~ "rc_needs_review",
      FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" & n_same_type == 1 ~ "major_donor",
      FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" & n_same_type > 1  ~ "md_needs_review"
    )
  )

# only auto-apply unambiguous single-ID matches
propagated_resolved <- propagated_effective %>%
  filter(n_same_type == 1) %>%
  mutate(propagated_id = as.character(propagated_id)) %>%
  distinct(contributor_name, propagated_id, FILER_TYPE, propagation_type)
```


- does this change aggregation at all? 

```{r}
# Run traversal for newly identified recipient committee PACs
new_rc_ids <- propagated_resolved %>%
  filter(propagation_type == "recipient_committee") %>%
  pull(propagated_id) %>%
  unique() %>%
  setdiff(pac_summaries$pac_id) # some may already be in the list

if (length(new_rc_ids) > 0) {
  new_result <- traverse_pac_graph(new_rc_ids, contributions_agg, id_map, max_depth = 5)
  new_result$over_5k_contributors <- new_result$over_5k_contributors %>%
    mutate(name_v5 = standardize_names(contributor_name) %>%
             str_remove_all("[[:punct:]]") %>% str_squish())
  pac_summaries        <- bind_rows(pac_summaries, new_result$pac_summaries) %>%
    distinct(pac_id, .keep_all = TRUE)   # original traversal takes priority on duplicates
  over_5k_contributors <- bind_rows(over_5k_contributors, new_result$over_5k_contributors) %>%
    distinct(parent_pac_id, entity_group, contributor_type, .keep_all = TRUE)
}

```


```{r}
# Apply propagated IDs to rows that currently have no contributor_id
# recipient_committee → pac; major_donor → major_donor_ie_committee
over_5k_contributors <- over_5k_contributors %>%
  left_join(
    propagated_resolved %>% select(contributor_name, propagated_id, propagation_type),
    by = "contributor_name"
  ) %>%
  mutate(
    contributor_type = case_when(
      !is.na(contributor_id)                    ~ contributor_type,
      propagation_type == "recipient_committee"  ~ "pac",
      propagation_type == "major_donor"          ~ "major_donor_ie_committee",
      TRUE                                       ~ contributor_type
    ),
    contributor_id = coalesce(contributor_id, propagated_id)
  ) %>%
  select(-propagated_id, -propagation_type)
```


- need to re-run id propagation? 

```{r}
# Rebuild lookups that depend on over_5k_contributors now that types and IDs are updated
federal_pac_ids <- union(
  federal_pac_ids,
  over_5k_contributors %>% filter(contributor_type == "federal_pac") %>%
    pull(contributor_id) %>% na.omit() %>% unique()
)
major_donor_ie_ids <- union(
  major_donor_ie_ids,
  over_5k_contributors %>% filter(contributor_type == "major_donor_ie_committee") %>%
    pull(contributor_id) %>% na.omit() %>% unique()
)

pac_summaries <- pac_summaries %>%
  mutate(
    is_federal_pac    = pac_id %in% federal_pac_ids,
    is_major_donor_ie = pac_id %in% major_donor_ie_ids
  )

type_priority <- c(federal_pac = 1, major_donor_ie_committee = 2, pac_unmatched_id = 3, pac_no_id = 4)
contributor_type_lookup <- over_5k_contributors %>%
  filter(!is.na(contributor_id), contributor_type %in% names(type_priority)) %>%
  distinct(contributor_id, contributor_type) %>%
  mutate(priority = type_priority[contributor_type]) %>%
  group_by(contributor_id) %>%
  slice_min(priority, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(contributor_id, contributor_type)
```

```{r}
pac_summaries %>% group_by(pac_id) %>% filter(n() >1)
```


### Add indicators for whether we are able to classify each PAC in theory


```{r}
UNITEMIZED_DUES_THRESHOLD <- 0.5
OVER_5K_COVERAGE_THRESHOLD <- 0.5

pac_classifiability <- pac_summaries %>%
  mutate(
    low_data_sufficiency     = !is.na(data_sufficiency) & data_sufficiency < DATA_SUFFICIENCY_THRESHOLD,
    unknown_data_sufficiency = is.na(data_sufficiency),
    data_ok                  = status == "resolved" & !low_data_sufficiency & !unknown_data_sufficiency,
    #pct_unitemized_dues      = (unitemized_amount + member_dues_amount) / total_received,
    pct_unitemized = unitemized_amount / total_received,
    pct_small_dollar = small_dollar_amount / total_received,
    pct_over_5k              = over_5k_amount / total_received,
    #classify_by_name         = !is.na(pct_unitemized_dues) & pct_unitemized_dues >= UNITEMIZED_DUES_THRESHOLD & data_ok, # getting rid of this
    data_ok                  = status == "resolved" & !is_federal_pac & !low_data_sufficiency & !unknown_data_sufficiency
  )

```

- This is only a preliminary indicator for whether we can classify that PAC, we need to see whether or not we can actually classify its contributors.

```{r}
min(pac_classifiability$pct_over_5k,na.rm = TRUE)
max(pac_classifiability$pct_over_5k,na.rm = TRUE)
```


### flag each $5k contributor as resolvable or not

```{r}
# look at the status of each contributor.
over_5k_tagged <- over_5k_contributors %>%
  left_join(
    pac_classifiability %>% select(pac_id,data_ok, pct_over_5k, status,low_data_sufficiency), # classify_by_name
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    is_resolvable = case_when(
      contributor_type == "individual_or_org" ~ TRUE,
      contributor_type == "major_donor_ie_committee" & !grepl(
        "\\bPAC\\b|POLITICAL ACTION|INDEPENDENT EXPENDITURE|\\bCOMMITTEE\\b",
        toupper(contributor_name)
      ) ~ TRUE,
      contributor_type == "pac" & data_ok & pct_over_5k >= OVER_5K_COVERAGE_THRESHOLD ~ TRUE,
      TRUE ~ FALSE
    ),
    unresolvable_reason = case_when(
      !is_resolvable & contributor_type == "federal_pac"                             ~ "federal_pac",
      !is_resolvable & contributor_type %in% c("pac_no_id", "pac_unmatched_id")     ~ contributor_type,
      !is_resolvable & contributor_type == "major_donor_ie_committee"                 ~ "major_donor_ie_committee",
      #!is_resolvable & contributor_type == "pac" & !data_ok                         ~ paste0("pac_", status),
      !is_resolvable & low_data_sufficiency  ~ "low_data_sufficiency",
      !is_resolvable & contributor_type == "pac" & data_ok                          ~ "pac_low_5k_coverage",
      TRUE                                                                           ~ NA_character_
    )
  )

```


```{r}
# flag contributor names that appear with more than one distinct contributor ID
# — same name, different IDs means contributions weren't aggregated as one entity and need review
name_id_conflicts <- contributions %>%
  filter(Recipient.Committee.ID %in% pac_summaries$pac_id,
         !is.na(std_contributor_id) & std_contributor_id != "",
         !is.na(standardized_name) & standardized_name != "") %>%
  distinct(standardized_name, std_contributor_id) %>%
  group_by(standardized_name) %>%
  filter(n_distinct(std_contributor_id) > 1) %>%
  ungroup() %>%
  distinct(standardized_name)

over_5k_tagged <- over_5k_tagged %>%
  mutate(has_conflicting_id = contributor_name %in% name_id_conflicts$standardized_name)
```

```{r}
over_5k_tagged
```


```{r}
name_id_conflicts
```


```{r}
over_5k_tagged %>% filter(has_conflicting_id == TRUE) %>% arrange(contributor_name)
```
CALIFORNIA REAL ESTATE POLITICAL ACTION COMMITTEE/FEDERAL - CALIFORNIA ASSOCIATION OF REALTORS
  - one recipient committee was terminated mid-cycle (December 2025)

NORTHERN CALIFORNIA CARPENTERS REGIONAL COUNCIL POWER PAC
  - one is a SCC and one is not, don't think we want to group
  
PACE OF CALIFORNIA SCHOOL EMPLOYEES ASSOCIATION
  - only one contributor id ended up in the data (other didn't meet threshold? )
  
Some cases where the same name could be associated with different contributor ids
Will need to check the mapping for this...
PACE OF CALIFORNIA SCHOOL EMPLOYEES ASSOCIATION -> could be federal pac or not
Mostly an issue for MISSING ids. Should probably just stick with the id it has. 

```{r}
contributions_sub <- contributions %>%
  filter(Recipient.Committee.ID %in% pac_summaries$pac_id)
contributions_sub %>% filter(grepl("PACE OF CALIFORNIA",toupper(Contributor.Name)))
```


### compute % of each PAC contributors that can be classified

```{r}
# pre-compute PAC contributor dependencies once — avoids filtering over_5k_tagged in every loop iteration
pac_deps <- over_5k_tagged %>%
  filter(contributor_type %in% c("pac", "major_donor_ie_committee"), !is.na(contributor_id)) %>%
  distinct(parent_pac_id, contributor_id) %>%
  group_by(parent_pac_id) %>%
  summarise(deps = list(contributor_id), .groups = "drop") %>%
  { setNames(.$deps, .$parent_pac_id) }

# topological sort with break on no-progress (handles cycles / missing deps)
processed     <- character(0)
ordered_pacs  <- character(0)
pacs_to_order <- pac_summaries %>% arrange(desc(depth)) %>% pull(pac_id)

while (length(pacs_to_order) > 0) {
  n_before <- length(pacs_to_order)
  ready <- Filter(function(pid) {
    deps <- pac_deps[[pid]]
    is.null(deps) || all(deps %in% processed)
  }, pacs_to_order)
  ordered_pacs  <- c(ordered_pacs, ready)
  processed     <- c(processed, ready)
  pacs_to_order <- setdiff(pacs_to_order, ready)
  if (length(pacs_to_order) == n_before) break  # no progress: cycles or missing deps
}

stuck_pacs <- pacs_to_order

# pre-split by parent_pac_id so inner lookup is O(1) instead of filtering 1.8M rows
over_5k_by_pac    <- split(over_5k_tagged, over_5k_tagged$parent_pac_id)
classifiability_idx <- setNames(seq_len(nrow(pac_classifiability)), pac_classifiability$pac_id)
resolvable_vec      <- setNames(rep(NA_real_,  nrow(pac_classifiability)), pac_classifiability$pac_id)

# process orderable PACs
for (pid in ordered_pacs) {
  summary_row <- pac_classifiability[classifiability_idx[[pid]], ]
  if (!summary_row$data_ok) { resolvable_vec[[pid]] <- 0; next }

  direct_contribs <- over_5k_by_pac[[pid]]
  if (is.null(direct_contribs) || nrow(direct_contribs) == 0) { resolvable_vec[[pid]] <- 0; next }

  direct_contribs <- direct_contribs %>%
    mutate(
      child_resolvable_fraction = resolvable_vec[contributor_id],
      resolvable_amount = case_when(
        contributor_type == "individual_or_org"                              ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" & is_resolvable       ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" & !is_resolvable      ~ 0,
        contributor_type == "pac"                                             ~ contributor_amount * coalesce(child_resolvable_fraction, 0),
        TRUE                                                                  ~ 0
      )
    )

  resolvable_vec[[pid]] <- sum(direct_contribs$resolvable_amount, na.rm = TRUE) / summary_row$total_received
}

# process stuck pacs
remaining <- stuck_pacs

compute_fraction <- function(pid) {
  summary_row <- pac_classifiability[classifiability_idx[[pid]], ]
  if (!summary_row$data_ok) { resolvable_vec[[pid]] <<- 0; return() }

  direct_contribs <- over_5k_by_pac[[pid]]
  if (is.null(direct_contribs) || nrow(direct_contribs) == 0) { resolvable_vec[[pid]] <<- 0; return() }

  direct_contribs <- direct_contribs %>%
    mutate(
      child_resolvable_fraction = coalesce(resolvable_vec[contributor_id], 0),
      resolvable_amount = case_when(
        contributor_type == "individual_or_org"                              ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" & is_resolvable       ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" & !is_resolvable      ~ 0,
        contributor_type == "pac"                                             ~ contributor_amount * child_resolvable_fraction,
        TRUE                                                                  ~ 0
      )
    )

  resolvable_vec[[pid]] <<- sum(direct_contribs$resolvable_amount, na.rm = TRUE) / summary_row$total_received
}

while (length(remaining) > 0) {
  # PACs whose stuck deps have all been resolved can now be processed normally
  ready <- Filter(function(pid) {
    deps <- pac_deps[[pid]]
    is.null(deps) || !any(deps %in% remaining)
  }, remaining)

  if (length(ready) == 0) {
    # nothing can be unblocked — true cycle remains; force all with cyclic deps = 0
    ready <- remaining
  }

  for (pid in ready) compute_fraction(pid)
  remaining <- setdiff(remaining, ready)
}

pac_resolvable_fractions <- pac_classifiability %>%
  mutate(resolvable_fraction = resolvable_vec[pac_id])

```


**look at order of PACs associated with multiple depths to make sure its working**


### coverage breakdown for each seed PAC

```{r}
# primary: most frequent contributor name from actual transactions, keyed by contributor ID
pac_names_from_contributions <- contributions_agg %>%
  filter(!is.na(std_contributor_id) & std_contributor_id != "") %>%
  group_by(std_contributor_id) %>%
  summarise(
    Contributor.Name = {
      ns <- standardized_name[!is.na(standardized_name) & standardized_name != ""]
      if (length(ns) > 0) names(sort(table(ns), decreasing = TRUE))[1] else NA_character_
    },
    .groups = "drop"
  ) %>%
  transmute(FILER_ID = std_contributor_id, Contributor.Name)

# fallback: registry name for PACs not seen as contributors
pac_names_registry <- id_map_full %>%
  group_by(FILER_ID) %>%
  slice(1) %>%
  ungroup() %>%
  transmute(FILER_ID = as.character(FILER_ID), Contributor.Name = NAML)

pac_names <- bind_rows(
  pac_names_from_contributions,
  pac_names_registry %>% filter(!FILER_ID %in% pac_names_from_contributions$FILER_ID)
)
```

```{r}
seed_coverage <- over_5k_tagged %>%
  filter(depth == 1, parent_pac_id %in% seed_ids) %>%
  left_join(
    pac_resolvable_fractions %>% select(pac_id, resolvable_fraction),
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    resolvable_amount = case_when(
      contributor_type == "individual_or_org"                             ~ contributor_amount,
      contributor_type == "major_donor_ie_committee" & is_resolvable      ~ contributor_amount,
      contributor_type == "major_donor_ie_committee" & !is_resolvable     ~ 0,
      contributor_type == "pac"                                            ~ contributor_amount * coalesce(resolvable_fraction, 0),
      TRUE                                                                 ~ 0
    )
  ) %>%
  group_by(parent_pac_id) %>%
  summarise(
    total_direct_over_5k  = sum(contributor_amount, na.rm = TRUE),
    total_resolvable      = sum(resolvable_amount, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    pac_classifiability %>% filter(pac_id %in% seed_ids) %>%
      select(pac_id, total_received, data_ok, status, data_sufficiency), # classify_by_name
    by = c("parent_pac_id" = "pac_id")
  ) %>%
  mutate(
    pct_resolvable      = total_resolvable / total_received,
    classification_status = case_when(
      #classify_by_name                              ~ "classify_by_name",
      data_ok & pct_resolvable >= OVER_5K_COVERAGE_THRESHOLD ~ "classify_by_contributors",
      TRUE                                          ~ "unresolvable"
    )
  ) 


seed_coverage <- seed_coverage %>% left_join(pac_names, by = c("parent_pac_id" = "FILER_ID"))

# get summary for all seed pacs
all_seed_status <- pac_classifiability %>%
  filter(pac_id %in% seed_ids) %>%
  left_join(
    seed_coverage %>%
      select(parent_pac_id, total_direct_over_5k, total_resolvable, pct_resolvable, classification_status),
    by = c("pac_id" = "parent_pac_id")
  ) %>%
  left_join(contributor_type_lookup, by = c("pac_id" = "contributor_id")) %>%
  left_join(pac_names, by = c("pac_id" = "FILER_ID")) %>%
  mutate(
    has_pac_language = is_major_donor_ie & grepl(
      "\\bPAC\\b|POLITICAL ACTION|INDEPENDENT EXPENDITURE|\\bCOMMITTEE\\b",
      toupper(coalesce(Contributor.Name, ""))
    ),
    classification_status = case_when(
      is_major_donor_ie & !has_pac_language                          ~ "classify_directly",
      !is.na(classification_status)                                   ~ classification_status,
      #classify_by_name                                                ~ "classify_by_name",
      status %in% c("unresolved", "low_data_sufficiency",
                    "unknown_data_sufficiency")                       ~ "unresolvable",
      TRUE                                                            ~ "unresolvable"
    ),
    pct_resolvable       = coalesce(pct_resolvable, 0),
    total_direct_over_5k = coalesce(total_direct_over_5k, 0),
    total_resolvable     = coalesce(total_resolvable, 0),
    unresolvable_reason  = case_when(
      classification_status != "unresolvable"  ~ NA_character_,
      is_federal_pac                           ~ "federal_pac",
      has_pac_language                         ~ "major_donor_ie_committee",
      !is.na(contributor_type)                 ~ contributor_type,
      status == "unresolved"                   ~ "no_contributor_data",
      low_data_sufficiency                     ~ "low_data_sufficieny",
      unknown_data_sufficiency                 ~ "unknown_data_sufficiency",
      total_direct_over_5k == 0               ~ "no_over_5k_contributors",
      pct_resolvable < OVER_5K_COVERAGE_THRESHOLD ~ "low_pct_resolvable",
      TRUE                                     ~ "other"
    )
  ) %>%
  select(-has_pac_language)
```



# Send PACs to review queue
  
```{r}
# flag member dues in over_5k_tagged (name- and employer-based)
over_5k_tagged <- over_5k_tagged %>%
  mutate(
    name_upper     = toupper(contributor_name),
    is_member_dues = grepl(
      "UNION MEMBER|MEMBER CONTRIBUTIONS|MEMBER DUES|LESS THAN \\$?100|UNDER \\$?100|CONTRIBUTIONS UNDER|UNITEMIZED DUES|LABORERS DUES|LABORER DUES|MEMBERS DUES|INTERMEDIARY FOR MEMBERS|LABORERS TRUST",
      name_upper
    ) 
  ) %>%
  select(-name_upper)


```


```{r}
over_5k_tagged %>%
  filter(contributor_type == "major_donor_ie_committee", !is_resolvable) %>%
  distinct(contributor_name, contributor_id, depth) %>%
  mutate(role = if_else(depth == 1, "direct", "pac_contributor")) %>%
  arrange(contributor_name)

```

```{r}
all_seed_status %>% filter(Contributor.Name == "YOUTH SAVE DEMOCRACY PAC")
```

```{r}
over_5k_contributors %>% filter(contributor_id == "1478478")
```


```{r}
contributor_review_queue <- over_5k_tagged %>%
  filter(
    (contributor_type == "major_donor_ie_committee" & !is_resolvable) |
    contributor_type %in% c("pac_no_id", "pac_unmatched_id", "federal_pac")
  ) %>%
  left_join(
    id_map %>% distinct(FILER_ID, FILER_TYPE),
    by = c("contributor_id" = "FILER_ID")
  ) %>%
  select(
    parent_pac_id,
    contributor_name,
    contributor_id,
    contributor_type,
    FILER_TYPE,
    contributor_amount,
    review_reason = unresolvable_reason
  )

```

```{r}
contributions_agg$Recipient.Committee.ID <- as.character(contributions_agg$Recipient.Committee.ID)
contributions$Recipient.Committee.ID <- as.character(contributions$Recipient.Committee.ID)
over_5k_name_pairs$Recipient.Committee.ID <- as.character(over_5k_name_pairs$Recipient.Committee.ID)
```


```{r}
# any pac that is not 100% resolvable by $5k+ contributors
review_pacs <- pac_resolvable_fractions %>%
  filter(is.na(resolvable_fraction) | resolvable_fraction < 1) %>%
  pull(pac_id)


# (recipient, name) pairs at $5k+ in aggregate — exclude from non-$5k list
over_5k_name_pairs <- contributions_agg %>%
  filter(Amount >= 5000) %>%
  select(Recipient.Committee.ID, standardized_name)

# sub-$5k contributor labels (org name or individual employer)
non_5k_labels <- contributions %>%
  filter(Recipient.Committee.ID %in% review_pacs, Amount > 0) %>%
  anti_join(over_5k_name_pairs, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  mutate(
    name_upper     = toupper(standardized_name),
    employer_upper = toupper(coalesce(standardized_employer_name, "")),
    entity_type    = if_else(sapply(standardized_name, is_individual), "individual", "organization"),
    is_dues        = 
      grepl("UNION MEMBER|MEMBER CONTRIBUTIONS|MEMBER DUES|LESS THAN \\$?100|UNDER \\$?100|CONTRIBUTIONS UNDER|UNITEMIZED DUES|LABORERS DUES|LABORER DUES|MEMBERS DUES|INTERMEDIARY FOR MEMBERS|LABORERS TRUST", name_upper) |
      grepl("UNION MEMBER|LABORERS|DISTRICT COUNCIL|EMPLOYEES|LOCAL\\s+\\d+|UNION\\s+\\d+|SMW\\s+\\d+|IBEW\\s+\\d+|SEIU\\s+\\d+|UFCW\\s+\\d+|OLCU\\s+\\d+|UNITE HERE\\s+\\d+|AFSCME\\s+\\d+|IRONWORKER\\s+\\d+|UPTE-CWA\\s+\\d+", employer_upper),
    label = case_when(
      is_dues ~ NA_character_,
      entity_type == "individual" &
        !is.na(standardized_employer_name) &
        !standardized_employer_name %in% c("", "NONE", "N/A", "NA", "RETIRED", "NOT EMPLOYED", "UNEMPLOYED", "SELF EMPLOYED") ~ str_squish(standardized_employer_name),
      TRUE ~ standardized_name
    )

  ) %>%
  filter(!is.na(label) & label != "") %>%
  group_by(Recipient.Committee.ID) %>%
  summarise(non_5k_contributors = paste(sort(unique(label)), collapse = " | "), .groups = "drop")

# unresolvable $5k+ contributors with reason (excluding member dues)
unresolvable_5k_labels <- over_5k_tagged %>%
  filter(parent_pac_id %in% review_pacs, !is_resolvable, !is_member_dues) %>%
  mutate(label = paste0(contributor_name, " (", coalesce(unresolvable_reason, "unknown"), ")")) %>%
  group_by(parent_pac_id) %>%
  summarise(unresolvable_5k_contributors = paste(label, collapse = " | "), .groups = "drop")

member_dues_amounts <- contributions_agg %>%
  filter(Amount > 0, is_dues) %>%
  group_by(Recipient.Committee.ID) %>%
  summarise(member_dues_amount = sum(Amount, na.rm = TRUE), .groups = "drop")

# generalize seed_coverage to ALL PACs in the graph
all_pac_coverage <- over_5k_tagged %>%
  left_join(
    pac_resolvable_fractions %>% select(pac_id, child_resolvable_fraction = resolvable_fraction),
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    resolvable_amount = case_when(
      contributor_type == "individual_or_org"                              ~ contributor_amount,
      contributor_type == "major_donor_ie_committee" & is_resolvable       ~ contributor_amount,
      contributor_type == "major_donor_ie_committee" & !is_resolvable      ~ 0,
      contributor_type == "pac"                                             ~ contributor_amount * coalesce(child_resolvable_fraction, 0),
      TRUE                                                                  ~ 0
    )
  ) %>%
  group_by(parent_pac_id) %>%
  summarise(
    total_direct_over_5k = sum(contributor_amount, na.rm = TRUE),
    total_resolvable     = sum(resolvable_amount,  na.rm = TRUE),
    .groups = "drop"
  )

```

```{r}
member_dues_amounts$Recipient.Committee.ID <- as.character(member_dues_amounts$Recipient.Committee.ID)
non_5k_labels$Recipient.Committee.ID <- as.character(non_5k_labels$Recipient.Committee.ID)
```




```{r}
pac_review_queue <- pac_resolvable_fractions %>%
  filter(is.na(resolvable_fraction) | resolvable_fraction < 1) %>%
  left_join(pac_names,        by = c("pac_id" = "FILER_ID")) %>%
  left_join(all_pac_coverage, by = c("pac_id" = "parent_pac_id")) %>%
  left_join(contributor_type_lookup, by = c("pac_id" = "contributor_id")) %>%
  mutate(
    is_seed              = pac_id %in% all_seed_status$pac_id,
    total_direct_over_5k = coalesce(total_direct_over_5k, 0),
    total_resolvable     = coalesce(total_resolvable, 0),
    pct_resolvable       = coalesce(resolvable_fraction, 0),
    classification_status = case_when(
      data_ok & pct_resolvable >= OVER_5K_COVERAGE_THRESHOLD ~ "classify_by_contributors",
      TRUE                                                    ~ "unresolvable"
    ),
    unresolvable_reason = case_when(
      classification_status != "unresolvable" ~ NA_character_,
      is_federal_pac                          ~ "federal_pac",
      is_major_donor_ie & grepl(
        "\\bPAC\\b|POLITICAL ACTION|INDEPENDENT EXPENDITURE|\\bCOMMITTEE\\b",
        toupper(coalesce(Contributor.Name, ""))
      )                                       ~ "major_donor_ie_committee",
      !is.na(contributor_type)                ~ contributor_type,
      status == "unresolved"                  ~ "no_contributor_data",
      low_data_sufficiency                    ~ "low_data_sufficieny",
      unknown_data_sufficiency                ~ "unknown_data_sufficiency",
      total_direct_over_5k == 0              ~ "no_over_5k_contributors",
      pct_resolvable < OVER_5K_COVERAGE_THRESHOLD ~ "low_pct_resolvable",
      TRUE                                    ~ "other"
    )
  ) %>%
  left_join(non_5k_labels,          by = c("pac_id" = "Recipient.Committee.ID")) %>%
  left_join(unresolvable_5k_labels, by = c("pac_id" = "parent_pac_id")) %>%
  left_join(member_dues_amounts,    by = c("pac_id" = "Recipient.Committee.ID")) %>%
  mutate(
    member_dues_amount = coalesce(member_dues_amount, 0),
    pct_member_dues    = member_dues_amount / total_received
  ) %>%
  #rename(Contributor.Name = name) %>%
  select(
    pac_id, Contributor.Name, is_seed,
    status, unresolvable_reason,
    is_federal_pac, is_major_donor_ie, low_data_sufficiency, unknown_data_sufficiency, data_ok,
    total_received, total_donated_all, data_sufficiency,
    unitemized_amount, small_dollar_amount, under_5k_amount, over_5k_amount,
    pct_unitemized, pct_small_dollar, pct_over_5k,
    total_direct_over_5k, total_resolvable, pct_resolvable,
    classification_status, unresolvable_reason,
    pct_member_dues, member_dues_amount,
    non_5k_contributors, unresolvable_5k_contributors
  )
nrow(pac_review_queue)
```




```{r}
# current
combined_review_queue <- bind_rows(
  pac_review_queue,
  contributor_review_queue %>%
    distinct(contributor_id, contributor_name, review_reason) %>%
    left_join(
      pac_resolvable_fractions %>% select(pac_id, status, total_donated_all),
      by = c("contributor_id" = "pac_id")
    ) %>%
    mutate(
      # NA (not FALSE) when contributor_id is NA -- unmatched, so genuinely unknown
      is_seed = if_else(is.na(contributor_id), NA, contributor_id %in% all_seed_status$pac_id)
    ) %>%
    transmute(
      pac_id              = contributor_id,
      Contributor.Name    = contributor_name,
      unresolvable_reason = review_reason,
      is_seed,
      status,
      total_donated_all
    )
)
combined_review_queue
```



```{r}
unique(combined_review_queue$unresolvable_reason)
combined_review_queue %>% filter(unresolvable_reason == "major_donor_ie_committee")
```


```{r}
#write.csv(combined_review_queue,"08_outputs/pac_review_queue_080326.csv")
```


# BUILD PAC CLASSIFICATION INPUT

```{r}
library(digest)
library(uuid)
```

```{r}
contributions_agg$Recipient.Committee.ID <- as.character(contributions_agg$Recipient.Committee.ID)
prepped$Recipient.Committee.ID <- as.character(prepped$Recipient.Committee.ID)
```



```{r}
hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16))
}

# get entity_group for each $5k+ individual/org contributor
# get entity_group for each $5k+ individual/org contributor
over_5k_indiv <- over_5k_tagged %>%
  filter(contributor_type == "individual_or_org" | (contributor_type == "major_donor_ie_committee" & is_resolvable == TRUE))

#over_5k_indiv <- over_5k_tagged %>%
#  filter(contributor_type == "individual_or_org") %>%
#  left_join(
#    contributions_agg %>%
#      select(Recipient.Committee.ID, standardized_name, entity_group, std_contributor_id),
#    by = c("parent_pac_id" = "Recipient.Committee.ID",
#           "contributor_name" = "standardized_name")
#  )

direct_for_classification <- over_5k_indiv %>%
  #select(-std_contributor_id) %>%   # prepped has its own std_contributor_id
  inner_join(
    prepped %>% filter(!is_pac, !is_fed_pac),
    by = c("parent_pac_id" = "Recipient.Committee.ID", "entity_group")
  ) %>%
  mutate(entity_type = if_else(
    sapply(Contributor.Name, is_individual), "individual", "organization"
  ))

# format to match 0801 classification_input schema
pac_classification_input <- direct_for_classification %>%
  mutate(
    entity_id = if_else(
      !is.na(std_contributor_id),
      paste0("id_", std_contributor_id),
      sapply(Contributor.Name, function(n) hash_id(n, "indiv"))
    ),
    unit_key = paste(entity_id,
                     coalesce(processed_employer_name, ""),
                     coalesce(processed_occupation, ""),
                     sep = "||"),
    classification_unit_id = sapply(unit_key, function(k) hash_id(k, "unit")),
    contribution_uuid      = replicate(n(), uuid::UUIDgenerate())
  ) %>%
  select(-unit_key)

```

```{r}
nrow(over_5k_indiv)
nrow(pac_classification_input)
```

```{r}
nrow(over_5k_indiv)
nrow(pac_classification_input)
```


```{r}
# does the raw-row total per entity match the aggregated amount it's supposed to represent?
check <- direct_for_classification %>%
  group_by(parent_pac_id, entity_group) %>%
  summarise(raw_sum = sum(Amount, na.rm = TRUE), n_raw = n(), .groups = "drop") %>%
  left_join(
    over_5k_indiv %>% select(parent_pac_id, entity_group, contributor_amount),
    by = c("parent_pac_id", "entity_group")
  ) %>%
  mutate(diff = raw_sum - contributor_amount)

# should be ~0 for every row (small rounding aside) — anything far off means duplication
check %>% filter(abs(diff) > 1) %>% arrange(desc(abs(diff)))

# see if the 32k is concentrated in a few high-frequency entities (expected) vs spread evenly (suspicious)
check %>% arrange(desc(n_raw)) %>% head(20)

```

```{r}
check %>% arrange(desc(n_raw)) %>% select(diff) %>% distinct()
```

```{r}
pac_classification_input %>% filter(parent_pac_id == "1468673" & Contributor.Name == "Aspen Skilled Healthcare Inc")
```


```{r}
prepped %>% filter(Recipient.Committee.ID == "1468673" & entity_group == 2112) #%>% select(Contributor.Name) %>% distinct()
```

```{r}
contributions_full %>% filter(Recipient.Committee.ID == "1468673" & Contributor.Name == "Aspen Skilled Healthcare Inc")
```


```{r}
nrow(over_5k_indiv)
nrow(direct_for_classification)
```


```{r}
unique(pac_classification_input$Contributor.ID)
pac_classification_input[which(!is.na(pac_classification_input$Contributor.ID)),]
```

ABBOTT LABORATORIES EMPLOYEE POLITICAL ACTION COMMITTEE - C00040279 got flagged as individual or org probs because pac label was taken away..

```{r}
pac_classification_input %>% select(Contributor.Name,Contributor.Employer) %>% distinct()
```



```{r}
contributions_full %>% filter(Contributor.Name == "Baray, Michael") %>% summarise(total = sum(Amount))
```



### run this 

```{r}
pac_classification_input$uuid <- pac_classification_input$contribution_uuid
```


```{r}
pac_classification_input$uuid
```


### combine with existing classification input - skip for now

```{r}
existing$rowid <- as.character(existing$rowid)
pac_classification_input$rowid <- as.character(pac_classification_input$rowid)
```

```{r}
unique(existing$Cycle)
existing$Cycle <- as.integer(existing$Cycle)
pac_classification_input$Cycle <- as.integer(pac_classification_input$Cycle)
```



```{r, eval = FALSE}
# combine with 0801's classification_input (run 0801 first so the file exists)
existing <- read_csv("08_outputs/classification_input.csv", col_types = cols(.default = "c"))
existing$Cycle <- as.integer(existing$Cycle)
existing$Amount <- as.numeric(existing$Amount)
existing$has_pac_language <- as.logical(existing$has_pac_language)
existing$rowid <- as.character(existing$rowid)
existing$row_id <- as.integer(existing$row_id)

combined_classification_input <- bind_rows(existing, pac_classification_input)

write_csv(combined_classification_input, "08_outputs/classification_input_combined.csv")
cat("Combined rows:", nrow(combined_classification_input),
    "  (PAC added:", nrow(pac_classification_input), ")\n")
```

### save files to use later

```{r}
write_csv(pac_classification_input,    "08_outputs/pac_classification_input.csv")
write_csv(over_5k_tagged,              "08_outputs/over_5k_tagged.csv")
write_csv(pac_classifiability,         "08_outputs/pac_classifiability.csv")

# save the traversal order — needed to replay topological sort for industry breakdown
write_csv(tibble(pac_id = ordered_pacs, order = seq_along(ordered_pacs)),
          "08_outputs/pac_traversal_order.csv")
write_csv(tibble(pac_id = stuck_pacs),  "08_outputs/pac_stuck_pacs.csv")

```



### generate individual review queue

```{r}
df <- read.csv("08_outputs/classified_contributors_combined.csv")
nrow(df)
```

missing classification unit_id

```{r}
classified <- read.csv("08_inputs/already_classified_contributions.csv")
```

```{r}
nrow(classified)
classified$
```



```{r}
new <- df %>% filter(naics_code == "100" & (!(Contributor.Name %in% classified$Contributor.Name))) #%>% select(data_source_1) #%>% distinct()
```


```{r}
distinct <- new %>%
  select(Contributor.Name, Contributor.Employer, Contributor.Occupation,
         Contributor.City, Contributor.State, Contributor.Zip.Code) %>%
  distinct(across(everything(), tolower), .keep_all = TRUE) %>%
  arrange(Contributor.Name)

#distinct <- new %>% select(Contributor.Name,Contributor.Employer,Contributor.Occupation,Contributor.City,Contributor.State,Contributor.Zip.Code) %>% distinct() %>% arrange(Contributor.Name)
```

```{r}
nrow(distinct)
```


```{r}
nrow(distinct)
```


```{r}
library(digest)
distinct <- distinct %>%
  mutate(name_id = sapply(Contributor.Name, function(x) substr(digest(x, algo = "xxhash32"), 1, 8)))

```


```{r}
distinct %>% group_by(Contributor.Name) %>% filter(n() > 1)
```

```{r}
write.csv(distinct,"08_outputs/individual_review_queue.csv",row.names = FALSE)
```


```{r}
unique(distinct$Contributor.Employer)
unique(distinct$Contributor.Occupation)
```


```{r}
which(new$Contributor.Name %in% classified$Contributor.Name)
```


332

