---
title: "R Notebook"
output: html_notebook
---

# TO DOS
- add process to propagate contributor.ids based on an exact name match
  - see how this works with major donors that are individuals
- determine what entity resolution to do on this data
  - entity resolution process will affect how contributors show up in the review queue data
- some PACs in the power search data may be missing a committee id; should fill in where we can
  - not for all PACs, but for the ones that show up in our data


# Read in packages and data

```{r}
library(dplyr)
library(readr)
library(stringr)
library(httr)
library(jsonlite)

source("standardization_helpers.R")
```


```{r}
#write.csv(contributions_full, "08_inputs/contributors_data_full.csv",row.names = FALSE)
```

```{r}
contributions_full <- read.csv("08_inputs/contributors_data_full.csv")
```

```{r}
# data to map contributor ids
id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv")
# get unique xref id / filer ids combos
id_map <- unique(id_map_full[,c("XREF_FILER_ID","FILER_ID","FILER_TYPE")])

# contributor data (will need to update with output from data processing file)
#x <- read.csv("08_inputs/power_search_governors_data_5k_071326.csv")
x <- read.csv("08_inputs/pac_input.csv")
```

filername_cd.tsv info: https://calaccess.californiacivicdata.org/documentation/raw-files/filername-cd/ 
- city, state, zip, phone, fax, email can differ across same filer id
- recipient committee id corresponds to FILER_ID, contributor.id corresponds to XREF_ID
- Map FILER_ID to x based on Contributor.ID matching xref_filer_id

```{r}
contributions_full$Amount <- as.numeric(contributions_full$Amount)
colnames(contributions_full)[13] <- "Ballot.Measure"
```


```{r,eval = FALSE}
id_map_full %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(NAML, FILER_ID, XREF_FILER_ID) %>% filter(grepl("POLITICAL", NAML)) %>% distinct()
```

Those i manually searched come up as major donor under entity type in cal-access
- USAA GROUP POLITICAL ACTION COMMITTEE comes up as major donor in cal-accesss
- COALITION FOR NATURAL HEALTH POLITICAL ACTION COMMITTEE, INC., THE: also major donor in cal access, says terminated on FEC website
- COUNTRYWIDE HOME LOANS, INC.; COUNTRYWIDE POLITICAL ACTION COMMITTEE (FEDERAL PAC) -> can't find on FEC even though name says federal PAC
  

# aggregate contributors data to only include $5k+ contributions
- this is imperfect for the first pass (will want to pull from entity-resulution data)
- not doing name match here

```{r}
x[which(x$Contributor.ID == "" | is.na(x$Contributor.ID)),] 
```

  
# Clean up data

```{r}
# convert all ids to character strings
id_map$FILER_ID         <- as.character(id_map$FILER_ID)   
id_map$XREF_FILER_ID <- as.character(id_map$XREF_FILER_ID)
x$Contributor.ID <- as.character(x$Contributor.ID)
contributions_full$Recipient.Committee.ID <- as.character(contributions_full$Recipient.Committee.ID)
contributions_full$Contributor.ID <- as.character(contributions_full$Contributor.ID)
```


```{r}
#x <- x[which(!is.na(x$Contributor.ID)),] # filter to only include committees

# map filer id to contributor id (different ids may be used internally vs externally)
x <- left_join(x, id_map, by = c("Contributor.ID" = "XREF_FILER_ID"))
#x %>% group_by(uuid) %>% summarise(count = n()) %>% filter(count > 1)

# check that each has a corresponding filer id
nrow(x[which(is.na(x$FILER_ID)),])
nrow(x[which(x$FILER_ID == ""),])
```


**PACs missing a committee id**
* Border Health Federal PAC and Laborer's Union Local No. 777 PAC don't have contributor ids and they should...
* International Brotherhood of Electrical Workers Political Action Committee also doesn't have a contributor id and well need to fill that in
* UA 290 Industry Advancement Fund
* there are probably other ones that we'll need to fill in as well
  * Border Health is from Texas
  
### contributors missing ids

Manually fix known cases of missing or incorrect ids. 
Contributor ids to fix: 
- 8222237 & California Independent Petroleum Association PAC -> 	822237
- Sammons Enterprises, Inc. Political Action Committee -> 1488229 (according to Cal Access) -> federal PAC according to cal-access
- 	Lockheed Martin Employees' Political Action Committee -> just missing in one case



- 	TOTAL CONTRIBUTIONS RECEIVED (DISCLOSED ON COMMITTEE'S FEC FORM 3X)
  -contributor to 1485832 / 	CALIFORNIA REAL ESTATE POLITICAL ACTION COMMITTEE/FEDERAL - CALIFORNIA ASSOCIATION OF REALTORS (FEDERAL PAC ID #C00083279)
  - check this is flagged as a federal PAC, but can remove this row... 
  - might not be b/c diff pattern (Federal Pac id)
  - need to make sure i'm grabbing anything w/ C#### as well
  - 1491433 another recipient that seems the same but i guess isn't? 
  - different ids b/c one was terminated at the end of 2025... 
- California Independent Petroleum Association PAC	8222237 -> 822237
- Drive committee flagged as unmatched_id because it has committee in the name i think? is it a pac?
  - this one is in washington, so guessing it should have the same ID? 
  - any instances where it wouldn't be the same? 
  - many have	880969 as Contributor.ID -> looks like a federal PAC
  - washington DC an indicator for a fedearl PAC? 
  - not seeing lower case "Drive" with this id?
- The Hartford Insurance Group, Inc. Political Action Committee (missing) -> 930174(XREF)/1058621
  - HARTFORD INSURANCE GROUP INC. PAC (AKA THE HARTFORD ADVOCATES FUND) (FED PAC ID# C00168864); THE	930174(XREF)/1058621
  - 	The Hartford Insurance Group, Inc. PAC	1303334
  - also associated with this id which looks like a lobbying client..(one rows, others match the other contrib id)
- THE HARTFORD INSURANCE GROUP, INC PAC (AKA THE HARTFORD ADVOCATES FUND) -> 930174 (also a federal pac)
- Sammons enterprises
  - SAMMONS ENTERPRISES INC. POLITICAL ACTION COMMITTEE (FED PAC ID #C00388777) -> 1488229  (only one in cal-access)
- LOCKHEED MARTIN EMPLOYEES POLITICAL ACTION COMMITTEE (many in the data that don't have ID) -> 	1431455
  - think it could be a major donor or a PAC... if just Lockheed Martin
  - Lockheed Martin Employees' Political Action Committee	1431455
  - cal-access doesn't call this a federal pac, but seems to be one? 
- ARROWHEAD CREDIT UNION (INDEP EXPENDITURE COMMITTEE) -> 1240437
  - do we also map plain "Arrowhead Credit Union" which doesn't note its an IE committee, but contributed to same recipient.
  - no received contributions under this id... 
- SENTRY INSURANCE COMPANY FEDERAL POLITICAL ACTION COMMITTEE -> 	1384599
  - also contributors that are just "Sentry Insurance Company" -> unclear if this is difference...
  - theres a record for Sentry Insurance Company in cal access as a lobbyist employer
- HDR Engineering, Inc. Employee Owners PAC
  - no other contributors to go off of
  - its a major donor, but just HDR Engineering, INc. 
  - can't find employee owners PAC
  - major donor id 498046
  **make sure this one is flagged in PAC review**
- DentaQuest PAC-TN-C,1408900 -> 1408990
- Oklahoma Victory Committee 
  - federal pac id is C00879072
  - can't find anything in cal-access
- Abbott Laboratories Employee Political Action Committee - C00040279 -> 1332968
  - several version of this missing an id
  - Abbott Laboratories Employee PAC
- General Motors Company Political Action Committee (GM PAC) -> 790461
  - also a fedearl PAC
  - assuming if one under this id is a fedearl PAC, they all are...
  - all from washington, DC, one is from arlington
  
  
```{r}
contributions <- contributions_full
```

```{r}
# manually change contributor ids when id has error or is missing

contributions[which(contributions$Contributor.Name == "California Independent Petroleum Association PAC" & contributions$Contributor.ID == "8222237"),]$Contributor.ID <- "822237"
contributions[which(contributions$Contributor.Name == "Drive Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "880969"
contributions[which(contributions$Contributor.Name == "The Hartford Insurance Group, Inc. Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "930174"
contributions[which(contributions$Contributor.Name == "The Hartford Insurance Group, Inc. PAC (AKA THE HARTFORD ADVOCATES FUND)" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "930174"
contributions[which(contributions$Contributor.Name == "Arrowhead Credit Union (Indep Expenditure Committee)" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1240437"
contributions[which(contributions$Contributor.Name == "Sammons Enterprises, Inc. Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1488229"
contributions[which(contributions$Contributor.Name == "Lockheed Martin Employees' Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1431455"
contributions[which(contributions$Contributor.Name == "Sentry Insurance Company Federal Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1384599"
contributions[which(contributions$Contributor.Name == "Sentry Insurance Company Federal PAC" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1384599"
contributions[which(contributions$Contributor.Name == "DentaQuest PAC TN-C" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1408990"
contributions[which(contributions$Contributor.Name == "DentaQuest PAC-TN-C" & contributions$Contributor.ID == "1408900"),]$Contributor.ID <- "1408990"
contributions[which(contributions$Contributor.Name == "Abbott Laboratories Employee Political Action Committee - C00040279" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1332968"
contributions[which(contributions$Contributor.Name == "Abbott Laboratories Employee Political Action Committee" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "1332968"
contributions[which(contributions$Contributor.Name == "General Motors Company Political Action Committee (GM PAC)" & is.na(contributions$Contributor.ID)),]$Contributor.ID <- "790461"

```

- was not able to find contributor ids for all the PACs missing ids -> flag those for review

### look at type of contributors

```{r}
unique(x$FILER_TYPE)
major_donor_ids <- x %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(FILER_ID)
major_donor_ids <- major_donor_ids$FILER_ID
#length(which(!(major_donor_ids %in% contributions$Recipient.Committee.ID))) / length(major_donor_ids)
#x %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(FILER_ID, Contributor.Name)
```

- There are 6 contributors that match to Major Donor / Independent expenditure committees
- None of those that are major donors / independent expenditures match to received contributions


```{r}
contributions_full %>% filter(Recipient.Committee.ID %in% major_donor_ids)
```


### standardize contributor info

```{r}
contributions <- contributions %>%
  mutate(
    standardized_name = standardize_names(Contributor.Name) %>% fix_typos() %>% str_squish(),
    #processed_name = standardized_name %>%
    #  remove_pac_info() %>% remove_unit_info() %>% replace_business_text() %>%
    #  str_remove_all("\\s*\\(.*?\\)") %>% str_squish(),
    standardized_employer_name = standardize_names(Contributor.Employer) %>% fix_typos() %>% str_squish(),
    processed_employer_name = standardized_employer_name %>% str_squish(),
    standardized_occupation = standardize_names(Contributor.Occupation),
    processed_occupation    = standardize_occupation_employer(standardized_occupation),
    standardized_city       = standardize_names(Contributor.City),
    standardized_state = standardize_names(Contributor.State) %>% str_squish(),
    zip_code_processed      = substr(Contributor.Zip.Code, 1, 5),
  
    entity_type      = if_else(sapply(standardized_name, is_individual), "individual", "organization"),
    row_id           = row_number()
  )
```

```{r}
#write.csv(contributions,"08_inputs/all_contributors_processed.csv")
contributions <- read.csv("08_inputs/all_contributors_processed.csv")
```


Don't think we need this (not currently taking city/state into account during entity resolution)

```{r,eval=FALSE}
# manually fix weird states
contributions <- contributions %>% mutate(
  standardized_state = if_else(standardized_state == "93" & standardized_city == "BAKERSFIELD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "C" & standardized_city == "LOS ANGELES", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "MJ" & standardized_city == "PARSIPPANY", "NJ",standardized_state),
  standardized_state = if_else(standardized_state == "A" & standardized_city == "FOLSOM", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "AC" & standardized_city == "GLENDALE", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "US" & standardized_city == "HERMITAGE", "TN",standardized_state),
  
  standardized_state = if_else(standardized_state == "" & standardized_city == "LOS ANGELES", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "SACRAMENTO", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "INGLEWOOD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "HAYWARD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "PASADENA", "CA",standardized_state),
)
```

- many contributions missing a city and/or state
  - could fill in some cases using zip code, but not sure that's needed


### Recipient.id error: 


```{r}
contributions %>% filter(Recipient.Committee.ID == std_contributor_id) %>% select(Recipient.Committee,Contributor.Name)#Recipient.Committee.ID,Contributor.ID)
```

- 152 cases of contributor and recipient having the same id
  - California Veterinary Medical Association PAC is clearly an error: all contributors (diff individuals) have the same contributor.id
  - hard to distinguish error from legit cases of someone giving to themselves
- 	International Brotherhood of Electrical Workers Local Union 40 PAC should be 1304081

How to handle this? 
- problem in aggregation
  - wipe these entirely? then some aggregation won't be quite right. is that okay? 
  - some pac will be missing ids... 
- problem when traversing the PAC graph
  - can just flag the issue then
  

# Aggregate contributions by contributor

CALIFORNIA REAL ESTATE POLITICAL ACTION COMMITTEE/FEDERAL - CALIFORNIA ASSOCIATION OF REALTORS (FEDERAL PAC ID #C00083279)
-> this should have been flagged as a federal pac, not sure why we grabbed its contributors...

```{r}
x$Contributor.ID <- as.character(x$Contributor.ID)
contributions$Contributor.ID <- as.character(contributions$Contributor.ID)
```


```{r}
contributions <- contributions %>% 
  left_join(
    id_map %>% select(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>% 
      rename(std_contributor_id = FILER_ID), 
    by = c("Contributor.ID" = "XREF_FILER_ID")
  )
```

```{r}
library(igraph)

# 1. flag and map contributor ids
prepped <- contributions %>%
  mutate(
    has_id       = !is.na(Contributor.ID) & Contributor.ID != "",
    is_fed_pac   = grepl("fed pac|federal pac|#c|id#|fec id|fec report|fed form|committee/federal", tolower(Contributor.Name)),
    is_pac_no_id = !has_id & !is_fed_pac &
                   grepl("\\bPAC\\b|COMMITTEE|POLITICAL ACTION", toupper(Contributor.Name))
  ) %>%
  #left_join(
  #  id_map %>% select(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>%
  #    rename(std_contributor_id = FILER_ID),
  #  by = c("Contributor.ID" = "XREF_FILER_ID")
  #) %>%
  mutate(
    is_self_contribution = !is.na(std_contributor_id) & std_contributor_id == Recipient.Committee.ID,
    unmatched_id        = has_id & is.na(std_contributor_id),
    std_contributor_id   = if_else(is_self_contribution, NA_character_, std_contributor_id),
    is_pac               = !is_self_contribution & has_id & !is_fed_pac
  )

# build per-PAC entity graph
name_nodes <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "") %>%
  distinct(Recipient.Committee.ID, standardized_name) %>%
  mutate(node = paste0("n_", row_number()))

id_nodes <- prepped %>%
  filter(!is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, std_contributor_id) %>%
  mutate(node = paste0("i_", row_number()))

# names that appear with >1 distinct id within the same PAC are ambiguous
ambiguous_within_pac <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "",
         !is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, standardized_name, std_contributor_id) %>%
  group_by(Recipient.Committee.ID, standardized_name) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  distinct(Recipient.Committee.ID, standardized_name)

# edges only for unambiguous 
edges <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "",
         !is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, standardized_name, std_contributor_id) %>%
  anti_join(ambiguous_within_pac, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(name_nodes, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(id_nodes,   by = c("Recipient.Committee.ID", "std_contributor_id")) %>%
  select(from = node.x, to = node.y)

all_nodes <- c(name_nodes$node, id_nodes$node)
g         <- graph_from_data_frame(edges, directed = FALSE, vertices = all_nodes)
comps     <- components(g)

component_df <- tibble(node = names(comps$membership), entity_group = comps$membership)

name_to_group <- name_nodes %>%
  left_join(component_df, by = "node") %>%
  select(Recipient.Committee.ID, standardized_name, name_entity_group = entity_group)

id_to_group <- id_nodes %>%
  left_join(component_df, by = "node") %>%
  select(Recipient.Committee.ID, std_contributor_id, id_entity_group = entity_group)

# assign entity group 
prepped <- prepped %>%
  left_join(name_to_group, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(id_to_group,   by = c("Recipient.Committee.ID", "std_contributor_id")) %>%
  mutate(
    entity_group = case_when(
      !is.na(std_contributor_id) & std_contributor_id != "" ~ id_entity_group,
      !is.na(name_entity_group)                             ~ name_entity_group,
      TRUE ~ as.numeric(max(comps$membership) + row_number()),
    ),
    name_upper     = toupper(standardized_name),
    employer_upper = toupper(coalesce(standardized_employer_name, "")),
    is_dues        = grepl("UNION MEMBER|MEMBER CONTRIBUTIONS|MEMBER DUES|LESS THAN \\$?100|UNDER \\$?100|CONTRIBUTIONS UNDER|UNITEMIZED DUES|LABORERS DUES|LABORER DUES|MEMBERS DUES|INTERMEDIARY FOR MEMBERS|LABORERS TRUST|UNION NO 78\\s+\\d+", name_upper) |
                     grepl("UNION MEMBER|LABORERS|DISTRICT COUNCIL|EMPLOYEES|LOCAL\\s+\\d+|UNION\\s+\\d+|SMW\\s+\\d+|IBEW\\s+\\d+|SEIU\\s+\\d+|UFCW\\s+\\d+|OLCU\\s+\\d+|UNITE HERE\\s+\\d+|AFSCME\\s+\\d+|IRONWORKER\\s+\\d+|UPTE-CWA\\s+\\d+|UNION NO 78\\s+\\d+", employer_upper)
    
    
  ) %>%
  select(-name_entity_group, -id_entity_group, -name_upper, -employer_upper)

#  aggregate 
contributions_agg <- prepped %>%
  group_by(Recipient.Committee.ID, entity_group) %>%
  summarise(
    Amount             = sum(Amount, na.rm = TRUE),
    standardized_name  = {
      ns <- standardized_name[!is.na(standardized_name) & standardized_name != ""]
      if (length(ns) > 0) names(sort(table(ns), decreasing = TRUE))[1] else NA_character_
    },
    std_contributor_id = {
      ids <- std_contributor_id[!is.na(std_contributor_id) & std_contributor_id != ""]
      if (length(ids) > 0) ids[1] else NA_character_
    },
    FILER_TYPE            = {
      types <- FILER_TYPE[!is.na(FILER_TYPE) & FILER_TYPE != ""]
      if (length(types) > 0) names(sort(table(types), decreasing = TRUE))[1] else NA_character_
    },
    is_fed_pac            = any(is_fed_pac,            na.rm = TRUE),
    is_pac_no_id          = any(is_pac_no_id,          na.rm = TRUE),
    is_pac                = any(is_pac,                na.rm = TRUE),
    unmatched_contributor_id         = any(unmatched_id,         na.rm = TRUE), # should change to them all missing this id? could propagate in some cases..
    has_self_contribution = any(is_self_contribution,  na.rm = TRUE),
    is_dues  = any(is_dues, na.rm = TRUE),
    .groups               = "drop"
  ) %>%
  filter(Amount >= 0) %>%
  mutate(is_pac_no_id = is_pac_no_id & is.na(std_contributor_id)) # address cases where some contribs in entity group do have id, even if otherd don't



```


1037597: typos are causing names not to aggregate: FOOTHILL DE ANZA COMMINITY COLLEGE FACULTY ASSOCIATION; FOOTHILL DE ANZA COMMUNITTY COLLEGE FACULTY ASSOCIATION
1486858: same name (missing, didn't aggregate)



### look at no id pacs

- Data is too big to do this step here. Should look at our PAC contributors once they are pulled to see which contributors are missing an id.

```{r}
contributions_agg %>% filter(is_pac_no_id == TRUE)
```

- some of these do have ids in their entity group but were flagged as not having an id..

Cases where name is an exact match to NAML in filer
- filter for recipipent committee and major donors
- nothing with namf - individual -> not important to us

```{r, eval = FALSE}
contributions <- contributions %>% mutate(
  name_v5 = standardize_names(Contributor.Name) %>%
           str_remove_all("\\s*\\(.*?\\)") %>%
          str_squish()
  )
```

```{r, eval = FALSE}
# Build name/ID lookup from full dataset
full_id_lookup <- id_map_full %>%
  filter(!is.na(FILER_ID), FILER_ID != "", (FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" | FILER_TYPE == "RECIPIENT COMMITTEE")) %>%
  mutate(name_v5 = standardize_names(NAML) %>% 
           str_remove_all("\\s*\\(.*?\\)") %>%
          str_squish()) %>%
  distinct(name_v5, FILER_ID) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(FILER_ID)) %>%
  ungroup()


```

```{r, eval = FALSE}
# Propagate IDs to orgs that are missing them
full_id_propagation <- contributions %>%
  filter(is.na(Contributor.ID)) %>%
  select(Contributor.Name, name_v5) %>%
  inner_join(full_id_lookup, by = "name_v5") %>%
  mutate(needs_review = n_ids > 1) %>%
  select(Contributor.Name, propagated_id = FILER_ID, n_ids, needs_review)
```

- some major donors and recipient committees have the same name -> can't tell which is which

```{r, eval = FALSE}
# apply unambiguous ones back to orgs
orgs <- orgs %>%
  left_join(
    full_id_propagation %>% filter(!needs_review) %>% distinct(row_id, propagated_id),
    by = "row_id"
  ) %>%
  mutate(
    Contributor.ID     = coalesce(Contributor.ID, propagated_id),
    has_contributor_id = !is.na(Contributor.ID) & str_trim(as.character(Contributor.ID)) != ""
  ) %>%
  select(-propagated_id)

# inspect review cases
full_id_propagation %>%
  filter(needs_review) %>%
  arrange(Contributor.Name, propagated_id)
```



# CREATE PAC GRAPH

### helper functions

```{r}
get_pac_contributors <- function(pac_id, contributions_agg) {
  contributions_agg %>%
    filter(Recipient.Committee.ID == pac_id)
}

```

```{r}
categorize_contributor <- function(name,amount) {
  name_upper <- toupper(name)
  case_when(
    name_upper == "UNITEMIZED CONTRIBUTIONS" ~ "unitemized",
    amount <= 100 ~ "small_dollar",
    amount < 5000   ~ "under_5k",
    TRUE  ~ "over_5k" 
  )
}
```

- could update to flag medium contributions (< $100)
- contributions <= $100 called unitemized, but should be referred to as small dollar.

```{r}
colnames(contributions_agg)[10] <- "unmatched_contributor_id"
```

```{r}
pac_total_donated <- contributions_agg %>%
  #filter(Contributor.ID != "" & !is.na(Contributor.ID)) %>%
  #filter(std_contributor_id != "" & !is.na(std_contributor_id)) %>%
  #left_join(
  #  id_map %>% select(XREF_FILER_ID, FILER_ID),
    #by = c("Contributor.ID" = "XREF_FILER_ID")
  #  by = c("std_contributor_id" == "FILER_ID")
  #) %>%
  #mutate(
    #unmatched_contributor_id = is.na(FILER_ID), # something wrong with contributor id
    #std_contributor_id       = as.character(FILER_ID)
  #) %>%
  group_by(std_contributor_id,unmatched_contributor_id) %>% #Contributor.ID, unmatched_contributor_id
  summarise(total_donated_all = sum(Amount, na.rm = TRUE), .groups = "drop")

```

- Need to check how much the PAC has donated vs how much it has received

```{r}
which(is.na(contributions_agg$std_contributor_id) & !is.na(contributions_agg$Contributor.ID))
```


### make graph


```{r}
DATA_SUFFICIENCY_THRESHOLD <- 0.75

traverse_pac_graph <- function(seed_ids, ref, id_map, max_depth = 5) {
  queue         <- lapply(seed_ids, function(id) list(id = id, depth = 0, path = id))
  visited       <- character(0)
  pac_summaries <- list()
  over_5k_contribs <- list()

  while (length(queue) > 0) {
    node   <- queue[[1]]; queue <- queue[-1]
    pac_id <- node$id; depth <- node$depth; path <- node$path

    if (pac_id %in% visited || depth >= max_depth) {
      if (depth >= max_depth && !(pac_id %in% visited)) {
        pac_summaries <- append(pac_summaries, list(tibble(
          pac_id = pac_id, depth = depth, path = path, status = "max_depth_reached",
          total_received = NA_real_, unitemized_amount = NA_real_,
          #member_dues_amount = NA_real_, under_5k_amount = NA_real_, over_5k_amount = NA_real_
        )))
      }
      next
    }
    visited <- c(visited, pac_id)
    
    contribs <- get_pac_contributors(pac_id, ref)

    total_donated_all <- pac_total_donated %>%
      filter(!unmatched_contributor_id, std_contributor_id == pac_id) %>%
      pull(total_donated_all)
    total_donated_all <- if (length(total_donated_all) == 0) NA_real_ else total_donated_all
    
    if (is.null(contribs) || nrow(contribs) == 0) {
      pac_summaries <- append(pac_summaries, list(tibble(
        pac_id = pac_id, depth = depth, path = path, status = "unresolved",
        total_received = NA_real_, total_donated_all = total_donated_all,
        data_sufficiency = NA_real_,
        unitemized_amount = NA_real_, #member_dues_amount = NA_real_,
        under_5k_amount = NA_real_, over_5k_amount = NA_real_
      )))
      next
    }
    
    # check data sufficiency
    total_received    <- sum(contribs$Amount, na.rm = TRUE)
    data_sufficiency  <- if (!is.na(total_donated_all) && total_donated_all > 0)
                           total_received / total_donated_all
                         else NA_real_

    
    #contribs <- get_pac_contributors(pac_id, ref)

    #if (is.null(contribs) || nrow(contribs) == 0) {
    #  pac_summaries <- append(pac_summaries, list(tibble(
    #    pac_id = pac_id, depth = depth, path = path, status = "unresolved",
    #    total_received = NA_real_, total_donated_all = NA_real_,
    #    data_sufficiency = NA_real_,
    #    unitemized_amount = NA_real_, #member_dues_amount = NA_real_,
    #    under_5k_amount = NA_real_, over_5k_amount = NA_real_
    #  )))
    #  next
    #}
    
    # check data sufficiency 
    #total_received    <- sum(contribs$Amount, na.rm = TRUE)
    #total_donated_all <- pac_total_donated %>%
    #  filter(!unmatched_contributor_id, std_contributor_id == pac_id) %>%
    #  pull(total_donated_all)
    #total_donated_all <- if (length(total_donated_all) == 0) NA_real_ else total_donated_all
    #data_sufficiency  <- if (!is.na(total_donated_all) && total_donated_all > 0)
    #                       total_received / total_donated_all
    #                     else NA_real_

    pac_status <- case_when(
      !is.na(data_sufficiency) & data_sufficiency < DATA_SUFFICIENCY_THRESHOLD ~ "low_data_sufficiency",
      is.na(data_sufficiency)                                                   ~ "unknown_data_sufficiency",
      TRUE                                                                      ~ "resolved" # change this from being resolved...
    )

    #contribs <- contribs %>%
    #  mutate(contributor_category = categorize_contributor(standardized_name, processed_employer_name, Amount))
    
    contribs <- contribs %>%
      mutate(contributor_category = categorize_contributor(standardized_name, Amount))
    
    # summarize PAC based on what kind of donations it received
    pac_summaries <- append(pac_summaries, list(tibble(
      pac_id             = pac_id,
      depth              = depth,
      path               = path,
      status             = "resolved",
      total_received     = sum(contribs$Amount, na.rm = TRUE), 
      total_donated_all = total_donated_all,
      data_sufficiency  = total_received / total_donated_all,   # NA if total_donated_all == 0
      unitemized_amount  = sum(contribs$Amount[contribs$contributor_category == "unitemized"],  na.rm = TRUE),
      small_dollar_amount = sum(contribs$Amount[contribs$contributor_category == "small_dollar"],  na.rm = TRUE),
      #member_dues_amount = sum(contribs$Amount[contribs$contributor_category == "member_dues"], na.rm = TRUE),
      under_5k_amount    = sum(contribs$Amount[contribs$contributor_category == "under_5k"],    na.rm = TRUE),
      over_5k_amount     = sum(contribs$Amount[contribs$contributor_category == "over_5k"],     na.rm = TRUE)
    )))


    # only process $5k+ contributors further - still want to add member dues even if its over $5k...
    #for(i in seq_len(nrow(contribs %>% filter(contribs$Amount > 5000)))){ # all contributors over $5k, even if union dues...
    for (i in seq_len(nrow(contribs %>% filter(contributor_category == "over_5k")))) {
      row <- (contribs %>% filter(contributor_category == "over_5k"))[i, ]

      if (isTRUE(row$is_fed_pac)) {
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id      = pac_id,
          path               = path,
          depth              = depth + 1,
          contributor_name   = row$standardized_name, # do i want original name here for matching purposes? Contributor.Name
          contributor_id     = row$std_contributor_id, #row$Contributor.ID,
          contributor_type   = "federal_pac",
          contributor_amount = row$Amount
        )))

      #} else if (isTRUE(row$FILER_TYPE == "RECIPIENT COMMITTEE")) { # do want to include independent expenditure committees
       } else if(isTRUE(!is.na(row$std_contributor_id))|| isTRUE(row$is_pac) || isTRUE(row$is_pac_no_id)) { # look at anything with a contributor id
        #new_id   <- if (!is.na(row$std_contributor_id)) row$std_contributor_id else row$Contributor.ID
        new_id <- row$std_contributor_id
        matched  <- !is.na(row$std_contributor_id)
        new_path <- paste0(path, " -> ", new_id)
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id      = pac_id,
          path               = new_path,
          depth              = depth + 1,
          contributor_name   = row$standardized_name, # Contributor.Name?
          contributor_id     = new_id,
          contributor_type   = if (!matched) "pac_unmatched_id" else if (isTRUE(row$FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE")) "major_donor_ie_committee" 
            else if (isTRUE(row$is_pac_no_id)) "pac_no_id" else "pac",
          #filer_type = if(!is.na(row$FILER_TYPE)) row$FILER_TYPE else NA_character_, # see which may be major donors / IE
          contributor_amount = row$Amount
        )))
        if (!is.na(new_id)) {
          queue <- append(queue, list(list(id = new_id, depth = depth + 1, path = new_path)))
        }

      } else {
        # individual or org — record for industry classification
        # could also miss some major donors - need to propagate ids
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id          = pac_id,
          path                   = path,
          depth                  = depth + 1,
          contributor_name       = row$standardized_name, #row$Contributor.Name,
          contributor_id         = NA_character_, # update this, they may have a major donor id
          contributor_type       = "individual_or_org",
          contributor_amount     = row$Amount,
          #contributor_employer   = row$processed_employer_name,
          #contributor_occupation = row$processed_occupation,
          #contributor_city       = row$standardized_city,
          #contributor_state      = row$standardized_state
        )))
      }
    }
  }

  list(
    pac_summaries    = bind_rows(pac_summaries),
    over_5k_contributors = bind_rows(over_5k_contribs)
  )
}

```

- What do do w/ client, slate mailer orgs, payment to influnce and NA?
- only need recipient committee? 


```{r}
# run PAC graph
seed_ids <- unique(x$FILER_ID)
seed_ids <- seed_ids[!is.na(seed_ids) & seed_ids != ""]

result           <- traverse_pac_graph(seed_ids, contributions_agg, id_map, max_depth = 5)
pac_summaries    <- result$pac_summaries
over_5k_contributors <- result$over_5k_contributors
```


```{r}
# add indicator for federal PAC into PAC summaries
fed_seed_ids <- x %>%
  filter(grepl("fed pac|federal pac|#c|id#|fec id|fec report|fed form", tolower(Contributor.Name))) %>%
  pull(FILER_ID) %>%
  unique()

pac_summaries <- pac_summaries %>%
  mutate(is_federal_pac = pac_id %in% fed_seed_ids)

```


```{r}
#write.csv(pac_summaries, "08_inputs/pac_summaries.csv",row.names = FALSE)
#write.csv(over_5k_contributors, "08_inputs/over_5k_contribs.csv",row.names = FALSE)
```


```{r}
contributor_type_lookup <- over_5k_contributors %>%
  filter(!is.na(contributor_id), contributor_type %in% c("major_donor_ie_committee", "pac_unmatched_id", "pac_no_id", "federal_pac")) %>%
  distinct(contributor_id, contributor_type) %>%
  group_by(contributor_id) %>%
  slice(1) %>%
  ungroup()

```

**would be better if we pulled this from contributions_agg?**


- Resolved just means the PAC was found (should change that language)


### check major donors/independent expenditure committees

```{r}
over_5k_contributors %>% filter(contributor_type == "major_donor_ie_committee")
```

- none of the major donors we're seeing have an contributors..

  
### check unmatched ids

```{r}
over_5k_contributors %>% filter(contributor_type == "pac_unmatched_id")
```

- some of these I couldn't find an ID for, others are new... 

```{r}
contributions_agg %>% filter(Recipient.Committee.ID == "1491433")  #filter(grepl("LOCKHEED MARTIN EMPLOYEES",toupper(Contributor.Name)))
contributions %>% filter(Recipient.Committee.ID == "1491433")
contributions %>% filter(Contributor.ID == "1491433")
#id_map_full %>% filter(grepl("LOCKHEED MARTIN",id_map_full$NAML)) %>% select(NAML, FILER_ID) %>% distinct()
```



### look for pacs that are missing contributor.ids

```{r}
# Build name/ID lookup from full dataset
full_id_lookup <- id_map_full %>%
  filter(!is.na(FILER_ID), FILER_ID != "", (FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" | FILER_TYPE == "RECIPIENT COMMITTEE")) %>%
  mutate(name_v5 = standardize_names(NAML) %>% 
           str_remove_all("\\s*\\(.*?\\)") %>%
          str_squish()) %>%
  distinct(name_v5, FILER_ID,FILER_TYPE) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(FILER_ID)) %>%
  ungroup()


```

- check for more than one filer type as well (although that should correspond to more than one id)

```{r}
over_5k_contributors <- over_5k_contributors %>% mutate(
  name_v5 = standardize_names(contributor_name) %>%
           str_remove_all("\\s*\\(.*?\\)") %>%
          str_squish()
  )
```


```{r}
# Propagate IDs to orgs that are missing them
full_id_propagation <- over_5k_contributors %>%
  filter(is.na(contributor_id)) %>%
  select(contributor_name, name_v5) %>%
  inner_join(full_id_lookup, by = "name_v5", relationship = "many-to-many") %>%
  mutate(needs_review = n_ids > 1) %>%
  select(contributor_name, name_v5, propagated_id = FILER_ID, n_ids, FILER_TYPE, needs_review)
```

```{r}
full_id_propagation #%>% filter(FILER_TYPE == "RECIPIENT COMMITTEE")
```

```{r}
full_id_propagation %>% filter(n_ids > 1) #filter(FILER_TYPE == "RECIPIENT COMMITTEE")
```

Matched to more than one potential id: 
- CALIFORNIA-NEVADA CONFERENCE OF OPERATING ENGINEERS
- CALIFORNIA CREDIT UNION LEAGUE (CCUL ADMIN)
- CALIFORNIA CREDIT UNION LEAGUE (CCUL INDEPENDENT EXPENDITURE ACCOUNT)
  - can assume this one IE is correct..
  
```{r}
ie_ids <- full_id_propagation %>% filter(FILER_TYPE == "RECIPIENT COMMITTEE")
```

  
```{r}
ie_ids <- full_id_propagation %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(propagated_id)
contributions %>% filter(Recipient.Committee.ID %in% ie_ids$propagated_id)
```


- not sure we should be taking parenthese out, but otherwise 
- some of these recipient committees don't show any recieved contributions.
- most are major donor/indepedent expenditure committee as expected
  - look to see if any of these ids have any received contributions
- 8 Recipient committees that didn't have an id
- WINE INSTITUTE, THE (major donor), WINE INSTITUTE (recipient committee) -> can't tell the different in name matching, it matches to recipient committee, but hard to say which is which..

Not sure these all make sense: 
- HOD CARRIERS AND LABORERS LOCAL 783 (none have ids) contributed to 	LABORERS LOCAL UNION NO. 783-PAC and these are that PACs only contributors
  - but recipient committee id is different from id associated with HOD CARRIERS AND LABORERS LOCAL 783 so maybe they're not the same? 
- CALIFORNIA CREDIT UNION LEAGUE (CCUL INDEPENDENT EXPENDITURE ACCOUNT)
  - treat as an indepedent expenditure? 

  
### Add indicators for whether we are able to classify each PAC in theory


```{r}
UNITEMIZED_DUES_THRESHOLD <- 0.5
OVER_5K_COVERAGE_THRESHOLD <- 0.5

pac_classifiability <- pac_summaries %>%
  mutate(
    low_data_sufficiency     = !is.na(data_sufficiency) & data_sufficiency < DATA_SUFFICIENCY_THRESHOLD,
    unknown_data_sufficiency = is.na(data_sufficiency),
    data_ok                  = status == "resolved" & !low_data_sufficiency & !unknown_data_sufficiency,
    #pct_unitemized_dues      = (unitemized_amount + member_dues_amount) / total_received,
    pct_unitemized = unitemized_amount / total_received,
    pct_small_dollar = small_dollar_amount / total_received,
    pct_over_5k              = over_5k_amount / total_received,
    #classify_by_name         = !is.na(pct_unitemized_dues) & pct_unitemized_dues >= UNITEMIZED_DUES_THRESHOLD & data_ok, # getting rid of this
    data_ok                  = status == "resolved" & !is_federal_pac & !low_data_sufficiency & !unknown_data_sufficiency
  )

```

- This is only a preliminary indicator for whether we can classify that PAC, we need to see whether or not we can actually classify its contributors.

```{r}
min(pac_classifiability$pct_over_5k,na.rm = TRUE)
max(pac_classifiability$pct_over_5k,na.rm = TRUE)
```

```{r}
pac_classifiability %>% filter(pct_over_5k >= 1)
```


### flag each $5k contributor as resolvable or not

```{r}
# look at the status of each contributor.
over_5k_tagged <- over_5k_contributors %>%
  left_join(
    pac_classifiability %>% select(pac_id,data_ok, pct_over_5k, status,low_data_sufficiency), # classify_by_name
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    is_resolvable = case_when(
      contributor_type == "individual_or_org"                                        ~ TRUE,
      #contributor_type == "major_donor_ie_committee" ~ TRUE, # change this to look for PAC/committee language in the name
      #contributor_type == "pac" & classify_by_name                                   ~ TRUE,
      contributor_type == "pac" & data_ok & pct_over_5k >= OVER_5K_COVERAGE_THRESHOLD ~ TRUE,
      TRUE                                                                           ~ FALSE
    ),
    unresolvable_reason = case_when(
      !is_resolvable & contributor_type == "federal_pac"                             ~ "federal_pac",
      !is_resolvable & contributor_type %in% c("pac_no_id", "pac_unmatched_id")     ~ contributor_type,
      !is_resolvable & contributor_type == "major_donor_ie_committee"                 ~ "major_donor_ie_committee",
      #!is_resolvable & contributor_type == "pac" & !data_ok                         ~ paste0("pac_", status),
      !is_resolvable & low_data_sufficiency  ~ "low_data_sufficiency",
      !is_resolvable & contributor_type == "pac" & data_ok                          ~ "pac_low_5k_coverage",
      TRUE                                                                           ~ NA_character_
    )
  )

```


### compute % of each PAC contributors that can be classified 

```{r}
# NEWER VERSION


# pre-compute PAC contributor dependencies once — avoids filtering over_5k_tagged in every loop iteration
pac_deps <- over_5k_tagged %>%
  filter(contributor_type %in% c("pac", "major_donor_ie_committee"), !is.na(contributor_id)) %>%
  distinct(parent_pac_id, contributor_id) %>%
  group_by(parent_pac_id) %>%
  summarise(deps = list(contributor_id), .groups = "drop") %>%
  { setNames(.$deps, .$parent_pac_id) }

# topological sort with break on no-progress (handles cycles / missing deps)
processed     <- character(0)
ordered_pacs  <- character(0)
pacs_to_order <- pac_summaries %>% arrange(desc(depth)) %>% pull(pac_id)

while (length(pacs_to_order) > 0) {
  n_before <- length(pacs_to_order)
  ready <- Filter(function(pid) {
    deps <- pac_deps[[pid]]
    is.null(deps) || all(deps %in% processed)
  }, pacs_to_order)
  ordered_pacs  <- c(ordered_pacs, ready)
  processed     <- c(processed, ready)
  pacs_to_order <- setdiff(pacs_to_order, ready)
  if (length(pacs_to_order) == n_before) break  # no progress: cycles or missing deps
}

stuck_pacs <- pacs_to_order

# pre-split by parent_pac_id so inner lookup is O(1) instead of filtering 1.8M rows
over_5k_by_pac    <- split(over_5k_tagged, over_5k_tagged$parent_pac_id)
classifiability_idx <- setNames(seq_len(nrow(pac_classifiability)), pac_classifiability$pac_id)
resolvable_vec      <- setNames(rep(NA_real_,  nrow(pac_classifiability)), pac_classifiability$pac_id)

# ── pass 1: process orderable PACs ───────────────────────────────────────────
for (pid in ordered_pacs) {
  summary_row <- pac_classifiability[classifiability_idx[[pid]], ]
  if (!summary_row$data_ok) { resolvable_vec[[pid]] <- 0; next }

  direct_contribs <- over_5k_by_pac[[pid]]
  if (is.null(direct_contribs) || nrow(direct_contribs) == 0) { resolvable_vec[[pid]] <- 0; next }

  direct_contribs <- direct_contribs %>%
    mutate(
      child_resolvable_fraction = resolvable_vec[contributor_id],
      resolvable_amount = case_when(
        contributor_type == "individual_or_org"        ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" ~ 0, #contributor_amount, # depends on how we end up dealing with these
        contributor_type == "pac"                      ~ contributor_amount * coalesce(child_resolvable_fraction, 0),
        TRUE                                           ~ 0
      )
    )

  resolvable_vec[[pid]] <- sum(direct_contribs$resolvable_amount, na.rm = TRUE) / summary_row$total_received
}

# process stuck pacs
remaining <- stuck_pacs

compute_fraction <- function(pid) {
  summary_row <- pac_classifiability[classifiability_idx[[pid]], ]
  if (!summary_row$data_ok) { resolvable_vec[[pid]] <<- 0; return() }

  direct_contribs <- over_5k_by_pac[[pid]]
  if (is.null(direct_contribs) || nrow(direct_contribs) == 0) { resolvable_vec[[pid]] <<- 0; return() }

  direct_contribs <- direct_contribs %>%
    mutate(
      child_resolvable_fraction = coalesce(resolvable_vec[contributor_id], 0),
      resolvable_amount = case_when(
        contributor_type == "individual_or_org"        ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" ~ 0, #contributor_amount,
        contributor_type == "pac"                      ~ contributor_amount * child_resolvable_fraction,
        TRUE                                           ~ 0
      )
    )

  resolvable_vec[[pid]] <<- sum(direct_contribs$resolvable_amount, na.rm = TRUE) / summary_row$total_received
}

while (length(remaining) > 0) {
  # PACs whose stuck deps have all been resolved can now be processed normally
  ready <- Filter(function(pid) {
    deps <- pac_deps[[pid]]
    is.null(deps) || !any(deps %in% remaining)
  }, remaining)

  if (length(ready) == 0) {
    # nothing can be unblocked — true cycle remains; force all with cyclic deps = 0
    ready <- remaining
  }

  for (pid in ready) compute_fraction(pid)
  remaining <- setdiff(remaining, ready)
}

pac_resolvable_fractions <- pac_classifiability %>%
  mutate(resolvable_fraction = resolvable_vec[pac_id])

```


### check PACs that were causing an infinite loop


```{r}
# PACs left unordered after the loop
stuck_pacs <- pacs_to_order

# For each stuck PAC, show which of its dependencies were never processed
stuck_details <- lapply(stuck_pacs, function(pid) {
  deps <- pac_deps[[pid]]
  tibble(
    pac_id          = pid,
    unmet_dep       = deps[!deps %in% processed]
  )
}) %>%
  bind_rows()

print(stuck_details)

```

```{r}
stuck_details %>%
  mutate(
    in_pac_list = unmet_dep %in% pac_summaries$pac_id,
    cause       = if_else(in_pac_list, "circular_dependency", "id_not_in_pac_list")
  )

```

```{r}
stuck_details <- stuck_details %>%
  mutate(in_pac_list = unmet_dep %in% pac_summaries$pac_id)

stuck_edges <- stuck_details %>%
  filter(in_pac_list, unmet_dep %in% stuck_pacs)

g_stuck <- graph_from_data_frame(
  stuck_edges %>% select(from = pac_id, to = unmet_dep),
  directed = TRUE,
  vertices = unique(c(stuck_edges$pac_id, stuck_edges$unmet_dep))
)

sccs <- components(g_stuck, mode = "strong")
scc_df <- tibble(
  pac_id   = names(sccs$membership),
  scc_id   = sccs$membership,
  scc_size = sccs$csize[sccs$membership]
)

stuck_details <- stuck_details %>%
  left_join(scc_df, by = "pac_id") %>%
  mutate(
    cause = case_when(
      !in_pac_list  ~ "id_not_in_pac_list",
      scc_size > 1  ~ "true_circular_dependency",
      TRUE          ~ "cascading_from_stuck_dep"
    )
  )

stuck_details %>% count(cause)


```

- Grove for Board of Equalization 2026 -> 	Reform California with Carl DeMaio - Ballot Measure Committee ~$6k (2025-2026)
  - Carl DeMaio for State Assembly 2026 -> Grove For Board Of Equalization 2026 ~$5k (2026)
- Housing Providers for Responsible Solutions (1446977) -> Committee to Support the Alliance of California's Farmers and Ranchers (1411956) ~ $21k
   - don't see the reverse contribution though... 
-	LCCC PAC, A Bi-Partisan Coalition of Municipal, County, Latino Appointed and Elected Leaders (1432932) -> Avila Farias for State Assembly 2026 (1478170) ~ $6k
  - do see this

PACs can be stuck for different reasons: 
1. circular dependency - PAC A lists PAC B as a contributor and vice versa
2. Contributor.ID not in PAC list (this shouldn't be happening)

- What happens to the circular dependency PACs when computing resolvability? 
  - can the circular part just get flag as unresolvable? 


**look at order of PACs associated with multiple depths to make sure its working**


### coverage breakdown for each seed PAC

```{r}
seed_coverage <- over_5k_tagged %>%
  filter(depth == 1, parent_pac_id %in% seed_ids) %>%
  left_join(
    pac_resolvable_fractions %>% select(pac_id, resolvable_fraction),
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    resolvable_amount = case_when(
      contributor_type == "individual_or_org" ~ contributor_amount,
      contributor_type == "major_donor_ie_committee" ~ 0, #contributor_amount, # again may need to change
      contributor_type == "pac"               ~ contributor_amount * coalesce(resolvable_fraction, 0),
      TRUE                                    ~ 0
    )
  ) %>%
  group_by(parent_pac_id) %>%
  summarise(
    total_direct_over_5k  = sum(contributor_amount, na.rm = TRUE),
    total_resolvable      = sum(resolvable_amount, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    pac_classifiability %>% filter(pac_id %in% seed_ids) %>%
      select(pac_id, total_received, data_ok, status, data_sufficiency), # classify_by_name
    by = c("parent_pac_id" = "pac_id")
  ) %>%
  mutate(
    pct_resolvable      = total_resolvable / total_received,
    classification_status = case_when(
      #classify_by_name                              ~ "classify_by_name",
      data_ok & pct_resolvable >= OVER_5K_COVERAGE_THRESHOLD ~ "classify_by_contributors",
      TRUE                                          ~ "unresolvable"
    )
  ) 

```

```{r}
pac_names <- x %>% group_by(FILER_ID) %>% slice(1) %>% ungroup() %>% select(FILER_ID, Contributor.Name)
seed_coverage <- seed_coverage %>% left_join(pac_names, by = c("parent_pac_id" = "FILER_ID"))
```




```{r}
# get summary for all seed pacs
all_seed_status <- pac_classifiability %>%
  filter(pac_id %in% seed_ids) %>%
  left_join(
    seed_coverage %>%
      select(parent_pac_id, total_direct_over_5k, total_resolvable, pct_resolvable, classification_status),
    by = c("pac_id" = "parent_pac_id")
  ) %>%
  left_join(contributor_type_lookup, by = c("pac_id" = "contributor_id")) %>%
  mutate(
    classification_status = case_when(
      !is.na(classification_status)                                   ~ classification_status,
      #classify_by_name                                                ~ "classify_by_name",
      status %in% c("unresolved", "low_data_sufficiency",
                    "unknown_data_sufficiency")                       ~ "unresolvable",
      TRUE                                                            ~ "unresolvable"
    ),
    pct_resolvable       = coalesce(pct_resolvable, 0),
    total_direct_over_5k = coalesce(total_direct_over_5k, 0),
    total_resolvable     = coalesce(total_resolvable, 0),
    unresolvable_reason  = case_when(
      classification_status != "unresolvable"  ~ NA_character_,
      is_federal_pac    ~ "federal_pac",    
      !is.na(contributor_type)                   ~ contributor_type, 
      status == "unresolved"  ~ "no_contributor_data",
      low_data_sufficiency                                ~ "low_data_sufficieny", 
      unknown_data_sufficiency                           ~ "unknown_data_sufficiency", 
      total_direct_over_5k == 0                                      ~ "no_over_5k_contributors",
      pct_resolvable < OVER_5K_COVERAGE_THRESHOLD                    ~ "low_pct_resolvable", 
      TRUE                                                            ~ "other"
    )
  )


all_seed_status <- all_seed_status %>% left_join(pac_names, by = c("pac_id" = "FILER_ID"))
```

```{r}
all_seed_status %>% filter(unresolvable_reason == "no_contributor_data") #%>% count()
all_seed_status %>% filter(unresolvable_reason == "low_data_sufficieny") #%>% count()
all_seed_status %>% filter(unresolvable_reason == "low_pct_resolvable") #%>% count()
all_seed_status %>% filter(unresolvable_reason == "no_over_5k_contributors")# %>% count()
all_seed_status %>% filter(unresolvable_reason == "federal_pac")# %>% count()
```


### check unmatched contributor id cases

```{r}
all_seed_status %>% filter(unresolvable_reason == "pac_unmatched_id")
```


```{r}
over_5k_tagged %>% filter(unresolvable_reason == "pac_unmatched_id") %>% select(contributor_name, parent_pac_id)
```

```{r}
contributions %>% filter(Recipient.Committee.ID == 	"1026324", grepl("general motors",tolower(Contributor.Name))) %>% select(Contributor.Name,Contributor.ID)
```


### check no over 5k contributor PACs

```{r}
all_seed_status %>% filter(unresolvable_reason == "no_over_5k_contributors") %>% select(Contributor.Name,pac_id)
```

```{r}
pac_summaries %>% filter(over_5k_amount == 0, is_federal_pac == FALSE)
```


```{r}
contributions_agg %>% filter(Recipient.Committee.ID == "1286381") %>% select(standardized_name,Amount) %>% arrange(desc(Amount))
```

- checked a handful manually and seems true

### manually check the no contributor data PACs

- some of these are major donor / IE
  - looking up a hanful of IE expender ids in cal-access -> cal-access calls the recipient committees

```{r}
all_seed_status %>% filter(unresolvable_reason == "no_contributor_data") %>% select(Contributor.Name,pac_id, depth)
```

Eloise Reyes for Senate 2024: none in 2025-26

- don't just want to check seed PACs that received no contributions, want to check all PACs that didn't receive any contributions


```{r}
seed_no_data <- pac_summaries %>%
  filter(!pac_id %in% contributions_agg$Recipient.Committee.ID) %>%
  select(pac_id) %>%
  left_join(
    over_5k_tagged %>% distinct(parent_pac_id, parent_pac_name),
    by = c("pac_id" = "parent_pac_id")
  )


cat("Seed PACs with no contributor data:", nrow(seed_no_data), "\n")
print(seed_no_data)


contributing_match <- over_5k_tagged %>%
  filter(contributor_type %in% c("pac", "major_donor_ie_committee"),
         !is.na(contributor_id)) %>%
  distinct(contributor_id, contributor_name, contributor_type) %>%
  mutate(has_contributor_data = contributor_id %in% contributions_agg$Recipient.Committee.ID)

contributing_match %>% count(contributor_type, has_contributor_data)

contributing_match %>%
  left_join(id_map %>% distinct(FILER_ID, FILER_TYPE), by = c("contributor_id" = "FILER_ID")) %>%
  count(contributor_type, FILER_TYPE, has_contributor_data) %>%
  arrange(contributor_type, FILER_TYPE)

```

```{r}
# collect names from all available sources
pac_names <- bind_rows(
  all_seed_status %>% distinct(pac_id, name = Contributor.Name),
  over_5k_tagged  %>% distinct(pac_id = contributor_id, name = contributor_name)
) %>%
  group_by(pac_id) %>%
  slice(1) %>%
  ungroup()

pacs_no_data <- pac_summaries %>%
  filter(!pac_id %in% contributions_agg$Recipient.Committee.ID) %>%
  left_join(pac_names,                                        by = "pac_id") %>%
  left_join(id_map %>% distinct(FILER_ID, FILER_TYPE),        by = c("pac_id" = "FILER_ID")) %>%
  mutate(is_seed = pac_id %in% all_seed_status$pac_id) %>%
  select(pac_id, name, FILER_TYPE, is_seed)

pacs_no_data %>% count(FILER_TYPE, is_seed)
print(pacs_no_data)

```


```{r}
contributions_agg %>% filter(Recipient.Committee.ID == "1030699") %>% arrange(desc(Amount))
```


```{r}
pacs_no_data
```
- client = lobbyist client
  - 1386983 (ILWU LOCAL 13 POLITICAL ACTION COMMITTEE FUND)
  - not sure what to do with these
- is there any chance those without received or made contributions are Fed PACS?

ONES WITH CONTRIBUTIONS RECEIVED: 
- check for any on dates 05/26 -> couldve gotten skipped somehow? 

- 1047877	Plumbers and Steamfitters Local Union No. 114 Political Action Committee
  - 05/26/26; filed 07/30
- 1047876	United Association of Journeymen & Apprentices of the Plumbing & Pipefitting Industry Local Union Number 345 Political Action Committee
  - 05/26/26; filed 07/30
- 1475124	Elect Dems Now PAC Fed ID# C00880062
  - 06/30/2026 (filed date 07/31/2026) - may not be in there yet? 
- 1490566	Californians for Local Affordable Solar and Storage (CLASS) PAC
  - all filed 07/31/26
  
  
  FOUND CONTRIBUTORS!!!
  
```{r}
id_map %>% filter(FILER_ID == "1059236")
```


```{r}
#contributions %>% filter(Recipient.Committee.ID == "890465")
contributions %>% filter(grepl("114",Recipient.Committee)) %>% select(Recipient.Committee,Recipient.Committee.ID) %>% distinct()
contributions %>% filter(Contributor.ID == "890465")
```


```{r}
contributions_full %>% filter(Recipient.Committee.ID == "890465")
#contributions_full %>% filter(grepl("plumbers",tolower(Recipient.Committee))) %>% select(Recipient.Committee,Recipient.Committee.ID) %>% distinct()
#contributions_full %>% filter(grepl("114",tolower(Contributor.Name))) %>% select(Contributor.Name,Contributor.ID,Recipient.Committee,Recipient.Committee.ID) %>% distinct()
```



-1475565: no contributions received found, but also no contributions made were found, which is weird
-930791: another example of this - can't find contributions made in cal-access, but they are in this data?
- what to do w/ slate mailer (1446747 - Reform CA Voter Guide)
- 

```{r}
contributing_match %>%
  filter(contributor_type == "major_donor_ie_committee", has_contributor_data) %>%
  select(contributor_id, contributor_name) %>%
  left_join(
    contributions %>%
      group_by(Recipient.Committee.ID) %>%
      summarise(n_contributors = n(), total_received = sum(Amount, na.rm = TRUE), .groups = "drop"),
    by = c("contributor_id" = "Recipient.Committee.ID")
  )

```

- Nothing thats an IE/major donor appears to match to any contributors in the data


### remaining


```{r}
all_seed_status %>% filter(unresolvable_reason == "federal_pac")
```

```{r}
table(all_seed_status$unresolvable_reason)
```

- 4 federal PACs (thought there were 6?)
- 42 PACs have contributed much more than they received 
- 14 PACs have <50% of contributions that came from $5k+
- 31 PACs have no contributor data in 2026
- 1 PAC has no contributions over $5k (all contributions greater than 100, but less than 5k - this is a dairies PAC, all its contributors are dairies)
- 16 PACs (10%) have more than 50% of their contributions come from $5k+ donations and have contributors we are able to classify (as opposed to unresolvable PACs)
  - For many of these, we'd still only be classifying around 60% of its received contributions
  

```{r}
#write.csv(all_seed_status,"08_inputs/classifiability_by_pac.csv")
```


### look at federal PACs

```{r}
fed_pac_ids <- all_seed_status %>% filter(is_federal_pac == TRUE) %>% select(pac_id)
x %>% filter(x$FILER_ID %in% fed_pac_ids$pac_id) %>% summarise(total = sum(Amount))
```
- $108k spent directly from federal PACs on races

```{r}
over_5k_contributors %>% filter(contributor_type == "federal_pac") %>% summarise(total = sum(contributor_amount))
```
- $63k spent by federal PACs to PAC contributors


# Send PACs to review queue
  
```{r}
# flag member dues in over_5k_tagged (name- and employer-based)
over_5k_tagged <- over_5k_tagged %>%
  mutate(
    name_upper     = toupper(contributor_name),
    is_member_dues = grepl(
      "UNION MEMBER|MEMBER CONTRIBUTIONS|MEMBER DUES|LESS THAN \\$?100|UNDER \\$?100|CONTRIBUTIONS UNDER|UNITEMIZED DUES|LABORERS DUES|LABORER DUES|MEMBERS DUES|INTERMEDIARY FOR MEMBERS|LABORERS TRUST",
      name_upper
    ) 
  ) %>%
  select(-name_upper)


```

```{r}
pac_names_from_transactions <- x %>% group_by(FILER_ID) %>% slice(1) %>% ungroup() %>% select(FILER_ID, Contributor.Name)

# fallback for PACs that never show up as a $5k+ contributor themselves in x
# (only as a recipient/parent PAC here) -- id_map_full is CalAccess's full
# filer registry, so NAML covers every registered committee regardless of
# whether it appears as a contributor in this data
pac_names_registry <- id_map_full %>%
  group_by(FILER_ID) %>%
  slice(1) %>%
  ungroup() %>%
  transmute(FILER_ID, Contributor.Name = NAML)

pac_names_registry$FILER_ID <- as.character(pac_names_registry$FILER_ID)
pac_names <- bind_rows(
  pac_names_from_transactions,
  pac_names_registry %>% filter(!FILER_ID %in% pac_names_from_transactions$FILER_ID)
)

```


```{r}
contributor_review_queue <- over_5k_tagged %>%
  filter(contributor_type %in% c("major_donor_ie_committee", "pac_no_id", "pac_unmatched_id", "federal_pac")) %>%
  left_join(
    id_map %>% distinct(FILER_ID, FILER_TYPE),
    by = c("contributor_id" = "FILER_ID")
  ) %>%
  select(
    parent_pac_id,
    contributor_name,
    contributor_id,
    contributor_type,
    FILER_TYPE,
    contributor_amount,
    review_reason = unresolvable_reason
  )

```

```{r}
#review_pacs <- all_seed_status %>%
#  filter(classification_status == "unresolvable" | pct_resolvable < 1) %>%
#  pull(pac_id)

review_pacs <- pac_resolvable_fractions %>%
  filter(is.na(resolvable_fraction) | resolvable_fraction < 1) %>%
  pull(pac_id)


# (recipient, name) pairs at $5k+ in aggregate — exclude from non-$5k list
over_5k_name_pairs <- contributions_agg %>%
  filter(Amount >= 5000) %>%
  select(Recipient.Committee.ID, standardized_name)

# column 1: sub-$5k contributor labels (org name or individual employer), excluding member dues
non_5k_labels <- contributions %>%
  filter(Recipient.Committee.ID %in% review_pacs, Amount > 0) %>%
  anti_join(over_5k_name_pairs, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  mutate(
    name_upper     = toupper(standardized_name),
    employer_upper = toupper(coalesce(standardized_employer_name, "")),
    entity_type    = if_else(sapply(standardized_name, is_individual), "individual", "organization"),
    is_dues        = 
      grepl("UNION MEMBER|MEMBER CONTRIBUTIONS|MEMBER DUES|LESS THAN \\$?100|UNDER \\$?100|CONTRIBUTIONS UNDER|UNITEMIZED DUES|LABORERS DUES|LABORER DUES|MEMBERS DUES|INTERMEDIARY FOR MEMBERS|LABORERS TRUST", name_upper) |
      grepl("UNION MEMBER|LABORERS|DISTRICT COUNCIL|EMPLOYEES|LOCAL\\s+\\d+|UNION\\s+\\d+|SMW\\s+\\d+|IBEW\\s+\\d+|SEIU\\s+\\d+|UFCW\\s+\\d+|OLCU\\s+\\d+|UNITE HERE\\s+\\d+|AFSCME\\s+\\d+|IRONWORKER\\s+\\d+|UPTE-CWA\\s+\\d+", employer_upper),
    label = case_when(
      is_dues ~ NA_character_,
      entity_type == "individual" &
        !is.na(standardized_employer_name) &
        !standardized_employer_name %in% c("", "NONE", "N/A", "NA", "RETIRED", "NOT EMPLOYED", "UNEMPLOYED", "SELF EMPLOYED") ~ str_squish(standardized_employer_name),
      TRUE ~ standardized_name
    )

  ) %>%
  filter(!is.na(label) & label != "") %>%
  group_by(Recipient.Committee.ID) %>%
  summarise(non_5k_contributors = paste(sort(unique(label)), collapse = " | "), .groups = "drop")

# column 2: unresolvable $5k+ contributors with reason (excluding member dues)
unresolvable_5k_labels <- over_5k_tagged %>%
  filter(parent_pac_id %in% review_pacs, !is_resolvable, !is_member_dues) %>%
  mutate(label = paste0(contributor_name, " (", coalesce(unresolvable_reason, "unknown"), ")")) %>%
  group_by(parent_pac_id) %>%
  summarise(unresolvable_5k_contributors = paste(label, collapse = " | "), .groups = "drop")

# member dues amounts: sum all contributions (any size) matching name or employer patterns
#member_dues_amounts <- contributions_agg %>% # this is an issue b/c its not aggregated, will over count... (w/ aggregation step and filtering out <0s)
#  filter(Amount > 0) %>%
#  mutate(
#    name_upper     = toupper(standardized_name),
#    employer_upper = toupper(coalesce(standardized_employer_name, "")),
#    is_dues        = grepl("UNION MEMBER|MEMBER CONTRIBUTIONS|MEMBER DUES|LESS THAN \\$?100|UNDER \\$?100|CONTRIBUTIONS UNDER|UNITEMIZED DUES|LABORERS DUES|LABORER DUES|MEMBERS DUES|INTERMEDIARY FOR MEMBERS|LABORERS #TRUST|UNION NO 78\\s+\\d+", name_upper) |
#                     grepl("UNION MEMBER|LABORERS|DISTRICT COUNCIL|EMPLOYEES|LOCAL\\s+\\d+|UNION\\s+\\d+|SMW\\s+\\d+|IBEW\\s+\\d+|SEIU\\s+\\d+|UFCW\\s+\\d+|OLCU\\s+\\d+|UNITE #HERE\\s+\\d+|AFSCME\\s+\\d+|IRONWORKER\\s+\\d+|UPTE-CWA\\s+\\d+|UNION NO 78\\s+\\d+", employer_upper)
#  ) %>%
#  filter(is_dues) %>%
#  group_by(Recipient.Committee.ID) %>%
#  summarise(member_dues_amount = sum(Amount, na.rm = TRUE), .groups = "drop")

member_dues_amounts <- contributions_agg %>%
  filter(Amount > 0, is_dues) %>%
  group_by(Recipient.Committee.ID) %>%
  summarise(member_dues_amount = sum(Amount, na.rm = TRUE), .groups = "drop")


# assemble review queue
#pac_review_queue <- all_seed_status %>%
#  filter(classification_status == "unresolvable" | pct_resolvable < 1) %>%
#  left_join(non_5k_labels,          by = c("pac_id" = "Recipient.Committee.ID")) %>%
#  left_join(unresolvable_5k_labels, by = c("pac_id" = "parent_pac_id")) %>%
#  left_join(member_dues_amounts,    by = c("pac_id" = "Recipient.Committee.ID")) %>%
#  mutate(
#    member_dues_amount = coalesce(member_dues_amount, 0),
#    pct_member_dues    = member_dues_amount / total_received
#  ) %>%
#  select(pac_id, Contributor.Name, unresolvable_reason,
#         pct_resolvable, total_received, total_donated_all, data_sufficiency,
#         pct_member_dues, pct_unitemized, pct_over_5k,
#         total_received, total_resolvable,
#         non_5k_contributors, unresolvable_5k_contributors)

#pac_review_queue <- pac_resolvable_fractions %>%
#  filter(is.na(resolvable_fraction) | resolvable_fraction < 1) %>%
#  left_join(pac_names, by = "pac_id") %>%
#  mutate(is_seed = pac_id %in% all_seed_status$pac_id) %>%
#  left_join(non_5k_labels,          by = c("pac_id" = "Recipient.Committee.ID")) %>%
#  left_join(unresolvable_5k_labels, by = c("pac_id" = "parent_pac_id")) %>%
#  left_join(member_dues_amounts,    by = c("pac_id" = "Recipient.Committee.ID")) %>%
#  mutate(
#    member_dues_amount = coalesce(member_dues_amount, 0),
#    pct_member_dues    = member_dues_amount / total_received
#  ) %>%
#  select(pac_id, name, is_seed, resolvable_fraction, total_received,
#         non_5k_contributors, unresolvable_5k_contributors,
#         pct_member_dues, member_dues_amount)

# generalize seed_coverage to ALL PACs in the graph
all_pac_coverage <- over_5k_tagged %>%
  left_join(
    pac_resolvable_fractions %>% select(pac_id, child_resolvable_fraction = resolvable_fraction),
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    resolvable_amount = case_when(
      contributor_type == "individual_or_org"        ~ contributor_amount,
      contributor_type == "major_donor_ie_committee" ~ 0, #contributor_amount,
      contributor_type == "pac"                      ~ contributor_amount * coalesce(child_resolvable_fraction, 0),
      TRUE                                           ~ 0
    )
  ) %>%
  group_by(parent_pac_id) %>%
  summarise(
    total_direct_over_5k = sum(contributor_amount, na.rm = TRUE),
    total_resolvable     = sum(resolvable_amount,  na.rm = TRUE),
    .groups = "drop"
  )

```

```{r}
pac_review_queue <- pac_resolvable_fractions %>%
  filter(is.na(resolvable_fraction) | resolvable_fraction < 1) %>%
  left_join(pac_names,        by = c("pac_id" = "FILER_ID")) %>%
  left_join(all_pac_coverage, by = c("pac_id" = "parent_pac_id")) %>%
  left_join(contributor_type_lookup, by = c("pac_id" = "contributor_id")) %>%
  mutate(
    is_seed              = pac_id %in% all_seed_status$pac_id,
    total_direct_over_5k = coalesce(total_direct_over_5k, 0),
    total_resolvable     = coalesce(total_resolvable, 0),
    pct_resolvable       = coalesce(resolvable_fraction, 0),
    classification_status = case_when(
      data_ok & pct_resolvable >= OVER_5K_COVERAGE_THRESHOLD ~ "classify_by_contributors",
      TRUE                                                    ~ "unresolvable"
    ),
    unresolvable_reason = case_when(
      classification_status != "unresolvable" ~ NA_character_,
      is_federal_pac                          ~ "federal_pac",
      !is.na(contributor_type)                  ~ contributor_type,
      status == "unresolved"                  ~ "no_contributor_data",
      low_data_sufficiency                    ~ "low_data_sufficieny",
      unknown_data_sufficiency                ~ "unknown_data_sufficiency",
      total_direct_over_5k == 0              ~ "no_over_5k_contributors",
      pct_resolvable < OVER_5K_COVERAGE_THRESHOLD ~ "low_pct_resolvable",
      TRUE                                    ~ "other"
    )
  ) %>%
  left_join(non_5k_labels,          by = c("pac_id" = "Recipient.Committee.ID")) %>%
  left_join(unresolvable_5k_labels, by = c("pac_id" = "parent_pac_id")) %>%
  left_join(member_dues_amounts,    by = c("pac_id" = "Recipient.Committee.ID")) %>%
  mutate(
    member_dues_amount = coalesce(member_dues_amount, 0),
    pct_member_dues    = member_dues_amount / total_received
  ) %>%
  #rename(Contributor.Name = name) %>%
  select(
    pac_id, Contributor.Name, is_seed,
    status, unresolvable_reason,
    is_federal_pac, low_data_sufficiency, unknown_data_sufficiency, data_ok,
    total_received, total_donated_all, data_sufficiency,
    unitemized_amount, small_dollar_amount, under_5k_amount, over_5k_amount,
    pct_unitemized, pct_small_dollar, pct_over_5k,
    total_direct_over_5k, total_resolvable, pct_resolvable,
    classification_status, unresolvable_reason,
    pct_member_dues, member_dues_amount,
    non_5k_contributors, unresolvable_5k_contributors
  )
nrow(pac_review_queue)
```


```{r}
# TEMPORARY PATCH: relabel seed-level major donor/IE committees that are currently labeled as no contributor data..
major_donor_seed_ids <- x %>%
  filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>%
  pull(FILER_ID) %>%
  unique()

pac_review_queue <- pac_review_queue %>%
  mutate(
    unresolvable_reason = if_else(
      pac_id %in% major_donor_seed_ids & unresolvable_reason == "no_contributor_data",
      "major_donor_ie_committee",
      unresolvable_reason
    )
  )


# TEMPORARY PATCH: reconcile is_federal_pac and unresolvable_reason for fed pacs

federal_pac_ids <- union(
  fed_seed_ids,
  over_5k_contributors %>%
    filter(contributor_type == "federal_pac") %>%
    pull(contributor_id) %>%
    na.omit() %>%
    unique()
)

pac_review_queue <- pac_review_queue %>%
  mutate(
    is_federal_pac = if_else(pac_id %in% federal_pac_ids, TRUE, is_federal_pac),
    unresolvable_reason = if_else(
      pac_id %in% federal_pac_ids & !is.na(unresolvable_reason),
      "federal_pac",
      unresolvable_reason
    )
  )

```



```{r}
# current
combined_review_queue <- bind_rows(
  pac_review_queue,
  contributor_review_queue %>%
    distinct(contributor_id, contributor_name, review_reason) %>%
    left_join(
      pac_resolvable_fractions %>% select(pac_id, status, total_donated_all),
      by = c("contributor_id" = "pac_id")
    ) %>%
    mutate(
      # NA (not FALSE) when contributor_id is NA -- unmatched, so genuinely unknown
      is_seed = if_else(is.na(contributor_id), NA, contributor_id %in% all_seed_status$pac_id)
    ) %>%
    transmute(
      pac_id              = contributor_id,
      Contributor.Name    = contributor_name,
      unresolvable_reason = review_reason,
      is_seed,
      status,
      total_donated_all
    )
)
combined_review_queue
```



```{r}
write.csv(combined_review_queue,"08_outputs/pac_review_queue_080326.csv")
```


# BUILD PAC CLASSIFICATION INPUT

```{r}
library(digest)
```

```{r}
hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16))
}

# get entity_group for each $5k+ individual/org contributor
over_5k_indiv <- over_5k_tagged %>%
  filter(contributor_type == "individual_or_org") %>%
  left_join(
    contributions_agg %>%
      select(Recipient.Committee.ID, standardized_name, entity_group, std_contributor_id),
    by = c("parent_pac_id" = "Recipient.Committee.ID",
           "contributor_name" = "standardized_name")
  )

over_5k_attrs <- prepped %>%
  filter(!is_pac, !is_fed_pac) %>%
  distinct(Recipient.Committee.ID, entity_group,
           processed_employer_name, processed_occupation,
           standardized_city, standardized_state, zip_code_processed)

# one row per alias
direct_for_classification <- over_5k_indiv %>%
  left_join(over_5k_attrs,
            by = c("parent_pac_id" = "Recipient.Committee.ID", "entity_group")) %>%
  mutate(entity_type = if_else(
    sapply(contributor_name, is_individual), "individual", "organization"
  ))

# format to match 0801 classification_input schema
pac_classification_input <- direct_for_classification %>%
  mutate(
    entity_id = if_else(
      !is.na(std_contributor_id),
      paste0("id_", std_contributor_id),
      sapply(contributor_name, function(n) hash_id(n, "indiv"))
    ),
    unit_key = paste(entity_id,
                     coalesce(processed_employer_name, ""),
                     coalesce(processed_occupation, ""),
                     sep = "||"),
    classification_unit_id = sapply(unit_key, function(k) hash_id(k, "unit"))
  ) %>%
  distinct(classification_unit_id, .keep_all = TRUE) %>%
  transmute(
    classification_unit_id,
    entity_id,
    entity_type,
    canonical_name = contributor_name,
    employer       = processed_employer_name,
    occupation     = processed_occupation,
    city           = standardized_city,
    state          = standardized_state,
    zip            = zip_code_processed,
    total_amount   = contributor_amount
  )

```


```{r}
#write_csv(pac_classification_input, "08_inputs/classification_intput_pacs.csv")
```


### combine with existing classification input

```{r}
# combine with 0801's classification_input (run 0801 first so the file exists)
existing <- read_csv("08_outputs/classification_input.csv", col_types = cols(.default = "c"))

combined_classification_input <- bind_rows(existing, pac_classification_input) %>%
  distinct(classification_unit_id, .keep_all = TRUE)

write_csv(combined_classification_input, "08_outputs/classification_input_combined.csv")
cat("Combined rows:", nrow(combined_classification_input),
    "  (PAC added:", nrow(pac_classification_input), ")\n")
```

### save files to use later

```{r}
write_csv(pac_classification_input,    "08_outputs/pac_classification_input.csv")
write_csv(over_5k_tagged,              "08_outputs/over_5k_tagged.csv")
write_csv(pac_classifiability,         "08_outputs/pac_classifiability.csv")

# save the traversal order — needed to replay topological sort for industry breakdown
write_csv(tibble(pac_id = ordered_pacs, order = seq_along(ordered_pacs)),
          "08_outputs/pac_traversal_order.csv")
write_csv(tibble(pac_id = stuck_pacs),  "08_outputs/pac_stuck_pacs.csv")

```








