---
title: "Process PAC Contributors"
output: html_notebook
---

# Summary 

**PAC classification process:**
PACs are classified by classifying each of their $5k+ contributors individually. 
If a PAC received contributions from entities that donated <$5k in aggregate, 
this PAC is sent to a manual review which determines how to classify its <$5k contributions. 
The <$5k contributors are all assigned a single category, while the 5k+ contributors each receive
their own category in the categorization pipeline. An industry breakdown is then computed for
each PAC. 

**Pulling PAC contributors:**

Input: 
(a) `08_inputs/power_search_contributions_normalized.csv` : full contributions data generated in 0801 script 
(b) `08_inputs/FILERNAME_CD.tsv` : data file from California Civic Data Coalition 
    (https://calaccess.californiacivicdata.org/documentation/raw-files/filername-cd/ ) that maps that PAC ids
    used when the PAC is a contributor (`XREF_FILER_ID`) to the id used when the PAC is a recipient
    (`FILER_ID`)
(c) `08_outputs/pac_input.csv` : PACs that gave $5k+ in aggregate to one of the tracker races (gov, insurance commissioner
    and props). These are the 'seed' PACs. 
(d) `../10_assign_classifications/10_inputs/pac_manual_review.csv` : feedback from the manual review of PACs. 

Output: 
  (a) a list of $5k+ orgs and individuals that will be classified in the classification pipeline
  (b) a list of PACs that need to be manually reviewed to determine how to classify
  
Process: 
Build a network of contributors to each PAC that goes 5 layers deep. Since PACs often give to other
PACs, we are not only pulling contributors to the 'seed' PACs, but also need to process PACs that
gave to our seed PACs by classifying their contributors. 

- layer 1: all $5k+ contributors to the seed PAC including orgs/indivduals and other PACs 
- layer 2: pulls all $5k+ contributors to the PACs that contributed to the 'seed' PACs.
- same process until you reach layer 5.

**PAC processing**
Candidates committees and federal PACs are not traversed in the graph. Federal PACs are manually
classified and candidate committees are classified as 'candidate / political party' - these PACs
are not classified based on their received contributions. PACs that are flagged as a federal PAC
or as a candidate committee are routed to the manual review. 

Summary statistics for each PAC are computed, including: 
- 'coverage' : how much the PAC received in contributions in 2025-26 compared to how much it gave
- amount received in unitemized contributions
- amount received in <$5k contributions
- amount received in >=$5k contributions
- Percent 'resolvable' : The % of a PACs received contributions that we can classify directly. 
  A received contribution is classified as 'resolvable' if: 
  (a) it came from an organization or individual we can classify directly
  (b) it came from a candidate committee we can classify directly
  (c) it came from another PAC that is fully resolvable (it can be classified in full by its
      $5k+ contributors)
      
A contribution is not counted as resolvable if: 
  - the contributing entity gave <$5k (these need to be classified manually; after manual
    review is completed for this PAC, the under $5k amount will then be counted as resolvable)
  - It came from a federal PAC
  - Only the 'resolvable' portion of a contributing PAC is flagged as 'resolvable' for the recipient PAC

**Manual review**
PACs are flagged for manual review if: 
- they have low coverage (received contributions < 75% of contributions it gave)
- the PAC is federal (meaning we can't reliably process its received contributions with CA data). 
  PACs are flagged as being a federal PAC if they have federal pac language in their name.
- the PAC has other signs of being a federal PAC
  - it appears to be based out of DC
  - all of its received contributions are from federal PACs (typically this is the federal
    branch of the PAC giving to its state counterpart)
- the PAC did not received 100% of its donations from $5k+ contributors that can be classified
  (meaning we need to determine how to classify its <$5k contributors).



# Read in packages and data

```{r}
library(dplyr)
library(readr)
library(stringr)
library(httr)
library(jsonlite)
library(igraph)

source("standardization_helpers.R")
```

```{r}
contributions_full <- read.csv("08_inputs/power_search_contributions_normalized.csv") # latest data pull from 0801 script
contributions_full <- contributions_full %>% rename(
  Ballot.Measure = Ballot.Measure.s.
)
nrow(contributions_full)
```

```{r}
# data to map contributor ids
id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv", col_types = cols(.default = "c"))
id_map_full <- id_map_full %>% filter(!is.na(FILER_ID) & !is.na(XREF_FILER_ID))

# get unique xref id / filer ids combos
id_map <- unique(id_map_full[,c("XREF_FILER_ID","FILER_ID","FILER_TYPE")])

# direct pac contributors (seed pacs)
new_pacs <- read.csv("08_outputs/new_prop_pacs.csv")
old_pacs <- read.csv("08_outputs/pac_input_09-07-26.csv")
x <- bind_rows(new_pacs, old_pacs)

# pac manual review feedback — matched first by pac_id, then by Contributor.Name for no-id rows
pac_manual_review <- read_csv(
  "../10_assign_classifications/10_inputs/pac_manual_review.csv",
  col_types = cols(.default = "c")
) %>%
  mutate(
    pac_id           = na_if(trimws(coalesce(pac_id, "")), ""), # pac_id may have been filled in by manual reviewer
    fec_id           = na_if(trimws(coalesce(fec_id, "")), ""),
    is_resolved      = !(
      coalesce(unresolvable_reason, "") == "additional_review_req" |
      is.na(industry_override_under_5k) |
      trimws(coalesce(industry_override_under_5k, "")) %in% c("", "PENDING")
    ),
    is_candidates_pp = trimws(coalesce(industry_override_under_5k, "")) == "Candidates/Political Parties",
    is_fec_federal   = !is.na(fec_id)
  )
```

```{r}
# load id corrections
library(googlesheets4)
gs4_deauth()

id_corrections <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1UnO3wldvEo84Md0ac5jbpg4SqWzvPle3yh15e_DGAVk",
  sheet = "corrected_pac_ids",
  col_types = "c"
) %>%
  mutate(
    Contributor.ID = coalesce(Contributor.ID, ""),
    resolved       = !is.na(resolved) & tolower(str_trim(resolved)) != "false",
    is_fed_pac     = tolower(str_trim(coalesce(is_fed_pac, ""))) %in% c("true", "1", "yes")
  )
```

## Clean up data 

### standardized col type

```{r}
# convert all ids to character strings
id_map$FILER_ID         <- as.character(id_map$FILER_ID)   
id_map$XREF_FILER_ID <- as.character(id_map$XREF_FILER_ID)
x$Contributor.ID <- as.character(x$Contributor.ID)
x$effective_Contributor.ID <- as.character(x$effective_Contributor.ID)
contributions_full$Recipient.Committee.ID <- as.character(contributions_full$Recipient.Committee.ID)
x$FILER_ID <- as.character(x$FILER_ID)
contributions_full$Contributor.ID <- as.character(contributions_full$Contributor.ID)
x$Recipient.Committee.ID <- as.character(x$Recipient.Committee.ID)
```

### apply fixes to pac ids (seed pacs)

Fix cases of missing or incorrect ids that have been reviewed manually. 

```{r}
# apply has_fed_pac_language so federal PACs are caught even if 0802 wasn't re-run
x <- x %>%
  mutate(is_fed_pac = is_fed_pac | has_fed_pac_language(Contributor.Name))

# only apply id corrections where the id was incorrect (don't propagated missing ids for now)
id_corrections_with_id <- id_corrections %>%
  filter(!is.na(Contributor.ID) & str_trim(Contributor.ID) != "")

x <- apply_contributor_corrections(x, id_corrections_with_id)

# re-derive FILER_ID and FILER_TYPE for any seeds whose effective_Contributor.ID changed
x <- x %>%
  left_join(
    id_map %>%
      distinct(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>%
      group_by(XREF_FILER_ID) %>% slice(1) %>% ungroup() %>%
      rename(derived_filer_id = FILER_ID, derived_filer_type = FILER_TYPE),
    by = c("effective_Contributor.ID" = "XREF_FILER_ID")
  ) %>%
  mutate(
    FILER_ID   = coalesce(na_if(as.character(FILER_ID), ""), derived_filer_id),
    FILER_TYPE = coalesce(na_if(FILER_TYPE, ""), derived_filer_type, "RECIPIENT COMMITTEE")
  ) %>%
  select(-derived_filer_id, -derived_filer_type)

# Apply is_fed_pac from all id_corrections entries, including name-only rows
fed_pac_name_keys <- toupper(str_squish(
  id_corrections %>% filter(resolved, is_fed_pac) %>% pull(Contributor.Name)
))
x <- x %>%
  mutate(is_fed_pac = coalesce(is_fed_pac, FALSE) |
           toupper(str_squish(Contributor.Name)) %in% fed_pac_name_keys)

```


### apply fixes to pac ids (all contributions)
  
```{r}
# apply corrections to contributions data
contributions <- contributions_full

#contributions <- apply_contributor_corrections(contributions, id_corrections) # fix incorrect ids + propagate missing ids
contributions <- apply_contributor_corrections(contributions, id_corrections_with_id) # only apply fixed to those with an incorrect id

contributions <- contributions %>%
  mutate(is_fed_pac = coalesce(is_fed_pac, FALSE) |
           toupper(str_squish(Contributor.Name)) %in% fed_pac_name_keys)

```


### Apply missing-ID corrections

Apply manually resolved corrections for PAC contributors with no FILER_ID.
The review queue (`missing_id_review_queue`) is generated later in this notebook with
one row per contribution. The reviewer fills in `corrected_id` (the correct XREF_FILER_ID
from CAL-ACCESS) after verifying each contribution, then saves the file as
`08_inputs/missing_id_review_queue_w_review.csv`.

IDs are joined on Contributor.Name, Recipient.Committee.ID, Amount, Start.Date — the same
fields used to identify a contribution in CAL-ACCESS Power Search.

```{r}
pac_no_id_corrections_path <- "08_inputs/missing_id_review_queue_w_review.csv"
if (file.exists(pac_no_id_corrections_path)) {
  pac_no_id_corrections <- read_csv(pac_no_id_corrections_path, col_types = cols(.default = "c")) %>%
    filter(!is.na(corrected_id) & str_trim(corrected_id) != "") %>%
    mutate(amount = as.numeric(amount)) %>%
    select(contributor_name, recipient_committee_id, amount, start_date,
           corrected_id, corrected_is_fed_pac = is_fed_pac)

  contributions <- contributions %>%
    left_join(
      pac_no_id_corrections,
      by = c(
        "Contributor.Name"       = "contributor_name",
        "Recipient.Committee.ID" = "recipient_committee_id",
        "Amount"                 = "amount",
        "Start.Date"             = "start_date"
      )
    ) %>%
    mutate(
      effective_Contributor.ID = coalesce(corrected_id, effective_Contributor.ID),
      is_fed_pac = coalesce(
        is_fed_pac,
        tolower(str_trim(coalesce(corrected_is_fed_pac, ""))) %in% c("true", "1", "yes")
      )
    ) %>%
    select(-corrected_id, -corrected_is_fed_pac)

  cat(sprintf("pac_no_id corrections applied: %d contributions with corrected IDs\n",
              nrow(pac_no_id_corrections)))
} else {
  cat("No pac_no_id corrections file found — skipping.\n")
}
```

```{r}
x$Contributor.ID <- as.character(x$Contributor.ID)
contributions$Contributor.ID <- as.character(contributions$Contributor.ID)

# join filer id to the contributions data
contributions <- contributions %>%
  left_join(
    id_map %>% select(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>%
      rename(std_contributor_id = FILER_ID),
    by = c("effective_Contributor.ID" = "XREF_FILER_ID")
  )
```



# Aggregate contributors 

For ALL PACs in the contributions data, we need to aggregate contributors to see how much each contributor gave.
The purpose of this is to determine which contributors meet the $5k threshold. All contributors that meet the 
$5k threshold (in aggregate) will be assigned an individual classification. 

There were several instances found where the Contributor.ID matched the Recipient.Committee.ID, implying that 
the PAC contributed to itself. Upoon review, most of these seem to be errors - the contributor and the recipient 
were typically the same entity, BUT the contributor didn't appear to be a PAC. When the Contributor.ID is found
to be the same as the Recipient.Committee.ID, it is flagged and the contributor id is REMOVED. An output file of 
all relevant cases of this is written later in the file.

**NOTE** the relevant `FILER_TYPEs` from id_map are `RECIPIENT COMMITTEE` and `MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE`
Despite the IE language in the `MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE` filer types, these all appear to be major 
donors and are processed as such. IE committees that received contributions have `FILER_TYPE` = `RECIPIENT COMMITTEE`


### 1. flag and map contributors to each PAC

```{r}
prepped <- contributions %>%
  mutate(
    has_blank_name   = is.na(Contributor.Name) | trimws(Contributor.Name) == "",
    has_id           = !is.na(effective_Contributor.ID) & effective_Contributor.ID != "",
    is_fed_pac       = has_fed_pac_language(Contributor.Name) | coalesce(is_fed_pac, FALSE), # apply function for identify federal pacs based on name
    is_candidate_pac = is_candidate_pac(Contributor.Name), # apply function for identifying candidate committees based on name 
    is_pac_no_id     = !has_id & !is_fed_pac & !is_candidate_pac & (
                       has_pac_language(Contributor.Name) |
                       grepl("\\bPAC\\b|COMMITTEE|POLITICAL ACTION|POLITICAL FUND|ACTION FUND", toupper(Contributor.Name))
                       )
  ) %>%
  mutate(
    is_self_contribution = !is.na(std_contributor_id) & std_contributor_id == Recipient.Committee.ID, # contributor.id and recipient.committee.id correspond to same entity
    unmatched_id        = has_id & is.na(std_contributor_id), # did not match a FILER_ID when the id_map was applied 
    std_contributor_id   = if_else(is_self_contribution, NA_character_, std_contributor_id), # remove contributor.ids when they are the same as the recipient.committee.id
    is_pac               = !is_self_contribution & has_id & !is_fed_pac &
                           coalesce(FILER_TYPE, "") != "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE" # don't count major donors as PACs (some major donors have contributor.ids)
  )

# build an entity graph for each PAC

# id and name nodes
name_nodes <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "") %>%
  distinct(Recipient.Committee.ID, standardized_name) %>% # distinct standardized_names within a recipient committee
  mutate(node = paste0("n_", row_number()))

id_nodes <- prepped %>%
  filter(!is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, std_contributor_id) %>%
  mutate(node = paste0("i_", row_number()))

# names that appear with >1 distinct (non-null) id within the same PAC are ambiguous
ambiguous_within_pac <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "",
         !is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, standardized_name, std_contributor_id) %>%
  group_by(Recipient.Committee.ID, standardized_name) %>% 
  filter(n() > 1) %>% # more than one contributor with this name
  ungroup() %>%
  distinct(Recipient.Committee.ID, standardized_name)

# edges only for unambiguous 
edges <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "",
         !is.na(std_contributor_id) & std_contributor_id != "") %>%
  distinct(Recipient.Committee.ID, standardized_name, std_contributor_id) %>%
  anti_join(ambiguous_within_pac, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(name_nodes, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(id_nodes,   by = c("Recipient.Committee.ID", "std_contributor_id")) %>%
  select(from = node.x, to = node.y)

all_nodes <- c(name_nodes$node, id_nodes$node)
g  <- graph_from_data_frame(edges, directed = FALSE, vertices = all_nodes)
comps     <- components(g)

component_df <- tibble(node = names(comps$membership), entity_group = comps$membership)

name_to_group <- name_nodes %>%
  left_join(component_df, by = "node") %>%
  select(Recipient.Committee.ID, standardized_name, name_entity_group = entity_group)

id_to_group <- id_nodes %>%
  left_join(component_df, by = "node") %>%
  select(Recipient.Committee.ID, std_contributor_id, id_entity_group = entity_group)

# assign entity group 
prepped <- prepped %>%
  left_join(name_to_group, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  left_join(id_to_group,   by = c("Recipient.Committee.ID", "std_contributor_id")) %>%
  mutate(
    entity_group = case_when(
      !is.na(std_contributor_id) & std_contributor_id != "" ~ id_entity_group,
      !is.na(name_entity_group)                             ~ name_entity_group,
      TRUE ~ as.numeric(max(comps$membership) + row_number()),
    )
  ) %>%
  select(-name_entity_group, -id_entity_group)

# propagate federal PAC flag by id: if ANY name variant for a given pac id
# is flagged as a federal PAC, flag all contributions with that ID as federal pacs
fed_pac_std_ids <- prepped %>%
  filter(!is.na(std_contributor_id), std_contributor_id != "", is_fed_pac) %>%
  distinct(std_contributor_id) %>%
  pull(std_contributor_id) # get all filer ids associated w/ a federal pac

fed_pac_raw_ids <- prepped %>%
  filter(!is.na(effective_Contributor.ID), effective_Contributor.ID != "", is_fed_pac) %>%
  distinct(effective_Contributor.ID) %>%
  pull(effective_Contributor.ID)

prepped <- prepped %>%
  mutate(
    is_fed_pac = is_fed_pac |
      (!is.na(std_contributor_id) & std_contributor_id %in% fed_pac_std_ids) |
      (!is.na(effective_Contributor.ID) & effective_Contributor.ID %in% fed_pac_raw_ids)
  )

# flag suspected federal PACs if contributor shares a Normalized name with a confirmed federal PAC
# `Normalized` name groups can contain multiple distinct PACs, so a Normalized name
# match does not confirm federal status. 

fed_pac_normalized_names <- prepped %>%
  filter(!is.na(Normalized) & Normalized != "", is_fed_pac) %>%
  distinct(Normalized) %>%
  pull(Normalized)

prepped <- prepped %>%
  mutate(
    entity_group_has_fed_pac = !is.na(Normalized) & Normalized != "" &
                               Normalized %in% fed_pac_normalized_names
  )
```


### 2. apply aggregation

PAC contribution amounts are aggregate by the contributors' `Normalized` name (`normalized_amount`) to determine if
they meet the $5k threshold. 

Sub-groups (`entity_group`) keep their own attributes (is_pac, is_fed_pac, std_contributor_id, etc.) because the
`Normalized` groups are much broader and may contain multiple different PACs or contain both PACs and non-PACs in the 
same entity. There are some flags that apply at the pac-level that may not apply at the normalized name level.

```{r}
# aggregate by entity_group — required base for contributions_agg below,
# is_pac/is_fed_pac/std_contributor_id/etc.
# should stay at this narrower aggregations since Normalized groups can span multiple PACs/types of PACs)
contributions_agg_by_entity <- prepped %>%
  group_by(Recipient.Committee.ID, entity_group) %>%
  summarise(
    Amount  = sum(Amount, na.rm = TRUE), # amount per entity-group 
    standardized_name  = { # choose the most common name per group 
      ns <- standardized_name[!is.na(standardized_name) & standardized_name != ""]
      if (length(ns) > 0) names(sort(table(ns), decreasing = TRUE))[1] else NA_character_
    },
    raw_name = { # most common raw name per group 
      rns <- Contributor.Name[!is.na(Contributor.Name) & Contributor.Name != ""]
      if (length(rns) > 0) names(sort(table(rns), decreasing = TRUE))[1] else NA_character_
    },
    std_contributor_id = { # fallback, but should never be the case where more than one non-null id is associated with an entity group. 
      ids <- std_contributor_id[!is.na(std_contributor_id) & std_contributor_id != ""]
      if (length(ids) > 0) ids[1] else NA_character_
    },
    FILER_TYPE            = {
      types <- FILER_TYPE[!is.na(FILER_TYPE) & FILER_TYPE != ""]
      if (length(types) > 0) names(sort(table(types), decreasing = TRUE))[1] else NA_character_
    },
    
    # apply flags at the entity-group level
    is_fed_pac            = any(is_fed_pac, na.rm = TRUE),
    is_candidate_pac      = any(is_candidate_pac, na.rm = TRUE),
    is_pac_no_id          = any(is_pac_no_id, na.rm = TRUE),
    is_pac                = any(is_pac, na.rm = TRUE),
    unmatched_contributor_id  = any(unmatched_id, na.rm = TRUE),
    raw_unmatched_id = {
      # preserve the ID for unmatched cases (has_id but not in id_map)
      ids <- effective_Contributor.ID[coalesce(unmatched_id, FALSE) &
                                      !is.na(effective_Contributor.ID) &
                                      effective_Contributor.ID != ""]
      if (length(ids) > 0) ids[1] else NA_character_
    },
    has_self_contribution = any(is_self_contribution, na.rm = TRUE),
    entity_group_has_fed_pac = any(entity_group_has_fed_pac, na.rm = TRUE),
    .groups               = "drop"
  ) %>%
  filter(Amount >= 0) %>% 
  mutate(is_pac_no_id = is_pac_no_id & is.na(std_contributor_id))
```

```{r}
blank_name_count <- sum(coalesce(prepped$has_blank_name, FALSE))
if (blank_name_count > 0) {
  cat(sprintf("Flagging %d contribution(s) with blank Contributor.Name — excluded from name grouping, amounts still counted.\n", blank_name_count))
}
```

Link entity_groups via a Normalized-name graph:
entity_groups are nodes, bridged by (a) any shared Normalized text and (b) `ambiguous_within_pac` pairs,
reused directly as edges, since those are likely the same entity with more than one
committee id. `ambiguous_within_pac` matches on
`standardized_name` (more processed) while default `Normalized` text is only `toupper()`, so these
two signals not always redundant — an unreviewed ambiguous pair can share
`standardized_name` without sharing `Normalized` text.

```{r}
# entity_group and Normalized edges
norm_text_pairs <- prepped %>%
  filter(!coalesce(has_blank_name, FALSE)) %>%
  mutate(Normalized = str_squish(coalesce(Normalized, ""))) %>%
  filter(Normalized != "") %>%
  distinct(Recipient.Committee.ID, entity_group, Normalized) # distinct Normalized names per entity group

eg_nodes <- prepped %>%
  distinct(Recipient.Committee.ID, entity_group) %>%
  mutate(node = paste0("eg_", row_number()))

normtext_nodes <- norm_text_pairs %>%
  distinct(Recipient.Committee.ID, Normalized) %>%
  mutate(node = paste0("nt_", row_number()))

norm_edges <- norm_text_pairs %>%
  left_join(eg_nodes,       by = c("Recipient.Committee.ID", "entity_group")) %>%
  left_join(normtext_nodes, by = c("Recipient.Committee.ID", "Normalized")) %>%
  select(from = node.x, to = node.y)

# entity_group and ambiguous-name edges (reusing ambiguous_within_pac directly
# same standardized_name, different ids -> like same entity with multiple committees)
ambig_entity_groups <- prepped %>%
  filter(!is.na(standardized_name) & standardized_name != "",
         !is.na(std_contributor_id) & std_contributor_id != "") %>%
  semi_join(ambiguous_within_pac, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  distinct(Recipient.Committee.ID, standardized_name, entity_group)

ambig_nodes <- ambig_entity_groups %>%
  distinct(Recipient.Committee.ID, standardized_name) %>%
  mutate(node = paste0("am_", row_number()))

ambig_edges <- ambig_entity_groups %>%
  left_join(eg_nodes,    by = c("Recipient.Committee.ID", "entity_group")) %>%
  left_join(ambig_nodes, by = c("Recipient.Committee.ID", "standardized_name")) %>%
  select(from = node.x, to = node.y)

norm_all_nodes <- c(eg_nodes$node, normtext_nodes$node, ambig_nodes$node)
norm_edges_all <- bind_rows(norm_edges, ambig_edges) 

g_norm  <- graph_from_data_frame(norm_edges_all, directed = FALSE, vertices = norm_all_nodes)
comps_norm <- components(g_norm)

entity_group_component_map <- eg_nodes %>%
  left_join(
    tibble(node = names(comps_norm$membership), normalized_component = comps_norm$membership),
    by = "node"
  ) %>%
  select(Recipient.Committee.ID, entity_group, normalized_component)

# display label only (most frequent raw Normalized text per entity_group) — not used as the merge key
entity_group_norm_label <- prepped %>%
  filter(!coalesce(has_blank_name, FALSE)) %>%
  mutate(Normalized = str_squish(coalesce(Normalized, ""))) %>%
  group_by(Recipient.Committee.ID, entity_group, Normalized) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Recipient.Committee.ID, entity_group) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(Recipient.Committee.ID, entity_group, Normalized)

# normalized_amount: sum of Amount across every entity_group in the same normalized_component
normalized_amount_by_component <- contributions_agg_by_entity %>%
  left_join(entity_group_component_map, by = c("Recipient.Committee.ID", "entity_group")) %>%
  group_by(Recipient.Committee.ID, normalized_component) %>%
  mutate(normalized_amount = sum(Amount, na.rm = TRUE)) %>%
  ungroup() %>%
  select(Recipient.Committee.ID, entity_group, normalized_component, normalized_amount)

contributions_agg <- contributions_agg_by_entity %>%
  left_join(entity_group_norm_label,        by = c("Recipient.Committee.ID", "entity_group")) %>%
  left_join(normalized_amount_by_component, by = c("Recipient.Committee.ID", "entity_group"))

# within each PAC x normalized_component, if ANY entity_group is confirmed federal,
# flag the others as suspected_fed_pac
norm_fed_pac <- contributions_agg %>%
  group_by(Recipient.Committee.ID, normalized_component) %>%
  summarise(norm_has_fed_pac = any(coalesce(is_fed_pac, FALSE)), .groups = "drop")

contributions_agg <- contributions_agg %>%
  left_join(norm_fed_pac, by = c("Recipient.Committee.ID", "normalized_component")) %>%
  mutate(suspected_fed_pac = !coalesce(is_fed_pac, FALSE) & coalesce(norm_has_fed_pac, FALSE)) %>%
  select(-norm_has_fed_pac)
```


### diagnostic (can skip this)

Check entity_groups that were previously separate (distinct entity_group ids from the
id/name graph above) but got pulled into the same normalized_component — i.e. actual merges
produced by the Normalized-name bridge or an ambiguous_within_pac edge, not just entity_groups
that happen to carry more than one Normalized variant on their own.

```{r}
entity_group_norm_variants <- norm_text_pairs %>%
  group_by(Recipient.Committee.ID, entity_group) %>%
  summarise(normalized_variants = paste(sort(unique(Normalized)), collapse = " | "), .groups = "drop")

# raw Contributor.Name(s) actually behind each entity_group — needed to tell apart
# "different raw names that were legitimately mapped to the same Normalized label"
# (expected) from "these are the same raw name and should already be one entity_group"
# (a red flag in the id/name graph, not in the Normalized bridge)
entity_group_raw_names <- prepped %>%
  filter(!coalesce(has_blank_name, FALSE)) %>%
  distinct(Recipient.Committee.ID, entity_group, Contributor.Name) %>%
  group_by(Recipient.Committee.ID, entity_group) %>%
  summarise(raw_names = paste(sort(unique(Contributor.Name)), collapse = " | "), .groups = "drop")

merged_by_normalized <- contributions_agg %>%
  left_join(entity_group_norm_variants, by = c("Recipient.Committee.ID", "entity_group")) %>%
  left_join(entity_group_raw_names,     by = c("Recipient.Committee.ID", "entity_group")) %>%
  group_by(Recipient.Committee.ID, normalized_component) %>%
  filter(n_distinct(entity_group) > 1) %>%
  ungroup() %>%
  # merge_case classifies entity_groups that share the SAME raw_names within a merged
  # normalized_component (i.e. would have been suspicious to see un-merged):
  #  - "same_raw_name_diff_id": they carry 2+ distinct non-null std_contributor_ids ->
  #     the expected ambiguous_within_pac signature (same entity, different committee id)
  #  - "same_raw_name_same_id_or_no_id": they share one id (or none at all) -> should
  #     already have been one entity_group before this graph ran; investigate the id/name
  #     graph itself, not the Normalized bridge
  # NA means this raw_names string is unique within the component (nothing to compare it to)
  group_by(Recipient.Committee.ID, normalized_component, raw_names) %>%
  mutate(
    n_ids_sharing_raw_name = n_distinct(std_contributor_id[!is.na(std_contributor_id) & std_contributor_id != ""]),
    merge_case = case_when(
      n() == 1                  ~ NA_character_,
      n_ids_sharing_raw_name > 1 ~ "same_raw_name_diff_id",
      TRUE                       ~ "same_raw_name_same_id_or_no_id"
    )
  ) %>%
  ungroup() %>%
  select(Recipient.Committee.ID, normalized_component, entity_group,
         raw_names, normalized_variants, std_contributor_id, merge_case,
         Amount, normalized_amount) %>%
  arrange(Recipient.Committee.ID, normalized_component, desc(Amount))

cat(sprintf(
  "%d entity_group(s) across %d normalized_component(s) merged with at least one other entity_group\n",
  nrow(merged_by_normalized),
  n_distinct(paste(merged_by_normalized$Recipient.Committee.ID, merged_by_normalized$normalized_component))
))

merged_by_normalized
```

```{r}
# expected pattern: same raw name, different committee ids (ambiguous_within_pac doing its job)
merged_by_normalized %>% filter(merge_case == "same_raw_name_diff_id")
merged_by_normalized %>% filter(merge_case == "same_raw_name_same_id_or_no_id") # should be 0 
```

# CREATE PAC GRAPH

### helper functions

```{r}
# pull all contributors to a PAC
get_pac_contributors <- function(pac_id, contributions_agg) {
  contributions_agg %>%
    filter(Recipient.Committee.ID == pac_id)
}

# determine which tier contributor falls into
categorize_contributor <- function(name,amount) {
  name_upper <- toupper(name)
  case_when(
    name_upper == "UNITEMIZED CONTRIBUTIONS" ~ "unitemized",
    amount <= 100 ~ "small_dollar",
    amount < 5000   ~ "under_5k",
    TRUE  ~ "over_5k" 
  )
}

# calculate how much each PAC donated (to compare to how much it received)
pac_total_donated <- contributions_agg %>%
  group_by(std_contributor_id,unmatched_contributor_id) %>% #Contributor.ID, unmatched_contributor_id
  summarise(total_donated_all = sum(Amount, na.rm = TRUE), .groups = "drop")

```


### define the pac graph 

```{r}
# constants
DATA_SUFFICIENCY_THRESHOLD <- 0.75 # data sufficieny compares how much pac received to how much it gave
```


```{r}
# function to grab all contributors in the PAC graph
traverse_pac_graph <- function(seed_ids, ref, id_map, max_depth = 5, federal_ids = character(0), candidate_ids = character(0)) {
  queue          <- lapply(seed_ids, function(id) list(id = id, depth = 0, path = id))
  visited        <- character(0)
  max_depth_seen <- character(0)
  pac_summaries  <- list()
  over_5k_contribs <- list()

  while (length(queue) > 0) {
    node   <- queue[[1]]; queue <- queue[-1]
    pac_id <- node$id; depth <- node$depth; path <- node$path

    if (pac_id %in% visited || depth >= max_depth) {
      # Only add a max_depth_reached row the first time this PAC hits the depth ceiling.
      if (depth >= max_depth && !(pac_id %in% visited) && !(pac_id %in% max_depth_seen)) {
        max_depth_seen <- c(max_depth_seen, pac_id)
        pac_summaries <- append(pac_summaries, list(tibble(
          pac_id = pac_id, depth = depth, path = path, status = "max_depth_reached",
          total_received = NA_real_, unitemized_amount = NA_real_,
        )))
      }
      next
    }
    visited <- c(visited, pac_id)

    # federal PACs are leaves — don't fetch their contributors
    if (pac_id %in% federal_ids) {
      pac_summaries <- append(pac_summaries, list(tibble(
        pac_id            = pac_id, depth = depth, path = path,
        status            = "federal_pac",
        total_received    = NA_real_, total_donated_all = NA_real_,
        data_sufficiency  = NA_real_,
        unitemized_amount = NA_real_, small_dollar_amount = NA_real_,
        under_5k_amount   = NA_real_, over_5k_amount = NA_real_
      )))
      next
    }

    # candidate PACs are classified directly — don't fetch their contributors
    if (pac_id %in% candidate_ids) {
      pac_summaries <- append(pac_summaries, list(tibble(
        pac_id            = pac_id, depth = depth, path = path,
        status            = "candidate_pac",
        total_received    = NA_real_, total_donated_all = NA_real_,
        data_sufficiency  = NA_real_,
        unitemized_amount = NA_real_, small_dollar_amount = NA_real_,
        under_5k_amount   = NA_real_, over_5k_amount = NA_real_
      )))
      next
    }

    contribs <- get_pac_contributors(pac_id, ref) # grab all contributors to this pac

    total_donated_all <- pac_total_donated %>% # pre-computed df w/ total contribution amnts for all PACs
      filter(!unmatched_contributor_id, std_contributor_id == pac_id) %>%
      pull(total_donated_all)
    total_donated_all <- if (length(total_donated_all) == 0) NA_real_ else total_donated_all
    
    if (is.null(contribs) || nrow(contribs) == 0) { # no contributors found
      pac_summaries <- append(pac_summaries, list(tibble(
        pac_id = pac_id, depth = depth, path = path, status = "unresolved",
        total_received = NA_real_, total_donated_all = total_donated_all,
        data_sufficiency = NA_real_,
        unitemized_amount = NA_real_, 
        under_5k_amount = NA_real_, over_5k_amount = NA_real_
      )))
      next
    }
    
    # check data sufficiency
    total_received    <- sum(contribs$Amount, na.rm = TRUE)
    data_sufficiency  <- if (!is.na(total_donated_all) && total_donated_all > 0)
                           total_received / total_donated_all
                         else NA_real_

    pac_status <- case_when(
      !is.na(data_sufficiency) & data_sufficiency < DATA_SUFFICIENCY_THRESHOLD ~ "low_data_sufficiency",
      is.na(data_sufficiency)                                                   ~ "unknown_data_sufficiency",
      TRUE                                                                      ~ "resolved" # change this from being resolved...
    )
    
    contribs <- contribs %>%
      mutate(contributor_category = categorize_contributor(standardized_name, normalized_amount)) # normalized_amount is the aggregated amount given by this entity
    
    # summarize PAC based on what kind of donations it received
    pac_summaries <- append(pac_summaries, list(tibble(
      pac_id             = pac_id,
      depth              = depth,
      path               = path,
      status             = "resolved", # change this language as well 
      total_received     = sum(contribs$Amount, na.rm = TRUE), 
      total_donated_all = total_donated_all,
      data_sufficiency  = total_received / total_donated_all,   # NA if total_donated_all == 0
      unitemized_amount  = sum(contribs$Amount[contribs$contributor_category == "unitemized"],  na.rm = TRUE),
      small_dollar_amount = sum(contribs$Amount[contribs$contributor_category == "small_dollar"],  na.rm = TRUE),
      under_5k_amount    = sum(contribs$Amount[contribs$contributor_category == "under_5k"],    na.rm = TRUE),
      over_5k_amount     = sum(contribs$Amount[contribs$contributor_category == "over_5k"],     na.rm = TRUE)
    )))

    # only process contributors that a part of an entity that contributed $5k+ (only these entities will be classified)
    for (i in seq_len(nrow(contribs %>% filter(contributor_category == "over_5k")))) {
      row <- (contribs %>% filter(contributor_category == "over_5k"))[i, ]

      # federal pacs and candidate committees
      # check both per-row flags AND the ID sets: the sets
      # include sources (prepped, pac_manual_review) that may not set the per-row flag.
      is_federal_contrib   <- isTRUE(row$is_fed_pac)   || (!is.na(row$std_contributor_id) && row$std_contributor_id %in% federal_ids)
      is_candidate_contrib <- isTRUE(row$is_candidate_pac) || (!is.na(row$std_contributor_id) && row$std_contributor_id %in% candidate_ids)
      if (is_federal_contrib || is_candidate_contrib) {
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id      = pac_id,
          path               = path,
          depth              = depth + 1,
          entity_group       = row$entity_group,
          contributor_name   = row$raw_name,
          contributor_id     = row$std_contributor_id,
          contributor_type   = if (is_federal_contrib) "federal_pac" else "candidate_pac",
          contributor_amount = row$Amount
        )))

      } 
      # pacs
      else if(isTRUE(!is.na(row$std_contributor_id))|| isTRUE(row$is_pac) || isTRUE(row$is_pac_no_id)) { 
        new_id    <- row$std_contributor_id     # used for traversal queuing only
        matched   <- !is.na(row$std_contributor_id) # whether contributor.id matched to a FILER_ID through id_map
        display_id <- coalesce(row$std_contributor_id, row$raw_unmatched_id) # for unmatched cases, fall back to the raw Contributor.ID for reference in review queue
        new_path  <- paste0(path, " -> ", display_id)
        
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id      = pac_id,
          path               = new_path,
          depth              = depth + 1,
          entity_group       = row$entity_group,
          contributor_name   = row$raw_name,
          contributor_id     = display_id,
          contributor_type   = if (isTRUE(row$is_pac_no_id)) "pac_no_id"
                               else if (!matched) "pac_unmatched_id"
                               else if (isTRUE(row$FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE")) "major_donor_ie_committee"
                               else "pac",
          contributor_amount = row$Amount
        )))
        if (!is.na(new_id)) {  # only queue matched std_contributor_ids 
          queue <- append(queue, list(list(id = new_id, depth = depth + 1, path = new_path)))
        }

      } 
      # individuals/orgs
      else {
        # individual or org — record for industry classification
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id          = pac_id,
          path                   = path,
          depth                  = depth + 1,
          entity_group           = row$entity_group,
          contributor_name       = row$raw_name,
          contributor_id         = NA_character_, # anything with a major donor id is processed in chunk above
          contributor_type       = "individual_or_org",
          contributor_amount     = row$Amount,
        )))
      }
    }
  }

  list(
    pac_summaries    = bind_rows(pac_summaries),
    over_5k_contributors = bind_rows(over_5k_contribs)
  )
}

```


### run pac graph 

```{r}
seed_ids <- unique(x$FILER_ID) # get all seeds pacs
seed_ids <- seed_ids[!is.na(seed_ids) & seed_ids != ""]

# federal seed pacs — require a non-NA FILER_ID so NAs don't pollute downstream ID sets
fed_seed_ids <- x %>%
  filter(coalesce(is_fed_pac, FALSE), !is.na(FILER_ID) & FILER_ID != "") %>%
  pull(FILER_ID) %>%
  unique()

# flag federal IDs for traversal (don't want to grab contributors to federal pacs)
# sources : is_fed_pac from contributions_agg + seed-level fed PACs + manual review overrides
federal_pac_ids <- union(
  contributions_agg %>%
    filter(coalesce(is_fed_pac, FALSE), !is.na(std_contributor_id)) %>%
    pull(std_contributor_id) %>%
    unique(),
  union(
    fed_seed_ids,
    pac_manual_review %>%
      filter(!is.na(pac_id), is_fec_federal) %>%
      pull(pac_id) %>%
      as.character()
  )
)

candidate_pac_recip_ids <- prepped %>%
  distinct(Recipient.Committee.ID, Recipient.Committee) %>%
  filter(is_candidate_pac(Recipient.Committee)) %>%
  pull(Recipient.Committee.ID) %>%
  unique()

# flag candidate pacs so the pac graph doesn't grab their contributors
# sources: is_candidate_pac from prepped (source of truth) + manual review overrides (is_candidates_pp)
contrib_candidate_ids <- union(
  prepped %>%
    filter(coalesce(is_candidate_pac, FALSE), !is.na(std_contributor_id) & std_contributor_id != "") %>%
    pull(std_contributor_id) %>%
    unique(),
  pac_manual_review %>%
    filter(!is.na(pac_id), is_candidates_pp) %>%
    pull(pac_id) %>%
    as.character()
)

candidate_pac_ids <- unique(c(contrib_candidate_ids,candidate_pac_recip_ids))

# traverse pac graph
result <- traverse_pac_graph(seed_ids, contributions_agg, id_map, max_depth = 5,
                                       federal_ids   = federal_pac_ids,
                                       candidate_ids = candidate_pac_ids)
pac_summaries <- result$pac_summaries
over_5k_contributors <- result$over_5k_contributors
```

```{r}
# flag contributor names that map to >1 traversed PAC ID
# — same name, different IDs across different recipient PACs' contributions (cross-recipient)
# — scoped to: recipient was traversed AND contributor ID was itself traversed
# erroneous IDs should be added to corrected_pac_ids and pipeline re-run
name_id_conflicts <- contributions %>%
  filter(
    Recipient.Committee.ID %in% pac_summaries$pac_id,
    std_contributor_id %in% pac_summaries$pac_id,
    !is.na(standardized_name) & standardized_name != ""
  ) %>%
  distinct(standardized_name, std_contributor_id) %>%
  group_by(standardized_name) %>%
  filter(n_distinct(std_contributor_id) > 1) %>%
  ungroup() %>%
  distinct(standardized_name)

if (nrow(name_id_conflicts) > 0) {
  cat(sprintf(
    "WARNING: %d contributor name(s) map to multiple traversed PAC IDs — add corrections to corrected_pac_ids before re-running.\n",
    nrow(name_id_conflicts)
  ))
  print(name_id_conflicts)
}

#over_5k_tagged <- over_5k_tagged %>%
#  mutate(has_conflicting_id = contributor_name %in% name_id_conflicts$standardized_name)
#name_id_conflicts

```

```{r}
# backfill pac_summaries for federal/candidate PAC contributors blocked from the
# traversal queue. otherwise the pac_classifiability join in over_5k_tagged
# returns NA for status/data_ok/low_data_sufficiency on those rows
untraversed_pacs <- over_5k_contributors %>%
  filter(
    contributor_type %in% c("federal_pac", "candidate_pac"),
    !is.na(contributor_id),
    !contributor_id %in% pac_summaries$pac_id
  ) %>%
  group_by(contributor_id) %>%
  slice_min(depth, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    pac_id              = contributor_id,
    depth               = depth,
    path                = NA_character_,
    status              = contributor_type,
    total_received      = NA_real_,
    total_donated_all   = NA_real_,
    data_sufficiency    = NA_real_,
    unitemized_amount   = NA_real_,
    small_dollar_amount = NA_real_,
    under_5k_amount     = NA_real_,
    over_5k_amount      = NA_real_
  )

if (nrow(untraversed_pacs) > 0)
  cat(sprintf("Backfilling %d federal/candidate PAC(s) into pac_summaries (blocked from traversal queue)\n",
              nrow(untraversed_pacs)))

pac_summaries <- bind_rows(pac_summaries, untraversed_pacs)

major_donor_ie_ids <- union(
  x %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% pull(FILER_ID) %>% unique(),
  over_5k_contributors %>% filter(contributor_type == "major_donor_ie_committee") %>% pull(contributor_id) %>% na.omit() %>% unique()
)

pac_summaries <- pac_summaries %>%
  mutate(
    is_federal_pac    = pac_id %in% federal_pac_ids,
    is_major_donor_ie = pac_id %in% major_donor_ie_ids,
    is_candidate_pac  = pac_id %in% candidate_pac_ids,
    status = if_else(is_federal_pac & status == "resolved", "federal_pac", status)
  )
```


```{r}
# data check
pac_summaries %>% group_by(pac_id) %>% filter(n() > 1) # should be 0 rows
pac_summaries %>% filter(status == "max_depth_reached") # ideally is 0 rows, but may not be
```


```{r}
type_priority <- c(federal_pac = 1, major_donor_ie_committee = 2, candidate_pac = 3, pac_unmatched_id = 4, pac_no_id = 5)

# only used to match back to join to pac_review_queue on pac_id
contributor_type_lookup <- over_5k_contributors %>%
  filter(!is.na(contributor_id), contributor_type %in% names(type_priority)) %>%
  distinct(contributor_id, contributor_type) %>%
  mutate(priority = type_priority[contributor_type]) %>%
  group_by(contributor_id) %>%
  slice_min(priority, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(contributor_id, contributor_type)

```

```{r, eval = FALSE}
#write.csv(pac_summaries, "08_outputs/pac_summaries.csv",row.names = FALSE)
#write.csv(over_5k_contributors, "08_outputs/over_5k_contribs.csv",row.names = FALSE)
```


### Add flags for federal PAC signals

Add flags for PACs that COULD be a federal PAC based on these signals. 

```{r}
# flag when every contributor to a PAC is a federal PAC
all_fed_contributor_pacs <- contributions_agg %>%
  group_by(Recipient.Committee.ID) %>%
  summarise(
    n_total = n(),
    n_fed   = sum(coalesce(is_fed_pac, FALSE)),
    .groups = "drop"
  ) %>%
  filter(n_total > 0, n_total == n_fed) %>%
  transmute(pac_id = Recipient.Committee.ID, suspected_federal_reason = "all_contributors_federal")

# flag when every received contribution came from the same entity 
# either the contributor's FILER_ID = Recipient.Committee.ID (self-contribution),
# or the contributor name matches the recipient committee name
rc_name_std <- prepped %>%
  distinct(Recipient.Committee.ID, Recipient.Committee) %>%
  mutate(recipient_name_std = toupper(str_squish(Recipient.Committee))) %>%
  group_by(Recipient.Committee.ID) %>%
  slice(1) %>%
  ungroup() %>%
  select(Recipient.Committee.ID, recipient_name_std)

self_contributing_pacs <- prepped %>%
  filter(Amount > 0) %>%
  left_join(rc_name_std, by = "Recipient.Committee.ID") %>%
  mutate(
    contributor_name_std = toupper(str_squish(Contributor.Name)),
    is_self_any = is_self_contribution | (contributor_name_std == recipient_name_std)
  ) %>%
  group_by(Recipient.Committee.ID) %>%
  summarise(
    n_total = n(),
    n_self  = sum(is_self_any, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(n_total > 0, n_total == n_self) %>%
  transmute(pac_id = Recipient.Committee.ID, suspected_federal_reason = "self_contributing")

suspected_federal_pacs <- bind_rows(all_fed_contributor_pacs, self_contributing_pacs) %>%
  group_by(pac_id) %>%
  summarise(
    suspected_federal_reason = paste(sort(unique(suspected_federal_reason)), collapse = "; "),
    .groups = "drop"
  )

pac_summaries <- pac_summaries %>%
  select(-any_of(c("suspected_federal_pac", "suspected_federal_reason"))) %>%
  left_join(suspected_federal_pacs, by = "pac_id") %>%
  mutate(suspected_federal_pac = !is_federal_pac & !is.na(suspected_federal_reason))
```


```{r, eval = FALSE}
# look as any suspected federal PACs
pac_summaries %>% filter(suspected_federal_pac == TRUE)
```


### Add flags to PACs and compute statistics

Add flags to PACs with a data sufficieny < 0.75 and for cases where contributors to PACs couldn't be found
for this cycle. 

Compute % of PACs received contributions came from unitemized contributions
and contributions from entities that contributed $5k+ in aggregate.

`over_5k_amount` is the sum of contributions from entities that gave over $5k
in aggregate, not the sum of individual contributions over $5k.

```{r}
# constants
OVER_5K_COVERAGE_THRESHOLD <- 0.5
```

```{r}
pac_classifiability <- pac_summaries %>%
  mutate(
    low_data_sufficiency     = !is.na(data_sufficiency) & data_sufficiency < DATA_SUFFICIENCY_THRESHOLD, # threshold is 0.75
    unknown_data_sufficiency = is.na(data_sufficiency), # contributors to PAC not found
    pct_unitemized = unitemized_amount / total_received,
    pct_small_dollar = small_dollar_amount / total_received,
    pct_over_5k  = over_5k_amount / total_received,
    data_ok = status == "resolved" & !is_federal_pac & !low_data_sufficiency & !unknown_data_sufficiency
  )

```


```{r, eval = FALSE}
# data check
min(pac_classifiability$pct_over_5k,na.rm = TRUE)
max(pac_classifiability$pct_over_5k,na.rm = TRUE)
```


### Flag each $5k contributor as resolvable or not

Determine how to handle each contributor that we are going to classify
because it was part of an entity that gave $5k+ in aggregate.

Contributors are marked with `is_resolvable` = TRUE when they can be classified
directly in the pipeline, which includes
  (a) individuals/orgs
  (b) major donors that do NOT contain any PAC language in their name
      (anything labeled a major donor shouldn't be a PAC, but want to check
       anything with pac language just in case)
  (c) candidate PACs

`over_5k_contributors` contains ALL contributors that gave at least $5k in 
aggregate, includes PACs, individuals, orgs and major donors. 


```{r}
# look at the status of each contributor
over_5k_tagged <- over_5k_contributors %>%
  left_join( # join pac statistics 
    pac_classifiability %>% select(pac_id,data_ok, pct_over_5k, status,low_data_sufficiency), 
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    is_resolvable = case_when(
      contributor_type == "individual_or_org" ~ TRUE,
      contributor_type == "candidate_pac"     ~ TRUE,
      contributor_type == "major_donor_ie_committee" & !grepl(
        "\\bPAC\\b|POLITICAL ACTION|INDEPENDENT EXPENDITURE|\\bCOMMITTEE\\b",
        toupper(contributor_name)
      ) ~ TRUE,
      contributor_type == "pac" & data_ok & pct_over_5k >= OVER_5K_COVERAGE_THRESHOLD ~ TRUE,
      TRUE ~ FALSE
    ),
    unresolvable_reason = case_when(
      !is_resolvable & contributor_type == "federal_pac"                             ~ "federal_pac", 
      !is_resolvable & contributor_type %in% c("pac_no_id", "pac_unmatched_id")     ~ contributor_type,
      !is_resolvable & contributor_type == "major_donor_ie_committee"                 ~ "major_donor_ie_committee",
      !is_resolvable & contributor_type == "pac" & status == "unresolved"           ~ "no_contributor_data",
      !is_resolvable & contributor_type == "pac" & status == "max_depth_reached"    ~ "max_depth_reached",
      !is_resolvable & low_data_sufficiency                                           ~ "low_data_sufficiency",
      !is_resolvable & contributor_type == "pac" & data_ok                           ~ "pac_low_5k_coverage",
      !is_resolvable                                                                   ~ "other_unresolvable",
      TRUE                                                                             ~ NA_character_
    )
  )

```


**Note:** federal and candidate PACs that are missing contributor ids have status, data_ok, etc. = NA.

```{r,eval= FALSE}
# data check for pac with no contributors found for this cycle
no_contrib_ids <- over_5k_tagged %>% filter(unresolvable_reason == "no_contributor_data") %>% pull(contributor_id)
contributions_full %>% filter(Recipient.Committee.ID %in% no_contrib_ids)
```


### compute % of each PAC contributors that can be classified

```{r}
# one known non-federal resolved PAC has no pac_id and will not receive override credit — expected
# if additional unexpected rows appear, add their pac_id to pac_manual_review
no_id_resolved <- pac_manual_review %>%
  filter(is_resolved, !coalesce(is_fec_federal, FALSE), !coalesce(is_candidates_pp, FALSE), is.na(pac_id))
if (nrow(no_id_resolved) > 1)
  cat(sprintf("WARNING: %d resolved non-federal PAC(s) have no pac_id — override credit will not be applied\n", nrow(no_id_resolved)))
no_id_resolved
```


```{r}
# pre-compute PAC contributor dependencies
# "pac" only: resolvable_amount for pac type uses child_resolvable_fraction, requiring ordering
# no child fraction dependency for candidate_pacs, fed pacs, pacs w/ no id : all processed differently
pac_deps <- over_5k_tagged %>%
  filter(contributor_type == "pac", !is.na(contributor_id)) %>%
  distinct(parent_pac_id, contributor_id) %>%
  group_by(parent_pac_id) %>%
  summarise(deps = list(contributor_id), .groups = "drop") %>%
  { setNames(.$deps, .$parent_pac_id) }

# topological sort with break on no-progress (handles cycles / missing deps)
processed     <- character(0)
ordered_pacs  <- character(0)
# federal and candidate PACs always short-circuit at !data_ok — exclude to avoid unnecessary iteration
pacs_to_order <- pac_summaries %>%
  filter(!coalesce(is_federal_pac, FALSE), !coalesce(is_candidate_pac, FALSE)) %>%
  arrange(desc(depth)) %>%
  pull(pac_id)

while (length(pacs_to_order) > 0) {
  n_before <- length(pacs_to_order)
  ready <- Filter(function(pid) {
    deps <- pac_deps[[pid]]
    is.null(deps) || all(deps %in% processed)
  }, pacs_to_order)
  ordered_pacs  <- c(ordered_pacs, ready)
  processed     <- c(processed, ready)
  pacs_to_order <- setdiff(pacs_to_order, ready)
  if (length(pacs_to_order) == n_before) break  # no progress: cycles or missing deps
}

stuck_pacs <- pacs_to_order # pacs not processed

# pre-split by parent_pac_id 
over_5k_by_pac    <- split(over_5k_tagged, over_5k_tagged$parent_pac_id)
classifiability_idx <- setNames(seq_len(nrow(pac_classifiability)), pac_classifiability$pac_id)
resolvable_vec      <- setNames(rep(NA_real_,  nrow(pac_classifiability)), pac_classifiability$pac_id)

# ── override credit ────────────────────────────────────────────────────────────
# apply override from pac_manual_review
# credits amounts toward resolvable fraction so the metric
# matches what we can actually classify.
#
# - Federal PACs (fec_id, by id or name) and Candidates/Political Parties PACs:
#   100% of total_received is classifiable (override covers the entire amount).
# - Other resolved PACs: only under_5k + unitemized counts 

override_resolvable_vec <- pac_classifiability %>%
  filter(pac_id %in% (pac_manual_review %>% filter(!is.na(pac_id), is_resolved) %>% pull(pac_id))) %>%
  left_join(
    pac_manual_review %>% filter(!is.na(pac_id)) %>%
      transmute(pac_id, is_fec_federal, is_candidates_pp, industry_override_under_5k),
    by = "pac_id"
  ) %>%
  transmute(
    pac_id,
    amount = case_when(
      # federal/candidate PACs with no classifiable industry get no credit
      (coalesce(is_fec_federal, FALSE) | coalesce(is_candidates_pp, FALSE)) &
        trimws(coalesce(industry_override_under_5k, "")) == "Uncategorized" ~ 0,
      coalesce(is_fec_federal, FALSE) | coalesce(is_candidates_pp, FALSE) ~
        coalesce(as.numeric(total_received), 0),
      TRUE ~
        coalesce(as.numeric(under_5k_amount), 0) + coalesce(as.numeric(unitemized_amount), 0)
    )
  ) %>%
  { setNames(.$amount, .$pac_id) }

# Federal PACs with no pac_id but a fec_id — matched by name in over_5k_tagged
fec_federal_pac_names <- pac_manual_review %>%
  filter(is.na(pac_id), is_fec_federal) %>%
  pull(Contributor.Name)


# ── end override ────────────────────────────────────────────────────────

# process orderable PACs
for (pid in ordered_pacs) {
  summary_row <- pac_classifiability[classifiability_idx[[pid]], ]
  if (!summary_row$data_ok) { resolvable_vec[[pid]] <- 0; next }

  direct_contribs <- over_5k_by_pac[[pid]]
  if (is.null(direct_contribs) || nrow(direct_contribs) == 0) { resolvable_vec[[pid]] <- 0; next }

  direct_contribs <- direct_contribs %>%
    mutate(
      child_resolvable_fraction = resolvable_vec[contributor_id],
      resolvable_amount = case_when(
        contributor_type == "individual_or_org"                              ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" & is_resolvable       ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" & !is_resolvable      ~ 0,
        contributor_type == "pac"                                             ~ contributor_amount * coalesce(child_resolvable_fraction, 0),
        contributor_type == "federal_pac" & contributor_name %in% fec_federal_pac_names ~ contributor_amount,
        contributor_type == "candidate_pac"                                   ~ contributor_amount,
        TRUE                                                                  ~ 0
      )
    )

  resolvable_vec[[pid]] <- (sum(direct_contribs$resolvable_amount, na.rm = TRUE) +
                             coalesce(override_resolvable_vec[pid], 0)) / summary_row$total_received
}

# process stuck pacs
remaining <- stuck_pacs

compute_fraction <- function(pid) {
  summary_row <- pac_classifiability[classifiability_idx[[pid]], ]
  if (!summary_row$data_ok) { resolvable_vec[[pid]] <<- 0; return() }

  direct_contribs <- over_5k_by_pac[[pid]]
  if (is.null(direct_contribs) || nrow(direct_contribs) == 0) { resolvable_vec[[pid]] <<- 0; return() }

  direct_contribs <- direct_contribs %>%
    mutate(
      child_resolvable_fraction = coalesce(resolvable_vec[contributor_id], 0),
      resolvable_amount = case_when(
        contributor_type == "individual_or_org"                              ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" & is_resolvable       ~ contributor_amount,
        contributor_type == "major_donor_ie_committee" & !is_resolvable      ~ 0,
        contributor_type == "pac"                                             ~ contributor_amount * child_resolvable_fraction,
        contributor_type == "federal_pac" & contributor_name %in% fec_federal_pac_names ~ contributor_amount,
        contributor_type == "candidate_pac"                                   ~ contributor_amount,
        TRUE                                                                  ~ 0
      )
    )

  resolvable_vec[[pid]] <<- (sum(direct_contribs$resolvable_amount, na.rm = TRUE) +
                              coalesce(override_resolvable_vec[pid], 0)) / summary_row$total_received
}

while (length(remaining) > 0) {
  # PACs whose stuck deps have all been resolved can now be processed normally
  ready <- Filter(function(pid) {
    deps <- pac_deps[[pid]]
    is.null(deps) || !any(deps %in% remaining)
  }, remaining)

  if (length(ready) == 0) {
    # nothing can be unblocked — true cycle remains; force all with cyclic deps = 0
    ready <- remaining
  }

  for (pid in ready) compute_fraction(pid)
  remaining <- setdiff(remaining, ready)
}

pac_resolvable_fractions <- pac_classifiability %>%
  mutate(resolvable_fraction = resolvable_vec[pac_id])

```

### audit stuck pacs

```{r,eval=FALSE}
# tracked_recipient_ids: used here and in pac_review_queue below
tracked_recipient_ids <- unique(x$Recipient.Committee.ID)

# audit stuck pacs — PACs that couldn't be topologically ordered (true cycles)
# is_tracked_recipient: stuck PAC is itself a tracked recipient committee in x
# — reviewer should apply that committee's known classification rather than treating it as unknown
stuck_pac_audit <- tibble(pac_id = stuck_pacs) %>%
  mutate(
    is_tracked_recipient = pac_id %in% tracked_recipient_ids,
    canonical_name       = x$Recipient.Committee[match(pac_id, x$Recipient.Committee.ID)],
    unresolved_deps      = sapply(pac_id, function(pid) paste(pac_deps[[pid]][pac_deps[[pid]] %in% stuck_pacs], collapse = ", "))
  )
stuck_pac_audit
```


# Compute coverage breakdowns for seed PACs


```{r}
# get names to use for pacs (for display)

# most frequent contributor name from transactions, keyed by contributor ID
pac_names_from_contributions <- contributions_agg %>%
  filter(!is.na(std_contributor_id) & std_contributor_id != "") %>%
  group_by(std_contributor_id) %>%
  summarise(
    Contributor.Name = {
      rns <- raw_name[!is.na(raw_name) & raw_name != ""]
      if (length(rns) > 0) names(sort(table(rns), decreasing = TRUE))[1] else NA_character_
    },
    .groups = "drop"
  ) %>%
  transmute(FILER_ID = std_contributor_id, Contributor.Name)

# fallback: registry name for PACs not seen as contributors
pac_names_registry <- id_map_full %>%
  group_by(FILER_ID) %>%
  slice(1) %>%
  ungroup() %>%
  transmute(FILER_ID = as.character(FILER_ID), Contributor.Name = NAML)

pac_names <- bind_rows(
  pac_names_from_contributions,
  pac_names_registry %>% filter(!FILER_ID %in% pac_names_from_contributions$FILER_ID)
)
```


# Send PACs to review queue


```{r}
over_5k_tagged <- over_5k_tagged %>%
  left_join(
    contributions_agg %>% select(Recipient.Committee.ID, entity_group, suspected_fed_pac),
    by = c("parent_pac_id" = "Recipient.Committee.ID", "entity_group")
  ) %>%
  mutate(
    suspected_fed_pac   = coalesce(suspected_fed_pac, FALSE),
    unresolvable_reason = case_when(
      suspected_fed_pac & is.na(unresolvable_reason) ~ "suspected_federal_pac",
      TRUE ~ unresolvable_reason
    )
  )
```



# Process PACs that are missing Contributor.IDs

Create a manual review queue for PACs that appear to be missing 
`Contributor.ID` ; these need to be reviewed at the transaction-level
to make sure we propagated the right id. 

```{r}
# build name/ID lookup from reference dataset
full_id_lookup <- id_map_full %>%
  filter(
    !is.na(FILER_ID), FILER_ID != "",
    FILER_TYPE %in% c("MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE", "RECIPIENT COMMITTEE"),
    STATUS == "ACTIVE" |
      (STATUS %in% c("INACTIVE", "TERMINATED") & !is.na(EFFECT_DT) & as.Date(EFFECT_DT) >= as.Date("2025-01-01"))
  ) %>%
  mutate(name_v5 = standardize_names(NAML) %>%
           str_remove_all("[[:punct:]]") %>%
          str_squish()) %>%
  distinct(name_v5, FILER_ID, FILER_TYPE, STATUS) %>%
  filter(name_v5 != "") %>%
  group_by(name_v5) %>%
  mutate(n_ids = n_distinct(FILER_ID)) %>%
  ungroup()


```

### proposed IDs based on name match

```{r}
# add name variant
over_5k_tagged <- over_5k_tagged %>% mutate(
  name_v5 = standardize_names(contributor_name) %>%
           str_remove_all("[[:punct:]]") %>%
          str_squish()
  )

# match no-id contributors to id_map entries by name_v5.
# uses distinct() because over_5k_tagged may have duplicate contributor_names
# (same contributor to multiple parent PACs).
full_id_propagation <- over_5k_tagged %>%
  filter(is.na(contributor_id)) %>%
  distinct(contributor_name, name_v5) %>%
  inner_join(full_id_lookup, by = "name_v5", relationship = "many-to-many") %>%
  mutate(needs_review = n_ids > 1) %>%
  select(contributor_name, name_v5, propagated_id = FILER_ID, n_ids, FILER_TYPE, STATUS, needs_review)
```

```{r}
# only care about propagating recipient committee IDs because that has implications for the traversal
# names matching both a recipient committee and a major donor are flagged as rc_needs_review.
names_also_matching_major_donor <- full_id_propagation %>%
  filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>%
  distinct(contributor_name)

propagated_effective <- full_id_propagation %>%
  filter(FILER_TYPE == "RECIPIENT COMMITTEE") %>%
  group_by(contributor_name) %>%
  mutate(n_rc_ids = n_distinct(propagated_id)) %>%
  ungroup() %>%
  mutate(
    also_matches_major_donor = contributor_name %in% names_also_matching_major_donor$contributor_name,
    propagation_type = case_when(
      n_rc_ids > 1           ~ "rc_needs_review",
      also_matches_major_donor ~ "rc_needs_review",
      TRUE                   ~ "recipient_committee"
    )
  )

# attach proposed_id/proposed_type to over_5k_tagged for all id_map matches
# (both unambiguous recipient_committee and rc_needs_review cases).
# Uses propagated_effective (all RC matches) so reviewers can decide on ambiguous
# cases manually via missing_id_review_queue below.
over_5k_tagged <- over_5k_tagged %>%
  left_join(
    propagated_effective %>%
      mutate(propagated_id = as.character(propagated_id)) %>%
      group_by(contributor_name) %>%
      summarise(
        proposed_id   = paste(sort(unique(propagated_id)), collapse = "; "),
        proposed_type = if_else(any(propagation_type == "rc_needs_review"),
                                "rc_needs_review", "recipient_committee"),
        .groups = "drop"
      ),
    by = "contributor_name"
  )
```

**NOTE:** sometimes a proposed_id is attached to a contributor that already has an id; 
          this happens when there are multiple contributors with the same name and some have
          and id while other don't.


### build review queue for pacs with missing ids

Record-level queue for (likely) pac contributors with no ID
includes any in-scope contributor (in over_5k_tagged: $5k+ to a traversal PAC) with:
   - PAC or federal-PAC language and no Contributor.ID, OR
   - a name matched to the id_map (proposed_id is not NA) - pac name doesnt alwasy have pac language

Reviewer will look at the contribution record and fill in the right id
in `corrected_id` after verification.

```{r}
missing_ids <- over_5k_tagged %>%
  filter(
    is.na(contributor_id),
    has_pac_language(coalesce(contributor_name, "")) |
    has_fed_pac_language(coalesce(contributor_name, "")) |
    !is.na(proposed_id)
  ) %>%
  distinct(parent_pac_id, entity_group, contributor_name, proposed_id, proposed_type)

n_pacs_by_name <- missing_ids %>%
  count(contributor_name, name = "n_other_recipients")

missing_id_review_queue <- missing_ids %>%
  left_join(
    prepped %>% # contribution-level records
      filter(is.na(std_contributor_id) | std_contributor_id == "") %>% 
      select(
        contribution_id, Contributor.ID, Recipient.Committee, Recipient.Committee.ID, entity_group,
        Contributor.Name, standardized_name, Amount, Start.Date, End.Date,
        Contributor.City, Contributor.State
      ),
    by = c("parent_pac_id" = "Recipient.Committee.ID", "entity_group")
  ) %>%
  left_join(n_pacs_by_name, by = "contributor_name") %>%
  transmute(
    contribution_id,
    contributor_id = Contributor.ID,
    contributor_name  = Contributor.Name,
    standardized_name,
    recipient_committee = Recipient.Committee,
    recipient_committee_id = parent_pac_id,
    amount              = Amount,
    start_date          = Start.Date,
    end_date            = End.Date,
    contributor_city    = Contributor.City,
    contributor_state   = Contributor.State,
    n_other_recipients,
    proposed_id,
    proposed_type,
    corrected_id        = NA_character_,
    is_fed_pac          = NA_character_,
    notes               = NA_character_
  ) %>%
  arrange(standardized_name, recipient_committee_id, start_date)
nrow(missing_id_review_queue)
```

```{r}
write_csv(missing_id_review_queue, sprintf("08_outputs/missing_id_review_queue_%s.csv", format(Sys.Date(), "%m-%d-%y")))
```


# Generate review queues for PACs

Any PAC that is not 100% classifiable by its $5k+
contributing entities needs to go to review so the reviewer
can determine how to classify the under $5k entities. 

### Generate review queue

```{r}
contributions_agg$Recipient.Committee.ID <- as.character(contributions_agg$Recipient.Committee.ID)
contributions$Recipient.Committee.ID <- as.character(contributions$Recipient.Committee.ID)

# all raw name variants seen for each PAC ID across the contribution data
# for reviewer to reference
pac_name_variants <- contributions %>%
  filter(!is.na(std_contributor_id) & std_contributor_id != "") %>%
  group_by(std_contributor_id) %>%
  summarise(
    name_variants = paste(sort(unique(Contributor.Name[!is.na(Contributor.Name) & Contributor.Name != ""])), collapse = " | "),
    .groups = "drop"
  ) %>%
  rename(pac_id = std_contributor_id)

# compute coverage totals for ALL PACs in the graph (seeds + deeper traversal)
# child_resolvable_fraction from pac_resolvable_fractions already incorporates overrides at child level
all_pac_coverage <- over_5k_tagged %>%
  left_join(
    pac_resolvable_fractions %>% select(pac_id, child_resolvable_fraction = resolvable_fraction),
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    resolvable_amount = case_when(
      contributor_type == "individual_or_org"                              ~ contributor_amount,
      contributor_type == "major_donor_ie_committee" & is_resolvable       ~ contributor_amount,
      contributor_type == "major_donor_ie_committee" & !is_resolvable      ~ 0,
      contributor_type == "candidate_pac"                                   ~ contributor_amount,
      contributor_type %in% c("pac", "federal_pac")                        ~ contributor_amount * coalesce(child_resolvable_fraction, 0),
      contributor_type %in% c("pac_no_id", "pac_unmatched_id")             ~ 0,
      TRUE                                                                  ~ 0
    )
  ) %>%
  group_by(parent_pac_id) %>%
  summarise(
    total_direct_over_5k = sum(contributor_amount, na.rm = TRUE),
    total_resolvable     = sum(resolvable_amount,  na.rm = TRUE),
    .groups = "drop"
  )

# DC PAC flag — check city/state from rows where the PAC appears as a CONTRIBUTOR.
# Uses a prefix match on "WASH" to catch common misspellings (Washongton, Washingotn, etc.),
# combined with state == "DC" 
dc_pac_ids <- prepped %>%
  filter(
    !is.na(std_contributor_id) & std_contributor_id != "",
    grepl("^WASH", toupper(str_squish(standardized_city))),
    toupper(str_squish(Contributor.State)) == "DC"
  ) %>%
  distinct(std_contributor_id) %>%
  rename(pac_id = std_contributor_id) %>%
  mutate(contributor_is_dc_based = TRUE)

# PACs already resolved in pac_manual_review will excluded from review queue
# is_resolved is FALSE for additional_review_req and PENDING/blank industry_override_under_5k,
# so those entries are not in already_reviewed_pac_ids/names and will still appear here.
already_reviewed_pac_ids <- pac_manual_review %>%
  filter(!is.na(pac_id), is_resolved) %>%
  pull(pac_id) %>% as.character()
already_reviewed_pac_names <- pac_manual_review %>%
  filter(is.na(pac_id), is_resolved) %>%
  pull(Contributor.Name)

# resolvable major donor/IE IDs — classified directly (name/occupation), not via PAC review
# unresolvable major donor/IE PACs (not in this set) are kept in the queue
major_donor_resolvable_ids <- over_5k_tagged %>%
  filter(contributor_type == "major_donor_ie_committee", is_resolvable, !is.na(contributor_id)) %>%
  pull(contributor_id) %>% unique()


unresolvable_contributor_ids <- over_5k_tagged %>%
  filter(
    (contributor_type == "major_donor_ie_committee" & !is_resolvable) |
    contributor_type %in% c("pac_unmatched_id", "pac_no_id", "federal_pac", "candidate_pac") |
    suspected_fed_pac
  ) %>%
  filter(!is.na(contributor_id)) %>%
  pull(contributor_id) %>% unique()

# no-ID seed PAC contributors
# these were excluded from seed_ids and never entered the traversal, so they don't appear
# in pac_summaries
no_id_seed_pacs <- x %>%
  filter(
    is.na(FILER_ID) | FILER_ID == "",
    coalesce(is_fed_pac, FALSE) |
      is_candidate_pac(Contributor.Name) |
      has_pac_language(Contributor.Name)
  ) %>%
  mutate(name_key = toupper(str_squish(coalesce(Contributor.Name, "")))) %>%
  group_by(name_key) %>%
  summarise(
    Contributor.Name = {
      rns <- Contributor.Name[!is.na(Contributor.Name) & Contributor.Name != ""]
      if (length(rns) > 0) names(sort(table(rns), decreasing = TRUE))[1] else NA_character_
    },
    is_federal_pac   = any(coalesce(is_fed_pac, FALSE)),
    is_candidate_pac = any(is_candidate_pac(Contributor.Name)),
    .groups = "drop"
  ) %>%
  transmute(
    pac_id              = NA_character_,
    Contributor.Name,
    is_seed             = TRUE,
    status              = "no_id",
    data_ok             = FALSE,
    pct_resolvable      = 0,
    is_federal_pac,
    is_candidate_pac,
    unresolvable_reason = case_when(
      is_federal_pac   ~ "federal_pac",
      is_candidate_pac ~ "candidate_pac",
      TRUE             ~ "no_id_pac"
    )
  )

# PAC-like contributors absent from pac_resolvable_fractions:
# - no-ID entries (pac_no_id / PAC language / suspected federal)
# - has-ID entries flagged as unresolvable but never traversed as a recipient PAC
no_id_pac_contributors <- over_5k_tagged %>%
  filter(
    (is.na(contributor_id) & (
      contributor_type %in% c("pac_no_id", "candidate_pac") |
      has_pac_language(coalesce(contributor_name, "")) |
      has_fed_pac_language(coalesce(contributor_name, "")) |
      suspected_fed_pac
    )) |
    (!is.na(contributor_id) &
      contributor_id %in% unresolvable_contributor_ids &
      !contributor_id %in% pac_resolvable_fractions$pac_id)
  ) %>%
  distinct(contributor_id, contributor_name, contributor_type, suspected_fed_pac) %>%
  filter(!(is.na(contributor_id) & contributor_name %in% no_id_seed_pacs$Contributor.Name)) %>%
  left_join(pac_names,  by = c("contributor_id" = "FILER_ID")) %>%
  left_join(dc_pac_ids, by = c("contributor_id" = "pac_id")) %>%
  transmute(
    pac_id               = contributor_id,
    Contributor.Name     = coalesce(Contributor.Name, contributor_name),
    name_variants        = NA_character_,
    is_seed              = coalesce(contributor_id %in% seed_ids, FALSE),
    is_tracked_recipient = coalesce(contributor_id %in% tracked_recipient_ids, FALSE),
    status               = if_else(is.na(contributor_id), "no_id", NA_character_),
    is_federal_pac       = coalesce(contributor_type == "federal_pac", FALSE),
    contributor_is_dc_based = coalesce(contributor_is_dc_based, FALSE),
    suspected_federal_pac = coalesce(suspected_fed_pac, FALSE) |
                            (contributor_is_dc_based & !coalesce(contributor_type == "federal_pac", FALSE)),
    suspected_federal_reason = if_else(contributor_is_dc_based, "dc_based", NA_character_),
    is_major_donor_ie    = coalesce(contributor_type == "major_donor_ie_committee", FALSE),
    is_candidate_pac     = coalesce(contributor_type == "candidate_pac", FALSE),
    low_data_sufficiency    = NA,
    unknown_data_sufficiency = NA,
    data_ok              = FALSE,
    total_received       = NA_real_,
    total_donated_all    = NA_real_,
    data_sufficiency     = NA_real_,
    unitemized_amount    = NA_real_,
    under_5k_amount      = NA_real_,
    over_5k_amount       = NA_real_,
    pct_over_5k          = NA_real_,
    total_direct_over_5k = 0,
    total_resolvable     = 0,
    pct_resolvable       = 0,
    unresolvable_reason  = case_when(
      coalesce(contributor_type == "federal_pac", FALSE) ~ "federal_pac",
      suspected_federal_pac                              ~ "suspected_federal_pac",
      contributor_type == "major_donor_ie_committee"     ~ "major_donor_ie_committee",
      contributor_type == "pac_unmatched_id"             ~ "pac_unmatched_id",
      contributor_type == "candidate_pac"                ~ "candidate_pac",
      is.na(contributor_id)                              ~ "no_id_pac",
      TRUE                                               ~ coalesce(contributor_type, "other")
    )
  )

combined_review_queue <- pac_resolvable_fractions %>%
  filter(
    is.na(resolvable_fraction) | resolvable_fraction < 1 |
    pac_id %in% dc_pac_ids$pac_id |
    coalesce(is_candidate_pac, FALSE) |
    coalesce(suspected_federal_pac, FALSE) |
    status == "unresolved" |
    coalesce(small_dollar_amount, 0) > 0 |
    coalesce(under_5k_amount, 0) > 0 |
    pac_id %in% unresolvable_contributor_ids
  ) %>%
  left_join(pac_names,              by = c("pac_id" = "FILER_ID")) %>%
  left_join(all_pac_coverage,       by = c("pac_id" = "parent_pac_id")) %>%
  left_join(contributor_type_lookup,by = c("pac_id" = "contributor_id")) %>%
  left_join(dc_pac_ids,             by = "pac_id") %>%
  left_join(pac_name_variants,      by = "pac_id") %>%
  # exclude resolvable major donor/IE without PAC language — classified directly, not via PAC review
  filter(!(is_major_donor_ie & !has_pac_language(coalesce(Contributor.Name, "")) &
           !coalesce(is_federal_pac, FALSE) & !coalesce(suspected_federal_pac, FALSE) &
           pac_id %in% major_donor_resolvable_ids)) %>%
  mutate(
    is_seed                  = pac_id %in% seed_ids,
    is_tracked_recipient     = pac_id %in% tracked_recipient_ids,
    total_direct_over_5k     = coalesce(total_direct_over_5k, 0),
    total_resolvable         = coalesce(total_resolvable, 0),
    pct_resolvable           = coalesce(resolvable_fraction, 0),
    contributor_is_dc_based  = coalesce(contributor_is_dc_based, FALSE),
    suspected_federal_pac    = coalesce(suspected_federal_pac, FALSE) |
                               (contributor_is_dc_based & !coalesce(is_federal_pac, FALSE)),
    suspected_federal_reason = case_when(
      contributor_is_dc_based & !coalesce(is_federal_pac, FALSE) & !is.na(suspected_federal_reason) ~
        paste(suspected_federal_reason, "dc_based", sep = "; "),
      contributor_is_dc_based & !coalesce(is_federal_pac, FALSE) ~ "dc_based",
      TRUE ~ suspected_federal_reason
    ),
    unresolvable_reason = case_when(
      coalesce(is_federal_pac, FALSE)                                      ~ "federal_pac",
      suspected_federal_pac                                                ~ "suspected_federal_pac",
      is_major_donor_ie & has_pac_language(coalesce(Contributor.Name, "")) ~ "major_donor_ie_committee",
      !is.na(contributor_type)                                             ~ contributor_type,
      status == "unresolved"                                               ~ "no_contributor_data",
      coalesce(low_data_sufficiency, FALSE)                                ~ "low_data_sufficiency",
      coalesce(unknown_data_sufficiency, FALSE)                            ~ "unknown_data_sufficiency",
      total_direct_over_5k == 0                                           ~ "no_over_5k_contributors",
      pct_resolvable < OVER_5K_COVERAGE_THRESHOLD                          ~ "low_pct_resolvable",
      TRUE                                                                 ~ "other"
    )
  ) %>%
  select(
    pac_id, Contributor.Name, name_variants, is_seed, is_tracked_recipient,
    status, unresolvable_reason,
    is_federal_pac, suspected_federal_pac, suspected_federal_reason,
    is_major_donor_ie, is_candidate_pac, low_data_sufficiency, unknown_data_sufficiency, data_ok,
    contributor_is_dc_based,
    total_received, total_donated_all, data_sufficiency,
    unitemized_amount, under_5k_amount, over_5k_amount, pct_over_5k,
    total_direct_over_5k, total_resolvable, pct_resolvable
  ) %>%
  bind_rows(no_id_seed_pacs, no_id_pac_contributors) %>%
  filter(
    !coalesce(pac_id, "") %in% already_reviewed_pac_ids,
    !coalesce(Contributor.Name, "") %in% already_reviewed_pac_names
  )

nrow(combined_review_queue)
```


```{r, eval = FALSE}
# data check
unique(combined_review_queue$unresolvable_reason)
combined_review_queue %>% filter(unresolvable_reason == "unknown_data_sufficiency")
combined_review_queue %>% filter(unresolvable_reason == "pac_unmatched_id")
combined_review_queue %>% filter(unresolvable_reason == "other")
combined_review_queue %>% filter(unresolvable_reason == "no_id_pac")
combined_review_queue %>% filter(unresolvable_reason == "pac_unmatched_id")
combined_review_queue %>% filter(unresolvable_reason == "no_contributor_data")
combined_review_queue %>% filter(unresolvable_reason == "suspected_federal_pac")
```

```{r}
today <- format(Sys.Date(), "%m-%d-%y")
PAC_REVIEW_OUTPUT_PATH <- paste0("08_outputs/pac_review_queue_", today, ".csv")
write_csv(combined_review_queue,PAC_REVIEW_OUTPUT_PATH)
```


### build review for all (in-scope) self-contributors

Check that all of these are either mistakes or seem like a non-PAC
arm of an org contributing to the org's PAC. 

```{r}
# self-contributions among all traversal PACs (seed PACs + every PAC encountered as a $5k contributor).
# catches these two cases:
#   1. ID match: std_contributor_id == Recipient.Committee.ID (already flagged as is_self_contribution,
#      which sets std_contributor_id to NA — use effective_Contributor.ID for the raw id)
#   2. Name match: Contributor.Name == Recipient.Committee (slightly processed)

traversal_pacs <- pac_summaries$pac_id

self_contributions_review <- prepped %>%
  filter(Recipient.Committee.ID %in% traversal_pacs, Amount > 0) %>%
  mutate(
    recipient_name_std   = toupper(str_squish(Recipient.Committee)),
    contributor_name_std = toupper(str_squish(Contributor.Name)),
    self_by_id           = coalesce(is_self_contribution, FALSE),
    self_by_name         = contributor_name_std == recipient_name_std
  ) %>%
  filter(self_by_id | self_by_name) %>%
  left_join(
    pac_classifiability %>% select(pac_id, total_received),
    by = c("Recipient.Committee.ID" = "pac_id")
  ) %>%
  transmute(
    Recipient.Committee.ID,
    Recipient.Committee,
    Contributor.Name,
    Contributor.ID,
    effective_Contributor.ID,
    Amount,
    self_contribution_reason = case_when(
      self_by_id & self_by_name ~ "id_and_name_match",
      self_by_id                ~ "id_match",
      TRUE                      ~ "name_match"
    ),
    pac_total_received = total_received
  ) %>%
  arrange(Recipient.Committee.ID, desc(Amount))

self_contributions_review
cat(sprintf("Self-contributions in traversal PACs: %d rows across %d PACs\n",
            nrow(self_contributions_review),
            n_distinct(self_contributions_review$Recipient.Committee.ID)))

write_csv(self_contributions_review, "08_outputs/self_contributions_review_queue.csv")
```


# BUILD PAC CLASSIFICATION INPUT

```{r}
library(digest)
library(uuid)
```

```{r}
contributions_agg$Recipient.Committee.ID <- as.character(contributions_agg$Recipient.Committee.ID)
prepped$Recipient.Committee.ID <- as.character(prepped$Recipient.Committee.ID)

hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16))
}

# get entity_group for each $5k+ individual/org contributor
over_5k_indiv <- over_5k_tagged %>%
  filter(contributor_type == "individual_or_org" | (contributor_type == "major_donor_ie_committee" & is_resolvable == TRUE))

direct_for_classification <- over_5k_indiv %>%
  inner_join(
    prepped %>% filter(!is_pac, !is_fed_pac, !is_candidate_pac),
    by = c("parent_pac_id" = "Recipient.Committee.ID", "entity_group")
  ) %>%
  mutate(entity_type = if_else(
    sapply(Contributor.Name, is_individual), "individual", "organization"
  ))

# load uuid map from 0802 (canonical_name -> entity_uuid)
# so PAC contributors get entity_id_normalized consistent with direct-contributor entity_id_normalized.
if (!exists("canonical_uuid_map")) {
  cumap_path <- "08_outputs/canonical_uuid_map.csv"
  if (file.exists(cumap_path)) {
    canonical_uuid_map <- read.csv(cumap_path, stringsAsFactors = FALSE)
    cat(sprintf("Loaded canonical_uuid_map: %d canonical names\n", nrow(canonical_uuid_map)))
  } else {
    canonical_uuid_map <- NULL
    cat("Warning: canonical_uuid_map.csv not found — entity_id_normalized will be NA for all PAC contributors. Run 0802 first.\n")
  }
}

# format to match 0801 classification_input schema
# entity_id: hash-based id preserved for downstream compatibility.
# entity_id_normalized: UUID from canonical_uuid_map via Normalized column — consistent with 0802's
#   direct-contributor entity_id_normalized; NA if the contributor has no Normalized mapping.
pac_classification_input <- direct_for_classification %>%
  mutate(
    Recipient.Committee.ID = parent_pac_id,
    entity_id = if_else(
      !is.na(std_contributor_id),
      paste0("id_", std_contributor_id),
      sapply(Contributor.Name, function(n) hash_id(n, "indiv"))
    ),
    unit_key = paste(entity_id,
                     coalesce(processed_employer_name, ""),
                     coalesce(processed_occupation, ""),
                     sep = "||"),
    classification_unit_id = sapply(unit_key, function(k) hash_id(k, "unit")),
    contribution_uuid      = replicate(n(), uuid::UUIDgenerate())
  ) %>%
  select(-unit_key)

if (!is.null(canonical_uuid_map)) {
  cumap_for_join <- canonical_uuid_map %>% select(canonical_name, entity_id_normalized = entity_uuid)
  duped_names <- cumap_for_join %>% group_by(canonical_name) %>% filter(n() > 1) %>% pull(canonical_name) %>% unique()
  if (length(duped_names) > 0) {
    cat(sprintf("WARNING: %d canonical_name(s) appear multiple times in canonical_uuid_map — using first match per name\n", length(duped_names)))
    cumap_for_join <- cumap_for_join %>% group_by(canonical_name) %>% slice(1) %>% ungroup()
  }
  pac_classification_input <- pac_classification_input %>%
    mutate(Normalized_clean = str_squish(coalesce(Normalized, ""))) %>%
    left_join(cumap_for_join, by = c("Normalized_clean" = "canonical_name")) %>%
    select(-Normalized_clean)
} else {
  pac_classification_input <- pac_classification_input %>%
    mutate(entity_id_normalized = NA_character_)
}

cat(sprintf("PAC contributor entity_id_normalized: %d / %d rows matched via Normalized\n",
            sum(!is.na(pac_classification_input$entity_id_normalized)),
            nrow(pac_classification_input)))

# exclude contributors to PACs already classified as Candidates/Political Parties -
# those PACs full amounts are assigned that category directly in match_pac_classifications.
# also exclude any PAC identified as a candidate committee by name pattern that
# isn't yet in pac_manual_review just in case
candidates_pp_pac_ids <- pac_manual_review %>%
  filter(!is.na(pac_id), is_candidates_pp) %>%
  pull(pac_id) %>% as.character()

candidate_pac_ids_by_name <- prepped %>%
  filter(is_candidate_pac, !is.na(std_contributor_id)) %>%
  pull(std_contributor_id) %>%
  unique()

all_candidate_pac_ids <- unique(c(candidates_pp_pac_ids, candidate_pac_ids_by_name))

pac_classification_input <- pac_classification_input %>%
  filter(!parent_pac_id %in% all_candidate_pac_ids)

```


### Match back UUIDs from prior classifications

```{r}
pad_zip <- function(z) {
  z <- trimws(as.character(z))
  needs_pad <- !is.na(z) & z != "" & str_detect(z, "^[0-9]+$") & nchar(z) < 5
  z[needs_pad] <- paste0(strrep("0", 5 - nchar(z[needs_pad])), z[needs_pad])
  substr(z, 1, 5)
}

normalize_missing <- function(x) {
  x <- trimws(as.character(x))
  if_else(x %in% c("", "N/A", "NA", "None", "n/a") | is.na(x), "N/A", x)
}
```

```{r}
already_classified_files <- list.files(
  "08_inputs",
  pattern = "^already_classified_contributions_.*\\.csv$",
  full.names = TRUE
)

if (length(already_classified_files) > 0) {
  most_recent <- already_classified_files[which.max(file.mtime(already_classified_files))]
  cat(sprintf("Loading %s for PAC contributor UUID match-back\n", basename(most_recent)))

  categorized_pac <- read_csv(most_recent, col_types = cols(.default = "c")) %>%
    mutate(Amount = as.numeric(Amount))

  # PAC rows in older already-classified files have Recipient.Committee.ID = NA because
  # pac_classification_input used parent_pac_id (not Recipient.Committee.ID) as the key.
  # Recover parent_pac_id via the prior pac_classification_input.csv output, joining on uuid.
  prior_pac_input_path <- "08_outputs/pac_classification_input.csv"
  if (any(is.na(categorized_pac$Recipient.Committee.ID)) && file.exists(prior_pac_input_path)) {
    prior_pac_input <- read_csv(prior_pac_input_path, col_types = cols(.default = "c")) %>%
      select(uuid, parent_pac_id) %>%
      mutate(parent_pac_id = sub("\\.0$", "", trimws(parent_pac_id)))
    categorized_pac <- categorized_pac %>%
      left_join(prior_pac_input, by = "uuid") %>%
      mutate(Recipient.Committee.ID = coalesce(
        na_if(trimws(Recipient.Committee.ID), ""),
        na_if(trimws(parent_pac_id), "")
      )) %>%
      select(-parent_pac_id)
    cat(sprintf("  Recovered Recipient.Committee.ID for %d PAC rows via prior pac_classification_input.csv\n",
                sum(!is.na(categorized_pac$Recipient.Committee.ID))))
  }

  # strip trailing .0 
  categorized_pac <- categorized_pac %>%
    mutate(Recipient.Committee.ID = sub("\\.0+$", "", trimws(coalesce(Recipient.Committee.ID, ""))))
  categorized_pac$Recipient.Committee.ID[categorized_pac$Recipient.Committee.ID == ""] <- NA_character_

  # sanity check 
  cat(sprintf("  %d rows still have Recipient.Committee.ID = NA after recovery (will be dropped)\n",
              sum(is.na(categorized_pac$Recipient.Committee.ID))))

  # check (b): every (Recipient.Committee.ID, Recipient.Name) combo in categorized_pac
  # should be a known combo in contributions_full. An unknown combo means the wrong ID
  # was propagated to a recipient that doesn't appear with that ID in the current data.
  known_id_name_combos <- contributions_full %>%
    filter(!is.na(Recipient.Committee.ID), !is.na(Recipient.Name)) %>%
    distinct(Recipient.Committee.ID, Recipient.Name)

  id_name_check <- categorized_pac %>%
    filter(!is.na(Recipient.Committee.ID), !is.na(Recipient.Name)) %>%
    distinct(Recipient.Committee.ID, Recipient.Name) %>%
    anti_join(known_id_name_combos, by = c("Recipient.Committee.ID", "Recipient.Name")) %>%
    arrange(Recipient.Committee.ID)

  cat(sprintf("  (b) ID→name consistency: %d (ID, name) combos not found in contributions_full — investigate id_name_check\n",
              nrow(id_name_check)))
  if (nrow(id_name_check) > 0) print(id_name_check)


  categorized_pac <- categorized_pac %>%
    
    filter(!is.na(Recipient.Committee.ID), Recipient.Committee.ID != "") %>%
    mutate(
      legacy_match_id = make_row_hash(
        Contributor.Name, Contributor.ID, Amount, Recipient.Committee.ID,
        Contributor.City, normalize_missing(Contributor.Employer),
        normalize_missing(Contributor.Occupation), pad_zip(Contributor.Zip.Code),
        Start.Date
      )
    ) %>%
    group_by(legacy_match_id) %>%
    mutate(dup_rank = row_number()) %>%
    ungroup()

  cat(sprintf("  %d previously classified PAC contribution rows available for match-back\n",
              nrow(categorized_pac)))
} else {
  categorized_pac <- NULL
  cat("No already_classified_contributions file found — all PAC contributor UUIDs will be fresh.\n")
}
```

```{r}
if (!is.null(categorized_pac)) {
  pac_classification_input <- pac_classification_input %>%
    mutate(
      legacy_match_id = make_row_hash(
        Contributor.Name, Contributor.ID, as.numeric(Amount), parent_pac_id,
        Contributor.City, normalize_missing(Contributor.Employer),
        normalize_missing(Contributor.Occupation), pad_zip(Contributor.Zip.Code),
        Start.Date
      )
    ) %>%
    group_by(legacy_match_id) %>%
    mutate(dup_rank = row_number()) %>%
    ungroup() %>%
    left_join(
      categorized_pac %>% select(legacy_match_id, dup_rank, prior_uuid = uuid),
      by = c("legacy_match_id", "dup_rank")
    ) %>%
    mutate(uuid = coalesce(prior_uuid, contribution_uuid)) %>%
    select(-dup_rank, -prior_uuid)

  n_matched <- sum(pac_classification_input$uuid != pac_classification_input$contribution_uuid,
                   na.rm = TRUE)
  cat(sprintf("PAC contributor UUID match-back: %d / %d matched from prior classifications\n",
              n_matched, nrow(pac_classification_input)))

  # check (a): for matched rows, verify key fields are consistent between pac_classification_input
  # and the categorized_pac row they matched to (catching any hash collision or wrong recovery).
  if (n_matched > 0) {
    matched_pci <- pac_classification_input %>%
      filter(uuid != contribution_uuid) %>%
      select(uuid, pci_name = Contributor.Name, pci_amount = Amount,
             pci_pac_id = parent_pac_id, pci_date = Start.Date, pci_contrib_id = Contributor.ID)

    field_check <- matched_pci %>%
      left_join(
        categorized_pac %>% select(uuid,
          cat_name = Contributor.Name, cat_amount = Amount,
          cat_pac_id = Recipient.Committee.ID, cat_date = Start.Date,
          cat_contrib_id = Contributor.ID
        ),
        by = "uuid"
      ) %>%
      filter(
        (!is.na(cat_name) & pci_name != cat_name) |
        (!is.na(cat_amount) & abs(as.numeric(pci_amount) - as.numeric(cat_amount)) > 0.01) |
        (!is.na(cat_pac_id) & pci_pac_id != cat_pac_id) |
        (!is.na(cat_date) & as.character(pci_date) != as.character(cat_date))
      )
    cat(sprintf("  (a) UUID field consistency: %d / %d matched rows have field mismatches (expect 0) — investigate field_check\n",
                nrow(field_check), n_matched))
  }
} else {
  pac_classification_input <- pac_classification_input %>%
    mutate(uuid = contribution_uuid)
}
```

### save files to use later

```{r}
# carry entity_id_normalized from pac_input.csv for seed PACs
# x (pac_input.csv) rows have entity_id (the UUID from 0802's Normalized-name
# entity resolution). One seed PAC -> one UUID; flag conflicts where FILER_ID
# maps to more than one distinct UUID (indicates a name-map gap).
seed_entity_ids <- x %>%
  filter(!is.na(FILER_ID) & trimws(as.character(FILER_ID)) != "") %>%
  mutate(FILER_ID = as.character(FILER_ID)) %>%
  group_by(FILER_ID) %>%
  summarise(
    entity_id_normalized = {
      ids <- na.omit(unique(trimws(entity_id)))
      if (length(ids) == 1) ids else NA_character_
    },
    n_entity_ids = n_distinct(na.omit(trimws(entity_id))),
    .groups = "drop"
  ) %>%
  rename(pac_id = FILER_ID)

n_conflict <- sum(seed_entity_ids$n_entity_ids > 1, na.rm = TRUE)
if (n_conflict > 0)
  cat(sprintf("Warning: %d seed PAC(s) map to >1 entity_id_normalized in pac_input.csv — entity_id_normalized set to NA for those. Fix name map upstream.\n", n_conflict))

pac_classifiability <- pac_classifiability %>%
  left_join(seed_entity_ids %>% select(pac_id, entity_id_normalized), by = "pac_id")

cat(sprintf("entity_id_normalized from pac_input.csv: %d / %d PACs matched\n",
            sum(!is.na(pac_classifiability$entity_id_normalized)), nrow(pac_classifiability)))
```

```{r}
# entity_id_normalized for non-seed PACs 
# seeds get their UUID from pac_input.csv 
# non-seeds: look up via canonical_uuid_map 
# then pac_entity_id_registry.csv (persistent store keyed by FILER_ID),
# then generate a new UUID and append to the registry

PAC_REGISTRY_PATH <- "08_inputs/pac_entity_id_registry.csv"

if (file.exists(PAC_REGISTRY_PATH)) {
  pac_entity_id_registry <- read.csv(PAC_REGISTRY_PATH, stringsAsFactors = FALSE,
                                     colClasses = c(pac_id = "character"))
} else {
  pac_entity_id_registry <- tibble(pac_id = character(), entity_uuid = character(),
                                    added_date = character())
}

# check that each FILER_ID must map to exactly one Normalized name.
# multiple Normalized values for the same PAC means the name map is inconsistent.
pac_normalized_check <- contributions_agg %>%
  filter(!is.na(std_contributor_id), std_contributor_id != "",
         !is.na(Normalized), str_squish(Normalized) != "") %>%
  mutate(Normalized = str_squish(Normalized)) %>%
  group_by(std_contributor_id) %>%
  summarise(n_normalized = n_distinct(Normalized),
            normalized_values = paste(sort(unique(Normalized)), collapse = " | "),
            .groups = "drop")

pac_name_conflicts <- pac_normalized_check %>% filter(n_normalized > 1)
if (nrow(pac_name_conflicts) > 0) {
  cat(sprintf("Note: %d PAC(s) have multiple Normalized names (unresolved name variants). Normalized-name lookup skipped for those; FILER_ID registry or fresh UUID used instead.\n",
              nrow(pac_name_conflicts)))
}

# PACs with multiple Normalized names get pac_normalized = NA — they skip the
# canonical_uuid_map lookup and fall through to the FILER_ID registry or a fresh UUID.
pac_normalized_names <- pac_normalized_check %>%
  mutate(pac_normalized = if_else(n_normalized > 1, NA_character_, normalized_values)) %>%
  select(pac_id = std_contributor_id, pac_normalized)

non_seed_pacs <- pac_classifiability %>%
  filter(is.na(entity_id_normalized)) %>%
  select(pac_id) %>%
  left_join(pac_normalized_names, by = "pac_id") %>%
  # conflicted seed PACs (>1 entity_id in x) skip canonical_uuid_map —
  # their Normalized name is ambiguous; the FILER_ID registry is the stable fallback.
  mutate(pac_normalized = if_else(
    pac_id %in% (seed_entity_ids %>% filter(n_entity_ids > 1) %>% pull(pac_id)),
    NA_character_, pac_normalized
  ))

# step 1 : canonical_uuid_map lookup
if (!is.null(canonical_uuid_map) && nrow(non_seed_pacs) > 0) {
  non_seed_pacs <- non_seed_pacs %>%
    left_join(canonical_uuid_map %>% select(canonical_name, entity_uuid),
              by = c("pac_normalized" = "canonical_name"))
} else {
  non_seed_pacs <- non_seed_pacs %>% mutate(entity_uuid = NA_character_)
}

# flag if any PAC matched multiple rows in canonical_uuid_map (indicates duplicate
# canonical_name entries in the map, which should never happen given distinct() at build time)
multi_uuid_match <- non_seed_pacs %>%
  filter(!is.na(entity_uuid)) %>%
  group_by(pac_id) %>%
  filter(n() > 1) %>%
  ungroup()
if (nrow(multi_uuid_match) > 0) {
  cat(sprintf("WARNING: %d PAC(s) matched multiple entries in canonical_uuid_map — collapsing to NA (will fall through to registry)\n",
              n_distinct(multi_uuid_match$pac_id)))
  print(multi_uuid_match)
  non_seed_pacs <- non_seed_pacs %>%
    mutate(entity_uuid = if_else(pac_id %in% unique(multi_uuid_match$pac_id), NA_character_, entity_uuid)) %>%
    distinct(pac_id, .keep_all = TRUE)
}

n_from_map <- sum(!is.na(non_seed_pacs$entity_uuid))

# step 2: persistent registry lookup by FILER_ID (stable across runs)
non_seed_pacs <- non_seed_pacs %>%
  left_join(pac_entity_id_registry %>% select(pac_id, registry_uuid = entity_uuid),
            by = "pac_id") %>%
  mutate(entity_uuid = coalesce(entity_uuid, registry_uuid)) %>%
  select(-registry_uuid)
n_from_registry <- sum(!is.na(non_seed_pacs$entity_uuid)) - n_from_map

# step 3: generate new UUIDs for still-unmatched PACs; persist to registry
needs_new <- is.na(non_seed_pacs$entity_uuid)
if (any(needs_new)) {
  non_seed_pacs$entity_uuid[needs_new] <- uuid::UUIDgenerate(n = sum(needs_new))

  new_entries <- non_seed_pacs[needs_new, c("pac_id", "entity_uuid")] %>%
    mutate(added_date = as.character(Sys.Date())) %>%
    filter(!pac_id %in% pac_entity_id_registry$pac_id)

  if (nrow(new_entries) > 0) {
    write.csv(bind_rows(pac_entity_id_registry, new_entries),
              PAC_REGISTRY_PATH, row.names = FALSE)
    cat(sprintf("Appended %d new PAC entity IDs to pac_entity_id_registry.csv\n",
                nrow(new_entries)))
  }
}

cat(sprintf(
  "Non-seed PAC entity_id_normalized: %d via canonical_uuid_map, %d via registry, %d newly generated\n",
  n_from_map, n_from_registry, sum(needs_new)
))

pac_classifiability <- pac_classifiability %>%
  left_join(non_seed_pacs %>% select(pac_id, non_seed_uuid = entity_uuid), by = "pac_id") %>%
  mutate(entity_id_normalized = coalesce(entity_id_normalized, non_seed_uuid)) %>%
  select(-non_seed_uuid)

cat(sprintf("entity_id_normalized total coverage: %d / %d PACs\n",
            sum(!is.na(pac_classifiability$entity_id_normalized)), nrow(pac_classifiability)))
```

```{r}
write_csv(pac_classification_input,    paste0("08_outputs/pac_classification_input_", today, ".csv"))
write_csv(over_5k_tagged,              paste0("08_outputs/over_5k_tagged_", today, ".csv"))
write_csv(pac_classifiability,         paste0("08_outputs/pac_classifiability_", today, ".csv"))

# save the traversal order — needed to replay topological sort for industry breakdown
write_csv(tibble(pac_id = ordered_pacs, order = seq_along(ordered_pacs)),
          paste0("08_outputs/pac_traversal_order_", today, ".csv"))
write_csv(tibble(pac_id = stuck_pacs),  paste0("08_outputs/pac_stuck_pacs_", today, ".csv"))

```


**NOTE**: 
UUID conflicts can arise when a ballot measure committee also contributes to
another PAC, causing it to be traversed in the PAC graph. Its incoming contributions
then appear in both paths (direct via 0802, pac via 0803). The dedup chunk below
resolves conflicts by keeping the direct row, but the upstream fix is to exclude
0802 target committees from pac_summaries in 0802b so they are never traversed.

```{r}
# combine with 0801's classification_input (run 0801 first so the file exists)
existing <- read_csv("08_outputs/classification_input_09-07-26.csv", col_types = cols(.default = "c"))
existing$Cycle <- as.integer(existing$Cycle)
existing$Amount <- as.numeric(existing$Amount)
existing$has_pac_language <- as.logical(existing$has_pac_language)
existing$rowid <- as.character(existing$rowid)
existing$row_id <- as.integer(existing$row_id)
existing$contributor_path <- "direct"

pac_classification_input$rowid <- as.character(pac_classification_input$rowid)
pac_classification_input$entity_type_ambiguous <- as.character(pac_classification_input$entity_type_ambiguous)
pac_classification_input$is_fed_pac <- as.character(pac_classification_input$is_fed_pac)
pac_classification_input$contributor_path <- "pac"

combined_classification_input <- bind_rows(existing, pac_classification_input)

# verigy
stopifnot(nrow(combined_classification_input) == nrow(existing) + nrow(pac_classification_input))
cat(sprintf("direct: %d rows | pac: %d rows | combined: %d rows\n",
            nrow(existing), nrow(pac_classification_input), nrow(combined_classification_input)))

# detect UUID conflicts — do NOT deduplicate automatically here; review uuid_conflicts first,
# then run the dedup chunk
uuid_conflicts <- combined_classification_input %>%
  group_by(uuid) %>% filter(n() > 1) %>% ungroup() %>% arrange(uuid)
cat(sprintf("UUID conflicts: %d rows across %d uuids appear in both paths — review uuid_conflicts before deduplicating\n",
            nrow(uuid_conflicts), n_distinct(uuid_conflicts$uuid)))
if (nrow(uuid_conflicts) > 0) {
  print(uuid_conflicts %>% select(uuid, contributor_path, Contributor.Name,
                                  Recipient.Committee.ID, Recipient.Name, Amount, parent_pac_id))
}
cat(sprintf("uuid uniqueness: %d distinct uuids across %d rows\n",
            n_distinct(combined_classification_input$uuid), nrow(combined_classification_input)))
```


```{r}
# run this chunk only after reviewing uuid_conflicts above
# keeps the direct row for each conflicted UUID
if (nrow(uuid_conflicts) > 0) {
  combined_classification_input <- combined_classification_input %>%
    arrange(uuid, contributor_path) %>%  # "direct" sorts before "pac"
    distinct(uuid, .keep_all = TRUE)
  cat(sprintf("After dedup: %d rows (%d removed)\n",
              nrow(combined_classification_input), nrow(uuid_conflicts) / 2))
} else {
  cat("No conflicts to resolve.\n")
}
```


```{r, eval = FALSE}
# data check 
combined_classification_input %>%
  group_by(uuid) %>%
  filter(n() > 1) %>%
  select(uuid, contributor_path, Contributor.Name, Contributor.ID,
         Recipient.Committee.ID, Recipient.Name, Amount, parent_pac_id) %>%
  arrange(uuid)
```

```{r}
write_csv(combined_classification_input, paste0("08_outputs/classification_input_combined_", today, ".csv"))
cat("Combined rows:", nrow(combined_classification_input),
    "  (PAC added:", nrow(pac_classification_input), ")\n")
```



