---
title: "Contributor Matching to OpenSecrets Data"
output: html_notebook
---

# To add
- rule for when the first name is just an initial
- add in crosswalk
- will read classified_contributors.csv (after EDD/H1B matches, not classification_input.csv)

# Load data and packages


```{r}
here::i_am("0806_match_opensecrets.Rmd")
library(here)
library(stringr)
library(stringdist)
library(fuzzyjoin)
library(dplyr)
library(readr)
library(reclin2)
library(phonics)
library(igraph)

source("standardization_helpers.R")
```


### load data

```{r}
x <- read.csv("08_outputs/classification_input.csv") #classified_contributors.csv
ref <- read.csv("../../data/03_input/masterfile/running_list_opensecrets_alt.csv")
```

```{r}
colnames(x)
```


### apply functions to normalize + process names

```{r}
# process contributors data
x <- x %>%
  mutate(
    # contributor name
    standardized_name = standardize_names(canonical_name),
    standardized_name = fix_typos(standardized_name),
    processed_name = remove_pac_ref(standardized_name),
    processed_name = remove_unit_ref(processed_name),
    #processed_name = replace_committee_text(processed_name),
    processed_name = replace_business_text(processed_name),
    processed_name = str_replace_all(processed_name, " CA ", " CALIFORNIA "),
    processed_name = str_remove_all(processed_name, "\\s*\\(.*?\\)"), # removes everything in parentheses
    processed_name = str_squish(processed_name),
    
    # contributor employer
    standardized_employer_name = standardize_names(employer),
    standardized_employer_name = fix_typos(standardized_employer_name),
    processed_employer_name = remove_pac_ref(standardized_employer_name),
    processed_employer_name = remove_unit_ref(processed_employer_name),
    #processed_employer_name = replace_committee_text(processed_employer_name),
    processed_employer_name = replace_business_text(processed_employer_name),
    processed_employer_name = standardize_occupation_employer(processed_employer_name), # standardize how 'none', 'not employed', etc are marked
    processed_employer_name = str_replace_all(processed_employer_name, " CA ", " CALIFORNIA "),
    processed_employer_name = str_remove_all(processed_employer_name, "\\s*\\(.*?\\)"), # removes everything in parentheses
    processed_employer_name = str_squish(processed_employer_name),
    
    # contributor occupation
    standardized_occupation = standardize_names(occupation),
    processed_occupation = standardize_occupation_employer(standardized_occupation),
    
    # contributor city
    standardized_city = standardize_names(city),
    
  ) 

#x$zip_code_processed <- substr(x$Contributor.Zip.Code,1,5) # first five digits of each zip code
```

- opensecrets columns were already processed and written to the csv


```{r,eval=FALSE}
# change N/As to blanks to be consistent with how occupation and employer is recorded for orgs - FOR ORGS ONLY (occupation & employer both N/A)
#x[which(x$employer == "N/A" & x$occupation == "N/A"),]$standardized_employer_name <- ""
#x[which(x$employer == "N/A" & x$occupation == "N/A"),]$processed_employer_name <- ""

#x[which(x$employer == "N/A" & x$occupation == "N/A"),]$standardized_occupation <- ""
#x[which(x$employer == "N/A" & x$occupation == "N/A"),]$processed_occupation <- ""

# Update employer name to 'Unknown' for individuals where employer is N/A, unless occupation specifies that they are self employed. In this case, the employer becomes 'self employed.'
#x <- x %>% 
#  mutate(standardized_employer_name = if_else((employer == "N/A" & x$occupation != "N/A"), if_else(standardized_occupation == "SELF EMPLOYED","SELF EMPLOYED","UNKNOWN"),processed_employer_name),
#         processed_employer_name = if_else((employer == "N/A" & x$occupation != "N/A"), if_else(standardized_occupation == "SELF EMPLOYED","SELF EMPLOYED","UNKNOWN"),processed_employer_name))

# check processing
#x[which(x$employer == "N/A" & x$occupation != "N/A"),]$standardized_employer_name #<- "UNKNOWN"
#x[which(x$employer == "N/A" & x$occupation != "N/A"),]$processed_employer_name #<- "UNKNOWN"
```


# Separate organizations from individuals and add row id

We only want to match orgs using this method. individuals will be matches using a different method. 

```{r}
# create row ids
x <- x %>% mutate(row_id = row_number())
ref <- ref %>% mutate(ref_id = row_number())

orgs <- x %>% filter(entity_type == "organization")
individuals <- x %>% filter(entity_type == "individual")
```

# Match organizations to OpenSecrets data


### look at matches from different processing of names


```{r}
# add name variants for both datasets
make_name_variants <- function(df, col) {
  df %>%
    mutate(
      name_v1 = standardize_names(.data[[col]]) %>% fix_typos(),
      name_v2 = remove_pac_ref(name_v1)          %>% str_squish(),
      name_v3 = remove_unit_ref(name_v2)          %>% str_squish(),
      #name_v4 = replace_committee_text(name_v3)    %>% str_squish(),
      name_v5 = replace_business_text(name_v3)     %>% str_squish(),  # = processed_name
      name_v6 = str_remove_all(name_v5, "\\s*\\(.*?\\)") %>% str_squish()  # + remove parens
    )
}

# add name variants
x_vars   <- make_name_variants(orgs,  "canonical_name") 
ref_vars <- make_name_variants(ref, "name")      

# look for exact matches using different processing of the name, then try fuzzy matching on what's left
run_exact <- function(pool, ref, col, label) {
  ref_slim <- ref %>% select(ref_id, all_of(col), level1_category, level2_category,level3_category)

  all_candidates <- inner_join(pool, ref_slim,
                               by = setNames(col, col),
                               relationship = "many-to-many") %>%
    mutate(match_method = label,
           matched_name = .data[[col]]) %>%
    group_by(row_id) %>%
    mutate(n_distinct_cats = n_distinct(level3_category, na.rm = TRUE),
           needs_review    = n_distinct_cats > 1) %>%
    ungroup()

  matched <- all_candidates %>%
    group_by(row_id) %>%
    slice_head(n = 1) %>%
    ungroup()

  list(matched = matched, all_candidates = all_candidates,
       remaining = pool %>% filter(!row_id %in% matched$row_id))
}


variants <- c("name_v1", "name_v2", "name_v3", "name_v5", "name_v6") # removed name_v4

all_matches     <- list()
all_candidates  <- list()
remaining       <- x_vars

for (v in variants) {
  result          <- run_exact(remaining, ref_vars, v, paste0("exact_", v))
  all_matches     <- append(all_matches, list(result$matched))
  all_candidates  <- append(all_candidates, list(result$all_candidates))
  remaining       <- result$remaining
}

candidates_df <- bind_rows(all_candidates)

ref_slim_fuzzy <- ref_vars %>% select(ref_id, name_v5, level1_category, level2_category,level3_category)

fuzzy_candidates <- stringdist_inner_join(
  remaining, ref_slim_fuzzy,
  by = c("name_v5" = "name_v5"),
  method = "jw", p = 0.1, max_dist = 0.08,
  distance_col = "jw_dist"
) %>%
  rename(matched_name = name_v5.y) %>%
  select(-name_v5.x) %>%
  rename(name_v5 = matched_name) %>%
  mutate(match_method = "fuzzy_jw") %>%
  group_by(row_id) %>%
  mutate(n_distinct_cats = n_distinct(level3_category, na.rm = TRUE),
         needs_review    = n_distinct_cats > 1) %>%
  ungroup()

fuzzy_matches <- fuzzy_candidates %>%
  group_by(row_id) %>%
  slice_min(jw_dist, n = 1, with_ties = FALSE) %>%
  ungroup()
```



### token jaccard similarity (alternate matching process to catch other matches)

```{r}
token_jaccard <- function(a, b) {
  tok_a <- str_split(str_squish(a), "\\s+")[[1]]
  tok_b <- str_split(str_squish(b), "\\s+")[[1]]
  n_intersect <- length(intersect(tok_a, tok_b))
  n_union     <- length(union(tok_a, tok_b))
  if (n_union == 0) return(0)
  n_intersect / n_union
}

# block on first token before cross-joining for Jaccard (otherwise takes forever to run)
remaining <- remaining %>% mutate(first_token = str_extract(name_v5, "^\\S+"))
ref_vars  <- ref_vars  %>% mutate(first_token = str_extract(name_v5, "^\\S+"))

ref_slim_jaccard <- ref_vars %>% select(ref_id, first_token, name_v5, level1_category, level2_category,level3_category)

jaccard_candidates <- inner_join(remaining, ref_slim_jaccard, by = "first_token") %>%
  rename(matched_name = name_v5.y) %>%
  rowwise() %>%
  mutate(jaccard = token_jaccard(name_v5.x, matched_name)) %>%
  ungroup() %>%
  filter(jaccard >= 0.6) %>%
  mutate(match_method = "fuzzy_jaccard") %>%
  group_by(row_id) %>%
  mutate(n_distinct_cats = n_distinct(level3_category, na.rm = TRUE),
         needs_review    = n_distinct_cats > 1) %>%
  ungroup()

jaccard_matches <- jaccard_candidates %>%
  group_by(row_id) %>%
  slice_max(jaccard, n = 1, with_ties = FALSE) %>%
  ungroup()


```

```{r}
# de-deuplicate when choosing final matches
final_matches <- bind_rows(
  bind_rows(all_matches) %>% mutate(method_priority = 1),
  fuzzy_matches          %>% mutate(method_priority = 2),
  jaccard_matches        %>% mutate(method_priority = 3)
) %>%
  group_by(row_id) %>%
  slice_min(method_priority, n = 1, with_ties = FALSE) %>%
  ungroup()

conflicted_ids <- final_matches %>% filter(needs_review) %>% pull(row_id)

all_candidates_combined <- bind_rows(
  bind_rows(all_candidates),
  fuzzy_candidates,
  jaccard_candidates
)
```


### review matches

```{r}
# match open secrets name to review for accuracy
final_matches <- final_matches %>%
  left_join(
    ref_vars %>% select(ref_id, name),  
    by = "ref_id"
  )

all_candidates_combined <- all_candidates_combined %>%
  left_join(
    ref_vars %>% select(ref_id, name),
    by = "ref_id"
  )
```

- why is matched_name missing for fuzzy_jw?

```{r}
final_matches %>% select(canonical_name,name,match_method,level1_category,)
```

### join matches back to original dataset


```{r}
x_matched <- x %>%
  left_join(
    final_matches %>%
      select(row_id, level1_category, level2_category,level3_category, match_method, needs_review, n_distinct_cats),
    by = "row_id"
  ) %>%
  mutate(matched = !is.na(level3_category))

```


```{r}
nrow(x)
nrow(x_matched)
nrow(x_matched[which(!is.na(x_matched$level1_category)),])
```



# Match individuals to OpenSecrets data

- Need to match individuals based on more than their name. Incorporate occupation, employer, zip code and city into matching process.

### parse names into first, middle, last + suffixes

```{r}
parsed <- parse_names_df(individuals$processed_name)
individuals <- individuals %>%
  mutate(
    last_name     = parsed$last,
    first_name    = parsed$first,
    middle_name   = parsed$middle,
    name_prefix   = parsed$title,
    name_suffix   = parsed$suffix,
    first_initial = substr(first_name, 1, 1)
  )

# parse additional ref data
parsed_ref <- parse_names_df(ref$name)
ref <- ref %>%
  mutate(
    last_name     = parsed_ref$last,
    first_name    = parsed_ref$first,
    middle_name   = parsed_ref$middle,
    name_prefix   = parsed_ref$title,
    name_suffix   = parsed_ref$suffix,
    first_initial = substr(first_name, 1, 1),
  )
```

```{r,eval=FALSE}
# check output of parsing function
individuals[which(individuals$individual_tag == TRUE & !is.na(individuals$middle_name)),c("standardized_name","last_name","first_name","middle_name","name_suffix","name_prefix","first_initial")]
#ref[which(!is.na(ref$middle_name)),c("standardized_name","last_name","first_name","middle_name","name_suffix","name_prefix","first_initial")]
```


### add metaphone key for blocking

```{r}
# deal with special characters in the names when computing metaphone
metaphone_func <- function(x) {
  clean <- str_replace_all(coalesce(x, ""), "[^A-Za-z]", "")
  ifelse(clean != "", metaphone(clean), NA_character_)
}

individuals <- individuals %>%
  mutate(
    last_meta = ifelse(
      !is.na(last_name) & last_name != "",
      metaphone_func(last_name),
      NA_character_
    )
  )

ref <- ref %>%
  mutate(
    last_meta = ifelse(
      !is.na(last_name) & last_name != "",
      metaphone_func(last_name),
      NA_character_
    )
  )
```

```{r}
which(is.na(ref$last_meta))
which(is.na(individuals$last_meta))
```


### deal with ids

```{r}
individuals <- individuals %>%
  rename(x_row_id = row_id) %>% # preserve this to match back to original data
  mutate(row_id = row_number())

ref <- ref %>%
  mutate(row_id = row_number())  # positional key for compare_and_score joins
```


### employer boilerplate stripping

Tokens that appear in >1% of employer strings across both datasets (e.g. "INC", "GROUP", "HOSPITAL") are too common to distinguish individuals. Strip them before cosine similarity so the score reflects the distinctive part of the employer name.

```{r}
all_emp_tokens <- c(
  str_split(coalesce(individuals$processed_employer_name, ""), "\\s+") %>% unlist(),
  str_split(coalesce(ref$processed_employer_name,         ""), "\\s+") %>% unlist()
)
n_emp_docs       <- nrow(individuals) + nrow(ref)
emp_token_counts <- table(all_emp_tokens)
BOILERPLATE_EMP_SHARE <- 0.01
boilerplate_emp_words <- names(emp_token_counts)[
  emp_token_counts > n_emp_docs * BOILERPLATE_EMP_SHARE & names(emp_token_counts) != ""
]

strip_boilerplate_emp <- function(name) {
  vapply(str_split(coalesce(name, ""), "\\s+"), function(toks) {
    kept <- toks[!toks %in% boilerplate_emp_words]
    paste(if (length(kept) == 0) toks else kept, collapse = " ")
  }, character(1))
}

individuals <- individuals %>% mutate(core_employer = strip_boilerplate_emp(processed_employer_name))
ref         <- ref         %>% mutate(core_employer = strip_boilerplate_emp(processed_employer_name))

cat("Boilerplate employer tokens (", length(boilerplate_emp_words), "):\n")
cat(paste(sort(boilerplate_emp_words), collapse = ", "), "\n")
```


## Record linkage function

- Function should handle cross dataset comparison and also detecting the same person within a dataset (but not applying this yet)

```{r}
bigram_cosine <- function(a, b) stringsim(a, b, method = "cosine", q = 2) # should help handle variation in word order
exact_match <- function(a, b) as.numeric(a == b & !is.na(a) & !is.na(b))


uninformative_vals <- c("UNKNOWN", "RETIRED", "NONE", "SELF EMPLOYED")

compare_and_score <- function(pairs_obj, df_a, df_b = NULL) {
  if (is.null(df_b)) df_b <- df_a

  # Uses core_employer (boilerplate-stripped) for cosine similarity so that
  # generic tokens like INC/GROUP/HOSPITAL don't inflate employer match scores.
  # The raw processed_employer_name is still joined in to check uninformativeness.
  pairs_obj %>%
    compare_pairs(
      on = c("last_name", "first_name", "zip_code_processed",
             "core_employer", "processed_occupation", "standardized_city"),
      default_comparator = cmp_jarowinkler(),
      comparators = list(
        core_employer        = function(a, b) stringsim(a, b, method = "cosine", q = 2),
        zip_code_processed   = cmp_identical(),
        processed_occupation = cmp_identical()
      )
    ) %>%
    as_tibble() %>%
    rename(employer_sim = core_employer) %>%
    left_join(
      df_a %>% select(row_id, processed_employer_name, processed_occupation) %>%
        rename(emp_a = processed_employer_name, occ_a = processed_occupation),
      by = c(".x" = "row_id")
    ) %>%
    left_join(
      df_b %>% select(row_id, processed_employer_name, processed_occupation) %>%
        rename(emp_b = processed_employer_name, occ_b = processed_occupation),
      by = c(".y" = "row_id")
    ) %>%
    mutate(
      emp_a_uninf = coalesce(emp_a, "") %in% uninformative_vals | is.na(emp_a) | emp_a == "",
      emp_b_uninf = coalesce(emp_b, "") %in% uninformative_vals | is.na(emp_b) | emp_b == "",
      occ_a_uninf = coalesce(occ_a, "") %in% uninformative_vals | is.na(occ_a) | occ_a == "",
      occ_b_uninf = coalesce(occ_b, "") %in% uninformative_vals | is.na(occ_b) | occ_b == "",
      w_emp   = if_else(emp_a_uninf & emp_b_uninf, 0, 0.25),
      w_occ   = if_else(occ_a_uninf & occ_b_uninf, 0, 0.10),
      w_total = 0.30 + 0.25 + w_emp + w_occ + 0.07 + 0.03,
      # zero out employer/occupation similarity when either side is uninformative
      emp_sim = if_else(emp_a_uninf | emp_b_uninf, 0, coalesce(employer_sim, 0)),
      occ_sim = if_else(occ_a_uninf | occ_b_uninf, 0, coalesce(processed_occupation, 0)),
      score   = (0.30 * coalesce(last_name, 0) +
                 0.25 * coalesce(first_name, 0) +
                 w_emp * emp_sim +
                 w_occ * occ_sim +
                 0.07 * coalesce(zip_code_processed, 0) +
                 0.03 * coalesce(standardized_city, 0)) / w_total,
      both_uninformative = (emp_a_uninf & emp_b_uninf) & (occ_a_uninf & occ_b_uninf)
    ) %>%
    # keep uninf flags + emp_sim so downstream can apply hard constraints and confidence
    select(-emp_a, -emp_b, -occ_a, -occ_b, -w_emp, -w_occ, -w_total)
}

```


### compute similarity scores


```{r}
### blocking for efficiey
pairs_cross_zip <- pair_blocking(individuals, ref,
                                 on = c("zip_code_processed", "last_meta"))
pairs_cross_init <- pair_blocking(individuals, ref,
                                   on = c("last_meta", "first_initial"))

# pass in both datasets to the compare and score dataset
scored_cross_zip   <- compare_and_score(pairs_cross_zip,   individuals, ref)
scored_cross_init  <- compare_and_score(pairs_cross_init,  individuals, ref)

pairs_cross <- bind_rows(scored_cross_zip, scored_cross_init) %>%
  group_by(.x, .y) %>%
  slice_max(score, n = 1, with_ties = FALSE) %>%
  ungroup()

# Hard employer constraint + confidence scoring.
# employer_conflict = TRUE when both sides have a real (informative) employer
# but those employers are too dissimilar (< EMP_SIM_MIN). These pairs are
# rejected regardless of name/location score — employer mismatch is a hard signal.
EMP_SIM_MIN <- 0.60

pairs_cross <- pairs_cross %>%
  mutate(
    employer_conflict = !emp_a_uninf & !emp_b_uninf & coalesce(emp_sim, 0) < EMP_SIM_MIN,
    confidence = case_when(
      employer_conflict                                             ~ "low",
      both_uninformative                                           ~ "low",
      emp_a_uninf | emp_b_uninf | occ_a_uninf | occ_b_uninf       ~ "medium",
      TRUE                                                         ~ "high"
    )
  )

```


```{r}
FIRST_NAME_MIN   <- 0.85
MATCH_THRESHOLD  <- 0.90
STRICT_THRESHOLD <- 0.94
UNCERTAIN_LOW    <- 0.70
STRICT_UNCERTAIN <- 0.75

matches_cross <- pairs_cross %>%
  filter(
    coalesce(first_name, 0) >= FIRST_NAME_MIN,
    !employer_conflict,                          # hard reject: different real employers
    (!both_uninformative & score >= MATCH_THRESHOLD) |
    ( both_uninformative & score >= STRICT_THRESHOLD)
  )

uncertain_cross <- pairs_cross %>%
  filter(
    coalesce(first_name, 0) >= FIRST_NAME_MIN,
    !employer_conflict,
    (!both_uninformative & score >= UNCERTAIN_LOW & score < MATCH_THRESHOLD) |
    ( both_uninformative & score >= STRICT_UNCERTAIN & score < STRICT_THRESHOLD)
  )

```


### look at matches 

```{r}
inspect_pairs <- function(pairs_df, df_a, df_b) {
  pairs_df %>%
    left_join(
      df_a %>%
        select(row_id, processed_name, processed_employer_name,
               processed_occupation, standardized_city, zip_code_processed) %>%
        rename(name_a     = processed_name,
               employer_a = processed_employer_name,
               occ_a      = processed_occupation,
               city_a     = standardized_city,
               zip_a      = zip_code_processed),
      by = c(".x" = "row_id")
    ) %>%
    left_join(
      df_b %>%
        select(row_id, processed_name, processed_employer_name,
               processed_occupation, standardized_city, zip_code_processed) %>%
        rename(name_b     = processed_name,
               employer_b = processed_employer_name,
               occ_b      = processed_occupation,
               city_b     = standardized_city,
               zip_b      = zip_code_processed),
      by = c(".y" = "row_id")
    ) %>%
    select(score, 
           name_a, name_b, employer_a, employer_b,
           occ_a, occ_b, city_a, city_b, zip_a, zip_b) %>%
    arrange(desc(score))
}

```

```{r}
#inspect_pairs(matches_cross, individuals, ref)
inspect_pairs(uncertain_cross, individuals, ref) # uncertain matches
```


- After looking at these, choosing a threshold of 0.75 for the current data. We may have to update this threshold later.

```{r}
# apply threshold
matches_cross_final <- pairs_cross %>%
  filter(
    coalesce(first_name, 0) >= FIRST_NAME_MIN,
    !employer_conflict,
    score >= 0.75
  ) %>%
  add_count(`.x`, name = "n_matches")

#matches_cross_final %>%
#  count(n_matches) %>%
#  mutate(match_type = if_else(n_matches == 1, "unique match", "multiple matches"))

# add 'ref' variables to the match candidates before grouping
matches_with_ref <- matches_cross_final %>%
  left_join(
    ref %>%
      select(row_id, open_secrets_contributor_id, level1_category,level2_category,level3_category, processed_name), # look for contradictions based on level3
    by = c(".y" = "row_id")
  )

# flag disagreement and keep best score
best_matches <- matches_with_ref %>%
  group_by(`.x`) %>%
  mutate(
    n_distinct_ids  = n_distinct(open_secrets_contributor_id, na.rm = TRUE),
    n_distinct_cats = n_distinct(level3_category, na.rm = TRUE)
  ) %>%
  slice_max(score, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(match_status = case_when(
    n_matches == 1                                              ~ "unique match",
    n_matches > 1 & n_distinct_ids <= 1 & n_distinct_cats <= 1 ~ "multiple matches - consistent",
    TRUE                                                        ~ "multiple matches - review"
  ),
  # downgrade confidence to "low" when multiple conflicting matches exist
  confidence = if_else(match_status == "multiple matches - review", "low", confidence))

```


### match back to data

```{r}
# dataset with matches
individuals_linked <- individuals %>%
  left_join(
    best_matches %>%
      select(`.x`, `.y`, score, confidence, n_matches, match_status,
             open_secrets_contributor_id, level1_category, level2_category, level3_category),
    by = c("row_id" = ".x")
  ) %>%
  rename(matched_ref_row      = `.y`,
         match_score          = score,
         opensecrets_confidence = confidence,
         n_ref_matches        = n_matches)

```


```{r}
# look at conflicts - match back candidates from the 'ref' dataset.
individuals_linked %>%
  filter(match_status == "multiple matches - review") %>%
  left_join(
    matches_with_ref %>%
      select(`.x`, `.y`, score, open_secrets_contributor_id, level1_category,level2_category,level3_category, processed_name),
    by = c("row_id" = ".x")
  ) %>%
  arrange(processed_name.x, desc(score)) %>%
  select(processed_name.x, score,
        open_secrets_contributor_id.y, level3_category.x,
          processed_name.y)   # ref name for comparison

```


```{r}
# join back to x using x_row_id
x_final <- x_matched %>%
  left_join(
    individuals_linked %>%
      select(x_row_id, level1_category, level2_category, level3_category,
             match_status, match_score, opensecrets_confidence),
    by = c("row_id" = "x_row_id")
  ) %>%
  mutate(
    level1_category = coalesce(level1_category.x, level1_category.y),
    level2_category = coalesce(level2_category.x, level2_category.y),
    level3_category = coalesce(level3_category.x, level3_category.y)
  ) %>%
  select(-ends_with(".x"), -ends_with(".y"))

#x_final <- x_matched %>%
#  left_join(
#    individuals_linked %>% select(x_row_id, open_secrets_contributor_id, level1_category,level2_category,level3_category, match_status, match_score),
#    by = c("row_id" = "x_row_id")
#  )
```


```{r}
# respect upstream classification when available
# could think about using this as an alternative sources instead and look for agreement
x_final <- x_final %>%
  mutate(
    # only use OpenSecrets result where 0803 produced nothing
    level1_category = if_else(!is.na(data_source_1), level1_from_0803, level1_category),
    level2_category = if_else(!is.na(data_source_1), level2_from_0803, level2_category),
    level3_category = if_else(!is.na(data_source_1), level3_from_0803, level3_category),
    data_source_final = coalesce(data_source_1, if_else(!is.na(match_score), "opensecrets", NA_character_))
  )

```



```{r}
x_final[which(x_final$processed_employer_name == "SELF EMPLOYED"),]
```


```{r}
unique(x_final[which(x_final$individual_tag == TRUE & !is.na(x_final$match_score)),])
```


```{r}
nrow(x_final[which(!is.na(x_final$level1_category)),])
nrow(x_final[which(!is.na(x_final$level1_category)),]) / nrow(x_final)
```

```{r}
nrow(x_final[which(!is.na(x_final$level1_category)),])
nrow(x_final[which(!is.na(x_final$level1_category)),]) / nrow(x_final)
```

- Matching about 14% of rows currently.


### map OpenSecrets categories to NAICS codes

Moved here from `10_assign_classifications/assign_final_classification.Rmd` --
this crosswalk is specific to OpenSecrets' own level3 taxonomy, so it belongs
with the OpenSecrets matching step rather than the final classification step.
Downstream consumers get `naics_code_os` for free instead of re-deriving it
from `level3_category` themselves.

```{r}
# read map: OpenSecrets level3 category -> custom NAICS code
crosswalk <- read.csv("NAICs, OpenSecrets Crosswalk - NAICS2_OpenSecretsLevel3.csv")
colnames(crosswalk)[1:2] <- c("level3_category", "naics_code_os")

# sub out special characters
crosswalk$naics_code_os <- gsub("\\*", "", crosswalk$naics_code_os)

x_final <- x_final %>%
  left_join(crosswalk[, c("level3_category", "naics_code_os")], by = "level3_category")
```

```{r}
unique(x_final[, c("level3_category", "naics_code_os")])
```


### write matches to csv

```{r}
write.csv(x_final,"08_outputs/contributions_os_matches.csv",row.names = FALSE)
```

