---
title: "R Notebook"
output: html_notebook
---

# TO DOS
- update so orgs are not aggregated by city
- some PACs in the power search data may be missing a committee id; should fill in where we can


# Read in packages and data

```{r}
library(dplyr)
library(readr)
library(stringr)

source("standardization_helpers.R")
```


```{r}
# data from datasette
url <- "https://calmatters-powersearch-2026.fly.dev/powersearch/contributions.csv?_stream=on&_size=max"
contributions_full <- read.csv(url)
```


```{r}
# data to map contributor ids
id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv")
# get unique xref id / filer ids combos
id_map <- unique(id_map_full[,c("XREF_FILER_ID","FILER_ID","FILER_TYPE")])

# contributor data (will need to update with output from data processing file)
x <- read.csv("08_inputs/power_search_governors_data_5k_071326.csv")
```

filername_cd.tsv info: https://calaccess.californiacivicdata.org/documentation/raw-files/filername-cd/ 
- city, state, zip, phone, fax, email can differ across same filer id
- recipient committee id corresponds to FILER_ID, contributor.id corresponds to XREF_ID
- Map FILER_ID to x based on Contributor.ID matching xref_filer_id

```{r,eval = FALSE}
id_map_full %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(NAML, FILER_ID, XREF_FILER_ID) %>% filter(grepl("POLITICAL", NAML)) %>% distinct()
```

Those i manually searched come up as major donor under entity type in cal-access
- USAA GROUP POLITICAL ACTION COMMITTEE comes up as major donor in cal-accesss
- COALITION FOR NATURAL HEALTH POLITICAL ACTION COMMITTEE, INC., THE: also major donor in cal access, says terminated on FEC website
- COUNTRYWIDE HOME LOANS, INC.; COUNTRYWIDE POLITICAL ACTION COMMITTEE (FEDERAL PAC) -> can't find on FEC even though name says federal PAC
  
# Clean up data


```{r}
# convert all ids to character strings
id_map$FILER_ID         <- as.character(id_map$FILER_ID)   
id_map$XREF_FILER_ID <- as.character(id_map$XREF_FILER_ID)
x$Contributor.ID <- as.character(x$Contributor.ID)
contributions_full$Recipient.Committee.ID <- as.character(contributions_full$Recipient.Committee.ID)
contributions_full$Contributor.ID <- as.character(contributions_full$Contributor.ID)
```

```{r}
x %>% filter(grepl("foster",tolower(Contributor.Name)))
```


```{r}
x <- x[which(!is.na(x$Contributor.ID)),] # filter to only include committees

# map filer id to contributor id (different ids may be used internally vs externally)
x <- left_join(x, id_map, by = c("Contributor.ID" = "XREF_FILER_ID"))
#x %>% group_by(uuid) %>% summarise(count = n()) %>% filter(count > 1)

# check that each has a corresponding filer id
nrow(x[which(is.na(x$FILER_ID)),])
nrow(x[which(x$FILER_ID == ""),])
```


**PACs missing a committee id**
* Border Health Federal PAC and Laborer's Union Local No. 777 PAC don't have contributor ids and they should...
* International Brotherhood of Electrical Workers Political Action Committee also doesn't have a contributor id and well need to fill that in
* UA 290 Industry Advancement Fund
* there are probably other ones that we'll need to fill in as well
  * Border Health is from Texas


### look at type of contributors

```{r}
unique(x$FILER_TYPE)
major_donor_ids <- x %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(FILER_ID)
major_donor_ids <- major_donor_ids$FILER_ID
#length(which(!(major_donor_ids %in% contributions$Recipient.Committee.ID))) / length(major_donor_ids)
x %>% filter(FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE") %>% select(FILER_ID, Contributor.Name)
```

- There are 6 contributors that match to Major Donor / Independent expenditure committees
- None of those that are major donors / independent exenditures match to received contributions
- Looked in Cal Access and Youth Save Democracy PAC is considered a major donor, not a recipient committee
- same with other 5 committees. 


### take major donors out of consideration

- will update upstream processing so these are not fed into this notebook

```{r}
x <- x[-which(x$FILER_TYPE == "MAJOR DONOR/INDEPENDENT EXPENDITURE COMMITTEE"),]
```



### standardize contributor info

```{r}
contributions <- contributions_full %>%
  mutate(
    standardized_name = standardize_names(Contributor.Name) %>% fix_typos() %>% str_squish(),
    #processed_name = standardized_name %>%
    #  remove_pac_info() %>% remove_unit_info() %>% replace_business_text() %>%
    #  str_remove_all("\\s*\\(.*?\\)") %>% str_squish(),
    standardized_employer_name = standardize_names(Contributor.Employer) %>% fix_typos() %>% str_squish(),
    processed_employer_name = standardized_employer_name %>% str_squish(),
    standardized_occupation = standardize_names(Contributor.Occupation),
    processed_occupation    = standardize_occupation_employer(standardized_occupation),
    standardized_city       = standardize_names(Contributor.City),
    standardized_state = standardize_names(Contributor.State) %>% str_squish(),
    zip_code_processed      = substr(Contributor.Zip.Code, 1, 5),
  
    entity_type      = if_else(sapply(standardized_name, is_individual), "individual", "organization"),
    row_id           = row_number()
  )

# manually fix weird states
contributions <- contributions %>% mutate(
  standardized_state = if_else(standardized_state == "93" & standardized_city == "BAKERSFIELD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "C" & standardized_city == "LOS ANGELES", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "MJ" & standardized_city == "PARSIPPANY", "NJ",standardized_state),
  standardized_state = if_else(standardized_state == "A" & standardized_city == "FOLSOM", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "AC" & standardized_city == "GLENDALE", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "US" & standardized_city == "HERMITAGE", "TN",standardized_state),
  
  standardized_state = if_else(standardized_state == "" & standardized_city == "LOS ANGELES", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "SACRAMENTO", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "INGLEWOOD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "HAYWARD", "CA",standardized_state),
  standardized_state = if_else(standardized_state == "" & standardized_city == "PASADENA", "CA",standardized_state),
)
```

- many contributions missing a city and/or state


# Aggregate contributions by contributor

**update to aggregate individuals by employer/occupation/city, but not companies/orgs**

```{r}
contributions_agg <- contributions %>%
  mutate(
    has_id       = !is.na(Contributor.ID) & Contributor.ID != "",
    is_fed_pac   = grepl("fed pac|federal pac|#c|id#|fec id|fec report|fed form", tolower(Contributor.Name)),
    is_pac       = has_id & !is_fed_pac,
    is_pac_no_id = !has_id & !is_fed_pac &
                   grepl("\\bPAC\\b|COMMITTEE|POLITICAL ACTION", toupper(Contributor.Name))
  ) %>%
  left_join(
    id_map %>% select(XREF_FILER_ID, FILER_ID, FILER_TYPE) %>% 
      rename(std_contributor_id = FILER_ID), 
    by = c("Contributor.ID" = "XREF_FILER_ID")
  ) %>%
  group_by(
    Recipient.Committee.ID,
    Contributor.ID,
    std_contributor_id,
    standardized_name,
    processed_employer_name,
    processed_occupation,
    standardized_city,
    standardized_state,
    is_pac, is_fed_pac, is_pac_no_id, FILER_TYPE 
  ) %>%
  summarise(Amount = sum(Amount, na.rm = TRUE), .groups = "drop")

```



# CREATE PAC GRAPH

### helper functions

```{r}
get_pac_contributors <- function(pac_id, contributions_agg) {
  contributions_agg %>%
    filter(Recipient.Committee.ID == pac_id)
}

```

```{r}
categorize_contributor <- function(name, employer, amount) {
  name_upper <- toupper(name)
  employer_upper <- toupper(employer)
  case_when(
    name_upper == "UNITEMIZED CONTRIBUTIONS" ~ "unitemized",
    amount <= 100 ~ "unitemized",
    grepl("UNION MEMBER|MEMBER CONTRIBUTIONS|MEMBER DUES|LESS THAN \\$?100|UNDER \\$?100|CONTRIBUTIONS UNDER|UNITEMIZED DUES|LABORERS DUES|LABORER DUES|MEMBERS DUES|
          INTERMEDIARY FOR MEMBERS|LABORERS TRUST",name_upper) ~ "member_dues",
    grepl("UNION MEMBER|LABORERS|DISTRICT COUNCIL|EMPLOYEES|LOCAL\\s+\\d+|UNION\\s+\\d+|SMW\\s+\\d+|IBEW\\s+\\d+|SEIU\\s+\\d+|UFCW\\s+\\d+|OLCU\\s+\\d+|UNITE HERE\\s+\\d+|AFSCME\\s+\\d+|IRONWORKER\\s+\\d+|UPTE-CWA\\s+\\d+",employer_upper) ~ "member_dues", # anything that has local unit number + other union indicators
    amount < 5000   ~ "under_5k",
    TRUE  ~ "over_5k" 
  )
}
```

- could update to flag medium contributions (< $100)
- contributions <= $100 called unitemized, but should be referred to as small dollar.


```{r}
pac_total_donated <- contributions %>%
  filter(Contributor.ID != "" & !is.na(Contributor.ID)) %>%
  left_join(
    id_map %>% select(XREF_FILER_ID, FILER_ID),
    by = c("Contributor.ID" = "XREF_FILER_ID")
  ) %>%
  mutate(
    unmatched_contributor_id = is.na(FILER_ID), # something wrong with contributor id
    std_contributor_id       = as.character(FILER_ID)
  ) %>%
  group_by(Contributor.ID, std_contributor_id, unmatched_contributor_id) %>%
  summarise(total_donated_all = sum(Amount, na.rm = TRUE), .groups = "drop")

```

- Need to check how much the PAC has donated vs how much it has received


# PAC GRAPH TO CONSIDER $5k THRESHOLD: 


```{r}
DATA_SUFFICIENCY_THRESHOLD <- 0.75

traverse_pac_graph <- function(seed_ids, ref, id_map, max_depth = 5) {
  queue         <- lapply(seed_ids, function(id) list(id = id, depth = 0, path = id))
  visited       <- character(0)
  pac_summaries <- list()
  over_5k_contribs <- list()

  while (length(queue) > 0) {
    node   <- queue[[1]]; queue <- queue[-1]
    pac_id <- node$id; depth <- node$depth; path <- node$path

    if (pac_id %in% visited || depth >= max_depth) {
      if (depth >= max_depth && !(pac_id %in% visited)) {
        pac_summaries <- append(pac_summaries, list(tibble(
          pac_id = pac_id, depth = depth, path = path, status = "max_depth_reached",
          total_received = NA_real_, unitemized_amount = NA_real_,
          member_dues_amount = NA_real_, under_5k_amount = NA_real_, over_5k_amount = NA_real_
        )))
      }
      next
    }
    visited <- c(visited, pac_id)
    
    contribs <- get_pac_contributors(pac_id, ref)

    if (is.null(contribs) || nrow(contribs) == 0) {
      pac_summaries <- append(pac_summaries, list(tibble(
        pac_id = pac_id, depth = depth, path = path, status = "unresolved",
        total_received = NA_real_, total_donated_all = NA_real_,
        data_sufficiency = NA_real_,
        unitemized_amount = NA_real_, member_dues_amount = NA_real_,
        under_5k_amount = NA_real_, over_5k_amount = NA_real_
      )))
      next
    }
    
    # check data sufficiency 
    total_received    <- sum(contribs$Amount, na.rm = TRUE)
    total_donated_all <- pac_total_donated %>%
      filter(!unmatched_contributor_id, std_contributor_id == pac_id) %>%
      pull(total_donated_all)
    total_donated_all <- if (length(total_donated_all) == 0) NA_real_ else total_donated_all
    data_sufficiency  <- if (!is.na(total_donated_all) && total_donated_all > 0)
                           total_received / total_donated_all
                         else NA_real_

    pac_status <- case_when(
      !is.na(data_sufficiency) & data_sufficiency < DATA_SUFFICIENCY_THRESHOLD ~ "low_data_sufficiency",
      is.na(data_sufficiency)                                                   ~ "unknown_data_sufficiency",
      TRUE                                                                      ~ "resolved"
    )

    contribs <- contribs %>%
      mutate(contributor_category = categorize_contributor(standardized_name, processed_employer_name, Amount))
    
    # summarize PAC based on what kind of donations it received
    pac_summaries <- append(pac_summaries, list(tibble(
      pac_id             = pac_id,
      depth              = depth,
      path               = path,
      status             = "resolved",
      total_received     = sum(contribs$Amount, na.rm = TRUE), 
      total_donated_all = total_donated_all,
      data_sufficiency  = total_received / total_donated_all,   # NA if total_donated_all == 0
      unitemized_amount  = sum(contribs$Amount[contribs$contributor_category == "unitemized"],  na.rm = TRUE),
      member_dues_amount = sum(contribs$Amount[contribs$contributor_category == "member_dues"], na.rm = TRUE),
      under_5k_amount    = sum(contribs$Amount[contribs$contributor_category == "under_5k"],    na.rm = TRUE),
      over_5k_amount     = sum(contribs$Amount[contribs$contributor_category == "over_5k"],     na.rm = TRUE)
    )))


    # only process $5k+ contributors further
    for (i in seq_len(nrow(contribs %>% filter(contributor_category == "over_5k")))) {
      row <- (contribs %>% filter(contributor_category == "over_5k"))[i, ]

      if (isTRUE(row$is_fed_pac)) {
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id      = pac_id,
          path               = path,
          depth              = depth + 1,
          contributor_name   = row$standardized_name, # do i want original name here for matching purposes? Contributor.Name
          contributor_id     = row$Contributor.ID,
          contributor_type   = "federal_pac",
          contributor_amount = row$Amount
        )))

      } else if (isTRUE(row$FILER_TYPE == "RECIPIENT COMMITTEE")) {
        new_id   <- if (!is.na(row$std_contributor_id)) row$std_contributor_id else row$Contributor.ID
        matched  <- !is.na(row$std_contributor_id)
        new_path <- paste0(path, " -> ", new_id)
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id      = pac_id,
          path               = new_path,
          depth              = depth + 1,
          contributor_name   = row$standardized_name, # Contributor.Name?
          contributor_id     = new_id,
          contributor_type   = if (!matched) "pac_unmatched_id" else if (isTRUE(row$is_pac_no_id)) "pac_no_id" else "pac",
          contributor_amount = row$Amount
        )))
        if (!is.na(new_id)) {
          queue <- append(queue, list(list(id = new_id, depth = depth + 1, path = new_path)))
        }

      } else {
        # individual or org — record for industry classification
        over_5k_contribs <- append(over_5k_contribs, list(tibble(
          parent_pac_id          = pac_id,
          path                   = path,
          depth                  = depth + 1,
          contributor_name       = row$standardized_name, #row$Contributor.Name,
          contributor_id         = NA_character_, # update this, they may have a major donor id
          contributor_type       = "individual_or_org",
          contributor_amount     = row$Amount,
          contributor_employer   = row$processed_employer_name,
          contributor_occupation = row$processed_occupation,
          contributor_city       = row$standardized_city,
          contributor_state      = row$standardized_state
        )))
      }
    }
  }

  list(
    pac_summaries    = bind_rows(pac_summaries),
    over_5k_contributors = bind_rows(over_5k_contribs)
  )
}

```

- What do do w/ client, slate mailer orgs, payment to influnce and NA?
- only need recipient committee? 


```{r}
# run PAC graph
seed_ids <- unique(x$FILER_ID)
seed_ids <- seed_ids[!is.na(seed_ids) & seed_ids != ""]

result           <- traverse_pac_graph(seed_ids, contributions_agg, id_map, max_depth = 5)
pac_summaries    <- result$pac_summaries
over_5k_contributors <- result$over_5k_contributors
```



```{r}
# add indicator for federal PAC into PAC summaries
fed_seed_ids <- x %>%
  filter(grepl("fed pac|federal pac|#c|id#|fec id|fec report|fed form", tolower(Contributor.Name))) %>%
  pull(FILER_ID) %>%
  unique()

pac_summaries <- pac_summaries %>%
  mutate(is_federal_pac = pac_id %in% fed_seed_ids)

```

- Resolved just means the PAC was found (should change tthat language)


### Add indicators for whether we are able to classify each PAC in theory

```{r}
UNITEMIZED_DUES_THRESHOLD <- 0.5
OVER_5K_COVERAGE_THRESHOLD <- 0.5

pac_classifiability <- pac_summaries %>%
  mutate(
    low_data_sufficiency     = !is.na(data_sufficiency) & data_sufficiency < DATA_SUFFICIENCY_THRESHOLD,
    unknown_data_sufficiency = is.na(data_sufficiency),
    data_ok                  = status == "resolved" & !low_data_sufficiency & !unknown_data_sufficiency,
    pct_unitemized_dues      = (unitemized_amount + member_dues_amount) / total_received,
    pct_over_5k              = over_5k_amount / total_received,
    classify_by_name         = !is.na(pct_unitemized_dues) & pct_unitemized_dues >= UNITEMIZED_DUES_THRESHOLD & data_ok,
    data_ok                  = status == "resolved" & !is_federal_pac & !low_data_sufficiency & !unknown_data_sufficiency
  )

```

- This is only a preliminary indicator for whether we can classify that PAC, we need to see whether or not we can actually classify its contributors.

### flag each $5k contributor as resolvable or not

```{r}
# look at the status of each contributor.
over_5k_tagged <- over_5k_contributors %>%
  left_join(
    pac_classifiability %>% select(pac_id, classify_by_name, data_ok, pct_over_5k, status),
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    is_resolvable = case_when(
      contributor_type == "individual_or_org"                                        ~ TRUE,
      contributor_type == "pac" & classify_by_name                                   ~ TRUE,
      contributor_type == "pac" & data_ok & pct_over_5k >= OVER_5K_COVERAGE_THRESHOLD ~ TRUE,
      TRUE                                                                           ~ FALSE
    ),
    unresolvable_reason = case_when(
      !is_resolvable & contributor_type == "federal_pac"                             ~ "federal_pac",
      !is_resolvable & contributor_type %in% c("pac_no_id", "pac_unmatched_id")     ~ contributor_type,
      !is_resolvable & contributor_type == "pac" & !data_ok                         ~ paste0("pac_", status),
      !is_resolvable & contributor_type == "pac" & data_ok                          ~ "pac_low_5k_coverage",
      TRUE                                                                           ~ NA_character_
    )
  )

```

```{r}
nrow(over_5k_contributors)
nrow(over_5k_tagged)
```

### compute % of each PAC contributors that can be classified 

```{r}
# try to build processing order: a PAC can only be processed after all its PAC contributors
processed <- character(0)
ordered_pacs <- character(0)

pacs_to_order <- pac_summaries %>% arrange(desc(depth)) %>% pull(pac_id) 

while (length(pacs_to_order) > 0) {
  for (pid in pacs_to_order) {
    pac_contributors <- over_5k_tagged %>%
      filter(parent_pac_id == pid, contributor_type == "pac") %>%
      pull(contributor_id) %>%
      na.omit()
    
    if (all(pac_contributors %in% processed)) {
      ordered_pacs <- c(ordered_pacs, pid)
      processed    <- c(processed, pid)
      pacs_to_order <- setdiff(pacs_to_order, pid)
    }
  }
}

# compute what fraction of its total_received is resolvable
pac_resolvable_fractions <- pac_classifiability %>%
  mutate(resolvable_fraction = NA_real_)

for (pid in ordered_pacs) { 
  summary_row <- pac_classifiability %>% filter(pac_id == pid)
  
  if (summary_row$classify_by_name) { 
    pac_resolvable_fractions <- pac_resolvable_fractions %>%
      mutate(resolvable_fraction = if_else(pac_id == pid, 1, resolvable_fraction))
    next
  }
  
  if (!summary_row$data_ok) {
    pac_resolvable_fractions <- pac_resolvable_fractions %>%
      mutate(resolvable_fraction = if_else(pac_id == pid, 0, resolvable_fraction))
    next
  }
  
  direct_contribs <- over_5k_tagged %>%
    filter(parent_pac_id == pid)
  
  if (nrow(direct_contribs) == 0) {
    pac_resolvable_fractions <- pac_resolvable_fractions %>%
      mutate(resolvable_fraction = if_else(pac_id == pid, 0, resolvable_fraction))
    next
  }
  
  direct_contribs <- direct_contribs %>%
    left_join(
      pac_resolvable_fractions %>% select(pac_id, child_resolvable_fraction = resolvable_fraction),
      by = c("contributor_id" = "pac_id")
    ) %>%
    mutate(
      resolvable_amount = case_when(
        contributor_type == "individual_or_org" ~ contributor_amount,
        contributor_type == "pac"               ~ contributor_amount * coalesce(child_resolvable_fraction, 0),
        TRUE                                    ~ 0
      )
    )
  
  frac <- sum(direct_contribs$resolvable_amount, na.rm = TRUE) / summary_row$total_received
  
  pac_resolvable_fractions <- pac_resolvable_fractions %>%
    mutate(resolvable_fraction = if_else(pac_id == pid, frac, resolvable_fraction))
}

```


**look at order of PACs associated with multiple depths to make sure its working**


### coverage breakdown for each seed PAC

```{r}
seed_coverage <- over_5k_tagged %>%
  filter(depth == 1, parent_pac_id %in% seed_ids) %>%
  left_join(
    pac_resolvable_fractions %>% select(pac_id, resolvable_fraction),
    by = c("contributor_id" = "pac_id")
  ) %>%
  mutate(
    resolvable_amount = case_when(
      contributor_type == "individual_or_org" ~ contributor_amount,
      contributor_type == "pac"               ~ contributor_amount * coalesce(resolvable_fraction, 0),
      TRUE                                    ~ 0
    )
  ) %>%
  group_by(parent_pac_id) %>%
  summarise(
    total_direct_over_5k  = sum(contributor_amount, na.rm = TRUE),
    total_resolvable      = sum(resolvable_amount, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    pac_classifiability %>% filter(pac_id %in% seed_ids) %>%
      select(pac_id, total_received, classify_by_name, data_ok, status, data_sufficiency),
    by = c("parent_pac_id" = "pac_id")
  ) %>%
  mutate(
    pct_resolvable      = total_resolvable / total_received,
    classification_status = case_when(
      classify_by_name                              ~ "classify_by_name",
      data_ok & pct_resolvable >= OVER_5K_COVERAGE_THRESHOLD ~ "classify_by_contributors",
      TRUE                                          ~ "unresolvable"
    )
  ) 

```

```{r}
pac_names <- x %>% group_by(FILER_ID) %>% slice(1) %>% ungroup() %>% select(FILER_ID, Contributor.Name)
seed_coverage <- seed_coverage %>% left_join(pac_names, by = c("parent_pac_id" = "FILER_ID"))
```


```{r}
# get summary for all seed pacs
all_seed_status <- pac_classifiability %>%
  filter(pac_id %in% seed_ids) %>%
  left_join(
    seed_coverage %>%
      select(parent_pac_id, total_direct_over_5k, total_resolvable, pct_resolvable, classification_status),
    by = c("pac_id" = "parent_pac_id")
  ) %>%
  mutate(
    classification_status = case_when(
      !is.na(classification_status)                                   ~ classification_status,
      classify_by_name                                                ~ "classify_by_name",
      status %in% c("unresolved", "low_data_sufficiency",
                    "unknown_data_sufficiency")                       ~ "unresolvable",
      TRUE                                                            ~ "unresolvable"
    ),
    pct_resolvable       = coalesce(pct_resolvable, 0),
    total_direct_over_5k = coalesce(total_direct_over_5k, 0),
    total_resolvable     = coalesce(total_resolvable, 0),
    unresolvable_reason  = case_when(
      classification_status != "unresolvable"  ~ NA_character_,
      is_federal_pac    ~ "federal_pac",                 
      status == "unresolved"  ~ "no_contributor_data",
      low_data_sufficiency                                ~ "low_data_sufficieny", 
      unknown_data_sufficiency                           ~ "unknown_data_sufficiency", 
      total_direct_over_5k == 0                                      ~ "no_over_5k_contributors",
      pct_resolvable < OVER_5K_COVERAGE_THRESHOLD                    ~ "low_pct_resolvable", 
      TRUE                                                            ~ "other"
    )
  )


all_seed_status <- all_seed_status %>% left_join(pac_names, by = c("pac_id" = "FILER_ID"))
```

```{r}
all_seed_status %>% filter(unresolvable_reason == "no_contributor_data") #%>% count()
all_seed_status %>% filter(unresolvable_reason == "low_data_sufficieny") #%>% count()
all_seed_status %>% filter(unresolvable_reason == "low_pct_resolvable") #%>% count()
all_seed_status %>% filter(unresolvable_reason == "no_over_5k_contributors")# %>% count()
all_seed_status %>% filter(unresolvable_reason == "federal_pac")# %>% count()
```

```{r}
all_seed_status %>% filter(unresolvable_reason == "federal_pac")
```

```{r}
all_seed_status %>% filter(classification_status == "classify_by_contributors")
```


```{r}
table(all_seed_status$unresolvable_reason)
```

- 4 federal PACs (thought there were 6?)
- 42 PACs have contributed much more than they received 
- 14 PACs have <50% of contributions that came from $5k+
- 31 PACs have no contributor data in 2026
- 1 PAC has no contributions over $5k (all contributions greater than 100, but less than 5k - this is a dairies PAC, all its contributors are dairies)
- 16 PACs (10%) have more than 50% of their contributions come from $5k+ donations and have contributors we are able to classify (as opposed to unresolvable PACs)
  - For many of these, we'd still only be classifying around 60% of its received contributions
  

```{r}
#write.csv(all_seed_status,"08_inputs/classifiability_by_pac.csv")
```


# Generate list of contributors to classify

```{r}
# PAC name lookup: seed PACs from filing data, intermediate PACs from graph traversal
pac_name_lookup <- bind_rows( 
  pac_names %>% rename(pac_id = FILER_ID, canonical_name = Contributor.Name),
  over_5k_tagged %>%
    filter(contributor_type %in% c("pac", "federal_pac")) %>%
    select(pac_id = contributor_id, canonical_name = contributor_name) %>%
    filter(!is.na(pac_id))
) %>%
  distinct(pac_id, .keep_all = TRUE)  # seed entry takes precedence

# list 1: PACs to classify by name (any PAC in the graph where classify_by_name = TRUE)
pacs_for_classification <- pac_classifiability %>%
  filter(classify_by_name) %>%
  left_join(pac_name_lookup, by = "pac_id") %>%
  transmute(
    contributor_id = pac_id,
    canonical_name,
    entity_type    = "organization",
    employer       = NA_character_,
    occupation     = NA_character_,
    city           = NA_character_,
    state          = NA_character_,
    total_amount   = total_received
  )

# list 2: all $5k+ individuals and orgs across the graph
# recover Contributor.ID (not preserved in BFS traversal) via join back to contributions_agg
direct_for_classification <- over_5k_tagged %>%
  filter(contributor_type == "individual_or_org") %>%
  left_join(
    contributions_agg %>%
      select(Recipient.Committee.ID, Contributor.ID, standardized_name,
             processed_employer_name, processed_occupation,
             standardized_city, standardized_state) %>%
      distinct(),
    by = c("parent_pac_id"          = "Recipient.Committee.ID",
           "contributor_name"       = "standardized_name",
           "contributor_employer"   = "processed_employer_name",
           "contributor_occupation" = "processed_occupation",
           "contributor_city"       = "standardized_city",
           "contributor_state"      = "standardized_state")
  ) %>%
  mutate(entity_type = if_else(sapply(contributor_name, is_individual),
                               "individual", "organization")) %>%
  group_by(contributor_name, entity_type, contributor_employer,
           contributor_occupation, contributor_city, contributor_state) %>%
  summarise(
    contributor_id = Contributor.ID[which(!is.na(Contributor.ID))[1]],
    total_amount   = sum(contributor_amount, na.rm = TRUE),
    .groups        = "drop"
  ) %>%
  transmute(
    contributor_id,
    canonical_name = contributor_name,
    entity_type,
    employer       = contributor_employer,
    occupation     = contributor_occupation,
    city           = contributor_city,
    state          = contributor_state,
    total_amount
  )

classification_list <- bind_rows(pacs_for_classification, direct_for_classification)

cat("PACs to classify by name:      ", nrow(pacs_for_classification), "\n")
cat("Individuals/orgs to classify:  ", nrow(direct_for_classification), "\n")
cat("Total:                         ", nrow(classification_list), "\n")

write.csv(classification_list, "08_inputs/classifier_input_from_pac_graph.csv", row.names = FALSE)
```




