---
title: "R Notebook"
output: html_notebook
---

```{r}
source("standardization_helpers.R")
```


```{r}
contributions_full
```


# Resolve same name, different individuals

Output ALL demographic info
- not sure I want to output the recipient committee tbh -> that would be a lot
- Could leave in Recipient committee (id or name?) just to flag which ones I've resolved. Look at ones relevant to me at least to start...


Don't forget to incorporate manual feedback: many of these people likely gave to DIFFERENT committees though, so may not need to resolve, but would be nice to have different ids for them...


```{r}
# make name list: generate output of people in THIS data with any of these names and flag manually

name_list <- c("SMITH, STEVE","DAVIS, JOHN","KING, JUSTIN","SPENCER, RICHARD","FISHER, ROBERT", "ROBERTSTON, CARL", "ROBERTSON, CHIP", "W ROBERTSON, CARL")
```


```{r}
contribs %>% filter(grepl("ROBERTSON, CHIP",toupper(Contributor.Name))) %>% arrange(Recipient.Committee,Contributor.Name) %>% select(Recipient.Committee,Contributor.Name,Contributor.Employer, Contributor.Occupation, Contributor.City,Contributor.State,Contributor.Zip.Code) %>% distinct()
```

```{r}
contribs %>% filter(grepl("DAVIS, JOHN",toupper(Contributor.Name))) %>% arrange(Recipient.Committee,Contributor.Name) %>% select(Recipient.Committee,Contributor.Name,Contributor.Employer, Contributor.Occupation, Contributor.City,Contributor.State,Contributor.Zip.Code) %>% distinct()
```

- Think I can ignore ActBlue as a recipient.committee - don't think we use that
- What committees can I leave out? 


- How to know how to aggregate contributors when there is MORE THAN ONE row for one individual


# NAME CHECK to fill in missing ids


```{r}
# read in id map
# data to map contributor ids
id_map_full <- read_tsv("08_inputs/FILERNAME_CD.tsv")
id_map <- unique(id_map_full[,c("XREF_FILER_ID","FILER_ID","FILER_TYPE")]) # get unique xref id / filer ids combos

# convert all ids to character strings
id_map$FILER_ID         <- as.character(id_map$FILER_ID)   
id_map$XREF_FILER_ID <- as.character(id_map$XREF_FILER_ID)
```



```{r}
# read in most up-to-date data
c <- read.csv("08_inputs/power_search_contributions_normalized.csv")
```

```{r}
nrow(c)
nrow(contributions_full)
```



```{r}
#c %>% filter(toupper(Contributor.Name) == "FRONTWAVE CREDIT UNION (INDEPENDENT EXPENDITURE ACCOUNT)")
#c %>% filter(grepl("lockheed martin",tolower(Contributor.Name))) %>% select(Contributor.Name, Contributor.ID, Contributor.City, Contributor.Zip.Code,Recipient.Committee, Recipient.Committee.ID) %>% distinct()
c %>% filter(grepl("international brotherhood of electrical workers",tolower(Contributor.Name))) %>% select(Contributor.Name, Contributor.ID, Contributor.City, Contributor.Zip.Code,Recipient.Committee, Recipient.Committee.ID) %>% distinct()
```


```{r}
c %>% filter()
```


Johnson & JOhnson
#C00010983


AbbVie Political Action Committee (Fed PAC ID C00536573)
ABBVIE PAC

```{r}
c %>% filter(Recipient.Committee.ID == "880969")
```


```{r}
c %>% filter(Recipient.Committee.ID == 1045835 & Recipient.Committee == Contributor.Name)
```



```{r}
c %>% filter(Contributor.State == "DC") %>% distinct(Contributor.City)
```

```{r}
c %>% filter(Contributor.City == "Washington") %>% distinct(Contributor.State)
```

```{r}
unique(c$Contributor.State)
```


880969
Should we flag this id more generally? 


DRIVE Committee
DRIVE Committee (FEC ID# C00032979)
DRIVE - Democrat, Republican, Independent Voter Education (The PAC of the International Brotherhood of Teamsters)
Drive- Democrat, Republican, Independent Voter Education
DRIVE - Dem, Rep, Ind Voter Education (The PAC of the Int'l Brhd of Teamsters)

```{r}
id_map_full %>% filter(grepl("DENTAQUEST",toupper(NAML))) %>% select(NAML,FILER_ID,XREF_FILER_ID,FILER_TYPE,STATUS,EFFECT_DT,NAMF) %>% distinct()
```


1049814	900161

```{r}
id_map_full %>% filter(FILER_ID == 880969 | XREF_FILER_ID == 880969)
```


```{r}
head(id_map_full)
head(id_map)
```

# LOOK AT EMPLOYERS in 

```{r}
contribs <- c %>% 
  rename(
    Ballot.Measure = `Ballot.Measure.s.`
  ) %>% 
  filter(
    Ballot.Measure == "OPPOSED: PROPOSITION 040 - IMPOSES ONE-TIME TAX ON CERTAIN INDIVIDUALS AND TRUSTS. INITIATIVE CONSTITUTIONAL AMENDMENT AND STATUTE." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 040 - IMPOSES ONE-TIME TAX ON CERTAIN INDIVIDUALS AND TRUSTS. INITIATIVE CONSTITUTIONAL AMENDMENT AND STATUTE." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 041 - REQUIRES AUDITS OF PROGRAMS FUNDED BY NEW STATE SPECIAL TAXES. PROHIBITS NEW STATE TAXES THAT ARE EXCLUDED FROM EXISTING VOTER-APPROVED STATE SPENDING LIMIT..." | 
    Ballot.Measure == "SUPPORTED: PROPOSITION 042 - PROHIBITS NEW STATE PERSONAL PROPERTY TAXES AND CERTAIN RETROACTIVE STATE TAXES. INITIATIVE CONSTITUTIONAL AMENDMENT." | 
    Recipient.Name == "BECERRA, XAVIER" | 
    Recipient.Name == "HILTON, STEVE" | 
    (Recipient.Name == "ALLEN, BEN" & Office == "Insurance Commissioner") | 
    Recipient.Name == "KIM, JANE" 
)
nrow(contribs)
```


```{r}
contribs %>% filter(entity_type == "organization") %>% distinct(Contributor.Occupation)
```


C00076810
C00219642
# C00128512
C00168864
ID# C00836866



# FLAG ORGS FUNCTION

- look at cases with an employer that were flagged as orgs
- look at cases without an employer that were flagged as individuals

```{r}
contributions %>% filter(grepl("JV",toupper(Contributor.Name)))
```


**issues**
- no good way to resolve individuals without a comma in their name


### apply updated function for comparison

```{r}
c <- contributions_full
```

```{r}
c <- contributions
nrow(c)
```



```{r}
c <- c %>% 
  mutate(entity_type_alt = if_else(
      mapply(is_individual, standardized_name, `Contributor.Employer`, `Contributor.Occupation`),
      "individual", "organization"),
      entity_type_ambiguous = entity_type_looks_ambiguous(entity_type, `Contributor.Employer`, `Contributor.Occupation`)
  )
```

```{r}
c %>% filter(entity_type_alt != entity_type) %>% select(Contributor.Name, entity_type_alt, Contributor.Employer, Contributor.Occupation, Contributor.City) %>% distinct()
```

**last round remaining errors (to address manually):**

Last0b6df6bd, First7606aee3
Lastcbea8f37, Firstd599d312
Associates, Sonny
Drilling, Dr. Patricia
Phan, The
Labor, Wilmot


- de los angeles last name
- vineyard last name
- Associates, Sonny -> confused about this
- Garcia, Sacramento
- Drilling, Dr. Patricia
- Phan, The
- Labor, Wilmot

more keywords: 
- motel
- would be good to figure out 'inn'
- inn & spa
- inn at


actually confused about what the centers are... 

- Nguyen, Dr. Loan
- Phan, Thuy-Loan
- Luu, Dr. Loan
- Huynh, Loan
 - remove loan as keyword I guess? 
- CENTER, SCOTT
- Center, Jay
- Center, Lynda
  - center as keyword, but has non-empty employer
- Drilling, Dr. Patricia
- Phan, The
  - dentist, outliter
- Manchester, including aggregated contributions and independent expenditures, Craig
  - don't know what this means

- Holdenried, Brent and Debra: now flagged as org even though theyre individuals, emplyer and occ empty though... 
- CENTER, LYNDA
- CENTER, ED
- CONGRESS, MILES
- CONGRESS, PATRICIA
  - seems like a last name? 
-STAHL FAMILY, THE
  - this and some other cases where ,the at the end doesn't work
  - but helpful in LOTS of cases
  - could put this in ambiguous and have employer decide
- CALLAWAY, THE
  - not sure what this is... 
- COUNCIL, SANDRA L
- RANCH, KELLEY
- HOENIGMAN FAMILY, THE
- PIMENTEL, CCTV OUTLET CORP

- Centene Corp, HealthNet: now and individual!!!

```{r}
c %>% filter(grepl("LABOR, ", toupper(Contributor.Name))) %>% select(Contributor.Name, Contributor.Employer) %>% distinct()
```

- can we not include "labor, "
- 

keywords to add: 
- associates
- business
- AFL-CIO
- investment
- professional
- DBA surrounded by spaces? 
- Corp at the end instead? 
- accountacy 
- centene
- healthnet (health?)

### look at remaining ambiguous cases

```{r}
unique(c$entity_type_ambiguous)
c %>% filter(entity_type_ambiguous == TRUE) %>% select(Contributor.Name, entity_type_alt, Contributor.Employer, Contributor.Occupation) %>% distinct()
```
```{r}
is_individual("BERRY & SONS, RICHARD")
```


```{r}
#contributions %>% filter(Contributor.Name == "Barr Co., Bill")
contributions %>% filter(grepl("INN AT",toupper(Contributor.Name))) %>% select(Contributor.Name, Contributor.Employer,Contributor.Occupation,Contributor.City) %>% distinct()
```

**remaining errors (to address manually)**
Goldberg & Loren, PA
Alice Valder Curran
LA3pez, Carlos
Bosch, Thermador, Gaggenau
Dr. B. Kamkar, A Prof Opto
Joseph Marchionna & Mike Delapena, OD's
B;izzard, Deb
Eric Michael Rudnick
Berry & Sons, Richard
Barr Co., Bill


UNCERTAIN Cases
- Arthur Friedman, O.D.
- Keith Lesar, Attorney at Law
- Joseph Metz, Broker
- Lisa B Stephenson, Realtor

- not sure what to do w/ family trusts
- all brokers now flagged as indivdiuals -> not sure what to do with them
  - could flag by looking for ,broker at tthe end
  
remaining errors: 
- HCSS - Sugarland, TX -> flagged as individual
- Goldberg & Loren, PA -> should be an org
- Rush Truck Center, Pico Rivera
- Alice Valder Curran -> not sure what to do w/ individuals w/out a comma
- LA3pez, Carlos -> not sure why flagged as an org? 
- Danilo & Laureta Cabanayan
- An York, dba Dr Lee Acupuncture -> should be an org? 
- Becker, Todd B Law Offices of
- Bosch, Thermador, Gaggenau -> flagged as individual
- Avi Kram
- Mai Lor
- Palmetto Digestive Disease, PA
- LOCKER, RICHARD 5
- The Ritz-Carlton, San Francisco
- B;izzard, Deb
- Gless - Caliente, John J. -> flagged as individual, but others are flagged orgs
  - Gless - P13 Ranch, John J. and Gless - Sunshine Ranch, John J.
- Meadowood, Napa Valley	individual			
- Meier Citrus, Steve -> looks like this is his name and not a citrus grower
- Berry & Sons, Richard -> should be org
- Poon and Son, P.A.
- Barr Co., Bill
- T-Mobile, USA

keywrods
- law offices & law office
- Optometrists
- neurology
- co. 
- change to vineyards

errors: 
- Young Vineyards, Robert
- Democrats in Sun City, Palm Desert
- Penny for Progress, Yes on Measure E
- Lifetime Green Coatings, Sacramento
- California Alliance for Progress and Education, an alliance of business organizations
E. Bruce Laycook : still an org

- update keywords
- run on contributions only
- decide what to do w/ ambiguous keywords like ', the' , 'center' and 'congress'
  - also look at centers specifically. 

keywords to add: 
- alliance
- business
- organization
- vineyards
- democrat
- republican 
- yes on measure
- attorney at law? not sure what to do with these if they are law offices or not
- sacramento, los angeles, california
- lifetime? coatings? 

### earlier pass

- look at contributor id - could be another indicator

```{r}
c %>% filter(Contributor.Name == "American Federation of State County & Municipal Employees, AFL-CIO")
```

American Federation of State County & Municipal Employees, AFL-CIO: is this a pac? should it have a contributor id? wouldn't impact how we classify I dont think



```{r}
contributions %>% filter(entity_type == "organization") %>% distinct(Contributor.Employer)
```
No Separate Business Name, Committee

```{r}
contributions %>% filter(Contributor.Employer == "Committee")
```


```{r}
contributions %>% filter(entity_type == "organization" & (Contributor.Employer != "" & Contributor.Employer != "." & Contributor.Employer != "N/A" )) %>% select(Contributor.Name,Contributor.Employer, Contributor.Occupation) %>% distinct()
```

```{r}
contributions %>% filter(entity_type == "individual" & (Contributor.Employer == "" | Contributor.Employer == ".")) %>% select(Contributor.Name,Contributor.Employer, Contributor.Occupation) %>% distinct()
```

keywords to add:
- resort
- coalition
- government
- sponsored
- pediatrics
- expenditures
- usa (eg t0mobile usa)
- joint venture
- insurance
- visit 
- farms
- institute
- center
- chapter
- federation
- employee
- industry
- conference
- council
- drilling
- ranch? / ranches
- buildings
- laboratory
- pawn shop
- realty
- ALC (a law corporation)
- chamber
- DBA
- office 
- law office
- medical
- congress
- loan
- jewelry
- attorney at law?
- apartment / apartments
- hotel 
- resort
- optometrists
- optometry
- dairy
- international
- sponsored
- ,the
- inn / spa
- properties
- residences
- maufacturing
 - aggregated
 - contributors

Orgs flagged as individuals: 
Powering California's Future, Sponsored by Edison International
Medici at South Hills, The
Fairmont Miramar Hotel, The
The Waterfront Beach Resort, a Hilton Hotel
A Cash Buyer Jewelry, Watch and Loan
- Drs. Holden, Knapp & Langsford AKA Family Optometry -> not sure what to do with this one
- American Federation of State County & Municipal Employees, AFL-CIO
- California Legislative Conference of the Plumbing, Heating & Piping Industry
- CA Legislative Conference of the Plumbing, Heating & Piping Industry - Admin
- Pearlman, Borska & Wax
- California Teamsters Public Affairs Council, Public Affairs
- Mahaffey Drilling Co., Don H.
- Campagnoli, Abelson & Campagnoli -> law firm or couple? 
- American Federation of State, County and Municipal Employees, AFL-CIO
- Borchard-DJB Ranch, David
- Brucker Ranch Co., Mike
- Road Ranch, Fairview
-HCSS - Sugarland, TX  
- Last, Faoro & Whitehorn -> law office
- Stacy and Witbeck/Herzog, a JV
- Eason & Tambornini, ALC
- Redwoods Retirement Community, The
- Karen Tabler, Broker -> unclear whether to flag these as orgs or individuals anyways...
- Hills of Corona - Afton Properties, The
- Mackaig Realty, R A
- Row at Redhill - Greystar, The
- Vilara - Greystar, The
- Southern Sierras Chapter, NECA
- USW 474 G Aluminum, Brick & Gl (intermediary for unitemized member dues)
- Hooper, Lundy & Bookman
Air-Conditioning, Heating & Refrigeration Institute				
Su, Huachang CA Medical Therapeutic Center
An York, dba Dr Lee Acupuncture
Becker, Todd B Law Offices of

Individuals flagged a org: 
- B;izzard, Deb
- Long Family Trust
- Corp, Craig
- E. Bruce Laycook	
- Alice Valder Curran
- CC\003B!rdenas, Jeffrey	; Democratic Party of Orange County
- BARBARA BRASS, PAT THOMPSON &
- LA3pez, Carlos ;	Werner
- Danilo & Laureta Cabanayan
- SPIEKER, JR., INCLUDING SPIEKER REALTY INVESTMENTS, WARREN E.
- Colatruglio, CFA, CFP, Mr. Frank ; wells fargo advisors
- Hone, Md, Ma, Face, Jennifer ;	Optum
- VLASAK P.C., BRACAMONTES	BRACAMONTES & VLASAK PC
- MCMORRIN COMMITTEE, YASMINE	CULVER CITY
- Assemblymember Schultz
- The Pera Nieh Trust (Charlotte & David)
- Dawniel Garcia
- Joseph Musquiz
- Stephen Brazeel
- Schreiber, LLP, Singleton ;	Singleton Schreiber Law Firm	
- Zaidal Obagi
- Trowbridge and affiliated entity Auburn manor Holding Corporation, Donald T.	Western Care Construction Company, Inc. -> weird case generally
- Avi Kram
- Heather Mulholland
- Mai Lor
- HOWARD, EDDIE&ALI
- G. Lowe Chang
- Quillin (Meadow Fund), Patty
- LOCKER, RICHARD 5
- YEE, EDISON& JANICE
- PETO~LEVICKI, SHEILA
- BENNETT, DAOM, L.AC, JOANN W.
SCHREEDER, MARIANNE (USE THIS A
YEE, EDISON& JANICE	NOT EMPLOYED	
TALIFER, T & N CHICAGO INC.	NOT EMPLOYED	
WIRELESS, 2X	2X WIRELESS	
ASKENASE, PHIL 2016	YALE U	
BENNETT, DAOM, L.AC, JOANN W.	SELF	
PETO~LEVICKI, SHEILA	GB	
SCHREEDER, MARIANNE (USE THIS A	NOT EMPLOYED	
SCHWARTZ, LISA &
MARTIN, ELAINE &EUGENE
HARTFORD, DAN&ELLEN	NOT EMPLOYED	
FIELDS, CHESTER& MARION
ABRONS, MD, MPH, HENRY
ABSEY, RN, HN-BC, CMT, JENNIFER
GILBERT, PHD, LMFT, KRISTA
Murphy~Bryner, Jayne
Peto~Levicki, Sheila


### check added names

1. no duplciate raw names
2. single uuid per canonical name
3. same # unique uuids and canonical names
4. one canonical name per raw name
5. no double spaces in canonical names
5. what else? 

```{r}
new_names <- read.csv("08_inputs/name_entity_mapping_rebuilt_to_add.csv")
nrow(new_names)
```

```{r}
# no duplicate raw names
new_names %>% group_by(raw_name) %>% filter(n() > 1) %>% arrange(raw_name)
```

- all other cases are those that have two canonical names per raw_name


```{r}
# one uuid per raw name
new_names %>% group_by(raw_name) %>% filter(n_distinct(canonical_name) > 1) %>% arrange(raw_name)
```

- these are the same 12 we already knew about.

```{r}
# one uuid per canonical
new_names %>% group_by(canonical_name) %>% filter(n_distinct(entity_uuid) > 1) %>% arrange(canonical_name)
```

```{r}
# no double spaces in canonical name
new_names %>% filter(grepl("  ",canonical_name))
new_names %>% filter(grepl("  ",raw_name))
```

```{r}
length(unique(new_names$canonical_name))
length(unique(new_names$entity_uuid))
```

```{r}
existing_names <- new_names %>% filter(source != "manual_confirmation")
added_names    <- new_names %>% filter(source == "manual_confirmation")
```

```{r}
# leading/trailing whitespace on raw_name or canonical_name
existing_names %>%
  filter(raw_name != trimws(raw_name) | canonical_name != trimws(canonical_name)) %>%
  select(raw_name, canonical_name, source)
```

```{r}
# leading/trailing whitespace on raw_name or canonical_name
added_names %>%
  filter(raw_name != trimws(raw_name) | canonical_name != trimws(canonical_name)) %>%
  select(raw_name, canonical_name, source)
```

```{r}
# new raw names already in existing map with a DIFFERENT canonical name (silent override)
added_names %>%
  inner_join(
    existing_names %>% select(raw_name, existing_canonical = canonical_name),
    by = "raw_name"
  ) %>%
  filter(canonical_name != existing_canonical) %>%
  select(raw_name, canonical_name, existing_canonical)
```

```{r}
# new canonical names that already exist in the map — UUID must match
added_names %>%
  inner_join(
    existing_names %>% distinct(canonical_name, existing_uuid = entity_uuid),
    by = "canonical_name"
  ) %>%
  filter(entity_uuid != existing_uuid) %>%
  select(raw_name, canonical_name, entity_uuid, existing_uuid)
```

```{r}
# raw_name == canonical_name (no-op entries) -> these are okay
added_names %>%
  filter(raw_name == canonical_name) %>%
  select(raw_name, canonical_name)
```

```{r}
# UUID format check
uuid_pattern <- "^[0-9a-f]{8}-[0-9a-f]{4}-[0-9a-f]{4}-[0-9a-f]{4}-[0-9a-f]{12}$"
added_names %>%
  filter(!grepl(uuid_pattern, entity_uuid, ignore.case = TRUE)) %>%
  select(raw_name, canonical_name, entity_uuid)
```

















