---
title: "R Notebook"
output: html_notebook
---

# Purpose
Configure descriptions for each custom NAICS code for the LLM to reference when assigning codes.

# TO DOs
- any other opensecrets categories we'd need to add in ?

```{r}
library(readxl)
library(dplyr)
library(stringr)
```

### Add more information to the titles

```{r}
codes <- read_excel("09_inputs/digit_2022_Codes.xlsx")

# remove blank rows and columns
codes <- codes[-1,]
codes <- codes[,-c(4,5)]

colnames(codes)[c(2,3)] <- c("Code","Title")

```

The codes with a dash got left out - weren't grabbed in the 2-digit filter...

```{r}
# expand two digit codes with a dash...will mess up sequence no

codes[which(codes$Code == "31-33"),]$Code <- "31"
codes <- rbind(codes,c(NA,"32","Manufacturing"))
codes <- rbind(codes,c(NA,"33","Manufacturing"))

codes[which(codes$Code == "44-45"),]$Code <- "44"
codes <- rbind(codes,c(NA,"45","Retail Trade"))

codes[which(codes$Code == "48-49"),]$Code <- "48"
codes <- rbind(codes,c(NA,"49","Transportation and Warehousing"))

tail(codes)
```


```{r}
sector_expanded <- codes %>%
  filter(nchar(Code) == 2) %>%
  mutate(expanded_title = 
    sapply(Code, function(code) {
      parent_title <- codes %>%
        filter(Code == code) %>%
        pull(Title)
      
      sub_titles <- codes %>%
        filter(str_starts(Code, code), nchar(Code) == 4) %>%
        distinct(Title) %>%
        pull(Title)
      
      if (length(sub_titles) == 0) return(parent_title)
      
      all_titles <- c(parent_title, sub_titles)
    
      paste0(parent_title, ", including: ", paste(sub_titles, collapse = "; "))
    })
  )

```


### Collapse descriptions for combined categories: 

40: Combines 53 and 23
50: Combines 54, 55 and 56a
60: Combines 62, 61 and lirbaries and archives
70: Combines 71 + parts of 51
76: Remove association description from relvant naics code? 
77: Remove assocition description from relevant naics code
80: 44-45, 42, 92, parts of 81
92: 92 + defense contractors, take out tribes

```{r}
# helper function to combine sector descriptions
combine_sectors <- function(new_code, new_title, source_codes = character(0), extra_subtitles = character(0)) {
  if (length(source_codes) > 0) {
    sub_titles <- codes %>%
      filter(sapply(Code, function(c) any(str_starts(c, source_codes))) & nchar(Code) == 4) %>%
      distinct(Title) %>%
      pull(Title)
  } else {
    sub_titles <- character(0)
  }

  all_subs <- c(sub_titles, extra_subtitles)

  description <- if (length(all_subs) == 0) {
    new_title
  } else {
    paste0(new_title, ", including: ", paste(all_subs, collapse = "; "))
  }

  data.frame(naics_sector = new_code, description = description, stringsAsFactors = FALSE)
}

```


```{r}
#x[which(x$naics_sector == "52"),]$description
codes %>% filter(grepl("^56",Code) & nchar(Code) == 4)
```

Natural gas distribution goes in oil and gas.... not in 22
Natural gas includes natural gas distribution? 

```{r}
x <- sector_expanded[,c("Code","expanded_title"),]
colnames(x) <- c("naics_sector","description")

x <- rbind(x,c("99","Unknown"))
x <- rbind(x,c("90","Candidates/Political Parties"))
x <- rbind(x,c("88","Political Actions Committees (PACs), contributor committees and political action funds"))
#x <- rbind(x,c("78","Defense contractors"))
#x <- rbind(x,c("79","Technology, software and AI companies"))
x <- rbind(x,c("76","Business, professional and trade associations"))
x <- rbind(x,c("77","Unions"))
x <- rbind(x,c("92a","Native American tribes"))
x <- rbind(x,c("100","Retired/Homemaker/Student/Unemployed"))
#x <- rbind(x,c("01","Candidate contributions to their own campaigns."))

# remove source codes that get merged into combined categories
x <- x %>% filter(!naics_sector %in% c("22","23", "53","42","44","45","48","49","51","53","54","55","56","61","62","71","72","81"))

# add combined categories
x <- rbind(x, combine_sectors("20", "Utilities, infrastructure, telecommunications, renewable energy, transportation and warehousing", 
                              source_codes = c("22","48","49"),
                              extra_subtitles = c("Wired and Wireless Telecommunications (except Satellite)",
                                                  "Satellite Telecommunications","All Other Telecommunications",
                                                  "Waste Collection","Waste Treatment and Disposal","Remediation and Other Waste Management Services")),
              combine_sectors("40", "Construction/Real Estate", source_codes = c("23", "53")),
              combine_sectors("50", "Professional/Legal", source_codes = c("54", "55"),
                              extra_subtitles = c("Management and Holding Companies",
                                                  "Office Administrative Services",
                                                  "Employment Services",
                                                  "Facilities Support Services",
                                                  "Business Support Services",
                                                  "Travel Arrangement and Reservation Services",
                                                  "Investigation and Security Services",
                                                  "Services to Buildings and Dwellings")),
              combine_sectors("60", "Healthcare/Education", source_codes = c("61", "62"),
                              extra_subtitles = c("Libraries and archives")),
              combine_sectors("70", "Entertainment, Media and Gambling", source_codes = c("71"),
                  extra_subtitles = c("Motion Picture and Video Industries", "Sound Recording Industries","Newspaper, Periodical, Book, and Directory Publishers",
                      "Radio and Television Broadcasting Stations")),
              combine_sectors("79", "Technology, software and AI companies", source_codes = character(0),
                  extra_subtitles = c("AI services","Software Publishers",
                                      "Media Streaming Distribution Services, Social Networks, and Other Media Networks and Content Providers",
                                      "Computing Infrastructure Providers, Data Processing, Web Hosting, and Related Services")),
              
              combine_sectors("80", "Other Industries",  source_codes = c("44","45","42","72"),
                  extra_subtitles = c("Automotive Repair and Maintenance","Electronic and Precision Equipment Repair and Maintenance",
                                      "Commercial and Industrial Machinery and Equipment (except Automotive and Electronic) Repair and Maintenance",
                                      "Personal and Household Goods Repair and Maintenance","Personal Care Services", "Death Care Services",
                                      "Drycleaning and Laundry Services","Other Personal Services","Religious Organizations","Grandmaking and Giving Services",
                                      "Social Advocacy Organizations","Civic and Social Organizations"))             
)


```

```{r}
x[which(x$naics_sector == "50"),]$description
```



```{r}
x %>% arrange((naics_sector))
```


```{r}
# write to file
write.csv(x,"09_inputs/naics_sector_title_expanded_with_custom_codes.csv",row.names = FALSE)
```


### write OS level2 categories to csv

```{r}
categories <- read.csv("../08_alternative_pipeline/08_inputs/ContributionOpenSecretsCategories.csv")
```


```{r}
# take out uncoded and unitemized contributions
level2_categories <- unique(categories$level2_category)
level2_categories <- level2_categories[-which(level2_categories == "Uncoded")]
level2_categories <- level2_categories[-which(level2_categories == "Unitemized Contributions")]
level2_categories <- level2_categories[-which(level2_categories == "No Employer Listed or Found")] # also have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Homemakers/Non-income earners")] # we have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Employer Listed/Category Unknown")] # we have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Generic Occupation/Category Unknown")] # we have a different category for this
```

```{r}
# add in relevant custom categories
level2_categories <- c(level2_categories,"Retired/Homemaker/Student/Unemployed","Unknown","Political Actions Committees (PACs), contributor committees and political action funds")
```


```{r}
write.csv(level2_categories, "../open_secrets_level2_categories.csv",row.names = FALSE)
```





