---
title: "R Notebook"
output: html_notebook
---

# Purpose
Configure descriptions for each custom NAICS code for the LLM to reference when assigning codes.

# TO DOs
- any other opensecrets categories we'd need to add in ?

```{r}
library(readxl)
library(dplyr)
library(stringr)
```

### Add more information to the titles

```{r}
codes <- read_excel("09_inputs/digit_2022_Codes.xlsx")

# remove blank rows and columns
codes <- codes[-1,]
codes <- codes[,-c(4,5)]

colnames(codes)[c(2,3)] <- c("Code","Title")

```

The codes with a dash got left out - weren't grabbed in the 2-digit filter...

```{r}
# expand two digit codes with a dash...will mess up sequence no

codes[which(codes$Code == "31-33"),]$Code <- "31"
codes <- rbind(codes,c(NA,"32","Manufacturing"))
codes <- rbind(codes,c(NA,"33","Manufacturing"))

codes[which(codes$Code == "44-45"),]$Code <- "44"
codes <- rbind(codes,c(NA,"45","Retail Trade"))

codes[which(codes$Code == "48-49"),]$Code <- "48"
codes <- rbind(codes,c(NA,"49","Transportation and Warehousing"))

tail(codes)
```


```{r}
sector_expanded <- codes %>%
  filter(nchar(Code) == 2) %>%
  mutate(expanded_title = 
    sapply(Code, function(code) {
      parent_title <- codes %>%
        filter(Code == code) %>%
        pull(Title)
      
      sub_titles <- codes %>%
        filter(str_starts(Code, code), nchar(Code) == 4) %>%
        distinct(Title) %>%
        pull(Title)
      
      if (length(sub_titles) == 0) return(parent_title)
      
      all_titles <- c(parent_title, sub_titles)
    
      paste0(parent_title, ", including: ", paste(sub_titles, collapse = "; "))
    })
  )

```


### Collapse descriptions for combined categories: 

40: Combines 53 and 23
50: Combines 54, 55 and 56a
60: Combines 62, 61 and lirbaries and archives
70: Combines 71 + parts of 51
76: Remove association description from relvant naics code? 
77: Remove assocition description from relevant naics code
80: 44-45, 42, 92, parts of 81
92: 92 + defense contractors, take out tribes

```{r}
# helper function to combine sector descriptions
combine_sectors <- function(new_code, new_title, source_codes = character(0), extra_subtitles = character(0)) {
  if (length(source_codes) > 0) {
    sub_titles <- codes %>%
      filter(sapply(Code, function(c) any(str_starts(c, source_codes))) & nchar(Code) == 4) %>%
      distinct(Title) %>%
      pull(Title)
  } else {
    sub_titles <- character(0)
  }

  all_subs <- c(sub_titles, extra_subtitles)

  description <- if (length(all_subs) == 0) {
    new_title
  } else {
    paste0(new_title, ", including: ", paste(all_subs, collapse = "; "))
  }

  data.frame(naics_sector = new_code, description = description, stringsAsFactors = FALSE)
}

```


```{r}
#x[which(x$naics_sector == "52"),]$description
codes %>% filter(grepl("^56",Code) & nchar(Code) == 4)
```

Natural gas distribution goes in oil and gas.... not in 22
Natural gas includes natural gas distribution? 

```{r}
x <- sector_expanded[,c("Code","expanded_title"),]
colnames(x) <- c("naics_sector","description")

x <- rbind(x,c("99","Unknown"))
x <- rbind(x,c("90","Candidates/Political Parties"))
x <- rbind(x,c("88","Political Actions Committees (PACs), contributor committees and political action funds"))
#x <- rbind(x,c("78","Defense contractors"))
#x <- rbind(x,c("79","Technology, software and AI companies"))
x <- rbind(x,c("76","Business, professional and trade associations"))
x <- rbind(x,c("77","Unions"))
x <- rbind(x,c("92a","Native American tribes"))
x <- rbind(x,c("100","Retired/Homemaker/Student/Unemployed"))
#x <- rbind(x,c("01","Candidate contributions to their own campaigns."))

# remove source codes that get merged into combined categories
x <- x %>% filter(!naics_sector %in% c("22","23", "53","42","44","45","48","49","51","53","54","55","56","61","62","71","72","81"))

# add combined categories
x <- rbind(x, combine_sectors("20", "Utilities, infrastructure, telecommunications, renewable energy, transportation and warehousing", 
                              source_codes = c("22","48","49"),
                              extra_subtitles = c("Wired and Wireless Telecommunications (except Satellite)",
                                                  "Satellite Telecommunications","All Other Telecommunications",
                                                  "Waste Collection","Waste Treatment and Disposal","Remediation and Other Waste Management Services")),
              combine_sectors("40", "Construction/Real Estate", source_codes = c("23", "53")),
              combine_sectors("50", "Professional/Legal", source_codes = c("54", "55"),
                              extra_subtitles = c("Management and Holding Companies",
                                                  "Office Administrative Services",
                                                  "Employment Services",
                                                  "Facilities Support Services",
                                                  "Business Support Services",
                                                  "Travel Arrangement and Reservation Services",
                                                  "Investigation and Security Services",
                                                  "Services to Buildings and Dwellings")),
              combine_sectors("60", "Healthcare/Education", source_codes = c("61", "62"),
                              extra_subtitles = c("Libraries and archives")),
              combine_sectors("70", "Entertainment, Media and Gambling", source_codes = c("71"),
                  extra_subtitles = c("Motion Picture and Video Industries", "Sound Recording Industries","Newspaper, Periodical, Book, and Directory Publishers",
                      "Radio and Television Broadcasting Stations")),
              combine_sectors("79", "Technology, software and AI companies", source_codes = character(0),
                  extra_subtitles = c("AI services","Software Publishers",
                                      "Media Streaming Distribution Services, Social Networks, and Other Media Networks and Content Providers",
                                      "Computing Infrastructure Providers, Data Processing, Web Hosting, and Related Services")),
              
              combine_sectors("80", "Other Industries",  source_codes = c("44","45","42","72"),
                  extra_subtitles = c("Automotive Repair and Maintenance","Electronic and Precision Equipment Repair and Maintenance",
                                      "Commercial and Industrial Machinery and Equipment (except Automotive and Electronic) Repair and Maintenance",
                                      "Personal and Household Goods Repair and Maintenance","Personal Care Services", "Death Care Services",
                                      "Drycleaning and Laundry Services","Other Personal Services","Religious Organizations","Grandmaking and Giving Services",
                                      "Social Advocacy Organizations","Civic and Social Organizations"))             
)


```

```{r}
x[which(x$naics_sector == "50"),]$description
```


### Behavioral corrections from manual-review feedback (2026-09-10)

- **21 (Oil and Gas):** oil/refining/distribution companies were sometimes classified
  by their formal manufacturing or wholesale NAICS code instead.
- **92 (Public Administration):** defense contractors were being classified as plain
  manufacturing (31-33) instead of grouped with national security.
- **88 (PACs):** far too many politically-active orgs (unions, associations, any org
  with an affiliated PAC) were being labeled 88 outright; 88 should be reserved for
  entities explicitly identified as a PAC, and unitemized/intermediary dues passed
  through a PAC should reflect the underlying org's industry, not "PAC" as a label.
- **76 (Associations):** associations were frequently mislabeled as 88 due to
  PAC-adjacent political activity.
- **79 (Technology):** tech/software/AI companies were sometimes classified by the
  industry vertical their product serves (e.g. health-tech -> 60) instead of 79.
- **100 (Retired/Homemaker/Student/Unemployed):** applies only when the contributor
  is literally the trust itself (e.g. "The Smith Family Trust") -- NOT to LLCs or
  holding companies merely owned by or affiliated with a family trust (a mistake this
  same correction round initially introduced and had to narrow back down).
- **11 (Agriculture):** vineyards were sometimes classified as Beverage Manufacturing
  (31) instead of by their underlying farming operation.
- **52 (Finance and Insurance):** integrated health systems that are both insurer and
  hospital operator (e.g. Kaiser, Sutter) were being classified as Healthcare (60).
- **92a (Tribes):** tribal casinos were sometimes classified as gambling/entertainment
  (70) instead of by tribal ownership.
- **92 (Public Administration):** ambulance services weren't consistently grouped here
  rather than under Healthcare (60).

```{r}
x$description[x$naics_sector == "21"] <- paste0(
  x$description[x$naics_sector == "21"],
  " Also use this code for any company primarily in the oil/gas/petroleum business -- ",
  "extraction, drilling, refining, or distribution -- even if its formal NAICS ",
  "classification would otherwise be manufacturing (31/32) or wholesale trade (80)."
)

x$description[x$naics_sector == "92"] <- paste0(
  x$description[x$naics_sector == "92"],
  " Also use this code for defense contractors and companies whose primary business is ",
  "military/defense products or services, even if their formal NAICS code would ",
  "otherwise be manufacturing (31-33). Also use this code for ambulance and ambulance ",
  "services, even though this might otherwise suggest Healthcare (60)."
)

x$description[x$naics_sector == "88"] <- paste0(
  "Entities EXPLICITLY registered or identified as a Political Action Committee (PAC), ",
  "contributor committee, or political action fund -- e.g. name contains 'PAC', ",
  "'Political Action Committee', 'COPE', or confirmed via research to be a distinct ",
  "registered political committee. Do NOT use this code for a union, trade association, ",
  "corporation, or other organization just because it is politically active, lobbies, or ",
  "has a separately-registered affiliated PAC -- classify the parent organization by its ",
  "own actual industry instead (e.g. a union is '77', not '88'). Unitemized dues or ",
  "intermediary contributions passed through a PAC should be classified by the underlying ",
  "organization's actual industry when it can be determined (e.g. a union's PAC dues -> ",
  "'77'), not as '88'."
)

x$description[x$naics_sector == "76"] <- paste0(
  x$description[x$naics_sector == "76"],
  " Use this code even if the association has an affiliated PAC or engages in political ",
  "or lobbying activity -- the association itself is '76', not '88'; only a separately-",
  "registered PAC entity gets '88'."
)

x$description[x$naics_sector == "79"] <- paste0(
  x$description[x$naics_sector == "79"],
  " Use this code regardless of which industry vertical the technology serves -- e.g. ",
  "health-tech, fintech, and ad-tech companies are '79', not the industry they serve."
)

x$description[x$naics_sector == "100"] <- paste0(
  x$description[x$naics_sector == "100"],
  " Family trusts -> '100' ONLY when the contributor is literally the trust itself ",
  "(name reads as a trust, e.g. 'The Smith Family Trust,' 'Jane Doe 1992 Revocable ",
  "Trust') -- narrow; do not extend to LLCs/holding companies merely owned by or ",
  "affiliated with a trust."
)

x$description[x$naics_sector == "11"] <- paste0(
  x$description[x$naics_sector == "11"],
  " Includes vineyards, even though winemaking might otherwise suggest Beverage ",
  "Manufacturing (31) -- classify a vineyard's primary grape-growing/farming ",
  "operation as agriculture."
)

x$description[x$naics_sector == "52"] <- paste0(
  x$description[x$naics_sector == "52"],
  " Also use this code for integrated health systems that are both a health insurer ",
  "and a hospital/care operator (e.g. Kaiser Permanente, Sutter Health) -- classify ",
  "by their insurance function, not '60'."
)

x$description[x$naics_sector == "92a"] <- paste0(
  x$description[x$naics_sector == "92a"],
  " Includes tribal casinos and other tribal gaming operations -- classify by tribal ",
  "ownership ('92a'), not as gambling/entertainment ('70')."
)
```


```{r}
x %>% arrange((naics_sector))
```


```{r}
# write to file
write.csv(x,"09_inputs/naics_sector_title_expanded_with_custom_codes.csv",row.names = FALSE)
```


### write OS level2 categories to csv

```{r}
categories <- read.csv("../08_alternative_pipeline/08_inputs/ContributionOpenSecretsCategories.csv")
```


```{r}
# take out uncoded and unitemized contributions
level2_categories <- unique(categories$level2_category)
level2_categories <- level2_categories[-which(level2_categories == "Uncoded")]
level2_categories <- level2_categories[-which(level2_categories == "Unitemized Contributions")]
level2_categories <- level2_categories[-which(level2_categories == "No Employer Listed or Found")] # also have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Homemakers/Non-income earners")] # we have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Employer Listed/Category Unknown")] # we have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Generic Occupation/Category Unknown")] # we have a different category for this
```

```{r}
# add in relevant custom categories
level2_categories <- c(level2_categories,"Retired/Homemaker/Student/Unemployed","Unknown","Political Actions Committees (PACs), contributor committees and political action funds")
```


```{r}
write.csv(level2_categories, "../open_secrets_level2_categories.csv",row.names = FALSE)
```





