---
title: "name matching"
format: html
editor: visual
---

## Matching OpenSecrets names to CA campaign contributions

-   Using data from PowerSearch on contributions to the gubernatorial election
-   National Institute on Money and Politics categorizations data from Digital Democracy

```{r}
### load packages
#here::i_am("name_matching.qmd")
library(here)
library(stringr)
library(stringdist)
library(fuzzyjoin)
library(dplyr)
library(readr)
```

```{r}
### read in data
x <- read.csv(here("../data/campaign_contribution_data/data-2026-03-20-14-51.csv")) # latest governor data
categories <- read.csv("../data/categorization_data/ContributionOpenSecretsCategories.csv") # open secrets categorized data
```

Note: the categories data has '?' for special characters. Will need to try to address this at somep point.

### Standardize names

Write functions for name standardization.

```{r}
standardize_names <- function(name) {
  name %>%
    str_to_upper() %>%
    str_replace_all("\\&", "AND") %>% # this may make some names weird e.g. M&D Inc -> MANDD Inc.
    str_replace_all("\\#", "NUMBER") %>%
    str_replace_all(" NO ", " NUMBER ") %>% # manually checked that these all should be # for this data
    str_replace_all("\\&", "AND") %>%
    #str_replace_all("\\ + ", " AND ") %>%
    str_replace_all("[.',]", "") %>% # only removing these for now, because there's some punctuation i wanted to preserve, but could try removing all 
    str_replace_all("[-]", " ") %>%
    str_squish()
}

fix_typos <- function(name) { # fix typos found in names 
  name %>%
    str_to_upper() %>%
    str_replace_all("ENTITITES", "AND") %>% # this may make some names weird e.g. M&D Inc -> MANDD Inc.
    str_replace_all("ENTITES", "NUMBER") %>%
    str_replace_all("COMMITTE ", "COMMITTEE ") %>% # manually checked that these all should be # for this data
    str_replace_all("CAMMITTEE", "COMMITTEE") %>%
    str_replace_all("VACINITY", "VICINITY") %>%
    str_replace_all("AND AFFILIATES ENTITIES", "AND AFFILIATED ENTITIES") %>%
    str_squish()
}

replace_business_text <- function(name){
  name %>%
    str_replace_all(" LTD LLLP$", "") %>%
    str_replace_all(" AND AFFILIATED ENTITIES$", "") %>% # sometimes removing this just leaves a name, can think more about whether to remove
    str_replace_all(" AND AFFILIATED COMPANIES$", "") %>%
    str_replace_all(" AND AFFILIATED ENTITIES INC$", "") %>%
    str_replace_all(" AND AFFILIATED$", "") %>%
    str_replace_all("AND SUBSIDIARIES", "") %>%
    str_replace_all("AND ITS SUBSIDIARIES", "") %>% # could also include 'and its subsidiaries affiliates'
    str_replace_all(" ASSOC ", " ASSOCIATION ") %>% # think it's always association and not associated/associates
    str_replace_all(" LTD LLC$", "") %>%
    str_replace_all(" CORP$", "") %>% # cuts off certain phrases like 'a law corp', but can play around with it
    str_replace_all(" INC$", "") %>%
    str_replace_all(" INCORPORATED$", "") %>% # could include incorporation, corporation
    str_replace_all(" LLC$", "") %>%
    str_replace_all(" LTD$", "") %>%
    str_replace_all(" LLP$", "") %>%
    str_replace_all(" LP$", "") %>%
    str_replace_all(" PA$", "") %>%
    str_replace_all(" AND CO$", "") %>%
    str_replace_all(" CO$", "") %>%
    str_replace_all(" MD$", "") %>%
    str_replace_all(" MR ", "") # sometimes included in names
}


replace_committee_text <- function(name){
  name %>% 
    str_replace_all("POLITICAL ACTION LEAGUE FOR", "") %>%
    str_replace_all("POLITICAL ACTION LEAGUE", "") %>%
    str_replace_all("POLITICAL ACTION COMMITTEE STATE PAC", "") %>%
    str_replace_all("POLITICAL ACTION COMMITTEE", "") %>%
    str_replace_all("POLITICAL ACTION CO$", "") %>%
    str_replace_all("POLITICAL ACTION LEAGUE FOR", "") %>%
    str_replace_all("POLITICAL FUND", "") %>%
    str_replace_all("POLITICAL ACTION FUND", "") %>%
    str_replace_all("PAC ALL PURPOSE", "") %>%
    str_replace_all("PAC ALL PURPOSE ACCOUNT", "") %>%
    str_replace_all("SCC", "") %>%
    str_replace_all("SMALL CONTRIBUTOR COMMITTEE", "") %>%
    str_replace_all("SMALL CONT COMMITTEE", "") %>%
    str_replace_all("SMALL COMMITTEE", "") %>%
    str_replace_all("PAC FED PAC", "") %>% # hoping to get some names that are missing a space e.g. 'CUMMINS INC PAC CIPAC FED PAC ID C00377952'
    str_replace_all("STATE PAC$", "") %>% 
    str_replace_all("STATEWIDE PAC$", "") %>% 
    str_replace_all("FEDERAL PAC$", "") %>% 
    str_remove("(\\s+(FED\\s+PAC|PAC))+$") %>%
    str_replace_all("MAJOR DONOR ACCOUNT", "") %>%
    str_replace_all("MAJOR DONOR COMMITTEE", "") %>%
    str_replace_all("MAJOR DONOR", "") %>%
    str_replace_all("FPPC", "") 
}

```

Write functions to remove PAC / major donor ids and local unit numbers.

```{r}
remove_pac_info <- function(name){ 
  name %>%
  str_remove_all("\\s*\\([^)]*(ID|FPPC|FEC)[^)]*\\)") %>%
  str_remove("\\s*(ID(?:\\s*NUMBER)?|FPPC|FEC)\\s*[A-Z0-9]+$") %>%
  #str_remove("(\\s+(FED\\s+PAC|PAC))+$") %>%
  str_squish()
}

remove_unit_info <- function(name) {
  name %>%
    str_replace_all(
      regex("\\s+LOCAL(\\s+UNION)?(\\s+NO|\\s+NUMBER)?\\s*\\d+", ignore_case = TRUE),
      ""
    ) %>%
    str_replace_all(
      regex("\\s+NUMBER\\s*\\d+", ignore_case = TRUE),
      ""
    ) %>%
    str_squish()
}
```

Apply functions to the contributor names.

```{r}
# apply standardization functions to names (basic standardization e.g. convert to all caps, remove some punctuation, etc.)
x$standardized_name <- standardize_names(x$Contributor.Name)
categories$standardized_name <- standardize_names(categories$name)

# fix typos
x$standardized_name <- fix_typos(x$standardized_name)
categories$standardized_name <- fix_typos(categories$standardized_name)

# remove pac / donor id info from names
x$processed_name <- remove_pac_info(x$standardized_name) # creating new variable to store this name
categories$processed_name <- remove_pac_info(categories$standardized_name)

# remove local unit id info from names
x$processed_name<- remove_unit_info(x$processed_name)
categories$processed_name<- remove_unit_info(categories$processed_name)

# take out text related to pac/contributor committee/major donor, etc. (doesn't directly take out all instances of 'PAC' right now)
x$processed_name <- replace_committee_text(x$processed_name)
categories$processed_name <- replace_committee_text(categories$processed_name)

# take out common business entity abbreviations (llc, ltd, etc.)
x$processed_name <- replace_business_text(x$processed_name)
categories$processed_name <- replace_business_text(categories$processed_name)

# change CA to california -- need to double check this doesn't mess any names up
x$processed_name <- str_replace_all(x$processed_name, " CA ", " CALIFORNIA ")
categories$processed_name <- str_replace_all(categories$processed_name, " CA ", " CALIFORNIA ")

# remove parentheses + stuff inside
x$processed_name <- str_remove_all(x$processed_name, "\\s*\\(.*?\\)")
categories$processed_name <- str_remove_all(categories$processed_name, "\\s*\\(.*?\\)")

# last squish
x$processed_name <- str_squish(x$processed_name)
categories$processed_name <- str_squish(categories$processed_name)
```

```{r}
#unique(x[grep("\\(", x$standardized_name),c("standardized_name","processed_name")])
#unique(categories[grep("\\(", categories$standardized_name),c("standardized_name","processed_name","name")])
```

**Edge cases i need to address**

-   Right now, I'm removing the parts in parentheses because it leads to more matches overall, but sometimes the part in parentheses is needed for a match e.g. only the part in parenthese would match, but main part of the string doesn't match
-   SUNSET VIEW PLAZA LLC(FARSHID SHOOSHANI) -- this gets messed up b/c looks for "id" in parentheses
-   POLITICAL ACTION COMMITTE LOS ANGELES 80-20 -- removing political action committee removes most meaningful info from this
-   Removing corp at the end replaces "A LAW CORP" -- may want to keep this. Other examples: A professional corp; investment corp; financial corp; a pro corp; a prof law corp; management corp; a professional law corp
-   Want to replace some '&' but not others. e.g. changing 'M&D' to "MANDD' is a little weird, but lots of other patterns to think about when replacing the '&' as well. should be consistent at least when matching?

**PAC, donor, etc. related**

-   'POLITICAL ACTION FUND / ALL PURPOSE ACCOUNT'
-   STATE MAJOR DONOR - PAC
-   SOUTHERN CALIFORNIA PIPE TRADES DISTRICT COUNCIL NO 16; the 'NO 16' didn't get removed, but meant to capture this pattern

**state related**

-   of the state of california -\> of california
-   and vacinity -\> could remove

**business related** Think about removing these endings

-   CORPORATION
-   INCORPORATION
-   DBA
-   PARTNERSHIP
-   MD
-   LTC
-   ELE
-   CRC
-   SERVICES
-   COMPANY
-   RE or RE LLC
-   GROUP
-   BPN
-   MANAGEMENT and MGMT
-   ASSOCIATES
-   MBA
-   MGMTCORP
-   INCLUDING AGGREGATED CONTRIBUTIONS AND INDEPENDENT EXPENDITURES; INCLUDING AGGREGATED CONTRIBUTIONS; INCLUDING AGGREGATED CONTRIBUTIONS OF ITS AFFILIATED ECONOMIC DEVELOPMENT ENTITIES; INCLUDING AGGREGATED CONTRIBUTIONS / INDEPENDENT EXPENDITURES

### Use fuzzy matching to identify matches in OpenSecrets data

-   I tried levenshtein distance as well, but thought jaro-winkler distance did a better job after looking at matches manually.

```{r}
# start with a subset of data
top_contributors <- x[which(x$Amount >= 10000),] # contributions $10k and above
top_contributors$id <- 1:nrow(top_contributors)
nrow(top_contributors)
```

```{r}
name_matches <- stringdist_left_join(
  top_contributors, categories,
  by = c("processed_name" = "processed_name"),
  method = "jw",    # jaro-winkler distance
  max_dist = 0.05,  # chose after looking at different thresholds -- could adjust
  distance_col = "distance", 
  p = 0.1 
)
```

\^ this takes a long time, should optimize.

```{r}
#name_matches %>% group_by(id) %>% mutate(n = n_distinct(level1_category)) %>% filter(n > 1) %>% select(processed_name.x,level1_category,id)
#name_matches[grep("DMCG", name_matches$processed_name.x),] # not sure why this one isn't matching
```

```{r}
# prefer matches that have an assigned category as opposed to matching to an OpenSecrets name that is "uncoded"
best_match <- name_matches %>%
  group_by(id) %>%
  arrange(
    level1_category == "Uncoded", # FALSE for preferred rows
    distance
  ) %>% slice(1) %>% ungroup()
nrow(best_match)
```

\^ this doesn't currently address any contradictions in the category, just looking for matches right now

```{r}
nrow(best_match)
nrow(best_match[which(!is.na(best_match$processed_name.y)),])/ nrow(best_match)
nrow(best_match[which(!is.na(best_match$processed_name.y) & best_match$level1_category != "Uncoded"),]) / nrow(best_match)
```

-   about 60% of names matched, but only 24% of names matched to a categorized contributor (most matched to 'uncoded' category)

Look at what proportion of the total contribution amount we're capturing

```{r}
sum(best_match$Amount)
sum(best_match[which(!is.na(best_match$processed_name.y)),]$Amount) / sum(best_match$Amount)
sum(best_match[which(!is.na(best_match$processed_name.y) & best_match$level1_category != "Uncoded"),]$Amount) / sum(best_match$Amount)
```

-   not capturing even close to a majority of the money

### manually add categories for certain cases

```{r}
best_match[which(best_match$processed_name.x == "STEYER TOM"),]$level1_category <- "Candidate Contributions" # think this is the category, but not positive
best_match[which(best_match$processed_name.x == "UNITEMIZED CONTRIBUTIONS"),]$level1_category <- "Small donor contributions"

#sum(best_match[which(best_match$processed_name.x == "STEYER TOM"),]$Amount)
#sum(best_match[which(best_match$processed_name.x == "UNITEMIZED CONTRIBUTIONS"),]$Amount)
#sum(best_match[which(!is.na(best_match$processed_name.y) & best_match$level1_category != "Uncoded"),]$Amount)
```

-   Tom Steyer + Unitemized contributions account for \$92,195,127 (like 70% ish of those over 10k i think)

```{r}
nrow(best_match[which(best_match$level1_category != "Uncoded"),]) / nrow(best_match)
sum(best_match[which(best_match$level1_category != "Uncoded"),]$Amount) / sum(best_match$Amount)
```

-   When you account for Tom Steyer and the unitemized contributions, we're covering more like 84% of the top contributions and \~30% of rows match (not the same as unique names)

#### names that didn't match, but do have a match

This is just a list of a few I noticed, sure there are others that have matches that didn't match

-   '21st Century Alliance - Silicon Valley Chapter' — the chapter part prevents match (although only coded name is '21ST CENTURY FOX AMERICA INC & AFFILIATED ENTITIES')
-   'KK Restaurants LLC (Jagan Reddy Balasundaram)' -\> want it to match to 'BALASUNDARAM, JAGAN' — need the part in parentheses to make the match
-   'Vinoda Basnayake and Affiliated Entities' (including Versus Equity and VS Global - LLC) -\> want it to match to 'BASNAYAKE, VINODA'
-   'AMERICAN RESORT DEVELOPMENT ASSOCIATION RESORT OWNERS COALITION PAC (ARDA-ROC PAC)' -\> think it matches 'American Resort Development Association'
-   'SHEET METAL, AIR, RAIL - AND TRANSPORTATION WORKERS INTERNATIONAL ASSOCIATION' and 'SHEET METAL, AIR, RAIL, TRANSPORTATION WORKERS INTERNATIONAL UNION (SMART) PAC' would match without int'l assc or intl union...
-   'DMCG INC DBA BAIL HOTLINE BAILS BONDS INCLUDING AGGREGATED CONTRIBUTIONS' -\> should have matched, not sure why it didn't
