---
title: "Assign Final Classifications"
output: html_notebook
---

Combines every source the pipeline has for a contributor's industry
classification -- covers all $5k+ contributors, both direct givers to
races/props and $5k+ givers to PACs -- and picks one best NAICS code and one
best category per `uuid` (the per-contribution record identifier), weighting sources by how
reliable they are and by whether they agree with each other. Anything picked
without solid support gets flagged in `manual_review` for a human to check.

Inputs (see config block below for exact paths/columns expected):

- **LLM classification** -- `09_llm_classifier/classify_web_results.py` output
- **OpenSecrets match** -- `08_alternative_pipeline/0805_contributor_matching_open_secrets.Rmd` output
- **Running-list / EDD+H1B / keyword match, and ML classifier** -- `08_alternative_pipeline/0804_classify_contributors.py`
  this also needs inputs from `07_ml_unmatched_classifier/0702_predict_unmatched.py`


# Load data and packages

```{r}
library(dplyr)
library(readr)
library(stringr)
library(tidyr)

source("../08_alternative_pipeline/standardization_helpers.R")
source("resolve_classifications.R")
```

```{r}
# file paths

# Direct contributor LLM — uuid-keyed output from 0902; UUIDs match the current rule file (1:1).
DIRECT_LLM_PATH <- "../09_llm_classifier/09_outputs/classification_input_combined_expanded_full_2026-08-16.csv"

# PAC contributor LLM files — entity-level output joined via classification_unit_id.
#UNIT_LLM_PATHS <- c(
#  "../09_llm_classifier/09_outputs/pac_classification_batch1_full_full_2026-08-13.csv",
#  "../09_llm_classifier/09_outputs/pac_md_classification_last_batch_full_2026-08-16.csv"
#)

RULE_PATH <- "../08_alternative_pipeline/08_outputs/classified_contributors_combined.csv"
OS_PATH   <- "../08_alternative_pipeline/08_outputs/combined_contributions_os_matches.csv"
ML_PATH   <- "../../data/07_output_ml_classification/08_entities_with_ml_2026-08-15.csv"

input <- read.csv("../08_alternative_pipeline/08_outputs/classification_input_combined.csv")
#df <- read.csv("../08_alternative_pipeline/08_outputs/classified_contributors_combined.csv")
#df <- read.csv("../09_llm_classifier/09_outputs/classification_inpute_combined_full_2026-08-16.csv")
```

```{r}
nrow(y)
nrow(df)

length(unique(y$uuid))
length(unique(df$uuid))
```


```{r}
x <- read.csv("../08_alternative_pipeline/08_outputs/classification_input_combined.csv")
classified <- read.csv("../08_alternative_pipeline/08_inputs/already_classified_contributions.csv")

label_url <- "https://docs.google.com/spreadsheets/d/11QHvNJsdtMlc1YKo_iNvMB_Jfn5Ui-iYdlWhFYjSm9g/export?format=csv"
labels <- read.csv(label_url)
```



# Configurations

```{r}
# update these according to how upstream sources name columns

LLM_COLS <- c(
  unit_id    = "uuid",
  name          = "standardized_name",
  employer      = "standardized_employer_name",
  code          = "naics_code_llm",
  naics_conf    = "naics_confidence",
  category      = "open_secrets_category",
  category_conf = "open_secrets_confidence",
  reasoning     = "naics_reasoning"
)

RULE_COLS <- c(
  unit_id         = "uuid",
  entity_id       = "entity_id",
  entity_type     = "entity_type",
  name            = "Contributor.Name",
  employer        = "employer",
  occupation      = "occupation",
  code            = "naics_code",
  label           = "naics_label",
  category        = "level2_category",
  source1         = "data_source_1",
  source2         = "data_source_2",
  ml_code         = "ml_naics_code",
  ml_conf         = "ml_naics_code_conf",
  ml_category     = "ml_level2_category",
  ml_category_conf = "ml_level2_category_conf"
)

OS_COLS <- c(
  unit_id    = "uuid",
  name       = "standardized_name",
  employer   = "standardized_employer_name",
  code       = "naics_code_os",
  category   = "level2_category",
  match_conf = "opensecrets_confidence"
)
```

```{r}
# require_cols() and apply_colmap() are defined in resolve_classifications.R (sourced above).
```


# Load sources

```{r}
rule_required <- unname(RULE_COLS[c("unit_id", "code", "source1")])
os_required   <- unname(OS_COLS[c("unit_id", "code")])

# Load rule first — needed to build join maps for LLM expansion
rule_raw <- read_csv(RULE_PATH, col_types = cols(.default = "c"))
os_raw   <- read_csv(OS_PATH, col_types = cols(.default = "c"))

# Direct contributor LLM: uuid join (1:1 exact — UUIDs are from the current data pull).
direct_llm_expanded <- read_csv(DIRECT_LLM_PATH, col_types = cols(.default = "c")) %>%
  filter(!is.na(naics_code_llm) & naics_code_llm != "")

cat(sprintf("DIRECT-LLM: %d rows\n", nrow(direct_llm_expanded)))

# PAC contributor LLM: entity-level files joined via classification_unit_id
# (PAC contributors share one classification_unit_id across contributions).
rule_uid_map <- rule_raw %>%
  filter(!is.na(classification_unit_id)) %>%
  select(uuid, classification_unit_id)

#unit_llm_raw <- bind_rows(lapply(UNIT_LLM_PATHS, \(p) read_csv(p, col_types = cols(.default = "c")))) %>%
#  filter(!is.na(naics_code_llm) & naics_code_llm != "") %>%
#  distinct(classification_unit_id, .keep_all = TRUE) %>%
#  select(-any_of("uuid"))  # drop PAC-traversal uuid before join adds rule uuid

#unit_llm_expanded <- unit_llm_raw %>%
#  inner_join(rule_uid_map, by = "classification_unit_id") %>%
#  select(-classification_unit_id)

#cat(sprintf("UNIT-LLM: %d unique entities -> %d contribution rows from %d file(s)\n",
#            nrow(unit_llm_raw), nrow(unit_llm_expanded), length(UNIT_LLM_PATHS)))

# Combine; direct takes priority when both cover the same uuid (shouldn't happen
# since direct/PAC are disjoint, but guards against edge cases).
llm_raw <- bind_rows(direct_llm_expanded) %>% # , unit_llm_expanded
  distinct(uuid, .keep_all = TRUE)
cat(sprintf("LLM total: %d classified uuids\n", nrow(llm_raw)))

llm_required <- unname(LLM_COLS[c("unit_id", "code")])
require_cols(llm_raw, llm_required, DIRECT_LLM_PATH)
require_cols(rule_raw, rule_required, RULE_PATH)
require_cols(os_raw, os_required, OS_PATH)

llm  <- apply_colmap(llm_raw, LLM_COLS)
rule <- apply_colmap(rule_raw, RULE_COLS)
os   <- apply_colmap(os_raw, OS_COLS)

# rule_code is already crosswalked to the custom sector scheme (0805); os_code
# and llm_code are not (see resolve_classifications.R)
llm$code <- apply_os_llm_crosswalk(llm$code)
os$code  <- apply_os_llm_crosswalk(os$code)

rule <- rule %>% mutate(
  rule_tier = source1,
  ml_conf = as.numeric(ml_conf),
  ml_category_conf = as.numeric(ml_category_conf)
)
```


```{r}
direct_llm_expanded %>% group_by(uuid) %>% filter(n() > 1)
```


```{r}
# join ML predictions
# ML output has one row per unique classification unit (unique org name, or unique employer+occupation pair for individuals)
# join key is derived from the standardized employer/occupation columns -- these match between the ML output and rule_raw, unlike the more-processed `employer` col that apply_colmap uses.

if (file.exists(ML_PATH)) {
  ml_raw <- read_csv(ML_PATH, col_types = cols(.default = "c"))

  # Build ml_unit_key in ML output: orgs -> Contributor.Name, individuals -> employer|occupation
  # (employer/occupation here are standardized_employer_name/standardized_occupation from the 08 pipeline)
  ml_keyed <- ml_raw %>%
    mutate(
      ml_unit_key = if_else(
        tolower(entity_type) == "individual",
        paste0(trimws(employer), "|", trimws(occupation)),
        trimws(`Contributor.Name`)
      )
    ) %>%
    select(ml_unit_key,
           ml_code          = ml_naics_code,
           ml_conf          = ml_naics_code_conf,
           ml_category      = ml_level2_category,
           ml_category_conf = ml_level2_category_conf)

  # Derive the same key from rule_raw (before apply_colmap stripped columns),
  # using standardized_employer_name/standardized_occupation which match the ML output.
  rule_keys <- rule_raw %>%
    select(uuid, entity_type,
           standardized_employer_name, standardized_occupation,
           `Contributor.Name`) %>%
    mutate(
      ml_unit_key = if_else(
        tolower(entity_type) == "individual",
        paste0(trimws(standardized_employer_name), "|", trimws(standardized_occupation)),
        trimws(`Contributor.Name`)
      )
    ) %>%
    select(uuid, ml_unit_key)

  rule <- rule %>%
    select(-ml_code, -ml_conf, -ml_category, -ml_category_conf) %>%
    left_join(rule_keys, by = c("unit_id" = "uuid")) %>%
    left_join(ml_keyed, by = "ml_unit_key") %>%
    mutate(
      ml_conf          = as.numeric(ml_conf),
      ml_category_conf = as.numeric(ml_category_conf)
    ) %>%
    select(-ml_unit_key)

  cat(sprintf("ML predictions joined: %d / %d rows have ml_code\n",
              sum(!is.na(rule$ml_code)), nrow(rule)))
} else {
  warning(sprintf(
    "ML output not found at %s -- ml_* columns will be NA.\nRun: python code/07_ml_unmatched_classifier/run_ml_on_08_output.py",
    ML_PATH
  ))
}
```


# Build the base spine and join sources


```{r}
# may need to re do this depending on how we end up tracking entities
base <- rule %>%
  select(unit_id, entity_id, entity_type, name, employer, occupation) %>%
  left_join(
    llm %>% select(unit_id, name, employer) %>% rename(llm_name = name, llm_employer = employer),
    by = "unit_id"
  ) %>%
  left_join(
    rule_raw %>% select(uuid, Recipient.Committee) %>% rename(recipient_committee = Recipient.Committee),
    by = c("unit_id" = "uuid")
  ) %>%
  mutate(
    # For PAC-traversal org rows the contributor name is often absent from the source
    # data; use the parent PAC name as a last-resort label so it's identifiable in review.
    name     = coalesce(
      name, llm_name,
      if_else(entity_type == "organization" & !is.na(recipient_committee),
              paste0("[contributor to PAC: ", recipient_committee, "]"), NA_character_)
    ),
    employer = coalesce(employer, llm_employer)
  ) %>%
  select(-llm_name, -llm_employer, -recipient_committee)

combined <- base %>%
  left_join(
    rule %>% select(unit_id, code, label, category, source1, source2, rule_tier, ml_code, ml_conf, ml_category, ml_category_conf),
    by = "unit_id"
  ) %>%
  left_join(
    llm %>% select(unit_id, code, naics_conf, category, category_conf, reasoning) %>%
      rename(llm_code = code, llm_conf = naics_conf, llm_category = category,
             llm_category_conf = category_conf, llm_reasoning = reasoning),
    by = "unit_id"
  ) %>%
  left_join(
    os %>% select(unit_id, code, category, match_conf) %>%
      rename(os_code = code, os_category = category, os_conf = match_conf),
    by = "unit_id"
  ) %>%
  rename(rule_code = code, rule_label = label, rule_category = category)
```


# Resolve one best code (and, separately, one best category) per unit

- conf_is_high(), resolve_one(), code_override(), resolve_code_one(), and
- resolve_category_one() are defined in resolve_classifications.R (sourced above).


```{r}
resolved <- combined %>%
  rowwise() %>%
  mutate(
    .resolved_code = list(
      resolve_code_one(rule_code, rule_tier, ml_code, ml_conf, os_code, os_conf, llm_code, llm_conf)
    ),
    .resolved_category = list(
      resolve_category_one(rule_category, rule_tier, ml_category_conf, os_category, os_conf, llm_category, llm_category_conf)
    )
  ) %>%
  ungroup() %>%
  unnest_wider(.resolved_code, names_sep = "_") %>%
  unnest_wider(.resolved_category, names_sep = "_") %>%
  rename(
    code_final            = .resolved_code_value,
    code_final_source     = .resolved_code_source,
    manual_review_code    = .resolved_code_review,
    manual_review_reason_code = .resolved_code_reason,
    category_final            = .resolved_category_value,
    category_final_source     = .resolved_category_source,
    manual_review_category    = .resolved_category_review,
    manual_review_reason_category = .resolved_category_reason
  ) %>%
  mutate(
    manual_review = manual_review_code,
    manual_review_reason = manual_review_reason_code
  )

# Flag if LLM disagrees with the chosen final code — LLM is the most reliable
# independent signal, so any divergence warrants a human check.
# Exempt the 100-family (not-employed/retired): code_override locks those and
# LLM often misclassifies retired individuals.
resolved <- resolved %>%
  mutate(
    .llm_final_disagree = !is.na(llm_code) & !is.na(code_final) &
                          code_final != llm_code &
                          !grepl("^100", coalesce(code_final, "")) &
                          coalesce(rule_tier, "") != "pre_classified",
    manual_review = if_else(.llm_final_disagree, TRUE, manual_review),
    manual_review_reason = case_when(
      .llm_final_disagree & is.na(manual_review_reason) ~ "llm_disagrees_with_final",
      .llm_final_disagree ~ paste0(manual_review_reason, " | llm_disagrees_with_final"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.llm_final_disagree)

# Flag 76 (Associations) and 77 (Unions) for subsector review — these top-level
# codes are placeholders; a reviewer must choose the appropriate sub-code.
resolved <- resolved %>%
  mutate(
    .needs_subsector = coalesce(code_final, "") %in% c("76", "77"),
    manual_review = if_else(.needs_subsector, TRUE, manual_review),
    manual_review_reason = case_when(
      .needs_subsector & is.na(manual_review_reason) ~ "subsector_needed: 76/77 requires more specific code",
      .needs_subsector ~ paste0(manual_review_reason, " | subsector_needed"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.needs_subsector)
```

```{r}
# conflicting aliases: same entity_id has name/employer variants whose final resolved code (code_final, after all sources are weighted) disagree.
conflict_entity_ids <- resolved %>%
  filter(!is.na(code_final) & code_final != FALLBACK_CODE) %>%
  group_by(entity_id) %>%
  summarise(n_codes = n_distinct(code_final), .groups = "drop") %>%
  filter(n_codes > 1) %>%
  pull(entity_id)

if (length(conflict_entity_ids) > 0) {
  cat(sprintf("%d entity_ids have alias conflicts (variants resolved to different final codes)\n",
              length(conflict_entity_ids)))
  resolved <- resolved %>%
    mutate(
      .conflict = entity_id %in% conflict_entity_ids,
      manual_review = if_else(.conflict, TRUE, manual_review),
      manual_review_reason = case_when(
        .conflict & is.na(manual_review_reason)
          ~ "alias_conflict: variants of this entity resolved to different codes",
        .conflict
          ~ paste0(manual_review_reason, " | alias_conflict"),
        TRUE ~ manual_review_reason
      )
    ) %>%
    select(-.conflict)
}

cat(sprintf("Flagged for review: %d / %d rows\n",
            sum(coalesce(resolved$manual_review, FALSE)),
            nrow(resolved)
))
```


## Patch: suspected entity-type mismatches

Orgs mislabeled as `entity_type = "individual"` end up in the individual review
queue because they have no employer. Detect them by name keywords and reroute to
the main review queue. This is a patch until 0802 entity-type detection is fixed.

- cvs health, including aggregated contributions
- majestic realty co. and edward p. roski, jr.
- southern sierras chapter, neca

```{r}
ORG_NAME_RE <- paste0(
  "LLC|\\bInc\\b|\\bCorp\\b|\\.L\\.P\\.|L\\.L\\.C\\.|",
  "Political Action|Contributions|",
  "\\bAssociation\\b|\\bFoundation\\b|\\bChapter\\b|",
  "\\bUnion\\b|\\bCouncil\\b|\\bCommittee\\b|\\bFund\\b|",
  "Holdings|Properties|\\bManagement\\b|",
  "\\bHospital\\b|Medical|\\bBank\\b|Insurance|Realty|Health|",
  "Construction|NECA|IBEW|\\bAFL\\b|\\bCIO\\b|\\bSEIU\\b|",
  "Contractors|Industries|Manufacturing|Supply|",
  "Resources|Energy|Electric|Communications|",
  "\\bMedia\\b|Networks|Technologies|Solutions|\\bServices\\b|",
  "\\bSystems\\b|\\bGroup\\b|Partners|Enterprises|\\bTrust\\b|",
  "International|\\bNational\\b|\\bCalifornia\\b|\\bAmerican\\b|FedEx"
)

resolved <- resolved %>%
  mutate(
    # entity_id prefix is unreliable (some orgs got "indiv_" IDs from 0802 detection),
    # so name keywords are the primary signal here.
    suspected_org = (tolower(entity_type) == "individual") &
      grepl(ORG_NAME_RE, coalesce(name, ""), ignore.case = TRUE, perl = TRUE),
    manual_review = if_else(suspected_org, TRUE, manual_review),
    manual_review_reason = case_when(
      suspected_org & is.na(manual_review_reason) ~ "suspected_entity_type_mismatch",
      suspected_org ~ paste0(manual_review_reason, " | suspected_entity_type_mismatch"),
      TRUE ~ manual_review_reason
    )
  )

cat(sprintf("Suspected entity-type mismatches: %d rows\n", sum(coalesce(resolved$suspected_org, FALSE))))
```


```{r}
resolved %>% filter(suspected_org == TRUE)
```


- FedExPAC, FedEx: shouldn't have been grabbed because parent pac is a federal PAC
  - make sure we update to not grabbing this




## Cross-source entity conflict detection

The same contributor can appear as both a direct contributor and a PAC contributor,
each getting a different `entity_id` because entity resolution runs independently
per path. Detect cross-source matches by normalized name (orgs) or
normalized name|employer (individuals) and flag rows where the matched group
resolves to different final codes.

```{r}
resolved <- resolved %>%
  mutate(
    entity_norm_key = tolower(trimws(coalesce(name, ""))) %>%
      str_replace_all("[^a-z0-9|]", " ") %>%
      str_squish()
  )

# 100/100a/100b are the same entity-type family; normalize to "100" before
# counting distinct codes so those sub-code differences don't trigger a false conflict.
cross_source_summary <- resolved %>%
  filter(entity_norm_key != "") %>%
  mutate(.code_cmp = sub("^100[ab]$", "100", coalesce(code_final, ""))) %>%
  group_by(entity_norm_key) %>%
  summarise(
    n_entity_ids  = n_distinct(entity_id, na.rm = TRUE),
    n_final_codes = n_distinct(.code_cmp[.code_cmp != "" & .code_cmp != FALLBACK_CODE]),
    .groups = "drop"
  ) %>%
  filter(n_entity_ids > 1 & n_final_codes > 1)

# If one entity in a conflict group was pre_classified by John, propagate that
# code to non-locked rows with a compatible employer. "Compatible" means the
# employers match (after normalization) OR at least one side has no employer 
# Rows in conflict groups whose employer doesn't
# match John's row are still flagged for manual review.
# (source2 carries the reviewer name from already_classified_contributions.csv)
norm_emp <- function(x) {
  tolower(trimws(coalesce(x, ""))) %>%
    str_replace_all("[^a-z0-9]", " ") %>%
    str_squish()
}

pre_classified_john_codes <- resolved %>%
  filter(
    entity_norm_key %in% cross_source_summary$entity_norm_key,
    rule_tier == "pre_classified",
    grepl("john", tolower(coalesce(source2, "")))
  ) %>%
  mutate(.john_emp = norm_emp(employer)) %>%
  group_by(entity_norm_key) %>%
  slice(1) %>%
  ungroup() %>%
  select(entity_norm_key, .propagated_code = code_final, .john_emp)

resolved <- resolved %>%
  mutate(.emp_norm = norm_emp(employer)) %>%
  left_join(pre_classified_john_codes, by = "entity_norm_key") %>%
  mutate(
    .employer_ok  = .emp_norm == "" | coalesce(.john_emp, "") == "" | .emp_norm == .john_emp,
    .in_conflict  = entity_norm_key %in% cross_source_summary$entity_norm_key,
    .not_locked   = coalesce(code_final_source, "") != "pre_classified (locked)",
    cross_source_conflict = .in_conflict & (is.na(.propagated_code) | !.employer_ok),
    .propagated   = .in_conflict & !is.na(.propagated_code) & .not_locked & .employer_ok,
    code_final        = if_else(.propagated, .propagated_code, code_final),
    code_final_source = if_else(.propagated, "pre_classified (propagated: John)", code_final_source),
    manual_review     = if_else(cross_source_conflict, TRUE, manual_review),
    manual_review_reason = case_when(
      cross_source_conflict & is.na(manual_review_reason) ~ "cross_source_entity_conflict",
      cross_source_conflict ~ paste0(manual_review_reason, " | cross_source_entity_conflict"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.propagated_code, -.in_conflict, -.not_locked, -.emp_norm, -.john_emp, -.employer_ok)

cat(sprintf(
  "Cross-source entity conflicts: %d unique norm keys → %d rows flagged | %d rows auto-resolved by John pre_classified propagation\n",
  nrow(cross_source_summary),
  sum(coalesce(resolved$cross_source_conflict, FALSE)),
  sum(coalesce(resolved$.propagated, FALSE))
))
```

```{r}
resolved %>% filter(cross_source_conflict == TRUE) %>% arrange(name) %>% select(name,employer,occupation, code_final, rule_code, os_code, llm_code, ml_code)
```


WEST, BENJAMIN 

```{r}
review_queue %>% filter(grepl("WEST, BENJAMIN",toupper(Contributor.Name)))
resolved %>% filter(grepl("WEST, BENJAMIN",toupper(name)))
```



```{r}
# Review: all rows where code was propagated from John's pre_classified entry.
# Check that the propagated code is correct before accepting.
propagated_review <- resolved %>%
  filter(coalesce(.propagated, FALSE)) %>%
  select(name, employer, occupation, entity_norm_key,
         code_final, code_final_source,
         rule_code, rule_tier, source2,
         llm_code, os_code, ml_code)

cat(sprintf("Propagated-code review: %d rows\n", nrow(propagated_review)))
propagated_review
```

```{r}
resolved <- resolved %>% select(-.propagated)
```

```{r}
# Propagate 100a / 100b sub-codes from already_classified to individuals who
# resolved to plain "100" (not-employed/retired) but whose exact Contributor.Name
# appears in the pre-classified data with a Megadonors / Major Donors sub-code.
sub_100_lookup <- classified %>%
  mutate(
    .match_name = trimws(toupper(coalesce(Contributor.Name, ""))),
    .sub_code = case_when(
      "revised_category" %in% names(.) &
        trimws(coalesce(revised_category, "")) == "Megadonors"   ~ "100a",
      "revised_category" %in% names(.) &
        trimws(coalesce(revised_category, "")) == "Major Donors" ~ "100b",
      trimws(coalesce(code_final, "")) == "100a" ~ "100a",
      trimws(coalesce(code_final, "")) == "100b" ~ "100b"
    )
  ) %>%
  filter(!is.na(.sub_code), .match_name != "") %>%
  distinct(.match_name, .keep_all = TRUE) %>%
  select(.match_name, .sub_code)

resolved <- resolved %>%
  mutate(.name_upper = trimws(toupper(coalesce(name, "")))) %>%
  left_join(sub_100_lookup, by = c(".name_upper" = ".match_name")) %>%
  mutate(
    .inherit_sub = code_final == NOT_EMPLOYED_CODE &
                   tolower(entity_type) == "individual" & !is.na(.sub_code),
    code_final        = if_else(.inherit_sub, .sub_code, code_final),
    code_final_source = if_else(.inherit_sub, "pre_classified (100 sub-code inherited)", code_final_source),
    manual_review     = if_else(.inherit_sub, FALSE, manual_review)
  ) %>%
  select(-.name_upper, -.sub_code)

cat(sprintf("100 sub-code inherited: %d rows\n",
            sum(coalesce(resolved$.inherit_sub, FALSE))))
```

```{r}
# Review: propagated 100→100a/100b rows
resolved %>%
  filter(coalesce(.inherit_sub, FALSE)) %>%
  select(name, employer, occupation, code_final, code_final_source, rule_code, llm_code)
```

```{r}
resolved <- resolved %>% select(-.inherit_sub)
```


```{r}
# pre_classified rows are already human-reviewed and should not be
# in the review queue in most cases. Exceptions:
#   - alias_conflict: a new name/employer variant may contain new information
#     that needs reconciling with the pre_classified code.
#   - cross_source_conflict: two pre_classified entries for the same entity
#     resolved to different codes — a prior-review contradiction worth flagging.
# All other review flags (LLM disagrees, 76/77 subsector, etc.) are cleared for
# pre_classified rows since the code has already been human-confirmed.
resolved <- resolved %>%
  mutate(
    manual_review = if_else(
      coalesce(rule_tier, "") == "pre_classified" &
        !grepl("alias_conflict|cross_source_entity_conflict",
               coalesce(manual_review_reason, ""),
               fixed = FALSE),
      FALSE,
      manual_review
    )
  )
cat(sprintf("After pre_classified guard — flagged: %d / %d rows\n",
            sum(coalesce(resolved$manual_review, FALSE)), nrow(resolved)))
```

DON'T include if difference is 100 vs 100a

TELACU Industries, Inc.
Telacu Industries Inc
Only one of these matches to rule_code and not sure why... 

Hernandez, Jose	TIERRA LUNA ENGINEERING	FARMER AND AEROSPACE ENGINEER	
Hernandez, Jose	NONE	CEO	
Hernandez, Jose	IDEATE CALIFORNIA	CEO

### data checkes

```{r}
# missing LLM and ML codes
resolved %>% filter(is.na(llm_code))
resolved %>% filter(is.na(ml_code))
```

```{r}
resolved %>% filter(grepl("telacu",tolower(name)))
```


```{r}
unique(resolved$code_final_source)
```

### misc sources

```{r}
resolved %>% filter(code_final_source == "rule (running_list)" & code_final != llm_code) %>% select(name, employer, rule_code, source1, llm_code, ml_code, manual_review) %>% distinct() 
```

### unresolved

```{r}
resolved %>% filter(code_final == "99")
```


```{r}
resolved %>% filter(code_final_source == "uncategorized fallback")
```


### political entities

```{r}
resolved %>% filter(code_final_source == "political-entity rule")
```

Nothing should end up as a PAC.... 
- Abbott Laboratories Employee Political Action Committee - C00040279 -> why is this in here? missing id? 

How did these end up in here? 
- the AFSCME General Fund, CA District Council 36 of
- Cope Healthcare Consulting, Inc.


### ML / LLM aggreement

- look at when this DOESN'T match final code

```{r}
resolved %>% filter(llm_code == ml_code)
```


```{r}
resolved %>% filter(llm_code == ml_code & llm_code != code_final) %>% select(name, employer, rule_code, source1, source2, llm_code, ml_code, manual_review, manual_review_code) %>% distinct() 
```

identity_override doesn't seem realible.. e.g. AI and Union

```{r}
resolved %>% filter(llm_code != rule_code & manual_review == FALSE & source1 != "pre_classified") %>% select(name, employer, rule_code, source1, llm_code, code_final, code_final_source, manual_review) %>% distinct() 
```

REMOVE from ID override
- software
- AI


### llm code as source

```{r}
resolved %>% filter(code_final_source == "llm (high confidence)" ) %>% select(name, rule_label, llm_code, ml_code, manual_review) %>% distinct() %>% select(rule_label) %>% distinct()
resolved %>% filter(code_final_source == "llm (high confidence)" ) %>% select(name, employer, rule_label, llm_code, ml_code, manual_review, manual_review_code) %>% distinct()
```

- LLM code ONLY used when ML and LLM disagree and no rule is available


### ml code as source

```{r}
resolved %>% filter(code_final_source == "ml only") %>% select(name, rule_label, llm_code, ml_code, manual_review) %>% distinct()
```

- ML code only used when no rule or LLM code available, manual_review is always true.

```{r}
unique(pac_final$os_conf)
```

```{r}
resolved %>% filter(name == "Dentaquest LLC")
```



### NA names

```{r}
resolved %>% filter(is.na(name))
```

All of these are the same entity. not sure what's happening.

```{r}
input %>% filter(uuid == "67e57677-2cf7-49d1-911c-38e77179168e")
input %>% filter(uuid == "1d751068-5c55-4b25-8873-fa41581fa154")
input %>% filter(uuid == "279f330b-8517-4391-a626-338efccefbd8")
```
- Not sure where these are coming from...

```{r}
resolved %>% filter(code_final_source == "rule (running_list)" & llm_code != rule_code)
```

```{r}
resolved %>% filter(code_final_source == "rule (identity_override)" & llm_code != rule_code)
```



```{r}
resolved %>% filter(code_final_source == "rule (employer_lookup)")
```

Everything else looks good
Huynh, Ethan	SELF	PHARMACIST -> classified by employer lookup???
Tuck, Nathaniel	SELF	PRODUCER -> same situation
Thomas, Elizabeth Allison	SELF	PRODUCER -> 11

```{r}
llm_raw %>% filter(uuid == "35bef780-2410-41e4-91dd-4347aa396e40")
```



```{r}
resolved %>% filter(code_final_source == "llm only") %>% select(name,employer,occupation,ml_code,llm_code,ml_conf)
```
- not that many right now that are LLM only, but will defs want to check these are correct... why no ML?
- should flag ML, LLM aggreement as being the code final source 
```{r}
length(which(resolved$manual_review == TRUE))
length(which(resolved$manual_review != TRUE))
```


```{r}
unique(resolved$manual_review_reason)
```

```{r}
resolved %>% filter(manual_review_reason == "no source produced a usable value" & !is.na(llm_code))
```

25k rows where no source had a usable value??? because of missing pac classifications? 
- almost all because LLM codes were missing

```{r}
resolved %>% filter(grepl("alias_conflict",manual_review_reason)) %>% arrange(entity_id)
```

Conflicts in aliases: 
- many because some rows show retired and others don't
- how to structure review queue so that aliases can be split if they are different? add a column to add groups to divide? 
- sometimes there are conflicts with pre-classified if employer is different, okay to flag those? probs have to


```{r}
unique(resolved$manual_review_reason)
```


```{r}
resolved %>% filter((rule_code != ml_code) & (ml_code == llm_code) &!is.na(llm_code)) %>% select(name, employer, source1,source2, entity_type,rule_code, ml_code, llm_code)
```

```{r}
unique(resolved$ml_code)
unique(resolved$source1)
unique(resolved$source2)
```
Not sure how much we can rely on employer regex

```{r}
colnames(resolved)
```

```{r}
resolved %>% filter(code_final == "99") %>% select(name,rule_code,)
```


```{r}
resolved %>% filter(manual_review == TRUE) %>% select(name,employer,occupation,manual_review_reason) %>% distinct()
```

- this will go down when we add more LLM classifications.. 

```{r}
length(unique(resolved[which(resolved$manual_review == TRUE),]$entity_id))
```

so many reviews to do... 

```{r}
resolved %>% filter(name == "Uber Technologies, Inc.")
```


# Attach the editorial description to the final code

```{r}
#desc <- read_csv(DESC_PATH, col_types = cols(.default = "c"))

# expects columns "Sector" (code) and "Custom.Label" 
#require_cols(desc, c("Sector", "Custom.Label"), DESC_PATH)
labels <- labels %>% select(sector,sector_description) %>% distinct()
desc <- labels %>% transmute(code_final = sector, code_final_description = sector_description)


resolved <- resolved %>%
  left_join(desc, by = "code_final")
```


### add category for unitemized contributions

```{r}
resolved %>% filter(grepl("unitemized",tolower(name))) %>% select(code_final,name)
```

Unitemized Contributions -> code 101 -> add to 0805??

### figure out computation as a part of the whole...that happens later


### check which categories were chosen

- rule_code
- source1
- source2
- rule_tier
- llm_code
- os_code

```{r}
unique(resolved$rule_code)
```


```{r}
resolved %>% filter(os_code != final_code) %>% select(unit_id, name, employer, code_final, rule_code,source1,source2, llm_code,os_code)
```

Need to map the pre-classified codes to our current codes. 
why does source1 = employer_lookup and source2 = John? 

```{r}
llm %>% filter(unit_id == "7738b3cd-eb0f-415a-8dc8-90b722af8cc4")
```


```{r}
colnames(resolved)
```



# Review + write output

```{r}
resolved %>% count(code_final_source, sort = TRUE)
resolved %>% count(category_final_source, sort = TRUE)
cat(sprintf("%d / %d units flagged for manual review (%.1f%%)\n",
            sum(resolved$manual_review), nrow(resolved),
            100 * mean(resolved$manual_review)))
```

```{r}
out_cols <- c(
  "unit_id", "entity_id", "entity_type", "name", "employer", "occupation",
  "code_final", "code_final_source", "code_final_description",
  "category_final", "category_final_source",
  "manual_review", "manual_review_reason",
  "suspected_org", "cross_source_conflict",
  "rule_code", "rule_label", "rule_tier", "rule_category",
  "os_code", "os_category", "os_conf",
  "llm_code", "llm_conf", "llm_category", "llm_category_conf", "llm_reasoning",
  "ml_code", "ml_conf", "ml_category", "ml_category_conf"
)
out_cols <- out_cols[out_cols %in% names(resolved)]

write_csv(resolved[, out_cols], "10_outputs/final_classifications_before_review.csv")
write_csv(resolved, "10_outputs/final_classifications_full_before_review.csv")
```


# Confidence review sample

Stratified random sample of confident (non-flagged) classifications, one group
per `code_final_source`, to check how reliable each source actually is in practice.

```{r}
SAMPLE_N_PER_SOURCE <- 20  # rows to draw per code_final_source bucket

set.seed(42)
confidence_sample <- resolved %>%
  filter(
    !manual_review,
    !is.na(code_final),
    code_final != FALLBACK_CODE,
    code_final != NOT_EMPLOYED_CODE,       # these go in the individual review queue
    code_final_source != "pre_classified (locked)"  # already human-reviewed
  ) %>%
  group_by(code_final_source) %>%
  slice_sample(n = SAMPLE_N_PER_SOURCE) %>%
  ungroup()

cat(sprintf("Confidence sample: %d rows across %d source buckets\n",
            nrow(confidence_sample), n_distinct(confidence_sample$code_final_source)))
confidence_sample %>% count(code_final_source, sort = TRUE)
```


```{r}
sample_cols <- c(
  "unit_id", "entity_id", "entity_type", "name", "employer", "occupation",
  "code_final", "code_final_source", "code_final_description",
  "rule_code", "rule_tier",
  "llm_code", "llm_conf", "llm_reasoning",
  "os_code", "os_conf",
  "ml_code", "ml_conf"
)

write_csv(
  confidence_sample %>% select(any_of(sample_cols)),
  "10_outputs/classification_confidence_sample.csv"
)
```



# Individual review queue (code = 100)

All individuals resolved to code 100 (not employed / retired / homemaker /
student), deduplicated across name + employer + occupation + location, with a
stable `name_id` hash of `Contributor.Name`. Excludes individuals already in
`already_classified_contributions.csv`.


```{r}
library(digest)

ind_review_raw <- resolved %>%
  filter(code_final == NOT_EMPLOYED_CODE,
         tolower(entity_type) == "individual",
         !coalesce(suspected_org, FALSE)) %>%
  left_join(
    rule_raw %>% select(uuid, Contributor.City, Contributor.State, Contributor.Zip.Code),
    by = c("unit_id" = "uuid")
  ) %>%
  filter(!(name %in% classified$Contributor.Name))

ind_review_queue <- ind_review_raw %>%
  select(name, employer, occupation,
         Contributor.City, Contributor.State, Contributor.Zip.Code) %>%
  distinct(across(everything(), tolower), .keep_all = TRUE) %>%
  arrange(name) %>%
  mutate(name_id = sapply(name, \(x) substr(digest(x, algo = "xxhash32"), 1, 8)))

cat(sprintf("Individual review queue: %d unique name/employer/occupation/location combos\n",
            nrow(ind_review_queue)))
```

```{r}
ind_review_queue %>% filter(name %in% tolower(classified$Contributor.Name))
```

```{r}
resolved %>% filter(tolower(name) == "quillin, patty") %>% select(code_final)
```


CVS Health, Including Aggregated Contributions... another indep expenditure example...
some with entity_type = org have a entity_id starting with indiv_...not sure why that would be, was it reassigned at some point? 
Just do a manual override? 
  - think that's the best for right now. send to manual review maybe? 
Need to update org flag... 


```{r}
write_csv(ind_review_queue, "10_outputs/individual_review_queue.csv")
```


### check individual overlap


```{r}
resolved %>% filter((tolower(name) %in% tolower(classified$Contributor.Name)) & source1 != "pre_classified")
```




# Combined manual review queue

All rows flagged for manual review plus the stratified confidence sample, in a
single file for a reviewer. Includes raw `Contributor.Name` / `Contributor.Employer`
fields (pre-standardization) alongside standardized columns — the raw fields are
the most stable keys for matching reviewed decisions back to future data pulls.

```{r}
# Step 1: derive recipient label and contribution type from raw contribution columns.
rule_raw_annotated <- rule_raw %>%
  mutate(
    # recipient: ballot measure name, candidate name, or committee name (in that priority)
    recipient_label = case_when(
      Ballot.Measure.Contribution == "Y" ~
        coalesce(na_if(trimws(Ballot.Measure), ""), na_if(trimws(Recipient.Committee), "")),
      !is.na(Recipient.Name) & trimws(Recipient.Name) != "" ~ trimws(Recipient.Name),
      TRUE ~ na_if(trimws(Recipient.Committee), "")
    ),
    # contribution type: pac_contributor (came via PAC traversal) / Ballot.Measure / office
    contribution_type = case_when(
      !is.na(parent_pac_id) & parent_pac_id != "" ~ "pac_contributor",
      Ballot.Measure.Contribution == "Y"           ~ "Ballot.Measure",
      Candidate.Contribution == "Y"                ~ "office",
      TRUE ~ NA_character_
    )
  )

# Step 2: join raw pre-standardization fields and build a stable dedup key.
# Key = raw Contributor.Name | raw Contributor.Employer | raw Contributor.Occupation
# for ALL entity types — avoids dependency on entity_type (which can change) and
# entity_id (which is reassigned each run). These raw fields come directly from
# the source data and are stable across processing and standardization changes.
review_queue_raw <- bind_rows(
  resolved %>%
    filter(coalesce(manual_review, FALSE)) %>%
    mutate(.queue_source = "flagged"),
  # code=100 individuals included alongside flagged rows so the same individual
  # who sometimes lists an employer (→ flagged for review) and sometimes doesn't
  # (→ code 100) can be resolved in one place. De-duplication below collapses them.
  resolved %>%
    filter(code_final == NOT_EMPLOYED_CODE,
           tolower(entity_type) == "individual",
           !coalesce(suspected_org, FALSE),
           !(name %in% classified$Contributor.Name),
           !coalesce(manual_review, FALSE)) %>%  # already included above if flagged
    mutate(.queue_source = "individual_100"),
  confidence_sample %>%
    mutate(.queue_source = "confidence_sample")
) %>%
  distinct(unit_id, .keep_all = TRUE) %>%
  left_join(
    rule_raw %>% select(uuid, any_of(c(
      "Contributor.Name", "Contributor.Employer", "Contributor.Occupation"
    ))),
    by = c("unit_id" = "uuid"),
    suffix = c("", ".raw")
  )

# Ensure raw columns exist even if missing from rule_raw schema
for (.col in c("Contributor.Name", "Contributor.Employer", "Contributor.Occupation")) {
  if (!.col %in% names(review_queue_raw)) review_queue_raw[[.col]] <- NA_character_
}

review_queue_raw <- review_queue_raw %>%
  mutate(
    dedup_key = paste0(
      trimws(coalesce(`Contributor.Name`, name, "")), "|",
      trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
      trimws(coalesce(`Contributor.Occupation`, occupation, ""))
    )
  )

# Step 3: per-dedup-group aggregates — pipe-join unique recipient, contribution type,
# and location variants so a reviewer sees every race/prop this entity appears in.
geo_variants <- review_queue_raw %>%
  select(unit_id, dedup_key) %>%
  left_join(
    rule_raw_annotated %>% select(uuid, recipient_label, contribution_type,
                                  Contributor.City, Contributor.State, Contributor.Zip.Code),
    by = c("unit_id" = "uuid")
  ) %>%
  group_by(dedup_key) %>%
  summarise(
    n_contributions    = n(),
    recipients         = paste(sort(unique(na.omit(trimws(recipient_label)))),   collapse = " | "),
    contribution_types = paste(sort(unique(na.omit(trimws(contribution_type)))), collapse = " | "),
    cities    = paste(sort(unique(na.omit(trimws(Contributor.City)))),     collapse = " | "),
    states    = paste(sort(unique(na.omit(trimws(Contributor.State)))),    collapse = " | "),
    zip_codes = paste(sort(unique(na.omit(trimws(Contributor.Zip.Code)))), collapse = " | "),
    .groups = "drop"
  )

# Step 4: dedup to one row per classification unit; prefer rows with the most source coverage.
# Then sort for reviewer: group by name so alias variants appear together, then by reason
# so alias_conflict rows cluster within a name, then by employer to sub-group within a name.
review_queue <- review_queue_raw %>%
  left_join(geo_variants, by = "dedup_key") %>%
  #arrange(desc(!is.na(llm_code)), desc(!is.na(rule_code)), desc(!is.na(ml_code))) %>%
  distinct(dedup_key, .keep_all = TRUE) %>%
  arrange(name, manual_review_reason, employer)

review_queue_cols <- c(
  "dedup_key",
  "Contributor.Name", "Contributor.Employer", "Contributor.Occupation",  # raw
  "name", "employer", "occupation",                                        # standardized
  "n_contributions", "contribution_types", "recipients",
  "cities", "states", "zip_codes",
  "code_final", "code_final_source", "code_final_description",
  "manual_review", "manual_review_reason",
  "rule_code", "rule_tier",
  "ml_code", "ml_conf",
  "llm_code", "llm_conf", "llm_reasoning",
  "os_code", "os_conf",
  ".queue_source"
)

```

```{r}
unique(review_queue$manual_review_reason)
```

```{r}
review_queue %>% filter(manual_review_reason == "no source produced a usable value")
```

```{r}
review_queue[which(review_queue$manual_review_reason == "llm only source"),]$manual_review <- FALSE
review_queue[which(review_queue$manual_review_reason == "llm only source"),]$manual_review_code <- FALSE
```



```{r}
review_queue %>% filter(manual_review_reason == "subsector_needed: 76/77 requires more specific code")
review_queue %>% filter(is.na(manual_review_reason))
```


```{r}
nrow(review_queue)
```

```{r}
unique(review_queue$manual_review_reason)
```

```{r}
review_queue %>% filter(grepl("west, benajamin",tolower(Contributor.Name)))
resolved %>% filter(grepl("west, benajamin",tolower(name)))
```


```{r}
review_queue %>% filter(manual_review_reason == "no rule; llm only") %>% select(name,ml_code, ml_conf,llm_code)
```


```{r}
write_csv(
  review_queue %>% select(any_of(review_queue_cols)),
  "10_outputs/manual_review_queue.csv"
)
cat(sprintf(
  "Review queue: %d unique entities (from %d raw rows; %d flagged + %d individual_100 + %d confidence sample)\n",
  nrow(review_queue),
  nrow(review_queue_raw),
  sum(coalesce(resolved$manual_review, FALSE)),
  sum(review_queue$.queue_source == "individual_100", na.rm = TRUE),
  nrow(confidence_sample)
))
```

