---
title: "Assign Final Classifications"
output: html_notebook
---

Combines every source the pipeline has for a contributor's industry
classification -- covers all $5k+ contributors, both direct givers to
races/props and $5k+ givers to PACs -- and picks one best NAICS code and one
best category per `classification_unit_id`, weighting sources by how
reliable they are and by whether they agree with each other. Anything picked
without solid support gets flagged in `manual_review` for a human to check.

Inputs (see config block below for exact paths/columns expected):

- **LLM classification** -- `09_llm_classifier/classify_web_results.py` output
- **OpenSecrets match** -- `08_alternative_pipeline/0805_contributor_matching_open_secrets.Rmd` output
- **Running-list / EDD+H1B / keyword match, and ML classifier** -- `08_alternative_pipeline/0804_classify_contributors.py`
  this also needs inputs from `07_ml_unmatched_classifier/0702_predict_unmatched.py`


# Config

```{r}
library(dplyr)
library(readr)
library(stringr)
library(tidyr)

source("../08_alternative_pipeline/standardization_helpers.R")
```

```{r}
# file paths (placeholders)
LLM_PATH  <- "09_inputs/classification_slim_full.csv"
RULE_PATH <- "../08_alternative_pipeline/08_outputs/classified_contributors.csv"
OS_PATH   <- "../08_alternative_pipeline/08_outputs/contributions_os_matches.csv"

# custom editoral label for each code
DESC_PATH <- "Campaign Finance Categories - Custom categories - updated.csv"

ML_HIGH_CONF_THRESHOLD <- 0.7

DETERMINISTIC_TIERS <- c("pre_classified", "employer_lookup", "running_list", "keyword match")

NOT_EMPLOYED_CODE <- "100"
FALLBACK_CODE     <- "99"
POLITICAL_CODES   <- c("88", "90") # pacs, candidate committees

# update these according to how upstream sources name columns
LLM_COLS <- c(
  unit_id    = "classification_unit_id", 
  name          = "canonical_name",
  employer      = "employer",
  code          = "naics_code_llm",
  naics_conf    = "naics_confidence",
  category      = "open_secrets_category",
  category_conf = "open_secrets_confidence",
  reasoning     = "naics_reasoning"
)

RULE_COLS <- c(
  unit_id         = "classification_unit_id",
  entity_id       = "entity_id",
  entity_type     = "entity_type",
  name            = "canonical_name",
  employer        = "employer",
  occupation      = "occupation",
  code            = "naics_code",
  label           = "naics_label",
  category        = "level2_category",
  source1         = "data_source_1",
  source2         = "data_source_2",
  ml_code         = "ml_naics_code",
  ml_conf         = "ml_naics_code_conf",
  ml_category     = "ml_level2_category",
  ml_category_conf = "ml_level2_category_conf"
)

# doesn't currently have a classification_unit_id
OS_COLS <- c(
  name       = "canonical_name",
  employer   = "employer",
  code       = "naics_code_os",
  category   = "level2_category",
  match_conf = "opensecrets_confidence"
)
```

```{r}
# fail if missing a required column
require_cols <- function(df, required_actual_names, path) {
  missing <- setdiff(required_actual_names, names(df))
  if (length(missing) > 0) {
    stop(sprintf(
      paste0(
        "%s is missing expected column(s): %s\n",
        "Available columns: %s\n",
        "If the upstream script renamed these, update the matching *_COLS ",
        "config block at the top of this file to match."
      ),
      path, paste(missing, collapse = ", "), paste(names(df), collapse = ", ")
    ))
  }
}

apply_colmap <- function(df, colmap) {
  for (logical_name in names(colmap)) {
    actual_name <- colmap[[logical_name]]
    if (actual_name %in% names(df)) {
      names(df)[names(df) == actual_name] <- logical_name
    } else {
      df[[logical_name]] <- NA
    }
  }
  df[, names(colmap), drop = FALSE]
}
```


# Load sources

```{r}
llm_required  <- unname(LLM_COLS[c("unit_id", "code")])
rule_required <- unname(RULE_COLS[c("unit_id", "code", "source1")])

llm_raw  <- read_csv(LLM_PATH, col_types = cols(.default = "c"))
rule_raw <- read_csv(RULE_PATH, col_types = cols(.default = "c"))
os_raw   <- read_csv(OS_PATH, col_types = cols(.default = "c"))

require_cols(llm_raw, llm_required, LLM_PATH)
require_cols(rule_raw, rule_required, RULE_PATH)

llm  <- apply_colmap(llm_raw, LLM_COLS)
rule <- apply_colmap(rule_raw, RULE_COLS)

rule <- rule %>% mutate(
  rule_tier = source1,
  ml_conf = as.numeric(ml_conf),
  ml_category_conf = as.numeric(ml_category_conf)
)
```

```{r}
# switch to a direct classification_unit_id join once 0805 carries it through

os_has_unit_id <- "classification_unit_id" %in% names(os_raw)

if (os_has_unit_id) { # join on classification_unit_id if available
  os_colmap <- c(OS_COLS, unit_id = "classification_unit_id")
  os <- apply_colmap(os_raw, os_colmap)
} else {
  # match on name + employer if classification_unit_id not available
  os <- apply_colmap(os_raw, OS_COLS) %>%
    mutate(
      name_key     = remove_unit_info(remove_pac_info(standardize_names(name))) %>% replace_business_text() %>% str_squish(),
      employer_key = standardize_occupation_employer(replace_business_text(remove_unit_info(remove_pac_info(standardize_names(employer)))))
    )
}
```


# Build the base spine and join sources

```{r}
# may need to re do this depending on how we end up tracking entities
base <- rule %>%
  select(unit_id, entity_id, entity_type, name, employer, occupation) %>%
  full_join(
    llm %>% select(unit_id, name, employer) %>% rename(llm_name = name, llm_employer = employer),
    by = "unit_id"
  ) %>%
  mutate(
    name     = coalesce(name, llm_name),
    employer = coalesce(employer, llm_employer)
  ) %>%
  select(-llm_name, -llm_employer)

combined <- base %>%
  left_join(
    rule %>% select(unit_id, code, label, category, source1, source2, rule_tier, ml_code, ml_conf, ml_category, ml_category_conf),
    by = "unit_id"
  ) %>%
  left_join(
    llm %>% select(unit_id, code, naics_conf, category, category_conf, reasoning) %>%
      rename(llm_code = code, llm_conf = naics_conf, llm_category = category,
             llm_category_conf = category_conf, llm_reasoning = reasoning),
    by = "unit_id"
  ) %>%
  rename(rule_code = code, rule_label = label, rule_category = category)

if (os_has_unit_id) {
  combined <- combined %>%
    left_join(
      os %>% select(unit_id, code, category, match_conf) %>%
        rename(os_code = code, os_category = category, os_conf = match_conf),
      by = "unit_id"
    )
} else {
  combined <- combined %>%
    mutate(
      name_key     = remove_unit_info(remove_pac_info(standardize_names(name))) %>% replace_business_text() %>% str_squish(),
      employer_key = standardize_occupation_employer(replace_business_text(remove_unit_info(remove_pac_info(standardize_names(employer)))))
    ) %>%
    left_join(
      os %>% select(name_key, employer_key, code, category, match_conf) %>%
        rename(os_code = code, os_category = category, os_conf = match_conf) %>%
        distinct(name_key, employer_key, .keep_all = TRUE),
      by = c("name_key", "employer_key")
    ) %>%
    select(-name_key, -employer_key)
}
```


# Resolve one best code (and, separately, one best category) per unit

```{r}
conf_is_high <- function(x) {
  if (length(x) == 0 || is.na(x)) return(FALSE)
  if (is.character(x)) return(tolower(x) == "high")
  is.numeric(x) && x >= ML_HIGH_CONF_THRESHOLD
}

resolve_one <- function(rule_val, rule_tier, rule_conf,
                         os_val, os_conf,
                         llm_val, llm_conf,
                         fallback_value = FALLBACK_CODE) {
  is_usable <- function(val) { # make sure its not the "unknown" fallback
    !is.na(val) && val != "" && (is.na(fallback_value) || val != fallback_value)
  }

  raw <- c(rule = rule_val, opensecrets = os_val, llm = llm_val)
  codes <- raw[vapply(raw, is_usable, logical(1))]

  rule_is_deterministic <- !is.na(rule_tier) && rule_tier %in% DETERMINISTIC_TIERS # came from reliable source
  rule_is_strong <- rule_is_deterministic ||
    (!is.na(rule_tier) && rule_tier == "ml prediction" && conf_is_high(rule_conf))

  # agreement between 2+ independent pipelines
  if (length(codes) >= 2) {
    tab <- table(codes)
    top <- names(tab)[which.max(tab)]
    if (max(tab) >= 2) {
      agreeing <- names(codes)[codes == top]
      return(list(value = top,
                  source = paste0("agreement (", paste(agreeing, collapse = "+"), ")"),
                  review = FALSE, reason = NA_character_))
    }
  }

  # single strong source wins outright
  # not sure we want to use high ML confidence outright - maybe update this
  if (is_usable(rule_val) && rule_is_strong) {
    tier_label <- if (rule_is_deterministic) rule_tier else "ml prediction, high confidence"
    return(list(value = rule_val, source = paste0("rule (", tier_label, ")"), review = FALSE, reason = NA_character_))
  }
  if (is_usable(os_val) && conf_is_high(os_conf)) {
    return(list(value = os_val, source = "opensecrets (high confidence)", review = FALSE, reason = NA_character_))
  }
  if (is_usable(llm_val) && conf_is_high(llm_conf)) {
    return(list(value = llm_val, source = "llm (high confidence)", review = FALSE, reason = NA_character_))
  }

  # exactly one low-confidence source available -- best guess, flag review
  if (length(codes) == 1) {
    src <- names(codes)[1]
    return(list(value = unname(codes[1]), source = paste0(src, " (low confidence, single source)"),
                review = TRUE, reason = paste0("only ", src, " produced a value, and confidence was not high")))
  }

  # 2+ sources but disagree and none is high-confidence -- take the most reliable pipeline available (rule > opensecrets > llm), flag for review
  # make sure this doesn't rank ML > LLM / opensecrets
  if (length(codes) > 1) {
    priority <- c("rule", "opensecrets", "llm")
    winner <- priority[priority %in% names(codes)][1]
    return(list(value = unname(codes[winner]),
                source = paste0(winner, " (disagreement, low confidence)"),
                review = TRUE,
                reason = paste0("sources disagreed: ",
                                 paste(names(codes), codes, sep = "=", collapse = ", "))))
  }

  # nothing reliable to use
  list(value = fallback_value, source = "uncategorized fallback", review = TRUE,
       reason = "no source produced a usable value")
}
```

```{r}
# hard overrides
code_override <- function(rule_val, os_val, llm_val, rule_tier) {
  vals <- c(rule = rule_val, opensecrets = os_val, llm = llm_val)

  # not-employed: any source saying so is authoritative
  if (any(vals == NOT_EMPLOYED_CODE, na.rm = TRUE)) {
    return(list(value = NOT_EMPLOYED_CODE, source = "not-employed override", review = FALSE, reason = NA_character_))
  }

  # political entities
  if (!is.na(rule_val) && !is.na(rule_tier) && rule_tier %in% DETERMINISTIC_TIERS && rule_val %in% POLITICAL_CODES) {
    return(list(value = rule_val, source = "political-entity rule", review = FALSE, reason = NA_character_))
  }

  NULL  # no override -- fall through to resolve_one()
}

# per-axis wrappers 
resolve_code_one <- function(rule_val, rule_tier, ml_conf, os_val, os_conf, llm_val, llm_conf) {
  override <- code_override(rule_val, os_val, llm_val, rule_tier)
  if (!is.null(override)) return(override)
  resolve_one(rule_val, rule_tier, ml_conf, os_val, os_conf, llm_val, llm_conf, fallback_value = FALLBACK_CODE)
}

resolve_category_one <- function(rule_val, rule_tier, ml_conf, os_val, os_conf, llm_val, llm_conf) {
  resolve_one(rule_val, rule_tier, ml_conf, os_val, os_conf, llm_val, llm_conf, fallback_value = NA_character_)
}
```

```{r}
resolved <- combined %>%
  rowwise() %>%
  mutate(
    .resolved_code = list(
      resolve_code_one(rule_code, rule_tier, ml_conf, os_code, os_conf, llm_code, llm_conf)
    ),
    .resolved_category = list(
      resolve_category_one(rule_category, rule_tier, ml_category_conf, os_category, os_conf, llm_category, llm_category_conf)
    )
  ) %>%
  ungroup() %>%
  unnest_wider(.resolved_code, names_sep = "_") %>%
  unnest_wider(.resolved_category, names_sep = "_") %>%
  rename(
    code_final            = .resolved_code_value,
    code_final_source     = .resolved_code_source,
    manual_review_code    = .resolved_code_review,
    manual_review_reason_code = .resolved_code_reason,
    category_final            = .resolved_category_value,
    category_final_source     = .resolved_category_source,
    manual_review_category    = .resolved_category_review,
    manual_review_reason_category = .resolved_category_reason
  ) %>%
  mutate(
    manual_review = manual_review_code | manual_review_category,
    manual_review_reason = case_when(
      manual_review_code & manual_review_category ~ paste0("code: ", manual_review_reason_code, " | category: ", manual_review_reason_category),
      manual_review_code                          ~ manual_review_reason_code,
      manual_review_category                      ~ manual_review_reason_category,
      TRUE                                         ~ NA_character_
    )
  )
```


# Attach the editorial description to the final code

```{r}
desc <- read_csv(DESC_PATH, col_types = cols(.default = "c"))

# expects columns "Sector" (code) and "Custom.Label" 
require_cols(desc, c("Sector", "Custom.Label"), DESC_PATH)
desc <- desc %>% transmute(code_final = Sector, code_final_description = Custom.Label)

resolved <- resolved %>%
  left_join(desc, by = "code_final")
```


# Review + write output

```{r}
resolved %>% count(code_final_source, sort = TRUE)
resolved %>% count(category_final_source, sort = TRUE)
cat(sprintf("%d / %d units flagged for manual review (%.1f%%)\n",
            sum(resolved$manual_review), nrow(resolved),
            100 * mean(resolved$manual_review)))
```

```{r}
out_cols <- c(
  "unit_id", "entity_id", "entity_type", "name", "employer", "occupation",
  "code_final", "code_final_source", "code_final_description",
  "category_final", "category_final_source",
  "manual_review", "manual_review_reason",
  "rule_code", "rule_label", "rule_tier", "rule_category",
  "os_code", "os_category", "os_conf",
  "llm_code", "llm_conf", "llm_category", "llm_category_conf", "llm_reasoning",
  "ml_code", "ml_conf", "ml_category", "ml_category_conf"
)
out_cols <- out_cols[out_cols %in% names(resolved)]
```

```{r}
#write_csv(resolved[, out_cols], "10_outputs/final_classifications.csv")
#write_csv(resolved, "10_outputs/final_classifications_full.csv")
```
