---
title: "Assign Final Classifications"
output: html_notebook
---

Combines every source the pipeline has for a contributor's industry
classification -- covers all $5k+ contributors, both direct givers to
races/props and $5k+ givers to PACs -- and picks one best NAICS code and one
best category per `uuid` (the per-contribution record identifier), weighting sources by how
reliable they are and by whether they agree with each other. Anything picked
without solid support gets flagged in `manual_review` for a human to check.

Inputs (see config block below for exact paths/columns expected):

- **LLM classification** -- `09_llm_classifier/classify_web_results.py` output
- **OpenSecrets match** -- `08_alternative_pipeline/0805_contributor_matching_open_secrets.Rmd` output
- **Running-list / EDD+H1B / keyword match, and ML classifier** -- `08_alternative_pipeline/0804_classify_contributors.py`
  this also needs inputs from `07_ml_unmatched_classifier/0702_predict_unmatched.py`


# Load data and packages

```{r}
library(dplyr)
library(readr)
library(stringr)
library(tidyr)

source("../08_alternative_pipeline/standardization_helpers.R")
source("resolve_classifications.R")
```

```{r}
# file paths

# Direct contributor LLM — uuid-keyed output from 0902; UUIDs match the current rule file (1:1).
#DIRECT_LLM_PATH <- "../09_llm_classifier/09_outputs/classification_input_combined_expanded_full_2026-08-16.csv"
DIRECT_LLM_PATH <- "../09_llm_classifier/09_outputs/classification_input_combined_expanded_full_2026-09-13.csv"
CLAUDE_PATH <- "../09_llm_classifier/09_outputs/claude_classification_results.csv"

# PAC contributor LLM files — entity-level output joined via classification_unit_id.
#UNIT_LLM_PATHS <- c(
#  "../09_llm_classifier/09_outputs/pac_classification_batch1_full_full_2026-08-13.csv",
#  "../09_llm_classifier/09_outputs/pac_md_classification_last_batch_full_2026-08-16.csv"
#)

RULE_PATH <- "../08_alternative_pipeline/08_outputs/classified_contributors_09-13-26.csv"
OS_PATH   <- "../08_alternative_pipeline/08_outputs/combined_contributions_os_matches_09-11-26.csv"
ML_PATH   <- "../../data/07_output_ml_classification/08_entities_with_ml_2026-08-15.csv"

input <- read.csv("../08_alternative_pipeline/08_outputs/classification_input_combined_09-08-26.csv")
```

```{r}
temp <- read.csv("../08_alternative_pipeline/08_outputs/classified_contributors_09-13-26.csv")
nrow(temp)
nrow(input)
```



```{r}
x <- read.csv("../08_alternative_pipeline/08_outputs/classification_input_combined_09-08-26.csv")
classified <- read.csv("../08_alternative_pipeline/08_inputs/already_classified_contributions_09-06-26.csv")

label_url <- "https://docs.google.com/spreadsheets/d/11QHvNJsdtMlc1YKo_iNvMB_Jfn5Ui-iYdlWhFYjSm9g/export?format=csv"
labels <- read.csv(label_url)
```

```{r}
nrow(classified)
nrow(input)
```


```{r}
classified %>% filter(Contributor.Name == "Alliance of California Ranchers and Farmers (Admin)")
```


# Configurations

```{r}
# update these according to how upstream sources name columns

LLM_COLS <- c(
  unit_id    = "uuid",
  name          = "standardized_name",
  employer      = "standardized_employer_name",
  code          = "naics_code_llm",
  naics_conf    = "naics_confidence",
  category      = "open_secrets_category",
  category_conf = "open_secrets_confidence",
  reasoning     = "naics_reasoning"
)

RULE_COLS <- c(
  unit_id         = "uuid",
  entity_id       = "entity_id",
  entity_type     = "entity_type",
  name            = "Contributor.Name",
  employer        = "employer",
  occupation      = "occupation",
  code            = "naics_code",
  label           = "naics_label",
  category        = "level2_category",
  source1         = "data_source_1",
  source2         = "data_source_2",
  ml_code         = "ml_naics_code",
  ml_conf         = "ml_naics_code_conf",
  ml_category     = "ml_level2_category",
  ml_category_conf = "ml_level2_category_conf"
)

OS_COLS <- c(
  unit_id    = "uuid",
  name       = "standardized_name",
  employer   = "standardized_employer_name",
  code       = "naics_code_os",
  category   = "level2_category",
  match_conf = "opensecrets_confidence"
)
```

```{r}
# require_cols() and apply_colmap() are defined in resolve_classifications.R (sourced above).
```


# Load sources


```{r}
rule_required <- unname(RULE_COLS[c("unit_id", "code", "source1")])
os_required   <- unname(OS_COLS[c("unit_id", "code")])

# Load rule first — needed to build join maps for LLM expansion
rule_raw <- read_csv(RULE_PATH, col_types = cols(.default = "c"))
os_raw   <- read_csv(OS_PATH, col_types = cols(.default = "c"))

# Direct contributor LLM: uuid join (1:1 exact — UUIDs are from the current data pull).
direct_llm_expanded <- read_csv(DIRECT_LLM_PATH, col_types = cols(.default = "c")) %>%
  filter(!is.na(naics_code_llm) & naics_code_llm != "")

cat(sprintf("DIRECT-LLM: %d rows\n", nrow(direct_llm_expanded)))

# PAC contributor LLM: entity-level files joined via classification_unit_id
# (PAC contributors share one classification_unit_id across contributions).
rule_uid_map <- rule_raw %>%
  filter(!is.na(classification_unit_id)) %>%
  select(uuid, classification_unit_id)

#unit_llm_raw <- bind_rows(lapply(UNIT_LLM_PATHS, \(p) read_csv(p, col_types = cols(.default = "c")))) %>%
#  filter(!is.na(naics_code_llm) & naics_code_llm != "") %>%
#  distinct(classification_unit_id, .keep_all = TRUE) %>%
#  select(-any_of("uuid"))  # drop PAC-traversal uuid before join adds rule uuid

#unit_llm_expanded <- unit_llm_raw %>%
#  inner_join(rule_uid_map, by = "classification_unit_id") %>%
#  select(-classification_unit_id)

#cat(sprintf("UNIT-LLM: %d unique entities -> %d contribution rows from %d file(s)\n",
#            nrow(unit_llm_raw), nrow(unit_llm_expanded), length(UNIT_LLM_PATHS)))

# Combine; direct takes priority when both cover the same uuid (shouldn't happen
# since direct/PAC are disjoint, but guards against edge cases).
llm_raw <- bind_rows(direct_llm_expanded) %>% # , unit_llm_expanded
  distinct(uuid, .keep_all = TRUE)
cat(sprintf("LLM total: %d classified uuids\n", nrow(llm_raw)))

llm_required <- unname(LLM_COLS[c("unit_id", "code")])
require_cols(llm_raw, llm_required, DIRECT_LLM_PATH)
require_cols(rule_raw, rule_required, RULE_PATH)
require_cols(os_raw, os_required, OS_PATH)

llm  <- apply_colmap(llm_raw, LLM_COLS)
rule <- apply_colmap(rule_raw, RULE_COLS)
os   <- apply_colmap(os_raw, OS_COLS)

# rule_code is already crosswalked to the custom sector scheme (0805); os_code
# and llm_code are not (see resolve_classifications.R)
llm$code <- apply_os_llm_crosswalk(llm$code)
os$code  <- apply_os_llm_crosswalk(os$code)

rule <- rule %>% mutate(
  rule_tier = source1,
  ml_conf = as.numeric(ml_conf),
  ml_category_conf = as.numeric(ml_category_conf)
)
```

```{r}
rule_raw %>% filter(data_source_1 != "pre_classified")
```

```{r}
rule_raw %>% filter(is.na(naics_code))
```


```{r}
nrow(rule_raw)
nrow(input)
```


```{r}
rule_raw %>% filter(data_source_1 != "pre_classified")
```

Really need to check Comcast Financial Agency Corporation -> so many rows



```{r}
direct_llm_expanded %>% group_by(uuid) %>% filter(n() > 1)
```

```{r}
# Bridge: uuid → search_key comes from the already-loaded expanded Gemini file.
# have to normalize search keys b/c gemini search keys contain city and state

norm_sk <- function(x) {
  parts <- strsplit(coalesce(x, ""), "|", fixed = TRUE)
  vapply(parts, function(p) paste0(p[1], "|", if (length(p) >= 2) p[2] else ""), character(1))
}

uuid_search_bridge <- direct_llm_expanded %>%
  select(uuid, search_key) %>%
  filter(!is.na(search_key) & search_key != "") %>%
  mutate(search_key_2 = norm_sk(search_key))

claude_raw <- read_csv(CLAUDE_PATH, col_types = cols(.default = "c")) %>%
  filter(!is.na(naics_code_claude) & naics_code_claude != "") %>%
  mutate(
    naics_code_claude = apply_os_llm_crosswalk(naics_code_claude),
    search_key_2 = norm_sk(search_key)
  )

claude_expanded <- uuid_search_bridge %>%
  left_join(
    claude_raw %>% select(search_key_2, claude_code = naics_code_claude,
                          claude_conf = naics_confidence, claude_reasoning = reasoning),
    by = "search_key_2"
  ) %>%
  select(-search_key_2)

cat(sprintf("CLAUDE: %d search_keys → %d / %d uuid rows covered\n",
            nrow(claude_raw), sum(!is.na(claude_expanded$claude_code)), nrow(claude_expanded)))


```

Claude classifications were only run on the most recent classifications. 

```{r}
# join ML predictions
# ML output has one row per unique classification unit (unique org name, or unique employer+occupation pair for individuals)
# join key is derived from the standardized employer/occupation columns -- these match between the ML output and rule_raw, unlike the more-processed `employer` col that apply_colmap uses.

if (file.exists(ML_PATH)) {
  ml_raw <- read_csv(ML_PATH, col_types = cols(.default = "c"))

  # Build ml_unit_key in ML output: orgs -> Contributor.Name, individuals -> employer|occupation
  # (employer/occupation here are standardized_employer_name/standardized_occupation from the 08 pipeline)
  ml_keyed <- ml_raw %>%
    mutate(
      ml_unit_key = if_else(
        tolower(entity_type) == "individual",
        paste0(trimws(employer), "|", trimws(occupation)),
        trimws(`Contributor.Name`)
      )
    ) %>%
    select(ml_unit_key,
           ml_code          = ml_naics_code,
           ml_conf          = ml_naics_code_conf,
           ml_category      = ml_level2_category,
           ml_category_conf = ml_level2_category_conf)

  # Derive the same key from rule_raw (before apply_colmap stripped columns),
  # using standardized_employer_name/standardized_occupation which match the ML output.
  rule_keys <- rule_raw %>%
    select(uuid, entity_type,
           standardized_employer_name, standardized_occupation,
           `Contributor.Name`) %>%
    mutate(
      ml_unit_key = if_else(
        tolower(entity_type) == "individual",
        paste0(trimws(standardized_employer_name), "|", trimws(standardized_occupation)),
        trimws(`Contributor.Name`)
      )
    ) %>%
    select(uuid, ml_unit_key)

  rule <- rule %>%
    select(-ml_code, -ml_conf, -ml_category, -ml_category_conf) %>%
    left_join(rule_keys, by = c("unit_id" = "uuid")) %>%
    left_join(ml_keyed, by = "ml_unit_key") %>%
    mutate(
      ml_conf          = as.numeric(ml_conf),
      ml_category_conf = as.numeric(ml_category_conf)
    ) %>%
    select(-ml_unit_key)

  cat(sprintf("ML predictions joined: %d / %d rows have ml_code\n",
              sum(!is.na(rule$ml_code)), nrow(rule)))
} else {
  warning(sprintf(
    "ML output not found at %s -- ml_* columns will be NA.\nRun: python code/07_ml_unmatched_classifier/run_ml_on_08_output.py",
    ML_PATH
  ))
}
```

Not sure which entities were not given an ML classification -> all self-employed/ 100? 

# Build the base spine and join sources

```{r}
# may need to re do this depending on how we end up tracking entities
base <- rule %>%
  select(unit_id, entity_id, entity_type, name, employer, occupation) %>%
  left_join(
    llm %>% select(unit_id, name, employer) %>% rename(llm_name = name, llm_employer = employer),
    by = "unit_id"
  ) %>%
  left_join(
    rule_raw %>% select(uuid, Recipient.Committee, Normalized) %>%
      rename(recipient_committee = Recipient.Committee, normalized_name = Normalized),
    by = c("unit_id" = "uuid")
  ) %>%
  mutate(
    # For PAC-traversal org rows the contributor name is often absent from the source
    # data; use the parent PAC name as a last-resort label so it's identifiable in review.
    name     = coalesce(
      name, llm_name,
      if_else(entity_type == "organization" & !is.na(recipient_committee),
              paste0("[contributor to PAC: ", recipient_committee, "]"), NA_character_)
    ),
    employer = coalesce(employer, llm_employer)
  ) %>%
  select(-llm_name, -llm_employer, -recipient_committee)

combined <- base %>%
  left_join(
    rule %>% select(unit_id, code, label, category, source1, source2, rule_tier, ml_code, ml_conf, ml_category, ml_category_conf),
    by = "unit_id"
  ) %>%
  left_join(
    llm %>% select(unit_id, code, naics_conf, category, category_conf, reasoning) %>%
      rename(llm_code = code, llm_conf = naics_conf, llm_category = category,
             llm_category_conf = category_conf, llm_reasoning = reasoning),
    by = "unit_id"
  ) %>%
  left_join(
    os %>% select(unit_id, code, category, match_conf) %>%
      rename(os_code = code, os_category = category, os_conf = match_conf),
    by = "unit_id"
  ) %>%
  rename(rule_code = code, rule_label = label, rule_category = category) %>%
  left_join(
    claude_expanded %>% select(uuid, claude_code, claude_conf, claude_reasoning),
    by = c("unit_id" = "uuid")
  )
```

```{r}
length(which(is.na(combined$llm_code)))
length(which(is.na(combined$claude_code)))
```

```{r}
combined %>% filter(is.na(combined$llm_code)) %>% distinct(rule_code)
```

```{r}
combined %>% filter(is.na(combined$llm_code) & (rule_code == "20" | is.na(rule_code)))
```
Not sure how these got left out of the data. 

# Resolve one best code (and, separately, one best category) per unit

- conf_is_high(), resolve_one(), code_override(), resolve_code_one(), and
- resolve_category_one() are defined in resolve_classifications.R (sourced above).

```{r}
resolved <- combined %>%
  rowwise() %>%
  mutate(
    .resolved_code = list(
      resolve_code_one(rule_code, rule_tier, ml_code, ml_conf, os_code, os_conf, llm_code, llm_conf)
    ),
    .resolved_category = list(
      resolve_category_one(rule_category, rule_tier, ml_category_conf, os_category, os_conf, llm_category, llm_category_conf)
    )
  ) %>%
  ungroup() %>%
  unnest_wider(.resolved_code, names_sep = "_") %>%
  unnest_wider(.resolved_category, names_sep = "_") %>%
  rename(
    code_final            = .resolved_code_value,
    code_final_source     = .resolved_code_source,
    manual_review_code    = .resolved_code_review,
    manual_review_reason_code = .resolved_code_reason,
    category_final            = .resolved_category_value,
    category_final_source     = .resolved_category_source,
    manual_review_category    = .resolved_category_review,
    manual_review_reason_category = .resolved_category_reason
  ) %>%
  mutate(
    manual_review = manual_review_code,
    manual_review_reason = manual_review_reason_code
  )

# Flag if LLM disagrees with the chosen final code — LLM is the most reliable
# independent signal, so any divergence warrants a human check.
# Exempt the 100-family (not-employed/retired): code_override locks those and
# LLM often misclassifies retired individuals.
resolved <- resolved %>%
  mutate(
    .llm_final_disagree = !is.na(llm_code) & !is.na(code_final) &
                          code_final != llm_code &
                          !grepl("^100", coalesce(code_final, "")) &
                          coalesce(rule_tier, "") != "pre_classified",
    manual_review = if_else(.llm_final_disagree, TRUE, manual_review),
    manual_review_reason = case_when(
      .llm_final_disagree & is.na(manual_review_reason) ~ "llm_disagrees_with_final",
      .llm_final_disagree ~ paste0(manual_review_reason, " | llm_disagrees_with_final"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.llm_final_disagree)

resolved <- resolved %>%
  mutate(
    .claude_gemini_disagree = !is.na(claude_code) & !is.na(llm_code) &
                              claude_code != llm_code &
                              !grepl("^100", coalesce(claude_code, "")),
    manual_review = if_else(.claude_gemini_disagree, TRUE, manual_review),
    manual_review_reason = case_when(
      .claude_gemini_disagree & is.na(manual_review_reason) ~ "claude_gemini_disagree",
      .claude_gemini_disagree ~ paste0(manual_review_reason, " | claude_gemini_disagree"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.claude_gemini_disagree)

# Flag 76 (Associations) and 77 (Unions) for subsector review — these top-level
# codes are placeholders; a reviewer must choose the appropriate sub-code.
resolved <- resolved %>%
  mutate(
    .needs_subsector = coalesce(code_final, "") %in% c("76", "77"),
    manual_review = if_else(.needs_subsector, TRUE, manual_review),
    manual_review_reason = case_when(
      .needs_subsector & is.na(manual_review_reason) ~ "subsector_needed: 76/77 requires more specific code",
      .needs_subsector ~ paste0(manual_review_reason, " | subsector_needed"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.needs_subsector)

cat(sprintf("Flagged for review: %d / %d rows\n",
            sum(coalesce(resolved$manual_review, FALSE)),
            nrow(resolved)))
```


```{r}
resolved %>% filter(!is.na(claude_code) & (claude_code != llm_code)) %>% select(name,employer,llm_code,claude_code) %>% distinct()
```

```{r}
# should be 0 rows 
resolved %>% filter(is.na(normalized_name))
```


```{r}
# conflicting aliases: flag rows where the same canonical entity (per normalized_name
# from 0802's Normalized column, covering all rows) or the same entity_id resolves
# to different final codes across its variants.
# Groups where every alias was already reviewed and accepted in the previous iteration
# (matched by name|employer|occupation) are skipped so they don't re-enter the queue.
.reviewed_alias_keys <- read_csv("10_inputs/manual_review_queue_reviewed.csv",
                                  col_types = cols(.default = "c")) %>%
  filter(grepl("alias_conflict", coalesce(manual_review_reason, "")),
         coalesce(manually_confirmed, "") == "TRUE") %>%
  mutate(.key = paste(name, employer, occupation, sep = "|")) %>%
  pull(.key) %>%
  unique()

.norm_conflict_summary <- resolved %>%
  filter(!is.na(code_final) & code_final != FALLBACK_CODE,
         !is.na(normalized_name)) %>%
  mutate(
    .code_cmp  = sub("^100[ab]$", "100", code_final),
    .match_key = paste(name, employer, occupation, sep = "|")
  ) %>%
  group_by(normalized_name) %>%
  summarise(
    n_codes        = n_distinct(.code_cmp),
    codes          = paste(sort(unique(.code_cmp)), collapse = ","),
    has_unreviewed = any(!(.match_key %in% .reviewed_alias_keys)),
    .groups        = "drop"
  ) %>%
  filter(n_codes > 1)

.norm_conflicts <- .norm_conflict_summary %>%
  filter(has_unreviewed) %>%
  pull(normalized_name)

.norm_conflicts_skipped <- .norm_conflict_summary %>%
  filter(!has_unreviewed)

if (nrow(.norm_conflicts_skipped) > 0) {
  cat(sprintf("Skipped %d normalized_name conflict(s) — all aliases previously reviewed:\n",
              nrow(.norm_conflicts_skipped)))
  print(.norm_conflicts_skipped %>% select(normalized_name, n_codes, codes))
}

.entity_conflict_summary <- resolved %>%
  filter(!is.na(code_final) & code_final != FALLBACK_CODE,
         !is.na(entity_id)) %>%
  mutate(
    .code_cmp  = sub("^100[ab]$", "100", code_final),
    .match_key = paste(name, employer, occupation, sep = "|")
  ) %>%
  group_by(entity_id) %>%
  summarise(
    n_codes        = n_distinct(.code_cmp),
    has_unreviewed = any(!(.match_key %in% .reviewed_alias_keys)),
    .groups        = "drop"
  ) %>%
  filter(n_codes > 1)

conflict_entity_ids <- .entity_conflict_summary %>%
  filter(has_unreviewed) %>%
  pull(entity_id)

cat(sprintf("%d normalized_name groups and %d entity_ids have alias conflicts\n",
            length(.norm_conflicts), length(conflict_entity_ids)))

resolved <- resolved %>%
  mutate(
    .conflict = (entity_id %in% conflict_entity_ids) |
                (!is.na(normalized_name) & normalized_name %in% .norm_conflicts),
    manual_review = if_else(.conflict, TRUE, manual_review),
    manual_review_reason = case_when(
      .conflict & is.na(manual_review_reason)
        ~ "alias_conflict: variants of this entity resolved to different codes",
      .conflict
        ~ paste0(manual_review_reason, " | alias_conflict"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.conflict)

cat(sprintf("Flagged for review: %d / %d rows\n",
            sum(coalesce(resolved$manual_review, FALSE)),
            nrow(resolved)
))
```

Flagged for review: 10188 / 42491 rows

```{r}
resolved %>%
  filter(grepl("alias_conflict", coalesce(manual_review_reason, "")),
         !grepl("cross_source", coalesce(manual_review_reason, ""))) %>%
  group_by(normalized_name) %>%
  summarise(n_rows = n(), n_codes = n_distinct(code_final), codes = paste(unique(code_final), collapse=",")) %>%
  filter(n_codes == 1) %>%
  arrange(desc(n_rows))

```


```{r}
# Is she in norm conflicts or entity_id conflicts?
cat("in .norm_conflicts:", "PEGGY JEROME" %in% .norm_conflicts, "\n")

resolved %>%
  filter(normalized_name == "PEGGY JEROME") %>%
  distinct(entity_id) %>%
  left_join(
    resolved %>%
      filter(!is.na(code_final) & code_final != FALLBACK_CODE) %>%
      mutate(.code_cmp = sub("^100[ab]$", "100", code_final)) %>%
      group_by(entity_id) %>%
      summarise(n_codes = n_distinct(.code_cmp), codes = paste(sort(unique(.code_cmp)), collapse=",")),
    by = "entity_id"
  )

# All codes across ALL PEGGY JEROME rows (not just flagged ones)
resolved %>%
  filter(normalized_name == "PEGGY JEROME") %>%
  count(code_final, rule_tier)

```


Grouping by the right entity_id here? 

```{r}
unique(resolved$manual_review_reason)
```


```{r}
resolved %>%
  filter(grepl("alias_conflict", coalesce(manual_review_reason, "")),
         !grepl("cross_source", coalesce(manual_review_reason, ""))) %>%
  group_by(normalized_name) %>%
  summarise(n_rows = n(), n_codes = n_distinct(code_final), codes = paste(unique(code_final), collapse=",")) %>%
  arrange(desc(n_rows))

```


## Patch: suspected entity-type mismatches

Orgs mislabeled as `entity_type = "individual"` end up in the individual review
queue because they have no employer. Detect them by name keywords and reroute to
the main review queue. This is a patch until 0802 entity-type detection is fixed.

- cvs health, including aggregated contributions
- majestic realty co. and edward p. roski, jr.
- southern sierras chapter, neca

```{r}
ORG_NAME_RE <- paste0(
  "LLC|\\bInc\\b|\\bCorp\\b|\\.L\\.P\\.|L\\.L\\.C\\.|",
  "Political Action|Contributions|",
  "\\bAssociation\\b|\\bFoundation\\b|\\bChapter\\b|",
  "\\bUnion\\b|\\bCouncil\\b|\\bCommittee\\b|\\bFund\\b|",
  "Holdings|Properties|\\bManagement\\b|",
  "\\bHospital\\b|Medical|\\bBank\\b|Insurance|Realty|Health|",
  "Construction|NECA|IBEW|\\bAFL\\b|\\bCIO\\b|\\bSEIU\\b|",
  "Contractors|Industries|Manufacturing|Supply|",
  "Resources|Energy|Electric|Communications|",
  "\\bMedia\\b|Networks|Technologies|Solutions|\\bServices\\b|",
  "\\bSystems\\b|\\bGroup\\b|Partners|Enterprises|\\bTrust\\b|",
  "International|\\bNational\\b|\\bCalifornia\\b|\\bAmerican\\b|FedEx"
)

resolved <- resolved %>%
  mutate(
    # entity_id prefix is unreliable (some orgs got "indiv_" IDs from 0802 detection),
    # so name keywords are the primary signal here.
    suspected_org = (tolower(entity_type) == "individual") &
      grepl(ORG_NAME_RE, coalesce(name, ""), ignore.case = TRUE, perl = TRUE),
    manual_review = if_else(suspected_org, TRUE, manual_review),
    manual_review_reason = case_when(
      suspected_org & is.na(manual_review_reason) ~ "suspected_entity_type_mismatch",
      suspected_org ~ paste0(manual_review_reason, " | suspected_entity_type_mismatch"),
      TRUE ~ manual_review_reason
    )
  )

cat(sprintf("Suspected entity-type mismatches: %d rows\n", sum(coalesce(resolved$suspected_org, FALSE))))
```


```{r}
resolved %>% filter(suspected_org == TRUE)
```


- FedExPAC, FedEx: shouldn't have been grabbed because parent pac is a federal PAC
  - make sure we update to not grabbing this

## Cross-source entity conflict detection

The same contributor can appear as both a direct contributor and a PAC contributor,
each getting a different `entity_id` because entity resolution runs independently
per path. Detect cross-source matches by normalized name (orgs) or
normalized name|employer (individuals) and flag rows where the matched group
resolves to different final codes.

```{r}
resolved <- resolved %>%
  mutate(
    entity_norm_key = tolower(trimws(coalesce(name, ""))) %>%
      str_replace_all("[^a-z0-9|]", " ") %>%
      str_squish()
  )

# 100/100a/100b are the same entity-type family; normalize to "100" before
# counting distinct codes so those sub-code differences don't trigger a false conflict.
cross_source_summary <- resolved %>%
  filter(entity_norm_key != "") %>%
  mutate(.code_cmp = sub("^100[ab]$", "100", coalesce(code_final, ""))) %>%
  group_by(entity_norm_key) %>%
  summarise(
    n_entity_ids  = n_distinct(entity_id, na.rm = TRUE),
    n_final_codes = n_distinct(.code_cmp[.code_cmp != "" & .code_cmp != FALLBACK_CODE]),
    .groups = "drop"
  ) %>%
  filter(n_entity_ids > 1 & n_final_codes > 1)

# If one entity in a conflict group was pre_classified by John, propagate that
# code to non-locked rows with a compatible employer. "Compatible" means the
# employers match (after normalization) OR at least one side has no employer 
# Rows in conflict groups whose employer doesn't
# match John's row are still flagged for manual review.
# (source2 carries the reviewer name from already_classified_contributions.csv)
norm_emp <- function(x) {
  tolower(trimws(coalesce(x, ""))) %>%
    str_replace_all("[^a-z0-9]", " ") %>%
    str_squish()
}

pre_classified_john_codes <- resolved %>%
  filter(
    entity_norm_key %in% cross_source_summary$entity_norm_key,
    rule_tier == "pre_classified",
    grepl("john", tolower(coalesce(source2, "")))
  ) %>%
  mutate(.john_emp = norm_emp(employer)) %>%
  group_by(entity_norm_key) %>%
  slice(1) %>%
  ungroup() %>%
  select(entity_norm_key, .propagated_code = code_final, .john_emp)

resolved <- resolved %>%
  mutate(.emp_norm = norm_emp(employer)) %>%
  left_join(pre_classified_john_codes, by = "entity_norm_key") %>%
  mutate(
    .employer_ok  = .emp_norm == "" | coalesce(.john_emp, "") == "" | .emp_norm == .john_emp,
    .in_conflict  = entity_norm_key %in% cross_source_summary$entity_norm_key,
    .not_locked   = coalesce(code_final_source, "") != "pre_classified (locked)",
    cross_source_conflict = .in_conflict & (is.na(.propagated_code) | !.employer_ok),
    .propagated   = .in_conflict & !is.na(.propagated_code) & .not_locked & .employer_ok,
    code_final        = if_else(.propagated, .propagated_code, code_final),
    code_final_source = if_else(.propagated, "pre_classified (propagated: John)", code_final_source),
    manual_review     = if_else(cross_source_conflict, TRUE, manual_review),
    manual_review_reason = case_when(
      cross_source_conflict & is.na(manual_review_reason) ~ "cross_source_entity_conflict",
      cross_source_conflict ~ paste0(manual_review_reason, " | cross_source_entity_conflict"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.propagated_code, -.in_conflict, -.not_locked, -.emp_norm, -.john_emp, -.employer_ok)

cat(sprintf(
  "Cross-source entity conflicts: %d unique norm keys → %d rows flagged | %d rows auto-resolved by John pre_classified propagation\n",
  nrow(cross_source_summary),
  sum(coalesce(resolved$cross_source_conflict, FALSE)),
  sum(coalesce(resolved$.propagated, FALSE))
))
```

```{r}
resolved %>%
  filter(grepl("alias_conflict", coalesce(manual_review_reason, "")),
         !grepl("cross_source", coalesce(manual_review_reason, ""))) %>%
  group_by(normalized_name) %>%
  summarise(n_rows = n(), n_codes = n_distinct(code_final), codes = paste(unique(code_final), collapse=",")) %>%
  filter(n_codes == 1) %>%
  arrange(desc(n_rows))

```

```{r}
resolved %>% filter(normalized_name == "ELIZABETH D SIMONS")
```
Maybe want her to be reviewed? skip for now. 

```{r}
input %>% filter(Normalized == "ELIZABETH D SIMONS")
```


```{r}
resolved %>% filter(cross_source_conflict == TRUE) %>% arrange(name) %>% select(name,employer,occupation, code_final, rule_code, os_code, llm_code, ml_code)
```


WEST, BENJAMIN 

```{r}
review_queue %>% filter(grepl("WEST, BENJAMIN",toupper(Contributor.Name)))
resolved %>% filter(grepl("WEST, BENJAMIN",toupper(name)))
```


```{r}
# Review: all rows where code was propagated from John's pre_classified entry.
# Check that the propagated code is correct before accepting.
propagated_review <- resolved %>%
  filter(coalesce(.propagated, FALSE)) %>%
  select(name, employer, occupation, entity_norm_key,
         code_final, code_final_source,
         rule_code, rule_tier, source2,
         llm_code, os_code, ml_code)

cat(sprintf("Propagated-code review: %d rows\n", nrow(propagated_review)))
propagated_review
```

```{r}
resolved <- resolved %>% select(-.propagated)
```


Don't need this anymore right? 
```{r}
# Propagate 100a / 100b sub-codes from already_classified to individuals who
# resolved to plain "100" (not-employed/retired) but whose exact Contributor.Name
# appears in the pre-classified data with a Megadonors / Major Donors sub-code.
sub_100_lookup <- classified %>%
  mutate(
    .match_name = trimws(toupper(coalesce(Contributor.Name, ""))),
    .sub_code = case_when(
      "revised_category" %in% names(.) &
        trimws(coalesce(revised_category, "")) == "Megadonors"   ~ "100a",
      "revised_category" %in% names(.) &
        trimws(coalesce(revised_category, "")) == "Major Donors" ~ "100b",
      trimws(coalesce(code_final, "")) == "100a" ~ "100a",
      trimws(coalesce(code_final, "")) == "100b" ~ "100b"
    )
  ) %>%
  filter(!is.na(.sub_code), .match_name != "") %>%
  distinct(.match_name, .keep_all = TRUE) %>%
  select(.match_name, .sub_code)

resolved <- resolved %>%
  mutate(.name_upper = trimws(toupper(coalesce(name, "")))) %>%
  left_join(sub_100_lookup, by = c(".name_upper" = ".match_name")) %>%
  mutate(
    .inherit_sub = code_final == NOT_EMPLOYED_CODE &
                   tolower(entity_type) == "individual" & !is.na(.sub_code),
    code_final        = if_else(.inherit_sub, .sub_code, code_final),
    code_final_source = if_else(.inherit_sub, "pre_classified (100 sub-code inherited)", code_final_source),
    manual_review     = if_else(.inherit_sub, FALSE, manual_review)
  ) %>%
  select(-.name_upper, -.sub_code)

cat(sprintf("100 sub-code inherited: %d rows\n",
            sum(coalesce(resolved$.inherit_sub, FALSE))))
```

```{r}
# Review: propagated 100→100a/100b rows
resolved %>%
  filter(coalesce(.inherit_sub, FALSE)) %>%
  select(name, employer, occupation, code_final, code_final_source, rule_code, llm_code)
```

```{r}
resolved <- resolved %>% select(-.inherit_sub)
```


### flag self-employed individuals:

```{r}
# Flag self-employed individuals for review: employer field contains "self-employed"
# (and common variants) or is exactly "self", OR the contributor's own name appears
# in the employer field. Suppress for contributors already reviewed and confirmed.
.reviewed_confirmed_keys <- read_csv("10_inputs/manual_review_queue_reviewed.csv",
                                      col_types = cols(.default = "c")) %>%
  filter(coalesce(manually_confirmed, "") == "TRUE") %>%
  mutate(.key = paste(name, employer, occupation, sep = "|")) %>%
  pull(.key) %>%
  unique()

resolved <- resolved %>%
  mutate(
    .name_key = str_squish(str_replace_all(tolower(trimws(coalesce(name, ""))),     "[^a-z0-9]", " ")),
    .emp_key  = str_squish(str_replace_all(tolower(trimws(coalesce(employer, ""))), "[^a-z0-9]", " ")),
    .se_key   = paste(name, employer, occupation, sep = "|"),
    .self_emp = tolower(entity_type) == "individual" &
      !(.se_key %in% .reviewed_confirmed_keys) & (
        grepl("\\bself[- ]?employ", coalesce(employer, ""), ignore.case = TRUE, perl = TRUE) |
        grepl("^self$", trimws(coalesce(employer, "")), ignore.case = TRUE) |
        (nchar(.name_key) >= 5 & .emp_key != "" & mapply(function(pat, txt) grepl(pat, txt, fixed = TRUE), .name_key, .emp_key, USE.NAMES = FALSE))
      ),
    manual_review = if_else(.self_emp, TRUE, manual_review),
    manual_review_reason = case_when(
      .self_emp & is.na(manual_review_reason) ~ "self_employed",
      .self_emp ~ paste0(manual_review_reason, " | self_employed"),
      TRUE ~ manual_review_reason
    )
  ) %>%
  select(-.name_key, -.emp_key, -.se_key, -.self_emp)

cat(sprintf("Self-employed flagged: %d rows\n",
            sum(grepl("self_employed", coalesce(resolved$manual_review_reason, "")))))
```


```{r}
resolved_test %>% filter(grepl("self_employed",manual_review_reason)) %>% distinct(employer)
```



### take pre-classified rows out of the review queue: 

```{r}
# pre_classified rows are already human-reviewed and should not be
# in the review queue in most cases. Exceptions:
#   - alias_conflict: a new name/employer variant may contain new information
#     that needs reconciling with the pre_classified code.
#   - cross_source_conflict: two pre_classified entries for the same entity
#     resolved to different codes — a prior-review contradiction worth flagging.
# All other review flags (LLM disagrees, 76/77 subsector, etc.) are cleared for
# pre_classified rows since the code has already been human-confirmed.
resolved <- resolved %>%
  mutate(
    manual_review = if_else(
      coalesce(rule_tier, "") == "pre_classified" &
        !grepl("alias_conflict|cross_source_entity_conflict|self_employed",
               coalesce(manual_review_reason, ""),
               fixed = FALSE),
      FALSE,
      manual_review
    ),
    manual_review_reason = if_else(
      !coalesce(manual_review, FALSE),
      NA_character_,
      manual_review_reason
    )
  )
cat(sprintf("After pre_classified guard — flagged: %d / %d rows\n",
            sum(coalesce(resolved$manual_review, FALSE)), nrow(resolved)))
```

```{r}
resolved %>% filter(normalized_name == "ELIZABETH D SIMONS")
```


DON'T include if difference is 100 vs 100a

TELACU Industries, Inc.
Telacu Industries Inc
Only one of these matches to rule_code and not sure why... 

Hernandez, Jose	TIERRA LUNA ENGINEERING	FARMER AND AEROSPACE ENGINEER	
Hernandez, Jose	NONE	CEO	
Hernandez, Jose	IDEATE CALIFORNIA	CEO

### data checkes

```{r}
# missing LLM and ML codes
resolved %>% filter(is.na(llm_code))
resolved %>% filter(is.na(ml_code))
```

```{r}
resolved %>% filter(grepl("telacu",tolower(name)))
```


```{r}
unique(resolved$code_final_source)
```

### misc sources

```{r}
resolved %>% filter(code_final_source == "rule (running_list)" & code_final != llm_code) %>% select(name, employer, rule_code, source1, llm_code, ml_code, manual_review) %>% distinct() 
```

### unresolved

```{r}
resolved %>% filter(code_final == "99" & is.na(rule_code)) %>% select(code_final,rule_code,claude_code,llm_code,llm_category_conf,llm_conf,manual_review_reason)
```

```{r}
resolved %>% filter(code_final == "99")
```

```{r}
colnames(resolved)
```


error: no rule; sources inconsistent | llm_disagrees_with_final
this shouldn't be labeled this way. 
so many without a rule_code -> is that right? 


```{r}
resolved %>% filter(code_final_source == "uncategorized fallback")
```


### political entities

```{r}
resolved %>% filter(code_final_source == "political-entity rule")
```

Nothing should end up as a PAC.... 
- Abbott Laboratories Employee Political Action Committee - C00040279 -> why is this in here? missing id? 

How did these end up in here? 
- the AFSCME General Fund, CA District Council 36 of
- Cope Healthcare Consulting, Inc.


### ML / LLM aggreement

- look at when this DOESN'T match final code

```{r}
resolved %>% filter(llm_code == ml_code)
```


```{r}
resolved %>% filter(llm_code == ml_code & llm_code != code_final) %>% select(name, employer, rule_code, source1, source2, llm_code, ml_code, manual_review, manual_review_code) %>% distinct() 
```

identity_override doesn't seem realible.. e.g. AI and Union

```{r}
resolved %>% filter(llm_code != rule_code & manual_review == FALSE & source1 != "pre_classified") %>% select(name, employer, rule_code, source1, llm_code, code_final, code_final_source, manual_review) %>% distinct() 
```

REMOVE from ID override
- software
- AI


### llm code as source

```{r}
resolved %>% filter(code_final_source == "llm (high confidence)" ) %>% select(name, rule_label, llm_code, ml_code, manual_review) %>% distinct() %>% select(rule_label) %>% distinct()
resolved %>% filter(code_final_source == "llm (high confidence)" ) %>% select(name, employer, rule_label, llm_code, ml_code, manual_review, manual_review_code) %>% distinct()
```

- LLM code ONLY used when ML and LLM disagree and no rule is available


### ml code as source

```{r}
resolved %>% filter(code_final_source == "ml only") %>% select(name, rule_label, llm_code, ml_code, manual_review) %>% distinct()
```

- ML code only used when no rule or LLM code available, manual_review is always true.

```{r}
unique(pac_final$os_conf)
```

```{r}
resolved %>% filter(name == "Dentaquest LLC")
```



### NA names

```{r}
resolved %>% filter(is.na(name))
```

All of these are the same entity. not sure what's happening.

```{r}
input %>% filter(uuid == "67e57677-2cf7-49d1-911c-38e77179168e")
input %>% filter(uuid == "1d751068-5c55-4b25-8873-fa41581fa154")
input %>% filter(uuid == "279f330b-8517-4391-a626-338efccefbd8")
```
- Not sure where these are coming from...

```{r}
resolved %>% filter(code_final_source == "rule (running_list)" & llm_code != rule_code)
```

```{r}
resolved %>% filter(code_final_source == "rule (identity_override)" & llm_code != rule_code)
```



```{r}
resolved %>% filter(code_final_source == "rule (employer_lookup)")
```

Everything else looks good
Huynh, Ethan	SELF	PHARMACIST -> classified by employer lookup???
Tuck, Nathaniel	SELF	PRODUCER -> same situation
Thomas, Elizabeth Allison	SELF	PRODUCER -> 11

```{r}
llm_raw %>% filter(uuid == "35bef780-2410-41e4-91dd-4347aa396e40")
```



```{r}
resolved %>% filter(code_final_source == "llm only") %>% select(name,employer,occupation,ml_code,llm_code,ml_conf)
```
- not that many right now that are LLM only, but will defs want to check these are correct... why no ML?
- should flag ML, LLM aggreement as being the code final source 
```{r}
length(which(resolved$manual_review == TRUE))
length(which(resolved$manual_review != TRUE))
```


```{r}
unique(resolved$manual_review_reason)
```

```{r}
resolved %>% filter(manual_review_reason == "no source produced a usable value" & !is.na(llm_code))
```

25k rows where no source had a usable value??? because of missing pac classifications? 
- almost all because LLM codes were missing

```{r}
resolved %>% filter(grepl("alias_conflict",manual_review_reason)) %>% arrange(entity_id)
```

Conflicts in aliases: 
- many because some rows show retired and others don't
- how to structure review queue so that aliases can be split if they are different? add a column to add groups to divide? 
- sometimes there are conflicts with pre-classified if employer is different, okay to flag those? probs have to


```{r}
unique(resolved$manual_review_reason)
```


```{r}
resolved %>% filter((rule_code != ml_code) & (ml_code == llm_code) &!is.na(llm_code)) %>% select(name, employer, source1,source2, entity_type,rule_code, ml_code, llm_code)
```

```{r}
unique(resolved$ml_code)
unique(resolved$source1)
unique(resolved$source2)
```
Not sure how much we can rely on employer regex

```{r}
colnames(resolved)
```

```{r}
resolved %>% filter(code_final == "99") %>% select(name,rule_code,)
```


```{r}
resolved %>% filter(manual_review == TRUE) %>% select(name,employer,occupation,manual_review_reason) %>% distinct()
```

- this will go down when we add more LLM classifications.. 

```{r}
length(unique(resolved[which(resolved$manual_review == TRUE),]$entity_id))
```

so many reviews to do... 

```{r}
resolved %>% filter(name == "Uber Technologies, Inc.")
```


# Attach the editorial description to the final code

```{r}
#desc <- read_csv(DESC_PATH, col_types = cols(.default = "c"))

# expects columns "Sector" (code) and "Custom.Label" 
#require_cols(desc, c("Sector", "Custom.Label"), DESC_PATH)
labels <- labels %>% select(sector,sector_description) %>% distinct()
desc <- labels %>% transmute(code_final = sector, code_final_description = sector_description)


resolved <- resolved %>%
  left_join(desc, by = "code_final")
```


### add category for unitemized contributions

```{r}
resolved %>% filter(tolower(name) == "unitemized contributions") %>% select(code_final,name)
```

```{r}
resolved %>% filter(grepl("unitemized",tolower(name))) %>% select(code_final,name)
```

!! note that some contributor names associated w/ unitemized contributions are labeled as association rather than union -> these should probs all be a union? 

### figure out computation as a part of the whole...that happens later


### check which categories were chosen

- rule_code
- source1
- source2
- rule_tier
- llm_code
- os_code

```{r}
unique(resolved$rule_code)
```


```{r}
resolved %>% filter(os_code != final_code) %>% select(unit_id, name, employer, code_final, rule_code,source1,source2, llm_code,os_code)
```

Need to map the pre-classified codes to our current codes. 
why does source1 = employer_lookup and source2 = John? 

```{r}
llm %>% filter(unit_id == "7738b3cd-eb0f-415a-8dc8-90b722af8cc4")
```


```{r}
colnames(resolved)
```



# Review + write output

```{r}
resolved %>% count(code_final_source, sort = TRUE)
resolved %>% count(category_final_source, sort = TRUE)
cat(sprintf("%d / %d units flagged for manual review (%.1f%%)\n",
            sum(resolved$manual_review), nrow(resolved),
            100 * mean(resolved$manual_review)))
```

```{r}
out_cols <- c(
  "unit_id", "entity_id", "entity_type", "name", "employer", "occupation",
  "code_final", "code_final_source", "code_final_description",
  "category_final", "category_final_source",
  "manual_review", "manual_review_reason",
  "suspected_org", "cross_source_conflict",
  "rule_code", "rule_label", "rule_tier", "rule_category",
  "os_code", "os_category", "os_conf",
  "llm_code", "llm_conf", "llm_category", "llm_category_conf", "llm_reasoning",
  "claude_code", "claude_conf", "claude_reasoning",
  "ml_code", "ml_conf", "ml_category", "ml_category_conf"
)
out_cols <- out_cols[out_cols %in% names(resolved)]

write_csv(resolved[, out_cols], "10_outputs/final_classifications_before_review_09-13-26.csv")
write_csv(resolved, "10_outputs/final_classifications_full_before_review_09-13-26.csv")
```


# Confidence review sample 

Stratified random sample of confident (non-flagged) classifications, one group
per `code_final_source`, to check how reliable each source actually is in practice.

```{r}
SAMPLE_N_PER_SOURCE <- 20  # rows to draw per code_final_source bucket

set.seed(42)
confidence_sample <- resolved %>%
  filter(
    !manual_review,
    !is.na(code_final),
    code_final != FALLBACK_CODE,
    code_final != NOT_EMPLOYED_CODE,       # these go in the individual review queue
    code_final_source != "pre_classified (locked)"  # already human-reviewed
  ) %>%
  group_by(code_final_source) %>%
  slice_sample(n = SAMPLE_N_PER_SOURCE) %>%
  ungroup()

cat(sprintf("Confidence sample: %d rows across %d source buckets\n",
            nrow(confidence_sample), n_distinct(confidence_sample$code_final_source)))
confidence_sample %>% count(code_final_source, sort = TRUE)
```


```{r}
sample_cols <- c(
  "unit_id", "entity_id", "entity_type", "name", "employer", "occupation",
  "code_final", "code_final_source", "code_final_description",
  "rule_code", "rule_tier",
  "llm_code", "llm_conf", "llm_reasoning",
  "claude_code", "claude_conf", "claude_reasoning",
  "os_code", "os_conf",
  "ml_code", "ml_conf"
)

write_csv(
  confidence_sample %>% select(any_of(sample_cols)),
  "10_outputs/classification_confidence_sample.csv"
)
```



# Individual review queue (code = 100) - SKIP (no longer separating into subgroups)

All individuals resolved to code 100 (not employed / retired / homemaker /
student), deduplicated across name + employer + occupation + location, with a
stable `name_id` hash of `Contributor.Name`. Excludes individuals already in
`already_classified_contributions.csv`.


```{r}
library(digest)

ind_review_raw <- resolved %>%
  filter(code_final == NOT_EMPLOYED_CODE,
         tolower(entity_type) == "individual",
         !coalesce(suspected_org, FALSE)) %>%
  left_join(
    rule_raw %>% select(uuid, Contributor.City, Contributor.State, Contributor.Zip.Code),
    by = c("unit_id" = "uuid")
  ) %>%
  filter(!(name %in% classified$Contributor.Name))

ind_review_queue <- ind_review_raw %>%
  select(name, employer, occupation,
         Contributor.City, Contributor.State, Contributor.Zip.Code) %>%
  distinct(across(everything(), tolower), .keep_all = TRUE) %>%
  arrange(name) %>%
  mutate(name_id = sapply(name, \(x) substr(digest(x, algo = "xxhash32"), 1, 8)))

cat(sprintf("Individual review queue: %d unique name/employer/occupation/location combos\n",
            nrow(ind_review_queue)))
```

```{r}
ind_review_queue %>% filter(name %in% tolower(classified$Contributor.Name))
```

```{r}
resolved %>% filter(tolower(name) == "quillin, patty") %>% select(code_final)
```


CVS Health, Including Aggregated Contributions... another indep expenditure example...
some with entity_type = org have a entity_id starting with indiv_...not sure why that would be, was it reassigned at some point? 
Just do a manual override? 
  - think that's the best for right now. send to manual review maybe? 
Need to update org flag... 


```{r}
write_csv(ind_review_queue, "10_outputs/individual_review_queue.csv")
```


### check individual overlap


```{r}
resolved %>% filter((tolower(name) %in% tolower(classified$Contributor.Name)) & source1 != "pre_classified")
```




# Combined manual review queue

All rows flagged for manual review plus the stratified confidence sample, in a
single file for a reviewer. Includes raw `Contributor.Name` / `Contributor.Employer`
fields (pre-standardization) alongside standardized columns — the raw fields are
the most stable keys for matching reviewed decisions back to future data pulls.

```{r}
# Step 1: derive recipient label and contribution type from raw contribution columns.
rule_raw_annotated <- rule_raw %>%
  mutate(
    # recipient: ballot measure name, candidate name, or committee name (in that priority)
    recipient_label = case_when(
      Ballot.Measure.Contribution == "Y" ~
        coalesce(na_if(trimws(Ballot.Measure), ""), na_if(trimws(Recipient.Committee), "")),
      !is.na(Recipient.Name) & trimws(Recipient.Name) != "" ~ trimws(Recipient.Name),
      TRUE ~ na_if(trimws(Recipient.Committee), "")
    ),
    # contribution type: pac_contributor (came via PAC traversal) / Ballot.Measure / office
    contribution_type = case_when(
      !is.na(parent_pac_id) & parent_pac_id != "" ~ "pac_contributor",
      Ballot.Measure.Contribution == "Y"           ~ "Ballot.Measure",
      Candidate.Contribution == "Y"                ~ "office",
      TRUE ~ NA_character_
    )
  )

# Step 2: join raw pre-standardization fields and build a stable dedup key.
# Key = raw Contributor.Name | raw Contributor.Employer | raw Contributor.Occupation
# for ALL entity types — avoids dependency on entity_type (which can change) and
# entity_id (which is reassigned each run). These raw fields come directly from
# the source data and are stable across processing and standardization changes.
review_queue_raw <- bind_rows(
  resolved %>%
    filter(coalesce(manual_review, FALSE)) %>%
    mutate(.queue_source = "flagged"),
  confidence_sample %>%
    mutate(.queue_source = "confidence_sample")
) %>%
  distinct(unit_id, .keep_all = TRUE) %>%
  left_join(
    rule_raw %>% select(uuid, any_of(c(
      "Contributor.Name", "Contributor.Employer", "Contributor.Occupation", "Contributor.Zip.Code"
    ))),
    by = c("unit_id" = "uuid"),
    suffix = c("", ".raw")
  )

# Ensure raw columns exist even if missing from rule_raw schema
for (.col in c("Contributor.Name", "Contributor.Employer", "Contributor.Occupation", "Contributor.Zip.Code")) {
  if (!.col %in% names(review_queue_raw)) review_queue_raw[[.col]] <- NA_character_
}

review_queue_raw <- review_queue_raw %>%
  mutate(
    dedup_key = if_else(
      tolower(entity_type) == "individual",
      paste0(
        trimws(coalesce(`Contributor.Name`, name, "")), "|",
        trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
        trimws(coalesce(`Contributor.Occupation`, occupation, "")), "|",
        trimws(coalesce(`Contributor.Zip.Code`, ""))
      ),
      paste0(
        trimws(coalesce(`Contributor.Name`, name, "")), "|",
        trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
        trimws(coalesce(`Contributor.Occupation`, occupation, ""))
      )
    )
  )

# Step 3: per-dedup-group aggregates — pipe-join unique recipient, contribution type,
# and location variants so a reviewer sees every race/prop this entity appears in.
geo_variants <- review_queue_raw %>%
  select(unit_id, dedup_key) %>%
  left_join(
    rule_raw_annotated %>% select(uuid, recipient_label, contribution_type,
                                  Contributor.City, Contributor.State, Contributor.Zip.Code),
    by = c("unit_id" = "uuid")
  ) %>%
  group_by(dedup_key) %>%
  summarise(
    n_contributions    = n(),
    recipients         = paste(sort(unique(na.omit(trimws(recipient_label)))),   collapse = " | "),
    contribution_types = paste(sort(unique(na.omit(trimws(contribution_type)))), collapse = " | "),
    cities    = paste(sort(unique(na.omit(trimws(Contributor.City)))),     collapse = " | "),
    states    = paste(sort(unique(na.omit(trimws(Contributor.State)))),    collapse = " | "),
    zip_codes = paste(sort(unique(na.omit(trimws(Contributor.Zip.Code)))), collapse = " | "),
    .groups = "drop"
  )

# Step 4: dedup to one row per classification unit; prefer rows with the most source coverage.
# Then sort for reviewer: group by name so alias variants appear together, then by reason
# so alias_conflict rows cluster within a name, then by employer to sub-group within a name.
review_queue <- review_queue_raw %>%
  left_join(geo_variants, by = "dedup_key") %>%
  #arrange(desc(!is.na(llm_code)), desc(!is.na(rule_code)), desc(!is.na(ml_code))) %>%
  distinct(dedup_key, .keep_all = TRUE) %>%
  arrange(name, manual_review_reason, employer)

review_queue_cols <- c(
  "dedup_key",
  "Contributor.Name", "Contributor.Employer", "Contributor.Occupation",  # raw
  "name", "employer", "occupation",                                        # standardized
  "n_contributions", "contribution_types", "recipients",
  "cities", "states", "zip_codes",
  "code_final", "code_final_source", "code_final_description",
  "manual_review", "manual_review_reason",
  "rule_code", "rule_tier",
  "ml_code", "ml_conf",
  "llm_code", "llm_conf", "llm_reasoning",
  "claude_code", "claude_conf", "claude_reasoning",
  "os_code", "os_conf",
  ".queue_source"
)

```

```{r}
nrow(review_queue)
```


```{r}
unique(review_queue$manual_review_reason)
```

```{r}
review_queue %>% count()
review_queue %>% filter(manual_review_reason == "llm only source") %>% count()
review_queue %>% filter(grepl("llm only source",manual_review_reason)) %>% count()
review_queue %>% filter(grepl("alias_conflict",manual_review_reason)) %>% count()
review_queue %>% filter(manual_review_reason == "subsector_needed: 76/77 requires more specific code") %>% count()
```
now only 52 cases of subsetor needed... 
llm only source as the only flag is 713 rows..
most is because subsector needed a specific source... 

What other ones did John say to flag?? 
- which other codes? 

```{r}
review_queue %>% filter(grepl("alias_conflict",manual_review_reason)) %>% arrange(normalized_name)
```

```{r}
review_queue %>% filter(manual_review_reason == "alias_conflict: variants of this entity resolved to different codes") %>% arrange(normalized_name)
```


difference b/n 
alias_conflict
and 
alias_conflict: variants of this entity resolved to different codes

Also check whether one resolution tracks most of them? 
All review codes being tracked as pre-classified and not an alias conflict? 



```{r}
nrow(review_queue)
```


```{r}
review_queue %>% filter(manual_review_reason == "no source produced a usable value")
```

```{r}
# RUN OR DONT? 
nrow(review_queue[which(review_queue$manual_review_reason == "llm only source"),])
review_queue[which(review_queue$manual_review_reason == "llm only source"),]$manual_review <- FALSE
review_queue[which(review_queue$manual_review_reason == "llm only source"),]$manual_review_code <- FALSE
```

Surprises me that there are no instances of the LLM being the only source

```{r}
review_queue %>% filter(manual_review_reason == "subsector_needed: 76/77 requires more specific code")
review_queue %>% filter(is.na(manual_review_reason))
```


```{r}
nrow(review_queue)
```

```{r}
unique(review_queue$manual_review_reason)
```

```{r}
review_queue %>% filter(grepl("west, benajamin",tolower(Contributor.Name)))
resolved %>% filter(grepl("west, benajamin",tolower(name)))
```


```{r}
review_queue %>% filter(manual_review_reason == "no rule; llm only") %>% select(name,ml_code, ml_conf,llm_code)
```


```{r}
write_csv(
  review_queue %>% select(any_of(review_queue_cols)),
  "10_outputs/manual_review_queue_09-13-26.csv"
)
cat(sprintf(
  "Review queue: %d unique entities (from %d raw rows; %d flagged + %d individual_100 + %d confidence sample)\n",
  nrow(review_queue),
  nrow(review_queue_raw),
  sum(coalesce(resolved$manual_review, FALSE)),
  sum(review_queue$.queue_source == "individual_100", na.rm = TRUE),
  nrow(confidence_sample)
))
```

Don't need to include individual_100 in these... 
How long was the review queue before? this seems crazy.. 
It was <2000 




# All-contributors readable output (pre-review)

One row per unique contributor. Two dedup strategies are written for comparison:
- **entity**: organizations deduped by name only; individuals by name|employer|occupation
- **3part**: all entity types deduped by name|employer|occupation

Includes `manual_review` and `manual_review_reason` so a reviewer can filter to
flagged rows without needing the separate review queue file.

```{r}
colnames(resolved)
```


```{r}
resolved_annotated_all <- resolved %>%
  left_join(
    rule_raw_annotated %>% select(uuid,
      Contributor.Name, Contributor.Employer, Contributor.Occupation,
      recipient_label, contribution_type,
      Contributor.City, Contributor.State, Contributor.Zip.Code
    ),
    by = c("unit_id" = "uuid")
  ) %>%
  mutate(
    # Option A: orgs deduped by name|employer|occupation; individuals add zip
    dedup_key_entity = case_when(
      tolower(entity_type) == "organization" ~
        paste0(
          trimws(coalesce(`Contributor.Name`, name, "")), "|",
          trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
          trimws(coalesce(`Contributor.Occupation`, occupation, ""))
        ),
      TRUE ~ paste0(
        trimws(coalesce(`Contributor.Name`, name, "")), "|",
        trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
        trimws(coalesce(`Contributor.Occupation`, occupation, "")), "|",
        trimws(coalesce(`Contributor.Zip.Code`, ""))
      )
    ),
    # Option B: uniform name|employer|occupation for all entity types
    dedup_key_3part = paste0(
      trimws(coalesce(`Contributor.Name`, name, "")), "|",
      trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
      trimws(coalesce(`Contributor.Occupation`, occupation, ""))
    )
  ) %>%
  left_join(
    read_csv("10_inputs/manual_review_queue_reviewed.csv", col_types = cols(.default = "c")) %>%
      filter(trimws(tolower(manually_confirmed)) == "true") %>%
      mutate(subsector             = na_if(trimws(subsector), ""),
             subsector_description = na_if(trimws(subsector_description), "")) %>%
      select(dedup_key, subsector, subsector_description) %>%
      distinct(dedup_key, .keep_all = TRUE),
    by = c("dedup_key_3part" = "dedup_key")
  )

all_readable_cols <- c(
  "normalized_name", "name", "employer", "occupation", "entity_type",
  "n_contributions", "contribution_types", "recipients",
  "cities", "states", "zip_codes",
  "code_final", "code_final_description", "subsector", "subsector_description",
  "code_final_source", "category_final",
  "manual_review", "manual_review_reason",
  "rule_code", "rule_tier", "source1", "source2",
  "llm_code", "llm_conf",
  "claude_code", "claude_conf",
  "ml_code", "ml_conf",
  "os_code", "os_conf"
)

# Helper: aggregate geo/recipient info by a given dedup key, then dedup to one row.
make_readable_all <- function(df, key_col) {
  geo <- df %>%
    group_by(.data[[key_col]]) %>%
    summarise(
      n_contributions    = n(),
      contribution_types = paste(sort(unique(na.omit(trimws(contribution_type)))), collapse = " | "),
      recipients         = paste(sort(unique(na.omit(trimws(recipient_label)))),   collapse = " | "),
      cities    = paste(sort(unique(na.omit(trimws(Contributor.City)))),     collapse = " | "),
      states    = paste(sort(unique(na.omit(trimws(Contributor.State)))),    collapse = " | "),
      zip_codes = paste(sort(unique(na.omit(trimws(Contributor.Zip.Code)))), collapse = " | "),
      .groups = "drop"
    )
  df %>%
    left_join(geo, by = key_col) %>%
    arrange(desc(!is.na(llm_code)), desc(!is.na(rule_code)), desc(!is.na(ml_code))) %>%
    distinct(.data[[key_col]], .keep_all = TRUE) %>%
    select(any_of(all_readable_cols)) %>%
    arrange(normalized_name, name, employer)
}

today_all <- format(Sys.Date(), "%m-%d-%y")

# Entity-type-aware dedup
contributors_entity <- make_readable_all(resolved_annotated_all, "dedup_key_entity")
cat(sprintf("Entity-aware dedup: %d rows (%d flagged, %.1f%%)\n",
            nrow(contributors_entity),
            sum(coalesce(contributors_entity$manual_review, FALSE)),
            100 * mean(coalesce(contributors_entity$manual_review, FALSE))))

# Uniform 3-part dedup
contributors_3part <- make_readable_all(resolved_annotated_all, "dedup_key_3part")
cat(sprintf("3-part dedup:       %d rows (%d flagged, %.1f%%)\n",
            nrow(contributors_3part),
            sum(coalesce(contributors_3part$manual_review, FALSE)),
            100 * mean(coalesce(contributors_3part$manual_review, FALSE))))

write_csv(contributors_entity, sprintf("10_outputs/contributors_classified_entity_%s.csv", today_all))
write_csv(contributors_entity, "10_outputs/contributors_classified_entity.csv")
write_csv(contributors_3part,  sprintf("10_outputs/contributors_classified_%s.csv", today_all))
write_csv(contributors_3part,  "10_outputs/contributors_classified.csv")
```


```{r}
contributors_entity %>% filter(manual_review == TRUE)
contributors_entity %>% filter(manual_review ==  FALSE)
```

```{r}
contributors_entity %>% filter(manual_review_reason == "subsector_needed: 76/77 requires more specific code")
```

About 500 come from subsector needing a code. 

# Apply manual review decisions

Run this section after completing `10_inputs/manual_review_queue_reviewed.csv`.
It can be run standalone (without re-running the full notebook) — the first chunk
reloads any objects not already in memory.

```{r}
library(dplyr)
library(readr)
library(stringr)
library(tidyr)

# Reload if running standalone (not available after a fresh R session).
if (!exists("resolved")) {
  resolved <- read_csv("10_outputs/final_classifications_full_before_review.csv",
                       col_types = cols(.default = "c")) %>%
    mutate(manual_review = as.logical(manual_review))
}
if (!exists("rule_raw")) {
  rule_raw <- read_csv("../08_alternative_pipeline/08_outputs/classified_contributors_combined.csv",
                       col_types = cols(.default = "c"))
}
if (!exists("desc") || !is.data.frame(desc)) {
  label_url <- "https://docs.google.com/spreadsheets/d/11QHvNJsdtMlc1YKo_iNvMB_Jfn5Ui-iYdlWhFYjSm9g/export?format=csv"
  labels <- read.csv(label_url) %>% select(sector, sector_description) %>% distinct()
  desc <- labels %>% transmute(code_final = sector, code_final_description = sector_description)
}
```

```{r}
reviewed_raw <- read_csv("10_inputs/manual_review_queue_reviewed.csv",
                         col_types = cols(.default = "c"))

reviewed_patched <- reviewed_raw %>%
  filter(trimws(tolower(manually_confirmed)) == "true") %>%
  mutate(
    code_final_reviewed     = na_if(trimws(code_final), ""),
    subsector_reviewed      = na_if(trimws(subsector), ""),
    subsector_desc_reviewed = na_if(trimws(subsector_description), "")
  ) %>%
  select(dedup_key, code_final_reviewed, subsector_reviewed, subsector_desc_reviewed)

cat(sprintf("Manually confirmed rows to patch: %d\n", nrow(reviewed_patched)))
```

```{r}
tail(reviewed_raw,10)
```


```{r}
reviewed_raw[which(reviewed_raw$Contributor.Name == ""  | is.na(reviewed_raw$Contributor.Name)),]
```


```{r}
# make sure no classification id has more the one code
reviewed_raw %>% group_by(Contributor.Name,Contributor.Employer,Contributor.Occupation) %>% filter(n_distinct(code_final) > 1)
```

```{r}
# Re-derive the same dedup_key used when building the review queue, so we can fan
# the one-row-per-entity reviewed decisions back to every uuid for that entity.
resolved_keyed <- resolved %>%
  left_join(
    rule_raw %>% select(uuid, Contributor.Name, Contributor.Employer, Contributor.Occupation, Contributor.Zip.Code),
    by = c("unit_id" = "uuid")
  ) %>%
  mutate(
    dedup_key = if_else(
      tolower(entity_type) == "individual",
      paste0(
        trimws(coalesce(`Contributor.Name`, name, "")), "|",
        trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
        trimws(coalesce(`Contributor.Occupation`, occupation, "")), "|",
        trimws(coalesce(`Contributor.Zip.Code`, ""))
      ),
      paste0(
        trimws(coalesce(`Contributor.Name`, name, "")), "|",
        trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
        trimws(coalesce(`Contributor.Occupation`, occupation, ""))
      )
    )
  )

resolved_final <- resolved_keyed %>%
  left_join(reviewed_patched, by = "dedup_key") %>%
  mutate(
    .has_review           = !is.na(code_final_reviewed),
    code_final            = if_else(.has_review, code_final_reviewed, code_final),
    code_final_source     = if_else(.has_review, "manual_review", code_final_source),
    subsector             = if_else(.has_review, subsector_reviewed, NA_character_),
    subsector_description = if_else(.has_review, subsector_desc_reviewed, NA_character_),
    manual_review         = if_else(.has_review, FALSE, manual_review)
  ) %>%
  select(-.has_review, -code_final_reviewed, -subsector_reviewed, -subsector_desc_reviewed,
         -Contributor.Name, -Contributor.Employer, -Contributor.Occupation, -dedup_key) %>%
  mutate(code_final = case_when(
    code_final %in% c("100a", "100b") ~ "100",
    code_final == "31-34"             ~ "31-33",
    TRUE                              ~ code_final
  )) %>%
  select(-any_of("code_final_description")) %>%
  left_join(desc, by = "code_final")

cat(sprintf("Rows patched from review: %d | still flagged: %d / %d (%.1f%%)\n",
            sum(resolved_final$code_final_source == "manual_review", na.rm = TRUE),
            sum(coalesce(resolved_final$manual_review, FALSE)), nrow(resolved_final),
            100 * mean(coalesce(resolved_final$manual_review, FALSE))))
```


```{r}
resolved_final %>%
  filter(coalesce(manual_review, FALSE)) %>%
  select(name, employer, occupation, entity_type,
         code_final, code_final_description, code_final_source,
         manual_review_reason,
         rule_code, llm_code, ml_code) %>%
  distinct() %>%
  arrange(name)
```


```{r}
# Breakdowns after incorporating manual review
resolved_final %>% count(code_final, code_final_description, sort = TRUE)
resolved_final %>% count(code_final_source, sort = TRUE)
```
### Look at what produced incorrect classifications

```{r}
nrow(resolved_final)
```

```{r}
resolved_final %>% filter(code_final != rule_code & source1 == "identity_override") %>% select(name, employer, occupation, code_final, code_final_source, rule_code, llm_code) %>% distinct() %>% arrange(name)
```

identity_override: 
- remove investment
- remove insurance
- credit union
- Coalition for Renewable Natural Gas
- software
- unsurance
- publishing? 
- palantir (going in 79 instead of 92)
- recyclying
- maybe casino? 
- general fund? (the AFSCME General Fund, CA District Council 36 of)

This shouldn't really be in here: 
FedExPAC, FedEx	FEDEX CORPORATION	NON-CA CONTRIBUTIONS	20

	SCRIPPS HEALTH
		UNIVERSITY OF CALIFORNIA DAVIS

```{r}
head(resolved_final)
```

```{r}
unique(resolved_final$source1)
unique(resolved_final$source2)
```


Why are some of the locked pre-clasified ones coded as 99??

Errors: 
- Airbnb as tech
- don't treat unitemized member dues as a pac
- ambulances going into gov/ public safety? 
- associations as assoc, not by their industry
- is real-estate investment 52 or 50? 
- why are self-emplloyed attorneys being classified as 100 and not as an attorney? 
- real estate development as 52 not as 40?? 
- lots of associations being classified as 88
- doordash is tech
- public university classified as 92 instead of 60? 
- make sure tech companies classified as tech even if they operate in another industry...
- anything oil/petroleum related, classify at 21 even if official NAICS disagrees... 
- family trust classify as 100
- anything defense related goes into 92....
- vineyards in 11 generally? 
- consulting to healthcare e.g. puruse health tagged as 60 instead of 50? 
- prior classifications had in 40 -> wuld need to revisit those
- uber is 79... 
- lyft was 20
- what to do with uber and lyft? 
- tribal casinos as tribes or as casios? 

What to do w/ something like Kaiser thats an insurer but also a medical provider? 
- local travel and toursim authorities go into 70, ot into 50 (idk why LLM put in 50)

**big change**
- make sure unitmiezed union member contributions are counted as a union
- association classified as that and not as an industry
- instructions on real estate developer / developments? 
- oil/refinining in 11 not in 31-33, including distributors...

- why is a lawyer for state of california classified as 50 and not as 92? 
- when do self-employed people go to 100 vs when do they not? 

NEED to do entity rez step b/c could have introduced inconssitentcies w/ manual review step..

rule_code
ml_code
llm_code
os_code
code_final


```{r}
contributions_full %>% filter(grepl("AMGEN",toupper(Contributor.Name)))
contributions_full %>% filter(grepl("AMGEN",toupper(Recipient.Committee)))
```

```{r}
contributions_full %>% filter(Contributor.ID == "1439642")
```

```{r}
contributions_full %>% filter(grepl("HILTON",toupper(Recipient.Committee)))
```




```{r}
final_out_cols <- c(
  "unit_id", "entity_id", "entity_type", "name", "employer", "occupation",
  "code_final", "code_final_source", "code_final_description",
  "subsector", "subsector_description",
  "category_final", "category_final_source",
  "manual_review", "manual_review_reason",
  "suspected_org", "cross_source_conflict",
  "rule_code", "rule_label", "rule_tier", "rule_category",
  "os_code", "os_category", "os_conf",
  "llm_code", "llm_conf", "llm_category", "llm_category_conf", "llm_reasoning",
  "claude_code", "claude_conf", "claude_reasoning",
  "ml_code", "ml_conf", "ml_category", "ml_category_conf"
)
final_out_cols <- final_out_cols[final_out_cols %in% names(resolved_final)]

write_csv(resolved_final[, final_out_cols], "10_outputs/final_classifications_08-23-25.csv")
write_csv(resolved_final, "10_outputs/final_classifications_full_08-23-25.csv")
cat("Wrote final_classifications.csv and final_classifications_full.csv\n")
```

```{r}
# Row-level classified output — same schema as already_classified_contributions.csv.
# Joins classification codes from resolved_final back to the raw contribution rows
# in rule_raw (one row per uuid). Written to a dated file; does not overwrite the
# existing already_classified_contributions.csv.

row_level_out_cols <- c(
  "uuid",
  "Contributor.Name", "Contributor.Employer", "Contributor.Occupation",
  "Contributor.City", "Contributor.State",
  "Start.Date", "Amount", "Recipient.Name", "Recipient.Committee","Recipient.Committee.ID","Office","Contributor.ID", "Contributor.Zip.Code",
  "entity_id", "entity_type",
  "code_final", "code_final_source", "code_final_description",
  "category_final", "category_final_source",
  "manual_review", "manual_review_reason",
  "rule_code", "rule_tier",
  "llm_code", "llm_conf", "llm_reasoning",
  "ml_code", "ml_conf"
)

already_classified_new <- rule_raw %>%
  left_join(
    resolved_final %>% select(any_of(c(
      unit_id = "unit_id",
      "entity_id", "entity_type",
      "code_final", "code_final_source", "code_final_description",
      "category_final", "category_final_source",
      "manual_review", "manual_review_reason",
      "rule_code", "rule_tier",
      "llm_code", "llm_conf", "llm_reasoning",
      "ml_code", "ml_conf"
    ))),
    by = c("uuid" = "unit_id")
  ) %>%
  select(any_of(row_level_out_cols))

already_classified_path <- sprintf(
  "../08_alternative_pipeline/08_inputs/already_classified_contributions_%s.csv",
  format(Sys.Date(), "%m-%d-%y")
)
write_csv(already_classified_new, already_classified_path)
cat(sprintf("Wrote row-level classified output: %d rows → %s\n",
            nrow(already_classified_new), already_classified_path))
```

```{r}
# One-row-per-entity readable summary — dedup on the same key used in the review queue,
# with pipe-separated geo/recipient/amount aggregates matching the review queue style.

# Join raw contributor + recipient + geo fields from rule_raw, then build dedup_key.
rule_annotated <- rule_raw %>%
  mutate(
    recipient_label = case_when(
      Ballot.Measure.Contribution == "Y" ~
        coalesce(na_if(trimws(Ballot.Measure), ""), na_if(trimws(Recipient.Committee), "")),
      !is.na(Recipient.Name) & trimws(Recipient.Name) != "" ~ trimws(Recipient.Name),
      TRUE ~ na_if(trimws(Recipient.Committee), "")
    ),
    contribution_type = case_when(
      !is.na(parent_pac_id) & parent_pac_id != "" ~ "pac_contributor",
      Ballot.Measure.Contribution == "Y"           ~ "Ballot.Measure",
      Candidate.Contribution == "Y"                ~ "office",
      TRUE ~ NA_character_
    )
  ) %>%
  select(uuid,
         Contributor.Name, Contributor.Employer, Contributor.Occupation,
         recipient_label, Recipient.Name, Recipient.Committee,
         contribution_type,
         Contributor.City, Contributor.State, Contributor.Zip.Code,
         any_of("Amount"))

# manually_confirmed flag from the reviewed file (TRUE for all rows that went
# through manual review, NA for rows that were never in the queue).
review_confirmed_flag <- reviewed_raw %>%
  select(dedup_key, manually_confirmed)

resolved_final_keyed <- resolved_final %>%
  left_join(rule_annotated, by = c("unit_id" = "uuid")) %>%
  mutate(
    # Option B: individuals split by zip; orgs by name|employer|occupation (used by review_confirmed_flag join)
    dedup_key = if_else(
      tolower(entity_type) == "individual",
      paste0(
        trimws(coalesce(`Contributor.Name`, name, "")), "|",
        trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
        trimws(coalesce(`Contributor.Occupation`, occupation, "")), "|",
        trimws(coalesce(`Contributor.Zip.Code`, ""))
      ),
      paste0(
        trimws(coalesce(`Contributor.Name`, name, "")), "|",
        trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
        trimws(coalesce(`Contributor.Occupation`, occupation, ""))
      )
    ),
    # Option A: orgs deduped by name only; individuals by name|employer|occupation
    dedup_key_entity = case_when(
      tolower(entity_type) == "organization" ~
        trimws(coalesce(`Contributor.Name`, name, "")),
      TRUE ~ paste0(
        trimws(coalesce(`Contributor.Name`, name, "")), "|",
        trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
        trimws(coalesce(`Contributor.Occupation`, occupation, ""))
      )
    )
  )

geo_agg <- resolved_final_keyed %>%
  group_by(dedup_key) %>%
  summarise(
    n_contributions      = n(),
    contribution_types   = paste(sort(unique(na.omit(trimws(contribution_type)))), collapse = " | "),
    recipients           = paste(sort(unique(na.omit(trimws(recipient_label)))),   collapse = " | "),
    recipient_names      = paste(sort(unique(na.omit(trimws(Recipient.Name)))),    collapse = " | "),
    recipient_committees = paste(sort(unique(na.omit(trimws(Recipient.Committee)))), collapse = " | "),
    amounts              = if ("Amount" %in% names(resolved_final_keyed))
                             paste(sort(unique(na.omit(trimws(Amount)))), collapse = " | ")
                           else NA_character_,
    cities    = paste(sort(unique(na.omit(trimws(Contributor.City)))),     collapse = " | "),
    states    = paste(sort(unique(na.omit(trimws(Contributor.State)))),    collapse = " | "),
    zip_codes = paste(sort(unique(na.omit(trimws(Contributor.Zip.Code)))), collapse = " | "),
    .groups = "drop"
  )

readable_cols <- c(
  "name", "employer", "occupation", "entity_type",
  "n_contributions", "contribution_types",
  "recipients", "recipient_names", "recipient_committees", "amounts",
  "cities", "states", "zip_codes",
  "manually_confirmed",
  "manual_review", "manual_review_reason",
  "code_final", "code_final_description", "subsector", "subsector_description",
  "code_final_source", "category_final",
  "rule_code", "rule_tier",
  "llm_code", "llm_conf",
  "claude_code", "claude_conf",
  "ml_code", "ml_conf",
  "os_code", "os_conf"
)

today <- format(Sys.Date(), "%m-%d-%y")

final_readable <- resolved_final_keyed %>%
  left_join(geo_agg, by = "dedup_key") %>%
  left_join(review_confirmed_flag, by = "dedup_key") %>%
  arrange(desc(!is.na(llm_code)), desc(!is.na(rule_code)), desc(!is.na(ml_code))) %>%
  distinct(dedup_key, .keep_all = TRUE) %>%
  select(any_of(readable_cols)) %>%
  arrange(name, employer)

cat(sprintf("Readable (3-part dedup): %d unique entities\n", nrow(final_readable)))
write_csv(final_readable, sprintf("10_outputs/final_classifications_readable_%s.csv", today))

# Entity-type-aware dedup version
geo_agg_entity <- resolved_final_keyed %>%
  group_by(dedup_key_entity) %>%
  summarise(
    n_contributions      = n(),
    contribution_types   = paste(sort(unique(na.omit(trimws(contribution_type)))), collapse = " | "),
    recipients           = paste(sort(unique(na.omit(trimws(recipient_label)))),   collapse = " | "),
    recipient_names      = paste(sort(unique(na.omit(trimws(Recipient.Name)))),    collapse = " | "),
    recipient_committees = paste(sort(unique(na.omit(trimws(Recipient.Committee)))), collapse = " | "),
    amounts              = if ("Amount" %in% names(resolved_final_keyed))
                             paste(sort(unique(na.omit(trimws(Amount)))), collapse = " | ")
                           else NA_character_,
    cities    = paste(sort(unique(na.omit(trimws(Contributor.City)))),     collapse = " | "),
    states    = paste(sort(unique(na.omit(trimws(Contributor.State)))),    collapse = " | "),
    zip_codes = paste(sort(unique(na.omit(trimws(Contributor.Zip.Code)))), collapse = " | "),
    .groups = "drop"
  )

final_readable_entity <- resolved_final_keyed %>%
  left_join(geo_agg_entity, by = "dedup_key_entity") %>%
  left_join(review_confirmed_flag, by = "dedup_key") %>%
  arrange(desc(!is.na(llm_code)), desc(!is.na(rule_code)), desc(!is.na(ml_code))) %>%
  distinct(dedup_key_entity, .keep_all = TRUE) %>%
  select(any_of(readable_cols)) %>%
  arrange(name, employer)

cat(sprintf("Readable (entity-aware dedup): %d unique entities\n", nrow(final_readable_entity)))
write_csv(final_readable_entity,
          sprintf("10_outputs/final_classifications_readable_entity_%s.csv", today))
```

```{r}
final_readable %>% select
```

```{r}
resolved$entity
```


# PAC-org entity group conflict check

Can look at this now, but subject to change after entities are reviewed. 

Flag org/individual rows in `resolved_final` (the post-review dataframe built in
"Apply manual review decisions") whose entity group also contains a PAC in
`pac_industry_breakdown.csv` that shows a classification inconsistency. Two
conflict types are detected:

- **code_mismatch**: PAC has non-99 industries that differ from the org's code.
  Partial-99 in PAC industries (from $5k+ contributors still going through
  review) is excluded from the mismatch check — that 99 will resolve itself once
  those contributors are reviewed.
- **pac_entirely_uncategorized**: ALL of the PAC's industries are 99, while the
  org has a real code. This signals a missing or PENDING `industry_override_under_5k`
  in `pac_review` for a PAC whose parent org is already classified.

Run order:
1. Full notebook including "Apply manual review decisions" → `final_classifications` written
2. `match_pac_classifications.Rmd` → `pac_industry_breakdown.csv` written
3. This chunk

Joins directly on `entity_id_normalized`: PAC rows carry it from
`pac_classifiability.csv` (via `match_pac_classifications.Rmd`); org/individual
rows use `entity_id` from `final_classifications`, which is the same UUID
namespace (both come from `canonical_uuid_map`).

Appends new rows and annotates existing rows in `manual_review_queue.csv`.

```{r}
PAC_BREAKDOWN_PATH <- "10_outputs/pac_industry_breakdown.csv"
REVIEW_QUEUE_PATH  <- "10_outputs/manual_review_queue.csv"

if (!file.exists(PAC_BREAKDOWN_PATH)) {
  message("Skipping PAC-org conflict check: pac_industry_breakdown.csv not found.",
          " Run match_pac_classifications.Rmd first.")
} else {

  # resolved_final is the dataframe produced by "Apply manual review decisions".
  # If running this chunk standalone (fresh session), load from the written CSV.
  .base <- if (exists("resolved_final") && is.data.frame(resolved_final)) {
    message("Using resolved_final (post-review classifications).")
    resolved_final
  } else {
    .path <- "10_outputs/final_classifications_full_08-23-25.csv"
    message("resolved_final not in memory; loading from ", .path)
    read_csv(.path, col_types = cols(.default = "c")) %>%
      mutate(manual_review = as.logical(manual_review))
  }

  if (!exists("rule_raw")) {
    rule_raw <- read_csv(
      "../08_alternative_pipeline/08_outputs/classified_contributors_combined.csv",
      col_types = cols(.default = "c")
    )
  }

  pac_bk <- read_csv(PAC_BREAKDOWN_PATH, col_types = cols(.default = "c"))

  # entity_id_normalized per PAC — carried directly from pac_classifiability via
  # match_pac_classifications. Each PAC row has the same entity_id_normalized, so
  # one distinct() call gives us the (pac_id, pac_name, entity_id_normalized) lookup.
  pac_uuids <- pac_bk %>%
    distinct(pac_id, pac_name, entity_id_normalized) %>%
    filter(!is.na(entity_id_normalized) & entity_id_normalized != "")

  cat(sprintf("PAC-org check: %d / %d PACs matched to entity_id_normalized\n",
              n_distinct(pac_uuids$pac_id), n_distinct(pac_bk$pac_id, na.rm = TRUE)))

  # All industries (including 99) aggregated per entity_id_normalized
  pac_by_uuid <- pac_bk %>%
    filter(!is.na(industry) & industry != "") %>%
    inner_join(pac_uuids %>% select(pac_id, entity_id_normalized), by = "pac_id") %>%
    group_by(entity_id_normalized) %>%
    summarise(
      pac_names      = paste(sort(unique(coalesce(pac_name, ""))), collapse = " | "),
      pac_industries = paste(sort(unique(industry)), collapse = " | "),
      .groups = "drop"
    )

  # entity_id_normalized per org/individual — use entity_id from final_classifications
  # directly; entity_id in that file IS the UUID from canonical_uuid_map (same namespace
  # as pac_classifiability$entity_id_normalized). Rows with code_final = 99 are skipped:
  # they are already in review and a re-run will catch conflicts once a real code is assigned.
  org_with_uuid <- .base %>%
    filter(
      !is.na(entity_id) & entity_id != "",
      !is.na(code_final), code_final != FALLBACK_CODE,
      tolower(coalesce(entity_type, "")) %in% c("organization", "individual")
    ) %>%
    select(unit_id, name, code_final, entity_id_normalized = entity_id)

  # Detect conflicts
  conflicts <- org_with_uuid %>%
    inner_join(pac_by_uuid, by = "entity_id_normalized") %>%
    rowwise() %>%
    mutate(
      .org_code       = sub("^100[ab]$", "100", code_final),
      .all_pac_codes  = strsplit(pac_industries, " \\| ")[[1]],
      .pac_only_99    = all(.all_pac_codes == FALLBACK_CODE),
      # Exclude 99 from mismatch check: partial-99 is from pending-review contributors
      .pac_real_codes = setdiff(.all_pac_codes, FALLBACK_CODE),
      .conflict_type  = case_when(
        .pac_only_99                              ~ "pac_entirely_uncategorized",
        any(.pac_real_codes != .org_code)         ~ "code_mismatch",
        TRUE                                      ~ "none"
      )
    ) %>%
    ungroup() %>%
    filter(.conflict_type != "none") %>%
    select(unit_id, entity_id_normalized, code_final, pac_names, pac_industries, .conflict_type)

  cat(sprintf("PAC-org entity group conflicts: %d rows across %d entity groups (%d code_mismatch, %d pac_entirely_uncategorized)\n",
              nrow(conflicts),
              n_distinct(conflicts$entity_id_normalized),
              sum(conflicts$.conflict_type == "code_mismatch"),
              sum(conflicts$.conflict_type == "pac_entirely_uncategorized")))

  if (nrow(conflicts) > 0 && file.exists(REVIEW_QUEUE_PATH)) {

    conflict_meta <- conflicts %>%
      transmute(
        unit_id,
        pac_names,
        pac_industries,
        .conflict_type,
        .reason = case_when(
          .conflict_type == "pac_entirely_uncategorized" ~ paste0(
            "pac_org_entity_group_conflict [pac_entirely_uncategorized]: org=", code_final,
            ", PAC(s)=", pac_names, " (all industries=99)"
          ),
          TRUE ~ paste0(
            "pac_org_entity_group_conflict [code_mismatch]: org=", code_final,
            ", PAC(s)=", pac_names, ", PAC industries=", pac_industries
          )
        )
      )

    # Derive dedup_key matching the construction in the main review queue build
    conflict_keyed <- conflict_meta %>%
      left_join(
        rule_raw %>% select(uuid, Contributor.Name, Contributor.Employer, Contributor.Occupation, Contributor.Zip.Code),
        by = c("unit_id" = "uuid")
      ) %>%
      left_join(
        .base %>% select(unit_id, name, employer, occupation, entity_type),
        by = "unit_id"
      ) %>%
      mutate(
        dedup_key = if_else(
          tolower(entity_type) == "individual",
          paste0(
            trimws(coalesce(`Contributor.Name`, name, "")), "|",
            trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
            trimws(coalesce(`Contributor.Occupation`, occupation, "")), "|",
            trimws(coalesce(`Contributor.Zip.Code`, ""))
          ),
          paste0(
            trimws(coalesce(`Contributor.Name`, name, "")), "|",
            trimws(coalesce(`Contributor.Employer`, employer, "")), "|",
            trimws(coalesce(`Contributor.Occupation`, occupation, ""))
          )
        )
      )

    existing_q <- read_csv(REVIEW_QUEUE_PATH, col_types = cols(.default = "c"))

    # Annotate existing rows already in the queue
    reason_patch <- conflict_keyed %>%
      distinct(dedup_key, .reason) %>%
      group_by(dedup_key) %>%
      summarise(.reason = paste(unique(.reason), collapse = " | "), .groups = "drop")

    updated_q <- existing_q %>%
      left_join(reason_patch, by = "dedup_key") %>%
      mutate(
        manual_review_reason = case_when(
          !is.na(.reason) & is.na(manual_review_reason)  ~ .reason,
          !is.na(.reason) ~ paste0(manual_review_reason, " | ", .reason),
          TRUE ~ manual_review_reason
        )
      ) %>%
      select(-.reason)

    n_annotated <- sum(conflict_keyed$dedup_key %in% existing_q$dedup_key, na.rm = TRUE)

    # Rows not yet in the queue at all
    new_keys <- setdiff(conflict_keyed$dedup_key, existing_q$dedup_key)
    new_rows <- conflict_keyed %>%
      filter(dedup_key %in% new_keys) %>%
      distinct(dedup_key, .keep_all = TRUE) %>%
      left_join(
        .base %>% select(
          unit_id, entity_id, entity_type,
          code_final, code_final_source, code_final_description,
          rule_code, rule_tier,
          llm_code, llm_conf, llm_reasoning,
          ml_code, ml_conf, os_code, os_conf
        ),
        by = "unit_id"
      ) %>%
      transmute(
        dedup_key,
        Contributor.Name, Contributor.Employer, Contributor.Occupation,
        name, employer, occupation,
        code_final, code_final_source, code_final_description,
        manual_review        = "TRUE",
        manual_review_reason = .reason,
        rule_code, rule_tier,
        llm_code, llm_conf, llm_reasoning,
        ml_code, ml_conf, os_code, os_conf,
        .queue_source = "pac_org_conflict"
      )

    write_csv(bind_rows(updated_q, new_rows), REVIEW_QUEUE_PATH)
    cat(sprintf(
      "Review queue updated: %d existing row(s) annotated, %d new row(s) added → %s\n",
      n_annotated, nrow(new_rows), REVIEW_QUEUE_PATH
    ))
  }
}
```

