---
title: "Match PAC Classifications"
output: html_notebook
---

# Summary
Computes an industry distribution for each PAC by:
- assigning under-$5k contributors the sector from `industry_override_under_5k` (pac_review)
- assigning federal PACs their entire amount to that same sector (over-$5k ignored)
- looking up individually classified over-$5k contributors from final_classifications
- recursively distributing PAC-type over-$5k contributors using their own distributions
- leaving PACs flagged `additional_review_req` or `PENDING` as UNCLASSIFIED

# To Do
- figure out what to do w/ unitemized that aren't directly labeled as unitemized
- pac_final_classifications.csv only covers PAC-graph donor contributors
  (pac_classification_input.csv). PACs' own direct giving to races/props
  (pac_input.csv) still needs its non-PAC rows classified -- see
  assign_pac_final_classification.Rmd's docstring.

# Load data and packages

```{r}
library(dplyr)
library(readr)
library(stringr)
library(digest)
library(googlesheets4)
```

```{r}
hash_id <- function(keys, prefix) {
  paste0(prefix, "_", substr(
    digest::digest(paste(sort(unique(keys)), collapse = "||"), algo = "xxhash32"), 1, 16
  ))
}
```

```{r}
over_5k_tagged      <- read_csv("../08_alternative_pipeline/08_outputs/over_5k_tagged.csv",
                                 col_types = cols(.default = "c", contributor_amount = "d"))
pac_classifiability <- read_csv("../08_alternative_pipeline/08_outputs/pac_classifiability.csv",
                                 col_types = cols(.default = "c", total_received = "d"))
pac_class_input     <- read_csv("../08_alternative_pipeline/08_outputs/pac_classification_input.csv",
                                 col_types = cols(.default = "c"))
ordered_pacs        <- read_csv("../08_alternative_pipeline/08_outputs/pac_traversal_order.csv") %>%
  arrange(order) %>% pull(pac_id) %>% as.character()
stuck_pacs          <- read_csv("../08_alternative_pipeline/08_outputs/pac_stuck_pacs.csv") %>%
  pull(pac_id) %>% as.character()
pac_final           <- read_csv("10_outputs/final_classifications_full_before_review.csv",
                                 col_types = cols(.default = "c"))
pac_review          <- read_csv("10_inputs/pac_manual_review.csv",
                                 col_types = cols(.default = "c", under_5k_amount = "d"))
```

```{r}
nrow(pac_final)
```


```{r}
# pac_review rows with no pac_id are PACs that contributed to other PACs but have no
# CAL-ACCESS committee ID. They cannot be looked up by id — they are matched by
# Contributor.Name throughout (see pac_name_override below).
cat(sprintf("pac_review rows with no pac_id (matched by name): %d\n",
            sum(is.na(pac_review$pac_id) | pac_review$pac_id == "")))
```

# Sector mapping

`industry_override_under_5k` is a sector or subsector description (e.g. "Misc
Unions") - need to map this to a sector code for consistency. Match against
the granular `subsector_description` first (unique per row, e.g. "Misc
Unions" -> subsector 77e -> sector 77), falling back to the broader
`sector_description` (repeated across a sector's subsectors, but never
ambiguous -- every sector_description maps to exactly one sector code).

```{r}
gs4_deauth()
sector_map_full <- read_sheet(
  "https://docs.google.com/spreadsheets/d/11QHvNJsdtMlc1YKo_iNvMB_Jfn5Ui-iYdlWhFYjSm9g/edit?gid=605391596#gid=605391596"
) %>%
  mutate(across(everything(), as.character))

subsector_lookup <- sector_map_full %>%
  filter(!is.na(subsector_description)) %>%
  distinct(subsector_description, .keep_all = TRUE) %>%
  select(subsector_description, sector)

sector_lookup <- sector_map_full %>%
  filter(!is.na(sector_description)) %>%
  distinct(sector_description, .keep_all = TRUE) %>%
  select(sector_description, sector)
```


# pac_review lookup tables

```{r}
# PACs (by id) that cannot be classified yet — leave their full amount as UNCLASSIFIED.
# Rows with no pac_id are excluded here; they are handled via pac_name_override below.
skip_pac_ids <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  filter(
    unresolvable_reason == "additional_review_req" |
    is.na(industry_override_under_5k)              |
    industry_override_under_5k == ""               |
    industry_override_under_5k == "PENDING"
  ) %>%
  pull(pac_id) %>%
  as.character()
```

```{r}
pac_review$unitemized_amount <- as.numeric(pac_review$unitemized_amount)
pac_review$small_dollar_amount <- as.numeric(pac_review$small_dollar_amount)

pac_override <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "", !pac_id %in% skip_pac_ids) %>%
  left_join(subsector_lookup, by = c("industry_override_under_5k" = "subsector_description")) %>%
  left_join(sector_lookup, by = c("industry_override_under_5k" = "sector_description"),
            suffix = c("", "_broad")) %>%
  mutate(pac_id = as.character(pac_id)) %>%
  # use pac_classifiability for amounts — it reflects the most recent data pull;
  # pac_review amounts may be from an earlier pull
  left_join(
    pac_classifiability %>% transmute(
      pac_id   = as.character(pac_id),
      under_5k = coalesce(as.numeric(under_5k_amount), 0) +
                 coalesce(as.numeric(unitemized_amount), 0) +
                 coalesce(as.numeric(small_dollar_amount), 0)
    ),
    by = "pac_id"
  ) %>%
  mutate(
    override_sector = coalesce(sector, sector_broad),
    is_federal      = !is.na(fec_id) & trimws(coalesce(fec_id, "")) != ""
  ) %>%
  select(pac_id, override_sector, industry_override_under_5k, under_5k, is_federal)

# warn on descriptions that didn't match any sector code
unmapped <- pac_override %>% filter(is.na(override_sector))
if (nrow(unmapped) > 0) {
  cat("Warning: no sector code found for these industry_override_under_5k values:\n")
  print(unmapped %>% select(pac_id, industry_override_under_5k))
}

pac_override_map <- setNames(seq_len(nrow(pac_override)), pac_override$pac_id)

# Name-keyed override for pac_review rows with no pac_id (PACs that contributed to other PACs
# but have no CAL-ACCESS committee ID). Matched by contributor_name in over_5k_tagged.
skip_pac_names <- pac_review %>%
  filter(is.na(pac_id) | pac_id == "") %>%
  filter(
    unresolvable_reason == "additional_review_req" |
    is.na(industry_override_under_5k)              |
    industry_override_under_5k == ""               |
    industry_override_under_5k == "PENDING"
  ) %>%
  pull(Contributor.Name)

pac_name_override <- pac_review %>%
  filter(is.na(pac_id) | pac_id == "",
         !Contributor.Name %in% skip_pac_names) %>%
  left_join(subsector_lookup, by = c("industry_override_under_5k" = "subsector_description")) %>%
  left_join(sector_lookup, by = c("industry_override_under_5k" = "sector_description"),
            suffix = c("", "_broad")) %>%
  mutate(
    override_sector = coalesce(sector, sector_broad),
    under_5k        = coalesce(under_5k_amount, 0) + coalesce(unitemized_amount, 0) + coalesce(small_dollar_amount, 0),
    is_federal      = !is.na(fec_id) & trimws(coalesce(fec_id, "")) != ""
  ) %>%
  filter(!is.na(override_sector)) %>%
  distinct(Contributor.Name, .keep_all = TRUE) %>%
  select(name = Contributor.Name, override_sector, industry_override_under_5k, under_5k, is_federal)

pac_name_override_map <- setNames(seq_len(nrow(pac_name_override)), pac_name_override$name)

unmapped_names <- pac_review %>%
  filter(is.na(pac_id) | pac_id == "",
         !Contributor.Name %in% skip_pac_names,
         !Contributor.Name %in% pac_name_override$name)
if (nrow(unmapped_names) > 0) {
  cat("Warning: no sector code for these name-only pac_review entries:\n")
  print(unmapped_names %>% select(Contributor.Name, industry_override_under_5k))
}
```

```{r}
pac_override
```

```{r}
pac_name_override
```


# Over-$5k individual/org classification lookup


```{r}
# over_5k_tagged has correct amounts (one row per contributor per PAC, totals
# match pac_classifiability$total_received). 
# Use over_5k_tagged for amounts; build a per-entity
# code lookup from pac_class_input + pac_final (best non-"99" code wins when
# the same entity_group maps to multiple uuids across different classification runs).
entity_code_lookup <- pac_class_input %>%
  filter(contributor_type != "pac") %>%
  mutate(parent_pac_id = as.character(parent_pac_id)) %>%
  left_join(pac_final %>% select(unit_id, code_final), by = c("uuid" = "unit_id")) %>%
  mutate(industry = coalesce(code_final, "99")) %>%
  group_by(parent_pac_id, entity_group) %>%
  arrange(industry == "99", .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  select(parent_pac_id, entity_group, industry)

over_5k_direct <- over_5k_tagged %>%
  filter(contributor_type != "pac") %>%
  mutate(
    contributor_amount = as.numeric(contributor_amount),
    parent_pac_id      = as.character(parent_pac_id)
  ) %>%
  left_join(entity_code_lookup, by = c("parent_pac_id", "entity_group")) %>%
  mutate(industry = coalesce(industry, "99"))

# amount per (pac, industry) for the direct portion
direct_by_pac_list <- over_5k_direct %>%
  group_by(parent_pac_id, industry) %>%
  summarise(amount = sum(contributor_amount, na.rm = TRUE), .groups = "drop") %>%
  split(~parent_pac_id)
```

- should we output entity_id instead of re-computing it here? 

```{r}
over_5k_direct
```

```{r}
unique(over_5k_direct$industry)
```


# Compute PAC industry distributions

```{r}
over_5k_by_pac <- split(
  over_5k_tagged %>% mutate(
    parent_pac_id  = as.character(parent_pac_id),
    contributor_id = as.character(contributor_id)
  ),
  ~ as.character(parent_pac_id)
)

pac_deps <- over_5k_tagged %>%
  filter(contributor_type %in% c("pac"), !is.na(contributor_id)) %>%
  mutate(parent_pac_id  = as.character(parent_pac_id),
         contributor_id = as.character(contributor_id)) %>%
  distinct(parent_pac_id, contributor_id) %>%
  group_by(parent_pac_id) %>%
  summarise(deps = list(contributor_id), .groups = "drop") %>%
  { setNames(.$deps, .$parent_pac_id) }
```




```{r}
industry_distrib <- list()

compute_industry_distrib <- function(pid) {
  pid <- as.character(pid)

  # (1) pending / needs review — entire amount stays uncategorized
  if (pid %in% skip_pac_ids) {
    pac_row <- pac_classifiability[pac_classifiability$pac_id == pid, ]
    total   <- if (nrow(pac_row) > 0) coalesce(pac_row$total_received[1], 0) else 0
    industry_distrib[[pid]] <<- if (total > 0) c("99" = total) else numeric(0)
    return(invisible(NULL))
  }

  ovr <- if (pid %in% names(pac_override_map)) pac_override[pac_override_map[[pid]], ] else NULL

  # (2) federal PAC — entire amount classified by override sector; over-$5k ignored if applicable
  if (!is.null(ovr) && isTRUE(ovr$is_federal)) {
    pac_row <- pac_classifiability[pac_classifiability$pac_id == pid, ]
    total   <- if (nrow(pac_row) > 0) coalesce(pac_row$total_received[1], 0) else 0
    code    <- coalesce(ovr$override_sector, "99")
    industry_distrib[[pid]] <<- if (total > 0) setNames(total, code) else numeric(0)
    return(invisible(NULL))
  }

  # (3) non-federal PAC: over-$5k direct + recursive PAC-type + under-$5k override

  direct    <- direct_by_pac_list[[pid]]
  ind_amounts <- if (!is.null(direct) && nrow(direct) > 0)
    setNames(direct$amount, direct$industry) else numeric(0)

  pac_rows <- over_5k_by_pac[[pid]]
  if (!is.null(pac_rows)) {
    all_pac_rows <- pac_rows %>%
      filter(contributor_type == "pac")

    # PAC contributors with a committee id: distribute recursively
    id_rows <- all_pac_rows %>% filter(!is.na(contributor_id) & contributor_id != "")
    for (i in seq_len(nrow(id_rows))) {
      child_id     <- id_rows$contributor_id[i]
      child_amount <- id_rows$contributor_amount[i]
      child_dist   <- industry_distrib[[child_id]]
      if (is.null(child_dist) || length(child_dist) == 0) {
        ind_amounts["99"] <- coalesce(ind_amounts["99"], 0) + child_amount
        next
      }
      child_total <- sum(child_dist)
      if (child_total <= 0) next
      scaled <- child_dist / child_total * child_amount
      for (ind in names(scaled))
        ind_amounts[ind] <- coalesce(ind_amounts[ind], 0) + scaled[ind]
    }

    # PAC contributors with no id: look up override sector by name
    no_id_rows <- all_pac_rows %>% filter(is.na(contributor_id) | contributor_id == "")
    for (i in seq_len(nrow(no_id_rows))) {
      child_name   <- no_id_rows$contributor_name[i]
      child_amount <- no_id_rows$contributor_amount[i]
      if (child_name %in% skip_pac_names || !child_name %in% names(pac_name_override_map)) {
        ind_amounts["99"] <- coalesce(ind_amounts["99"], 0) + child_amount
        next
      }
      code <- pac_name_override$override_sector[pac_name_override_map[[child_name]]]
      code <- coalesce(code, "99")
      ind_amounts[code] <- coalesce(ind_amounts[code], 0) + child_amount
    }
  }

  # under-$5k override
  if (!is.null(ovr) && !is.na(ovr$under_5k) && ovr$under_5k > 0) {
    code <- coalesce(ovr$override_sector, "99")
    ind_amounts[code] <- coalesce(ind_amounts[code], 0) + ovr$under_5k
  } else {
    # PAC not yet in pac_review (no manual override): classify all under-$5k as 99
    pac_row <- pac_classifiability[pac_classifiability$pac_id == pid, ]
    if (nrow(pac_row) > 0) {
      under_5k_amt <- coalesce(as.numeric(pac_row$under_5k_amount[1]), 0) +
                      coalesce(as.numeric(pac_row$unitemized_amount[1]), 0) +
                      coalesce(as.numeric(pac_row$small_dollar_amount[1]), 0)
      if (under_5k_amt > 0)
        ind_amounts["99"] <- coalesce(ind_amounts["99"], 0) + under_5k_amt
    }
  }

  industry_distrib[[pid]] <<- ind_amounts
}
```

```{r}
for (pid in ordered_pacs) compute_industry_distrib(pid)

remaining <- stuck_pacs
while (length(remaining) > 0) {
  ready <- Filter(function(pid) {
    deps <- pac_deps[[pid]]
    is.null(deps) || !any(deps %in% remaining)
  }, remaining)
  if (length(ready) == 0) ready <- remaining
  for (pid in ready) compute_industry_distrib(pid)
  remaining <- setdiff(remaining, ready)
}
```

# Output

```{r}
pac_name_lookup <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  mutate(pac_id = as.character(pac_id)) %>%
  group_by(pac_id) %>%
  slice(1) %>%
  ungroup() %>%
  select(pac_id, pac_name = Contributor.Name)

# id-based PAC distributions
id_breakdown <- lapply(names(industry_distrib), function(pid) {
  ind <- industry_distrib[[pid]]
  if (length(ind) == 0) return(NULL)
  tibble(pac_id = pid, industry = names(ind), amount = unname(ind))
}) %>%
  bind_rows() %>%
  left_join(pac_name_lookup, by = "pac_id") %>%
  left_join(
    pac_classifiability %>% transmute(pac_id = as.character(pac_id), total_received),
    by = "pac_id"
  ) %>%
  mutate(pct_of_total = amount / total_received)

# no-id PAC distributions (PACs that contributed to other PACs but have no committee id)
name_breakdown <- pac_name_override %>%
  filter(!is.na(under_5k), under_5k > 0) %>%
  transmute(
    pac_id       = NA_character_,
    pac_name     = name,
    industry     = coalesce(override_sector, "99"),
    amount       = under_5k,
    total_received = NA_real_,
    pct_of_total = NA_real_
  )

pac_industry_breakdown <- bind_rows(id_breakdown, name_breakdown) %>%
  select(pac_id, pac_name, industry, amount, total_received, pct_of_total) %>%
  arrange(pac_id, pac_name, desc(amount))
```


### manual data checks to confirm processing
```{r}
pac_industry_breakdown %>% group_by(pac_id) %>% mutate(total = sum(pct_of_total)) %>% arrange(pac_id) 
```
1305436	CALIFORNIA ASSOCIATION OF DENTAL PLANS PAC -> was wrong

```{r}
over_5k_tagged %>%filter(parent_pac_id == "1305436")
```

```{r}
pac_industry_breakdown %>% filter(pac_id == "1408990")
```


```{r}
over_5k_direct %>%filter(parent_pac_id == "1305436") #%>% summarise(total = sum(contributor_amount))
```

```{r}
pac_industry_breakdown %>% group_by(pac_id) %>% mutate(total = sum(pct_of_total)) %>% arrange(pac_id) %>% filter(total < 0.999999)
```

```{r}
pac_review %>% filter(pac_id == "1305436")
```

```{r}
pac_classifiability %>% filter(pac_id == "1305436")
```


```{r}
over_5k_direct %>% filter(grepl("empower",contributor_name))
```

```{r}
over_5k_direct %>% filter(parent_pac_id == "1018355", industry == "76") %>% summarise(total = sum(contributor_amount))
over_5k_direct %>% filter(parent_pac_id == "1018355", industry == "76") %>% summarise(total = sum(contributor_amount))
```

```{r}
#1.182035e+01
#1.394942e+00
#1.321530e+01	
1.321530e+01
2187603.11	
(1.975310e+06	+ 1.623288e+05	)	
```

```{r}
over_5k_tagged %>% filter(contributor_id == "1467150")
```


Under $5k is ALL 77

```{r}
52737.13 +	6205.85	+ 520134.1	
```


unitemized, small dollar, under 5k, over 5k
52737.13	6205.85	520134.1	1371173.59

```{r}
ids <- pac_class_input %>% filter(parent_pac_id == "1408990") %>% select(uuid)
ids <- ids$uuid
length(unique(ids))

pac_final %>% filter(unit_id %in% ids) 
```


llm code is medium -> ml code was chosen??
os_code was there,


Remove regex for code 76 -> associations that are actually unions..


```{r}
pac_final %>% filter(is.na(os_conf) & !is.na(os_code))
```


```{r}
pac_inputs
```


# WRITE OUTPUT
```{r}
write_csv(pac_industry_breakdown, "10_outputs/pac_industry_breakdown.csv")
```


```{r}
pac_input <- read.csv("../08_alternative_pipeline/08_outputs/pac_input.csv")
```

pac input = all direct contributor pacs... 


```{r}
pac_ids   <- pac_classifiability %>%
  pull(pac_id) %>% as.character()
```

```{r}
pac_input <- pac_input %>%
  mutate(is_pac = FILER_ID %in% pac_ids)
```


```{r}
pac_input %>% filter(!(is_pac))
```


```{r}
full <- read.csv("../08_alternative_pipeline/08_inputs/power_search_contributions_normalized.csv")
nrow(full)
```

```{r}
full %>% filter(toupper(Recipient.Name) == "HILTON, STEVE"& toupper(Contributor.Name) == "UNITEMIZED CONTRIBUTIONS") %>% summarise(total = sum(Amount))
```

```{r}
full %>% filter(toupper(Recipient.Name) == "BECERRA, XAVIER" & toupper(Contributor.Name) == "UNITEMIZED CONTRIBUTIONS") %>% summarise(total = sum(Amount))
```

