---
title: "R Notebook"
output: html_notebook
---

# Purpose
Configure descriptions for each custom NAICS code for the LLM to reference when assigning codes.


```{r}
library(readxl)
library(dplyr)
library(stringr)
```

### Add more information to the titles

```{r}
codes <- read_excel("03_inputs/digit_2022_Codes.xlsx")

# remove blank rows and columns
codes <- codes[-1,]
codes <- codes[,-c(4,5)]

colnames(codes)[c(2,3)] <- c("Code","Title")

```

```{r}
# expand two digit codes with a dash

codes[which(codes$Code == "31-33"),]$Code <- "31"
codes <- rbind(codes,c(NA,"32","Manufacturing"))
codes <- rbind(codes,c(NA,"33","Manufacturing"))

codes[which(codes$Code == "44-45"),]$Code <- "44"
codes <- rbind(codes,c(NA,"45","Retail Trade"))

codes[which(codes$Code == "48-49"),]$Code <- "48"
codes <- rbind(codes,c(NA,"49","Transportation and Warehousing"))

tail(codes)
```


```{r}
sector_expanded <- codes %>%
  filter(nchar(Code) == 2) %>%
  mutate(expanded_title = 
    sapply(Code, function(code) {
      parent_title <- codes %>%
        filter(Code == code) %>%
        pull(Title)
      
      sub_titles <- codes %>%
        filter(str_starts(Code, code), nchar(Code) == 4) %>%
        distinct(Title) %>%
        pull(Title)
      
      if (length(sub_titles) == 0) return(parent_title)
      
      all_titles <- c(parent_title, sub_titles)
    
      paste0(parent_title, ", including: ", paste(sub_titles, collapse = "; "))
    })
  )

```


### Collapse descriptions for combined categories: 

40: Combines 53 and 23
50: Combines 54, 55 and 56a
60: Combines 62, 61 and lirbaries and archives
70: Combines 71 + parts of 51
76: Remove association description from relvant naics code? 
77: Remove assocition description from relevant naics code
80: 44-45, 42, 92, parts of 81
92: 92 + defense contractors, take out tribes

```{r}
# helper function to combine sector descriptions
combine_sectors <- function(new_code, new_title, source_codes = character(0), extra_subtitles = character(0)) {
  if (length(source_codes) > 0) {
    sub_titles <- codes %>%
      filter(sapply(Code, function(c) any(str_starts(c, source_codes))) & nchar(Code) == 4) %>%
      distinct(Title) %>%
      pull(Title)
  } else {
    sub_titles <- character(0)
  }

  all_subs <- c(sub_titles, extra_subtitles)

  description <- if (length(all_subs) == 0) {
    new_title
  } else {
    paste0(new_title, ", including: ", paste(all_subs, collapse = "; "))
  }

  data.frame(naics_sector = new_code, description = description, stringsAsFactors = FALSE)
}

```


```{r}
x <- sector_expanded[,c("Code","expanded_title"),]
colnames(x) <- c("naics_sector","description")

x <- rbind(x,c("99","Unknown"))
x <- rbind(x,c("90","Candidates/Political Parties"))
x <- rbind(x,c("88","Political Actions Committees (PACs), contributor committees and political action funds"))
x <- rbind(x,c("76","Business, professional and trade associations"))
x <- rbind(x,c("77","Unions"))
x <- rbind(x,c("92a","Native American tribes"))
x <- rbind(x,c("100","Retired/Homemaker/Student/Unemployed"))

# remove source codes that get merged into combined categories
x <- x %>% filter(!naics_sector %in% c("22","23", "53","42","44","45","48","49","51","53","54","55","56","61","62","71","72","81"))

# add combined categories
x <- rbind(x, combine_sectors("20", "Utilities, infrastructure, telecommunications, renewable energy, transportation and warehousing", 
                              source_codes = c("22","48","49"),
                              extra_subtitles = c("Wired and Wireless Telecommunications (except Satellite)",
                                                  "Satellite Telecommunications","All Other Telecommunications",
                                                  "Waste Collection","Waste Treatment and Disposal","Remediation and Other Waste Management Services")),
              combine_sectors("40", "Construction/Real Estate", source_codes = c("23", "53")),
              combine_sectors("50", "Professional/Legal", source_codes = c("54", "55"),
                              extra_subtitles = c("Management and Holding Companies",
                                                  "Office Administrative Services",
                                                  "Employment Services",
                                                  "Facilities Support Services",
                                                  "Business Support Services",
                                                  "Travel Arrangement and Reservation Services",
                                                  "Investigation and Security Services",
                                                  "Services to Buildings and Dwellings")),
              combine_sectors("60", "Healthcare/Education", source_codes = c("61", "62"),
                              extra_subtitles = c("Libraries and archives")),
              combine_sectors("70", "Entertainment, Media and Gambling", source_codes = c("71"),
                  extra_subtitles = c("Motion Picture and Video Industries", "Sound Recording Industries","Newspaper, Periodical, Book, and Directory Publishers",
                      "Radio and Television Broadcasting Stations")),
              combine_sectors("79", "Technology, software and AI companies", source_codes = character(0),
                  extra_subtitles = c("AI services","Software Publishers",
                                      "Media Streaming Distribution Services, Social Networks, and Other Media Networks and Content Providers",
                                      "Computing Infrastructure Providers, Data Processing, Web Hosting, and Related Services")),
              
              combine_sectors("80", "Other Industries",  source_codes = c("44","45","42","72"),
                  extra_subtitles = c("Automotive Repair and Maintenance","Electronic and Precision Equipment Repair and Maintenance",
                                      "Commercial and Industrial Machinery and Equipment (except Automotive and Electronic) Repair and Maintenance",
                                      "Personal and Household Goods Repair and Maintenance","Personal Care Services", "Death Care Services",
                                      "Drycleaning and Laundry Services","Other Personal Services","Religious Organizations","Grandmaking and Giving Services",
                                      "Social Advocacy Organizations","Civic and Social Organizations"))             
)


```


### Expand descriptions after manual review feedback 

```{r}
x$description[x$naics_sector == "21"] <- paste0(
  x$description[x$naics_sector == "21"],
  " Also use this code for any company primarily in the oil/gas/petroleum business -- ",
  "extraction, drilling, refining, or distribution -- even if its formal NAICS ",
  "classification would otherwise be manufacturing (31/32) or wholesale trade (80)."
)

x$description[x$naics_sector == "92"] <- paste0(
  x$description[x$naics_sector == "92"],
  " Also use this code for defense contractors and companies whose primary business is ",
  "military/defense products or services, even if their formal NAICS code would ",
  "otherwise be manufacturing (31-33). Also use this code for ambulance and ambulance ",
  "services, even though this might otherwise suggest Healthcare (60)."
)

x$description[x$naics_sector == "88"] <- paste0(
  "Entities EXPLICITLY registered or identified as a Political Action Committee (PAC), ",
  "contributor committee, or political action fund -- e.g. name contains 'PAC', ",
  "'Political Action Committee', 'COPE', or confirmed via research to be a distinct ",
  "registered political committee. Do NOT use this code for a union, trade association, ",
  "corporation, or other organization just because it is politically active, lobbies, or ",
  "has a separately-registered affiliated PAC -- classify the parent organization by its ",
  "own actual industry instead (e.g. a union is '77', not '88'). Unitemized dues or ",
  "intermediary contributions passed through a PAC should be classified by the underlying ",
  "organization's actual industry when it can be determined (e.g. a union's PAC dues -> ",
  "'77'), not as '88'."
)

x$description[x$naics_sector == "76"] <- paste0(
  x$description[x$naics_sector == "76"],
  " Use this code even if the association has an affiliated PAC or engages in political ",
  "or lobbying activity -- the association itself is '76', not '88'; only a separately-",
  "registered PAC entity gets '88'."
)

x$description[x$naics_sector == "79"] <- paste0(
  x$description[x$naics_sector == "79"],
  " Use this code regardless of which industry vertical the technology serves -- e.g. ",
  "health-tech, fintech, and ad-tech companies are '79', not the industry they serve."
)

x$description[x$naics_sector == "100"] <- paste0(
  x$description[x$naics_sector == "100"],
  " Family trusts -> '100' ONLY when the contributor is literally the trust itself ",
  "(name reads as a trust, e.g. 'The Smith Family Trust,' 'Jane Doe 1992 Revocable ",
  "Trust') -- narrow; do not extend to LLCs/holding companies merely owned by or ",
  "affiliated with a trust."
)

x$description[x$naics_sector == "11"] <- paste0(
  x$description[x$naics_sector == "11"],
  " Includes vineyards, even though winemaking might otherwise suggest Beverage ",
  "Manufacturing (31) -- classify a vineyard's primary grape-growing/farming ",
  "operation as agriculture."
)

x$description[x$naics_sector == "52"] <- paste0(
  x$description[x$naics_sector == "52"],
  " Also use this code for integrated health systems that are both a health insurer ",
  "and a hospital/care operator (e.g. Kaiser Permanente, Sutter Health) -- classify ",
  "by their insurance function, not '60'."
)

x$description[x$naics_sector == "92a"] <- paste0(
  x$description[x$naics_sector == "92a"],
  " Includes tribal casinos and other tribal gaming operations -- classify by tribal ",
  "ownership ('92a'), not as gambling/entertainment ('70')."
)
```

```{r}
# write to file
write_csv(x,"03_inputs/naics_sector_title_expanded_with_custom_codes.csv")
```


### write OS level2 categories to csv

```{r}
categories <- read.csv("../01_contributor_pipeline/01_inputs/ContributionOpenSecretsCategories.csv")
```

```{r}
# take out uncoded and unitemized contributions
level2_categories <- unique(categories$level2_category)
level2_categories <- level2_categories[-which(level2_categories == "Uncoded")]
level2_categories <- level2_categories[-which(level2_categories == "Unitemized Contributions")]
level2_categories <- level2_categories[-which(level2_categories == "No Employer Listed or Found")] # also have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Homemakers/Non-income earners")] # we have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Employer Listed/Category Unknown")] # we have a different category for this
level2_categories <- level2_categories[-which(level2_categories == "Generic Occupation/Category Unknown")] # we have a different category for this
```

```{r}
# add in relevant custom categories
level2_categories <- c(level2_categories,"Retired/Homemaker/Student/Unemployed","Unknown","Political Actions Committees (PACs), contributor committees and political action funds")
```

```{r}
write_csv(level2_categories, "../open_secrets_level2_categories.csv")
```





